### 2019-06-25 manuscript/japan_ferns_diversity_ms.Rmd
Create Rmd for ms by copying from word document
JPR日本産シダReview190624.docx received from Atushi Ebihara on 2019-06-24

---
title: "An update and reassessment of fern and lycophyte diversity data in the Japanese Archipelago"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

An update and reassessment of fern and lycophyte diversity data in the Japanese Archipelago

Atsushi Ebihara & Joel H. Nitta

Introduction
Ferns and lycophytes or seed-free vascular plants, also traditionally known as ‘pteridophytes’ are a group of plants which have been paid strong attention as research material of biology. Although estimated number of extant species of ferns and lycophytes is only 11916 (Pteridophyte Phylogeny Group, 2016), much less (approximately one twelfth) than seed plant diversity, the frequency of publication of studies used ferns and lycophytes especially in the field of systematics and evolution seems to be high comparing their proportion on number of species. Thus, ferns and lycophytes are relatively rich organisms for information on their diversity. In last centenary, the diversity of ferns and lycophytes in Japan was quite thoroughly studied, and is probably the third best studied area after Europe and North America. Guo et al. (2003) already noticed that the ferns and lycophyte flora in Japan accompanied with rich trait information was worth analyzing determinants of species diversity, abundance and distribution. Even today, there are scarce examples of analyses on fern diversity pattern in a country or wider scale (e.g. Bogonovich et al. 2014, Mountier et al. 2018). A sole serious problem in the analyses by Gao et al. (2003) was the lack of phylogenetic information at that time, but it was completely solved by subsequent studies (Ebihara 2011, Ebihara et al. 2010). In addition, we should note that the occurrence data they used based on Kurata and Nakaike (1979, 1981, 1983, 1985, 1987, 1990, 1994, 1997) did not covered all the species. Reflecting numerous new occurrences and recircumscription of the species based on the results of recent studies, Ebihara (2016, 2017) provided updated distribution maps for all the native taxa in Japan. We intend to publicize up-to-date diversity datasets on ferns and lycophytes in Japan combined with a review of recent progresses and reassessment of the results in Guo et al. (2003).

Flora
 The studies on fern and lycophyte flora of the Japanese Archipelago initiated with 44 species enumerated in Thunberg (1786)’s “Flora Japonica”. Thereafter the number of recognized species native to there are gradually increasing: 460 species including subspecies and varieties (Makino and Nemoto 1925), ca. 400 species excluding those in Ogasawara (Bonin) and Ryukyu Islands (Tagawa 1959) and 630 species (Iwatsuki 1995). Our latest checklist of native species of Japan ‘FernGreenList ver. 1.01’ (Ebihara et al. 2017a) comprises 691 species (721 taxa including subspecies and varieties [Electronic supplementary material 1]; hereafter call ‘taxa’) belonging to 37 of the 51 families accepted in PPGI (Pteridophyte Phylogeny Group 2016). As almost no uninvestigated area remains in the archipelago, most of the new additions to the flora after the 1990’s are newly segregated species from known ones based on evidences from ploidy levels, enzyme analyses and/or DNA sequences (e.g. Ebihara et al. 2019a, Fujiwara et al. 2018, Hori et al. 2015, 2018a, 2018c, Lin et al. 1994, 1996, Masuyama and Watano 2010, Murakami et al. 1999, Ohta and Takamiya 1999, Serizawa 2015, Takamiya et al. 1997) along with only small number of brand-new records (Ebihara et al. 2014, 2019a). The number of taxa in Japan is nearly identical to that in Taiwan (721 taxa excluding forms and hybrids, Knapp 2014). The strongest floristic affinity is found to the continental part of China sharing 487 taxa (67.5%), almost corresponding to two thirds of the native taxa in Japan (Ebihara 2016, 2017). Accumulated up-to-date information identified 125 endemic taxa (17.3% of the native taxa) to Japan (Ebihara 2018).

Interspecific hybrids
 Hybrid ferns have been widely utilized as material of evolutionary biological studies (Sigel 2016), and recognition of interspecific hybrids is always their starting points. Besides those 721 taxa, a huge number of interspecific natural hybrid taxa (371 combinations are listed in Ebihara 2017) is an outstanding characteristic of the flora. This is much more than known hybrid combinations in North America (xxx), China (10 were formerly accepted in Flora of China [Wu et al. 2013] and 20 or more combinations were suggested in the notes) and Taiwan (13 listed in Knapp 2014). Such high diversity is most likely a result of efforts on seeking new hybrid combinations by Satoru Kurata (1922-1978) and his pupils, but might be, at least in part, by a geographical effect of the long archipelago extending north to south which enable sympatric occurrence of species with different habitat preferences according climate fluctuation (any example???). It is highly likely that most of the interspecific hybrid individuals in a large part of the world are unrecognized yet, including the cases of miscategorization to non-hybrid taxa. We should mind that only a few combinations of hybridity were verified by artificial crossing, enzyme patterns or nuclear DNA markers and the others are still putative hybrids on morphological basis. An expedient method observing irregular size and shape of spores is often adopted for recognizing interspecific F1 hybrids, but is not always true represented by normal-shaped regular spores produced by homoploid hybrids, Botrychium x silvicola (Ebihara et al. 2015, Sahashi 1983) and other putative homoploid hybrids of Botrychium species in Japan (N. Sahashi, personal communication). An incomplete establishment of reproductive isolation was observed between Osmunda lancea, a rheophytic species endemic to Japan and its closest relative O. japonica, and their ‘hybrid’ O. x intermedia is actually hybrid swarms including not only F1 but also F2 or later-generation hybrid (Yatabe et al. 2009, 2011). Fern interspecific hybrids have long been believed to occur in places where sporophytes of their two parent species are sympatric, and this idea was an important ground for identification of putative parentages. However, our studies demonstrated a number of examples of sterile interspecific hybrid unaccompanied by their parental species. One of the extreme cases is found in a dennstaedtioid fern Monachosorum x arakii which was originally considered as an endemic species to Japan. Our study (Ebihara et al. 2016) concluded that it is a sterile hybrid between Monachosorum henryi and M. nipponica, the former is not currently distributed in Japan but in Taiwan, continental China and Himalaya. Although it is somewhat a special case with strong vegetative reproduction ability not only with rhizomes but also with bulbils making relict distribution possible, we should always hypothesize any possible combination of hybrid parentages as flexible as possible. Apart from relict distribution, we obtained strong evidences for contribution of independent gametophytes to formation of hybrid sporophytes in Vandenboschia (Ebihara et al. 2009).

Occurrence data and regional species richness
Eight volumes of "Illustrations of Pteridophytes of Japan" (Kurata and Nakaike, 1979, 1981, 1983, 1985, 1987, 1990, 1994, 1997) were a pioneer work of a nationwide organism distribution survey by citizens in Japan. Along with brief description about each species, these books featured 10 km-grid based distribution maps vouchered by ca. 180000 specimens collected by the members of the Nippon Fernist Club. A large part of the specimens presently deposited in the herbarium of National Museum of Nature and Science (TNS) were recently re-identified accommodating up-to-date species circumscription, and ca. 35000 specimens which are new records on the 10 km-grids were newly collected by the club members. Occurrences based on specimens are usually regarded as presence-only data, but present grid maps can be virtually regarded as presence/absence data as the results of the members’ extensive searching in the field. For the convenience in reusing the occurrence data, a list of specimens which assure the distribution maps in Ebihara (2016, 2017) are publicized here, and it contains 216687 herbarium specimen records (Electronic supplementary material 2). We should note that our policy for the dataset is to include specimens or specimen-images examined by ourselves (only exception is Isoetes occurrences which cited from a specimens list provided by Takamiya et al. [1997]). At least one species is recorded in 4343 of 4853 grids (89.5 %), and the species occurred in the maximum number of grids is Pteridium aquilinum subsp. japonicum (Dennstaedtiaceae) recorded in 2777 grids (57.2%).
Even not included in present dataset, much larger number of specimen data without images are available on-line via Science Museum Net (http://science-net.kahaku.go.jp/) which is a portal of natural history specimens deposited in museums in Japan, and the occurrence data are also searchable in the GBIF data portal (https://www.gbif.org/). As of May 23, 2019, Japan is the country of locality with the largest number of fern and lycophyte specimens in GBIF, totally 531311 occurrences.
In the summary of species richness in 10 km-grids excluding nothotaxa (Fig. 1a), the grids with the highest (212 taxa, grid code: 453034) and the second highest diversity (210 taxa, grid code: 453044) are found in the mountain area of Yakushima Island, located south of Kyushu including the island’s highest peak (Mt. Miyanoura, 1936 m in elevation). The third highest grid (202 taxa, grid code 503547) is located in the area covering Nachi-no-taki Falls in southern part of Wakayama Prefecture, southern Honshu.
By adding phylogenetic information, we are able to overview the diversity distribution patterns by a different measure from species number. We updated the rbcL sequence dataset to cover 97.8 % of the native taxa by making several additions and corrections to that prepared by Ebihara et al. (2010) (Electronic supplementary material 1), and then Phylogenetic Diversity (PD, “the minimum total length of all the phylogenetic branches required to span a given set of taxa on the phylogenetic tree”, Faith 1992, 2006) on a tree for every grid was calculated using Phylogenetic Diversity Analyzer ver. 1.0.3 (Chernomor et al. 2015). In the analysis with a dataset covering all ferns and lycophyte taxa considering the tree unrooted (Fig. 1b), the highest PD value was obtained in a grid in Wakayama Prefecture (grid code: 503547), the second highest one in the northern part of the Okinawa Island (grid code: 402801) and third largest one in Iriomote Island (grid code: 362346) located in the southwesternmost part of the Archipelago. These rankings were not significantly changed when we analyzed a dataset with only fern occurrences rooted by a lycophyte (Lycopodium clavatum var. nipponicum) (Fig. 1c). In the result of PD analyses, the Yakushima grid (grid code: 453034) with maximum species richness was ranked 12th in the PD analysis using ferns+lycophytes, and 10th in that only using ferns, respectively. The disparity of the results might suggest the difference in their constituent species – species richness in Yakushima Island is resulted from presence of some species rich genera (e.g. 19 taxa of Dryopteris, 16 taxa of Diplazium, 14 taxa of Athyrium).
Our updated occurrence data supported the trend that species with seasonal-green phenology showed wider distribution than evergreen species suggested by Guo et al. (2003), but it is probably an artifact caused by small land areas at lower altitudes in the country where evergreen species dominate. Guo et al. (2003) also suggested species with multiple reproductive modes showed wider distribution areas than species with a single reproductive mode. However, our data did not support this trend (mean numbers of grids, sex.+apo. = 356.63 (n = 8), sex. = 367.01 (n = 413), apo. = 314.59 (n = 79), none of the combination differs significantly in paired t-test) – the result seems to be strongly affected by species circumscription and presumably some “species with multiple reproductive modes” in Guo et al. (2003) are separated into multiple taxa with each single mode in present treatment. Even though their distribution area sizes are not significantly different, apomictic taxa show significantly narrower latitudinal range than sexual ones (mean degree of ranges: sex. = 7.73 and apo. 6.01, p < 0.001 in paired t-test), and it coincides with the result by Tanaka et al. (2014) that represented smaller range sizes of apomicitic ones both in latitude and elevation. Also the southerly distribution phenomenon of apomicitic species which was presumed a result of poor cold-hardiness of gametophytes of apomictic species having larger cell sizes (Tanaka et al. 2014), was confirmed in present data (Fig. 1d, e). The trend that narrower distribution of polyploid taxa (Guo et al. 2003) is not supported by present data excluding apomictic species (average number of grids, sexual diploid = 337.14, sexual polyploid = 370.51) – the trend probably came out by narrower distribution of apomictic species which are dominated by triploid.

Cytotaxonomic data
Even today chromosome numbers of ferns and lycophytes which are important clues for knowing the plant’s ploidy level and reproductive mode as well as basic chromosome numbers, which were used as information to infer phylogenetic affinities before the introduction of DNA sequencing. Researchers in Japan stimulated by Irene Manton’s leading work “Problems of Cytology and Evolution in the Pteridophyta” (Manton 1950) actively made chromosome counts of ferns and lycophytes using the squash technique in the second half of the 20th century, and more than 150 papers had been published – all the counts on Japanese material before 1996 was indexed by Takamiya (1996). Additions have been continuously made, and Nakato and Ebihara (2016) summarized that the information drawn from Japanese individuals is available for 74 % of the native taxa (61 %, if nothotaxa are counted) – it is a much higher coverage compared with other areas of East Asia (e.g. ca. 27 % in China, Cheng and Zhang [2010]) and the second highest coverage in Asia following India (ca. 89 %, Bir and Verma [2010]). Further addition is highly needed for ascertaining some early records without preserving voucher specimen, and for detecting infraspecific polyploidy and/or multiple reproductive modes overlooked by previous superficial samplings as well as unsampled taxa. Flow cytometry is recently used for inferring ploidy levels of large number of samples belonging to a closely related species groups (e.g. Ebihara et al. 2005, Fujiwara et al. 2018, Hori et al. 2015, Nitta et al. 2011), but the method always requires at least one control sample whose ploidy level was confirmed by chromosome counts. Flow cytometry analyses using not only fresh leaves but also recently collected spores in herbarium specimens, succeeded in some ferns (Kuo and Huang 2017), might accelerate data accumulation.

Threatened species
 Latest national red list of Japan (Ministry of the Environment, Japan 2019) included more than one third of the native fern and lycophyte taxa (260 in total, 6 EX, 3 EW, 85 CR, 59 EN, 68 VU, 38 NT and 1 DD; see Fig. 1f for the distribution of richness). The largest threat for ferns is herbivory of the sika deer distributed throughout the archipelago which obviously increased after 1980’s (Takatsuki 2009). In particular, species growing on the forest floor (e.g. Athyrium, Terada and Takamiya 2006) undergo severe damage, and several narrowly distributed species in Kyushu and Yakushima Island became almost extinct or only found inside deer fences. Although national law for endangered species is recently expanding objective fern and lycophyte species which are proscribed collection and sellout (28 species as of May 2019), it is apparently not directly effective for preventing herbivory of deer. There are still a few reports of conservation activities and researches on ferns and lycophytes in Japan: ubiquitous genotyping of Athyrium viridescentipes (Izuno et al. 2012) and clarifying origin and propagation from spore in Dryopteris shibipedis (Ebihara et al. 2012) which was extinct in the wild. Despite 70 of the 260 taxa in the latest national red list are endemic species to Japan, they are neither assessed nor listed in the IUCN Red List as globally threatened species as of May 2019 (IUCN 2019).

Introduced species
Compared with flowering plants, number of naturalized species of ferns and lycophytes is quite small in the Japanese Archipelago, and they are almost uninvolved with the invasive species issue. Selaginella uncinata (Selaginellaceae) and Pityrogramma calomelanos (Pteridaceae) are two of scarce examples of relatively widespread introduced species in the country (Ebihara 2016). Dryopteris intermedia (ref), D. carthusiana (ref) (Dryopteridaceae) and Dennstaedtia punctilobula (Dennstaedtiaceae) (Ueno et al. 2019) were probably introduced with sprayed seeds on banks and only collected in each one or a few localities. Azolla cristata (Salviniaceae) is the only invasive fern species controlled by national low (Invasive Alien Species Act), but is often troubled in identification due to morphologically similarlity to an endangered and endemic species, A. japonica. Another similar species, Azolla filiculoides is still controversial on its natural distribution in Japan (Ebihara 2016, Suzuki et al. 2005). Selaginella moellendolffii (Selaginellaceae), Psilotum nudum (Psilotaceae), Adiantum capillis-veneris, Pteris vittata (Pteridaceae) and Thelypteris dentata (Thelypteridaceae) have certain natural distribution in the subtropical areas of the country, but are more frequently found as escaped plants from greenhouses. Murakami et al. (2008) demonstrated a range expansion of Thelypteris dentata facilitated by urbanization in Kinki District, Japan. The natural habitat of Cyrtomium falcatum subsp. falcatum is restricted to coastal areas (Ebihara et al. 2017a, Matsumoto 2003), and scattered occurrences in urban environments (Ebihara 2017) in inlands are probably introduced ones.

Reticulate evolution
 Species complexes or aggregate species exhibiting wide ranges of continuous morphological variation are quite commonly known in ferns and lycophytes (Haufler 1996), and this trend seems to be true for the flora of Japan. In case of that interspecific F1 hybrids make the species boundaries ambiguous, we could usually find distinct gaps between them by removing their hybrid individuals suggested by irregular-shaped spores with low germination rates (e.g. Ohta and Takamiya 1999 in the Diplazium mettenianum complex). When interspecific hybrids overcome sterility by chromosome doubling, the situation becomes more complicated and the biological units composing the complex are not easily clarifiable solely by morphological approach. The allopolyploids of hybrid origins often re-hybridize with their parental species --- such "reticulate evolution" process became well-known after the study on Appalachain Asplenium (Wagner 1954). Apparently similar approach combining chromosome and morphological observations was rarely adopted to Japanese species. After 1990's, allozyme analyses were frequently applied to the fern complexes in Japan, and plastid sequences often accompanied as indicators of maternal lineage (e.g. Darnaedi et al. 1990 in Dryopteris yakusilvicola and its related species, Lin et al. 1996 in Odontosoria, Shinohara et al. 2010 in Lepisorus, Takamiya and Ohta 2001 in Diplazium). In 2000's, single or low copy nuclear DNA regions became popularly used as biparently inherited markers (e.g. Ishikawa et al. 2002), and sequence polymorphisms were efficiently isolated by cloning or the SSCP (Single Strand Conformation Polymorphism) method (initiated by Ebihara et al. 2005 in the Vandenboschia radicans complex, thereafter applied to various fern groups including Ceratopteris [Adjie et al. 2007]), the Pteris cretica complex [Jaruwattanaphan et al. 2013], the Davallia repens complex [Chen et al. 2014], the Asplenium normale complex [Fujiwara et al. 2017]). Although next-generation sequencing is widely used in phylogenetic and population genetic studies on various organisms in recent years, it is still not popularly applied to analyses on species complexes of ferns and lycophytes mainly due to technical difficulties in removing chimeral contigs identifying and homologous copies from the obtained sequences. More recently Rothfels et al. (2017) succeeded in phylogenetic analysis on a polyploid complex of cystoptreroid ferns using PacBio sequences, and the method is highly expected to apply wide ranges of fern complexes.

Species complexes involved with apomixis
 Further complexity is provided when apomictic reproduction (reviewed in Grusz et al. 2016) is concerned. The topic, apomixis or apogamy in ferns has attracted researchers in Japan as the reproductive mode is known in a higher rate in the native taxa of the country (13 % of the species whose reproductive modes are known in Takamiya 1996; 12 % in present study) than the average rate in the world (ca. 10 %, Grusz et al. 2016). Apomixis plays a role not only in overcoming sterility of interspecific hybrids (e.g. Park and Kato 2003) but also making secondary hybridization by fertilization between a sperm produced by the ‘hybrid’ and an egg of sexually reproducing races. This process can generates countless apomict races with various genomic combinations which are morphologically continuous, and the races are often controversial about circumscription of ‘species’ (a polymorphic species versus numerous morphologically ill-defined species). One of the best known apomict complexes troublesome in identification, the Dryopteris varia complex was recently clarified their internal relationships using single-copy nuclear DNA markers (Hori et al. 2014, 2015, 2016), and analyses on several other complexes are in progress. However, even using molecular markers, we often meet an obstacle that we cannot find progenitors or sexual diploids of the complex. Hori et al. (2018d) discovered several sexual diploid individuals of Dryopteris in southern China, and as the progenitors the newly found diploids well explained the reticulate evolution processes of the D. erythrosora complex in Japan which had bothered researchers for many years (Hori et al. 2018b). This case supported the hypothesis by Gao et al (2003): “The ancestral sexual species of many Japanese apogamous species occur in China, suggesting that these species might have migrated from China”. In contrast to these successful clarifications, we are still struggling with more challenging complexes. The species of Cyrtomium (Dryopteridaceae) in Japan typically match this pattern: 12 native species and subspecies are all triploid apomicts except for two subspecies of C. falcatum (subsp. australe and subsp. littorale, sexual diploid) and C. devexiscapulae (sexual tetraploid) (Ebihara et al. 2017a, Matsumoto 2003). Despite careful exploration of sexually reproducing individuals by examining spore number per sporangium in herbarium specimens collected in China, a large part of them are apomictic and sexual ones were found less frequently in geographically narrow ranges (Mitsuta 1986, 1988). The Pteris fauriei complex is probably in similar situation, and it includes several endemic ‘species’ to Japan which are diploid or triploid apomict (Ebihara 2016). The “Diploids-First Approach” (Beck et al. 2010) is a principle for the studies on fern reticulate complexes, but we should mind that diploids are not always present nearby including possible extinction in complexes involved with apomixis.

Species concept
 It is obvious that recent increase of recoded taxa in Japan is largely owing to transition from morphological species concept to biological or evolutional ones. In many cases, subsequent taxonomic revisions were published after analyses on certain species complexes. Although we have not reached any final consensus on the concept, strict application of neither biological species concept nor evolutionary species one seem to be preferred by recent studies. In particular, an autotetraploid, which should be treated as an independent species from a diploid sharing same genome in the biological species concept, is more often treated as infraspecific polyploidy. Ebihara et al. (2005, 2009) circumscribed a new allotetraploid species Vandenboschia hokurikuensis including recurrently originated (at least twice, based on maternally inherited plastid sequences). In practice, every allopolyploid with different combination of diploid progenitor species requires a name. Among the 195 taxa in Japan which sexual polyploid (tetraploid or higher) was reported for, evidences for judging autopolyploidy/allopolyploidy are available in 34 taxa (17.4 %). Only four of them (Asplenium boreale [Fujiwara et al. 2017], Asplenium pekinense [Lin et al. 2013], Pteris deltodon [Jaruwattanaphan et al. 2013], and Lepisorus tosaensis [Fujiwara et al. 2018]) are probable autopolyploid (autotetraploid) and the remaining 30 taxa are allopolyploid. Some of currently accepted infraspecific taxa defined to correspond to ploidy levels (e.g. diploid Lycopodium clavatum var. nipponicum vs. tetraploid var. asiaticum [Takamiya 1989]; tetraploid Deparia petersenii var. petersenii vs. hexaploid var. yakusimensis [Shinohara et al. 2003]) might be classified into different species if future studies clarify differences in their genomic combinations.
 We are often faced with a problem about taxonomic treatment for apomictic races. As they are asexually reproducing, the biological species concept is not applicable, and in fact several studies had no other choice than suspension of corresponding taxonomic treatment (e.g. Chen et al. 2014, Nitta et al. 2011). Another trouble is on their nomenclatural categories: i.e. should we describe apomictic races of hybrid origin as nothotaxa? Recircumscription of species recently made by Hori et al. (2018a) for the Dryopteris varia complex is probably the most easily acceptable decision, which is, providing a species name at non-hybrid category for every apomictic race with different combination of sexually reproducing progenitors. This treatment resulted a split of Dryopteris hikonensis to three apomictic species namely D. hikonensis s.s. (= D. protobissetiana x D. varia), D. erythrovaria (= D. caudipinna x D. protobissetiana x D. varia) and D. subhikonensis (= D. protobissetiana x D. saxifraga x D. varia). It should be noted that we treated some individuals originated from hybridization between an apomictic race and sexual species which showing lower spore germination rates and/or growing not vigorously in the wild as nothotaxa in FernGreenList (Ebihara et al. 2016). The categorization largely depends on our observation and experiences, and we could not always strictly border them.
 There is no standard taxonomic treatment for the case that both sexual race and apomictic race coexist in a single species. While in many cases both reproductive modes are not distinguished taxonomically (e.g. Pteris oshimensis, Phegopteris connectilis and Dryopteris chinensis), the apomicitc race of Pteris terminalis are distinguished at varietal rank (var. fauriei, Ebihara et al. 2017b) considering its morphological distinctness and failure of finding unique allele in the apomictic race.

Gametophytes
Our knowledge about gametophytes of ferns and lycophytes is considerably limited compared with that in sporophytes, nevertheless Japan is one of the world’s most information-rich areas for fern gametophytes. Two dennstaedtioid genera, Coptodipteris (‘Coptidipteris’) and Fuziifilix described by Nakai and Momose (1937) are scarce examples segregated based on gametophyte characters, even though they are probably merged into one of the two clades of Dennstaedtia (Perrie et al. 2015). Momose (1967) observed and illustrated gametophytes of ca. 330 species mostly native to the country using cultured ones from spores. Although he succeeded in culture almost only for cordate gametophytes, subsequent researchers observed non-cordate gametophytes which grow and maturate usually slower than cordate ones (e.g. Hymenophyllaceae [Yoroi 1972, 1976, 1992], Haplopteris (Pteridaceae) [Yoroi 1975], Pleurosoriopsis (Polypodiaceae) [Ebihara et al. 2019b, Masuyama 1975] and Botrychium (Ophioglossaceae) [Takahashi and Imaichi 2007]).
 Ebihara et al. (2010) completed setting up a country scale DNA barcode library (plastid rbcL and trnH-psbA regions, covering ca. 94 % of native fern and lycophyte taxa) which is quite useful for identification of gametophytes in the wild. Our preliminary sampling demonstrated an interesting phenomenon that non-cordate gametophytes frequently ‘independent’ or occur far from their counterpart sporophytes localities while cordate ones are almost always sympatric with conspecific sporophytes (Ebihara et al. 2013). This is also true for independent filamentous gametophytes of Vandenboschia as mentioned above, and some of the independent gametophytes might be solely gametophytes distributed in the country (Haplopteris sp. in Yakushima Island, Kuo et al. 2017). Expanded samplings of gametophytes in the wild will shed light on more and more independent ones. In another aspect, identification using DNA barcode enabled detection of fungal association of wild fern gametophytes (Ogura-Tsujita et al. 2013, 2016). Although they discussed possible correlation between gametophyte cushion structure and arbuscular mycorrhizal association, much wider taxonomic and geographical samplings are necessary for understanding the evolutional trend as well as for comparison with symbiotic fungi in sporophytes (reviewed by Lenhert et al. 2016, Pressel et al. 2016).

Taxonomic key characters
Ebihara (2016, 2017) publicized matrices for major morphological characters frequently used for identification keys for all the 721 native ferns and lycophytes in Japan. All the quantitative characters were newly measured by the members of the Nippon Fernist Club in ten randomly selected herbarium specimens per taxon. As these data are much more useful in an electric form than a printed version, we provide the data file (Electronic supplementary material 3) in the format for Lucid 3.3, which is a free version of the software for interactive keys available at http://www.lucidcentral.com/en-us/home.aspx.

Future perspective
 Albeit none of the genus accepted in PPGI (Pteridophyte Phylogeny Group 2016) is endemic to the archipelago, recent studies revealed several infrageneric clades distributed centered in and/or representing the highest diversity there. Spicantopsis, the most recently resurrected genus in Blechnaceae is now subendemic to Japan (2 endemic species and 1 species in Japan and Taiwan) and its ancestor estimated to migrate from North America to East Asia in ca. 85 mya (Gabriel y Galan et al., in press). Apart from Spicantopsis, we could find at least three groups exhibiting significant diversity in there: Deparia subsect. Athyriopsis (Kuo et al. 2016, 2018), “Clades VI + VII” of Lepisorus (Wang et al. 2009), “the HYSUFI clade” of Polystichum (Le Pechon et al. 2016). Clarification of their speciation and biogeographic histories is expected to contribute for understanding flora formation in East Asia and the Far East. 
 Regional flora should never conclude in itself but be compared with those of neighboring regions toward a global view. In widespread species or species complexes, we often troubled when applying the results of evidential revisional works with limited geographical coverage of samplings to regional flora of unsampled areas. As Shrestha and Zhang (2014) demonstrated in the Chinese populations, Huperzia serrata sensu lato (Lycopodiaceae) is never a single species but an aggregation of multiple biological units, and it was supported by DNA content analyses using populations in Aichi Prefecture of Japan (Aman et al. 2011). Unfortunately, these findings were not reflected but recognized only Huperzia serrata s.l. in FernGreenList 1.0.1 (Ebihara et al. 2016) as we could not drew reliable synthesis of these results for applying Japanese populations. There are a number of native species or groups in Japan which are awaiting careful comparison with published results mostly generated in China and other Asian countries (e.g. Pteridium [Zhou et al. 2015], Hymenasplenium [Xu et al. 2018]) and Leptochilus [Zhang et al. 2019].
 Contrasting to a consensus on familial classification of ferns and lycophytes almost achieved in PPGI (Pteridophyte Phylogeny Group 2016), generic classification proposed in the system requires a good deal of update and emendation based on newly generated evidences. For example, the 30-generic system of Thelypteridaceae adopted in the PPGI system seems hardly applicable to the species in Japan due to generation of a nothogenus x Chrinephrium (x C. insulare) and due to uncertain placement for the species formerly included in Parathelypteris (e.g. the material of “Parathelypteris nipponica” from China used in the phylogeny of He and Zhang (2012) and Almeida et al. (2016) is in question for the identification). It is important to publicize data on East Asian material with accurate identification working as bases for generic recircumscription in forthcoming “PPGII”.
 Accumulation of data on ploidy and reproductive modes along with recognition of proper taxonomic units and their distribution data enables us to discuss about various characteristics of ferns and lycophytes as an organism group such as a proportion of polyploids and a proportion of apomict. In a country scale, combined analyses between the occurrence data provided by the present study and various kinds of geographic variables may yield further new findings but we here confine preliminary ones. We finally note an interesting finding which is not exposed in present maps: a hotspot of diploid progenitors which are probable palaeoendemics on the riverside in Amami-oshima Island, central Ryukyu. The hotspot is represented by at least four examples: 1) Polystichum obae (Dryopteridaceae), an endemic diploid species (Takamiya 1996) with a quite small population, and is sister to a clade consisting of tetraploid P. polyblepharon (Takamiya 1996) + diploid P. parvipinnulum (Tsai and Shieh 1985) in plastid phylogeny (Le Pechon et al. 2016); 2) Coniogramme gracilis (Pteridaceae), a diploid species (Kurita 1972) endemic to the island, and is presumably a progenitor of widespread tetraploid C. japonica (Kurita 1963); 3) Diplazium amamiana (Athyriaceae), a diploid species (Takamiya et al. 1999) endemic to the island, and sister to widespread apomictic species D. hachijoense in plastid phylogeny (Ebihara 2011, Wei et al. 2013); 4) Pteris oshimensis (Pteridaceae), a widespread species, and sexual diploid is only found in Amami Islands in contrast to widespread apomictic races (Nakato and Ebihara 2016). The pattern, sexual diploid confined to Amami Islands vs. widespread polyploids, shared by these four pairs strongly suggests that the island is a relic for ancestral diploids. However, this may be one of many outcomes from our present datasets of ferns and lycophytes in Japan. It is highly expected to be drawn evolutionary scenarios for formation of flora and diversity by accumulation and integration of huge but reliable information.

References
Adjie B, Masuyama S, Ishikawa H, Watano Y (2007) Independent origins of tetraploid cryptic species in the fern Ceratopteris thalictroides. J Plant Res 120:129-138
Almeida TE, Hennequin S, Schneider H, Smith AR, Batista JAN, Ramalho AJ, Proite K, Salino A (2016) Towards a phylogenetic generic classification of Thelypteridaceae: Additional sampling suggests alterations of neotropical taxa and further study of paleotropical genera. Molec Phylogen Evol 94: 688-700
Aman A, Kato J, Serizawa S (2011) Studies on the Huperzia serrata group (Lycopodiaceae) in Japan (1). Nuclear DNA contents in the populations of Aichi prefecture. Shidekobushi 2:23-32 (in Japanese with English summary)
Beck JB, Windham MD, Yatskievych G, Pryer KM (2010) A diploids-first approach to species delimitation and interpreting polyploid evolution in the fern genus Astrolepis (Pteridaceae). Syst Bot 35:223-234
Bir SS, Verma SC (2010) Chromosome Atlas of the Indian Pteridophytes (1951-2009). Bishen Singh Mahendra Pal Singh, Dehra Dun
Bogonovich M, Robeson S, Watson M (2014) Patterns of North American fern and lycophyte richness at three taxonomic levels. Amer Fern J 103:193-214
Chen CW, Ngan LT, Hidayat A, Evangelista L, Nooteboom HP, Chiou WL (2014) First insights into the evolutionary history of the Davallia repens complex. Blumea 59:49-58
Cheng X, Zhang SZ (2010) Index to chromosome numbers of Chinese Pteridophyta (1969-2009). J Fairylake Bot Gard 9:1-58
Chernomor O, Minh BQ, Forest F, Klaere S, Ingram T, Henzinger M, von Haeseler A (2015) Split diversity in constrained conservation prioritization using integer linear programming. Methods Ecol Evol 6:83-91 DOI: 10.1111/2041-210X.12299
Darnaedi D, Kato M, IwatsukiK (1990) Electrophoretic evidence for the origin of Dryopteris yakusilvicola (Dryopteridaceae). Bot Mag (Tokyo) 103:1-10
Ebihara A (2011) RbcL phylogeny of Japanese pteridophyte flora and implications on infrafamilial systematics. Bull Nat Mus Nat Sci B 37:63-74
Ebihara A (2016) The Standard of Ferns and Lcyophytes in Japan 1. Gakken Plus, Tokyo (in Japanese) 
Ebihara A (2017) The Standard of Ferns and Lcyophytes in Japan 2. Gakken Plus, Tokyo (in Japanese)
Ebihara A (2018) A checklist of ferns and lycophytes endemic to Japan (edition March 2018) http://www.kahaku.go.jp/research/activities/project/hotspot_japan/endemic_list/
Ebihara A, Ishikawa H, Matsumoto S, Lin SJ, Iwatsuki K, Takamiya M, Watano Y, Ito M (2005) Nuclear DNA, chloroplast DNA, and ploidy analysis clarified biological complexity of the Vandenboschia radicans complex (Hymenophyllaceae) in Japan and adjacent areas. Amer J Bot 92:1535-1547.
Ebihara A, Ito M, Nagamasu H, Fujii S, Katsuyama T, Yonekura K, Yahara, T (2016) Fern GreenList ver. 1.01 http://www.rdplants.org/gl/
Ebihara A, Matsumoto S, Ito M (2009) Hybridization involving independent gametophytes in the Vandenboschia radicans complex (Hymenophyllaceae): a new perspective on the distribution of fern hybrids. Molec Ecol 18:4904-4911
Ebihara A, Matsumoto S, Kato M (2012) Origin of Dryopteris shibipedis (Dryopteridaceae), a fern species extinct in the wild. J Plant Res 125:499-505
Ebihara A, Matsumoto S, Mazumdar J, Yamamoto K (2017a) Updates of taxonomic treatments for ferns of Japan 2. Athyrium and Cyrtomium. Bull Natl Mus Nat Sci B 43:19-25
Ebihara A, Morita H, Yamamuro K (2019a) Diplazium megaphyllum (Athyriaceae) new to Japan, discovered in Tokunoshima Island. J Jpn Bot 93:404-406 (in Japanese with English summary)
Ebihara A, Nakato N, Amoroso VB, Hidayat A, Kuo LY (2016) Monachosorum arakii Tagawa (Dennstaedtiaceae) is a relict "international" hybrid: A reassessment of the Monachosorum species. Syst Bot 41:586-595
Ebihara A, Nakato N, Jaruwattanaphan T (2017b) A new taxonomic treatment for the apogamous counterpart of Pteris terminalis (Pteridaceae). Phytotaxa 314:73-80
Ebihara A, Nakato N, Kuo LY, Miyazaki H, Serizawa S (2019b) Allopolyploid origin and distribution range of Acystopteris taiwaniana (Cystopteridaceae: Polypodiales). Acta Phytotax Geobot 70:19-28
Ebihara A, Nakato N, Matsumoto S (2015) Updates of taxonomic treatments for ferns of Japan 1. Botrychium, Osmolindsaea and Pteris. Bull Natl Mus Nat Sci B 41:15-24
Ebihara A, Nakato N, Saito Y, Oka T, Minamitani T (2014) New records of Asplenium varians (Aspleniaceae) and two new hybrids in Japan. Acta Phytotax Geobot 65:53-65
Ebihara A, Nitta JH, Ito M (2010) Molecular species identification with rich floristic sampling: DNA barcoding the pteridophyte flora of Japan. PLoS ONE 5(12):e15136
Ebihara A, Nitta JH, Matsumoto Y, Fukazawa Y, Kurihara M, Yokote H, Sakuma K, Azakami O, Hirayama Y, Imaichi R (2019b) Growth dynamics of independent gametophytes of Pleurosoriopsis makinoi (Polypodiaceae). Bull Nat Mus Nat Sci B 45:77-86
Ebihara A, Yamaoka A, Mizukami N, Sakoda A, Nitta JH, Imaichi R (2013) A survey of the fern gametophyte flora of Japan: Frequent independent occurrences of noncordiform gametophytes. Amer J Bot 100:735-743
Faith DP (1992) Conservation evaluation and phylogenetic diversity. Biol Conserv 61:1-10
Faith DP, Baker AM (2006) Phylogenetic diversity (PD) and biodiversity conservation: some bioinformatics challenges. Evol Bioinform Online 2006:2 121-128
Fujiwara T, Serizawa S, Watano Y (2018) Phylogenetic analysis reveals the origins of tetraploid and hexaploid species in the Japanese Lepisorus thunbergianus (Polypodiaceae) complex. J Plant Res 131:945-959
Fujiwara T, Uehara A, Iwashina T, Matsumoto S, Chang YH, Chao YS, Watano Y (2017) Allotetraploid cryptic species in Asplenium normale in the Japanese Archipelago, detected by chemotaxonomic and multi-locus genotype approaches. Amer J Bot 104:1390-1406
Gabriel y Galan JM, Molino S, Sessa EB, Wasowicz P (in press) A multi-character analysis of Struthiopteris leads to the rescue of Spicantopsis (Blechnaceae, Polypodiopsida). Taxon
Grusz AL (2016) A current perspective on apomixis in ferns. J Syst Evol 54:656-665
Haufler CH (1996) Species concepts and speciation in pteridophytes.In: Camus JM, Gibby M, Johns RJ eds. Pteridology in perspective. Proceedings of the Holttum Memorial Pteridophyte Symposium, Kew 1995. Kew: Royal Botanical Gardens. pp. 291-305
Guo Q, Kato M, Ricklefs RE (2003) Life history, diversity and distribution: a study of Japanese pteridophytes. Ecography 26:129-138
He LJ, Zhang XC (2012) Exploring generic delimitation within the fern family Thelypteridaceae. Molec Phylogen Evol 65:757-764 
Hori K, Ebihara A, Murakami N (2018a) Revised classification of the species within the Dryopteris varia complex (Dryopteridaceae) in Japan. Acta Phytotax Geobot 69:77-108
Hori K, Ebihara A, Nakato N, Murakami N (2015) Dryopteris protobissetiana (Dryopteridaceae), a new diploid sexual species of the Dryopteris varia complex (subg. Erythrovariae, sect. Variae) from Yakushima, Kagoshima, Japan. Acta Phyotax Geobot 66:47-57
Hori K, Tono A, Fujimoto K, Kato J, Ebihara A, Watano Y,  Murakami N (2014) Reticulate evolution in the apogamous Dryopteris varia complex (Dryopteridaceae, subg. Erythrovariae, sect. Variae) and its related sexual species in Japan. J Plant Res 127:661-684
Hori K, Watano Y, Murakami N (2016) Hybrid origin of the apogamous fern Dryopteris hondoensis (Dryopteridaceae). Acta Phytotax Geobot 67:133-146
Hori K, Zhou XL, Shao W, Yan YH, Ebihara A, Wang RX, Ishikawa H, Watano Y, Murakami N (2018b) Hybridization of the Dryopteris erythrosora complex (Dryopteridaceae, Polypodiidae) in Japan and adjacent areas. Hikobia 17:299-313
Hori K, Zhou XL, Shao W, Yan YH, Wang RX, Murakami N (2018d) New diploid sexual cytotypes of Dryopteris sect. Erythrovariae (Dryopteridaceae) in China. Acta Phytotax Geobot 69:127-133
Hori K, Zhou XL, Yan YH, Inoue Y, Murakami N (2018c) Evidence for maternal ability in hybridization of apogamous fern species: Dryopteris tsushimense K. Hori & N. Murak. and D. subtsushimense K. Hori & N. Murak. (Dryopteridaceae), new tetraploid apogamous pteridophytes of hybrid origin from Tsushima, Japan. Acta Phytotax Geobot 69:143-160
Ishikawa H, Watano Y, Kano K, Ito M, Kurita S (2002) Development of primer sets for PCR amplification of the PgiC gene in ferns. J Plant Res 115:65-70
IUCN (2019) The IUCN Red List of Threatened Species. Version 2019-1. http://www.iucnredlist.org. Downloaded on 21 March 2019
Iwatsuki K, Yamazaki T, Boufford DE, Ohba H (1995) Flora of Japan, vol I. Kodansha, Tokyo
Izuno A, Takamiya M, Kaneko S, Isagi Y (2012) Genetic variation and structure of the endangered lady fern Athyrium viridescentipes based on ubiquitous genotyping. J Plant Res 125:613-618
Jaruwattanaphan T, Matsumoto S, Watano Y (2013) Reconstructing hybrid speciation events in the Pteris cretica group (Pteridaceae) in Japan and adjacent regions. Syst Bot 38:15-27
Knapp R (2014) Commented list of Pteridophytes of Taiwan (ed. 3.1) https://sites.google.com/site/fernsoftaiwan/
Kuo LY, Chen CW, Shinohara W, Ebihara A, Kudoh H, Sato H, Huang YM, Chiou WL (2017) Not only in the temperate zone: independent gametophytes of two vittarioid ferns (Pteridaceae, Polypodiales) in East Asian subtropics. J Plant Res 130:255-262
Kuo LY, Ebihara A, Hsu TC, Rouhan G, Huang YM, Wang CN, Chiou WL, Kato M (2018) Infrageneric revision of the fern genus Deparia (Athyriaceae, Aspleniineae, Polypodiales). Syst Bot 43:645-655
Kuo LY, Ebihara A, Shinohara W, Rouhan G, Wood KR, Wang CN, Chiou WL (2016) Historical biogeography of the fern genus Deparia (Athyriaceae) and its relation with polyploidy. Molec Phylogen Evol 104:123-134
Kuo LY, Huang YM (2017) Determining genome size from spores of seedless vascular plants. Bio Prtocol 7(11):e2322
Kurata S, Nakaike T (1979) Illustrations of Pteridophytes of Japan, vol 1. University of Tokyo Press, Tokyo (in Japanese)
Kurata S, Nakaike T (1981) Illustrations of Pteridophytes of Japan, vol 2. University of Tokyo Press, Tokyo (in Japanese)
Kurata S, Nakaike T (1983) Illustrations of Pteridophytes of Japan, vol 3. University of Tokyo Press, Tokyo (in Japanese)
Kurata S, Nakaike T (1985) Illustrations of Pteridophytes of Japan, vol 4. University of Tokyo Press, Tokyo (in Japanese)
Kurata S, Nakaike T (1987) Illustrations of Pteridophytes of Japan, vol 5. University of Tokyo Press, Tokyo (in Japanese)
Kurata S, Nakaike T (1990) Illustrations of Pteridophytes of Japan, vol 6. University of Tokyo Press, Tokyo (in Japanese)
Kurata S, Nakaike T (1994) Illustrations of Pteridophytes of Japan, vol 7. University of Tokyo Press, Tokyo (in Japanese)
Kurata S, Nakaike T (1997) Illustrations of Pteridophytes of Japan, vol 8. University of Tokyo Press, Tokyo (in Japanese)
Kurita S (1963) Cytotaxonomy of Japanese Coniogramme Fée. J Jpn Bot 38:42-46
Kurita S (1972) Chromosome numbers of some Japanese ferns (8). Ann Rep For Stud　Coll Chiba Univ 7:47-53
Le Péchon T, He H, Zhang L, Zhou XM, Gao XF, Zhang LB (2016) Using a multilocus phylogeny to test morphology-based classifications of Polystichum (Dryopteridaceae), one of the largest fern genera. BMC Evol Biol 16:55
Lehnert M, Krug M, Kessler M (2017) A review of symbiotic fungal endophytes in lycophytes and ferns--a global phylogenetic and ecological perspective. Symbiosis 71:77-89
Lin SJ, Kato M, Iwatsuki K (1994) A taxonomic study of the fern genus Sphenomeris (Lindsaeaceae) in Japan. J Jpn Bot 69:127-141
Lin SJ, Kato M, Iwatsuki K (1996) Electrophoretic evidence for the origins of tetraploids and hybrids of Sphenomeris chinensis sensu lato (Lindsaeaceae) in Japan. J Plant Res 109:201-209
Lin YX, Viane R (2013) Aspleniaceae. Pp. 267-316 in Wu ZY, Raven PH, Hong DY eds., Flora of China, Vol. 2-3 (Pteridophytes). Beijing: Science Press; St. Louis: Missouri Botanical Garden Press
Makino T, Nemoto K (1925) Nippon-Shokubutsu-Soran (Flora of Japan) with Descriptions of Every Plant Phanerograms and Higher Cryptogams Indigenous to, Introduced into and Cultivated in the Empire of Japan Kabafuto, Hokkaido, Honshu, Shikoku, Kiushiu, Riukiu and Taiwan. Nippon-Shokubutsu-Soran Kankokai, Tokyo (in Japanese)
Manton I (1950) Problems of Cytology and Evolution in the Pteridophyta. Cambridge University Press
Masuyama S (1983) The gametophyte of Pleurosoriopsis makinoi (Maxim.) Fomin. J Jpn Bot 50:105-114
Masuyama S, Watano Y (2010) Cryptic species in the fern Ceratopteris thalictroides (L.) Brongn. (Parkeriaceae). IV. Taxonomic revision. Acta Phytotax Geobot 61:75-86
Matsumoto S (2003) Species ecological study on reproductive systems and speciation of Cyrtomium falcatum complex (Dryopteridaceae) in Japanese Archipelago. Ann Tsukuba Bot Gard 22:1-141 (in Japanese with English summary)
Ministry of the Environment, Japan (2019) Red List of Japan, 2019 https://www.env.go.jp/press/106383.html (in Japanese)
Mitsuta S (1986) A preliminary report on reproductive type of Cyrtonium (Dryopteridaceae). Acta Phytotax Geobot 37:117-122 (in Japanese with English summary)
Momose S (1967) Prothallia of the Japanese Ferns (Filicales). University of Tokyo Press, Tokyo (in Japanese)
Mountier CF, Case BS, Perrie L, Brownsey P, Paterson AM, Curran TJ, Buckley HL (2018) Patterns of range size in New Zealand ferns and lycophytes. NZ J Ecol 42:248-261
Murakami K, Matsui R, Morimoto Y (2007) Northward invasion and range expansion of the invasive fern Thelypteris dentata (Forssk.) St. John into the urban matrix of three prefectures in Kinki District, Japan. Amer Fern J 97:186-199
Murakami N, Watanabe M, Yokoyama J, Yatabe Y, Iwasaki H, Serizawa S (1999) Molecular taxonomic study and revision of the three Japanese species of Asplenium sect. Thamnopteris. J Plant Res 112:15-25
Nakai T, Momose S (1937) A morphological and taxonomical study on Japanese Microlepia. Cytologia Fujii Jubilaei:360-365
Nakato N, Ebihara A (2016) Chromosome numbers of 18 ferns in Japan: toward completion of chromosome information in Japanese ferns. Bull Natl Mus Nat Sci B 42:25-40
Nitta JH, Ebihara A, Ito M (2011) Reticulate evolution in the Crepidomanes minutum species complex (Hymenophyllaceae). Amer J Bot 98:1782-1800
Ogura-Tsujita Y, Sakoda A, Ebihara A, Yukawa T, Imaichi R (2013) Arbuscular mycorrhiza formation in cordate gametophytes of two ferns, Angiopteris lygodiifolia and Osmunda japonica. J Plant Res 126:41-50
Ogura-Tsujita Y, Hirayama Y, Sakoda A, Suzuki A, Ebihara A, Morita N, Imaichi R (2016) Arbuscular mycorrhizal colonization in field-collected terrestrial cordate gametophytes of pre-polypod leptosporangiate ferns (Osmundaceae, Gleicheniaceae, Plagiogyriaceae, Cyatheaceae). Mycorrhiza  26:87-97
Ohta N, Takamiya M (1999) Taxonomic studies of the Diplazium mettenianum complex (Woodsiaceae; Pteridophyta) in Japan: Morphology, cytology and taxonomy of plants with normal-shaped spores. J Plant Res 112:67-86
Park CH, Kato M (2003) Apomixis in the interspecific triploid hybrid fern Cornopteris christenseniana (Woodsiaceae). J Plant Res 116:93-103
Perrie LR, Shepherd LD, Brownsey PJ (2015) An expanded phylogeny of the Dennstaedtiaceae ferns: Oenotrichia falls within a non-monophyletic Dennstaedtia, and Saccoloma is polyphyletic. Austral Syst Bot 28:256-264
Pressel S, Bidartondo MI, Field KJ, Rimington WR, Duckett JG (2016) Pteridophyte fungal associations: current knowledge and future perspectives. J Syst Evol 54:666-678
The Pteridophyte Phylogeny Group (2016) A community-derived classification for extant lycophytes and ferns. J Syst Evol 54:563-606
Rothfels CJ, Pryer KM, Li FW (2017) Next-generation polyploid phylogenetics: rapid resolution of hybrid polyploid complexes using PacBio single-molecule sequencing. New Phytol 213:413-429
Sahashi N (1983) Morphological and taxonomical studies on Ophioglossales in Japan and the adjacent regions (9). Additional notes on Sceptridium in Isl. Oshima, Izu Islands. J Jpn Bot 58:338-344
Serizawa S (2015) Misecellaneous notes on Japanese pteridophytes (6). Shidekobushi 3:39-59
Shinohara W, Takamiya M, Murakami N (2003) Taxonomic study of Japanese Deparia petersenii (Woodsiaceae) based on cytological and molecular information. Acta Phytotax Geobot 54:137-148
Shinohara W, Ushio Y, Seo A, Nakato N, Kono M, Kudoh H, Tobe H, Murakami N (2010) Evidence for hybrid origin and segmental allopolyploidy in eutetraploid and aneutetraploid Lepisorus thunbergianus (Polypodiaceae). Syst Bot 35: 20-29
Shrestha N, Zhang XC (2015) Recircumscription of Huperzia serrata complex in China using morphological and climatic data. J Syst Evol 53:88-103
Sigel EM (2016) Genetic and genomic aspects of hybridization in ferns. J Syst Evol 54:638-655
Suzuki T, Watanabe I, Shiraiwa T (2005) Allozyme types of water fern Azolla japonica and its relatives (Azollaceae) growing in Japan. Acta Phytotax Geobot 56:71-83
Tagawa M (1959) Coloured Illustrations of the Japanese Pteridophyta. Hoikusha, Osaka (in Japanese)
Namegata T, Kurata S (1961) Enumeration of the Japanese Pteridophytes. In: Namegata T (ed) Collection and 
Takahashi N, Imaichi R (2007) Developmental morphology of young gametophytes of Botrychium microphyllum in axenic culture. J Jpn Women's Univ Fac Sci 15:45-49
Takamiya M (1989) Cytological and ecological studies on the speciation of Lycopodium clavatum L. in the Japanese Archipelago. J Sci Hiroshima Univ B2 22:353-430
Takamiya M (1996) Index to chromosomes of Japanese Pteridophyta (1910-1996). Japan Pteridological Society, Tokyo
Takamiya M, Ohta N (2001) Cytological and reproductive studies of Japanese Diplazium (Woodsiaceae; Pteridophyta). III. The cytological complexity of species groups with simply pinnate to bipinnatifid leaves. J Plant Res 114:443-457
Takamiya M, Watanabe M, Ono K (1997) Biosystematic studies on the genus Isoetes (Isoetaceae) in Japan IV. Morphology and anatomy of sporophytes, phytogeograohy and taxonomy. Acta Phytotax Geobot 48:89-122
Takatsuki S (2009) Effects of sika deer on vegetation in Japan: A review. Biol Conserv 142:1922-1929
Tanaka T, Isaka Y, Hattori M, Sato T (2014) Ecological and phylogenetic approaches for diversification of apogamous ferns in Japan. Plant Syst Evol 300: 2041-2050 
Terada Y, Takamiya M (2006) Cytological and genetic study of two putative hybrids and their parents of Athyrium (Woodsiaceae; Pteridophyta) in Tsai JL, Shieh WC (1985) A cytotaxonomic survey of the fern family Aspidiaceae (sensu Copeland) in Taiwan. J Sci Engin 22:121-144
Yakushima Island, southwestern Japan. Acta Phytotax Geobot 57:95-106
Thunberg CP (1786) Flora Japonica sistens Plantas Insularum Japonicarum. In Bibliopolio I. G. Muelleriano, Leipzig
Ueno K, Ueno Y, Ebihara A (2019) Dennstaedtia punctilobula (Dennstaedtiaceae), newly naturalized in Nagano Prefecture, Japan. J Jpn Bot 94:58-60
Wagner WH Jr (1954) Reticulate evolution in the Appalachian Aspleniums. Evolution 8:103-118
Wang L, Qi XP, Xiang QP, Heinrichs J, Schneider H, Zhang XC (2010) Phylogeny of the paleotropical fern genus Lepisorus (Polypodiaceae, Polypodiopsida) inferred from four chloroplast DNA regions. Molec Phylogen Evol 54:211-225
Wei R, Schneider H, Zhang XC (2013) Toward a new circumscription of the twinsorus-fern genus Diplazium (Athyriaceae): A molecular phylogeny with morphological implications and infrageneric taxonomy. Taxon 62:441-457
Wu ZY, Raven PH, Hong D (2013) Flora of China, Volumes 2 - 3, Lycopodiaceae through Polypodiaceae. Science Press & Missouri Botanical Garden Press
Xu KW, Zhang L, Lu NT, Zhou XM, He H, Luong TT, Knapp R, Liao WB, Zhang LB (2018) Nine new species of Hymenasplenium (Aspleniaceae) from Asia. Phytotaxa 358:1-25
Yatabe Y, Tsutsumi C, Hirayama Y, Mori K, Murakami N, Kato M (2009) Genetic population structure of Osmunda japonica, rheophilous Osmunda lancea and their hybrids. J Plant Res 122:585-595
Yatabe Y, Yamamoto K, Tsutsumi C, Shinohara W, Murakami N, Kato M (2011) Fertility and precocity of Osmunda × intermedia offspring in culture. J  Plant Res 124:265-268
Yoroi R (1972) Studies on spore germination and gametophyte of Japanese Hymenophyllaceae. Sci Rep Tokyo Kyoiku Daigaku B 225:81-110
Yoroi R (1975) Spore germination and gametophyte development of the Vittariaceae (1). J Jpn Bot 50:33-43
Yoroi R (1976) Gametophyte and apogamous embryo of Crepidomanes latemarginale from Isl. Ishigaki, Ryukyu. J Jpn Bot 51:257-267
Yoroi R (1992) Gametophyte and Embryo of Microgonium tahitense (Hymenophyllaceae). J Jpn Bot 67:169-176
Zhang L, Lu NT, Zhou XM, Chen DK, Knapp R, Zhou L, Guo L, Luong TT, Sun H, Gao XF, Zhang LB (2019) A plastid phylogeny of the Old World fern genus Leptochilus (Polypodiaceae): Implications for cryptic speciation and progressive colonization from lower to higher latitudes. Molec Phylogen Evol 134:311-322
Zhou S, Dong W, Chen X, Zhang X, Wen J, Schneider H (2014) How many species of bracken (Pteridium) are there? Assessing the Chinese brackens using molecular evidence. Taxon 63:509-521
Fig. 1. Maps indicating fern and lycophyte diversity in Japan on 10-km grid maps. a. Richness of native fern and lycophyte taxa. b. Phylogenetic diversity of fern and lycophyte taxa on an unrooted Bayesian tree. c. Phylogenetic diversity of fern taxa on an Bayesian tree rooted by a lycophyte (Lycopodium clavatum var. nipponicum). d. Richness of apomictic fern taxa. e. Proportion of apomicitic taxa to all the taxa. f. Richness of threatened taxa included in the national red list of Japan.

Electronic supplementary material 1. A list of native fern and lycophyte taxa (species, subspecies and varieties) in Japan accepted in this study. Taxon ID refers that in FernGreenList ver.1.0.1 (http://www.rdplants.org/gl/). Otherwise noted, rbcL GenBank accession numbers are those used in Ebihara et al. (2010). Information on reproductive modes and ploidy levels follow those in Ebihara et al. (2016, 2017), and only the records based on material collected in Japan are accepted. For reproductive mode, irregular meiosis is not considered, 0 = no information, 1 = sexual, 2 = apomictic and 3 = sexual + apomictic.

Electronic supplementary material 2. A list of herbarium specimens of 720 native fern and lycophyte taxa in Japan which the distribution maps in ca. 10-km grids in Ebihara et al. (2016, 2017) based on. Small number of corrections on identification and location after the publication were applied, but findings (e.g. new species and new records) after the publication are not reflected. Taxon ID refers that in FernGreenList ver.1.0.1 (http://www.rdplants.org/gl/). Secondary grid code of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and coordinates of their center points are listed. Herbarium acronyms follow Index Herbariorum (http://sweetgum.nybg.org/science/ih/) but the following abbreviations are used for those not listed in Index Herbariorum; Gifu: Gifu Prefectural Museum; Gunma: Gunma Museum of Natural History; Iwate: Iwate Prefectural Museum; Kobe: Kobe University; Kurashiki: Kurashiki Museum of Natural History; Miyajima: Miyajima Natural Botanical Garden, Hiroshima University, Niigata: Niigata University; Sendai City Wild Plants Garden; Tochigi: Tochigi Prefectural Museum; Yamaguchi: Yamaguchi Prefectural Museum; Yokohama; Yokohama City Children’s Botanical Garden; Yokosuka: Yokosuka City Museum. All the listed specimens were seen by the authors except for a part of Isoetes specimens which are cited from Takamiya et al. (1997).

Electronic supplementary material 3. JpFernKeys ver. 1.0: An interactive key file for identification of all the native fern and lycophyte taxa (species, subspecies and varieties) in Japan plus a few common naturalized species for the software Lucid v3.3 (free edition) available at http://www.lucidcentral.com. The data are excerpts from matrices of morphological characters included in Ebihara (2016, 2017). The measurements were based on each ten fertile herbarium specimens collected in Japan which were selected randomly with a few exceptions for the species bearing large fronds (e.g. tree ferns). For some rare species with less than ten specimens available, those collected outside Japan were used.

### 2019-06-25 manuscript/japan_ferns_diversity_ms.Rmd
Add headers, output to word

@@
 title: "An update and reassessment of fern and lycophyte diversity data in the Japanese Archipelago"
-output: html_document
+output: word_document
 ---
@@
 
-Introduction
+## Introduction
+
 Ferns and lycophytes or seed-free vascular plants, also traditionally known as ‘pteridophytes’ are a group of plants which have been paid strong attention as research material of biology.
@@
 
-Flora
+## Flora
+
 The studies on fern and lycophyte flora of the Japanese Archipelago initiated with 44 species enumerated in Thunberg (1786)’s “Flora Japonica”.
@@
 
-Interspecific hybrids
+## Interspecific hybrids
+
 Hybrid ferns have been widely utilized as material of evolutionary biological studies (Sigel 2016), and recognition of interspecific hybrids is always their starting points.
@@
 
-Occurrence data and regional species richness
+## Occurrence data and regional species richness
+
 Eight volumes of "Illustrations of Pteridophytes of Japan" (Kurata and Nakaike, 1979, 1981, 1983, 1985, 1987, 1990, 1994, 1997) were a pioneer work of a nationwide organism distribution survey by citizens in Japan.
@@
 The third highest grid (202 taxa, grid code 503547) is located in the area covering Nachi-no-taki Falls in southern part of Wakayama Prefecture, southern Honshu.
+
 By adding phylogenetic information, we are able to overview the diversity distribution patterns by a different measure from species number.
@@
 19 taxa of Dryopteris, 16 taxa of Diplazium, 14 taxa of Athyrium).
+
 Our updated occurrence data supported the trend that species with seasonal-green phenology showed wider distribution than evergreen species suggested by Guo et al.
@@
 
-Cytotaxonomic data
+## Cytotaxonomic data
+
 Even today chromosome numbers of ferns and lycophytes which are important clues for knowing the plant’s ploidy level and reproductive mode as well as basic chromosome numbers, which were used as information to infer phylogenetic affinities before the introduction of DNA sequencing.
@@
 
-Threatened species
+## Threatened species
+
 Latest national red list of Japan (Ministry of the Environment, Japan 2019) included more than one third of the native fern and lycophyte taxa (260 in total, 6 EX, 3 EW, 85 CR, 59 EN, 68 VU, 38 NT and 1 DD; see Fig.
@@
 
-Introduced species
+## Introduced species
+
 Compared with flowering plants, number of naturalized species of ferns and lycophytes is quite small in the Japanese Archipelago, and they are almost uninvolved with the invasive species issue.
@@
 
-Reticulate evolution
+## Reticulate evolution
+
 Species complexes or aggregate species exhibiting wide ranges of continuous morphological variation are quite commonly known in ferns and lycophytes (Haufler 1996), and this trend seems to be true for the flora of Japan.
@@
 
-Species complexes involved with apomixis
+## Species complexes involved with apomixis
+
 Further complexity is provided when apomictic reproduction (reviewed in Grusz et al.
@@
 
-Species concept
+## Species concept
+
 It is obvious that recent increase of recoded taxa in Japan is largely owing to transition from morphological species concept to biological or evolutional ones.
@@
 2003]) might be classified into different species if future studies clarify differences in their genomic combinations.
+
 We are often faced with a problem about taxonomic treatment for apomictic races.
@@
 
-Gametophytes
+## Gametophytes
+
 Our knowledge about gametophytes of ferns and lycophytes is considerably limited compared with that in sporophytes, nevertheless Japan is one of the world’s most information-rich areas for fern gametophytes.
@@
 2019b, Masuyama 1975] and Botrychium (Ophioglossaceae) [Takahashi and Imaichi 2007]).
+
 Ebihara et al.
@@
 
-Taxonomic key characters
+## Taxonomic key characters
+
 Ebihara (2016, 2017) publicized matrices for major morphological characters frequently used for identification keys for all the 721 native ferns and lycophytes in Japan.
@@
 
-Future perspective
+## Future perspective
+
 Albeit none of the genus accepted in PPGI (Pteridophyte Phylogeny Group 2016) is endemic to the archipelago, recent studies revealed several infrageneric clades distributed centered in and/or representing the highest diversity there.
@@
 Clarification of their speciation and biogeographic histories is expected to contribute for understanding flora formation in East Asia and the Far East.
+
 Regional flora should never conclude in itself but be compared with those of neighboring regions toward a global view.
@@
 2019].
+
 Contrasting to a consensus on familial classification of ferns and lycophytes almost achieved in PPGI (Pteridophyte Phylogeny Group 2016), generic classification proposed in the system requires a good deal of update and emendation based on newly generated evidences.
@@
 It is important to publicize data on East Asian material with accurate identification working as bases for generic recircumscription in forthcoming “PPGII”.
+
 Accumulation of data on ploidy and reproductive modes along with recognition of proper taxonomic units and their distribution data enables us to discuss about various characteristics of ferns and lycophytes as an organism group such as a proportion of polyploids and a proportion of apomict.
@@
 References
+
 Adjie B, Masuyama S, Ishikawa H, Watano Y (2007) Independent origins of tetraploid cryptic species in the fern Ceratopteris thalictroides. J Plant Res 120:129-138
@@
 Zhou S, Dong W, Chen X, Zhang X, Wen J, Schneider H (2014) How many species of bracken (Pteridium) are there? Assessing the Chinese brackens using molecular evidence. Taxon 63:509-521
+
 Fig.

### 2019-06-27 reports/japan_pteridos_biodiv.Rmd
Add report on PD

---
title: "Biodiversity of Japanese Pteridophytes"
author: "Joel Nitta"
date: "6/27/2019"
output: 
  bookdown::html_document2:
    fig_caption: yes
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load-data}
loadd(richness_pteridos_map)
loadd(alpha_div_pteridos)
loadd(alpha_div_ferns_north)
loadd(ses_pd_pteridos_map)
loadd(ses_pd_ferns_north_map)
loadd(ses_pd_highlight_fern_north_map)
loadd(occ_data_ferns_north)
loadd(occ_data_ferns)
```

### Richness

Richness of pteridophytes in Japan roughly follows the latitudinal species gradient, with more species in the south (Fig. \@ref(fig:richness-pteridos)).

### Phylogenetic diversity (PD) calculations

Phylogenetic diversity (PD) is the total phylogenetic branch length of all species in a community  (a "community" is defined here as co-occurring taxa within a 1 km^2^ grid cell). Observed PD is highly correlated with species richness (Fig. \@ref(fig:compare-obs-pd-richness)), as has been demonstrated in many other studies, and by itself is not very informative.

To determine whether communities are significantly phylogenetically overdispersed (i.e., diverse) or clustered (i.e., depauperate), we compared observed PD to a null distribution of 999 randomly generated PD values for each community. Null values were generated by randomizing the names on the tips of the tree using `picante::tipShuffle()`, and calculating PD with the randomized tree. The standard effect size (SES) of PD was then calculated by comparing observed PD to the null distribution using the following equation:

SES PD = (observed PD - mean of null PD values) / standard deviation of null PD values

Thus, large positive values of SES of PD indicate phylogenetic overdispersion, and large negative values indicate phylogenetic clustering. Values near 0 are not significantly different from the null distribution.

### SES of PD: all pteridophytes

SES of PD shows a much weaker correlation with species richness than observed PD (Fig. \@ref(fig:compare-ses-pd-richness)). Likewise, spatial distribution of SES of PD differs strongly from species richness (Fig. \@ref(fig:pd-pteridos)). Mountainous areas on the main island of Japan tend to be clustered (blue), and smaller islands in the south (especially the Okinawan islands) and areas on the coast tend to be overdispersed (red).

However, SES of PD values for pteridophytes tend to skew towards clustering overall (median `r median(alpha_div_pteridos$ses_pd, na.rm = TRUE) %>% round(2)`, Fig. \@ref(fig:pd-pteridos-hist)). This could be due to the fact that a large number of primarily tropical fern lineages only occur in the far-southern islands of Japan: `r setdiff(occ_data_ferns$genus, occ_data_ferns_north$genus) %>% length` genera and `r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% length` families (`r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% paste(collapse = ", ")`) do not occur north of 30.1 degrees latitude.

When these lineages are included in the randomization procedure to calculate SES of PD, they likely contribute long branches that bias the northern communities towards clustering. Lycophytes, which occur on long branches, may have a similar effect. Therefore, we also calculated SES of PD for a dataset including only ferns north of 30.1 degrees latitude (from Yakushima island northwards). This includes a more or less contiguous land area that ferns could likely disperse across, unlike the farth-south islands which are separated by multiple straits from the main islands of Japan.

```{r pd-by-el-summary}
# Run linear model of SES of PD vs. elevation in northern ferns dataset
pd_by_el_summary <- lm(ses_pd ~ elevation, alpha_div_ferns_north) %>% summary

# Extract R2 and pvalue for mentioning in text.
pd_by_el_r2 <- pd_by_el_summary[['adj.r.squared']] %>% round(2)

pd_by_el_pval <- pd_by_el_summary %>% tidy %>% filter(term == "elevation") %>% pull(p.value)

```

### SES of PD: ferns on main islands of Japan only

Standard effect size of PD is less skewed in the dataset including only ferns north of 30.1° (median `r median(alpha_div_ferns_north$ses_pd, na.rm = TRUE) %>% round(2)`, Fig. \@ref(fig:pd-ferns-north-hist)). A similar overall pattern is observed in the spatial distribution of SES of PD for ferns north of 30.1° compared to all pteridophytes, but the number of clustered cells is fewer (Fig. \@ref(fig:pd-ferns-north)). This is especially clear when plotting only communities that are significantly overdispersed or clustered (Fig. \@ref(fig:pd-ferns-north-highlight)). Standard effect size of PD decreases significantly with elevation (Fig. \@ref(fig:pd-ferns-north-el), linear model; *R*^2^ = `r pd_by_el_r2`, *P* << 0.05).

## Figures

```{r richness-pteridos, fig.cap = "Richness of native Japanese pteridophyte taxa."}
richness_pteridos_map
```

```{r compare-obs-pd-richness, fig.cap = "Correlation between observed phylogenetic diversity (PD) and species richness, for all native Japanese pteridophyte taxa excluding hybrids."}
ggplot(alpha_div_pteridos, aes(richness, pd_obs)) +
  geom_point(alpha = 0.2) +
  jntools::standard_theme() +
  labs(
    y = "Observed PD",
    x = "Richness"
  )
```

```{r compare-ses-pd-richness, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) vs. species richness, for all native Japanese pteridophyte taxa excluding hybrids."}
ggplot(alpha_div_pteridos, aes(richness, ses_pd)) +
  geom_point(alpha = 0.2) +
  jntools::standard_theme() +
  labs(
    y = "SES of PD",
    x = "Richness"
  )
```

```{r pd-pteridos, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) of native Japanese pteridophyte taxa, excluding hybrids."}
ses_pd_pteridos_map
```

```{r pd-pteridos-hist, fig.cap = "Histogram of standard effect size (SES) of phylogenetic diversity (PD) values of native Japanese pteridophyte taxa, excluding hybrids. Red line shows median."}
ggplot(alpha_div_pteridos, aes(ses_pd)) +
  geom_histogram() +
  geom_vline(xintercept = median(alpha_div_pteridos$ses_pd, na.rm = TRUE), color = "red") +
  labs(x = "SES of PD")
```

```{r pd-ferns-north-hist, fig.cap = "Histogram of standard effect size (SES) of phylogenetic diversity (PD) values of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids. Red line shows median."}
ggplot(alpha_div_ferns_north, aes(ses_pd)) +
  geom_histogram() +
  geom_vline(xintercept = median(alpha_div_ferns_north$ses_pd, na.rm = TRUE), color = "red") +
  labs(x = "SES of PD")
```

```{r pd-ferns-north, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids."}
ses_pd_ferns_north_map
```

```{r pd-ferns-north-highlight, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa excluding hybrids north of 30.1 degrees latitude, only showing cells that are significantly overdispersed (red) or clustered (blue)."}
ses_pd_highlight_fern_north_map
```

```{r pd-ferns-north-el, fig.cap = "Relationship between standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa excluding hybrids north of 30.1 degrees latitude and elevation (m). Blue trendline shows significant linear model fit with confidence interval in pink."}

lm(ses_pd ~ elevation, alpha_div_ferns_north) %>% augment %>%
ggplot(aes(x = elevation)) +
  geom_point(aes(y = ses_pd), alpha = 0.2) +
  geom_ribbon(aes(ymin=.fitted-1.96*.se.fit, ymax=.fitted+1.96*.se.fit), alpha = 0.4, fill = "pink") +
  geom_line(aes(y = .fitted), color = "blue") +
  annotate("text", 
           x = Inf, y = Inf, 
           label = glue::glue("italic(R) ^ 2 == {pd_by_el_r2}"),
           parse = TRUE,
           hjust = 1.2,
           vjust = 1.2) +
  jntools::standard_theme() +
  labs(
    y = "SES of PD",
    x = "Elevation (m)"
  )

```

### 2019-06-27 code/functions.R

@@ -367,12 +367,19 @@ ses_pd <- function (comm, phy, n_reps) {
 #' observed PD, SES of PD, etc.
 #' @param richness Species richness of communities (1km2 grid
 #' cells)
+#' @param all_cells Longitude, latitude, and elevation of all 
+#' grid cells (including those not in `all_pd` or `richness`)
 #'
-#' @return Tibble
-merge_metrics <- function (all_pd, richness) {
+#' @return Tibble including rows for all grid cells in 1km2 grid cells
+#' of Japan dataset. NA (pd values) or 0 (richness) will be entered
+#' for grid cells that weren't in `all_pd` or `richness`.
+merge_metrics <- function (all_pd, richness, all_cells) {
   all_pd %>%
     rename(secondary_grid_code = site) %>%
-    left_join(richness)
+    left_join(richness) %>%
+    select(-latitude, -longitude) %>%
+    right_join(all_cells) %>%
+    mutate(richness = replace_na(richness, 0))
 }
 
 #' Match community data and tree
@@ -424,6 +431,46 @@ match_comm_and_tree <- function (comm, phy, return = c("comm", "tree")) {
 
 # Plotting ----
 
+#' Get the lower, upper, or absolute maximum value
+#' of a variable in a dataframe
+#' 
+#' For setting plotting limits manually
+#'
+#' @param data Dataframe
+#' @param var Name of variable (column) in dataframe
+#' @param digits Number of digits desired in output
+#' @param type Type of limit to calculate: "min" is lower,
+#' "max" is upper, and "abs" is the absolute greatest value.
+#'
+#' @return Number
+#'
+#' @examples
+#' get_limit(mtcars, disp, "max")
+get_limit <- function (data, var, type = c("min", "max", "abs"), digits = 2) {
+  
+  var <- enquo(var)
+  
+  switch(type,
+         
+         max = data %>% 
+           pull(!!var) %>% 
+           max(na.rm = TRUE) %>% 
+           multiply_by(10^digits) %>% ceiling %>% divide_by(10^digits),
+         
+         min = data %>% 
+           pull(!!var) %>% 
+           min(na.rm = TRUE) %>% 
+           multiply_by(10^digits) %>% floor %>% divide_by(10^digits),
+         
+         abs = c(
+           data %>% pull(!!var) %>% max(na.rm = TRUE),
[…]

### 2019-06-27 code/packages.R

@@ -5,12 +5,14 @@ library(readxl)
 library(janitor)
 library(drake)
 library(maps)
-library(viridis)
+# library(viridis)
 library(scales)
 library(jntools)
 library(broom)
 library(gghighlight)
 library(assertr)
+library(scico)
+library(magrittr)
 library(tidyverse)
 
 conflict_prefer("map", "purrr")

### 2019-06-28 reports/japan_pteridos_biodiv.Rmd
Move all plots to report, plot SES of PD for ferns in N and S separately

@@
 author: "Joel Nitta"
-date: "6/27/2019"
+date: "6/28/2019"
 output: 
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide")
 ```
@@
 ```{r load-data}
-loadd(richness_pteridos_map)
+loadd(world_map)
+loadd(occ_data_pteridos)
+loadd(occ_data_ferns_north)
+loadd(occ_data_ferns)
 loadd(alpha_div_pteridos)
 loadd(alpha_div_ferns_north)
-loadd(ses_pd_pteridos_map)
-loadd(ses_pd_ferns_north_map)
-loadd(ses_pd_highlight_fern_north_map)
-loadd(occ_data_ferns_north)
-loadd(occ_data_ferns)
+loadd(alpha_div_ferns_north_south)
 ```
@@
 
+```{r pd-by-rich-summary}
+# Run exponential model of raw PD vs. richness across all pteridophytes and
+# extract R2 and pvalue for mentioning in text.
+
+# Two ways to fit the model:
+# GLM
+glm_sum <- glm(pd_obs + 0.0001 ~ richness, data=alpha_div_pteridos, family=Gamma(link="log")) %>%
+  summary
+
+# Linear model after log transform. This one gives us an R-squared.
+lm_sum <- lm(-log(pd_obs + 0.0001) ~ richness, data = alpha_div_pteridos) %>% summary
+
+# Extract R-squared.
+pd_by_rich_r2 <- lm_sum[['adj.r.squared']] %>% round(2)
+
+```
+
+```{r ses_pd-by-rich-summary}
+# Run linear model of ses PD vs. richness across all pteridophytes and
+# extract R2 and pvalue for mentioning in text.
+
+# Linear model after log transform. This one gives us an R-squared.
+ses_pd_rich_sum <- lm(ses_pd ~ richness, data = alpha_div_pteridos) %>% summary
+
+# Extract R-squared.
+ses_pd_by_rich_r2 <- ses_pd_rich_sum[['adj.r.squared']] %>% round(2)
+
+# P is very very small
+tidy(ses_pd_rich_sum)
+
+```
 ### Phylogenetic diversity (PD) calculations
@@
 Observed PD is highly correlated with species richness (Fig.
-\@ref(fig:compare-obs-pd-richness)), as has been demonstrated in many other studies, and by itself is not very informative.
+\@ref(fig:compare-obs-pd-richness), exponential model; *R*^2^ = `r pd_by_rich_r2`, *P* << 0.05), as has been demonstrated in many other studies, and by itself is not very informative.
 
 To determine whether communities are significantly phylogenetically overdispersed (i.e., diverse) or clustered (i.e., depauperate), we compared observed PD to a null distribution of 999 randomly generated PD values for each community.
-Null values were generated by randomizing the names on the tips of the tree using `picante::tipShuffle()`, and calculating PD with the randomized tree.
+Null values were generated by randomizing the names on the tips of the tree, and calculating PD with the randomized tree.
 The standard effect size (SES) of PD was then calculated by comparing observed PD to the null distribution using the following equation:
 
-SES PD = (observed PD - mean of null PD values) / standard deviation of null PD values
+$PD_{SES} = \frac{(PD_{obs} - mean_{null})}{SD_{null}}$
 
@@
 Values near 0 are not significantly different from the null distribution.
+Significance was assesed using a two-tailed test: observed PD values at the 0.025% lower or upper extremes of the null distribution were considered significant.
 
@@
 SES of PD shows a much weaker correlation with species richness than observed PD (Fig.
-\@ref(fig:compare-ses-pd-richness)).
+\@ref(fig:compare-ses-pd-richness), linear model; *R*^2^ = `r ses_pd_by_rich_r2`, *P* << 0.05).
 Likewise, spatial distribution of SES of PD differs strongly from species richness (Fig.
@@
 Lycophytes, which occur on long branches, may have a similar effect.
-Therefore, we also calculated SES of PD for a dataset including only ferns north of 30.1 degrees latitude (from Yakushima island northwards).
-This includes a more or less contiguous land area that ferns could likely disperse across, unlike the farth-south islands which are separated by multiple straits from the main islands of Japan.
+Therefore, we also calculated SES of PD separately for ferns  north and south of 30.1 degrees latitude (corresponding to the Tokara strait).
 
@@
 # Run linear model of SES of PD vs. elevation in northern ferns dataset
-pd_by_el_summary <- lm(ses_pd ~ elevation, alpha_div_ferns_north) %>% summary
+pd_by_el_summary <- lm(ses_pd ~ elevation, alpha_div_ferns_north_south) %>% summary
 
@@
 
-pd_by_el_pval <- pd_by_el_summary %>% tidy %>% filter(term == "elevation") %>% pull(p.value)
+# P is very very small
+pd_by_el_summary %>% tidy %>% filter(term == "elevation") %>% pull(p.value)
 
@@
 
-### SES of PD: ferns on main islands of Japan only
+### SES of PD: ferns only, calculated separately in N and S
 
-Standard effect size of PD is less skewed in the dataset including only ferns north of 30.1° (median `r median(alpha_div_ferns_north$ses_pd, na.rm = TRUE) %>% round(2)`, Fig.
+Standard effect size of PD is less skewed in the ferns dataset when N and S regions are calculated separately (median `r median(alpha_div_ferns_north_south$ses_pd, na.rm = TRUE) %>% round(2)`, Fig.
 \@ref(fig:pd-ferns-north-hist)).
-A similar overall pattern is observed in the spatial distribution of SES of PD for ferns north of 30.1° compared to all pteridophytes, but the number of clustered cells is fewer (Fig.
-\@ref(fig:pd-ferns-north)).
-This is especially clear when plotting only communities that are significantly overdispersed or clustered (Fig.
-\@ref(fig:pd-ferns-north-highlight)).
+A similar overall pattern is observed in the spatial distribution of SES of PD for all pteridophytes, but the number of clustered cells is fewer (Fig.
+\@ref(fig:pd-ferns-ns)).
+
+Significantly clustered areas include Aomori prefecture near Misawa, Yamaguchi prefecture, and many other mountainous areas; significantly overdispersed areas include the southern tip of Kyushu, Wakayama prefecture near Tanabe, and the Ogasawara islands (Fig.
+\@ref(fig:pd-ferns-ns-highlight)).
 Standard effect size of PD decreases significantly with elevation (Fig.
-\@ref(fig:pd-ferns-north-el), linear model; *R*^2^ = `r pd_by_el_r2`, *P* << 0.05).
+\@ref(fig:pd-ferns-ns-el), linear model; *R*^2^ = `r pd_by_el_r2`, *P* << 0.05).
 
@@
 ```{r richness-pteridos, fig.cap = "Richness of native Japanese pteridophyte taxa."}
-richness_pteridos_map
+make_diversity_map(
+    div_data = alpha_div_pteridos, 
+    world_map = world_map, 
+    occ_data = occ_data_pteridos, 
+    div_metric = "richness", 
+    metric_title = "Richness"
+  ) + scale_fill_scico(palette = "bilbao", na.value="white")
 ```
 
-```{r compare-obs-pd-richness, fig.cap = "Correlation between observed phylogenetic diversity (PD) and species richness, for all native Japanese pteridophyte taxa excluding hybrids."}
-ggplot(alpha_div_pteridos, aes(richness, pd_obs)) +
+```{r compare-obs-pd-richness, fig.cap = "Correlation between observed phylogenetic diversity (PD) and species richness, for all native Japanese pteridophyte taxa excluding hybrids.
+Blue trendline shows significant exponential model fit with confidence interval in pink."}
+# Shift pd_obs up by an arbitrarily small amount so the exponential curve can be fit.
+
+ggplot(alpha_div_pteridos, aes(richness, pd_obs + 0.0001)) +
   geom_point(alpha = 0.2) +
-  jntools::standard_theme() +
+  geom_smooth(
+    formula = -log(pd_obs + 0.0001) ~ richness,
+    size = 0.5,
+    fill = "pink") +
+  annotate("text", 
+           x = 0, y = Inf, 
+           label = glue::glue("italic(R) ^ 2 == {pd_by_rich_r2}"),
+           parse = TRUE,
+           hjust = 0,
+           vjust = 1.2) +
   labs(
@@
     x = "Richness"
-  )
+  ) +
+  jntools::standard_theme()
+
 ```
 
-```{r compare-ses-pd-richness, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) vs. species richness, for all native Japanese pteridophyte taxa excluding hybrids."}
+```{r compare-ses-pd-richness, fig.cap = "Correlation between standard effect size (SES) of phylogenetic diversity (PD) vs.
+species richness, for all native Japanese pteridophyte taxa excluding hybrids.
+Blue trendline shows significant linear model fit with confidence interval in pink."}
 ggplot(alpha_div_pteridos, aes(richness, ses_pd)) +
   geom_point(alpha = 0.2) +
+  geom_smooth(
+    size = 0.5,
+    fill = "pink",
+    method = "lm") +
+  annotate("text", 
+           x = 0, y = Inf, 
+           label = glue::glue("italic(R) ^ 2 == {ses_pd_by_rich_r2}"),
+           parse = TRUE,
+           hjust = 0,
+           vjust = 1.2) +
   jntools::standard_theme() +
@@
 ```{r pd-pteridos, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) of native Japanese pteridophyte taxa, excluding hybrids."}
-ses_pd_pteridos_map
+make_diversity_map(
+    div_data = alpha_div_pteridos, 
+    world_map = world_map, 
+    occ_data = occ_data_pteridos, 
+    div_metric = "ses_pd", 
+    metric_title = "PD"
+  ) + scale_fill_scico(
+    palette = "vik", 
+    na.value="white",
+    limits = c(
+      -get_limit(alpha_div_pteridos, ses_pd, "abs", 3), 
+      get_limit(alpha_div_pteridos, ses_pd, "abs", 3)
+    ))
 ```
@@
   geom_vline(xintercept = median(alpha_div_pteridos$ses_pd, na.rm = TRUE), color = "red") +
-  labs(x = "SES of PD")
+  labs(x = "SES of PD") +
+  scale_x_continuous(
+    limits = c(
+      -get_limit(alpha_div_pteridos, ses_pd, "abs", 3), 
+      get_limit(alpha_div_pteridos, ses_pd, "abs", 3)
+    )
+  )
 ```
 
-```{r pd-ferns-north-hist, fig.cap = "Histogram of standard effect size (SES) of phylogenetic diversity (PD) values of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids.
+```{r pd-ferns-north-hist, fig.cap = "Histogram of standard effect size (SES) of phylogenetic diversity (PD) values of native Japanese fern taxa calculated separately north and south of 30.1 degrees latitude, excluding hybrids.
 Red line shows median."}
-ggplot(alpha_div_ferns_north, aes(ses_pd)) +
+ggplot(alpha_div_ferns_north_south, aes(ses_pd)) +
   geom_histogram() +
   geom_vline(xintercept = median(alpha_div_ferns_north$ses_pd, na.rm = TRUE), color = "red") +
-  labs(x = "SES of PD")
+  labs(x = "SES of PD") +
+  scale_x_continuous(
+    limits = c(
+      -get_limit(alpha_div_ferns_north_south, ses_pd, "abs", 3), 
+      get_limit(alpha_div_ferns_north_south, ses_pd, "abs", 3)
+    )
+  )
 ```
 
-```{r pd-ferns-north, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids."}
-ses_pd_ferns_north_map
+```{r pd-ferns-ns, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids."}
+make_diversity_map(
+    div_data = alpha_div_ferns_north_south, 
+    world_map = world_map, 
+    occ_data = occ_data_ferns, 
+    div_metric = "ses_pd", 
+    metric_title = "SES of PD"
+  ) + scale_fill_scico(
+    palette = "vik", 
+    na.value="white",
+    limits = c(
+      -get_limit(alpha_div_ferns_north_south, ses_pd, "abs", 3), 
+      get_limit(alpha_div_ferns_north_south, ses_pd, "abs", 3)
+    ))
 ```
 
-```{r pd-ferns-north-highlight, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa excluding hybrids north of 30.1 degrees latitude, only showing cells that are significantly overdispersed (red) or clustered (blue)."}
-ses_pd_highlight_fern_north_map
+```{r pd-ferns-ns-highlight, fig.cap = "Standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa calculated separately north and south of 30.1 degrees latitude, excluding hybrids.
+Only showing cells that are significantly overdispersed (red) or clustered (blue)."}
+make_pd_highlight_map(
+    div_data = alpha_div_ferns_north_south, 
+    world_map = world_map, 
+    occ_data = occ_data_ferns
+  ) + scale_fill_scico(
+    palette = "vik", 
+    na.value="white",
+    limits = c(
+      -get_limit(alpha_div_ferns_north, ses_pd, "abs", 3), 
+      get_limit(alpha_div_ferns_north, ses_pd, "abs", 3)
+    ))
 ```
 
-```{r pd-ferns-north-el, fig.cap = "Relationship between standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa excluding hybrids north of 30.1 degrees latitude and elevation (m).
+```{r pd-ferns-ns-el, fig.cap = "Relationship between standard effect size (SES) of phylogenetic diversity (PD) of native Japanese fern taxa calculated separately north and south of 30.1 degrees latitude, excluding hybrids, and elevation (m).
 Blue trendline shows significant linear model fit with confidence interval in pink."}
 
-lm(ses_pd ~ elevation, alpha_div_ferns_north) %>% augment %>%
+lm(ses_pd ~ elevation, alpha_div_ferns_north_south) %>% augment %>%
   ggplot(aes(x = elevation)) +

### 2019-06-28 code/plan.R

@@ -46,6 +46,9 @@ plan <- drake_plan (
   occ_data_ferns_north = occ_data_ferns %>%
     filter(latitude > 30.1),
   
+  occ_data_ferns_south = occ_data_ferns %>%
+    filter(latitude < 30.1),
+  
   # Read in raw phylogenetic tree of all non-hybrid pteridophyte
   # taxa based on rbcL gene.
   japan_pterido_tree_raw = read.nexus("data/PD170708Bayes2.nxs"),
@@ -64,6 +67,11 @@ plan <- drake_plan (
   japan_fern_tree_north = drop.tip(
     japan_pterido_tree, 
     setdiff(japan_pterido_tree$tip.lab, occ_data_ferns_north$taxon_id)
+  ),
+  
+  japan_fern_tree_south = drop.tip(
+    japan_pterido_tree, 
+    setdiff(japan_pterido_tree$tip.lab, occ_data_ferns_south$taxon_id)
   ),
   
   # Basic world map.
@@ -128,6 +136,8 @@ plan <- drake_plan (
   
   richness_ferns_north = make_richness_matrix(occ_data_ferns_north),
   
+  richness_ferns_south = make_richness_matrix(occ_data_ferns_south),
+  
   # Make community matrix (presence/absence of each species in
   # 1km2 grid cells), trim to only species in tree.
   comm_pteridos = make_comm_matrix(occ_data_pteridos) %>% 
@@ -138,6 +148,9 @@ plan <- drake_plan (
   
   comm_ferns_north = make_comm_matrix(occ_data_ferns_north) %>% 
     match_comm_and_tree(japan_fern_tree_north, "comm"),
+  
+  comm_ferns_south = make_comm_matrix(occ_data_ferns_south) %>% 
+    match_comm_and_tree(japan_fern_tree_south, "comm"),
   
   ### Calculate phylogenetic diversity
   #
@@ -172,6 +185,13 @@ plan <- drake_plan (
     transform = map(i = !!seq_len(4))
   ),
   
+  comm_ferns_south_split = target(
+    drake_slice(column_to_rownames(comm_ferns_south, "species"), 
+                slices = 4, index = i, margin = 2),
+    transform = map(i = !!seq_len(4))
+  ),
+  
+  
   # Caclulate standard effect size of Faith's phylogenetic
   # diversity on each slice of the dataset.
   pd_pteridos = target(
@@ -198,6 +218,14 @@ plan <- drake_plan (
[…]

### 2019-06-28 make.R

@@ -29,13 +29,13 @@ source("code/plan.R")
 make(plan) 
 
 # OR, uncomment out the below lines to run in parallel.
-
-# future::plan(future::multiprocess) 
+# 
+# future::plan(future::multiprocess)
 # 
 # tictoc::tic()
 # make(
-#   plan, 
-#   parallelism = "future", 
+#   plan,
+#   parallelism = "future",
 #   jobs = 4,
 #   prework = list(
 #     quote(conflict_prefer("map", "purrr")),

### 2019-06-28 reports/japan_pteridos_biodiv.Rmd
Simplify extraction of r2 with glance()

@@
 # GLM
-glm_sum <- glm(pd_obs + 0.0001 ~ richness, data=alpha_div_pteridos, family=Gamma(link="log")) %>%
-  summary
+pd_by_rich_glm_mod <- glm(pd_obs + 0.0001 ~ richness, data=alpha_div_pteridos, family=Gamma(link="log"))
+summary(pd_by_rich_glm_mod)
 
 # Linear model after log transform. This one gives us an R-squared.
-lm_sum <- lm(-log(pd_obs + 0.0001) ~ richness, data = alpha_div_pteridos) %>% summary
+pd_by_rich_lm_mod <- lm(-log(pd_obs + 0.0001) ~ richness, data = alpha_div_pteridos)
 
 # Extract R-squared.
-pd_by_rich_r2 <- lm_sum[['adj.r.squared']] %>% round(2)
+pd_by_rich_r2 <- pd_by_rich_lm_mod %>% glance %>% pull(r.squared) %>% round(3)
 
+# P is very very small, effectively 0
+pd_by_rich_lm_mod %>% glance
 ```
@@
 # extract R2 and pvalue for mentioning in text.
-
-# Linear model after log transform. This one gives us an R-squared.
-ses_pd_rich_sum <- lm(ses_pd ~ richness, data = alpha_div_pteridos) %>% summary
+ses_pd_rich_mod <- lm(ses_pd ~ richness, data = alpha_div_pteridos)
 
 # Extract R-squared.
-ses_pd_by_rich_r2 <- ses_pd_rich_sum[['adj.r.squared']] %>% round(2)
+ses_pd_by_rich_r2 <- ses_pd_rich_mod %>% glance %>% pull(r.squared) %>% round(3)
 
 # P is very very small
-tidy(ses_pd_rich_sum)
+ses_pd_rich_mod %>% glance
+```
 
-```
 ### Phylogenetic diversity (PD) calculations
@@
 # Run linear model of SES of PD vs. elevation in northern ferns dataset
-pd_by_el_summary <- lm(ses_pd ~ elevation, alpha_div_ferns_north_south) %>% summary
+pd_by_el_mod<- lm(ses_pd ~ elevation, alpha_div_ferns_north_south)
 
 # Extract R2 and pvalue for mentioning in text.
-pd_by_el_r2 <- pd_by_el_summary[['adj.r.squared']] %>% round(2)
+pd_by_el_r2 <- pd_by_el_mod %>% glance %>% pull(r.squared) %>% round(3)
 
 # P is very very small
-pd_by_el_summary %>% tidy %>% filter(term == "elevation") %>% pull(p.value)
-
+pd_by_el_mod %>% glance
 ```
@@
 \@ref(fig:pd-ferns-ns-el), linear model; *R*^2^ = `r pd_by_el_r2`, *P* << 0.05).
+
+### Breeding system
+
+```{r ses-pd-vs-perc-sex-dip-summary}
+# Run linear model of SES of PD vs. percentage of sexual diploid ferns
+ses_pd_vs_repro_ferns_mod <- lm(ses_pd ~ percent_sex_dip, ses_pd_vs_repro_ferns)
+
+# Extract R2 and pvalue for mentioning in text.
+ses_pd_vs_repro_ferns_r2 <- ses_pd_vs_repro_ferns_mod %>% glance %>% pull(r.squared) %>% round(3)
+
+# P is very very small
+ses_pd_vs_repro_ferns_mod %>% glance
+```
+
+Sexual diploid taxa tend to be distributed in northern latitudes (Fig. \@ref(fig:sex-dip-ferns-map)).
+
+Somewhat surprisingly, SES of PD shows a negative relationship with the percentage of sexual diploid taxa (Fig.
+\@ref(fig:ses-pd-vs-perc-sex-dip); *R*^2^ = `r ses_pd_vs_repro_ferns_r2`, *P* << 0.05).
+We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed.
+It seems that communities with many non-sexual taxa actually tend to include distantly related taxa rather than sister pairs of sexual and non-sexual species.
 

### 2019-06-28 reports/japan_pteridos_biodiv.Rmd
Add analysis of breeding system

@@
 loadd(alpha_div_ferns_north_south)
+loadd(percent_sex_dip_ferns)
+loadd(ses_pd_vs_repro_ferns)
 ```
@@
 ```
+
+```{r sex-dip-ferns-map, fig.cap = "Percentage of sexual diploid fern taxa, excluding hybrids."}
+make_diversity_map(
+  div_data = percent_sex_dip_ferns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "percent_sex_dip",
+  metric_title = "% sex. diploid"
+) + scale_fill_scico(palette = "bilbao", na.value="white")
+```
+
+```{r ses-pd-vs-perc-sex-dip, fig.cap = "Correlation between percentage of sexual diploid fern taxa (excluding hybrids) and standard effect size (SES) of phylogenetic diversity (PD)."}
+ggplot(ses_pd_vs_repro_ferns, aes(percent_sex_dip, ses_pd)) +
+  geom_point(alpha = 0.4) +
+  geom_smooth(
+    method = "lm",
+    size = 0.5,
+    fill = "pink"
+  ) +
+  annotate("text", 
+           x = Inf, y = Inf, 
+           label = glue::glue("italic(R) ^ 2 == {ses_pd_vs_repro_ferns_r2}"),
+           parse = TRUE,
+           hjust = 1.2,
+           vjust = 1.2) +
+  labs(
+    x = "% sexual diploid taxa",
+    y = "SES PD"
+  ) +
+  jntools::standard_theme()
+```

### 2019-06-28 code/plan.R

@@ -262,6 +262,27 @@ plan <- drake_plan (
     filter(alpha_div_ferns_south, secondary_grid_code %in% all_pd_ferns_south$site)
   ),
   
+  # Analyze percentage of sexual diploid ferns.
+  percent_sex_dip_ferns =
+    right_join(
+      occ_data_ferns, 
+      select(repro_data, -taxon_name),
+      by = "taxon_id") %>%
+    group_by(secondary_grid_code, latitude, longitude) %>%
+    summarize(
+      num_sex_dip = sum(sexual_diploid),
+      num_total = n()
+    ) %>%
+    ungroup %>%
+    mutate(percent_sex_dip = num_sex_dip / num_total),
+  
+  # Compare SES PD vs. percentage of sexual diploid ferns.
+  ses_pd_vs_repro_ferns =
+    left_join(
+      select(alpha_div_ferns_north_south, secondary_grid_code, latitude, longitude, ses_pd),
+      select(percent_sex_dip_ferns, secondary_grid_code, percent_sex_dip)
+    ),
+  
   # Write out report ----
   report = rmarkdown::render(
     knitr_in(here::here("reports/japan_pteridos_biodiv.Rmd")),

### 2019-06-30 reports/japan_pteridos_biodiv.Rmd
Use rbind instead of bind_rows for dataframes
bind_rows drops rownames

--- 2019-06-28 reports/japan_pteridos_biodiv.Rmd
@@
 loadd(alpha_div_ferns_north_south)
+loadd(percent_sex_dip_ferns)
+loadd(alpha_div_vs_repro_ferns)
 ```
@@
 # Run linear model of SES of PD vs. percentage of sexual diploid ferns
-ses_pd_vs_repro_ferns_mod <- lm(ses_pd ~ percent_sex_dip, ses_pd_vs_repro_ferns)
+alpha_div_vs_repro_ferns_mod <- lm(ses_pd ~ percent_sex_dip, alpha_div_vs_repro_ferns)
 
 # Extract R2 and pvalue for mentioning in text.
-ses_pd_vs_repro_ferns_r2 <- ses_pd_vs_repro_ferns_mod %>% glance %>% pull(r.squared) %>% round(3)
+alpha_div_vs_repro_ferns_r2 <- alpha_div_vs_repro_ferns_mod %>% glance %>% pull(r.squared) %>% round(3)
 
 # P is very very small
-ses_pd_vs_repro_ferns_mod %>% glance
+alpha_div_vs_repro_ferns_mod %>% glance
 ```
@@
 Somewhat surprisingly, SES of PD shows a negative relationship with the percentage of sexual diploid taxa (Fig.
-\@ref(fig:ses-pd-vs-perc-sex-dip); *R*^2^ = `r ses_pd_vs_repro_ferns_r2`, *P* << 0.05).
+\@ref(fig:ses-pd-vs-perc-sex-dip); *R*^2^ = `r alpha_div_vs_repro_ferns_r2`, *P* << 0.05).
 We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed.
@@
 ```
+
+```{r sex-dip-ferns-map, fig.cap = "Percentage of sexual diploid fern taxa, excluding hybrids."}
+make_diversity_map(
+  div_data = percent_sex_dip_ferns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "percent_sex_dip",
+  metric_title = "% sex. diploid"
+) + scale_fill_scico(palette = "bilbao", na.value="white")
+```
+
+```{r ses-pd-vs-perc-sex-dip, fig.cap = "Correlation between percentage of sexual diploid fern taxa (excluding hybrids) and standard effect size (SES) of phylogenetic diversity (PD)."}
+ggplot(alpha_div_vs_repro_ferns, aes(percent_sex_dip, ses_pd)) +
+  geom_point(alpha = 0.4) +
+  geom_smooth(
+    method = "lm",
+    size = 0.5,
+    fill = "pink"
+  ) +
+  annotate("text", 
+           x = Inf, y = Inf, 
+           label = glue::glue("italic(R) ^ 2 == {alpha_div_vs_repro_ferns_r2}"),
+           parse = TRUE,
+           hjust = 1.2,
+           vjust = 1.2) +
+  labs(
+    x = "% sexual diploid taxa",
+    y = "SES PD"
+  ) +
+  jntools::standard_theme()
+```
+
+```{r mntd-ferns-ns, fig.cap = "Standard effect size (SES) of MNTD of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids."}
+make_diversity_map(
+    div_data = alpha_div_ferns_north_south, 
+    world_map = world_map, 
+    occ_data = occ_data_ferns, 
+    div_metric = "mntd.obs.z", 
+    metric_title = "MNTD"
+  ) + scale_fill_scico(
+    palette = "vik", 
+    na.value="white",
+    limits = c(
+      -get_limit(alpha_div_ferns_north_south, mntd.obs.z, "abs", 3), 
+      get_limit(alpha_div_ferns_north_south, mntd.obs.z, "abs", 3)
+    ))
+```
+
+```{r mpd-ferns-ns, fig.cap = "Standard effect size (SES) of MPD of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids."}
+make_diversity_map(
+    div_data = alpha_div_ferns_north_south, 
+    world_map = world_map, 
+    occ_data = occ_data_ferns, 
+    div_metric = "mpd.obs.z", 
+    metric_title = "mpd"
+  ) + scale_fill_scico(
+    palette = "vik", 
+    na.value="white",
+    limits = c(
+      -get_limit(alpha_div_ferns_north_south, mpd.obs.z, "abs", 3), 
+      get_limit(alpha_div_ferns_north_south, mpd.obs.z, "abs", 3)
+    ))
+```
+
+```{r mntd-vs-perc-sex-dip, fig.cap = "Correlation between percentage of sexual diploid fern taxa (excluding hybrids) and MNTD."}
+
+# Run linear model of SES of PD vs. percentage of sexual diploid ferns
+mtnd_vs_percent_sex_mod <- lm(mntd.obs.z ~ percent_sex_dip, alpha_div_vs_repro_ferns)
+
+# Extract R2 and pvalue for mentioning in text.
+mtnd_vs_percent_sex_mod_r2 <- mtnd_vs_percent_sex_mod %>% glance %>% pull(r.squared) %>% round(3)
+
+ggplot(alpha_div_vs_repro_ferns, aes(percent_sex_dip, mntd.obs.z)) +
+  geom_point(alpha = 0.4) +
+  geom_smooth(
+    method = "lm",
+    size = 0.5,
+    fill = "pink"
+  ) +
+  annotate("text",
+           x = Inf, y = Inf,
+           label = glue::glue("italic(R) ^ 2 == {mtnd_vs_percent_sex_mod_r2}"),
+           parse = TRUE,
+           hjust = 1.2,
+           vjust = 1.2) +
+  labs(
+    x = "% sexual diploid taxa",
+    y = "mntd.obs.z"
+  ) +
+  jntools::standard_theme()
+```
+
+```{r mntd-vs-el, fig.cap = "Correlation between elevation and MNTD."}
+
+# Run linear model of SES of PD vs. percentage of sexual diploid ferns
+mtnd_vs_el_mod <- lm(mntd.obs.z ~ elevation, alpha_div_vs_repro_ferns)
+
+# Extract R2 and pvalue for mentioning in text.
+mtnd_vs_el_r2 <- mtnd_vs_el_mod %>% glance %>% pull(r.squared) %>% round(3)
+
+ggplot(alpha_div_vs_repro_ferns, aes(elevation, mntd.obs.z)) +
+  geom_point(alpha = 0.4) +
+  geom_smooth(
+    method = "lm",
+    size = 0.5,
+    fill = "pink"
+  ) +
+  annotate("text",
+           x = Inf, y = Inf,
+           label = glue::glue("italic(R) ^ 2 == {mtnd_vs_el_r2}"),
+           parse = TRUE,
+           hjust = 1.2,
+           vjust = 1.2) +
+  labs(
+    x = "Elevation (m)",
+    y = "mntd.obs.z"
+  ) +
+  jntools::standard_theme()
+```
+
+```{r mpd-vs-el, fig.cap = "Correlation between elevation and mpd."}
+
+# Run linear model of SES of PD vs. percentage of sexual diploid ferns
+mtnd_vs_el_mod <- lm(mpd.obs.z ~ elevation, alpha_div_vs_repro_ferns)
+
+# Extract R2 and pvalue for mentioning in text.
+mtnd_vs_el_r2 <- mtnd_vs_el_mod %>% glance %>% pull(r.squared) %>% round(3)
+
+ggplot(alpha_div_vs_repro_ferns, aes(elevation, mpd.obs.z)) +
+  geom_point(alpha = 0.4) +
+  geom_smooth(
+    method = "lm",
+    size = 0.5,
+    fill = "pink"
+  ) +
+  annotate("text",
+           x = Inf, y = Inf,
+           label = glue::glue("italic(R) ^ 2 == {mtnd_vs_el_r2}"),
+           parse = TRUE,
+           hjust = 1.2,
+           vjust = 1.2) +
+  labs(
+    x = "Elevation (m)",
+    y = "mpd.obs.z"
+  ) +
+  jntools::standard_theme()
+```

### 2019-06-30 code/plan.R

@@ -286,22 +286,22 @@ plan <- drake_plan (
   ),
   
   all_mpd_ferns_north = target(
-    bind_rows(mpd_ferns_north),
+    rbind(mpd_ferns_north),
     transform = combine(mpd_ferns_north)
   ),
   
   all_mpd_ferns_south = target(
-    bind_rows(mpd_ferns_south),
+    rbind(mpd_ferns_south),
     transform = combine(mpd_ferns_south)
   ),
   
   all_mntd_ferns_north = target(
-    bind_rows(mntd_ferns_north),
+    rbind(mntd_ferns_north),
     transform = combine(mntd_ferns_north)
   ),
   
   all_mntd_ferns_south = target(
-    bind_rows(mntd_ferns_south),
+    rbind(mntd_ferns_south),
     transform = combine(mntd_ferns_south)
   ),
   
@@ -311,9 +311,11 @@ plan <- drake_plan (
   alpha_div_pteridos = merge_metrics(all_pd_pteridos, richness_pteridos, all_cells),
   alpha_div_ferns = merge_metrics(all_pd_ferns, richness_ferns, all_cells),
   alpha_div_ferns_north = merge_metrics(all_pd_ferns_north, richness_ferns_north, all_cells) %>%
-    left_join(rownames_to_column(all_mntd_ferns_north, "secondary_grid_code")),
+    left_join(rownames_to_column(all_mntd_ferns_north, "secondary_grid_code") %>% select(-ntaxa, -runs)) %>%
+    left_join(rownames_to_column(all_mpd_ferns_north, "secondary_grid_code") %>% select(-ntaxa, -runs)),
   alpha_div_ferns_south = merge_metrics(all_pd_ferns_south, richness_ferns_south, all_cells)  %>%
-    left_join(rownames_to_column(all_mntd_ferns_south, "secondary_grid_code")),
+    left_join(rownames_to_column(all_mntd_ferns_south, "secondary_grid_code") %>% select(-ntaxa, -runs)) %>%
+    left_join(rownames_to_column(all_mpd_ferns_south, "secondary_grid_code") %>% select(-ntaxa, -runs)),
   
   # Combine alpha diversity for SES of PD measured
   # in North and South regions separately
@@ -337,9 +339,9 @@ plan <- drake_plan (
     mutate(percent_sex_dip = num_sex_dip / num_total),
   
   # Compare SES PD vs. percentage of sexual diploid ferns.
[…]

### 2019-07-02 ms/japan_pteridos_biodiv.Rmd
Move manuscript rmd to ms/

---
title: "Biodiversity of Japanese Pteridophytes"
author: "Joel Nitta, Atushi Ebihara"
bibliography: "references.bib"
output: 
  bookdown::html_document2:
    fig_caption: yes
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide")
```

```{r load-data}
loadd(list = c("world_map",
  "occ_data_pteridos",
  "occ_data_ferns_north",
  "occ_data_ferns",
  "alpha_div_pteridos",
  "alpha_div_ferns",
  "alpha_div_ferns_ns",
  "alpha_div_vs_repro_ferns"), 
  cache = ja_fern_cache)
```

## Methods

### Phylogenetic diversity

We measured phylogenetic diversity using mean phylogenetic distance (MPD), the mean distance between all co-occurring taxa in a community, and mean nearest taxon distance (MNTD), the mean distance between co-occurring sister taxa in a community [@Webb2000; @Webb2002]. A "community" is defined here as co-occurring taxa within a 1 km^2^ grid cell. Since observed values of MPD and MNTD are known to correlate with species richness, we determined whether communities are phylogenetically overdispersed (i.e., diverse) or clustered (i.e., depauperate), by calculating the standard effect size (SES) of MPD and MNTD. We compared observed values to those of a null distribution of 999 random communities with the `ses.pd` function in the `picante` R package [@Kembel2010]. Random communities were each generated using 10,000 iterations of the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities. Positive values of SES of MPD and MNTD (hereafter referred to simply as "MPD" or "MTND" respectively unless indicated otherwise) indicate phylogenetic overdispersion, and negative values indicate phylogenetic clustering. Values at the 2.5% lower or upper extremes of the null distribution were considered statistically significant.

### Datasets

A large number of primarily tropical fern lineages only occur in the far-southern islands of Japan: `r setdiff(occ_data_ferns$genus, occ_data_ferns_north$genus) %>% length` genera and `r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% length` families (`r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% paste(collapse = ", ")`) do not occur north of 30.1 degrees latitude.

When these lineages are included in the randomization procedure to calculate SES of MPD and MNTD, they likely contribute long branches that bias the northern communities towards clustering. Lycophytes, which occur on long branches, may have a similar effect. Therefore, we calculated phylogenetic diversity for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns separately north and south of 30.1 degrees latitude (corresponding to the Tokara strait, hereafter referred to as "ferns_ns").

## Results

### Richness

Richness of pteridophytes in Japan roughly follows the latitudinal species gradient, with more species in the south (Fig. \@ref(fig:richness-pteridos)).

### Phylogenetic diversity

Standard effect sizes of mean phylogenetic diversity (MPD) and mean nearest taxon distance (MNTD) were either not correlated or only very weakly correlated with species richness (SI Fig. \@ref(fig:compare-pd-richness)), indicating that our null models successfully decoupled measures of phylogenetic diversity and richness. The distribution of MPD and MNTD values were very similar across datasets (SI Fig. \@ref(fig:alpha-div-ridge)).

Values of MPD are clustered in mountainous areas and overdispersed along the coasts of the main islands of Japan, with similar patterns between pteridophytes and ferns (Fig. \@ref(fig:mpd)a, b). The pattern is similar in the "ferns_NS" dataset, except that the strength of both clustering and overdispersion is weaker in the main islands of Japan, and the pattern of overdispresion on the islands of Amami-oshima in sourthern Japan reverses to clustering (Fig. \@ref(fig:mpd)c).

The signal of clustering is much stronger when measured with MNTD, especially in mountainous areas of central Honshu (Fig. \@ref(fig:mntd)). Patterns are similar between all three datasets except that MNTD values are smaller in the "ferns_NS" dataset (Fig. \@ref(fig:mntd)c).

Values of MPD and MNTD decrease significantly with elevation across all datasets (Fig. \@ref(fig:compare-pd-el), linear model).

### Breeding system

Sexual diploid taxa tend to be distributed in northern latitudes (Fig. \@ref(fig:sex-dip-ferns-map)).

Somewhat surprisingly, MPD and MNTD show a weak negative relationship with the percentage of sexual diploid taxa (Fig. \@ref(fig:compare-pd-sexdip); linear model). We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed. It seems that communities with many non-sexual taxa actually tend to include distantly related taxa rather than sister pairs of sexual and non-sexual species.

## Figures

```{r richness-pteridos, fig.cap = "Richness of native Japanese pteridophyte taxa."}
make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "richness", 
    metric_title = "Richness"
  ) + scale_fill_scico(palette = "bilbao", na.value="white")
```

```{r compare-pd-richness, fig.cap = "Correlation between phylogenetic diversity and richness. Phylogenetic diversity measured as standard effect size (SES) of mean phylogenetic diversity (MPD) or mean nearest taxon distance (MNTD) in native pteridophytes, ferns, or ferns calculated separately north and south of 30.1 latitude of Japan. Blue trendline shows significant linear model fit (*P* < 0.05) with 96% confidence interval in pink."}
a <- make_scatter_with_lm(alpha_div_pteridos, "richness", "mpd_obs_z", "Richness", "MPD") +
  labs(title = "Pteridophytes", x = "")
b <- make_scatter_with_lm(alpha_div_ferns, "richness", "mpd_obs_z", "Richness", "MPD") +
  labs(title = "Ferns", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "mpd_obs_z", "Richness", "MPD") +
  labs(title = "Ferns (NS separate)", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_pteridos, "richness", "mntd_obs_z", "Richness", "MNTD")
e <- make_scatter_with_lm(alpha_div_ferns, "richness", "mntd_obs_z", "Richness", "MNTD") +
  labs(y = "")
f <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "mntd_obs_z", "Richness", "MNTD") +
  labs(y = "")
a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
```

```{r alpha-div-ridge, fig.cap = "Ridge plots showing distribution of the standard effect size of a) mean phylogenetic distance (MPD) and b) mean nearest taxon distance (MNTD) for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns when the data were split into northern and southern portions and calculated separately. Black vertical lines shows median."}
# Combine mpd and mntd calculations from the three datasets for comparison
comp_data <- 
  bind_rows(
  transmute(alpha_div_pteridos, dataset = "pteridos", mpd_obs_z, mntd_obs_z),
  transmute(alpha_div_ferns, dataset = "ferns", mpd_obs_z, mntd_obs_z),
  transmute(alpha_div_ferns_ns, dataset = "ferns_ns", mpd_obs_z, mntd_obs_z)
  )

# MPD
a <- ggplot(comp_data, aes(x = mpd_obs_z, y = dataset, fill = dataset)) + geom_density_ridges(
  scale = 1,
  quantile_lines = TRUE, quantiles = 2)  +
  labs(
    x = "MPD",
    y = "Dataset"
  ) +
  jntools::standard_theme() +
  theme(legend.position = "none")

# MNTD
b <- ggplot(comp_data, aes(x = mntd_obs_z, y = dataset, fill = dataset)) + geom_density_ridges(
  scale = 1,
  quantile_lines = TRUE, quantiles = 2) +
  theme(legend.position = "none") +
   labs(
    x = "MNTD",
    y = "Dataset"
  ) +
  jntools::standard_theme() +
  theme(legend.position = "none")

# final plot
a + b + plot_layout(ncol = 1) + plot_annotation(tag_levels = "a")

```

```{r mpd, fig.cap = "Standard effect size (SES) of mean phylogenetic distance (MPD) of native Japanese pteridophytes (ferns and lycophytes), excluding hybrids: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude."}
a <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "mpd_obs_z", 
    metric_title = "MPD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_pteridos, mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, mpd_obs_z, "abs", 3)
    )) +
  theme(legend.position = "bottom")

b <- make_diversity_map(
    div_data = alpha_div_ferns, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "mpd_obs_z", 
    metric_title = "MPD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns, mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, mpd_obs_z, "abs", 3)
    )) +
  theme(legend.position = "bottom")

c <- make_diversity_map(
    div_data = alpha_div_ferns_ns, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "mpd_obs_z", 
    metric_title = "MPD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns_ns, mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns_ns, mpd_obs_z, "abs", 3)
    )) +
  theme(legend.position = "bottom")

a + b + c + plot_layout(nrow = 1) + plot_annotation(tag_levels = "a")
```

```{r mntd, fig.cap = "Standard effect size (SES) of mean nearest taxon distance (MNTD) of native Japanese pteridophytes (ferns and lycophytes), excluding hybrids: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude."}
a <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "mntd_obs_z", 
    metric_title = "mntd"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_pteridos, mntd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, mntd_obs_z, "abs", 3)
    )) +
  theme(legend.position = "none")

b <- make_diversity_map(
    div_data = alpha_div_ferns, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "mntd_obs_z", 
    metric_title = "mntd"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns, mntd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, mntd_obs_z, "abs", 3)
    )) +
  theme(legend.position = "none")

c <- make_diversity_map(
    div_data = alpha_div_ferns_ns, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "mntd_obs_z", 
    metric_title = "mntd"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns_ns, mntd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns_ns, mntd_obs_z, "abs", 3)
    )) +
  theme(legend.position = "none")

a + b + c + plot_layout(nrow = 1)
```

```{r compare-pd-el, fig.cap = "Relationship between standard effect size (SES) of mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) of native Japanese pteridophytes and elevation (m). Blue trendline shows significant linear model fit with confidence interval in pink."}
a <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Pteridophytes", x = "")
b <- make_scatter_with_lm(alpha_div_ferns, "elevation", "mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Ferns", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns_ns, "elevation", "mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Ferns (NS separate)", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "mntd_obs_z", "Elevation (m)", "MNTD")
e <- make_scatter_with_lm(alpha_div_ferns, "elevation", "mntd_obs_z", "Elevation (m)", "MNTD") +
  labs(y = "")
f <- make_scatter_with_lm(alpha_div_ferns_ns, "elevation", "mntd_obs_z", "Elevation (m)", "MNTD") +
  labs(y = "")
a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
```

```{r sex-dip-ferns-map, fig.cap = "Percentage of sexual diploid fern taxa, excluding hybrids."}
make_diversity_map(
  div_data = alpha_div_ferns, 
  world_map = world_map, 
  occ_data = occ_data_ferns, 
  div_metric = "percent_sex_dip",
  metric_title = "% sex. diploid"
) + scale_fill_scico(palette = "bilbao", na.value="white")
```

```{r compare-pd-sexdip, fig.cap = "Relationship between standard effect size (SES) of mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) of native Japanese pteridophytes and percent sexual diploid taxa in each community. Blue trendline shows significant linear model fit with confidence interval in pink."}
a <- make_scatter_with_lm(alpha_div_pteridos, "percent_sex_dip", "mpd_obs_z", "% sexual diploid", "MPD") +
  labs(title = "Pteridophytes", x = "")
b <- make_scatter_with_lm(alpha_div_ferns, "percent_sex_dip", "mpd_obs_z", "% sexual diploid", "MPD") +
  labs(title = "Ferns", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns_ns, "percent_sex_dip", "mpd_obs_z", "% sexual diploid", "MPD") +
  labs(title = "Ferns (NS separate)", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_pteridos, "percent_sex_dip", "mntd_obs_z", "% sexual diploid", "MNTD")
e <- make_scatter_with_lm(alpha_div_ferns, "percent_sex_dip", "mntd_obs_z", "% sexual diploid", "MNTD") +
  labs(y = "")
f <- make_scatter_with_lm(alpha_div_ferns_ns, "percent_sex_dip", "mntd_obs_z", "% sexual diploid", "MNTD") +
  labs(y = "")
a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
```

### 2019-07-02 ms/japan_pteridos_biodiv.Rmd
Add ms to plan

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide")
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)
 ```
 
-```{r load-data}
+```{r load-data, cache = FALSE}
 loadd(list = c("world_map",
@@
 Since observed values of MPD and MNTD are known to correlate with species richness, we determined whether communities are phylogenetically overdispersed (i.e., diverse) or clustered (i.e., depauperate), by calculating the standard effect size (SES) of MPD and MNTD.
-We compared observed values to those of a null distribution of 999 random communities with the `ses.pd` function in the `picante` R package [@Kembel2010].
+We compared observed values to those of a null distribution of 999 random communities with the "ses.pd" function in the "picante" R package [@Kembel2010].
 Random communities were each generated using 10,000 iterations of the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities.
@@
 Lycophytes, which occur on long branches, may have a similar effect.
-Therefore, we calculated phylogenetic diversity for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns separately north and south of 30.1 degrees latitude (corresponding to the Tokara strait, hereafter referred to as "ferns_ns").
+Therefore, we calculated phylogenetic diversity for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns separately north and south of 30.1 degrees latitude (corresponding to the Tokara strait, hereafter referred to as "ferns_NS").
 
@@
 
-```{r richness-pteridos, fig.cap = "Richness of native Japanese pteridophyte taxa."}
+```{r richness-pteridos, fig.cap = "Richness of native Japanese pteridophyte taxa, excluding hybrids."}
 make_diversity_map(
@@
 
-```{r mpd, fig.cap = "Standard effect size (SES) of mean phylogenetic distance (MPD) of native Japanese pteridophytes (ferns and lycophytes), excluding hybrids: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude."}
+```{r mpd, fig.asp = 2.0, fig.cap = "Standard effect size (SES) of mean phylogenetic distance (MPD) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude.
+All datasets are native taxa of Japan excluding hybrids."}
 a <- make_diversity_map(
@@
     )) +
-  theme(legend.position = "bottom")
+  labs(title = "a, pteridophytes") +
+  theme(legend.position = c(0.1, 0.80))
 
@@
     )) +
-  theme(legend.position = "bottom")
+  labs(title = "b, ferns") +
+  theme(legend.position = c(0.1, 0.80))
 
@@
     )) +
-  theme(legend.position = "bottom")
+  labs(title = "c, ferns (NS separate)") +
+  theme(legend.position = c(0.1, 0.80))
 
-a + b + c + plot_layout(nrow = 1) + plot_annotation(tag_levels = "a")
+a + b + c + plot_layout(ncol = 1)
 ```
 
-```{r mntd, fig.cap = "Standard effect size (SES) of mean nearest taxon distance (MNTD) of native Japanese pteridophytes (ferns and lycophytes), excluding hybrids: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude."}
+```{r mntd, fig.asp = 2.0, fig.cap = "Standard effect size (SES) of mean nearest taxon distance (MNTD) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude.
+All datasets are native taxa of Japan excluding hybrids."}
 a <- make_diversity_map(
@@
     div_metric = "mntd_obs_z", 
-    metric_title = "mntd"
+    metric_title = "MNTD"
   ) + scale_fill_scico(
@@
     )) +
-  theme(legend.position = "none")
+  labs(title = "a, pteridophytes") +
+  theme(legend.position = c(0.1, 0.80))
 
@@
     )) +
-  theme(legend.position = "none")
+  labs(title = "b, ferns") +
+  theme(legend.position = c(0.1, 0.80))
 
@@
     )) +
-  theme(legend.position = "none")
+  labs(title = "c, ferns (NS separate)") +
+  theme(legend.position = c(0.1, 0.80))
 
-a + b + c + plot_layout(nrow = 1)
+a + b + c + plot_layout(ncol = 1)
 ```
@@
 ```
+
+## References

### 2019-07-02 code/plan.R

@@ -288,11 +288,11 @@ plan <- drake_plan (
                 clean_names %>%
                 select(-ntaxa, -runs)) %>%
     left_join(select(percent_sex_dip_pteridos, secondary_grid_code, percent_sex_dip)) %>%
-    mutate(richness = replace_na(richness, 0))
+    mutate(richness = replace_na(richness, 0)),
   
   # Write out manuscript ----
   ms = rmarkdown::render(
-    knitr_in(here::here("reports/japan_pteridos_biodiv.Rmd")),
-    output_file = file_out(here::here("reports/japan_pteridos_biodiv.html")),
+    knitr_in(here::here("ms/japan_pteridos_biodiv.Rmd")),
+    output_file = file_out(here::here("ms/japan_pteridos_biodiv.html")),
     quiet = TRUE)
 )

### 2019-08-31 report/japan_pteridos_biodiv.Rmd
Move html report to report/

---
title: "Biodiversity of Japanese Pteridophytes"
author: "Joel Nitta, Atushi Ebihara"
bibliography: "references.bib"
output: 
  bookdown::html_document2:
    fig_caption: yes
    number_sections: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)
```

```{r load-data, cache = FALSE}
loadd(list = c("world_map",
  "occ_data_pteridos",
  "occ_data_ferns_north",
  "occ_data_ferns",
  "alpha_div_pteridos",
  "alpha_div_ferns",
  "alpha_div_ferns_ns",
  "alpha_div_vs_repro_ferns"), 
  cache = ja_fern_cache)
```

## Methods

### Phylogenetic diversity

We measured phylogenetic diversity using mean phylogenetic distance (MPD), the mean distance between all co-occurring taxa in a community, and mean nearest taxon distance (MNTD), the mean distance between co-occurring sister taxa in a community [@Webb2000; @Webb2002]. A "community" is defined here as co-occurring taxa within a 1 km^2^ grid cell. Since observed values of MPD and MNTD are known to correlate with species richness, we determined whether communities are phylogenetically overdispersed (i.e., diverse) or clustered (i.e., depauperate), by calculating the standard effect size (SES) of MPD and MNTD. We compared observed values to those of a null distribution of 999 random communities with the "ses.pd" function in the "picante" R package [@Kembel2010]. Random communities were each generated using 10,000 iterations of the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities. Positive values of SES of MPD and MNTD (hereafter referred to simply as "MPD" or "MTND" respectively unless indicated otherwise) indicate phylogenetic overdispersion, and negative values indicate phylogenetic clustering. Values at the 2.5% lower or upper extremes of the null distribution were considered statistically significant.

### Datasets

A large number of primarily tropical fern lineages only occur in the far-southern islands of Japan: `r setdiff(occ_data_ferns$genus, occ_data_ferns_north$genus) %>% length` genera and `r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% length` families (`r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% paste(collapse = ", ")`) do not occur north of 30.1 degrees latitude.

When these lineages are included in the randomization procedure to calculate SES of MPD and MNTD, they likely contribute long branches that bias the northern communities towards clustering. Lycophytes, which occur on long branches, may have a similar effect. Therefore, we calculated phylogenetic diversity for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns separately north and south of 30.1 degrees latitude (corresponding to the Tokara strait, hereafter referred to as "ferns_NS").

## Results

### Richness

Richness of pteridophytes in Japan roughly follows the latitudinal species gradient, with more species in the south (Fig. \@ref(fig:richness-pteridos)).

### Phylogenetic diversity

Standard effect sizes of mean phylogenetic diversity (MPD) and mean nearest taxon distance (MNTD) were either not correlated or only very weakly correlated with species richness (SI Fig. \@ref(fig:compare-phylo-div-richness)), indicating that our null models successfully decoupled measures of phylogenetic diversity and richness. The distribution of MPD and MNTD values were very similar across datasets (SI Fig. \@ref(fig:alpha-div-ridge)).

Values of MPD are clustered in mountainous areas and overdispersed along the coasts of the main islands of Japan, with similar patterns between pteridophytes and ferns (Fig. \@ref(fig:phy-mpd)a, b). The pattern is similar in the "ferns_NS" dataset, except that the strength of both clustering and overdispersion is weaker in the main islands of Japan, and the pattern of overdispresion on the islands of Amami-oshima in sourthern Japan reverses to clustering (Fig. \@ref(fig:phy-mpd)c).

The signal of clustering is much stronger when measured with MNTD, especially in mountainous areas of central Honshu (Fig. \@ref(fig:phy-mntd)). Patterns are similar between all three datasets except that MNTD values are smaller in the "ferns_NS" dataset (Fig. \@ref(fig:phy-mntd)c).

Values of MPD and MNTD decrease significantly with elevation across all datasets (Fig. \@ref(fig:compare-phy-div-el), linear model).

### Breeding system

Sexual diploid taxa tend to be distributed in northern latitudes (Fig. \@ref(fig:sex-dip-ferns-map)).

Somewhat surprisingly, MPD and MNTD show a weak negative relationship with the percentage of sexual diploid taxa (Fig. \@ref(fig:compare-pd-sexdip); linear model). We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed. It seems that communities with many non-sexual taxa actually tend to include distantly related taxa rather than sister pairs of sexual and non-sexual species.

## Figures

```{r richness-pteridos, fig.cap = "Richness of native Japanese pteridophyte taxa, excluding hybrids."}
make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "richness", 
    metric_title = "Richness"
  ) + scale_fill_scico(palette = "bilbao", na.value="white")
```

```{r compare-phylo-div-richness, fig.cap = "Correlation between phylogenetic diversity and richness. Phylogenetic diversity measured as standard effect size (SES) of mean phylogenetic diversity (MPD) or mean nearest taxon distance (MNTD) in native pteridophytes, ferns, or ferns calculated separately north and south of 30.1 latitude of Japan. Blue trendline shows significant linear model fit (*P* < 0.05) with 96% confidence interval in pink."}
a <- make_scatter_with_lm(alpha_div_pteridos, "richness", "phy_mpd_obs_z", "Richness", "MPD") +
  labs(title = "Pteridophytes", x = "")
b <- make_scatter_with_lm(alpha_div_ferns, "richness", "phy_mpd_obs_z", "Richness", "MPD") +
  labs(title = "Ferns", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "phy_mpd_obs_z", "Richness", "MPD") +
  labs(title = "Ferns (NS separate)", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_pteridos, "richness", "phy_mntd_obs_z", "Richness", "MNTD")
e <- make_scatter_with_lm(alpha_div_ferns, "richness", "phy_mntd_obs_z", "Richness", "MNTD") +
  labs(y = "")
f <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "phy_mntd_obs_z", "Richness", "MNTD") +
  labs(y = "")
a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
```

```{r compare-func-div-richness, fig.cap = "Correlation between functional diversity and richness. Functional diversity measured as standard effect size (SES) of mean phylogenetic diversity (MPD) or mean nearest taxon distance (MNTD) using a trait distance matrix in native pteridophytes, ferns, or ferns calculated separately north and south of 30.1 latitude of Japan. Blue trendline shows significant linear model fit (*P* < 0.05) with 96% confidence interval in pink."}
a <- make_scatter_with_lm(alpha_div_pteridos, "richness", "func_mpd_obs_z", "Richness", "MPD") +
  labs(title = "Pteridophytes", x = "")
b <- make_scatter_with_lm(alpha_div_ferns, "richness", "func_mpd_obs_z", "Richness", "MPD") +
  labs(title = "Ferns", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "func_mpd_obs_z", "Richness", "MPD") +
  labs(title = "Ferns (NS separate)", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_pteridos, "richness", "func_mntd_obs_z", "Richness", "MNTD")
e <- make_scatter_with_lm(alpha_div_ferns, "richness", "func_mntd_obs_z", "Richness", "MNTD") +
  labs(y = "")
f <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "func_mntd_obs_z", "Richness", "MNTD") +
  labs(y = "")
a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
```

```{r alpha-div-ridge, fig.cap = "Ridge plots showing distribution of the standard effect size of a) mean phylogenetic distance (MPD) and b) mean nearest taxon distance (MNTD) for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns when the data were split into northern and southern portions and calculated separately. Black vertical lines shows median."}
# Combine mpd and mntd calculations from the three datasets for comparison
comp_data <- 
  bind_rows(
  transmute(alpha_div_pteridos, dataset = "pteridos", phy_mpd_obs_z, phy_mntd_obs_z),
  transmute(alpha_div_ferns, dataset = "ferns", phy_mpd_obs_z, phy_mntd_obs_z),
  transmute(alpha_div_ferns_ns, dataset = "ferns_ns", phy_mpd_obs_z, phy_mntd_obs_z)
  )

# MPD
a <- ggplot(comp_data, aes(x = phy_mpd_obs_z, y = dataset, fill = dataset)) + geom_density_ridges(
  scale = 1,
  quantile_lines = TRUE, quantiles = 2)  +
  labs(
    x = "MPD",
    y = "Dataset"
  ) +
  jntools::standard_theme() +
  theme(legend.position = "none")

# MNTD
b <- ggplot(comp_data, aes(x = phy_mntd_obs_z, y = dataset, fill = dataset)) + geom_density_ridges(
  scale = 1,
  quantile_lines = TRUE, quantiles = 2) +
  theme(legend.position = "none") +
   labs(
    x = "MNTD",
    y = "Dataset"
  ) +
  jntools::standard_theme() +
  theme(legend.position = "none")

# final plot
a + b + plot_layout(ncol = 1) + plot_annotation(tag_levels = "a")

```

```{r phy-mpd, fig.asp = 2.0, fig.cap = "Phylogenetic standard effect size (SES) of mean phylogenetic distance (MPD) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude. All datasets are native taxa of Japan excluding hybrids."}
a <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "phy_mpd_obs_z", 
    metric_title = "MPD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    )) +
  labs(title = "a, pteridophytes") +
  theme(legend.position = c(0.1, 0.80))

b <- make_diversity_map(
    div_data = alpha_div_ferns, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "phy_mpd_obs_z", 
    metric_title = "MPD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
    )) +
  labs(title = "b, ferns") +
  theme(legend.position = c(0.1, 0.80))

c <- make_diversity_map(
    div_data = alpha_div_ferns_ns, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "phy_mpd_obs_z", 
    metric_title = "MPD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns_ns, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns_ns, phy_mpd_obs_z, "abs", 3)
    )) +
  labs(title = "c, ferns (NS separate)") +
  theme(legend.position = c(0.1, 0.80))

a + b + c + plot_layout(ncol = 1)
```

```{r phy-mntd, fig.asp = 2.0, fig.cap = "Phylogenetic standard effect size (SES) of mean nearest taxon distance (MNTD) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude. All datasets are native taxa of Japan excluding hybrids."}
a <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "phy_mntd_obs_z", 
    metric_title = "MNTD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mntd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mntd_obs_z, "abs", 3)
    )) +
  labs(title = "a, pteridophytes") +
  theme(legend.position = c(0.1, 0.80))

b <- make_diversity_map(
    div_data = alpha_div_ferns, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "phy_mntd_obs_z", 
    metric_title = "MNTD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns, phy_mntd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, phy_mntd_obs_z, "abs", 3)
    )) +
  labs(title = "b, ferns") +
  theme(legend.position = c(0.1, 0.80))

c <- make_diversity_map(
    div_data = alpha_div_ferns_ns, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "phy_mntd_obs_z", 
    metric_title = "MNTD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns_ns, phy_mntd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns_ns, phy_mntd_obs_z, "abs", 3)
    )) +
  labs(title = "c, ferns (NS separate)") +
  theme(legend.position = c(0.1, 0.80))

a + b + c + plot_layout(ncol = 1)
```

```{r func-mpd, fig.asp = 2.0, fig.cap = "Functional standard effect size (SES) of mean phylogenetic distance (MPD) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude. All datasets are native taxa of Japan excluding hybrids."}
a <- make_diversity_map(
  div_data = alpha_div_pteridos, 
  world_map = world_map, 
  occ_data = occ_data_pteridos, 
  div_metric = "func_mpd_obs_z", 
  metric_title = "MPD"
) + scale_fill_scico(
  palette = "vik", 
  na.value="white",
  limits = c(
    -get_limit(alpha_div_pteridos, func_mpd_obs_z, "abs", 3), 
    get_limit(alpha_div_pteridos, func_mpd_obs_z, "abs", 3)
  )) +
  labs(title = "a, pteridophytes") +
  theme(legend.position = c(0.1, 0.80))

b <- make_diversity_map(
  div_data = alpha_div_ferns, 
  world_map = world_map, 
  occ_data = occ_data_ferns, 
  div_metric = "func_mpd_obs_z", 
  metric_title = "MPD"
) + scale_fill_scico(
  palette = "vik", 
  na.value="white",
  limits = c(
    -get_limit(alpha_div_ferns, func_mpd_obs_z, "abs", 3), 
    get_limit(alpha_div_ferns, func_mpd_obs_z, "abs", 3)
  )) +
  labs(title = "b, ferns") +
  theme(legend.position = c(0.1, 0.80))

c <- make_diversity_map(
  div_data = alpha_div_ferns_ns, 
  world_map = world_map, 
  occ_data = occ_data_ferns, 
  div_metric = "func_mpd_obs_z", 
  metric_title = "MPD"
) + scale_fill_scico(
  palette = "vik", 
  na.value="white",
  limits = c(
    -get_limit(alpha_div_ferns_ns, func_mpd_obs_z, "abs", 3), 
    get_limit(alpha_div_ferns_ns, func_mpd_obs_z, "abs", 3)
  )) +
  labs(title = "c, ferns (NS separate)") +
  theme(legend.position = c(0.1, 0.80))

a + b + c + plot_layout(ncol = 1)
```

```{r func-mntd, fig.asp = 2.0, fig.cap = "Functional standard effect size (SES) of mean nearest taxon distance (MNTD) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude. All datasets are native taxa of Japan excluding hybrids."}
a <- make_diversity_map(
  div_data = alpha_div_pteridos, 
  world_map = world_map, 
  occ_data = occ_data_pteridos, 
  div_metric = "func_mntd_obs_z", 
  metric_title = "MNTD"
) + scale_fill_scico(
  palette = "vik", 
  na.value="white",
  limits = c(
    -get_limit(alpha_div_pteridos, func_mntd_obs_z, "abs", 3), 
    get_limit(alpha_div_pteridos, func_mntd_obs_z, "abs", 3)
  )) +
  labs(title = "a, pteridophytes") +
  theme(legend.position = c(0.1, 0.80))

b <- make_diversity_map(
  div_data = alpha_div_ferns, 
  world_map = world_map, 
  occ_data = occ_data_ferns, 
  div_metric = "func_mntd_obs_z", 
  metric_title = "MNTD"
) + scale_fill_scico(
  palette = "vik", 
  na.value="white",
  limits = c(
    -get_limit(alpha_div_ferns, func_mntd_obs_z, "abs", 3), 
    get_limit(alpha_div_ferns, func_mntd_obs_z, "abs", 3)
  )) +
  labs(title = "b, ferns") +
  theme(legend.position = c(0.1, 0.80))

c <- make_diversity_map(
  div_data = alpha_div_ferns_ns, 
  world_map = world_map, 
  occ_data = occ_data_ferns, 
  div_metric = "func_mntd_obs_z", 
  metric_title = "MNTD"
) + scale_fill_scico(
  palette = "vik", 
  na.value="white",
  limits = c(
    -get_limit(alpha_div_ferns_ns, func_mntd_obs_z, "abs", 3), 
    get_limit(alpha_div_ferns_ns, func_mntd_obs_z, "abs", 3)
  )) +
  labs(title = "c, ferns (NS separate)") +
  theme(legend.position = c(0.1, 0.80))

a + b + c + plot_layout(ncol = 1)
```

```{r compare-phy-div-el, fig.cap = "Relationship between phylogenetic standard effect size (SES) of mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) of native Japanese pteridophytes and elevation (m). Blue trendline shows significant linear model fit with confidence interval in pink."}
a <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "phy_mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Pteridophytes", x = "")
b <- make_scatter_with_lm(alpha_div_ferns, "elevation", "phy_mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Ferns", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns_ns, "elevation", "phy_mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Ferns (NS separate)", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "phy_mntd_obs_z", "Elevation (m)", "MNTD")
e <- make_scatter_with_lm(alpha_div_ferns, "elevation", "phy_mntd_obs_z", "Elevation (m)", "MNTD") +
  labs(y = "")
f <- make_scatter_with_lm(alpha_div_ferns_ns, "elevation", "phy_mntd_obs_z", "Elevation (m)", "MNTD") +
  labs(y = "")
a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
```

```{r compare-func-div-el, fig.cap = "Relationship between functional standard effect size (SES) of mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) of native Japanese pteridophytes and elevation (m). Blue trendline shows significant linear model fit with confidence interval in pink."}
a <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "func_mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Pteridophytes", x = "")
b <- make_scatter_with_lm(alpha_div_ferns, "elevation", "func_mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Ferns", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns_ns, "elevation", "func_mpd_obs_z", "Elevation (m)", "MPD") +
  labs(title = "Ferns (NS separate)", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "func_mntd_obs_z", "Elevation (m)", "MNTD")
e <- make_scatter_with_lm(alpha_div_ferns, "elevation", "func_mntd_obs_z", "Elevation (m)", "MNTD") +
  labs(y = "")
f <- make_scatter_with_lm(alpha_div_ferns_ns, "elevation", "func_mntd_obs_z", "Elevation (m)", "MNTD") +
  labs(y = "")
a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
```

```{r sex-dip-ferns-map, fig.cap = "Percentage of sexual diploid fern taxa, excluding hybrids."}
make_diversity_map(
  div_data = alpha_div_ferns, 
  world_map = world_map, 
  occ_data = occ_data_ferns, 
  div_metric = "percent_sex_dip",
  metric_title = "% sex. diploid"
) + scale_fill_scico(palette = "bilbao", na.value="white")
```

```{r compare-pd-sexdip, fig.cap = "Relationship between standard effect size (SES) of mean phylogenetic distance (MPD) and mean nearest taxon distance (MNTD) of native Japanese pteridophytes and percent sexual diploid taxa in each community. Blue trendline shows significant linear model fit with confidence interval in pink."}
a <- make_scatter_with_lm(alpha_div_pteridos, "percent_sex_dip", "phy_mpd_obs_z", "% sexual diploid", "MPD") +
  labs(title = "Pteridophytes", x = "")
b <- make_scatter_with_lm(alpha_div_ferns, "percent_sex_dip", "phy_mpd_obs_z", "% sexual diploid", "MPD") +
  labs(title = "Ferns", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns_ns, "percent_sex_dip", "phy_mpd_obs_z", "% sexual diploid", "MPD") +
  labs(title = "Ferns (NS separate)", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_pteridos, "percent_sex_dip", "phy_mntd_obs_z", "% sexual diploid", "MNTD")
e <- make_scatter_with_lm(alpha_div_ferns, "percent_sex_dip", "phy_mntd_obs_z", "% sexual diploid", "MNTD") +
  labs(y = "")
f <- make_scatter_with_lm(alpha_div_ferns_ns, "percent_sex_dip", "phy_mntd_obs_z", "% sexual diploid", "MNTD") +
  labs(y = "")
a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
```

## References

### 2019-08-31 ms/manuscript.Rmd
Add manuscript Rmd and associated files

---
title: "Exploring multiple dimensions of biodiversity in Japanese ferns and lycophytes"
bibliography: "references.bib"
mainfont: Times New Roman.ttf
fontsize: 12pt
mainfontoptions: 
- BoldFont=Times New Roman Bold.ttf
- ItalicFont=Times New Roman Italic.ttf
- BoldItalicFont=Times New Roman Bold Italic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
    includes:
      in_header: newphyt.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  chunk_output_type: console
# nocite: | 
#   @Vogel1968, @Wollenweber2000
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)
```

```{r load-packages, cache = FALSE}
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(tidyverse)
```

```{r load-data, cache = FALSE}
# Set cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

source(here::here("code/functions.R"))

# Load targets
loadd(list = c(
  "all_cells",
  "occ_data_ferns",
  "occ_data_ferns_north",
  "occ_data_pteridos",
  "alpha_div_pteridos",
  "world_map"), 
  cache = ja_fern_cache)
```

<!-- FOR CONVERSION TO MS WORD ONLY: remove page numbers -->
\pagenumbering{gobble}

<!-- Set left justification (better conversion to Word than default full justification) -->
\raggedright

<!-- Don't indent authors -->
\setlength{\parindent}{0ex}

Joel H. Nitta^1^ and Atsushi Ebihara^2^

^1^Department of Botany, National Museum of Natural History, Smithsonian Institute, Washington, D.C., 20013, U.S.A.

^2^Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba-shi, Ibaraki, 305-0005, Japan

Author for correspondence:  
*Joel H. Nitta*  
*Tel: +1 6174960515*  
*Email: joelnitta@gmail.com*

<!-- Set indent to about 3 characters hereafter -->
\setlength{\parindent}{3ex}

\clearpage

## Abstract

The pteridophytes (ferns and lycophytes) of Japan are one of the best-characterized regional floras in the world thanks to decades of collection and taxonomic work. Here, we combine phylogenetic, functional, and environmental datasets to characterize the spatial distribution of biodiversity in the pteridophytes of Japan.

**Key words:** biodiversity, biogeography, phylogenetic diversity.

\clearpage

## Introduction

## Materials and Methods

### Study Site

Our data include all ferns and lycophytes known to occur within the political boundaries of Japan (20° to 45° N, 122° to 153° E). Japan consists of four main islands and thousands of smaller ones. Our data exlude islands smaller than XX km^2^. The climate is subtropical in the south and highly seasonal in other parts of the country.

### Occurrence data

We used occurrence data of ferns and lycophytes in Japan compiled from > 200,000 vouchered specimens including 721 non-hybrid taxa (species, subspecies, and varieties) and 371 interspecific hybrids (1090 taxa total) [@Ebihara2019b]. The data are formatted as taxon presences in 10 km × 10 km grid cells (hereafter '10 km cells'). The 10 km grid cells correspond to the secondary level of basic grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) (`r nrow(all_cells) %>% comma` 10 km cells total).

### Trait data

We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017; @Ebihara2019b]. Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens. All qualitative traits formatted in binary format; e.g., a single trait with three states is formatted as three binary traits. Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately. Quantitative traits include aspects of leaf size (e.g., frond length, frond width, stipe length, etc.). Qualitative traits include aspects of leaf morphology (e.g., margin shape, leaf color, indusia shape, spore arrangement, etc.).

### Statistical analysis

We used two distance-based metrics to quantify alpha biodiversity, mean phylogenetic distance (MPD; the mean distance between all co-occurring taxa in a community) and mean nearest taxon distance (MNTD; the mean distance between closest taxon pairs in a community) [@Webb2000; @Webb2002]. Although MPD and MNTD were originally formulated for phylogenetic distances, they can be applied to any type of distance matrix. We calculated phylogenetic distances from the phylogenetic tree using the "cophenetic" function in R. We calculated functional distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011]. We weighted quantitative and qualitative traits equally, and weighted each binary component trait of each qualitative trait equally.

Since observed values of MPD and MNTD are known to correlate with species richness [@Swenson2014], we determined whether communities are overdispersed (i.e., diverse) or clustered (i.e., depauperate), by calculating the standard effect size (SES) of MPD and MNTD. We compared observed values to those of a null distribution of 999 random communities with the \function{ses.mpd} and \function{ses.mntd} functions in the \package{picante} R package [@Kembel2010]. Random communities were each generated using 10,000 iterations of the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities. Positive values of SES of MPD and MNTD indicate overdispersion, and negative values indicate clustering. We hereafter refer to the SES of phylogenetic and functional MPD and MNTD as MPD~phy~ and MNTD~phy~, and MPD~func~ and MTND~func~, respectively.

A large number of primarily tropical fern lineages only occur in the far-southern islands of Japan: `r setdiff(occ_data_ferns$genus, occ_data_ferns_north$genus) %>% length` genera and `r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% length` families (`r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% paste(collapse = ", ")`) do not occur north of 30.1 degrees latitude. When these lineages are included in the randomization procedure to calculate SES of MPD and MNTD, they likely contribute long branches that bias the northern communities towards clustering. Lycophytes, which occur on long branches, may have a similar effect. Therefore, we calculated phylogenetic diversity for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns separately north and south of 30.1 degrees latitude (corresponding to the Tokara strait, hereafter referred to as "ferns_NS").

### Data accessibility

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. Raw and processed data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.

### Richness

Richness of pteridophytes in Japan roughly follows the latitudinal species gradient, with more species in the south (\fig \@ref(fig:div-pteridos)a). Richness tended to increase with elevation ((\fig \@ref(fig:div-el)a).

### Phylogenetic diversity

Standard effect sizes of mean phylogenetic diversity (MPD~phy~) and mean nearest taxon distance (MNTD~phy~) were either not correlated or only very weakly correlated with species richness (SI Fig. ), indicating that our null models successfully decoupled measures of phylogenetic diversity and richness.

Values of MPD~phy~ are clustered in mountainous areas and overdispersed along the coasts of the main islands of Japan (\fig \@ref(fig:div-pteridos)b), with similar patterns between pteridophytes and ferns (SI Fig. ). The pattern is similar in the "ferns_NS" dataset, except that the strength of both clustering and overdispersion is weaker in the main islands of Japan, and the pattern of overdispersion on the islands of Amami-oshima in sourthern Japan reverses to clustering (SI Fig. ).

The signal of clustering is much stronger when measured with MNTD~phy~, especially in mountainous areas of central Honshu (\fig \@ref(fig:div-pteridos)c). Patterns are similar between all three datasets except that MNTD~phy~ values are smaller in the "ferns_NS" dataset (SI Fig. ).

Values of MPD and MNTD decrease significantly with elevation across all datasets (\fig \@ref(fig:div-el)b, c; linear model).

### Functional diversity

Spatial patterns of functional diversity differed from phylogenetic diversity and by metric. When measured with MPD~func~, pteridophytes were overdispersed in Hokkaido, northern Honshu, and the smaller islands south of Kyushu, but clustered in southern Honshu, Shikoku, and Kyushu (\fig \@ref(fig:div-pteridos)d). Less spatial structure was evident in MNTD~func~, with more clustering overall (\fig \@ref(fig:div-pteridos)e).

Functional diversity increased with elevation when measured with MPD~func~, but not MNTD~func~ (\fig \@ref(fig:div-el)d, e; linear model)

### Breeding system

Sexual diploid taxa tend to be distributed in northern latitudes (\fig \@ref(fig:div-pteridos)f).

Somewhat surprisingly, MPD and MNTD show a weak negative relationship with the percentage of sexual diploid taxa (linear model). We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed. It seems that communities with many non-sexual taxa actually tend to include distantly related taxa rather than sister pairs of sexual and non-sexual species.

## Discussion

## Acknowledgements

We thank people

## Author Contribution

Who did what

## References
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

\clearpage

## Supporting Information

Additional Supporting Information may be found online in the Supporting Information section at the end of the article:

**Table S1**

**Fig. S1**

\clearpage

**\fig \@ref(fig:div-pteridos).** Alpha diversity of ferns and lycophytes in Japan. **a** species richness. **b** Phylogenetic mean phylogenetic distance (MPD~phy~). **c** Phylogenetic mean nearest taxon distance (MNTD~phy~). **d** Morphological MPD (MPD~morph~). **e** Morphological MNTD (MNTD~morph~).

**\fig \@ref(fig:div-el).** Relationship between alpha diversity metrics and elevation (m) in native Japanese pteridophytes. Blue trendline shows significant linear model fit with confidence interval in pink. **a** species richness. **b** Phylogenetic mean phylogenetic distance (MPD~phy~). **c** Phylogenetic mean nearest taxon distance (MNTD~phy~). **d** Morphological MPD (MPD~morph~). **e** Morphological MNTD (MNTD~morph~).

\clearpage

<!-- Tables -->

\clearpage

<!-- Figures -->

```{r div-pteridos, fig.height = 9, fig.width = 7, fig.cap = "Alpha diversity of ferns and lycophytes in Japan"}

# pterido richness
  a <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "richness", 
    metric_title = "No. species ",
    label = "a") +
    scale_fill_scico(palette = "bamako", na.value="grey")

  # pterido_phy_mpd
  b <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "phy_mpd_obs_z", 
    metric_title = "SES of MPD",
    label = "b"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    ))

  # pterido_phy_mntd
  c <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "phy_mntd_obs_z", 
    metric_title = "SES of MNTD",
    label = "c"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    ))

  # pterido_func_mpd
  d <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "func_mpd_obs_z", 
    metric_title = "SES of MPD",
    label = "d"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    ))

  # pterido_func_mntd
  e <- make_diversity_map(
    div_data = alpha_div_pteridos, 
    world_map = world_map, 
    occ_data = occ_data_pteridos, 
    div_metric = "func_mntd_obs_z", 
    metric_title = "SES of MNTD",
    label = "e"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
    ))

  f <- make_diversity_map(
  div_data = alpha_div_pteridos, 
  world_map = world_map, 
  occ_data = occ_data_pteridos, 
  div_metric = "percent_sex_dip",
  metric_title = "% sex. diploid",
  label = "f"
)  +
    scale_fill_scico(palette = "bamako", na.value="grey")

  a + b + c + d + e + f + plot_layout(ncol = 2, nrow = 3)

ggsave(here::here("ms/fig_1.pdf"), width = 7, height = 9, units = "in")
```

```{r div-el, fig.height = 3.5, fig.width = 7, fig.cap = "Relationship between alpha diversity metrics and elevation"}
a <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "richness", "Elevation (m)", "Richness") +
  labs(title = "a", x = "Elevation (m)")
b <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "phy_mpd_obs_z", "Elevation (m)", "MPDphy") +
  labs(title = "b", x = "Elevation (m)")
c <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "phy_mntd_obs_z", "Elevation (m)", "MNTDphy") +
  labs(title = "c", x = "Elevation (m)")
d <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "func_mpd_obs_z", "Elevation (m)", "MPDfunc") +
  labs(title = "d", x = "Elevation (m)")
e <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "func_mntd_obs_z", "Elevation (m)", "MNTDfunc") +
  labs(title = "e", x = "Elevation (m)")
a + b + c + d + e + plot_layout(ncol = 3, nrow = 2)

ggsave(here::here("ms/fig_2.pdf"), width = 7, height = 3.5, units = "in")
```

### 2019-08-31 render.R

@@ -1,0 +1,4 @@
+rmarkdown::render(
+  here::here("ms/manuscript.Rmd"),
+  output_file = here::here("ms/manuscript.pdf"),
+  quiet = TRUE)

### 2019-09-01 ms/manuscript.Rmd
Update manuscript
- Add supplemental info

@@
 ---
-title: "Exploring multiple dimensions of biodiversity in Japanese ferns and lycophytes"
+title: "Exploring dimensions of biodiversity in Japanese ferns and lycophytes"
 bibliography: "references.bib"
@@
   "all_cells",
+  "alpha_div_pteridos",
+  "alpha_div_ferns",
+  "alpha_div_ferns_ns",
   "occ_data_ferns",
@@
   "occ_data_pteridos",
-  "alpha_div_pteridos",
+  "traits_for_dist",
   "world_map"), 
   cache = ja_fern_cache)
+
 ```
 
-<!-- FOR CONVERSION TO MS WORD ONLY: remove page numbers -->
+<!-- Remove page numbers -->
 \pagenumbering{gobble}
 
-<!-- Set left justification (better conversion to Word than default full justification) -->
+<!-- Set left justification -->
 \raggedright
 
-<!-- Don't indent authors -->
+<!-- Don't indent -->
 \setlength{\parindent}{0ex}
@@
 *Joel H. Nitta*  
-*Tel: +1 6174960515*  
 *Email: joelnitta@gmail.com*
-
-<!-- Set indent to about 3 characters hereafter -->
-\setlength{\parindent}{3ex}
 
@@
 
-The pteridophytes (ferns and lycophytes) of Japan are one of the best-characterized regional floras in the world thanks to decades of collection and taxonomic work.
-Here, we combine phylogenetic, functional, and environmental datasets to characterize the spatial distribution of biodiversity in the pteridophytes of Japan.
+The pteridophytes (ferns and lycophytes) of Japan are one of the best-characterized regional floras in the world due to decades of intensive collection and taxonomic revision.
+We combined phylogenetic, functional, and environmental datasets to explore the spatial distribution of biodiversity in the pteridophytes of Japan.
 
@@
 Japan consists of four main islands and thousands of smaller ones.
-Our data exlude islands smaller than XX km^2^.
+Our data exclude islands smaller than XX km^2^.
 The climate is subtropical in the south and highly seasonal in other parts of the country.
@@
 
-We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
+We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017; @Ebihara2019b] (\tab \@ref(tab:traits-table)).
 Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens.
-All qualitative traits formatted in binary format; e.g., a single trait with three states is formatted as three binary traits.
+All qualitative traits in binary format; e.g., a single trait with three states is formatted as three binary traits.
 Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately.
@@
 Qualitative traits include aspects of leaf morphology (e.g., margin shape, leaf color, indusia shape, spore arrangement, etc.).
+
+We used ploidy level and breeding system data in @Ebihara2019b.
 
@@
 Although MPD and MNTD were originally formulated for phylogenetic distances, they can be applied to any type of distance matrix.
-We calculated phylogenetic distances from the phylogenetic tree using the "cophenetic" function in R.
+We calculated phylogenetic distances from the phylogenetic tree using the \function{cophenetic} function in R v.
+3.6.0 [@RCoreTeam2019].
 We calculated functional distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011].
 We weighted quantitative and qualitative traits equally, and weighted each binary component trait of each qualitative trait equally.
+We log-transformed and scaled quantitative traits prior to calculating distances.
 
@@
 A large number of primarily tropical fern lineages only occur in the far-southern islands of Japan: `r setdiff(occ_data_ferns$genus, occ_data_ferns_north$genus) %>% length` genera and `r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% length` families (`r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% paste(collapse = ", ")`) do not occur north of 30.1 degrees latitude.
-When these lineages are included in the randomization procedure to calculate SES of MPD and MNTD, they likely contribute long branches that bias the northern communities towards clustering.
+When these lineages are included in the randomization procedure to calculate SES of MPD and MNTD, they may contribute long branches that bias the northern communities towards clustering.
 Lycophytes, which occur on long branches, may have a similar effect.
@@
 
+## Results
+
 ### Richness
@@
 Richness of pteridophytes in Japan roughly follows the latitudinal species gradient, with more species in the south (\fig \@ref(fig:div-pteridos)a).
-Richness tended to increase with elevation ((\fig \@ref(fig:div-el)a).
+Richness tended to increase with elevation (\fig \@ref(fig:div-el)a).
 
@@
 
-Standard effect sizes of mean phylogenetic diversity (MPD~phy~) and mean nearest taxon distance (MNTD~phy~) were either not correlated or only very weakly correlated with species richness (SI Fig.
-), indicating that our null models successfully decoupled measures of phylogenetic diversity and richness.
+Standard effect sizes of mean phylogenetic diversity (MPD~phy~) and mean nearest taxon distance (MNTD~phy~) were either not correlated or only very weakly correlated (\rval^2^ < 0.005) with species richness (\fig \@ref(fig:phy-div-rich)), indicating that our null models successfully decoupled measures of phylogenetic diversity and richness.
 
-Values of MPD~phy~ are clustered in mountainous areas and overdispersed along the coasts of the main islands of Japan (\fig \@ref(fig:div-pteridos)b), with similar patterns between pteridophytes and ferns (SI Fig.
-).
-The pattern is similar in the "ferns_NS" dataset, except that the strength of both clustering and overdispersion is weaker in the main islands of Japan, and the pattern of overdispersion on the islands of Amami-oshima in sourthern Japan reverses to clustering (SI Fig.
-).
+Values of MPD~phy~ are clustered in mountainous areas and overdispersed along the coasts of the main islands of Japan (\fig \@ref(fig:div-pteridos)b), with similar patterns between pteridophytes and ferns (\fig \@ref(fig:mpd-phy)).
+The pattern is similar in the "ferns_NS" dataset, except that the strength of both clustering and overdispersion is weaker in the main islands of Japan, and the pattern of overdispersion on the islands of Amami-oshima in sourthern Japan reverses to clustering (\fig \@ref(fig:mpd-phy)).
 
 The signal of clustering is much stronger when measured with MNTD~phy~, especially in mountainous areas of central Honshu (\fig \@ref(fig:div-pteridos)c).
-Patterns are similar between all three datasets except that MNTD~phy~ values are smaller in the "ferns_NS" dataset (SI Fig.
-).
+Patterns are similar between all three datasets except that MNTD~phy~ values are smaller in the "ferns_NS" dataset (\fig \@ref(fig:mntd-phy)).
 
-Values of MPD and MNTD decrease significantly with elevation across all datasets (\fig \@ref(fig:div-el)b, c; linear model).
+Values of MPD and MNTD decrease significantly with elevation across all datasets (\fig \@ref(fig:div-el)b, c).
 
@@
 
-Spatial patterns of functional diversity differed from phylogenetic diversity and by metric.
+Values of (MPD~func~) and (MNTD~func~) were very weakly negatively correlated (\rval^2^ < 0.01) with species richness (\fig \@ref(fig:func-div-rich)).
+
 When measured with MPD~func~, pteridophytes were overdispersed in Hokkaido, northern Honshu, and the smaller islands south of Kyushu, but clustered in southern Honshu, Shikoku, and Kyushu (\fig \@ref(fig:div-pteridos)d).
-Less spatial structure was evident in MNTD~func~, with more clustering overall (\fig \@ref(fig:div-pteridos)e).
+Patterns were similar across datasets except for a reversal in MPD values on the islands of Amami-oshima in the "ferns_NS" dataset (\fig \@ref(fig:mpd-func)), similar to that observed for phylogenetic diversity (\fig \@ref(fig:mpd-phy)).
 
-Functional diversity increased with elevation when measured with MPD~func~, but not MNTD~func~ (\fig \@ref(fig:div-el)d, e; linear model)
+Less spatial structure was evident in MNTD~func~, with more clustering overall (\fig \@ref(fig:div-pteridos)e). Patterns were similar across datasets (\fig \@ref(fig:mntd-func)).
+
+Functional diversity increased with elevation when measured with MPD~func~, but not MNTD~func~ (\fig \@ref(fig:div-el)d, e).
 
@@
 
-Somewhat surprisingly, MPD and MNTD show a weak negative relationship with the percentage of sexual diploid taxa (linear model).
+Somewhat surprisingly, MPD and MNTD show a weak negative relationship with the percentage of sexual diploid taxa (\fig \@ref(fig:compare-pd-sexdip)).
 We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed.
@@
 
-We thank people
-
 ## Author Contribution
-
-Who did what
 
@@
 
-## Supporting Information
-
-Additional Supporting Information may be found online in the Supporting Information section at the end of the article:
-
-**Table S1**
-
-**Fig. S1**
-
-\clearpage
+<!-- Figure Legends -->
 
@@
 **e** Morphological MNTD (MNTD~morph~).
-
-\clearpage
-
-<!-- Tables -->
 
@@
 ```
+
+\clearpage
+
+\beginsupplement
+
+## Supporting Information
+
+\clearpage
+
+```{r traits-table, results = "asis"}
+make_trait_summary(traits_for_dist) %>%
+knitr::kable(
+  booktabs = TRUE,
+  caption = 'Traits used in this study.'
+)
+```
+
+\clearpage
+
+```{r phy-div-rich, fig.height = 4, fig.width = 7, fig.cap = "Relationship between phylogenetic diversity and richness.
+Phylogenetic diversity metrics include standard effect size (SES) of mean phylogenetic diversity (MPD~phy~) or mean nearest taxon distance (MNTD~phy~) using functional trait distances in native pteridophytes, ferns, or ferns calculated separately north and south of 30.1 latitude of Japan.
+Blue trendline shows significant linear model fit (*P* < 0.05) with 96% confidence interval in pink."}
+a <- make_scatter_with_lm(alpha_div_pteridos, "richness", "phy_mpd_obs_z", "Richness", "MPDphy") +
+  labs(title = "Pteridophytes", x = "")
+b <- make_scatter_with_lm(alpha_div_ferns, "richness", "phy_mpd_obs_z", "Richness", "MPDphy") +
+  labs(title = "Ferns", y = "", x = "")
+c <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "phy_mpd_obs_z", "Richness", "MPDphy") +
+  labs(title = "Ferns (NS separate)", y = "", x = "")
+d <- make_scatter_with_lm(alpha_div_pteridos, "richness", "phy_mntd_obs_z", "Richness", "MNTDphy")
+e <- make_scatter_with_lm(alpha_div_ferns, "richness", "phy_mntd_obs_z", "Richness", "MNTDphy") +
+  labs(y = "")
+f <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "phy_mntd_obs_z", "Richness", "MNTDphy") +
+  labs(y = "")
+a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
+```
+
+```{r mpd-phy, fig.height = 7, fig.width = 7, fig.cap = "Phylogenetic standard effect size of mean phylogenetic distance (MPD~phy~) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude.
+Native taxa of Japan excluding hybrids."}
+# pterido mpd
+a <- make_diversity_map(
+  div_data = alpha_div_pteridos, 
+  world_map = world_map, 
+  occ_data = occ_data_pteridos, 
+  div_metric = "phy_mpd_obs_z", 
+  metric_title = "MPDphy",
+  label = "a") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
+    ))
+
+b <- make_diversity_map(
+  div_data = alpha_div_ferns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "phy_mpd_obs_z", 
+  metric_title = "MPDphy",
+  label = "b") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
+    ))
+
+c <- make_diversity_map(
+  div_data = alpha_div_ferns_ns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "phy_mpd_obs_z", 
+  metric_title = "MPDphy",
+  label = "c") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns_ns, phy_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns_ns, phy_mpd_obs_z, "abs", 3)
+    ))
+
+a + b + c + plot_layout(ncol = 2, nrow = 2)
+```
+
+```{r mntd-phy, fig.height = 7, fig.width = 7, fig.cap = "Phylogenetic standard effect size of mean nearest taxon distance (MNTD~phy~) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude.
+Native taxa of Japan excluding hybrids."}
+# pterido mpd
+a <- make_diversity_map(
+  div_data = alpha_div_pteridos, 
+  world_map = world_map, 
+  occ_data = occ_data_pteridos, 
+  div_metric = "phy_mntd_obs_z", 
+  metric_title = "MNTDphy",
+  label = "a") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
+    ))
+
+b <- make_diversity_map(
+  div_data = alpha_div_ferns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "phy_mntd_obs_z", 
+  metric_title = "MNTDphy",
+  label = "b") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
+    ))
+
+c <- make_diversity_map(
+  div_data = alpha_div_ferns_ns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "phy_mntd_obs_z", 
+  metric_title = "MNTDphy",
+  label = "c") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns_ns, phy_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns_ns, phy_mpd_obs_z, "abs", 3)
+    ))
+
+a + b + c + plot_layout(ncol = 2, nrow = 2)
+```
+
+```{r func-div-rich, fig.height = 4, fig.width = 7, fig.cap = "Relationship between functional diversity and richness.
+Functional diversity metrics include standard effect size (SES) of mean phylogenetic diversity (MPD~func~) or mean nearest taxon distance (MNTD~func~) using functional trait distances in native pteridophytes, ferns, or ferns calculated separately north and south of 30.1 latitude of Japan.
+Blue trendline shows significant linear model fit (*P* < 0.05) with 96% confidence interval in pink."}
+a <- make_scatter_with_lm(alpha_div_pteridos, "richness", "func_mpd_obs_z", "Richness", "MPDfunc") +
+  labs(title = "Pteridophytes", x = "")
+b <- make_scatter_with_lm(alpha_div_ferns, "richness", "func_mpd_obs_z", "Richness", "MPDfunc") +
+  labs(title = "Ferns", y = "", x = "")
+c <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "func_mpd_obs_z", "Richness", "MPDfunc") +
+  labs(title = "Ferns (NS separate)", y = "", x = "")
+d <- make_scatter_with_lm(alpha_div_pteridos, "richness", "func_mntd_obs_z", "Richness", "MNTDfunc")
+e <- make_scatter_with_lm(alpha_div_ferns, "richness", "func_mntd_obs_z", "Richness", "MNTDfunc") +
+  labs(y = "")
+f <- make_scatter_with_lm(alpha_div_ferns_ns, "richness", "func_mntd_obs_z", "Richness", "MNTDfunc") +
+  labs(y = "")
+a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
+```
+
+```{r mpd-func, fig.height = 7, fig.width = 7, fig.cap = "Functional standard effect size of mean phylogenetic distance (MPD~func~) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude.
+Native taxa of Japan excluding hybrids."}
+# pterido mpd
+a <- make_diversity_map(
+  div_data = alpha_div_pteridos, 
+  world_map = world_map, 
+  occ_data = occ_data_pteridos, 
+  div_metric = "func_mpd_obs_z", 
+  metric_title = "MPDfunc",
+  label = "a") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_pteridos, func_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_pteridos, func_mpd_obs_z, "abs", 3)
+    ))
+
+b <- make_diversity_map(
+  div_data = alpha_div_ferns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "func_mpd_obs_z", 
+  metric_title = "MPDfunc",
+  label = "b") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns, func_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns, func_mpd_obs_z, "abs", 3)
+    ))
+
+c <- make_diversity_map(
+  div_data = alpha_div_ferns_ns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "func_mpd_obs_z", 
+  metric_title = "MPDfunc",
+  label = "c") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns_ns, func_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns_ns, func_mpd_obs_z, "abs", 3)
+    ))
+
+a + b + c + plot_layout(ncol = 2, nrow = 2)
+```
+
+```{r mntd-func, fig.height = 7, fig.width = 7, fig.cap = "Functional standard effect size of mean nearest taxon distance (MNTD~func~) of: a) pteridophytes, b) ferns, c) ferns analyzed separately north and south of 30.1 degrees latitude.
+Native taxa of Japan excluding hybrids."}
+# pterido mpd
+a <- make_diversity_map(
+  div_data = alpha_div_pteridos, 
+  world_map = world_map, 
+  occ_data = occ_data_pteridos, 
+  div_metric = "func_mntd_obs_z", 
+  metric_title = "MNTDfunc",
+  label = "a") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_pteridos, func_mntd_obs_z, "abs", 3), 
+      get_limit(alpha_div_pteridos, func_mntd_obs_z, "abs", 3)
+    ))
+
+b <- make_diversity_map(
+  div_data = alpha_div_ferns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "func_mntd_obs_z", 
+  metric_title = "MNTDfunc",
+  label = "b") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns, func_mntd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns, func_mntd_obs_z, "abs", 3)
+    ))
+
+c <- make_diversity_map(
+  div_data = alpha_div_ferns_ns, 
+  world_map = world_map, 
+  occ_data = occ_data_ferns, 
+  div_metric = "func_mntd_obs_z", 
+  metric_title = "MNTDfunc",
+  label = "c") + 
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns_ns, func_mntd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns_ns, func_mntd_obs_z, "abs", 3)
+    ))
+
+a + b + c + plot_layout(ncol = 2, nrow = 2)
+```
+
+```{r compare-pd-sexdip, fig.cap = "Relationship between standard effect size of mean phylogenetic distance (MPD~phy~) and mean nearest taxon distance (MNTD~phy~) of native Japanese pteridophytes and percent sexual diploid taxa in each community.
+Blue trendline shows significant linear model fit with confidence interval in pink."}
+a <- make_scatter_with_lm(alpha_div_pteridos, "percent_sex_dip", "phy_mpd_obs_z", "% sexual diploid", "MPDphy") +
+  labs(title = "Pteridophytes", x = "")
+b <- make_scatter_with_lm(alpha_div_ferns, "percent_sex_dip", "phy_mpd_obs_z", "% sexual diploid", "MPD") +
+  labs(title = "Ferns", y = "", x = "")
+c <- make_scatter_with_lm(alpha_div_ferns_ns, "percent_sex_dip", "phy_mpd_obs_z", "% sexual diploid", "MPDphy") +
+  labs(title = "Ferns (NS separate)", y = "", x = "")
+d <- make_scatter_with_lm(alpha_div_pteridos, "percent_sex_dip", "phy_mntd_obs_z", "% sexual diploid", "MNTDphy")
+e <- make_scatter_with_lm(alpha_div_ferns, "percent_sex_dip", "phy_mntd_obs_z", "% sexual diploid", "MNTD") +
+  labs(y = "")
+f <- make_scatter_with_lm(alpha_div_ferns_ns, "percent_sex_dip", "phy_mntd_obs_z", "% sexual diploid", "MNTDphy") +
+  labs(y = "")
+a + b + c + d + e + f + plot_layout(ncol = 3, nrow = 2)
+```

### 2019-10-02 reports/ecostructure_geo/ecostructure_geo.Rmd
Add geographic motifs report

---
title: "Ecostructure Global Geographic Motifs, v2"
author: "Joel Nitta"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    html_preview: false
  html_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE, eval = TRUE, cache = TRUE)
```

```{r library, cache = FALSE, warning = FALSE, message = FALSE}
# Load packages
library(here)
library(scales)
library(ecostructure)
library(assertr)
library(raster)
library(drake)
library(patchwork)
library(tidyverse)

# Load cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

# Load functions
source(here::here("code/functions.R"))

# Resolve conflicts
source(here::here("code/resolve_conflicts.R"))
```

## Geographic motifs

Geographic motifs are the clusters generated by analyzing dispersion fields, i.e., the global richness for all species present in a given site.

Settings used for `ecos_fit()` were `tol` = 0.1, `K` from 2 to 10, and `num_trials` = 1.

Read in the results, check their dimensions, and have a quick peak at the data:

```{r read-results, warning = FALSE, message = FALSE}
results_vector <- glue::glue(here("results/ecosfit_motif=geo_dataset=pteridos-no-obs_K={2:10}_tol=0.1_trials=1.RDS"))

results_list <- purrr::map(results_vector, readRDS)

dim(results_list[[1]][["theta"]])
dim(results_list[[1]][["omega"]])

head(results_list[[1]][["theta"]])
head(results_list[[1]][["omega"]])
```

Theta is the contribution of each dispersion field to each of the clusters. Omega is the degree of membership to each of the clusters for each grid cell. The reason the omega and theta matrices have the same dimensions is that I dropped all cells with zero species from the dispersion fields prior to analysis.

## Omega plots

```{r omega-plots, warning = FALSE, message = FALSE}
# For easy tweaking of plot size to get pie charts
# to not overlap.
tweak <- 1.0

# Loadd all_cells to get max and min lat/longs
loadd(all_cells, cache = ja_fern_cache)

# Read in high-resolution map of Japan, convert to sf
japan <- rnaturalearth::ne_countries(country = 'japan', scale = 'large') %>% sf::st_as_sf()

# Loop through results and output pie plots
for (i in 1:length(results_list)) {
  ecos_plot_pie2(
    omega = results_list[[i]][["omega"]],
    image_width = 6000*tweak,
    image_height = 4800*tweak,
    radius = 0.05,
    bgmap_path = japan,
    long_lim = c(pull(all_cells, longitude) %>% min %>% floor, 
                 pull(all_cells, longitude) %>% max %>% ceiling),
    lat_lim =  c(pull(all_cells, latitude) %>% min %>% floor, 
                 pull(all_cells, latitude) %>% max %>% ceiling),
    path = glue::glue(here("reports/ecostructure_geo/geo_omega_k={i+1}.png"))
  )
}

```

K = 2

![](geo_omega_k=2.png)

K = 3

![](geo_omega_k=3.png)

K = 4

![](geo_omega_k=4.png)

K = 5

![](geo_omega_k=5.png)

K = 6

![](geo_omega_k=6.png)

K = 7

![](geo_omega_k=7.png)

K = 8

![](geo_omega_k=8.png)

K = 9

![](geo_omega_k=9.png)

K = 10

![](geo_omega_k=10.png)

## Theta plots

```{r theta-ggplot, warning = FALSE, message = FALSE}
# Function to make single theta plot
make_geo_theta_plot <- function (geo_theta_long, motif) {
  ggplot(ggplot2::map_data("world"), aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill = "light grey") +
  geom_raster(data = geo_theta_long, aes(fill = theta)) +
  scale_fill_gradient(low = "darkseagreen3", high = "red") +
  coord_sf(ylim = c(-50, 90), datum = NA) +
  labs(subtitle = motif) +
  jntools::blank_x_theme() +
  jntools::blank_y_theme() +
  theme_void() +
  theme(
    legend.position = "none",
    plot.background = element_rect(fill = "transparent"),
    panel.background = element_rect(fill = "transparent")
  )
}

# Convert data to long format, nested. Set K = 6
k_select = 6
geo_theta_nested <- results_list[[k_select - 1]][["theta"]] %>%
  as.data.frame %>%
  rownames_to_column("long_lat") %>%
  as_tibble %>%
  separate(long_lat, c("long", "lat"), sep = "_") %>%
  gather(motif, theta, -long, -lat) %>%
  mutate_at(vars(long, lat), as.numeric) %>%
# Check that longitude is between -180 and 180,
# that latitude is between -90 and 90, and that
# each long-lat-motif combination is unique
  verify(long > -180 & long < 180 & lat > -90 & lat < 90) %>%
  assert(is_uniq(c(.$long, .$lat, .$motif))) %>%
  group_by(motif) %>%
  nest

# Map over each theta plot separately so they each get their own
# scale.
geo_theta_plot_list <- map2(geo_theta_nested$data, geo_theta_nested$motif, make_geo_theta_plot)

# Combine into final plot and save
geo_theta_plot <- patchwork::wrap_plots(geo_theta_plot_list, ncol = 2)

ggsave(plot = geo_theta_plot, filename = here(glue::glue("reports/ecostructure_geo/geo_theta_k={k_select}.png")), width = 12.5, height = 8, units = "in")

```

![](geo_theta_k=6.png)

### 2019-10-04 reports/ecostructure_geo/ecostructure_geo.Rmd
Output combined figures for ecostructure geo report

@@
 ```{r read-results, warning = FALSE, message = FALSE}
+# Load ecostructure results
 results_vector <- glue::glue(here("results/ecosfit_motif=geo_dataset=pteridos-no-obs_K={2:10}_tol=0.1_trials=1.RDS"))
@@
 head(results_list[[1]][["omega"]])
+
+# Load all_cells to get max and min lat/longs
+loadd(all_cells, cache = ja_fern_cache)
+
+# Read in high-resolution map of Japan, convert to sf
+japan <- rnaturalearth::ne_countries(country = 'japan', scale = 'large') %>% sf::st_as_sf()
 ```
@@
 tweak <- 1.0
-
-# Loadd all_cells to get max and min lat/longs
-loadd(all_cells, cache = ja_fern_cache)
-
-# Read in high-resolution map of Japan, convert to sf
-japan <- rnaturalearth::ne_countries(country = 'japan', scale = 'large') %>% sf::st_as_sf()
 
@@
 
+make_geo_theta_plot_ja <- function (geo_theta_long, motif, occ_data) {
+  ggplot(ggplot2::map_data("world"), aes(x = long, y = lat)) +
+  geom_polygon(aes(group = group), fill = "light grey") +
+  geom_raster(data = geo_theta_long, aes(fill = theta)) +
+  scale_fill_gradient(low = "darkseagreen3", high = "red") +
+  coord_quickmap(
+      xlim = c(pull(occ_data, longitude) %>% min %>% floor, 
+               pull(occ_data, longitude) %>% max %>% ceiling),
+      ylim = c(pull(occ_data, latitude) %>% min %>% floor, 
+               pull(occ_data, latitude) %>% max %>% ceiling)
+    ) +
+  labs(subtitle = motif) +
+  jntools::blank_x_theme() +
+  jntools::blank_y_theme() +
+  theme_void() +
+  theme(
+    legend.position = "none",
+    plot.background = element_rect(fill = "transparent"),
+    panel.background = element_rect(fill = "transparent")
+  )
+}
+
 # Convert data to long format, nested. Set K = 6
 k_select = 6
+
+# Global
 geo_theta_nested <- results_list[[k_select - 1]][["theta"]] %>%
@@
 # scale.
+# - Global plots
 geo_theta_plot_list <- map2(geo_theta_nested$data, geo_theta_nested$motif, make_geo_theta_plot)
+
+# - Japan-focused plots
+geo_theta_plot_list_js <- 
+  list(
+    geo_theta_long = geo_theta_nested$data,
+    motif = geo_theta_nested$motif,
+    occ_data = list(all_cells)
+  ) %>%
+  pmap(make_geo_theta_plot_ja)
 
@@
 ![](geo_theta_k=6.png)
+
+## Combined plots
+
+```{r combined-plot}
+make_geo_theta_plot_color <- function (geo_theta_long, border_color) {
+  ggplot(ggplot2::map_data("world"), aes(x = long, y = lat)) +
+  geom_polygon(aes(group = group), fill = "light grey") +
+  geom_raster(data = geo_theta_long, aes(fill = theta)) +
+  scale_fill_gradient(low = "darkseagreen3", high = "red") +
+  coord_sf(ylim = c(-50, 90), datum = NA) +
+  jntools::blank_x_theme() +
+  jntools::blank_y_theme() +
+  theme_void() +
+  theme(
+    legend.position = "none",
+    plot.background = element_rect(fill = "transparent"),
+    panel.background = element_rect(fill = "transparent"),
+    panel.border = element_rect(colour = border_color, fill=NA, size=2)
+  )
+}
+
+k_select <- 4
+
+cols <- RColorBrewer::brewer.pal(k_select, name = "Set1")
+
+ecos_plot_pie2(
+    omega = results_list[[k_select - 1]][["omega"]],
+    image_width = 6000*tweak,
+    image_height = 4800*tweak,
+    radius = 0.05,
+    color = cols,
+    bgmap_path = japan,
+    long_lim = c(pull(all_cells, longitude) %>% min %>% floor, 
+                 pull(all_cells, longitude) %>% max %>% ceiling),
+    lat_lim =  c(pull(all_cells, latitude) %>% min %>% floor, 
+                 pull(all_cells, latitude) %>% max %>% ceiling),
+    path = glue::glue(here("reports/ecostructure_geo/geo_omega_combined_k={k_select}.png"))
+  )
+
+geo_theta_nested <- results_list[[k_select - 1]][["theta"]] %>%
+  as.data.frame %>%
+  rownames_to_column("long_lat") %>%
+  as_tibble %>%
+  separate(long_lat, c("long", "lat"), sep = "_") %>%
+  gather(motif, theta, -long, -lat) %>%
+  mutate_at(vars(long, lat), as.numeric) %>%
+# Check that longitude is between -180 and 180,
+# that latitude is between -90 and 90, and that
+# each long-lat-motif combination is unique
+  verify(long > -180 & long < 180 & lat > -90 & lat < 90) %>%
+  assert(is_uniq(c(.$long, .$lat, .$motif))) %>%
+  group_by(motif) %>%
+  nest
+
+geo_theta_plot_list_colors <- 
+  list(
+    geo_theta_long = geo_theta_nested$data,
+    border_color = cols
+  ) %>%
+  pmap(make_geo_theta_plot_color)
+
+# Combine into final plot and save
+geo_theta_plot_colors <- patchwork::wrap_plots(geo_theta_plot_list_colors, ncol = 1)
+
+ggsave(plot = geo_theta_plot_colors, filename = here(glue::glue("reports/ecostructure_geo/geo_theta_colors_k={k_select}.png")), width = 7, height = 8, units = "in")
+
+```

### 2019-10-04 reports/ecostructure_geo/ecostructure_geo.Rmd
Output k=3 for the combined plot

@@
 ```{r combined-plot}
+# For easy tweaking of plot size to get pie charts
+# to not overlap.
+tweak <- 1.5
+
 make_geo_theta_plot_color <- function (geo_theta_long, border_color) {
@@
 
-k_select <- 4
+k_select <- 3
 
-cols <- RColorBrewer::brewer.pal(k_select, name = "Set1")
+cols <- RColorBrewer::brewer.pal(k_select, name = "Dark2")
 
@@
     image_height = 4800*tweak,
-    radius = 0.05,
+    radius = 0.04,
     color = cols,

### 2020-02-04 presentations/jsps_2020-03-01/index.Rmd
Add Rmd to generate slides for JSPS conference talk

---
output:
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom.css"]
    lib_dir: libs
    mathjax: default
    seal: false
    nature:
      slideNumberFormat: "" # To turn off slide numbers
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["left", "middle", "inverse"]
editor_options:
  chunk_output_type: console
---
class: inverse, middle, title-slide, big-margin
<!-- background-position: 50% 50%
background-image: url(images/title_slide_background.svg)
background-size: contain -->

# 日本のシダ植物多様性の多次元解析
<br>
## 新田ジョエル^1^、海老原淳^2^

^1^スミソニアン国立自然史博物館植物研究部<br>
<span style = 'font-size: 80%;'>https://joelnitta.com</span>

^2^国立科学博物館植物研究部

日本植物分類学会第19回大会<br><span style = 'font-size: 80%;'>2020.03.01</span>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE, eval = TRUE, cache = TRUE, dev = "svg")
```

```{r load-packages, cache = FALSE}
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(tidyverse)
```

```{r load-data, cache = FALSE}
# Set cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

source(here::here("code/functions.R"))

# Load targets
loadd(list = c(
  "all_cells",
  "alpha_div_ferns",
  "occ_data_ferns",
  "ja_bioregions_ferns",
  "biodiv_results_pteridos",
  "species_motifs_ferns_6L",
  "percent_sex_dip_ferns"), 
  cache = ja_fern_cache)

```

---
class: big-margin

## Biodiversity is in danger from climate change

---
class: big-margin

## We must understand the distribution of biodiversity to save it

- Most analyses of BD only include richness
- But BD is more complicated: evo history also important

---
class: big-margin

## Japan is an ideal site to study fern BD

- Extremely well sampled

---
class: big-margin

## Goals

- Map biodiversity (species richness, PD, endemism)
- Identify biogeographical realms
- Understand drivers of biogeo structure

---
class: big-margin

## Data

- Occurrence data of Japanese fern specimens
- Phylogenetic tree (*rbcL*)
- Morphological traits
- Climate data

---

## Species richness peaks on Yakushima Isl.

```{r richness-plot, out.width = "120%", fig.asp = 0.7}
alpha_div_ferns <-
  alpha_div_ferns %>%
  filter(latitude > 21)

alpha_div_ferns_binned <-
  alpha_div_ferns %>%
  mutate(
    latitude = cut_width(latitude, width = 0.1, closed = "left") %>%
      str_split(",") %>%
      map_chr(1) %>%
      parse_number) %>%
  group_by(latitude) %>%
  summarize_at(
    vars(richness, phy_mpd_obs_z, func_mpd_obs_z), mean, na.rm = TRUE
  )

# Fern richness
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_scico(palette = "bamako", na.value="grey") +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

b <-
  ggplot(alpha_div_ferns_binned, aes(y = latitude, x = richness)) +
  geom_path() +
  labs(
    x = "No. species"
  ) + 
  ylim(24, 46) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )

a + b + plot_layout(widths = c(3,1))

```

---

## Phylogenetic diversity is high in N and S

```{r mpd-plot, out.width = "120%", fig.asp = 0.7}

# pterido_phy_mpd
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "phy_mpd_obs_z", 
  metric_title = "MPD",
  label = "Mean Phylogenetic Distance") +
  scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
    )) +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

b <- ggplot(
  alpha_div_ferns_binned, 
  aes(y = latitude, x = phy_mpd_obs_z)) +
  geom_path() +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    position = "right") +
  labs(
    subtitle = "",
    x = "MPD"
  ) + 
  annotate(
    "segment", 
    x = 0, xend = 0, 
    y = -Inf, yend = Inf, 
    linetype = 2) + 
  ylim(24, 46) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )

a + b + plot_layout(widths = c(3,1))

```

---

## Highest bioregional diversity in S islands

```{r mapequation-prep, include = FALSE}
# Convert all_cells dataframe to sf object
all_cells_sf <-
  purrr::map2(all_cells$longitude, all_cells$latitude, ~st_point(c(.x, .y))) %>%
  st_sfc(crs = 4326) %>%
  st_sf(all_cells, .)

# Join secondary grid cells with bioregions
all_cells_regions <-sf::st_join(
  all_cells_sf,
  ja_bioregions_ferns) %>% 
  select(secondary_grid_code, longitude, latitude, bioregion) %>% 
  as_tibble %>%
  mutate(bioregion = as.factor(bioregion)) %>%
  filter(secondary_grid_code %in% occ_data_ferns$secondary_grid_code)
```

```{r mapequation-plot, out.width = "120%", fig.asp = 0.7}
ggplot(all_cells_regions, aes(x = longitude, y = latitude, fill = bioregion)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_brewer(palette = "Set1") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Bioregion") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )

```

---

## Honshu has significant ecological structure

```{r ecostructure}
# For easy manipulation of plot size
tweak = 0.5

# Output pie plot without basemap
ecos_plot_pie2(
  omega =  species_motifs_ferns_6L$omega,
  image_width = 6000*tweak,
  image_height = 4800*tweak,
  radius = 0.05,
  no_map = TRUE,
  blank_bg = "white",
  long_lim = c(122, 146),
  lat_lim = c(24, 46),
  path = here::here("presentations/jsps_2020-03-01/images/species_motifs_ferns_6L.png")
)

# crop image, save as species_motifs_ferns_6L_cropped.png
```

.center[<img src="images/species_motifs_ferns_6L_cropped.png" alt="drawing" height="600">]

---

## S islands have high rates of endemism

```{r canape-plot, out.width = "120%", fig.asp = 0.7}
ggplot(biodiv_results_pteridos, aes(x = longitude, y = latitude, fill = canape)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_brewer(palette = "Set1", na.value = "grey80") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Endemism") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )
```

---

## Neoendemic areas have more sexual diploids

```{r canape-sex, out.width = "120%", fig.asp = 0.7}

percent_sex_dip_ferns_round <-
  percent_sex_dip_ferns %>% 
  mutate(latitude = round(latitude, 3)) %>%
  mutate(longitude = round(longitude, 3))

canape_sex_mode <-
  biodiv_results_pteridos %>%
  mutate(latitude = round(latitude, 3)) %>%
  mutate(longitude = round(longitude, 3)) %>%
  inner_join(
    percent_sex_dip_ferns_round,
    by = c("latitude", "longitude")
  ) %>%
  assert(is_uniq, secondary_grid_code) %>%
  filter(!is.na(canape))

# Run analysis of variance (AOV) on
# latitudinal breadth by reproductive mode.
canape_by_repro_model = aov(
  percent_sex_dip ~ canape,
  data = canape_sex_mode)

canape_by_repro_model_summary = tidy(canape_by_repro_model)

# AOV of latidudinal breadth by reproductive mode
# showed a significant difference, so
# run Tukey HSD test on results.
canape_by_repro_tukey = TukeyHSD(canape_by_repro_model) %>% tidy

ggplot(canape_sex_mode, aes(x = canape, y = percent_sex_dip)) +
  geom_jitter() +
  geom_boxplot() +
  coord_flip()

```

### 2020-02-05 presentations/jsps_2020-03-01/index.Rmd
Update slides
- remove infomaps bioregions

@@
 class: inverse, middle, title-slide, big-margin
-<!-- background-position: 50% 50%
-background-image: url(images/title_slide_background.svg)
-background-size: contain -->
+background-color: #000000
+background-position: 50% 50%
+background-image: url(images/title_background.png)
+background-size: contain
 
-# 日本のシダ植物多様性の多次元解析
+# 日本のシダ植物多様性の<br>多次元解析
 <br>
-## 新田ジョエル^1^、海老原淳^2^
+## 新田ジョエル<sup>1</sup>、海老原淳<sup>2</sup>
 
-^1^スミソニアン国立自然史博物館植物研究部<br>
+<sup>1</sup>スミソニアン国立自然史博物館・植物研究部<br>
 <span style = 'font-size: 80%;'>https://joelnitta.com</span>
 
-^2^国立科学博物館植物研究部
+<sup>2</sup>国立科学博物館・植物研究部
 
@@
 ```{r load-packages, cache = FALSE}
+library(magick)
+library(gt)
 library(drake)
@@
 loadd(list = c(
+  "taxon_id_map",
   "all_cells",
@@
   "ja_bioregions_ferns",
-  "biodiv_results_pteridos",
+  "biodiv_results_ferns",
   "species_motifs_ferns_6L",
@@
   label = "Richness") +
-  scale_fill_scico(palette = "bamako", na.value="grey") +
+  scale_fill_viridis_c() +
   scale_y_continuous(
@@
 a + b + plot_layout(widths = c(3,1))
+
+```
+
+```{r title-slide-image}
+a <- make_diversity_map(
+  div_data = alpha_div_ferns, 
+  occ_data = occ_data_ferns, 
+  div_metric = "richness", 
+  metric_title = "No. species ",
+  label = "Richness") +
+  scale_fill_viridis_c() +
+  theme_void() +
+  theme(
+    plot.title = element_blank(),
+    plot.subtitle = element_blank(),
+    legend.position = "none", 
+     panel.background = element_rect(fill = "black"),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+  )
+
+ggsave(plot = a, here::here("presentations/jsps_2020-03-01/images/title_slide_part_a.png"))
+
+b <- make_diversity_map(
+  div_data = alpha_div_ferns, 
+  occ_data = occ_data_ferns, 
+  div_metric = "phy_mpd_obs_z", 
+  metric_title = "MPD",
+  label = "Mean Phylogenetic Distance") +
+  scale_fill_scico(
+    palette = "vik", 
+    na.value="dark grey",
+    limits = c(
+      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
+      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
+    )) +
+  scale_y_continuous(
+    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
+    limits = c(24, 46)) +
+  theme_void() +
+  theme(
+    plot.title = element_blank(),
+    plot.subtitle = element_blank(),
+    legend.position = "none", 
+     panel.background = element_rect(fill = "black"),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),
+  )
+
+ggsave(plot = b, here::here("presentations/jsps_2020-03-01/images/title_slide_part_b.png"))
 
@@
 
-## Highest bioregional diversity in S islands
+## S islands have high rates of endemism
 
-```{r mapequation-prep, include = FALSE}
-# Convert all_cells dataframe to sf object
-all_cells_sf <-
-  purrr::map2(all_cells$longitude, all_cells$latitude, ~st_point(c(.x, .y))) %>%
-  st_sfc(crs = 4326) %>%
-  st_sf(all_cells, .)
+```{r canape-plot, out.width = "120%", fig.asp = 0.7}
+canape_cols <- RColorBrewer::brewer.pal(4, "Dark2") %>%
+  set_names(c("mixed", "neo", "palaeo", "super"))
 
-# Join secondary grid cells with bioregions
-all_cells_regions <-sf::st_join(
-  all_cells_sf,
-  ja_bioregions_ferns) %>% 
-  select(secondary_grid_code, longitude, latitude, bioregion) %>% 
-  as_tibble %>%
-  mutate(bioregion = as.factor(bioregion)) %>%
-  filter(secondary_grid_code %in% occ_data_ferns$secondary_grid_code)
-```
-
-```{r mapequation-plot, out.width = "120%", fig.asp = 0.7}
-ggplot(all_cells_regions, aes(x = longitude, y = latitude, fill = bioregion)) +
+ggplot(biodiv_results_ferns, aes(x = longitude, y = latitude, fill = canape)) +
   geom_tile() +
@@
   ) +
-  scale_fill_brewer(palette = "Set1") +
-    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  labs(fill = "Bioregion") +
-  theme(
-    panel.background = element_blank(),
-      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
-      panel.grid.minor = element_blank(),
-      axis.title = element_blank(),
-    axis.ticks = element_blank()
-  )
-
-```
-
----
-
-## Honshu has significant ecological structure
-
-```{r ecostructure}
-# For easy manipulation of plot size
-tweak = 0.5
-
-# Output pie plot without basemap
-ecos_plot_pie2(
-  omega =  species_motifs_ferns_6L$omega,
-  image_width = 6000*tweak,
-  image_height = 4800*tweak,
-  radius = 0.05,
-  no_map = TRUE,
-  blank_bg = "white",
-  long_lim = c(122, 146),
-  lat_lim = c(24, 46),
-  path = here::here("presentations/jsps_2020-03-01/images/species_motifs_ferns_6L.png")
-)
-
-# crop image, save as species_motifs_ferns_6L_cropped.png
-```
-
-.center[<img src="images/species_motifs_ferns_6L_cropped.png" alt="drawing" height="600">]
-
----
-
-## S islands have high rates of endemism
-
-```{r canape-plot, out.width = "120%", fig.asp = 0.7}
-ggplot(biodiv_results_pteridos, aes(x = longitude, y = latitude, fill = canape)) +
-  geom_tile() +
-  coord_quickmap(
-    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
-             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
-    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
-             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
-  ) +
-  scale_fill_brewer(palette = "Set1", na.value = "grey80") +
+  scale_fill_manual(values = canape_cols, na.value = "grey80") +
     scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
@@
 
-```{r canape-sex, out.width = "120%", fig.asp = 0.7}
+<br>
+
+.center[
+```{r canape-sex, fig.asp = 0.7}
 
@@
 canape_sex_mode <-
-  biodiv_results_pteridos %>%
+  biodiv_results_ferns %>%
   mutate(latitude = round(latitude, 3)) %>%
@@
 
-ggplot(canape_sex_mode, aes(x = canape, y = percent_sex_dip)) +
-  geom_jitter() +
-  geom_boxplot() +
-  coord_flip()
+canape_sex_mode %>%
+  group_by(canape) %>%
+  summarize(
+    mean_sex_dip = mean(percent_sex_dip, na.rm = TRUE),
+    sd_sex_dip = sd(percent_sex_dip, na.rm = TRUE)
+  ) %>%
+  mutate(
+    upper = mean_sex_dip + sd_sex_dip,
+    lower = mean_sex_dip - sd_sex_dip
+  ) %>%
+  mutate(canape = fct_reorder(canape, mean_sex_dip)) %>%
+  ggplot(aes(x = canape, y = mean_sex_dip, fill = canape)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::percent) +
+  scale_fill_manual(values = canape_cols, na.value = "grey80", guide = FALSE) +
+  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25) +
+  labs(
+    x = "Type of endemicity",
+    y = "% sexual diploids"
+  ) +
+  jntools::standard_theme()
 
 ```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+## Ecological structure of Japanese ferns, *K* = 6
+
+```{r ecos-map-prep}
+# Write out ecostructure plots using modified version of ecos_plot_pie(),
+# then crop them to the right size.
+# (ggscatterpie is too slow, can't map aesthetics)
+
+# Set a base tweaking factor for easy manipulation of plot size
+tweak = 0.5
+
+# Use color brewer "Set1" for the 6 motifs (k=6)
+ecos_cols <- RColorBrewer::brewer.pal(6, "Set1")
+
+# Define a function to dim all other colors (using alpha)
+# besides the focus color
+apply_focus <- function (cols, focus, alpha = 0.15) {
+  new_cols <- scales::alpha(cols, alpha)
+  new_cols[[focus]] <- cols[[focus]]
+  new_cols
+}
+
+# Make list of colors to use for writing out plots
+cols_list <- purrr::map(1:6, ~apply_focus(ecos_cols, .)) %>%
+  c(list(ecos_cols), .) %>%
+  set_names(c("all", paste0("focus_", 1:6)))
+
+# Loop over outputting pie plot without basemap
+purrr::walk2(
+  cols_list,
+  names(cols_list),
+  ~ecos_plot_pie2(
+    omega =  species_motifs_ferns_6L$omega,
+    color = .x,
+    image_width = 6000*tweak,
+    image_height = 4800*tweak,
+    radius = 0.05,
+    no_map = TRUE,
+    blank_bg = "black",
+    long_lim = c(122, 146),
+    lat_lim = c(24, 46),
+    path = here::here(
+      glue::glue("presentations/jsps_2020-03-01/images/species_motifs_ferns_{.y}.png")
+    )
+  )
+)
+
+# Define a function to crop ecostructure map plot
+crop_ecos_plot <- function (image_path) {
+
+  image_read(image_path) %>%
+    # Crops to 2410x2410, starting 525 to the right and 150 down from UL corner
+    image_crop("2140x2140+525+150") %>%
+    image_write(
+      path = image_path %>% 
+        fs::path_ext_remove() %>%
+        paste0("_crop.png")
+    )
+}
+
+# Loop over plots and crop them
+here::here(glue::glue("presentations/jsps_2020-03-01/images/species_motifs_ferns_{names(cols_list)}.png")) %>%
+  walk(crop_ecos_plot)
+```
+
+```{r ecos-species-prep}
+# Make overall dataframe of species contributions to motifs based on theta
+species_ranking <-
+species_motifs_ferns_6L$theta %>%
+  as.data.frame %>%
+  rownames_to_column("taxon_id") %>%
+  as_tibble %>%
+  pivot_longer(-taxon_id, names_to = "motif", values_to = "score") %>%
+  left_join(taxon_id_map, by = "taxon_id") %>% 
+  select(motif, taxon, score) %>%
+  mutate(taxon = str_replace_all(taxon, "_", " ")) %>%
+  mutate(score = round(score, 2))
+
+# Define function for making species ranking table
+make_species_ranking_table <- function (df, motif_select, slice_n = 15) {
+
+  df %>%
+  filter(motif == motif_select) %>%
+  arrange(desc(score), taxon) %>% 
+  slice(1:slice_n) %>%
+  select(taxon, score) %>%
+  gt() %>%
+  tab_options(
+    row.striping.include_table_body = FALSE,
+    row.striping.include_stub = FALSE,
+    table.background.color = "#000000",
+    column_labels.background.color = "#000000") %>%
+  tab_style(
+    style = list(
+      cell_fill(color = "#000000"),
+      cell_text(color = "white", size = "medium")
+      ),
+    locations = cells_column_labels(columns = everything())
+    ) %>%
+  tab_style(
+    style = list(
+      cell_fill(color = "#000000"),
+      cell_text(color = "white", size = "small")
+      ),
+    locations = cells_data(
+      columns = everything())
+  ) %>%
+  tab_style(
+    style = list(
+      cell_text(style = "italic")
+      ),
+    locations = cells_data(
+      columns = vars(taxon))
+  )
+}
+
+```
+
+.center[<img src="images/species_motifs_ferns_all_crop.png" alt="drawing" height = "600">]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_2_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-ranking-2}
+make_species_ranking_table(species_ranking, 2)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_3_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-ranking-3}
+make_species_ranking_table(species_ranking, 3)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_4_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-ranking-4}
+make_species_ranking_table(species_ranking, 4)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_5_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-ranking-5}
+make_species_ranking_table(species_ranking, 5)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_6_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-ranking-6}
+make_species_ranking_table(species_ranking, 6)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_1_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-ranking-1}
+make_species_ranking_table(species_ranking, 1)
+```
+]

### 2020-02-05 code/functions.R

@@ -1739,7 +1739,7 @@ ecos_plot_pie2 <- function (
   invisible(lapply(1:dim(omega)[1], function(r) do.call(mapplots::add.pie,
                                                         append(list(z = as.integer(100 * omega[r, ]), x = coords[r,
                                                                                                                  1], y = coords[r, 2], labels = c("", "", ""), radius = radius,
-                                                                    col = sapply(color, scales::alpha, intensity)), pie_control))))
+                                                                    color), pie_control))))
   invisible(dev.off())
 }
 

### 2020-02-06 presentations/jsps_2020-03-01/index.Rmd
Update presentation slides
- add inset distribution maps for ecos motifs

@@
 library(gt)
+library(ggtext)
 library(drake)
@@
   "occ_data_ferns",
+  "occ_point_data_raw",
   "ja_bioregions_ferns",
@@
 
----
-class: big-margin
-
-## Biodiversity is in danger from climate change
-
----
-class: big-margin
-
-## We must understand the distribution of biodiversity to save it
-
-- Most analyses of BD only include richness
-- But BD is more complicated: evo history also important
-
----
-class: big-margin
-
-## Japan is an ideal site to study fern BD
-
-- Extremely well sampled
-
----
-class: big-margin
-
-## Goals
-
-- Map biodiversity (species richness, PD, endemism)
-- Identify biogeographical realms
-- Understand drivers of biogeo structure
-
----
-class: big-margin
-
-## Data
-
-- Occurrence data of Japanese fern specimens
-- Phylogenetic tree (*rbcL*)
-- Morphological traits
-- Climate data
-
----
-
-## Species richness peaks on Yakushima Isl.
-
-```{r richness-plot, out.width = "120%", fig.asp = 0.7}
+```{r alpha-div-ferns-prep}
+# Add ecostructure motif scores and canape results to alpha diversity,
+# filter out a single location south of 21 lat that had no ferns
 alpha_div_ferns <-
-  alpha_div_ferns %>%
+  species_motifs_ferns_6L$omega %>%
+  as.data.frame %>%
+  rownames_to_column("site") %>%
+  as_tibble %>%
+  separate(site, c("longitude", "latitude"), sep = "_") %>%
+  left_join(
+    mutate_at(alpha_div_ferns, vars(longitude, latitude), as.character),
+    by = c("longitude", "latitude")) %>%
+  left_join(
+    mutate_at(biodiv_results_ferns, vars(longitude, latitude), as.character),
+    by = c("longitude", "latitude")) %>%
+  mutate_at(vars(longitude, latitude), as.numeric) %>%
+  rename_at(vars(1:6 %>% as.character), ~paste0("motif_", .)) %>%
   filter(latitude > 21)
-
-alpha_div_ferns_binned <-
-  alpha_div_ferns %>%
-  mutate(
-    latitude = cut_width(latitude, width = 0.1, closed = "left") %>%
-      str_split(",") %>%
-      map_chr(1) %>%
-      parse_number) %>%
-  group_by(latitude) %>%
-  summarize_at(
-    vars(richness, phy_mpd_obs_z, func_mpd_obs_z), mean, na.rm = TRUE
-  )
-
-# Fern richness
-a <- make_diversity_map(
-  div_data = alpha_div_ferns, 
-  occ_data = occ_data_ferns, 
-  div_metric = "richness", 
-  metric_title = "No. species ",
-  label = "Richness") +
-  scale_fill_viridis_c() +
-  scale_y_continuous(
-    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
-    limits = c(24, 46)) +
-  theme(
-    plot.subtitle = element_blank())
-
-b <-
-  ggplot(alpha_div_ferns_binned, aes(y = latitude, x = richness)) +
-  geom_path() +
-  labs(
-    x = "No. species"
-  ) + 
-  ylim(24, 46) +
-  theme(
-    panel.background = element_blank(),
-    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
-    panel.grid.minor = element_blank(),
-    axis.title.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.y = element_blank()
-  )
-
-a + b + plot_layout(widths = c(3,1))
-
 ```
 
-```{r title-slide-image}
+```{r title-slide-image, eval = FALSE}
+# Code to make title slide image, don't need to run when rendering slides
+
 a <- make_diversity_map(
@@
 ---
+class: big-margin
+
+## Biodiversity is in danger from climate change
+
+---
+class: big-margin
+
+## We must understand the distribution of biodiversity to save it
+
+- Most analyses of BD only include richness
+- But BD is more complicated: evo history also important
+
+---
+class: big-margin
+
+## Japan is an ideal site to study fern BD
+
+- Extremely well sampled
+
+---
+class: big-margin
+
+## Goals
+
+- Map biodiversity (species richness, PD, endemism)
+- Identify biogeographical realms
+- Understand drivers of biogeo structure
+
+---
+class: big-margin
+
+## Data
+
+- Occurrence data of Japanese fern specimens
+- Phylogenetic tree (*rbcL*)
+- Morphological traits
+- Climate data
+
+---
+
+## Species richness peaks on Yakushima Isl.
+
+```{r richness-plot, out.width = "120%", fig.asp = 0.7}
+# Fern richness
+# - map
+a <- make_diversity_map(
+  div_data = alpha_div_ferns, 
+  occ_data = occ_data_ferns, 
+  div_metric = "richness", 
+  metric_title = "No. species ",
+  label = "Richness") +
+  scale_fill_viridis_c() +
+  scale_y_continuous(
+    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
+    limits = c(24, 46)) +
+  theme(
+    plot.subtitle = element_blank())
+
+# - line tracking mean richness
+alpha_div_ferns_binned <-
+  alpha_div_ferns %>%
+  mutate(
+    latitude = cut_width(latitude, width = 0.1, closed = "left") %>%
+      str_split(",") %>%
+      map_chr(1) %>%
+      parse_number) %>%
+  group_by(latitude) %>%
+  summarize_at(
+    vars(richness, phy_mpd_obs_z, func_mpd_obs_z), mean, na.rm = TRUE
+  )
+
+b <-
+  ggplot(alpha_div_ferns_binned, aes(y = latitude, x = richness)) +
+  geom_path() +
+  labs(
+    x = "No. species"
+  ) + 
+  ylim(24, 46) +
+  theme(
+    panel.background = element_blank(),
+    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank()
+  )
+
+a + b + plot_layout(widths = c(3,1))
+```
+
+---
 
@@
 ```{r mpd-plot, out.width = "120%", fig.asp = 0.7}
-
-# pterido_phy_mpd
+# Phylogenetic diversity (MPD)
+# - map
 a <- make_diversity_map(
@@
 
+# line tracking mean MPD
 b <- ggplot(
@@
 a + b + plot_layout(widths = c(3,1))
-
 ```
@@
   mutate(taxon = str_replace_all(taxon, "_", " ")) %>%
-  mutate(score = round(score, 2))
+  mutate(score = round(score, 2)) %>%
+  group_by(taxon) %>%
+  mutate(score_sd = sd(score)) %>%
+  ungroup
 
 # Define function for making species ranking table
-make_species_ranking_table <- function (df, motif_select, slice_n = 15) {
+make_species_ranking_table <- function (ranking_data, motif_select, taxon_select = NULL, slice_n = 10) {
 
-  df %>%
+  ranking_data_sliced <- ranking_data %>%
     filter(motif == motif_select) %>%
-  arrange(desc(score), taxon) %>% 
+    # Select top-scoring species with high SD
+    # (high SD should indicate mostly contributing to
+    # a single motif)
+    arrange(desc(score), desc(score_sd)) %>%
     slice(1:slice_n) %>%
     select(taxon, score) %>%
+    mutate(score = as.factor(score)) %>%
+    # Arrange alphabetically within score
+    arrange(desc(score), taxon)
+
+  table <-
+    ranking_data_sliced %>%
     gt() %>%
@@
     )
+
+  if(!is.null(taxon_select))
+    table <-
+      table %>%
+      tab_style(
+        style = list(
+          cell_fill(
+            color = "red",
+            alpha = 1.0
+          )
+        ),
+        locations = cells_data(
+          columns = everything(),
+          rows = taxon == ranking_data_sliced$taxon[[taxon_select]])
+      )
+
+  table
+
 }
 
+# Define function to make simple distribution map for a species
+plot_distribution <- function (ranking_data, motif_select, taxon_select, occ_data, show_title = TRUE) {
+
+   ranking_data <- ranking_data %>%
+    filter(motif == motif_select) %>%
+    arrange(desc(score), desc(score_sd)) %>%
+    select(taxon, score) %>%
+    mutate(score = as.factor(score)) %>%
+    arrange(desc(score), taxon)
+
+  cells_present <-
+    occ_data %>%
+    filter(taxon_name == ranking_data$taxon[[taxon_select]]) %>%
+    select(secondary_grid_code) %>%
+    unique %>%
+    mutate(present = TRUE)
+
+  plot_data <-
+    occ_data %>%
+    select(secondary_grid_code, latitude, longitude) %>%
+    left_join(cells_present) %>%
+    mutate(present = replace_na(present, FALSE))
+
+  plot <- ggplot(plot_data, aes(x = longitude, y = latitude, fill = present)) +
+    geom_tile() +
+    coord_quickmap(
+      xlim = c(pull(occ_data, longitude) %>% min %>% floor,
+               pull(occ_data, longitude) %>% max %>% ceiling),
+      ylim = c(pull(occ_data, latitude) %>% min %>% floor,
+               pull(occ_data, latitude) %>% max %>% ceiling)
+    ) +
+    scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "grey80"), na.value = "grey80", guide = FALSE) +
+    theme_void()
+
+  if(isTRUE(show_title))
+    plot <- plot +
+    labs(
+      title = ranking_data$taxon[[taxon_select]]
+    )
+
+  plot
+}
+```
+
+```{r write_out_motif_maps, eval = FALSE}
+# Don't run when making slides, but it's a useful practice to check the
+# distributions for the top 10 taxa for each motif
+plot_dist_by_motif <- function(species_ranking, motif, occ_data_ferns) {
+  dist_maps <- purrr::map(1:10, ~plot_distribution(species_ranking, motif, ., occ_data_ferns)) %>%
+    patchwork::wrap_plots(ncol = 2, nrow = 5)
+
+  ggsave(plot = dist_maps, file = glue::glue("motif_{motif}_maps.png"), width = 6, height = 12, units = "in")
+}
+
+walk(1:6, ~plot_dist_by_motif(species_ranking, ., occ_data_ferns))
 ```
@@
 
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_1_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-1}
+make_species_ranking_table(species_ranking, 1)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_1_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-1-focus-3}
+make_species_ranking_table(species_ranking, 1, 3)
+plot_distribution(species_ranking, 1, 3, occ_data_ferns, show_title = FALSE)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_1_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-1-focus-7}
+make_species_ranking_table(species_ranking, 1, 7)
+plot_distribution(species_ranking, 1, 7, occ_data_ferns, show_title = FALSE)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
 .pull-left-wide[<img src="images/species_motifs_ferns_focus_2_crop.png" alt="drawing">]
@@
 .pull-right-narrow[
-```{r ecos-ranking-2}
+```{r ecos-motif-2}
 make_species_ranking_table(species_ranking, 2)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_2_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-2-focus-1}
+make_species_ranking_table(species_ranking, 2, 1)
+plot_distribution(species_ranking, 2, 1, occ_data_ferns, show_title = FALSE)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_2_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-2-focus-2}
+make_species_ranking_table(species_ranking, 2, 2)
+plot_distribution(species_ranking, 2, 2, occ_data_ferns, show_title = FALSE)
 ```
@@
 .pull-right-narrow[
-```{r ecos-ranking-3}
+```{r ecos-motif-3}
 make_species_ranking_table(species_ranking, 3)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_3_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-3-focus-3}
+make_species_ranking_table(species_ranking, 3, 3)
+plot_distribution(species_ranking, 3, 3, occ_data_ferns, show_title = FALSE)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_3_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-3-focus-7}
+make_species_ranking_table(species_ranking, 3, 7)
+plot_distribution(species_ranking, 3, 7, occ_data_ferns, show_title = FALSE)
 ```
@@
 .pull-right-narrow[
-```{r ecos-ranking-4}
+```{r ecos-motif-4}
 make_species_ranking_table(species_ranking, 4)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_4_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-4-focus-2}
+make_species_ranking_table(species_ranking, 4, 2)
+plot_distribution(species_ranking, 4, 2, occ_data_ferns, show_title = FALSE)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_4_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-4-focus-8}
+make_species_ranking_table(species_ranking, 4, 8)
+plot_distribution(species_ranking, 4, 8, occ_data_ferns, show_title = FALSE)
 ```
@@
 .pull-right-narrow[
-```{r ecos-ranking-5}
+```{r ecos-motif-5}
 make_species_ranking_table(species_ranking, 5)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_5_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-5-focus-3}
+make_species_ranking_table(species_ranking, 5, 3)
+plot_distribution(species_ranking, 5, 3, occ_data_ferns, show_title = FALSE)
+```
+]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_5_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-5-focus-7}
+make_species_ranking_table(species_ranking, 5, 7)
+plot_distribution(species_ranking, 5, 7, occ_data_ferns, show_title = FALSE)
 ```
@@
 .pull-right-narrow[
-```{r ecos-ranking-6}
+```{r ecos-motif-6}
 make_species_ranking_table(species_ranking, 6)
@@
 
-.pull-left-wide[<img src="images/species_motifs_ferns_focus_1_crop.png" alt="drawing">]
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_6_crop.png" alt="drawing">]
 
 .pull-right-narrow[
-```{r ecos-ranking-1}
-make_species_ranking_table(species_ranking, 1)
+```{r ecos-motif-6-focus-1}
+make_species_ranking_table(species_ranking, 6, 1)
+plot_distribution(species_ranking, 6, 1, occ_data_ferns, show_title = FALSE)
 ```
 ]
+
+---
+class: inverse, middle
+background-color: #000000
+
+.pull-left-wide[<img src="images/species_motifs_ferns_focus_6_crop.png" alt="drawing">]
+
+.pull-right-narrow[
+```{r ecos-motif-6-focus-2}
+make_species_ranking_table(species_ranking, 6, 2)
+plot_distribution(species_ranking, 6, 2, occ_data_ferns, show_title = FALSE)
+```
+]
+
+---
+
+## Motif 2 tends to cluster, motif 6 tends to overdisperse
+
+.pull-left[<img src="images/species_motifs_ferns_all_crop.png" alt="drawing">]
+
+.pull-right[
+```{r mpd-vs-motifs}
+# Plot omega score vs motif
+a <- make_scatter_with_lm(alpha_div_ferns, "motif_1", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[1]]) +
+  labs(x = "", title = "Motif 1")
+b <- make_scatter_with_lm(alpha_div_ferns, "motif_2", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[2]]) +
+  labs(title = "Motif 2", y = "", x = "")
+c <- make_scatter_with_lm(alpha_div_ferns, "motif_3", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[3]]) +
+  labs(title = "Motif 3", y = "", x = "")
+d <- make_scatter_with_lm(alpha_div_ferns, "motif_4", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[4]]) +
+  labs(title = "Motif 4")
+e <- make_scatter_with_lm(alpha_div_ferns, "motif_5", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[5]]) +
+  labs(title = "Motif 5", y = "")
+f <- make_scatter_with_lm(alpha_div_ferns, "motif_6", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[6]]) +
+  labs(title = "Motif 6", y = "")
+a + b + c + d + e + f + plot_layout(ncol = 2, nrow = 3)
+```
+]

### 2020-02-06 code/functions.R

@@ -8,13 +8,13 @@
 #' @return List
 #'
 read_nexus_in_zip <- function (zip_folder, nexus_file) {
-
+  
   temp_dir <- tempdir()
-
+  
   unzip(zip_folder, exdir = temp_dir)
-
+  
   ape::read.nexus(fs::path(temp_dir, nexus_file))
-
+  
 }
 
 #' Read in Catalog of Life plants data
@@ -24,12 +24,12 @@ read_nexus_in_zip <- function (zip_folder, nexus_file) {
 #' @return Tibble
 #'
 read_col_plants <- function (path_to_col) {
-
+  
   read_tsv(path_to_col,
-    col_types = rep("c", 31) %>% paste(collapse = "")
+           col_types = rep("c", 31) %>% paste(collapse = "")
   ) %>%
     mutate(taxonID = as.numeric(taxonID), acceptedNameUsageID = as.numeric(acceptedNameUsageID))
-
+  
 }
 
 
@@ -39,9 +39,9 @@ read_col_plants <- function (path_to_col) {
 #'
 #' @return Tibble
 process_repro_data <- function (data) {
-
+  
   data %>%
-  clean_names %>%
+    clean_names %>%
     mutate(
       reproductive_mode = case_when(
         reproductive_mode == 0 ~ "unknown",
@@ -61,7 +61,7 @@ process_repro_data <- function (data) {
     ) %>%
     rename(rbcl_genbank_no = rbc_l_gen_bank_accession_no) %>%
     mutate(rbcl_genbank_no = str_remove_all(rbcl_genbank_no, "\\*"))
-
+  
 }
 
 #' Get a list of URLs to download data for a kokudosuuchi dataset
@@ -140,9 +140,9 @@ tidy_japan_names <- function (data) {
 #' @return tibble
 count_species_per_cell <- function (occ_data, repro_data) {
   occ_data %>%
-  filter(taxon_id %in% repro_data$taxon_id) %>%
-  group_by(secondary_grid_code) %>%
-  count(sort = TRUE)
+    filter(taxon_id %in% repro_data$taxon_id) %>%
+    group_by(secondary_grid_code) %>%
+    count(sort = TRUE)
 }
 
 #' Count grid cells per species
@@ -152,8 +152,8 @@ count_species_per_cell <- function (occ_data, repro_data) {
 #' @return tibble
 count_cells_per_species <- function (occ_data) {
   occ_data %>%
[…]

### 2020-02-06 presentations/jsps_2020-03-01/index.Rmd
Use PNG for species dist maps, much faster to load

@@
 # Define function for making species ranking table
-make_species_ranking_table <- function (ranking_data, motif_select, taxon_select = NULL, slice_n = 10) {
+make_species_ranking_table <- function (ranking_data, motif_select, taxon_select = NULL, taxon_include = "", slice_n = 10) {
 
@@
     filter(motif == motif_select) %>%
+    mutate(include = ifelse(taxon %in% taxon_include, 1, 0)) %>%
     # Select top-scoring species with high SD
@@
     # a single motif)
-    arrange(desc(score), desc(score_sd)) %>%
+    arrange(desc(include), desc(score), desc(score_sd)) %>%
     slice(1:slice_n) %>%
@@
     arrange(desc(score), taxon)
+
+  if(is.numeric(taxon_select)) taxon_select <- ranking_data_sliced$taxon[[taxon_select]]
 
@@
           columns = everything(),
-          rows = taxon == ranking_data_sliced$taxon[[taxon_select]])
+          rows = taxon == taxon_select)
       )
@@
 # Define function to make simple distribution map for a species
-plot_distribution <- function (ranking_data, motif_select, taxon_select, occ_data, show_title = TRUE) {
+plot_distribution <- function (ranking_data, motif_select, taxon_select, occ_data, show_title = FALSE, taxon_include = "") {
 
@@
     filter(motif == motif_select) %>%
-    arrange(desc(score), desc(score_sd)) %>%
+    mutate(include = ifelse(taxon %in% taxon_include, 1, 0)) %>%
+    arrange(desc(include), desc(score), desc(score_sd)) %>%
     select(taxon, score) %>%
@@
 
+  if(is.numeric(taxon_select)) taxon_select <- ranking_data$taxon[[taxon_select]]
+
   cells_present <-
     occ_data %>%
-    filter(taxon_name == ranking_data$taxon[[taxon_select]]) %>%
+    filter(taxon_name == taxon_select) %>%
     select(secondary_grid_code) %>%
@@
     labs(
-      title = ranking_data$taxon[[taxon_select]]
+      title = taxon_select
     )
@@
 plot_dist_by_motif <- function(species_ranking, motif, occ_data_ferns) {
-  dist_maps <- purrr::map(1:10, ~plot_distribution(species_ranking, motif, ., occ_data_ferns)) %>%
+  dist_maps <- purrr::map(1:10, ~plot_distribution(species_ranking, motif, ., occ_data_ferns, show_title = TRUE)) %>%
     patchwork::wrap_plots(ncol = 2, nrow = 5)
@@
 .pull-right-narrow[
-```{r ecos-motif-1-focus-3}
-make_species_ranking_table(species_ranking, 1, 3)
-plot_distribution(species_ranking, 1, 3, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-1-focus-3, dev = "png"}
+make_species_ranking_table(species_ranking, 1, "Equisetum arvense")
+plot_distribution(species_ranking, 1, "Equisetum arvense", occ_data_ferns, show_title = FALSE)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-1-focus-7}
-make_species_ranking_table(species_ranking, 1, 7)
-plot_distribution(species_ranking, 1, 7, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-1-focus-7, dev = "png"}
+make_species_ranking_table(species_ranking, 1, "Athyrium vidalii")
+plot_distribution(species_ranking, 1, "Athyrium vidalii", occ_data_ferns)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-2-focus-1}
-make_species_ranking_table(species_ranking, 2, 1)
-plot_distribution(species_ranking, 2, 1, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-2-focus-1, dev = "png"}
+make_species_ranking_table(species_ranking, 2, "Adiantum pedatum")
+plot_distribution(species_ranking, 2, "Adiantum pedatum", occ_data_ferns)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-2-focus-2}
-make_species_ranking_table(species_ranking, 2, 2)
-plot_distribution(species_ranking, 2, 2, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-2-focus-2, dev = "png"}
+make_species_ranking_table(species_ranking, 2, "Arachniodes borealis")
+plot_distribution(species_ranking, 2, "Arachniodes borealis", occ_data_ferns)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-3-focus-3}
-make_species_ranking_table(species_ranking, 3, 3)
-plot_distribution(species_ranking, 3, 3, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-3-focus-3, dev = "png"}
+make_species_ranking_table(species_ranking, 3, "Dryopteris erythrosora")
+plot_distribution(species_ranking, 3, "Dryopteris erythrosora", occ_data_ferns)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-3-focus-7}
-make_species_ranking_table(species_ranking, 3, 7)
-plot_distribution(species_ranking, 3, 7, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-3-focus-7, dev = "png"}
+make_species_ranking_table(species_ranking, 3, "Odontosoria chinensis")
+plot_distribution(species_ranking, 3, "Odontosoria chinensis", occ_data_ferns)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-4-focus-2}
-make_species_ranking_table(species_ranking, 4, 2)
-plot_distribution(species_ranking, 4, 2, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-4-focus-2, dev = "png"}
+make_species_ranking_table(species_ranking, 4, "Blechnum niponicum")
+plot_distribution(species_ranking, 4, "Blechnum niponicum", occ_data_ferns)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-4-focus-8}
-make_species_ranking_table(species_ranking, 4, 8)
-plot_distribution(species_ranking, 4, 8, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-4-focus-8, dev = "png"}
+make_species_ranking_table(species_ranking, 4, "Polystichum retrosopaleaceum")
+plot_distribution(species_ranking, 4, "Polystichum retrosopaleaceum", occ_data_ferns)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-5-focus-3}
-make_species_ranking_table(species_ranking, 5, 3)
-plot_distribution(species_ranking, 5, 3, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-5-focus-3, dev = "png"}
+make_species_ranking_table(species_ranking, 5, "Dicranopteris linearis")
+plot_distribution(species_ranking, 5, "Dicranopteris linearis", occ_data_ferns)
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-5-focus-7}
-make_species_ranking_table(species_ranking, 5, 7)
-plot_distribution(species_ranking, 5, 7, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-5-focus-7, dev = "png"}
+make_species_ranking_table(species_ranking, 5, "Polystichum tagawanum")
+plot_distribution(species_ranking, 5, "Polystichum tagawanum", occ_data_ferns, show_title = FALSE)
 ```
@@
 ```{r ecos-motif-6}
-make_species_ranking_table(species_ranking, 6)
+make_species_ranking_table(species_ranking, 6, taxon_include = c("Angiopteris lygodiifolia", "Ctenitis subglandulosa"))
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-6-focus-1}
-make_species_ranking_table(species_ranking, 6, 1)
-plot_distribution(species_ranking, 6, 1, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-6-focus-2, dev = "png"}
+make_species_ranking_table(species_ranking, 6, "Angiopteris lygodiifolia", taxon_include = c("Angiopteris lygodiifolia", "Ctenitis subglandulosa"))
+plot_distribution(species_ranking, 6, "Angiopteris lygodiifolia", occ_data_ferns, taxon_include = "Angiopteris lygodiifolia")
 ```
@@
 .pull-right-narrow[
-```{r ecos-motif-6-focus-2}
-make_species_ranking_table(species_ranking, 6, 2)
-plot_distribution(species_ranking, 6, 2, occ_data_ferns, show_title = FALSE)
+```{r ecos-motif-6-focus-1, dev = "png"}
+make_species_ranking_table(species_ranking, 6, "Arachniodes sporadosora", taxon_include = c("Angiopteris lygodiifolia", "Ctenitis subglandulosa"))
+plot_distribution(species_ranking, 6, "Arachniodes sporadosora", occ_data_ferns)
 ```

### 2020-02-08 presentations/jsps_2020-03-01/index.Rmd
Update slides

@@
   "species_motifs_ferns_6L",
-  "percent_sex_dip_ferns"), 
+  "percent_sex_dip_ferns",
+  "repro_data"),
   cache = ja_fern_cache)
@@
   mutate_at(vars(longitude, latitude), as.numeric) %>%
+  # Make canape 'NA' an explicit factor level
+  mutate(canape = forcats::fct_explicit_na(canape, "NA")) %>%
   rename_at(vars(1:6 %>% as.character), ~paste0("motif_", .)) %>%
@@
 
-## Biodiversity is in danger from climate change
+## 生物多様性は危機にさらされている
+
+- 地球温暖化をはじめ、人間の活動によって生物多様性が失われている
+
+- 保全対策が必要
 
@@
 
-## We must understand the distribution of biodiversity to save it
+## 生物多様性の保全には、まずその分布を理解する必要がある
 
-- Most analyses of BD only include richness
-- But BD is more complicated: evo history also important
+- 今まで生物多様性は主に種数だけで計れてきた
+
+- しかし、生物多様性には様々な側面（次元）がある
+
+ - 系統的多様性
+
+ - 機能的多様性
+
+ - など
 
@@
 
-## Japan is an ideal site to study fern BD
+## 日本のシダ植物
 
-- Extremely well sampled
+.pull-left[
+- どの国よりも豊富なサンプリング
+
+- 日本シダの会の動力の甲斐
+
+- 生物多様性の多次元解析の最適？
+]
+
+.pull-right[
+```{r gbif-colls-per-country}
+# Read in pteridophyte GBIF collections per country data from
+# Iwasaki lab presentation
+collections_by_country <- read_csv(here::here("data_raw/gbif_pterido_coll_by_country.csv")) %>%
+  mutate(country = factor(country)) %>%
+  mutate(country = fct_reorder(country, n))
+
+# Plot showing top 10
+ggplot(collections_by_country, aes(x = country, y = n)) +
+  geom_col() +
+  coord_flip() +
+  labs(
+    title = "GBIF pteridophyte \ncollections (top 10)",
+    y = "No. collections") +
+  scale_y_continuous(labels = scales::comma) +
+  ggplot2::theme_bw(base_size = 24) +
+  theme(
+    axis.title.y = element_blank(),
+    panel.grid.minor = ggplot2::element_blank(),
+    panel.grid.major.y = ggplot2::element_blank(),
+      axis.text.x = ggplot2::element_text(colour="black"),
+      axis.text.y = ggplot2::element_text(colour="black")
+  )
+```
+]
 
@@
 
-## Goals
+## 日本のシダ植物
 
-- Map biodiversity (species richness, PD, endemism)
-- Identify biogeographical realms
-- Understand drivers of biogeo structure
+.pull-left[
+- どの国よりも豊富なサンプリング
+
+- 日本シダの会の動力の甲斐
+
+- 生物多様性の多次元解析の最適？
+]
+
+.pull-right[
+```{r ja-pterido-coll-curve}
+# Format pteridophyte abundances (no hybrids) as numeric vector
+pterido_abun <-
+occ_point_data_raw %>%
+  group_by(species) %>%
+  count(sort = TRUE) %>%
+  pull(n)
+
+# Run iNEXT
+inext_out <- iNEXT::iNEXT(pterido_abun, q=0, datatype="abundance")
+
+# create subset of observed points to add to plot as dots
+observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]
+
+# Make iNEXT plot
+inext_out$iNextEst %>%
+  filter(m < 500000) %>%
+  ggplot(aes(x=m, y=qD)) +
+  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70", alpha=0.2) +
+  geom_line(aes(linetype=method), size=0.8) +
+  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
+  geom_point(data=observed_data, size=5) +
+  scale_y_continuous("No. spp.") +
+  scale_x_continuous("No. collections") +
+  labs(
+    title = "Species collection curve",
+    subtitle = "Japanese pteridophytes (no hybrids)"
+  ) +
+  ggplot2::theme_bw(base_size = 24) +
+  theme(
+    axis.title.y = element_blank(),
+    panel.grid.minor = ggplot2::element_blank(),
+    axis.text.x = ggplot2::element_text(colour="black"),
+    axis.text.y = ggplot2::element_text(colour="black"),
+    legend.position="bottom", 
+    legend.title=element_blank()
+    )
+
+```
+]
 
@@
 
-## Data
+## 本研究の目的
 
-- Occurrence data of Japanese fern specimens
-- Phylogenetic tree (*rbcL*)
-- Morphological traits
-- Climate data
+**日本産シダ類** において
+
+- 多様性を複数の次元から把握すること
+
+- 生態的構造を検討すること
+
+- 以上のパターンを生み出している要素を探ること
+
+---
+class: big-margin
+
+```{r calc-species-num}
+# Count number of fern species, no hybrids
+
+# Inner-join fern occurrence data (incl. hybrids)
+# with repro data (no hybrids)
+n_ferns_no_hybrids <-
+occ_data_ferns %>%
+  select(taxon_id) %>%
+  unique %>%
+  inner_join(repro_data) %>%
+  pull(taxon_name) %>%
+  n_distinct()
+
+```
+
+## データ
+
+- 日本産シダ類の採取データ
+
+  - `r nrow(occ_data_ferns) %>% scales::comma()`点
+
+  - `r n_ferns_no_hybrids` 種類（雑種を除く）
+
+- 系統樹 (*rbcL*)
+
+- 形態的特徴
+
+- 気候データ
+
+.footnote[Ebihara and Nitta. 2019. *Journal of Plant Research* 132:723–738]
 
@@
 
-## Species richness peaks on Yakushima Isl.
+## 屋久島は種数の頂点
 
@@
 ---
+class: big-margin
 
-## Phylogenetic diversity is high in N and S
+## 系統的多様性
+
+- Mean phylogenetic distance (MPD): ある地点に生育する種の**平均的系統距離**
+
+- 生値をランダムな分布と比較して、ランダムより高いか、低いか、解析する
+
+.footnote[Webb, et al. 2002. *Annual Review of Ecology and Systematics* 33:475–505.]
+
+---
+
+## 系統的多様性は南北に高い
 
@@
 ```
+---
+class: big-margin
+
+## Categorical Analyses of Neo- and Paleo-Endemism (CANAPE)
+
+- 分布域が狭く、系統樹の枝が短い：Neoendemic
+
+- 分布域が狭く、系統樹の枝が長い：Paleoendemic
+
+- 両方混じっている地域：Mixed
+
+- 生値をランダムな分布と比較して、ランダムと有意義な差があれば**固有地域**として認める
+
+.footnote[Mishler, et al. 2014. *Nature Communications* 5:4473]
 
@@
 
-## S islands have high rates of endemism
+## 南の島々は固有種が多い
 
 ```{r canape-plot, out.width = "120%", fig.asp = 0.7}
-canape_cols <- RColorBrewer::brewer.pal(4, "Dark2") %>%
-  set_names(c("mixed", "neo", "palaeo", "super"))
+canape_cols <- RColorBrewer::brewer.pal(4, "Set1") %>%
+  c(., "grey80") %>%
+  set_names(c("neo", "palaeo", "super", "mixed", "NA"))
 
-ggplot(biodiv_results_ferns, aes(x = longitude, y = latitude, fill = canape)) +
+biodiv_results_ferns %>%
+  mutate(canape = fct_relevel(canape, "palaeo", "mixed", "neo")) %>%
+ggplot(aes(x = longitude, y = latitude, fill = canape)) +
   geom_tile() +
@@
 
-## Neoendemic areas have more sexual diploids
+## Neoendemicな地域は有性生殖・二倍体種が多い
 
@@
 ```{r canape-sex, fig.asp = 0.7}
+# Run analysis of variance (AOV) on
+# endemism type by reproductive mode.
+canape_by_repro_model <- aov(
+  percent_sex_dip ~ canape,
+  data = alpha_div_ferns)
 
-percent_sex_dip_ferns_round <-
-  percent_sex_dip_ferns %>% 
-  mutate(latitude = round(latitude, 3)) %>%
-  mutate(longitude = round(longitude, 3))
+# Inspect results:
 
-canape_sex_mode <-
-  biodiv_results_ferns %>%
-  mutate(latitude = round(latitude, 3)) %>%
-  mutate(longitude = round(longitude, 3)) %>%
-  inner_join(
-    percent_sex_dip_ferns_round,
-    by = c("latitude", "longitude")
-  ) %>%
-  assert(is_uniq, secondary_grid_code) %>%
-  filter(!is.na(canape))
+# - p-value
+# tidy(canape_by_repro_model)
 
-# Run analysis of variance (AOV) on
-# latitudinal breadth by reproductive mode.
-canape_by_repro_model = aov(
-  percent_sex_dip ~ canape,
-  data = canape_sex_mode)
-
-canape_by_repro_model_summary = tidy(canape_by_repro_model)
-
-# AOV of latidudinal breadth by reproductive mode
+# - AOV of endemism type by reproductive mode
 # showed a significant difference, so
 # run Tukey HSD test on results.
-canape_by_repro_tukey = TukeyHSD(canape_by_repro_model) %>% tidy
+# TukeyHSD(canape_by_repro_model) %>% tidy
 
-canape_sex_mode %>%
+# Assign tukey groups for plotting
+tukey_groups <-
+    multcomp::glht(
+    canape_by_repro_model, 
+    linfct = multcomp::mcp(canape = "Tukey")) %>%
+    multcomp::cld() %>%
+    purrr::pluck("mcletters", "Letters") %>%
+    tibble(canape = names(.), group = .) %>%
+    mutate(
+      y_pos = 1.1
+    )
+
+alpha_div_ferns %>%
   group_by(canape) %>%
@@
   ) %>%
-  mutate(canape = fct_reorder(canape, mean_sex_dip)) %>%
+  mutate(canape = fct_relevel(canape, "palaeo", "mixed", "neo")) %>%
   ggplot(aes(x = canape, y = mean_sex_dip, fill = canape)) +
   geom_col() +
-  scale_y_continuous(labels = scales::percent) +
+  geom_text(data = tukey_groups, aes(y = y_pos, label = group), color = "black") +
+  scale_y_continuous(labels = scales::percent, breaks = c(0, 0.25, 0.5, 0.75, 1.0)) +
   scale_fill_manual(values = canape_cols, na.value = "grey80", guide = FALSE) +
@@
 
+.footnote['a', 'b', 'c' = Tukey's HSDテストによる有意義な差があったグループ]
+
+---
+class: big-margin
+
+## 生態的構造
+
+STRUCTURE (Pritchard 2000)と同じようなアルゴリズムで生態的構造を解析する。
+
+STRUCTUREでは、**遺伝子型**のデータから**個体レベル**の遺伝的構造を推定する
+
+Ecostructure (White 2019)では、**種の分布**のデータから**群集レベル**の生態的構造を推定する
+
+.footnote[Pritchard, et al. 2000. *Genetics* 155: 945-959<br>
+White, et al. 2019. *Nature Communications* 10:2646.]
+
 ---
@@
 
-## Ecological structure of Japanese ferns, *K* = 6
+## 日本産シダ類の生態的構造 (*K* = 6)
 
@@
 
-## Motif 2 tends to cluster, motif 6 tends to overdisperse
+## 生態的構造のグループによって、系統的多様性が違う
 

### 2020-06-24 reports/figures/figures.Rmd
Add working report to generate figures

---
title: "figures"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)

# Load packages and functions
source(here::here("R/packages.R"))
source(here::here("R/functions.R"))
library(patchwork)
library(rcartocolor)

# Load objects from cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

loadd(list = c(
  "ses_div_ferns",
  "shape_ferns",
  "comm_scaled_coverage_0.2",
  "ses_div_ferns_endemic",
  "k_taxonomy",
  "k_phylogeny",
  "regions_taxonomy",
  "regions_phylogeny",
  "species_motifs_ferns"
),
cache = ja_fern_cache)
```

```{r plot-prep}
tax_regions_data <- regions_taxonomy %>%
  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

phy_regions_data <- regions_phylogeny %>%
  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

ses_div_ferns_spatial <-
  shape_ferns %>%
  left_join(categorize_endemism(ses_div_ferns), by = c(grids = "site")) %>%
  left_join(comm_scaled_coverage_0.2, by = c(grids = "site")) %>%
  mutate(redundancy = 1 - (richness/abundance)) %>%
  left_join(tax_regions_data, by = "grids") %>%
  left_join(phy_regions_data, by = "grids")

ses_div_ferns_endemic_spatial <-
shape_ferns %>%
  left_join(categorize_endemism(ses_div_ferns_endemic), by = c(grids = "site"))

map_theme <- theme(
    # panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "white", size = 0.2),
    axis.ticks = element_blank()
  )

```

```{r k-taxonomy-plot, include = FALSE}
k_taxonomy$df %>%
  as_tibble() %>%
  mutate(optimal = ifelse(k == k_taxonomy$optimal$k, TRUE, FALSE)) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal)) +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  labs(
    y = "Explained variances",
    x = "Number of clusters") +
  annotate("text", x = Inf, y = Inf, label = glue::glue("K = {k_taxonomy$optimal$k}"), hjust = 1, vjust = 1) +
  theme(legend.position = "none")
```

```{r k-phylogeny-plot, include = FALSE}
k_phylogeny$df %>%
  as_tibble() %>%
  mutate(optimal = ifelse(k == k_phylogeny$optimal$k, TRUE, FALSE)) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal)) +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  labs(
    y = "Explained variances",
    x = "Number of clusters") +
  annotate("text", x = Inf, y = Inf, label = glue::glue("K = {k_phylogeny$optimal$k}"), hjust = 1, vjust = 1) +
  theme(legend.position = "none")
```

Redundancy (number of specimens per species) is generally high throughout the country, but lower in Hokkaido and some coastal areas and small islands. Species richness is highest in southern Kyushu and on Yakushima Island, and lowest on Hokkaido and low-lying coastal areas. Raw MPD is uniform throughout, but SES of MPD shows clustering in mountainous areas of Honshu and overdispersion in souther Kyushu and islands further south.

```{r diversity-plots, fig.width = 8, fig.height = 8}
# Richness
a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() +
  labs(fill = "Richness \n (n spp.)") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# Sampling rendundancy
b <- ggplot(ses_div_ferns_spatial, aes(fill = redundancy)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Redundancy") +
  map_theme +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank()
  )

# MPD
c <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_obs)) +
  geom_sf(color = "transparent") +
  labs(fill = "MPD") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  scale_fill_viridis_c() +
  map_theme

# SES MPD
d <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES MPD") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.y = element_blank()
  )

a + b + c + d  +
  plot_annotation(tag_levels = 'A')

```

**Figure 1**. Taxonomic and phylogenetic diversity of the ferns of Japan. A, richness. B, sampling redundancy. C, mean phylogenetic distance (MPD). D, Standard effect size of MPD.

```{r endem-plots, fig.width = 8, fig.height = 10}
# PE
a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  map_theme

# SES PE
b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, pe_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, pe_obs_z, "abs")
    )) +
  map_theme

# PE (endemic)
c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  map_theme

# SES PE (endemic)
d <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs"),
      get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs")
    )) +
  map_theme

# CANAPE
e <- ggplot(ses_div_ferns_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  map_theme

# CANAPE (endemic)
f <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  map_theme

a + b + c + d + e + f +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')
```

**Figure 2**. Phylogenetic endemism of the ferns of Japan. A, phylogenetic endemism (PE). B, Standard effect size (SES) of PE. C, PE of endemic taxa only. D, SES of PE of endemic taxa only. E, Categorical analysis of neo- and paleo- endemism (CANAPE) results. F, CANAPE results for endemic taxa only.

```{r bioregion-plots, fig.width = 8}
# Taxonomic bioregions
a <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
  geom_sf(color = "transparent") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Safe") +
  map_theme

# Phylogenetic bioregions
b <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
  geom_sf(color = "transparent") +
  labs(fill = "Cluster") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Safe") +
  map_theme

a + b +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')
```

**Figure 3**. Regionalization of the ferns of Japan. A, taxonomic regionalization. B, phylogenetic regionalization.

```{r ecostructure}
tweak <- 0.75

sites <- ses_div_ferns_spatial %>%
  pull(grids)

centroids <-
ses_div_ferns_spatial %>%
  select(geometry) %>%
  mutate(centroid = sf::st_centroid(geometry)) %>%
  sf::st_coordinates(x = .$centroid) %>%
  as_tibble() %>%
  mutate(site = sites)

omega <-
species_motifs_ferns$omega %>%
  as.data.frame %>%
  rownames_to_column("site") %>%
  left_join(centroids, by = "site") %>%
  column_to_rownames("site")

omega_centroids <- select(omega, X, Y) %>% # longitude, latitude
  as.matrix()

omega <- select(omega, -X, -Y)

ecos_plot_pie2(
    omega = omega,
    image_width = 6000*tweak,
    image_height = 4800*tweak,
    radius = 0.08,
    no_map = TRUE,
    coords = omega_centroids,
    long_lim = c(121, 148),
    lat_lim = c(23, 47),
    color = carto_pal(8, "Safe"),
    blank_bg = "grey10",
    path = here::here(
      "reports/figures/species_motifs_ferns.png"
    )
)

```

![](species_motifs_ferns.png)

**Figure 3**. Ecological structure of the ferns of Japan. K = 8.

### 2020-06-24 reports/figures/figures.Rmd
Update figures report

@@
 ---
-title: "figures"
+title: "Results"
 output: html_document
@@
 
-Redundancy (number of specimens per species) is generally high throughout the country, but lower in Hokkaido and some coastal areas and small islands.
-Species richness is highest in southern Kyushu and on Yakushima Island, and lowest on Hokkaido and low-lying coastal areas.
-Raw MPD is uniform throughout, but SES of MPD shows clustering in mountainous areas of Honshu and overdispersion in souther Kyushu and islands further south.
+Redundancy (number of specimens per species) is generally high throughout the country, but lower in Hokkaido and some coastal areas and small islands (Fig 1B).
+Species richness is highest in southern Kyushu and on Yakushima Island, and lowest on Hokkaido and low-lying coastal areas (Fig 1A).
+Raw MPD is mostly uniform throughout (Fig 1C), but SES of MPD shows clustering in mountainous areas of Honshu and overdispersion in souther Kyushu and islands further south (Fig 1D).
 
-```{r diversity-plots, fig.width = 8, fig.height = 8}
+Phylogenetic endemism (PE) is extremeley low throughout except for in some of the southern islands (Fig 2A).
+SES of PE is high starting in southern Kyushu and southwards, including the Ogasawara islands (Fig 2B).
+Results of CANAPE (Categorical Analysis of Palaeo and Neo Endemism) indicate hotspots of mixed endemism in the southern islands, but no neo- or paleo-endemism (Fig 2E).
+It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (the "border effect").
+This is especially the case for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country.
+To account for this, we also analyzed a dataset including only species endemic to Japan (Fig 2C, D, and F).
+The patterns are generally similar to the full dataset, indicating that the southern islands are genuine hotspots of endemicity.
+
+Analysis of biogeographic regions using taxonomy (species names) resulted in the vast majority of grid cells classified into four regions: Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Okinawa), and Cluster 4 (Ogasawara Islands) (Fig 3A).
+Using phylogenetic distances produced similar results, except that most grid cells in Okinawa and the Ogawara Islands grouped into the same cluster.
+
+Analysis of ecological structure using species names with the same number of clusters as the taxonomic bioregions (K = 8) revealed high amounts of turnover throughout Honshu that mostly tracks latitude and elevation.
+
+```{r diversity-plots, fig.width = 10, fig.height=7.5}
 # Richness
@@
 
-```{r endem-plots, fig.width = 8, fig.height = 10}
+```{r endem-plots, fig.width = 10, fig.height=11.5}
 # PE
@@
 
-```{r bioregion-plots, fig.width = 8}
+```{r bioregion-plots, fig.width = 10, fig.height=4.5}
 # Taxonomic bioregions
@@
   geom_sf(color = "transparent") +
-  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Safe") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
   map_theme
@@
   labs(fill = "Cluster") +
-  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Safe") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
   map_theme
@@
 
-**Figure 3**. Regionalization of the ferns of Japan. A, taxonomic regionalization. B, phylogenetic regionalization.
+**Figure 3**. Bioregions of the ferns of Japan. A, taxonomic bioregions B, phylogenetic bioregions.
 
@@
     lat_lim = c(23, 47),
-    color = carto_pal(8, "Safe"),
-    blank_bg = "grey10",
+    color = carto_pal(8, "Bold"),
+    blank_bg = "grey90",
     path = here::here(
@@
 
-**Figure 3**. Ecological structure of the ferns of Japan. K = 8.
+**Figure 4**. Ecological structure of the ferns of Japan based on species' occurrences. K = 8.

### 2020-06-28 ms/SI.Rmd
Refactor manuscript
- Use new analysis results
- Add SI
- Make captions with captioner

---
title: "Exploring dimensions of biodiversity in Japanese ferns"
bibliography: "main_library.bib"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
- BoldFont=Roboto-Bold.ttf
- ItalicFont=Roboto-Italic.ttf
- BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
    includes:
      in_header: newphyt.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  chunk_output_type: console
# nocite: | 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)
```

```{r load-data, cache = FALSE}
# NOTE: This Rmd is meant to be knit as part of the drake plan (R/plan.R).
# If using interactively, run `source("_drake.R")` first.

loadd(list = c(
  "occ_point_data_ferns",
  "ses_phy_ferns",
  "ses_phy_ferns_endemic",
  "ses_traits_ferns",
  "shape_ferns",
  "k_taxonomy",
  "k_phylogeny",
  "regions_taxonomy",
  "regions_phylogeny",
  "species_motifs_ferns"
),
cache = ja_fern_cache)

phylo_cache <- new_cache(here::here("phylo_cache"))

loadd(list = c(
  "plastome_calibration_dates",
  "plastome_tree"
),
cache = phylo_cache)

# Load captions
source(here::here("ms/captions.R"))
```

<!-- Remove page numbers -->
\pagenumbering{gobble}

<!-- Set left justification -->
\raggedright

<!-- Don't indent -->
\setlength{\parindent}{0ex}

# Supplemental Information

```{r plot-prep}
tax_regions_data <- regions_taxonomy %>%
  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

phy_regions_data <- regions_phylogeny %>%
  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

ses_div_ferns_spatial <-
  shape_ferns %>%
  left_join(categorize_endemism(ses_phy_ferns), by = c(grids = "site")) %>%
  left_join(ses_traits_ferns, by = c(grids = "site")) %>%
  mutate(redundancy = 1 - (richness/abundance)) %>%
  left_join(tax_regions_data, by = "grids") %>%
  left_join(phy_regions_data, by = "grids")

ses_div_ferns_endemic_spatial <-
shape_ferns %>%
  left_join(categorize_endemism(ses_phy_ferns_endemic), by = c(grids = "site"))

map_theme <- theme(
    panel.grid.major = element_line(color = "white", size = 0.2),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
  )
```

```{r k-plot, fig.width = 7, fig.height = 3}
k_taxonomy_select <- 8

a <- k_taxonomy$df %>%
  as_tibble() %>%
  mutate(
    optimal_auto = ifelse(k == k_taxonomy$optimal$k, TRUE, FALSE),
    optimal_manual = ifelse(k == k_taxonomy_select, TRUE, FALSE),
    label = ifelse(k == k_taxonomy_select, glue::glue("k = {k}"), "")
  ) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal_manual, shape = optimal_auto), size = 2.5) +
  geom_text(aes(label = label), vjust = 0, nudge_y = 0.02, size = 4) +
  scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
  scale_shape_manual(values = c("TRUE" = 17, "FALSE" = 16)) +
  labs(
    y = "Explained variance",
    x = "Number of clusters") +
  theme(
    legend.position = "none"
  )

k_phylogeny_select <- 7

b <- k_phylogeny$df %>%
  as_tibble() %>%
  mutate(
    optimal_auto = ifelse(k == k_phylogeny$optimal$k, TRUE, FALSE),
    optimal_manual = ifelse(k == k_phylogeny_select, TRUE, FALSE),
    label = ifelse(k == k_phylogeny_select, glue::glue("k = {k}"), "")
  ) %>%
  ggplot(aes(x = k, y = ev)) +
  geom_path() +
  geom_point(aes(color = optimal_manual, shape = optimal_auto), size = 2.5) +
  geom_text(aes(label = label), vjust = 0, nudge_y = 0.02, size = 4) +
  scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
  scale_shape_manual(values = c("TRUE" = 17, "FALSE" = 16)) +
  labs(
    y = "Explained variance",
    x = "Number of clusters",
    shape = "Optimal k\n(auto)",
    color = "Optimal k\n(manual)") +
  theme(
    axis.title.y = element_blank()
  )

a + b +
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')

```

**`r s_figure("k-plot")`**. `r s_figure_cap("k-plot")` **A**, taxonomic bioregions. **B**, phylogenetic bioregions. Shape indicates optimal *k* selected by \function{optimal{\_}phyloregion} function in the R package \package{phyloregion} (circle, non-optimal; triangle, optimal); color indicates optimal *k* selected manually (black, non-optimal; red, optimal; see Methods).

<!-- (Add k plot for ecostructure) -->

\clearpage

```{r endemism-restricted, fig.width = 7, fig.height = 6}
# PE (endemic)
a <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_viridis_c() + 
  labs(fill = "PE") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.x = element_blank()
  )

# SES PE (endemic)
b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES PE") +
  scale_fill_scico(
    palette = "romaO",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs"),
      get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs")
    )) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme +
  theme(
    axis.text.y = element_blank()
  )

# CANAPE (endemic)
c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = endem_type)) +
  geom_sf(color = "transparent") +
  labs(fill = "Endemism\ntype") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme

a + b + c + 
  plot_layout(ncol = 2) +
  plot_annotation(tag_levels = 'A')

ggsave(here::here("results/Fig_S1.pdf"), width = 7, height = 6, units = "in")
```

**`r s_figure("endemism-restricted")`**. `r s_figure_cap("endemism-restricted")` **A**, phylogenetic endemism (PE). **B**, Standard effect size (SES) of PE. **C**, Categorical analysis of neo- and paleo-endemism (CANAPE).

### 2020-06-28 ms/manuscript.Rmd
Refactor manuscript
- Use new analysis results
- Add SI
- Make captions with captioner

--- 2019-08-31 ms/manuscript.Rmd
@@
 ---
-title: "Exploring multiple dimensions of biodiversity in Japanese ferns and lycophytes"
-bibliography: "references.bib"
-mainfont: Times New Roman.ttf
-fontsize: 12pt
+title: "Exploring dimensions of biodiversity in Japanese ferns"
+bibliography: "main_library.bib"
+mainfont: Roboto-Regular.ttf
+fontsize: 11pt
 mainfontoptions: 
-- BoldFont=Times New Roman Bold.ttf
-- ItalicFont=Times New Roman Italic.ttf
-- BoldItalicFont=Times New Roman Bold Italic.ttf
+- BoldFont=Roboto-Bold.ttf
+- ItalicFont=Roboto-Italic.ttf
+- BoldItalicFont=Roboto-BoldItalic.ttf
 output: 
@@
 # nocite: | 
-#   @Vogel1968, @Wollenweber2000
 ---
@@
 
-```{r load-packages, cache = FALSE}
-library(drake)
-library(knitr)
-library(patchwork)
-library(scales)
-library(scico)
-library(magrittr)
-library(broom)
-library(tidyverse)
+```{r load-data, cache = FALSE}
+# NOTE: This Rmd is meant to be knit as part of the drake plan (R/plan.R).
+# If using interactively, run `source("_drake.R")` first.
+
+loadd(list = c(
+  "occ_point_data_ferns",
+  "ses_phy_ferns",
+  "ses_phy_ferns_endemic",
+  "ses_traits_ferns",
+  "shape_ferns",
+  "k_taxonomy",
+  "k_phylogeny",
+  "regions_taxonomy",
+  "regions_phylogeny",
+  "species_motifs_ferns"
+),
+cache = ja_fern_cache)
+
+phylo_cache <- new_cache(here::here("phylo_cache"))
+
+loadd(list = c(
+  "plastome_calibration_dates",
+  "plastome_tree"
+),
+cache = phylo_cache)
+
+# Load captions
+source(here::here("ms/captions.R"))
 ```
 
-```{r load-data, cache = FALSE}
-# Set cache
-ja_fern_cache <- new_cache(here::here("ja_fern_cache"))
-
-source(here::here("code/functions.R"))
-
-# Load targets
-loadd(list = c(
-  "all_cells",
-  "occ_data_ferns",
-  "occ_data_ferns_north",
-  "occ_data_pteridos",
-  "alpha_div_pteridos",
-  "world_map"), 
-  cache = ja_fern_cache)
-```
-
-<!-- FOR CONVERSION TO MS WORD ONLY: remove page numbers -->
+<!-- Remove page numbers -->
 \pagenumbering{gobble}
 
-<!-- Set left justification (better conversion to Word than default full justification) -->
+<!-- Set left justification -->
 \raggedright
 
-<!-- Don't indent authors -->
+<!-- Don't indent -->
 \setlength{\parindent}{0ex}
@@
 
-^1^Department of Botany, National Museum of Natural History, Smithsonian Institute, Washington, D.C., 20013, U.S.A.
+^1^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo 113-0032, Japan
 
@@
 
-The pteridophytes (ferns and lycophytes) of Japan are one of the best-characterized regional floras in the world thanks to decades of collection and taxonomic work.
-Here, we combine phylogenetic, functional, and environmental datasets to characterize the spatial distribution of biodiversity in the pteridophytes of Japan.
+The ferns of Japan are one of the best-characterized regional floras in the world due to decades of intensive collection and taxonomic revision.
+We combined phylogenetic, functional, and environmental datasets to explore the spatial distribution of biodiversity in the ferns of Japan.
 
@@
 
+Characterizing biodiversity is important for two reasons: to understand the processes generating it, and to conserve it.
+
+Most past studies of biodiversity only focus on one dimension: richness.
+
+Recently tools and data are available to analyze additional dimensions (in particular phylogenetic dimensions).
+Furthermore, better ways to characterize biological regions ("bioregions") are also available, rather than simply using expert opinion or other ad-hoc characterizations.
+
+Incorporating two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.
+
+However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data or can be lead astray.
+
+Ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are very densely sampled.
+Japan is situated along a latitudinal gradient spanning highly season temperate areas in the north to subtropical, mostly aseasonal islands in the south.
+
+Here we leverage this exceptionally rich dataset to analyze the geographic distribution of biodiversity, reveal its spatial structure, and assess drivers of community assembly.
+
 ## Materials and Methods
@@
 
-Our data include all ferns and lycophytes known to occur within the political boundaries of Japan (20° to 45° N, 122° to 153° E).
+Our data include all ferns known to occur within the political boundaries of Japan (20° to 45° N, 122° to 153° E).
 Japan consists of four main islands and thousands of smaller ones.
-Our data exlude islands smaller than XX km^2^.
 The climate is subtropical in the south and highly seasonal in other parts of the country.
@@
 
-We used occurrence data of ferns and lycophytes in Japan compiled from > 200,000 vouchered specimens including 721 non-hybrid taxa (species, subspecies, and varieties) and 371 interspecific hybrids (1090 taxa total) [@Ebihara2019b].
-The data are formatted as taxon presences in 10 km × 10 km grid cells (hereafter '10 km cells').
-The 10 km grid cells correspond to the secondary level of basic grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) (`r nrow(all_cells) %>% comma` 10 km cells total).
+We used occurrence data of ferns in Japan comprising `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` vouchered specimens representing `r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` non-hybrid taxa (species, subspecies, and varieties).
+All specimens are housed at herbaria in Japan, and taxonomy follows the Japan Green List v.
+1.01 (http://www.rdplants.org/gl/).
+
+Any analysis of biodiversity must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling.
+We tested using 0.1, 0.2, or 0.4 degree grid-cells for defining communities (co-occurring species).
+At each size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell).
+We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling.
+Preliminary analysis indicated that 0.2 degree grid-cells are optimal for our dataset, so we used this for all subsequent analyses.
 
@@
 Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens.
-All qualitative traits formatted in binary format; e.g., a single trait with three states is formatted as three binary traits.
+All qualitative traits in binary format; e.g., a single trait with three states is formatted as three binary traits.
 Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately.
@@
 
-### Statistical analysis
+We used ploidy level and breeding system data in @Ebihara2019b.
 
-We used two distance-based metrics to quantify alpha biodiversity, mean phylogenetic distance (MPD; the mean distance between all co-occurring taxa in a community) and mean nearest taxon distance (MNTD; the mean distance between closest taxon pairs in a community) [@Webb2000; @Webb2002].
-Although MPD and MNTD were originally formulated for phylogenetic distances, they can be applied to any type of distance matrix.
-We calculated phylogenetic distances from the phylogenetic tree using the "cophenetic" function in R.
-We calculated functional distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R pacakge \package{FD} [@Laliberte2011].
-We weighted quantitative and qualitative traits equally, and weighted each binary component trait of each qualitative trait equally.
+### Phylogenetic analysis
 
-Since observed values of MPD and MNTD are known to correlate with species richness [@Swenson2014], we determined whether communities are overdispersed (i.e., diverse) or clustered (i.e., depauperate), by calculating the standard effect size (SES) of MPD and MNTD.
-We compared observed values to those of a null distribution of 999 random communities with the \function{ses.mpd} and \function{ses.mntd} functions in the \package{picante} R package [@Kembel2010].
-Random communities were each generated using 10,000 iterations of the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities.
-Positive values of SES of MPD and MNTD indicate overdispersion, and negative values indicate clustering.
-We hereafter refer to the SES of phylogenetic and functional MPD and MNTD as MPD~phy~ and MNTD~phy~, and MPD~func~ and MTND~func~, respectively.
+Sequences for the plastid, coding *rbcL* gene are available for ca.
+97% of the Japanese fern flora [@Ebihara2019b].
+However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils availabsle for molecular dating.
+Because one goal of the study is to understand the distribution of paleo- vs.
+neo-endemic areas (defined in units of time), we require an ultrametric tree.
+To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total).
+We conducted phylogenetic analysis with IQTREE [@Nguyen2015] and molecular dating with treePL [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a.
+We then trimmed the tree to Japanese taxa only and used this for all further analyses.
 
-A large number of primarily tropical fern lineages only occur in the far-southern islands of Japan: `r setdiff(occ_data_ferns$genus, occ_data_ferns_north$genus) %>% length` genera and `r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% length` families (`r setdiff(occ_data_ferns$family, occ_data_ferns_north$family) %>% paste(collapse = ", ")`) do not occur north of 30.1 degrees latitude.
-When these lineages are included in the randomization procedure to calculate SES of MPD and MNTD, they likely contribute long branches that bias the northern communities towards clustering.
-Lycophytes, which occur on long branches, may have a similar effect.
-Therefore, we calculated phylogenetic diversity for three datasets: pteridophytes (ferns and lycophytes), ferns, and ferns separately north and south of 30.1 degrees latitude (corresponding to the Tokara strait, hereafter referred to as "ferns_NS").
+### Analysis of biodiversity and endemism
+
+We measured taxanomic diversity using species richness.
+
+We measured phylogenetic diversity using the mean phylogenetic distance between all co-occurring taxa in a community (MPD) [@Webb2000; @Webb2002].
+We calculated distances from the phylogenetic tree using the \function{cophenetic} function in R v.
+4.0.0 [@RCoreTeam2019] and MPD using the \function{mpd} function in the R \package{picante}.
+
+For morphological diversity, we first calculated morphological distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011].
+We weighted quantitative and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled quantitative traits prior to calculating distances.
+We then used the morphological distances as input into the \function{mpd} function to calculate mean morphological distance (MMD).
+
+We measured phylogenetic endemism (PE) [@Rosauer2009], which is the sum of all branches at a site inversely weighted by their range, using the \function{phylo{\_}endemism} function in the R package \package{phyloregion}.
+
+Observed values of MPD, MMD, and PE are known to be closely related to species richness, and tend to saturate at high values of richness.
+Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of species by calculating their standard effect sizes (SES), defined as the difference between the observed value and the mean of a pool of random values, divided by the standard deviation of the random values.
+We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then calculated SES of each metric.
+
+We tested whether endemic areas have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
+Low RPE indicates short branchlengths (neoendemism) and high RPE indicates long branchlengths (paleoendemism).
+Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable to those calculated for the null distribution.
+We used the same set of null communities as for SES of PE, and carried out CANAPE using custom R scripts.
+
+It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (the "border effect").
+This is especially expected for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country.
+To account for this, we also calculated all endemism-related metrics using a dataset including only species endemic to Japan.
+
+### Analysis of bioregions and ecological structure
+
+We analyzed bioregions using the framework of @Daru2020 and their R package \package{phyloregion}.
+Briefly, this involves first calculating beta diversity (change in composition between communities) using either species names (taxonomic bioregions) or phylogenetic distances (phylogenetic bioregions).
+For taxonomic bioregions, we calculated the turnover in species between sites due to species replacement [@Baselga2012] with the Sørensen index.
+For phylogenetic bioregions, we calculated phylogenetic dissimilarities based on the PhyloSor index [@Bryant2008].
+The distances are used to construct a dendrogram, which is then split into into *k* bioregions.
+As the optimal number of bioregions cannot be known *a-priori*, we tested values of *k* from 1 to 20.
+\package{phyloregion} provides a function to automatically select *k* from a range of test values, but we found the results to be unreliable.
+We therefore evaluated the explained variance at each value of *k*, then selected *k* that provided the greatest increase in variance relative to the previous value (`r s_figure("k-plot")`).
+
+We analyzed ecological structure using the framework of @White2019 and their R package \package{ecostructure}.
+This method involves fitting the community matrix to a grade of membership model, which allows communities have partial membership in multiple groups [@White2019].
+Like bioregion analysis, this method also requires selecting the number of clusters (in this context called "motifs" after @White2019) to fit the model.
+We initially tested values of *k* from 1 to 20 and inspected the log-likelihood for each, but found that it continually improved without reaching a clear optimum.
+We therefore used the same *k* that was selected for taxonomic bioregions.
 
@@
 
-### Richness
+## Results
 
-Richness of pteridophytes in Japan roughly follows the latitudinal species gradient, with more species in the south (\fig \@ref(fig:div-pteridos)a).
-Richness tended to increase with elevation ((\fig \@ref(fig:div-el)a).
+### Taxonomic, phylogenetic, and morphological diversity
 
-### Phylogenetic diversity
+Species richness is highest in southern Kyushu and on Yakushima Island, and lowest on Hokkaido and low-lying coastal areas (`r figure("diversity")`A).
+Sampling redundancy is generally high throughout the country, with a peak just north of Tokyo, but lower in Hokkaido and some coastal areas and small islands (`r figure("diversity")`B).
+Raw MPD is mostly uniform throughout (`r figure("diversity")`C).
+SES of MPD shows clustering in some parts of central Honshu, with overdispersion both in the northern half of the country and from southern Kyushu to the south (`r figure("diversity")`D).
+Morphological diversity shows a similar pattern to phylogenetic diversity, but with more extreme overdispersion and clustering (`r figure("diversity")`F).
 
-Standard effect sizes of mean phylogenetic diversity (MPD~phy~) and mean nearest taxon distance (MNTD~phy~) were either not correlated or only very weakly correlated with species richness (SI Fig.
-), indicating that our null models successfully decoupled measures of phylogenetic diversity and richness.
+### Phylogenetic endemism
 
-Values of MPD~phy~ are clustered in mountainous areas and overdispersed along the coasts of the main islands of Japan (\fig \@ref(fig:div-pteridos)b), with similar patterns between pteridophytes and ferns (SI Fig.
-).
-The pattern is similar in the "ferns_NS" dataset, except that the strength of both clustering and overdispersion is weaker in the main islands of Japan, and the pattern of overdispersion on the islands of Amami-oshima in sourthern Japan reverses to clustering (SI Fig.
-).
+Phylogenetic endemism (PE) is extremeley low throughout except for in some of the southern islands (`r figure("endemism")`A).
+SES of PE is high starting in southern Kyushu and southwards, including the Ogasawara islands (`r figure("endemism")`B).
+Results of CANAPE (Categorical Analysis of Palaeo and Neo Endemism) indicate hotspots of mixed endemism in the southern islands, but no neo- and very little paleo-endemism (`r figure("endemism")`C).
+When we analyzed patterns of endemicity with a reduced dataset including only taxa endemic to Japan to account for the border effect, patterns were generally similar to the full dataset, indicating that the southern islands are genuine hotspots of endemicity (`r s_figure("endemism-restricted")`).
 
-The signal of clustering is much stronger when measured with MNTD~phy~, especially in mountainous areas of central Honshu (\fig \@ref(fig:div-pteridos)c).
-Patterns are similar between all three datasets except that MNTD~phy~ values are smaller in the "ferns_NS" dataset (SI Fig.
-).
+### Bioregions
 
-Values of MPD and MNTD decrease significantly with elevation across all datasets (\fig \@ref(fig:div-el)b, c; linear model).
+Analysis of biogeographic regions using taxonomy (species names) resulted in the vast majority of grid cells classified into four regions: Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Okinawa), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A).
+Using phylogenetic distances produced similar results, but Okinawa and the Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B).
 
-### Functional diversity
+### Ecological structure
 
-Spatial patterns of functional diversity differed from phylogenetic diversity and by metric.
-When measured with MPD~func~, pteridophytes were overdispersed in Hokkaido, northern Honshu, and the smaller islands south of Kyushu, but clustered in southern Honshu, Shikoku, and Kyushu (\fig \@ref(fig:div-pteridos)d).
-Less spatial structure was evident in MNTD~func~, with more clustering overall (\fig \@ref(fig:div-pteridos)e).
-
-Functional diversity increased with elevation when measured with MPD~func~, but not MNTD~func~ (\fig \@ref(fig:div-el)d, e; linear model)
-
-### Breeding system
-
-Sexual diploid taxa tend to be distributed in northern latitudes (\fig \@ref(fig:div-pteridos)f).
-
-Somewhat surprisingly, MPD and MNTD show a weak negative relationship with the percentage of sexual diploid taxa (linear model).
-We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed.
-It seems that communities with many non-sexual taxa actually tend to include distantly related taxa rather than sister pairs of sexual and non-sexual species.
+Ecological structure partly resembled bioregions, with most sites in Hokkaido and northern Honshu belonging to the same motifs (dark-green and purple), and sites in Okinawa also mostly belonging to another motif (grey) (`r figure("ecostructure")`).
+However, the grade-of-membership model also revealed complex ecological structure that mostly tracks latitude and elevation, with particularly high turnover in the southern half of Honshu.
+A single motif united low-richness sites (dark purple).
 
@@
 
+TBD
+
 ## Acknowledgements
 
-We thank people
-
 ## Author Contribution
-
-Who did what
 
@@
 
-## Supporting Information
+<!-- Figures -->
 
-Additional Supporting Information may be found online in the Supporting Information section at the end of the article:
+```{r plot-prep}
+tax_regions_data <- regions_taxonomy %>%
+  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
 
-**Table S1**
+phy_regions_data <- regions_phylogeny %>%
+  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
 
-**Fig. S1**
+ses_div_ferns_spatial <-
+  shape_ferns %>%
+  left_join(categorize_endemism(ses_phy_ferns), by = c(grids = "site")) %>%
+  left_join(ses_traits_ferns, by = c(grids = "site")) %>%
+  mutate(redundancy = 1 - (richness/abundance)) %>%
+  left_join(tax_regions_data, by = "grids") %>%
+  left_join(phy_regions_data, by = "grids")
+
+ses_div_ferns_endemic_spatial <-
+shape_ferns %>%
+  left_join(categorize_endemism(ses_phy_ferns_endemic), by = c(grids = "site"))
+
+map_theme <- theme(
+    panel.grid.major = element_line(color = "white", size = 0.2),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
+  )
+
+```
+
+```{r diversity, fig.width = 7, fig.height = 9}
+# Richness
+a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
+  geom_sf(color = "transparent") +
+  scale_fill_viridis_c() +
+  labs(fill = "Richness \n (n spp.)") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.x = element_blank()
+  )
+
+# Sampling rendundancy
+b <- ggplot(ses_div_ferns_spatial, aes(fill = redundancy)) +
+  geom_sf(color = "transparent") +
+  scale_fill_viridis_c() +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  labs(fill = "Redundancy") +
+  map_theme +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank()
+  )
+
+# MPD
+c <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_obs)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "MPD") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  scale_fill_viridis_c() +
+  map_theme +
+  theme(
+    axis.text.x = element_blank()
+  )
+
+# SES MPD
+d <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES MPD") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank()
+  )
+
+# Raw morph diversity
+e <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_morph_obs)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "MMD") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  scale_fill_viridis_c() +
+  map_theme
+
+# SES of morph diversity
+f <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_morph_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES MMD") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, mpd_morph_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, mpd_morph_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.y = element_blank()
+  )
+
+a + b + c + d + e + f +
+  plot_layout(ncol = 2) +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(here::here("results/Fig_1.pdf"), width = 7, height = 9, units = "in")
+```
+
+**`r figure("diversity")`**.
+`r figure_cap("diversity")` **A**, richness.
+**B**, sampling redundancy.
+**C**, mean phylogenetic distance (MPD).
+**D**, Standard effect size (SES) of MPD.
+**E**, mean morphological distance (MMD) **F**, SES of MMD.
 
@@
 
-**\fig \@ref(fig:div-pteridos).** Alpha diversity of ferns and lycophytes in Japan.
-**a** species richness.
-**b** Phylogenetic mean phylogenetic distance (MPD~phy~).
-**c** Phylogenetic mean nearest taxon distance (MNTD~phy~).
-**d** Morphological MPD (MPD~morph~).
-**e** Morphological MNTD (MNTD~morph~).
+```{r endemism, fig.width = 7, fig.height = 6}
+# PE
+a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs)) +
+  geom_sf(color = "transparent") +
+  scale_fill_viridis_c() + 
+  labs(fill = "PE") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.x = element_blank()
+  )
 
-**\fig \@ref(fig:div-el).** Relationship between alpha diversity metrics and elevation (m) in native Japanese pteridophytes.
-Blue trendline shows significant linear model fit with confidence interval in pink.
-**a** species richness.
-**b** Phylogenetic mean phylogenetic distance (MPD~phy~).
-**c** Phylogenetic mean nearest taxon distance (MNTD~phy~).
-**d** Morphological MPD (MPD~morph~).
-**e** Morphological MNTD (MNTD~morph~).
+# SES PE
+b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES PE") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, pe_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, pe_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.y = element_blank()
+  )
+
+# CANAPE
+c <- ggplot(ses_div_ferns_spatial, aes(fill = endem_type)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "Endemism\ntype") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme
+
+a + b + c +
+  plot_layout(ncol = 2) +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(here::here("results/Fig_2.pdf"), width = 7, height = 6, units = "in")
+```
+
+**`r figure("endemism")`**. `r figure_cap("endemism")` **A**, phylogenetic endemism (PE). **B**, Standard effect size (SES) of PE. **C**, Categorical analysis of neo- and paleo-endemism (CANAPE).
 
@@
 
-<!-- Tables -->
+```{r bioregions, fig.width = 7, fig.height = 3}
+# Taxonomic bioregions
+a <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme
+
+# Phylogenetic bioregions
+b <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "Cluster") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme
+
+a + b +
+  plot_layout(ncol = 2) +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(here::here("results/Fig_3.pdf"), width = 7, height = 3, units = "in")
+```
+
+**`r figure("bioregions")`**.
+`r figure_cap("bioregions")` **A**, taxonomic bioregions.
+**B**, phylogenetic bioregions.
+Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances into *k* = 8 or 7 groups, respectively, using the R package \package{phyloregion}.
 
@@
 
-<!-- Figures -->
+```{r ecostructure}
+tweak <- 0.75
 
-```{r div-pteridos, fig.height = 9, fig.width = 7, fig.cap = "Alpha diversity of ferns and lycophytes in Japan"}
+sites <- ses_div_ferns_spatial %>%
+  pull(grids)
 
-# pterido richness
-  a <- make_diversity_map(
-    div_data = alpha_div_pteridos, 
-    world_map = world_map, 
-    occ_data = occ_data_pteridos, 
-    div_metric = "richness", 
-    metric_title = "No. species ",
-    label = "a") +
-    scale_fill_scico(palette = "bamako", na.value="grey")
+centroids <-
+ses_div_ferns_spatial %>%
+  select(geometry) %>%
+  mutate(centroid = sf::st_centroid(geometry)) %>%
+  sf::st_coordinates(x = .$centroid) %>%
+  as_tibble() %>%
+  mutate(site = sites)
 
-  # pterido_phy_mpd
-  b <- make_diversity_map(
-    div_data = alpha_div_pteridos, 
-    world_map = world_map, 
-    occ_data = occ_data_pteridos, 
-    div_metric = "phy_mpd_obs_z", 
-    metric_title = "SES of MPD",
-    label = "b"
-  ) + scale_fill_scico(
-    palette = "vik", 
-    na.value="dark grey",
-    limits = c(
-      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
-      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
-    ))
+omega <-
+species_motifs_ferns$omega %>%
+  as.data.frame %>%
+  rownames_to_column("site") %>%
+  left_join(centroids, by = "site") %>%
+  column_to_rownames("site")
 
-  # pterido_phy_mntd
-  c <- make_diversity_map(
-    div_data = alpha_div_pteridos, 
-    world_map = world_map, 
-    occ_data = occ_data_pteridos, 
-    div_metric = "phy_mntd_obs_z", 
-    metric_title = "SES of MNTD",
-    label = "c"
-  ) + scale_fill_scico(
-    palette = "vik", 
-    na.value="dark grey",
-    limits = c(
-      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
-      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
-    ))
+omega_centroids <- select(omega, X, Y) %>% # longitude, latitude
+  as.matrix()
 
-  # pterido_func_mpd
-  d <- make_diversity_map(
-    div_data = alpha_div_pteridos, 
-    world_map = world_map, 
-    occ_data = occ_data_pteridos, 
-    div_metric = "func_mpd_obs_z", 
-    metric_title = "SES of MPD",
-    label = "d"
-  ) + scale_fill_scico(
-    palette = "vik", 
-    na.value="dark grey",
-    limits = c(
-      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
-      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
-    ))
+omega <- select(omega, -X, -Y)
 
-  # pterido_func_mntd
-  e <- make_diversity_map(
-    div_data = alpha_div_pteridos, 
-    world_map = world_map, 
-    occ_data = occ_data_pteridos, 
-    div_metric = "func_mntd_obs_z", 
-    metric_title = "SES of MNTD",
-    label = "e"
-  ) + scale_fill_scico(
-    palette = "vik", 
-    na.value="dark grey",
-    limits = c(
-      -get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3), 
-      get_limit(alpha_div_pteridos, phy_mpd_obs_z, "abs", 3)
-    ))
+ecos_plot_pie2(
+    omega = omega,
+    image_width = 6000*tweak,
+    image_height = 4800*tweak,
+    radius = 0.08,
+    no_map = TRUE,
+    coords = omega_centroids,
+    long_lim = c(121, 148),
+    lat_lim = c(23, 47),
+    color = carto_pal(8, "Bold"),
+    blank_bg = "grey90",
+    path = here::here("results/species_motifs_ferns.png")
+)
 
-  f <- make_diversity_map(
-  div_data = alpha_div_pteridos, 
-  world_map = world_map, 
-  occ_data = occ_data_pteridos, 
-  div_metric = "percent_sex_dip",
-  metric_title = "% sex. diploid",
-  label = "f"
-)  +
-    scale_fill_scico(palette = "bamako", na.value="grey")
-
-  a + b + c + d + e + f + plot_layout(ncol = 2, nrow = 3)
-
-ggsave(here::here("ms/fig_1.pdf"), width = 7, height = 9, units = "in")
 ```
 
-```{r div-el, fig.height = 3.5, fig.width = 7, fig.cap = "Relationship between alpha diversity metrics and elevation"}
-a <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "richness", "Elevation (m)", "Richness") +
-  labs(title = "a", x = "Elevation (m)")
-b <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "phy_mpd_obs_z", "Elevation (m)", "MPDphy") +
-  labs(title = "b", x = "Elevation (m)")
-c <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "phy_mntd_obs_z", "Elevation (m)", "MNTDphy") +
-  labs(title = "c", x = "Elevation (m)")
-d <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "func_mpd_obs_z", "Elevation (m)", "MPDfunc") +
-  labs(title = "d", x = "Elevation (m)")
-e <- make_scatter_with_lm(alpha_div_pteridos, "elevation", "func_mntd_obs_z", "Elevation (m)", "MNTDfunc") +
-  labs(title = "e", x = "Elevation (m)")
-a + b + c + d + e + plot_layout(ncol = 3, nrow = 2)
+![](`r here::here("results/species_motifs_ferns.png")`)
 
-ggsave(here::here("ms/fig_2.pdf"), width = 7, height = 3.5, units = "in")
-```
+**`r figure("ecostructure")`**.
+`r figure_cap("ecostructure")`.
+Pie charts represent the proportional contribution of each site to each of *k* = 8 spatial motifs, fit with a grade-of-membership model based on the community matrix using the R package \package{ecostructure}.

### 2020-06-28 R/packages.R

@@ -21,6 +21,8 @@ library(scico)
 library(clustermq)
 library(magrittr)
 library(glue)
+library(patchwork)
+library(rcartocolor)
 library(tidyverse)
 
 # Resolve conflicts

### 2020-06-28 R/plan.R

@@ -68,12 +68,6 @@ plan <- drake_plan (
     transform = map(scale = c(0.2, 0.4))
   ),
   
-  # Check coverage at each scale
-  # comm_scaled_coverage = target(
-  #   calculate_coverage(comm_scaled_list),
-  #   transform = map(comm_scaled_list, .id = scale)
-  # ),
-  
   # Decide that 0.2 scale is optimal, use this for downstream analyses
   comm_ferns = comm_from_points2comm(comm_scaled_list_0.2),
   
@@ -150,14 +144,25 @@ plan <- drake_plan (
   regions_phylogeny = cluster_phylo_regions(
     comm_df = comm_ferns,
     phy = japan_fern_tree,
-    k = 12
+    k = 7
   ),
 
   # Ecostructure ----
 
   species_motifs_ferns = ecostructure::ecos_fit(
       dat = comm_ferns,
-      K = 8, tol = 0.1, num_trials = 1)
+      K = 8, tol = 0.1, num_trials = 1),
+  
+  # Write out manuscript ----
+  ms = rmarkdown::render(
+    knitr_in("ms/manuscript.Rmd"),
+    output_dir = here::here("results"),
+    quiet = TRUE),
+  
+  si = rmarkdown::render(
+    knitr_in("ms/SI.Rmd"),
+    output_dir = here::here("results"),
+    quiet = TRUE)
 
   # # Make richness matrix (number of species per
   # # 10 km grid cell).

### 2020-06-28 ms/captions.R

@@ -1,0 +1,35 @@
+# Set up captions
+
+# - Figures
+fig_nums <- captioner::captioner(prefix = "Fig.")
+fig_nums(name = "diversity", caption = "Taxonomic, phylogenetic, and morphological diversity of the ferns of Japan.")
+fig_nums(name = "endemism", caption = "Phylogenetic endemism of the ferns of Japan.")
+fig_nums(name = "bioregions", caption = "Bioregions of the ferns of Japan.")
+fig_nums(name = "ecostructure", caption = "Ecological structure of the ferns of Japan.")
+
+# - Tables
+table_nums <- captioner::captioner(prefix = "Table")
+table_nums(name = "traits-used", caption = "Fern traits used in this study")
+
+# - SI figures
+s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
+s_fig_nums(name = "k-plot", "Selection of *k* for bioregions.")
+s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan, restricted dataset including only taxa endemic to Japan.")
+
+# - SI tables
+# s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
+# s_table_nums(name = "si_table", "")
+
+# Make short versions of citation functions
+
+# - Just the number
+figure <- pryr::partial(fig_nums, display = "cite")
+table <- pryr::partial(table_nums, display = "cite")
+s_figure <- pryr::partial(s_fig_nums, display = "cite")
+# s_table <- pryr::partial(s_table_nums, display = "cite")
+
+# - Just the caption
+figure_cap <- function(x) {fig_nums(x) %>% stringr::str_match(": (.*)$") %>% magrittr::extract(,2)}
+table_cap <- function(x) {table_nums(x) %>% stringr::str_match(": (.*)$") %>% magrittr::extract(,2)}
+s_figure_cap <- function(x) {s_fig_nums(x) %>% stringr::str_match(": (.*)$") %>% magrittr::extract(,2)}
+# s_table_cap <- function(x) {s_table_nums(x) %>% stringr::str_match(": (.*)$") %>% magrittr::extract(,2)}

### 2020-06-28 ms/manuscript.Rmd
Update plan
- add back in analysis of multiple k values for ecostructure
- track phylogeny file
- move categorize_endemism() to immediately after running ses analysis

@@
 tax_regions_data <- regions_taxonomy %>%
-  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
+  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  select(grids, taxonomic_cluster)
 
 phy_regions_data <- regions_phylogeny %>%
-  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
+  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  select(grids, phylo_cluster)
 
@@
   shape_ferns %>%
-  left_join(categorize_endemism(ses_phy_ferns), by = c(grids = "site")) %>%
+  left_join(ses_phy_ferns, by = c(grids = "site")) %>%
   left_join(ses_traits_ferns, by = c(grids = "site")) %>%
-  mutate(redundancy = 1 - (richness/abundance)) %>%
   left_join(tax_regions_data, by = "grids") %>%
@@
 shape_ferns %>%
-  left_join(categorize_endemism(ses_phy_ferns_endemic), by = c(grids = "site"))
+  left_join(ses_phy_ferns_endemic, by = c(grids = "site"))
 
@@
 Pie charts represent the proportional contribution of each site to each of *k* = 8 spatial motifs, fit with a grade-of-membership model based on the community matrix using the R package \package{ecostructure}.
+
+\clearpage
+
+```{r diversity-by-cluster}
+ses_div_ferns_spatial %>%
+  filter(taxonomic_cluster %in% c(1,2,3,4)) %>%
+  as_tibble %>%
+  select(mpd_obs_z, mpd_morph_obs_z, pe_obs_z, taxonomic_cluster) %>%
+  pivot_longer(names_to = "var", values_to = "value", -taxonomic_cluster) %>%
+  ggplot(aes(x = taxonomic_cluster, y = value)) +
+  geom_jitter() +
+  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.5) +
+  facet_wrap(vars(var)) +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+  )
+```
+
+Standard effect size (SES) of mean phylogenetic distance (MPD), mean morphological distance (MMD), and phylogenetic endemism (PE) by taxonomic bioregion (top four regions with the most sites only shown).

### 2020-06-28 R/plan.R

@@ -1,5 +1,5 @@
 # Set vector of k-values to use for ecostructure
-k_vals <- 2:10
+k_vals <- 2:20
 
 # Define analysis plan
 plan <- drake_plan (
@@ -69,10 +69,16 @@ plan <- drake_plan (
   ),
   
   # Decide that 0.2 scale is optimal, use this for downstream analyses
+  
+  # - extract community matrix
   comm_ferns = comm_from_points2comm(comm_scaled_list_0.2),
   
-  shape_ferns = shape_from_points2comm(comm_scaled_list_0.2),
+  # - extract geographic shapes, richness, and number of specimens
+  shape_ferns = shape_from_points2comm(comm_scaled_list_0.2) %>%
+    # calculate redundancy
+    mutate(redundancy = 1 - (richness/abundance)),
     
+  # - make community matrix subset to taxa endemic to Japan
   comm_ferns_endemic = subset_comm_to_endemic(
     comm = comm_ferns,
     green_list = green_list
@@ -80,7 +86,9 @@ plan <- drake_plan (
   
   # Read in ultrametric phylogenetic tree of all pteridophytes,
   # not including hybrids (706 taxa total)
-  japan_pterido_tree = ape::read.tree("data_raw/japan_pterido_tree_dated.tre"),
+  japan_pterido_tree_path = target("data_raw/japan_pterido_tree_dated.tre", format = "file"),
+  
+  japan_pterido_tree = ape::read.tree(japan_pterido_tree_path),
   
   # - subset to only ferns
   japan_fern_tree = subset_tree(
@@ -106,7 +114,7 @@ plan <- drake_plan (
   # Analyze standard effect size (SES) of diversity metrics ----
 
   ses_phy = target(
-    run_ses_analysis(comm, japan_fern_tree, n_reps = 999, metrics),
+    run_ses_analysis(comm, japan_fern_tree, n_reps = 999, metrics) %>% categorize_endemism,
     transform = map(
       comm = c(comm_ferns, comm_ferns_endemic),
       metrics = c(
@@ -149,9 +157,12 @@ plan <- drake_plan (
 
   # Ecostructure ----
 
-  species_motifs_ferns = ecostructure::ecos_fit(
-      dat = comm_ferns,
-      K = 8, tol = 0.1, num_trials = 1),
+  species_motifs_ferns = target(
+      ecostructure::ecos_fit(
+        comm_ferns,
[…]

### 2020-06-29 ms/manuscript.Rmd
Update MS

@@
 
-Characterizing biodiversity is important for two reasons: to understand the processes generating it, and to conserve it.
+Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology for two reasons: to understand the processes generating it, and to conserve it.
+The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on one metric, species richness.
+Much progress has been made documenting hotspots of exceptional species richness, the processes generating them, and enabling conservation strategies to preserve them.
+However, richness alone cannot provide a complete picture of biodiversity.
+All organisms are related by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of biodiversity.
 
-Most past studies of biodiversity only focus on one dimension: richness.
-
-Recently tools and data are available to analyze additional dimensions (in particular phylogenetic dimensions).
-Furthermore, better ways to characterize biological regions ("bioregions") are also available, rather than simply using expert opinion or other ad-hoc characterizations.
-
-Incorporating two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.
+Presumably, the main reason phylogeny has not been taken into account more promimently in biodiversity studies is because data (DNA sequences) and analytical tools have only become available relatively recently.
+These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism.
+Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations.
+Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.
 

### 2020-06-30 ms/SI.Rmd
Update SI

@@
 **C**, Categorical analysis of neo- and paleo-endemism (CANAPE).
+
+```{r biplots}
+
+loadd(ses_phy_ferns, cache = ja_fern_cache)
+loadd(ses_traits_ferns, cache = ja_fern_cache)
+loadd(shape_ferns, cache = ja_fern_cache)
+
+# Helper function for making a scatterplot
+scatterplot <- function(data, x_var, y_var, x_lab, y_lab) {
+  ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
+    geom_point(alpha = 0.5) +
+    labs(
+      x = x_lab,
+      y = y_lab
+    )
+}
+
+# Join data
+ses_comp <-
+ses_phy_ferns %>%
+  select(site, mpd_obs, pe_obs, mpd_obs_z, pe_obs_z) %>%
+  left_join(
+    select(ses_traits_ferns, site, mpd_morph_obs, mpd_morph_obs_z),
+    by = "site"
+  ) %>%
+  left_join(
+    as_tibble(shape_ferns) %>% select(site = grids, richness),
+    by = "site"
+  )
+
+# Raw metrics ----
+
+# Make dataframe of arguments to feed into plotting function
+# - includes all non-duplicated combinations of raw diversity metrics
+raw_comparisons <-
+  cross_df(
+    list(
+      x_var = c("mpd_obs", "pe_obs", "mpd_morph_obs", "richness"),
+      y_var = c("mpd_obs", "pe_obs", "mpd_morph_obs")
+    ),
+    .filter = `==` # filter out exact duplicates
+  ) %>%
+  # Remove reversed duplicates
+  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
+  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
+  # Use factors to keep the preferred combination of x and y
+  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs", "mpd_morph_obs", "richness")) %>%
+  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs", "pe_obs", "mpd_morph_obs")) %>%
+  arrange(x_var, y_var) %>%
+  slice(1) %>%
+  ungroup() %>%
+  select(-grp) %>%
+  # Convert x_var and y_var back to character for plotting function
+  mutate(across(c(x_var, y_var), as.character)) %>%
+  # Add labels
+  mutate(
+    x_lab = case_when(
+      x_var == "mpd_morph_obs" ~ "MMD",
+      x_var == "mpd_obs" ~ "MPD",
+      x_var == "pe_obs" ~ "PE",
+      x_var == "richness" ~ "Richness"
+    ),
+    y_lab = case_when(
+      y_var == "mpd_morph_obs" ~ "MMD",
+      y_var == "mpd_obs" ~ "MPD",
+      y_var == "pe_obs" ~ "PE",
+      y_var == "richness" ~ "Richness"
+    )
+  )
+
+# Make plot
+pmap(raw_comparisons, scatterplot, data = ses_comp) %>%
+wrap_plots() +
+  plot_annotation(tag_levels = 'A')
+
+# SES metrics ----
+
+# Make dataframe of arguments to feed into plotting function
+# - includes all non-duplicated combinations of raw diversity metrics
+ses_comparisons <-
+  cross_df(
+    list(
+      x_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z", "richness"),
+      y_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")
+    ),
+    .filter = `==` # filter out exact duplicates
+  ) %>%
+  # Remove reversed duplicates
+  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
+  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
+  # Use factors to keep the preferred combination of x and y
+  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs_z", "mpd_morph_obs_z", "richness")) %>%
+  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")) %>%
+  arrange(x_var, y_var) %>%
+  slice(1) %>%
+  ungroup() %>%
+  select(-grp) %>%
+  # Convert x_var and y_var back to character for plotting function
+  mutate(across(c(x_var, y_var), as.character)) %>%
+  # Add labels
+  mutate(
+    x_lab = case_when(
+      x_var == "mpd_morph_obs_z" ~ "SES MMD",
+      x_var == "mpd_obs_z" ~ "SES MPD",
+      x_var == "pe_obs_z" ~ "SES PE",
+      x_var == "richness" ~ "Richness"
+    ),
+    y_lab = case_when(
+      y_var == "mpd_morph_obs_z" ~ "SES MMD",
+      y_var == "mpd_obs_z" ~ "SES MPD",
+      y_var == "pe_obs_z" ~ "SES PE",
+      y_var == "richness" ~ "Richness"
+    )
+  )
+
+# Make plot
+pmap(ses_comparisons, scatterplot, data = ses_comp) %>%
+  wrap_plots() +
+  plot_annotation(tag_levels = 'A')
+```

### 2020-06-30 ms/manuscript.Rmd
Use species_motifs_ferns_8L instead of species_motifs_ferns

@@
   "regions_phylogeny",
-  "species_motifs_ferns"
+  "species_motifs_ferns_8L"
 ),
@@
 omega <-
-species_motifs_ferns$omega %>%
+species_motifs_ferns_8L$omega %>%
   as.data.frame %>%

### 2020-06-30 ms/manuscript.Rmd
Update manuscript

@@
 Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology for two reasons: to understand the processes generating it, and to conserve it.
-The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on one metric, species richness.
-Much progress has been made documenting hotspots of exceptional species richness, the processes generating them, and enabling conservation strategies to preserve them.
+The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness.
+For example, hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots and motivating conservation strategies to preserve them.
 However, richness alone cannot provide a complete picture of biodiversity.
@@
 
-Presumably, the main reason phylogeny has not been taken into account more promimently in biodiversity studies is because data (DNA sequences) and analytical tools have only become available relatively recently.
+Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences) and analytical tools have only become available relatively recently.
 These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism.

### 2020-06-30 ms/SI.Rmd
Update SI
- add function for automatically generating results files based on caption name
- add raw and SES diversity biplots

@@
 
-phylo_cache <- new_cache(here::here("phylo_cache"))
-
-loadd(list = c(
-  "plastome_calibration_dates",
-  "plastome_tree"
-),
-cache = phylo_cache)
-
 # Load captions
@@
 
+ggsave(result_file("k-plot", "pdf"), width = 7, height = 3, units = "in")
 ```
@@
 
-ggsave(here::here("results/Fig_S1.pdf"), width = 7, height = 6, units = "in")
+ggsave(result_file("endemism-restricted", "pdf"), width = 7, height = 6, units = "in")
 ```
@@
 
-```{r biplots}
-
-loadd(ses_phy_ferns, cache = ja_fern_cache)
-loadd(ses_traits_ferns, cache = ja_fern_cache)
-loadd(shape_ferns, cache = ja_fern_cache)
-
+```{r biplot-prep}
 # Helper function for making a scatterplot
@@
 
-# Join data
-ses_comp <-
-ses_phy_ferns %>%
-  select(site, mpd_obs, pe_obs, mpd_obs_z, pe_obs_z) %>%
-  left_join(
-    select(ses_traits_ferns, site, mpd_morph_obs, mpd_morph_obs_z),
-    by = "site"
-  ) %>%
-  left_join(
-    as_tibble(shape_ferns) %>% select(site = grids, richness),
-    by = "site"
-  )
+```
 
-# Raw metrics ----
+```{r raw-biplot, fig.width = 7, fig.height = 6}
+## Biplots of raw metrics ##
 
@@
 # Make plot
-pmap(raw_comparisons, scatterplot, data = ses_comp) %>%
+pmap(raw_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
 wrap_plots() +
@@
 
-# SES metrics ----
+ggsave(result_file("raw-biplot", "pdf"), width = 7, height = 6, units = "in")
+```
+
+**`r s_figure("raw-biplot")`**. `r s_figure_cap("raw-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism.
+
+```{r ses-biplot, fig.width = 7, fig.height = 6}
+## Biplots of SES metrics ##
 
@@
 # Make plot
-pmap(ses_comparisons, scatterplot, data = ses_comp) %>%
+pmap(ses_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
   wrap_plots() +
   plot_annotation(tag_levels = 'A')
+
+ggsave(result_file("ses-biplot", "pdf"), width = 7, height = 6, units = "in")
 ```
+
+**`r s_figure("ses-biplot")`**. `r s_figure_cap("ses-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism.

### 2020-06-30 ms/SI.Rmd
Update MS and SI
add modified t-test to biplots

--- 2020-06-28 ms/SI.Rmd
@@
 ```{r plot-prep}
+# Combine bioregion results, biodiversity metrics, and shape data into single dataframe
 tax_regions_data <- regions_taxonomy %>%
-  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
+  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  select(grids, taxonomic_cluster)
 
 phy_regions_data <- regions_phylogeny %>%
-  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
+  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  select(grids, phylo_cluster)
 
@@
 
+# - same for endemic species
 ses_div_ferns_endemic_spatial <-
 shape_ferns %>%
-  left_join(categorize_endemism(ses_phy_ferns_endemic), by = c(grids = "site"))
+  left_join(ses_phy_ferns_endemic, by = c(grids = "site"))
 
+# Define ggplot2 theme for maps
 map_theme <- theme(
@@
   )
+
 ```
@@
 **C**, Categorical analysis of neo- and paleo-endemism (CANAPE).
+
+```{r biplot-prep}
+# Extract centroids
+centroids <- sf::st_centroid(ses_div_ferns_spatial$geometry) %>%
+  sf::st_coordinates()
+
+# Define helper function for making a scatterplot
+scatterplot <- function(data, x_var, y_var, x_lab, y_lab, label, annotate) {
+  p <- ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
+    geom_point(alpha = 0.5) +
+    labs(
+      x = x_lab,
+      y = y_lab
+    )
+
+  # Add annotation and trendline if annotate = TRUE
+  if(annotate) p <- p +
+      annotate("label", x = Inf, y = Inf, label = label, hjust = 1, vjust = 1, label.size = 0, fill = "transparent") +
+      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
+
+  p
+}
+```
+
+```{r raw-biplot, fig.width = 7, fig.height = 6}
+## Biplots of raw metrics ##
+
+# Make dataframe of arguments to feed into plotting function
+# - includes all non-duplicated combinations of raw diversity metrics
+raw_comparisons <-
+  cross_df(
+    list(
+      x_var = c("mpd_obs", "pe_obs", "mpd_morph_obs", "richness"),
+      y_var = c("mpd_obs", "pe_obs", "mpd_morph_obs")
+    ),
+    .filter = `==` # filter out exact duplicates
+  ) %>%
+  # Remove reversed duplicates
+  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
+  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
+  # Use factors to keep the preferred combination of x and y
+  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs", "mpd_morph_obs", "richness")) %>%
+  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs", "pe_obs", "mpd_morph_obs")) %>%
+  arrange(x_var, y_var) %>%
+  slice(1) %>%
+  ungroup() %>%
+  select(-grp) %>%
+  # Convert x_var and y_var back to character for plotting function
+  mutate(across(c(x_var, y_var), as.character)) %>%
+  # Add labels
+  mutate(
+    x_lab = case_when(
+      x_var == "mpd_morph_obs" ~ "MMD",
+      x_var == "mpd_obs" ~ "MPD",
+      x_var == "pe_obs" ~ "PE",
+      x_var == "richness" ~ "Richness"
+    ),
+    y_lab = case_when(
+      y_var == "mpd_morph_obs" ~ "MMD",
+      y_var == "mpd_obs" ~ "MPD",
+      y_var == "pe_obs" ~ "PE",
+      y_var == "richness" ~ "Richness"
+    )
+  ) %>%
+  # Run modified t-test
+  mutate(
+    test_res = map2(
+      .x = x_var, 
+      .y = y_var, 
+      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
+    # Extract relevant stats from results
+    cor = map_dbl(test_res, pluck("corr")),
+    pval = map_dbl(test_res, pluck("p.value"))
+    ) %>%
+  # Generate label
+  mutate(
+    cor_lab = glue("r = {round(cor, 3)}"),
+    pval_lab = glue("p = {round(pval, 3)}"),
+    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
+  ) %>%
+  # Specify which plots to annotate
+  mutate(
+    annotate = case_when(
+      x_var == "mpd_morph_obs" & y_var == "mpd_obs" & pval < 0.05 ~ TRUE,
+      x_var == "mpd_morph_obs" & y_var == "pe_obs" & pval < 0.05 ~ TRUE,
+      x_var == "mpd_obs" & y_var == "pe_obs" & pval < 0.05 ~ TRUE,
+      x_var == "richness" & y_var == "pe_obs" & pval < 0.05  ~ TRUE,
+      TRUE ~ FALSE
+    )
+  ) %>%
+  select(x_var, y_var, x_lab, y_lab, label, annotate)
+
+# Make plot
+pmap(raw_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
+wrap_plots() +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(result_file("raw-biplot", "pdf"), width = 7, height = 6, units = "in")
+```
+
+**`r s_figure("raw-biplot")`**.
+`r s_figure_cap("raw-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism.
+Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05; correlations between MPD and MMD with richness not expected to be linear, so were not tested.
+
+```{r ses-biplot, fig.width = 7, fig.height = 6}
+## Biplots of SES metrics ##
+
+# Make dataframe of arguments to feed into plotting function
+# - includes all non-duplicated combinations of raw diversity metrics
+ses_comparisons <-
+  cross_df(
+    list(
+      x_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z", "richness"),
+      y_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")
+    ),
+    .filter = `==` # filter out exact duplicates
+  ) %>%
+  # Remove reversed duplicates
+  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
+  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
+  # Use factors to keep the preferred combination of x and y
+  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs_z", "mpd_morph_obs_z", "richness")) %>%
+  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")) %>%
+  arrange(x_var, y_var) %>%
+  slice(1) %>%
+  ungroup() %>%
+  select(-grp) %>%
+  # Convert x_var and y_var back to character for plotting function
+  mutate(across(c(x_var, y_var), as.character)) %>%
+  # Add labels
+  mutate(
+    x_lab = case_when(
+      x_var == "mpd_morph_obs_z" ~ "SES MMD",
+      x_var == "mpd_obs_z" ~ "SES MPD",
+      x_var == "pe_obs_z" ~ "SES PE",
+      x_var == "richness" ~ "Richness"
+    ),
+    y_lab = case_when(
+      y_var == "mpd_morph_obs_z" ~ "SES MMD",
+      y_var == "mpd_obs_z" ~ "SES MPD",
+      y_var == "pe_obs_z" ~ "SES PE",
+      y_var == "richness" ~ "Richness"
+    )
+  ) %>%
+  # Run modified t-test
+  mutate(
+    test_res = map2(
+      .x = x_var, 
+      .y = y_var, 
+      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
+    # Extract relevant stats from results
+    cor = map_dbl(test_res, pluck("corr")),
+    pval = map_dbl(test_res, pluck("p.value"))
+    ) %>%
+  # Generate label
+  mutate(
+    cor_lab = glue("r = {round(cor, 3)}"),
+    pval_lab = glue("p = {round(pval, 3)}"),
+    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
+  ) %>%
+  # Specify which plots to annotate
+  mutate(
+    annotate = case_when(
+      pval < 0.05 ~ TRUE,
+      TRUE ~ FALSE
+    )
+  ) %>%
+  select(x_var, y_var, x_lab, y_lab, label, annotate)
+
+# Make plot
+pmap(ses_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
+  wrap_plots() +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(result_file("ses-biplot", "pdf"), width = 7, height = 6, units = "in")
+```
+
+**`r s_figure("ses-biplot")`**.
+`r s_figure_cap("ses-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism.
+Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; all combinations tested, but *p* value and correlation (r) only shown for *p* < 0.05.

### 2020-06-30 ms/manuscript.Rmd
Update MS and SI
add modified t-test to biplots

@@
 A single motif united low-richness sites (dark purple).
+
+### Relationships between metrics
+
+Raw values of MPD and MMD showed high variance and low species richness, then merged towards a single value at high species richness, as expected (`r s_figure("raw-biplot")`D, E).
+Raw PE was linearly correlated with species richness (`r s_figure("raw-biplot")`F).
+Both raw and SES values of MPD and MMD were highly correlated (`r s_figure("raw-biplot")`A, `r s_figure("ses-biplot")`A, B).
+SES values of MPD, MMD, and PE were all positively correlated (`r s_figure("ses-biplot")`A, B, C).
+None of the SES values were correlated with richness (`r s_figure("ses-biplot")`D, E, F).
 

### 2020-06-30 ms/SI.Rmd
Change k for phylo bioregions from 7 to 6
Add k plot for ecostructure

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
 ```
@@
   "regions_phylogeny",
-  "species_motifs_ferns"
+  "species_motifs_ferns",
+  "species_motifs_ferns_combined"
 ),
@@
 
-k_phylogeny_select <- 7
+k_phylogeny_select <- 6
 
@@
 
-<!-- (Add k plot for ecostructure) -->
+\clearpage
+
+```{r k-plot-ecos, fig.width = 4, fig.height = 3}
+# Extract log likelihoods for each value of K from ecostructure results
+log_l <- map_dbl(species_motifs_ferns_combined, ~pluck(., "L")) %>%
+  tibble(k = names(.) %>% parse_number, log_l = .) %>%
+  mutate(
+    optimal_manual = ifelse(k == k_taxonomy_select, TRUE, FALSE)
+  )
+
+ggplot(log_l, aes(x = k, y = log_l)) +
+  geom_line() +
+  geom_point(aes(color = optimal_manual), size = 2.5) +
+  scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
+  labs(
+    y = "Log likelihood",
+    x = "Number of clusters") +
+  theme(legend.position = "none")
+
+ggsave(result_file("k-plot-ecos", "pdf"), width = 4, height = 3, units = "in")
+```
+
+**`r s_figure("k-plot-ecos")`**. `r s_figure_cap("k-plot-ecos")` Red indicates optimal *k* selected manually (*k* = 8, same as for taxonomic bioregions).
 

### 2020-06-30 ms/manuscript.Rmd
Add k plot for ecostructure

@@
 Like bioregion analysis, this method also requires selecting the number of clusters (in this context called "motifs" after @White2019) to fit the model.
-We initially tested values of *k* from 1 to 20 and inspected the log-likelihood for each, but found that it continually improved without reaching a clear optimum.
+We initially tested values of *k* from 2 to 20 and inspected the log-likelihood for each, but found that it continually improved without reaching a clear optimum (`r s_figure("k-plot-ecos")`).
 We therefore used the same *k* that was selected for taxonomic bioregions.

### 2020-06-30 R/plan.R

@@ -164,6 +164,11 @@ plan <- drake_plan (
       transform = map(K = !!k_vals)
     ),
   
+  species_motifs_ferns_combined = target(
+    list(species_motifs_ferns) %>% set_names(k_vals),
+    transform = combine(species_motifs_ferns)
+  ),
+  
   # Write out manuscript ----
   ms = rmarkdown::render(
     knitr_in("ms/manuscript.Rmd"),

### 2020-06-30 ms/captions.R

@@ -14,6 +14,7 @@ table_nums(name = "traits-used", caption = "Fern traits used in this study")
 # - SI figures
 s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
 s_fig_nums(name = "k-plot", "Selection of *k* for bioregions.")
+s_fig_nums(name = "k-plot-ecos", "Selection of *k* for ecological structure analysis.")
 s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan, restricted dataset including only taxa endemic to Japan.")
 s_fig_nums(name = "raw-biplot", "Relationships between raw biodiversity metrics of the ferns of Japan.")
 s_fig_nums(name = "ses-biplot", "Relationships between standard effect size (SES) of biodiversity metrics of the ferns of Japan.")

### 2020-07-01 ms/SI.Rmd
Subset traits to ferns only, add NMDS plot

@@
   "species_motifs_ferns",
-  "species_motifs_ferns_combined"
+  "species_motifs_ferns_combined",
+  "traits_nmds",
+  "ppgi"
 ),
@@
 Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; all combinations tested, but *p* value and correlation (r) only shown for *p* < 0.05.
+
+```{r trait-space, fig.width = 8, fig.height = 8}
+# Extract points, join to PPGI
+nmds_points <- traits_nmds[["points"]] %>%
+  as.data.frame %>%
+  rownames_to_column("taxon") %>%
+  as_tibble %>%
+  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
+  left_join(ppgi, by = "genus") %>%
+  mutate(
+    order = factor(order),
+    family = factor(family))
+
+# A: all ferns
+a <- ggplot(nmds_points, aes(x = MDS1, y = MDS2, color = order, shape = order)) +
+  geom_point() +
+  scale_shape_manual(values = 1:n_distinct(nmds_points$order)) +
+  labs(title = "A. Ferns")
+
+# B: Polypods only
+nmds_points_polypod <- filter(nmds_points, order == "Polypodiales")
+
+b <- ggplot(nmds_points_polypod, aes(x = MDS1, y = MDS2, color = family, shape = family)) +
+  geom_point() +
+  scale_shape_manual(values = 1:n_distinct(nmds_points_polypod$family)) +
+  labs(title = "B. Polypodiales") +
+  guides(
+    color = guide_legend(ncol = 2),
+    shape = guide_legend(ncol = 2))
+
+a + b + 
+  plot_layout(nrow = 2)
+
+ggsave(result_file("traits-nmds", "pdf"), width = 8, height = 8, units = "in")
+```
+
+**`r s_figure("traits-nmds")`**.
+`r s_figure_cap("traits-nmds")` Interspecific trait distances calculated using Gower's method with 16 traits, then projected into trait space using nonmetric multidimensional scaling (NMDS).
+**A**, All ferns.
+**B**, Polypodiales.
+Taxonomy follows PPGI.

### 2020-07-01 R/functions.R

@@ -161,18 +161,18 @@ process_occ_data <- function (occ_data_raw, all_cells, ppgi) {
   
 }
 
-#' Subset occurrence point data to only ferns
+#' Subset data to only ferns
 #'
-#' @param occ_point_data_raw Raw occurrence point data (one row per specimen)
+#' @param data Input dataframe with column "taxon"
 #' @param ppgi Pteridophyte phylogeny group I taxonomy
 #'
 #' @return Dataframe subset to ferns only
 #' 
-subset_occ_point_data <- function(occ_point_data_raw, ppgi) {
+subset_to_ferns <- function(data, ppgi) {
   
-  cols_keep <- colnames(occ_point_data_raw)
+  cols_keep <- colnames(data)
   
-  occ_point_data_raw %>%
+  data %>%
     mutate(genus = str_split(taxon, " |_") %>% map_chr(1)) %>%
     assert(not_na, genus) %>%
     # Add higher-level taxonomy
@@ -180,8 +180,6 @@ subset_occ_point_data <- function(occ_point_data_raw, ppgi) {
     assert(not_na, class) %>%
     # Filter to only ferns
     filter(class == "Polypodiopsida") %>%
-    # Check for missing data
-    assert(not_na, longitude, latitude, taxon) %>%
     select(all_of(cols_keep))
   
 }
@@ -849,15 +847,14 @@ make_trait_dist_matrix <- function (traits_for_dist) {
 #'
 #' @return GGPlot object
 #'
-make_trait_nmds_plot <- function (nmds, ppgi, taxon_id_map) {
+make_trait_nmds_plot <- function (nmds, ppgi) {
   
   nmds_points = nmds[["points"]] %>%
     as.data.frame %>%
-    rownames_to_column("taxon_id") %>%
+    rownames_to_column("taxon") %>%
     as_tibble %>%
-    left_join(taxon_id_map) %>%
     mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
-    left_join(ppgi) %>%
+    left_join(ppgi, by = "genus") %>%
     mutate(
       order = factor(order),
       family = factor(family))
@@ -865,7 +862,7 @@ make_trait_nmds_plot <- function (nmds, ppgi, taxon_id_map) {
   a <- ggplot(nmds_points, aes(x = MDS1, y = MDS2, color = order, shape = order)) +
     geom_point() +
[…]

### 2020-07-01 R/plan.R

@@ -49,11 +49,10 @@ plan <- drake_plan (
   # - standardize names to Green List
   occ_point_data = rename_taxa(occ_point_data_raw, green_list),
   
-  # Subset to just ferns (674 taxa)
-  occ_point_data_ferns = subset_occ_point_data(
-    occ_point_data_raw = occ_point_data,
-    ppgi = ppgi
-  ),
+  # - subset to just ferns (674 taxa)
+  occ_point_data_ferns = subset_to_ferns(occ_point_data, ppgi) %>%
+    # check for missing data
+    assert(not_na, longitude, latitude, taxon),
   
   # Generate datasets at different spatial scales: 
   # 0.2 degree ~ 20 km x 20 km,
@@ -95,12 +94,13 @@ plan <- drake_plan (
     phy = japan_pterido_tree, 
     ppgi = ppgi),
   
-  # Format trait data
+  # Format trait data, subset to ferns
   raw_trait_data_path = target("data_raw/JpFernLUCID_forJoel.xlsx", format = "file"),
   
   traits_for_dist = format_traits(
     path_to_lucid_traits = raw_trait_data_path,
-    taxon_id_map = green_list),
+    taxon_id_map = green_list) %>%
+    subset_to_ferns(ppgi),
 
   # Make trait distance matrix using taxon IDs as labels
   trait_distance_matrix = make_trait_dist_matrix(traits_for_dist),
@@ -169,6 +169,11 @@ plan <- drake_plan (
     transform = combine(species_motifs_ferns)
   ),
   
+  # Traits ----
+  
+  # Run NMDS on traits
+  traits_nmds = vegan::metaMDS(trait_distance_matrix, k = 3),
+  
   # Write out manuscript ----
   ms = rmarkdown::render(
     knitr_in("ms/manuscript.Rmd"),
@@ -178,7 +183,7 @@ plan <- drake_plan (
   si = rmarkdown::render(
     knitr_in("ms/SI.Rmd"),
     output_dir = here::here("results"),
-    quiet = TRUE)
+    quiet = TRUE),
 
   # # Make richness matrix (number of species per
   # # 10 km grid cell).

### 2020-07-01 ms/captions.R

@@ -18,6 +18,7 @@ s_fig_nums(name = "k-plot-ecos", "Selection of *k* for ecological structure anal
 s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan, restricted dataset including only taxa endemic to Japan.")
 s_fig_nums(name = "raw-biplot", "Relationships between raw biodiversity metrics of the ferns of Japan.")
 s_fig_nums(name = "ses-biplot", "Relationships between standard effect size (SES) of biodiversity metrics of the ferns of Japan.")
+s_fig_nums(name = "traits-nmds", "Trait space of the ferns of Japan.")
 
 # - SI tables
 # s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")

### 2020-07-02 ms/SI.Rmd
Add PD and RPE

@@
 
-ggsave(result_file("k-plot", "pdf"), width = 7, height = 3, units = "in")
+ggsave(result_file(s_figure("k-plot"), "pdf"), width = 7, height = 3, units = "in")
 ```
@@
 
-ggsave(result_file("k-plot-ecos", "pdf"), width = 4, height = 3, units = "in")
+ggsave(result_file(s_figure("k-plot-ecos"), "pdf"), width = 4, height = 3, units = "in")
 ```
@@
 
-```{r endemism-restricted, fig.width = 7, fig.height = 6}
+```{r endemism-restricted, fig.width = 7, fig.height = 8}
 # PE (endemic)
@@
   theme(
-    axis.text.y = element_blank()
+    axis.text.y = element_blank(),
+    axis.text.x = element_blank()
   )
 
+# RPE
+c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs)) +
+  geom_sf(color = "transparent") +
+  scale_fill_viridis_c() + 
+  labs(fill = "RPE") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.x = element_blank()
+  )
+
+# SES RPE
+d <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES RPE") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs"),
+      get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(axis.text.y = element_blank())
+
 # CANAPE (endemic)
-c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = endem_type)) +
+e <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = endem_type)) +
   geom_sf(color = "transparent") +
@@
 
-a + b + c + 
+a + b + c + d + e +
   plot_layout(ncol = 2) +
@@
 
-ggsave(result_file("endemism-restricted", "pdf"), width = 7, height = 6, units = "in")
+ggsave(result_file(s_figure("endemism-restricted"), "pdf"), width = 7, height = 8, units = "in")
 ```
@@
 **B**, Standard effect size (SES) of PE.
-**C**, Categorical analysis of neo- and paleo-endemism (CANAPE).
+**C**, relative phylogenetic endemism (RPE).
+**D**, SES of RPE.
+**E**, Categorical analysis of neo- and paleo-endemism (CANAPE).
 
@@
 # Define helper function for making a scatterplot
-scatterplot <- function(data, x_var, y_var, x_lab, y_lab, label, annotate) {
+scatterplot <- function(data, x_var, y_var, x_lab, y_lab, label, annotate, loc) {
   p <- ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
@@
   # Add annotation and trendline if annotate = TRUE
-  if(annotate) p <- p +
-      annotate("label", x = Inf, y = Inf, label = label, hjust = 1, vjust = 1, label.size = 0, fill = "transparent") +
+  if(annotate & loc == "UL") p <- p +
+      annotate("label", x = -Inf, y = Inf, label = label, hjust = 0, vjust = 1, label.size = 0, size = 3, fill = "transparent") +
+      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
+
+  if(annotate & loc == "UR") p <- p +
+      annotate("label", x = Inf, y = Inf, label = label, hjust = 1, vjust = 1, label.size = 0, size = 3, fill = "transparent") +
+      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
+
+  if(annotate & loc == "LL") p <- p +
+      annotate("label", x = -Inf, y = -Inf, label = label, hjust = 0, vjust = 0, label.size = 0, size = 3, fill = "transparent") +
+      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
+
+  if(annotate & loc == "LR") p <- p +
+      annotate("label", x = Inf, y = -Inf, label = label, hjust = 1, vjust = 0, label.size = 0, size = 3, fill = "transparent") +
       geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
@@
 
-```{r raw-biplot, fig.width = 7, fig.height = 6}
+```{r raw-biplot, fig.width = 8, fig.height = 10}
 ## Biplots of raw metrics ##
+
+# Set up x and y variables as factors so they can be ordered appropriately in the plot
+x_vars <-
+  c("mpd_obs", "pd_obs", "pe_obs", "rpe_obs", "mpd_morph_obs", "richness") %>%
+  factor() %>%
+  fct_relevel(as.character(.))
+
+y_vars <- c("mpd_obs", "pd_obs", "pe_obs", "rpe_obs", "mpd_morph_obs") %>%
+  factor() %>%
+  fct_relevel(as.character(.))
 
@@
     list(
-      x_var = c("mpd_obs", "pe_obs", "mpd_morph_obs", "richness"),
-      y_var = c("mpd_obs", "pe_obs", "mpd_morph_obs")
+      x_var = x_vars,
+      y_var = y_vars
     ),
@@
   # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
-  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
-  # Use factors to keep the preferred combination of x and y
-  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs", "mpd_morph_obs", "richness")) %>%
-  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs", "pe_obs", "mpd_morph_obs")) %>%
+  group_by(grp = paste(pmax(as.character(x_var), as.character(y_var)), pmin(as.character(x_var), as.character(y_var)), sep = "_")) %>%
   arrange(x_var, y_var) %>%
@@
   select(-grp) %>%
+  arrange(x_var, y_var) %>%
   # Convert x_var and y_var back to character for plotting function
@@
       x_var == "mpd_obs" ~ "MPD",
+      x_var == "pd_obs" ~ "PD",
+      x_var == "rpe_obs" ~ "RPE",
       x_var == "pe_obs" ~ "PE",
@@
       y_var == "mpd_obs" ~ "MPD",
+      y_var == "pd_obs" ~ "PD",
+      y_var == "rpe_obs" ~ "RPE",
       y_var == "pe_obs" ~ "PE",
-      y_var == "richness" ~ "Richness"
-    )
-  ) %>%
-  # Run modified t-test
-  mutate(
-    test_res = map2(
-      .x = x_var, 
-      .y = y_var, 
-      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
-    # Extract relevant stats from results
-    cor = map_dbl(test_res, pluck("corr")),
-    pval = map_dbl(test_res, pluck("p.value"))
-    ) %>%
-  # Generate label
-  mutate(
-    cor_lab = glue("r = {round(cor, 3)}"),
-    pval_lab = glue("p = {round(pval, 3)}"),
-    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
-  ) %>%
-  # Specify which plots to annotate
-  mutate(
-    annotate = case_when(
-      x_var == "mpd_morph_obs" & y_var == "mpd_obs" & pval < 0.05 ~ TRUE,
-      x_var == "mpd_morph_obs" & y_var == "pe_obs" & pval < 0.05 ~ TRUE,
-      x_var == "mpd_obs" & y_var == "pe_obs" & pval < 0.05 ~ TRUE,
-      x_var == "richness" & y_var == "pe_obs" & pval < 0.05  ~ TRUE,
-      TRUE ~ FALSE
-    )
-  ) %>%
-  select(x_var, y_var, x_lab, y_lab, label, annotate)
-
-# Make plot
-pmap(raw_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
-wrap_plots() +
-  plot_annotation(tag_levels = 'A')
-
-ggsave(result_file("raw-biplot", "pdf"), width = 7, height = 6, units = "in")
-```
-
-**`r s_figure("raw-biplot")`**.
-`r s_figure_cap("raw-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism.
-Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05; correlations between MPD and MMD with richness not expected to be linear, so were not tested.
-
-```{r ses-biplot, fig.width = 7, fig.height = 6}
-## Biplots of SES metrics ##
-
-# Make dataframe of arguments to feed into plotting function
-# - includes all non-duplicated combinations of raw diversity metrics
-ses_comparisons <-
-  cross_df(
-    list(
-      x_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z", "richness"),
-      y_var = c("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")
-    ),
-    .filter = `==` # filter out exact duplicates
-  ) %>%
-  # Remove reversed duplicates
-  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
-  group_by(grp = paste(pmax(x_var, y_var), pmin(x_var, y_var), sep = "_")) %>%
-  # Use factors to keep the preferred combination of x and y
-  mutate(x_var = as.factor(x_var) %>% fct_relevel("pe_obs_z", "mpd_morph_obs_z", "richness")) %>%
-  mutate(y_var = as.factor(y_var) %>% fct_relevel("mpd_obs_z", "pe_obs_z", "mpd_morph_obs_z")) %>%
-  arrange(x_var, y_var) %>%
-  slice(1) %>%
-  ungroup() %>%
-  select(-grp) %>%
-  # Convert x_var and y_var back to character for plotting function
-  mutate(across(c(x_var, y_var), as.character)) %>%
-  # Add labels
-  mutate(
-    x_lab = case_when(
-      x_var == "mpd_morph_obs_z" ~ "SES MMD",
-      x_var == "mpd_obs_z" ~ "SES MPD",
-      x_var == "pe_obs_z" ~ "SES PE",
-      x_var == "richness" ~ "Richness"
-    ),
-    y_lab = case_when(
-      y_var == "mpd_morph_obs_z" ~ "SES MMD",
-      y_var == "mpd_obs_z" ~ "SES MPD",
-      y_var == "pe_obs_z" ~ "SES PE",
       y_var == "richness" ~ "Richness"
@@
       TRUE ~ FALSE
+    ),
+    # annotate in upper-right corner for mpd_morph_obs + rpe_obs,
+    # upper-left otherwise
+    loc = case_when(
+      y_var == "mpd_morph_obs" & x_var == "rpe_obs" ~ "UR",
+      TRUE ~ "UL"
     )
   ) %>%
-  select(x_var, y_var, x_lab, y_lab, label, annotate)
+  select(x_var, y_var, x_lab, y_lab, label, annotate, loc)
+
+# Make plot
+pmap(raw_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
+wrap_plots() +
+  plot_layout(ncol = 3) +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(result_file(s_figure("raw-biplot"), "pdf"), width = 8, height = 10, units = "in")
+```
+
+**`r s_figure("raw-biplot")`**.
+`r s_figure_cap("raw-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PD, phylogenetic diversity; PE, phylogenetic endemism; RPE, relative phylogenetic endemism.
+Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05.
+
+```{r ses-biplot, fig.width = 8, fig.height = 10}
+## Biplots of SES metrics ##
+
+# Set up x and y variables as factors so they can be ordered appropriately in the plot
+x_vars <-
+  c("mpd_obs_z", "pd_obs_z", "pe_obs_z", "rpe_obs_z", "mpd_morph_obs_z", "richness") %>%
+  factor() %>%
+  fct_relevel(as.character(.))
+
+y_vars <- c("mpd_obs_z", "pd_obs_z", "pe_obs_z", "rpe_obs_z", "mpd_morph_obs_z") %>%
+  factor() %>%
+  fct_relevel(as.character(.))
+
+# Make dataframe of arguments to feed into plotting function
+# - includes all non-duplicated combinations of raw diversity metrics
+ses_comparisons <-
+  cross_df(
+    list(
+      x_var = x_vars,
+      y_var = y_vars
+    ),
+    .filter = `==` # filter out exact duplicates
+  ) %>%
+  # Remove reversed duplicates
+  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
+  group_by(grp = paste(pmax(as.character(x_var), as.character(y_var)), pmin(as.character(x_var), as.character(y_var)), sep = "_")) %>%
+  arrange(x_var, y_var) %>%
+  slice(1) %>%
+  ungroup() %>%
+  select(-grp) %>%
+  arrange(x_var, y_var) %>%
+  # Convert x_var and y_var back to character for plotting function
+  mutate(across(c(x_var, y_var), as.character)) %>%
+  # Add labels
+  mutate(
+    x_lab = case_when(
+      x_var == "mpd_morph_obs_z" ~ "SES MMD",
+      x_var == "mpd_obs_z" ~ "SES MPD",
+      x_var == "pd_obs_z" ~ "SES PD",
+      x_var == "pe_obs_z" ~ "SES PE",
+      x_var == "rpe_obs_z" ~ "SES RPE",
+      x_var == "richness" ~ "Richness"
+    ),
+    y_lab = case_when(
+      y_var == "mpd_morph_obs_z" ~ "SES MMD",
+      y_var == "mpd_obs_z" ~ "SES MPD",
+      y_var == "pd_obs_z" ~ "SES PD",
+      y_var == "pe_obs_z" ~ "SES PE",
+      y_var == "rpe_obs_z" ~ "SES RPE",
+      y_var == "richness" ~ "Richness"
+    )
+  ) %>%
+  # Run modified t-test
+  mutate(
+    test_res = map2(
+      .x = x_var, 
+      .y = y_var, 
+      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
+    # Extract relevant stats from results
+    cor = map_dbl(test_res, pluck("corr")),
+    pval = map_dbl(test_res, pluck("p.value"))
+    ) %>%
+  # Generate label
+  mutate(
+    cor_lab = glue("r = {round(cor, 3)}"),
+    pval_lab = glue("p = {round(pval, 3)}"),
+    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
+  ) %>%
+  # Specify which plots to annotate
+  mutate(
+    annotate = case_when(
+      pval < 0.05 ~ TRUE,
+      TRUE ~ FALSE
+    ),
+    # annotate in upper-right corner for mpd_morph_obs + rpe_obs,
+    # upper-left otherwise
+    loc = case_when(
+      y_var == "mpd_morph_obs_z" & x_var == "pd_obs_z" ~ "LR",
+      y_var == "rpe_obs_z" & x_var == "pe_obs_z" ~ "LR",
+      y_var == "mpd_morph_obs_z" & x_var == "pe_obs_z" ~ "LR",
+      y_var == "pd_obs_z" & x_var == "richness" ~ "UR",
+      TRUE ~ "UL"
+    )
+  ) %>%
+  select(x_var, y_var, x_lab, y_lab, label, annotate, loc)
 
@@
 wrap_plots() +
+  plot_layout(ncol = 3) +
   plot_annotation(tag_levels = 'A')
 
-ggsave(result_file("ses-biplot", "pdf"), width = 7, height = 6, units = "in")
+ggsave(result_file(s_figure("ses-biplot"), "pdf"), width = 8, height = 10, units = "in")
 ```
@@
 **`r s_figure("ses-biplot")`**.
-`r s_figure_cap("ses-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PE, phylogenetic endemism.
-Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; all combinations tested, but *p* value and correlation (r) only shown for *p* < 0.05.
+`r s_figure_cap("ses-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PD, phylogenetic diversity; PE, phylogenetic endemism; RPE, relative phylogenetic endemism.
+Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05.
 
@@
 
-ggsave(result_file("traits-nmds", "pdf"), width = 8, height = 8, units = "in")
+ggsave(result_file(s_figure("traits-nmds"), "pdf"), width = 8, height = 8, units = "in")
 ```

### 2020-07-02 ms/manuscript.Rmd
Add PD and RPE

@@
 
-Joel H. Nitta^1^ and Atsushi Ebihara^2^
+Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1^, and Atsushi Ebihara^3^
 
@@
 
-^2^Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba-shi, Ibaraki, 305-0005, Japan
+^2^University and Jepson Herbaria, University of California, Berkeley, CA 94720‐2465, USA
+
+^3^Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba-shi, Ibaraki, 305-0005, Japan
 
@@
 
-Raw values of MPD and MMD showed high variance and low species richness, then merged towards a single value at high species richness, as expected (`r s_figure("raw-biplot")`D, E).
-Raw PE was linearly correlated with species richness (`r s_figure("raw-biplot")`F).
-Both raw and SES values of MPD and MMD were highly correlated (`r s_figure("raw-biplot")`A, `r s_figure("ses-biplot")`A, B).
-SES values of MPD, MMD, and PE were all positively correlated (`r s_figure("ses-biplot")`A, B, C).
-None of the SES values were correlated with richness (`r s_figure("ses-biplot")`D, E, F).
+Raw values of MPD and MMD showed high variance and low species richness, then merged towards a single value at high species richness, as expected (`r s_figure("raw-biplot")`K, O).
+Raw PE was linearly correlated with species richness (`r s_figure("raw-biplot")`M).
+Both raw and SES values of MPD and MMD were highly correlated (`r s_figure("raw-biplot")`D, `r s_figure("ses-biplot")`D).
+SES values of MPD, MMD, and PE were all positively correlated (`r s_figure("ses-biplot")`).
+None of the SES values were correlated with richness (`r s_figure("ses-biplot")`).
 
@@
   scale_fill_viridis_c() +
-  map_theme
+  map_theme +
+  theme(
+    axis.text.x = element_blank()
+  )
 
@@
   theme(
+    axis.text.x = element_blank(),
     axis.text.y = element_blank()
@@
 
-a + b + c + d + e + f +
+g <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "MMD") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  scale_fill_viridis_c() +
+  map_theme
+
+h <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES PD") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.y = element_blank()
+  )
+
+a + b + c + d + e + f + g + h +
   plot_layout(ncol = 2) +
@@
 
-ggsave(here::here("results/Fig_1.pdf"), width = 7, height = 9, units = "in")
+ggsave(result_file(figure("diversity"), "pdf"), width = 7, height = 9, units = "in")
 ```
@@
 **E**, mean morphological distance (MMD) **F**, SES of MMD.
+**G**, phylogenetic diversity (PD) **H**, SES of PD.
 
@@
 
-```{r endemism, fig.width = 7, fig.height = 6}
+```{r endemism, fig.width = 7, fig.height = 8}
 # PE
@@
 
+# RPE
+c <- ggplot(ses_div_ferns_spatial, aes(fill = rpe_obs)) +
+  geom_sf(color = "transparent") +
+  scale_fill_viridis_c() + 
+  labs(fill = "RPE") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.x = element_blank()
+  )
+
+# SES RPE
+d <- ggplot(ses_div_ferns_spatial, aes(fill = rpe_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES RPE") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, rpe_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, rpe_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme +
+  theme(
+    axis.text.y = element_blank()
+  )
+
 # CANAPE
-c <- ggplot(ses_div_ferns_spatial, aes(fill = endem_type)) +
+e <- ggplot(ses_div_ferns_spatial, aes(fill = endem_type)) +
   geom_sf(color = "transparent") +
@@
 
-a + b + c +
+a + b + c + d + e +
   plot_layout(ncol = 2) +
@@
 
-ggsave(here::here("results/Fig_2.pdf"), width = 7, height = 6, units = "in")
+ggsave(result_file(figure("endemism"), "pdf"), width = 7, height = 8, units = "in")
 ```
 
-**`r figure("endemism")`**. `r figure_cap("endemism")` **A**, phylogenetic endemism (PE). **B**, Standard effect size (SES) of PE. **C**, Categorical analysis of neo- and paleo-endemism (CANAPE).
+**`r figure("endemism")`**.
+`r figure_cap("endemism")` **A**, phylogenetic endemism (PE).
+**B**, Standard effect size (SES) of PE.
+**C**, relative phylogenetic endemism (RPE).
+**D**, SES of RPE.
+**E**, Categorical analysis of neo- and paleo-endemism (CANAPE).
 
@@
 
-ggsave(here::here("results/Fig_3.pdf"), width = 7, height = 3, units = "in")
+ggsave(result_file(figure("bioregions"), "pdf"), width = 7, height = 3, units = "in")
 ```

### 2020-07-02 R/functions.R

@@ -3430,16 +3430,16 @@ remove_dup_seqs <- function (
 #' Only works for figures or tables cited in text, and outputs
 #' files to the "results" folder
 #'
-#' @param name Name of cited figure or table, as defined in `captions.R`
+#' @param result_num Number of result, e.g. "Fig. 2"
 #' @param extension Extension to use for file
 #'
 #' @return String.
 #' 
-result_file <- function (name, extension) {
+result_file <- function (result_num, extension) {
   
   fs::path(
     here::here("results"),
-    s_figure(name) %>% 
+    result_num %>%
       str_remove_all("\\.") %>% 
       str_replace_all(" ", "_")
   ) %>%

### 2020-07-04 ms/manuscript.Rmd
Update MS
add optional highlighting

@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  # highlight only significant cells
+  # gghighlight(mpd_morph_obs_p > 0.975 | mpd_morph_obs_p < 0.025) +
   map_theme +
@@
 
+# PD
 g <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
@@
 
+# SES of PD
 h <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  # highlight only significant cells
+  # gghighlight(pd_obs_p > 0.975 | pd_obs_p < 0.025) +
   map_theme +
@@
   )
+
+# To add: RPD
+# ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
+#   geom_sf(color = "transparent") +
+#   labs(fill = "SES RPD") +
+#   scale_fill_scico(
+#     palette = "romaO",
+#     na.value="grey80",
+#     direction = -1,
+#     limits = c(
+#       -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
+#       get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
+#     )) +
+#   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+#   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+#   # highlight only significant cells
+#   # gghighlight(rpd_obs_p > 0.95 | rpd_obs_p < 0.05) +
+#   map_theme +
+#   theme(
+#     axis.text.y = element_blank()
+#   )
 
@@
 # SES PE
-b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_z)) +
+b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_p)) +
   geom_sf(color = "transparent") +
   labs(fill = "SES PE") +
-  scale_fill_scico(
-    palette = "romaO",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_spatial, pe_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, pe_obs_z, "abs")
-    )) +
+  scale_fill_viridis_c() +
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +

### 2020-07-04 ms/SI.Rmd
Change metrics to all richness-based measures
no MPD or MNTD

@@
 map_theme <- theme(
-    panel.grid.major = element_line(color = "white", size = 0.2),
+    panel.grid.major = element_line(color = "grey95", size = 0.1),
+    panel.background = element_rect(fill = "grey60"),
     axis.ticks = element_blank(),
@@
 
+map_theme_light <- theme(
+    panel.grid.major = element_line(color = "white", size = 0.1),
+    panel.background = element_rect(fill = "grey90"),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
+  )
 ```
@@
 # PE (endemic)
+# FIXME: cells with zero species should be NA, not zero PE
 a <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
   geom_sf(color = "transparent") +
-  scale_fill_viridis_c() + 
+  scale_fill_scico(palette = "lajolla") + 
   labs(fill = "PE") +
@@
 # SES PE (endemic)
-b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_z)) +
+# FIXME: cells with zero species should be NA, not zero PE
+b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_rank)) +
   geom_sf(color = "transparent") +
-  labs(fill = "SES PE") +
-  scale_fill_scico(
-    palette = "romaO",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs"),
-      get_limit(ses_div_ferns_endemic_spatial, pe_obs_z, "abs")
-    )) +
+  labs(fill = "Rank PE") +
+  scale_fill_scico(palette = "lajolla") +
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
@@
   theme(
-    axis.text.y = element_blank(),
-    axis.text.x = element_blank()
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank()
   )
@@
   geom_sf(color = "transparent") +
-  scale_fill_viridis_c() + 
+  scale_fill_scico(palette = "lajolla") + 
   labs(fill = "RPE") +
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme +
+  map_theme_light +
   theme(
@@
   scale_fill_scico(
-    palette = "romaO",
+    palette = "roma",
     na.value="grey80",
@@
 
-# CANAPE (endemic)
-e <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = endem_type)) +
+# CANAPE, restricted dataset
+# set color scheme
+canape_cols <- c(
+  "neo" = "#E41A1C", # Color brewer set1 red
+  "paleo" = "#377EB8", # Color brewer blue
+  "mixed" = "#984EA3", # Color brewer purple
+  "none" = "#FFFFCC" # Scico lajolla faintest color
+)
+
+e <- ses_div_ferns_endemic_spatial %>%
+  mutate(endem_type = replace_na(endem_type, "none")) %>%
+  ggplot(aes(fill = endem_type)) +
   geom_sf(color = "transparent") +
   labs(fill = "Endemism\ntype") +
+  scale_fill_manual(values = canape_cols, breaks = c("neo", "paleo", "mixed", "none"), labels = c("Neo", "Paleo", "Mixed", "Not endemic")) + 
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme
+  map_theme_light
 
@@
 **E**, Categorical analysis of neo- and paleo-endemism (CANAPE).
+
+\clearpage
 
@@
 
-```{r raw-biplot, fig.width = 8, fig.height = 10}
+```{r raw-biplot, fig.width = 8, fig.height = 10.5}
 ## Biplots of raw metrics ##
@@
 x_vars <-
-  c("mpd_obs", "pd_obs", "pe_obs", "rpe_obs", "mpd_morph_obs", "richness") %>%
+  c("pd_obs", "rpd_obs", "fd_obs", "rfd_obs", "pe_obs", "rpe_obs", "richness") %>%
   factor() %>%
@@
 
-y_vars <- c("mpd_obs", "pd_obs", "pe_obs", "rpe_obs", "mpd_morph_obs") %>%
+y_vars <- c("pd_obs", "rpd_obs", "fd_obs", "rfd_obs", "pe_obs", "rpe_obs") %>%
   factor() %>%
@@
     x_lab = case_when(
-      x_var == "mpd_morph_obs" ~ "MMD",
-      x_var == "mpd_obs" ~ "MPD",
       x_var == "pd_obs" ~ "PD",
+      x_var == "rpd_obs" ~ "RPD",
+      x_var == "fd_obs" ~ "FD",
+      x_var == "rfd_obs" ~ "RFD",
+      x_var == "pe_obs" ~ "PE",
       x_var == "rpe_obs" ~ "RPE",
-      x_var == "pe_obs" ~ "PE",
       x_var == "richness" ~ "Richness"
@@
     y_lab = case_when(
-      y_var == "mpd_morph_obs" ~ "MMD",
-      y_var == "mpd_obs" ~ "MPD",
       y_var == "pd_obs" ~ "PD",
-      y_var == "rpe_obs" ~ "RPE",
+      y_var == "rpd_obs" ~ "RPD",
+      y_var == "fd_obs" ~ "FD",
+      y_var == "rfd_obs" ~ "RFD",
       y_var == "pe_obs" ~ "PE",
-      y_var == "richness" ~ "Richness"
+      y_var == "rpe_obs" ~ "RPE"
     )
@@
     cor_lab = glue("r = {round(cor, 3)}"),
-    pval_lab = glue("p = {round(pval, 3)}"),
-    label = glue::glue("{cor_lab}\n{ifelse(pval < 0.001, 'p < 0.001', pval_lab)}")
+    pval_lab = case_when(
+      pval < 0.001 ~ "***",
+      pval < 0.01 ~ "**",
+      pval < 0.05 ~ "*",
+      TRUE ~ ""
+    ),
+    label = glue::glue("{cor_lab}{pval_lab}")
   ) %>%
@@
 wrap_plots() +
-  plot_layout(ncol = 3) +
-  plot_annotation(tag_levels = 'A')
+  plot_layout(ncol = 3)
 
-ggsave(result_file(s_figure("raw-biplot"), "pdf"), width = 8, height = 10, units = "in")
+ggsave(result_file(s_figure("raw-biplot"), "pdf"), width = 8, height = 10.5, units = "in")
 ```
@@
 
-```{r ses-biplot, fig.width = 8, fig.height = 10}
+\clearpage
+
+```{r ses-biplot, fig.width = 8, fig.height = 10.5}
 ## Biplots of SES metrics ##
@@
 x_vars <-
-  c("mpd_obs_z", "pd_obs_z", "pe_obs_z", "rpe_obs_z", "mpd_morph_obs_z", "richness") %>%
+  c("pd_obs_z", "rpd_obs_z", "fd_obs_z", "rfd_obs_z", "pe_obs_z", "rpe_obs_z", "richness") %>%
   factor() %>%
@@
 
-y_vars <- c("mpd_obs_z", "pd_obs_z", "pe_obs_z", "rpe_obs_z", "mpd_morph_obs_z") %>%
+y_vars <- c("pd_obs_z", "rpd_obs_z", "fd_obs_z", "rfd_obs_z", "pe_obs_z", "rpe_obs_z") %>%
   factor() %>%
@@
     x_lab = case_when(
-      x_var == "mpd_morph_obs_z" ~ "SES MMD",
-      x_var == "mpd_obs_z" ~ "SES MPD",
       x_var == "pd_obs_z" ~ "SES PD",
+      x_var == "rpd_obs_z" ~ "SES RPD",
+      x_var == "fd_obs_z" ~ "SES FD",
+      x_var == "rfd_obs_z" ~ "SES RFD",
       x_var == "pe_obs_z" ~ "SES PE",
@@
     y_lab = case_when(
-      y_var == "mpd_morph_obs_z" ~ "SES MMD",
-      y_var == "mpd_obs_z" ~ "SES MPD",
       y_var == "pd_obs_z" ~ "SES PD",
+      y_var == "rpd_obs_z" ~ "SES RPD",
+      y_var == "fd_obs_z" ~ "SES FD",
+      y_var == "rfd_obs_z" ~ "SES RFD",
       y_var == "pe_obs_z" ~ "SES PE",
-      y_var == "rpe_obs_z" ~ "SES RPE",
-      y_var == "richness" ~ "Richness"
+      y_var == "rpe_obs_z" ~ "SES RPE"
     )
@@
     loc = case_when(
-      y_var == "mpd_morph_obs_z" & x_var == "pd_obs_z" ~ "LR",
       y_var == "rpe_obs_z" & x_var == "pe_obs_z" ~ "LR",
-      y_var == "mpd_morph_obs_z" & x_var == "pe_obs_z" ~ "LR",
+      y_var == "fd_obs_z" & x_var == "richness" ~ "UR",
       y_var == "pd_obs_z" & x_var == "richness" ~ "UR",
@@
 wrap_plots() +
-  plot_layout(ncol = 3) +
-  plot_annotation(tag_levels = 'A')
+  plot_layout(ncol = 3)
 
-ggsave(result_file(s_figure("ses-biplot"), "pdf"), width = 8, height = 10, units = "in")
+ggsave(result_file(s_figure("ses-biplot"), "pdf"), width = 8, height = 10.5, units = "in")
 ```
@@
 Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05.
+
+\clearpage
 

### 2020-07-04 ms/manuscript.Rmd
Change metrics to all richness-based measures
no MPD or MNTD

--- 2020-07-02 ms/manuscript.Rmd
@@
 map_theme <- theme(
-    panel.grid.major = element_line(color = "white", size = 0.2),
+    panel.grid.major = element_line(color = "grey95", size = 0.1),
+    panel.background = element_rect(fill = "grey60"),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
+  )
+
+map_theme_light <- theme(
+    panel.grid.major = element_line(color = "white", size = 0.1),
+    panel.background = element_rect(fill = "grey90"),
     axis.ticks = element_blank(),
@@
 
-```{r diversity, fig.width = 7, fig.height = 9}
+```{r richness, width = 7, height = 4}
 # Richness
@@
   geom_sf(color = "transparent") +
-  scale_fill_viridis_c() +
-  labs(fill = "Richness \n (n spp.)") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(fill = "Richness") +
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme
+
+# Sampling rendundancy
+b <- ggplot(ses_div_ferns_spatial, aes(fill = redundancy)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  labs(fill = "Redundancy") +
+  map_theme +
+  theme(
+    axis.text.y = element_blank()
+  )
+
+a + b + 
+  plot_layout(ncol = 2) +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(result_file(figure("richness"), "pdf"), width = 7, height = 4, units = "in")
+```
+
+**`r figure("richness")`**. `r figure_cap("diversity")` **A**, richness. **B**, sampling redundancy.
+
+```{r diversity, fig.width = 7, fig.height = 9}
+# PD
+a <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "PD") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  scale_fill_scico(palette = "lajolla") + 
   map_theme +
@@
 
-# Sampling rendundancy
-b <- ggplot(ses_div_ferns_spatial, aes(fill = redundancy)) +
+# SES of PD
+b <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
   geom_sf(color = "transparent") +
-  scale_fill_viridis_c() +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  labs(fill = "Redundancy") +
-  map_theme +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank()
-  )
-
-# MPD
-c <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_obs)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "MPD") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  scale_fill_viridis_c() +
-  map_theme +
-  theme(
-    axis.text.x = element_blank()
-  )
-
-# SES MPD
-d <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_obs_z)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "SES MPD") +
+  labs(fill = "SES PD") +
   scale_fill_scico(
@@
 
-# Raw morph diversity
-e <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_morph_obs)) +
+# RPD
+c <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs)) +
   geom_sf(color = "transparent") +
-  labs(fill = "MMD") +
+  labs(fill = "RPD") +
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  scale_fill_viridis_c() +
+  scale_fill_scico(palette = "lajolla") + 
   map_theme +
@@
 
-# SES of morph diversity
-f <- ggplot(ses_div_ferns_spatial, aes(fill = mpd_morph_obs_z)) +
+# SES of RPD
+d <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
   geom_sf(color = "transparent") +
-  labs(fill = "SES MMD") +
+  labs(fill = "SES RPD") +
   scale_fill_scico(
-    palette = "romaO",
+    palette = "roma",
     na.value="grey80",
@@
     limits = c(
-      -get_limit(ses_div_ferns_spatial, mpd_morph_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, mpd_morph_obs_z, "abs")
+      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
     )) +
@@
 
-g <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
+# FD
+e <- ggplot(ses_div_ferns_spatial, aes(fill = fd_obs)) +
   geom_sf(color = "transparent") +
-  labs(fill = "MMD") +
+  labs(fill = "FD") +
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  scale_fill_viridis_c() +
-  map_theme
+  scale_fill_scico(palette = "lajolla") + 
+  map_theme +
+  theme(
+    axis.text.x = element_blank()
+  )
 
-h <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
+# SES of FD
+f <- ggplot(ses_div_ferns_spatial, aes(fill = fd_obs_z)) +
   geom_sf(color = "transparent") +
-  labs(fill = "SES PD") +
+  labs(fill = "SES FD") +
   scale_fill_scico(
-    palette = "romaO",
+    palette = "roma",
     na.value="grey80",
@@
     limits = c(
-      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
+      -get_limit(ses_div_ferns_spatial, fd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, fd_obs_z, "abs")
     )) +
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  # highlight only significant cells
+  # gghighlight(mpd_morph_obs_p > 0.975 | mpd_morph_obs_p < 0.025) +
+  map_theme +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank()
+  )
+
+# RFD
+g <- ggplot(ses_div_ferns_spatial, aes(fill = rfd_obs)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "RFD") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  scale_fill_scico(palette = "lajolla") + 
+  map_theme
+
+# SES of RFD
+h <- ggplot(ses_div_ferns_spatial, aes(fill = rfd_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES RFD") +
+  scale_fill_scico(
+    palette = "roma",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  # highlight only significant cells
+  # gghighlight(rfd_obs_p > 0.975 | rfd_obs_p < 0.025) +
   map_theme +
@@
 **`r figure("diversity")`**.
-`r figure_cap("diversity")` **A**, richness.
-**B**, sampling redundancy.
-**C**, mean phylogenetic distance (MPD).
-**D**, Standard effect size (SES) of MPD.
-**E**, mean morphological distance (MMD) **F**, SES of MMD.
-**G**, phylogenetic diversity (PD) **H**, SES of PD.
+`r figure_cap("diversity")` **A**, Phylogenetic diversity (PD).
+**B**, Standard effect size (SES) of PD.
+**C**, Relative phylogenetic diversity (RPD).
+**D**, SES of RPD.
+**E**, Functional diversity (FD).
+**F**, SES of FD **G**, Relative functional diversity (RFD) **H**, SES of RFD.
 
@@
   geom_sf(color = "transparent") +
-  scale_fill_viridis_c() + 
+  scale_fill_scico(palette = "lajolla") + 
   labs(fill = "PE") +
@@
 
-# SES PE
-b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_z)) +
+# PE rank
+b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_rank)) +
   geom_sf(color = "transparent") +
-  labs(fill = "SES PE") +
-  scale_fill_scico(
-    palette = "romaO",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_spatial, pe_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, pe_obs_z, "abs")
-    )) +
+  labs(fill = "Rank PE") +
+  scale_fill_scico(palette = "lajolla") +
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
@@
   geom_sf(color = "transparent") +
-  scale_fill_viridis_c() + 
+  scale_fill_scico(palette = "lajolla") +
   labs(fill = "RPE") +
@@
 
-# CANAPE
-e <- ggplot(ses_div_ferns_spatial, aes(fill = endem_type)) +
+# CANAPE, full dataset
+# set color scheme
+canape_cols <- c(
+  "neo" = "#E41A1C", # Color brewer set1 red
+  "paleo" = "#377EB8", # Color brewer blue
+  "mixed" = "#984EA3", # Color brewer purple
+  "none" = "#FFFFCC" # Scico lajolla faintest color
+)
+
+e <- 
+  ses_div_ferns_spatial %>%
+  mutate(endem_type = replace_na(endem_type, "none")) %>%
+  ggplot(aes(fill = endem_type)) +
   geom_sf(color = "transparent") +
   labs(fill = "Endemism\ntype") +
+  scale_fill_manual(values = canape_cols, breaks = c("neo", "paleo", "mixed", "none"), labels = c("Neo", "Paleo", "Mixed", "Not endemic")) + 
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme
+  map_theme_light
 
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme
+  map_theme_light
 
@@
 
-```{r diversity-by-cluster}
+```{r diversity-by-cluster, fig.width = 7, fig.height = 7}
+data <-
 ses_div_ferns_spatial %>%
@@
   as_tibble %>%
-  select(mpd_obs_z, mpd_morph_obs_z, pe_obs_z, taxonomic_cluster) %>%
-  pivot_longer(names_to = "var", values_to = "value", -taxonomic_cluster) %>%
-  ggplot(aes(x = taxonomic_cluster, y = value)) +
-  geom_jitter() +
-  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.5) +
-  facet_wrap(vars(var)) +
+  select(pd_obs_z, rpd_obs_z, fd_obs_z, rfd_obs_z, pe_obs_rank, taxonomic_cluster)
+
+cluster_cols <- carto_pal(8, "Bold")[1:4] %>%
+  set_names(1:4)
+
+a <- ggplot(data, aes(x = taxonomic_cluster, y = pd_obs_z)) +
+  geom_jitter(aes(color = taxonomic_cluster)) +
+  scale_color_manual(values = cluster_cols, breaks = 1:4) +
+  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "SES PD",
+    x = "Cluster",
+    color = "Cluster"
+  ) +
+  theme_bw() +
   theme(
@@
     panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank()
   )
+
+b <- ggplot(data, aes(x = taxonomic_cluster, y = rpd_obs_z)) +
+  geom_jitter(aes(color = taxonomic_cluster)) +
+  scale_color_manual(values = cluster_cols, breaks = 1:4) +
+  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "SES RPD",
+    x = "Cluster",
+    color = "Cluster"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+
+c <- ggplot(data, aes(x = taxonomic_cluster, y = fd_obs_z)) +
+  geom_jitter(aes(color = taxonomic_cluster)) +
+  scale_color_manual(values = cluster_cols, breaks = 1:4) +
+  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "SES FD",
+    x = "Cluster",
+    color = "Cluster"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+
+d <- ggplot(data, aes(x = taxonomic_cluster, y = rfd_obs_z)) +
+  geom_jitter(aes(color = taxonomic_cluster)) +
+  scale_color_manual(values = cluster_cols, breaks = 1:4) +
+  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "SES RFD",
+    x = "Cluster",
+    color = "Cluster"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.title.x = element_blank()
+  )
+
+e <- ggplot(data, aes(x = taxonomic_cluster, y = pe_obs_rank)) +
+  geom_jitter(aes(color = taxonomic_cluster)) +
+  scale_color_manual(values = cluster_cols, breaks = 1:4) +
+  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "Rank PE",
+    x = "Cluster",
+    color = "Cluster"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+
+  )
+
+a + b + c + d + e +
+  plot_layout(ncol = 2) +
+plot_annotation(
+  tag_levels = 'A'
+) &
+  theme(legend.position = "none")
+
+ggsave(result_file(figure("div-by-tax-region"), "pdf"), width = 7, height = 7, units = "in")
+
 ```
 
-Standard effect size (SES) of mean phylogenetic distance (MPD), mean morphological distance (MMD), and phylogenetic endemism (PE) by taxonomic bioregion (top four regions with the most sites only shown).
+**`r figure("div-by-tax-region")`**.
+`r figure_cap("div-by-tax-region")`.
+**A**, Standard effect size (SES) of phylogenetic diversity (PD).
+**B**, SES of relative PD (RPD).
+**C**, SES of functional diversity (FD).
+**D**, SES of relative FD (RFD).
+**E**, rank of observed phylogenetic endemism (PE) out of 999 random communities.
+Only top four clusters with the most sites shown.

### 2020-07-04 R/packages.R

@@ -23,6 +23,7 @@ library(magrittr)
 library(glue)
 library(patchwork)
 library(rcartocolor)
+library(ggtext)
 library(tidyverse)
 
 # Resolve conflicts

### 2020-07-04 R/plan.R

@@ -112,22 +112,27 @@ plan <- drake_plan (
     repro_data = repro_data),
 
   # Analyze standard effect size (SES) of diversity metrics ----
-
+  
   ses_phy = target(
-    run_ses_analysis(comm, japan_fern_tree, n_reps = 999, metrics) %>% categorize_endemism,
+    run_ses_analysis(
+      comm = comm, 
+      phy = japan_fern_tree, 
+      n_reps = 999, 
+      metrics = metrics) %>% 
+      categorize_endemism,
     transform = map(
       comm = c(comm_ferns, comm_ferns_endemic),
       metrics = c(
-        c("mpd", "pd", "rpd", "pe", "rpe"),
-        c("pd", "pe", "rpe")
+        c("pd", "rpd", "pe", "rpe"),
+        c("pe", "rpe")
       ),
       .names = c("ses_phy_ferns", "ses_phy_ferns_endemic"))
   ),
 
   ses_traits_ferns = run_ses_analysis(
-    comm_ferns,
+    comm = comm_ferns,
     n_reps = 999,
-    metrics = "mpd_morph",
+    metrics = c("fd", "rfd"),
     trait_distances = trait_distance_matrix),
 
   # Analyze phyloregions
@@ -179,11 +184,11 @@ plan <- drake_plan (
     knitr_in("ms/manuscript.Rmd"),
     output_dir = here::here("results"),
     quiet = TRUE),
-  
+   
   si = rmarkdown::render(
     knitr_in("ms/SI.Rmd"),
     output_dir = here::here("results"),
-    quiet = TRUE),
+    quiet = TRUE)
 
   # # Plot and write out traits dendrogram
   # # traits_dendrogram = make_traits_dendrogram(trait_distance_matrix, taxon_id_map),

### 2020-07-04 ms/captions.R

@@ -2,7 +2,8 @@
 
 # - Figures
 fig_nums <- captioner::captioner(prefix = "Fig.")
-fig_nums(name = "diversity", caption = "Taxonomic, phylogenetic, and morphological diversity of the ferns of Japan.")
+fig_nums(name = "richness", caption = "Taxonomic diversity of the ferns of Japan.")
+fig_nums(name = "diversity", caption = "Phylogenetic and morphological diversity of the ferns of Japan.")
 fig_nums(name = "endemism", caption = "Phylogenetic endemism of the ferns of Japan.")
 fig_nums(name = "bioregions", caption = "Bioregions of the ferns of Japan.")
 fig_nums(name = "ecostructure", caption = "Ecological structure of the ferns of Japan.")

### 2020-07-06 ms/SI.Rmd
Analyze redundancy by grain size with custom function

@@
 loadd(list = c(
+  "richness_by_res",
   "occ_point_data_ferns",
@@
 ```
+
+```{r grain-size, fig.width = 7, fig.height = 3}
+# Helper function to format resolution labels
+degree_lab <- function (x) {
+  as.numeric(x) %>% 
+    scales::number(suffix = "\u00b0", accuracy = 0.1)
+}
+
+# Summarize redundancy statistics by grain size
+redundancy_summary <-
+  richness_by_res %>%
+  mutate(above_50 = ifelse(redundancy > 0.5, 1, 0)) %>%
+  mutate(singleton = ifelse(redundancy == 0, 1, 0)) %>%
+  group_by(res) %>%
+  summarize(
+    n_total = n(),
+    n_above_50 = sum(above_50),
+    n_singletons = sum(singleton),
+    percent_above_50 = n_above_50/n_total,
+    percent_singletons = n_singletons/ n_total,
+    .groups = "drop"
+  )
+
+# Percent of grid cells with redundancy > 0.5 by grain size
+a <- ggplot(redundancy_summary, aes(x = res, y = percent_above_50)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::percent) +
+  scale_x_discrete(labels = degree_lab) +
+  labs(
+    x = "Resolution",
+    y = "Redundancy > 0.5"
+  )
+
+# Percent of grid cells with redundancy = 0 by grain size
+b <- ggplot(redundancy_summary, aes(x = res, y = percent_singletons)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::percent) +
+  scale_x_discrete(labels = degree_lab) +
+  labs(
+    x = "Resolution",
+    y = "Redundancy = 0"
+  )
+
+# Distribution of redundancy by grain size
+c <- ggplot(redund_by_res, aes(x = redundancy, y = res)) +
+  geom_density_ridges(scale = 1.5) + 
+  scale_y_discrete(expand = c(0, 0), labels = degree_lab) +     # will generally have to set the `expand` option
+  scale_x_continuous(expand = c(0, 0)) +   # for both axes to remove unneeded padding
+  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
+  labs(
+    x = "Redundancy",
+    y = "Resolution"
+  )
+
+a + b + c +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(result_file(s_figure("grain-size"), "pdf"), width = 7, height = 3, units = "in")
+```
+
+**`r s_figure("grain-size")`**.
+`r s_figure_cap("grain-size")` **A**, Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
+**B**, Percent of grid cells with redundancy = 0 by grain size.
+**C**, Distribution of redundancy values by grain size.
+
+\clearpage
 

### 2020-07-06 ms/captions.R

@@ -14,6 +14,7 @@ table_nums(name = "traits-used", caption = "Fern traits used in this study")
 
 # - SI figures
 s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
+s_fig_nums(name = "grain-size", "Analysis of different grain sizes on sampling redundancy.")
 s_fig_nums(name = "k-plot", "Selection of *k* for bioregions.")
 s_fig_nums(name = "k-plot-ecos", "Selection of *k* for ecological structure analysis.")
 s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan, restricted dataset including only taxa endemic to Japan.")

### 2020-07-06 ms/manuscript.Rmd
Update MS

@@
 
-Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology for two reasons: to understand the processes generating it, and to conserve it.
+Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it, and to conserve it.
 The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness.
-For example, hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots and motivating conservation strategies to preserve them.
+Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots and motivating conservation strategies to preserve them.
 However, richness alone cannot provide a complete picture of biodiversity.

### 2020-07-06 ms/SI.Rmd
Use custom function for generating comm matrix at different grain sizes
uses less memory than phyloregion::points2comm

@@
 loadd(list = c(
+  "comm_scaled_list_0.1",
+  "comm_scaled_list_0.2",
+  "comm_scaled_list_0.3",
+  "comm_scaled_list_0.4",
   "richness_by_res",
@@
 
+# Combine results of building community matrices at different grain sizes
+redundancy_by_res <- 
+  list(
+    "0.1" = comm_scaled_list_0.1$poly_shp,
+    "0.2" = comm_scaled_list_0.2$poly_shp,
+    "0.3" = comm_scaled_list_0.3$poly_shp,
+    "0.4" = comm_scaled_list_0.4$poly_shp) %>%
+  map_df(~as_tibble(.) %>% select(grids, abundance, richness), .id = "res") %>%
+  # Calculate redundancy
+  mutate(redundancy = 1 - (richness/abundance))
+
 # Summarize redundancy statistics by grain size
 redundancy_summary <-
-  richness_by_res %>%
+  redundancy_by_res %>%
   mutate(above_50 = ifelse(redundancy > 0.5, 1, 0)) %>%
@@
 # Distribution of redundancy by grain size
-c <- ggplot(redund_by_res, aes(x = redundancy, y = res)) +
+c <- ggplot(redundancy_by_res, aes(x = redundancy, y = res)) +
   geom_density_ridges(scale = 1.5) + 

### 2020-07-06 R/plan.R

@@ -3,7 +3,7 @@ k_vals <- 2:20
 
 # Define analysis plan
 plan <- drake_plan (
-
+  
   # Load and process data ----
   
   # Load basic world map.
@@ -35,7 +35,7 @@ plan <- drake_plan (
   repro_data_raw = read_csv(
     file_in("data_raw/ebihara_2019/ESM1.csv"),
     col_types = "cccccnnnnn"),
-
+  
   repro_data = process_repro_data(repro_data_raw) %>%
     rename_taxa(green_list),
   
@@ -45,7 +45,7 @@ plan <- drake_plan (
     col_types = c("text", "numeric", "numeric", "text", "text", "text", "text"),
     col_names = c("species", "longitude", "latitude", "date", "tns_barcode", "herbarium_code", "taxon_id"),
     skip = 1),
-
+  
   # - standardize names to Green List
   occ_point_data = rename_taxa(occ_point_data_raw, green_list),
   
@@ -54,20 +54,19 @@ plan <- drake_plan (
     # check for missing data
     assert(not_na, longitude, latitude, taxon),
   
-  # Calculate richness, abundance, and redundancy at three scales: 
+  # Calculate richness, abundance, and redundancy at four scales: 
   # 0.1, 0.2, 0.3, and 0.4 degrees
-  richness_by_res = map_df(
-    c("0.1" = 0.1, "0.2" = 0.2, "0.3" = 0.3, "0.4" = 0.4), 
-    ~richness_from_points(occ_point_data_ferns, .), .id = "res"),
+  comm_scaled_list = target(
+    comm_from_points(
+      species_coods = occ_point_data_ferns,
+      res = scale,
+      lon = "longitude",
+      lat = "latitude",
+      species = "taxon"),
+    transform = map(scale = c(0.1, 0.2, 0.3, 0.4))
+  ),
   
   # Decide that 0.2 scale is optimal, use this for downstream analyses
-  comm_scaled_list_0.2 = phyloregion::points2comm(
-    dat = occ_point_data_ferns,
-    res = 0.2,
-    lon = "longitude",
-    lat = "latitude",
-    species = "taxon"),
-  
   # - extract community matrix
   comm_ferns = comm_from_points2comm(comm_scaled_list_0.2),
   
@@ -75,7 +74,7 @@ plan <- drake_plan (
   shape_ferns = shape_from_points2comm(comm_scaled_list_0.2) %>%
     # calculate redundancy
[…]

### 2020-07-07 ms/SI.Rmd
Remove ecostructure analysis from plan and MS

@@
 Shape indicates optimal *k* selected by \function{optimal{\_}phyloregion} function in the R package \package{phyloregion} (circle, non-optimal; triangle, optimal); color indicates optimal *k* selected manually (black, non-optimal; red, optimal; see Methods).
-
-\clearpage
-
-```{r k-plot-ecos, fig.width = 4, fig.height = 3}
-# Extract log likelihoods for each value of K from ecostructure results
-log_l <- map_dbl(species_motifs_ferns_combined, ~pluck(., "L")) %>%
-  tibble(k = names(.) %>% parse_number, log_l = .) %>%
-  mutate(
-    optimal_manual = ifelse(k == k_taxonomy_select, TRUE, FALSE)
-  )
-
-ggplot(log_l, aes(x = k, y = log_l)) +
-  geom_line() +
-  geom_point(aes(color = optimal_manual), size = 2.5) +
-  scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
-  labs(
-    y = "Log likelihood",
-    x = "Number of clusters") +
-  theme(legend.position = "none")
-
-ggsave(result_file(s_figure("k-plot-ecos"), "pdf"), width = 4, height = 3, units = "in")
-```
-
-**`r s_figure("k-plot-ecos")`**. `r s_figure_cap("k-plot-ecos")` Red indicates optimal *k* selected manually (*k* = 8, same as for taxonomic bioregions).
 

### 2020-07-07 ms/manuscript.Rmd
Remove ecostructure analysis from plan and MS

@@
   "regions_taxonomy",
-  "regions_phylogeny",
-  "species_motifs_ferns_8L"
+  "regions_phylogeny"
 ),
@@
 
-```{r ecostructure}
-tweak <- 0.75
-
-sites <- ses_div_ferns_spatial %>%
-  pull(grids)
-
-centroids <-
-ses_div_ferns_spatial %>%
-  select(geometry) %>%
-  mutate(centroid = sf::st_centroid(geometry)) %>%
-  sf::st_coordinates(x = .$centroid) %>%
-  as_tibble() %>%
-  mutate(site = sites)
-
-omega <-
-species_motifs_ferns_8L$omega %>%
-  as.data.frame %>%
-  rownames_to_column("site") %>%
-  left_join(centroids, by = "site") %>%
-  column_to_rownames("site")
-
-omega_centroids <- select(omega, X, Y) %>% # longitude, latitude
-  as.matrix()
-
-omega <- select(omega, -X, -Y)
-
-ecos_plot_pie2(
-    omega = omega,
-    image_width = 6000*tweak,
-    image_height = 4800*tweak,
-    radius = 0.08,
-    no_map = TRUE,
-    coords = omega_centroids,
-    long_lim = c(121, 148),
-    lat_lim = c(23, 47),
-    color = carto_pal(8, "Bold"),
-    blank_bg = "grey90",
-    path = here::here("results/species_motifs_ferns.png")
-)
-
-```
-
-![](`r here::here("results/species_motifs_ferns.png")`)
-
-**`r figure("ecostructure")`**.
-`r figure_cap("ecostructure")`.
-Pie charts represent the proportional contribution of each site to each of *k* = 8 spatial motifs, fit with a grade-of-membership model based on the community matrix using the R package \package{ecostructure}.
-
-\clearpage
-
 ```{r diversity-by-cluster, fig.width = 7, fig.height = 7}

### 2020-07-07 R/plan.R

@@ -1,6 +1,3 @@
-# Set vector of k-values to use for ecostructure
-k_vals <- 2:20
-
 # Define analysis plan
 plan <- drake_plan (
   
@@ -158,20 +155,6 @@ plan <- drake_plan (
     k = 6
   ),
 
-  # Ecostructure ----
-
-  species_motifs_ferns = target(
-      ecostructure::ecos_fit(
-        comm_ferns,
-        K = K, tol = 0.1, num_trials = 1),
-      transform = map(K = !!k_vals)
-    ),
-  
-  species_motifs_ferns_combined = target(
-    list(species_motifs_ferns) %>% set_names(k_vals),
-    transform = combine(species_motifs_ferns)
-  ),
-  
   # Traits ----
   
   # Run NMDS on traits

### 2020-07-07 ms/manuscript.Rmd
Remove ecostructure from MS

@@
 
-We analyzed ecological structure using the framework of @White2019 and their R package \package{ecostructure}.
-This method involves fitting the community matrix to a grade of membership model, which allows communities have partial membership in multiple groups [@White2019].
-Like bioregion analysis, this method also requires selecting the number of clusters (in this context called "motifs" after @White2019) to fit the model.
-We initially tested values of *k* from 2 to 20 and inspected the log-likelihood for each, but found that it continually improved without reaching a clear optimum (`r s_figure("k-plot-ecos")`).
-We therefore used the same *k* that was selected for taxonomic bioregions.
-
 ### Data accessibility
@@
 Using phylogenetic distances produced similar results, but Okinawa and the Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B).
-
-### Ecological structure
-
-Ecological structure partly resembled bioregions, with most sites in Hokkaido and northern Honshu belonging to the same motifs (dark-green and purple), and sites in Okinawa also mostly belonging to another motif (grey) (`r figure("ecostructure")`).
-However, the grade-of-membership model also revealed complex ecological structure that mostly tracks latitude and elevation, with particularly high turnover in the southern half of Honshu.
-A single motif united low-richness sites (dark purple).
 

### 2020-07-07 ms/SI.Rmd
Update MS

@@
 
+```{r sampling-curve, fig.width = 4, fig.height = 3}
+# Format fern abundances (no hybrids) as numeric vector
+fern_abun <-
+  occ_point_data_ferns %>%
+  group_by(taxon) %>%
+  count(sort = TRUE) %>%
+  pull(n)
+
+# Run iNEXT
+inext_out <- iNEXT::iNEXT(fern_abun, q=0, datatype="abundance")
+
+# create subset of observed points to add to plot as dots
+observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]
+
+# check estimated completeness
+est_completeness <- inext_out$iNextEst %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
+rich_interval <- glue::glue("{observed_data$qD.LCL %>% round} to {observed_data$qD.UCL %>% round}")
+completeness_interval <- glue::glue("{observed_data$SC.LCL %>% scales::percent()} to {observed_data$SC.UCL %>% scales::percent()}")
+
+# Make iNEXT plot
+inext_out$iNextEst %>%
+  filter(method != "observed") %>%
+  ggplot(aes(x=m, y=qD)) +
+  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70") +
+  geom_line(aes(linetype=method), size=0.8) +
+  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
+  geom_point(data=observed_data, size=3) +
+  scale_y_continuous("Richness") +
+  scale_x_continuous("No. specimens") +
+  theme_minimal() +
+  theme(
+    legend.position="bottom", 
+    legend.title=element_blank()
+  )
+
+ggsave(result_file(s_figure("sampling-curve"), "pdf"), width = 4, height = 3, units = "in")
+```
+
+**`r s_figure("sampling-curve")`**.
+`r s_figure_cap("sampling-curve")` Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line).
+Point indicates observed taxonomic richness (`r observed_data$qD` taxa, 95% confidence interval `r rich_interval` taxa).
+Taxa include species, subspecies, and varieties, excluding hybrids.
+Sampling completeness is estimated to be `r est_completeness` (95% confidence interval `r completeness_interval`).
+
+\clearpage
+
 ```{r grain-size, fig.width = 7, fig.height = 3}
@@
 **C**, Distribution of redundancy values by grain size.
+There is a sudden improvement in redundancy values from 0.1 to 0.2 degree grid-cells, but less improvement as grain size is increased beyond 0.2 degrees.
 
@@
 ```{r k-plot, fig.width = 7, fig.height = 3}
-k_taxonomy_select <- 8
-
 a <- k_taxonomy$df %>%
@@
     optimal_auto = ifelse(k == k_taxonomy$optimal$k, TRUE, FALSE),
-    optimal_manual = ifelse(k == k_taxonomy_select, TRUE, FALSE),
-    label = ifelse(k == k_taxonomy_select, glue::glue("k = {k}"), "")
+    label = ifelse(k == k_taxonomy$optimal$k, glue::glue("k = {k}"), "")
   ) %>%
@@
   geom_path() +
-  geom_point(aes(color = optimal_manual, shape = optimal_auto), size = 2.5) +
+  geom_point(aes(color = optimal_auto), size = 2.5) +
   geom_text(aes(label = label), vjust = 0, nudge_y = 0.02, size = 4) +
   scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
-  scale_shape_manual(values = c("TRUE" = 17, "FALSE" = 16)) +
   labs(
@@
 
-k_phylogeny_select <- 6
+k_phylogeny_select <- 8
 
@@
     optimal_auto = ifelse(k == k_phylogeny$optimal$k, TRUE, FALSE),
-    optimal_manual = ifelse(k == k_phylogeny_select, TRUE, FALSE),
-    label = ifelse(k == k_phylogeny_select, glue::glue("k = {k}"), "")
+    label = ifelse(k == k_phylogeny$optimal$k, glue::glue("k = {k}"), "")
   ) %>%
@@
   geom_path() +
-  geom_point(aes(color = optimal_manual, shape = optimal_auto), size = 2.5) +
+  geom_point(aes(color = optimal_auto), size = 2.5) +
   geom_text(aes(label = label), vjust = 0, nudge_y = 0.02, size = 4) +
   scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
-  scale_shape_manual(values = c("TRUE" = 17, "FALSE" = 16)) +
   labs(
     y = "Explained variance",
-    x = "Number of clusters",
-    shape = "Optimal k\n(auto)",
-    color = "Optimal k\n(manual)") +
+    x = "Number of clusters") +
   theme(
-    axis.title.y = element_blank()
+    legend.position = "none"
   )
@@
 **B**, phylogenetic bioregions.
-Shape indicates optimal *k* selected by \function{optimal{\_}phyloregion} function in the R package \package{phyloregion} (circle, non-optimal; triangle, optimal); color indicates optimal *k* selected manually (black, non-optimal; red, optimal; see Methods).
+Color indicates optimal *k* selected selected by \function{optimal{\_}phyloregion} function in the R package \package{phyloregion} (black, non-optimal; red, optimal; see Methods).
 

### 2020-07-07 ms/manuscript.Rmd
Update MS

@@
   "occ_point_data_ferns",
+  "occ_point_data_ferns_unfiltered",
   "ses_phy_ferns",
@@
 
-Here we leverage this exceptionally rich dataset to analyze the geographic distribution of biodiversity, reveal its spatial structure, and assess drivers of community assembly.
+Here we leverage this exceptionally rich dataset to analyze the geographic distribution of biodiversity in detail.
+We ask the following questions in our study system, the ferns of Japan: 1.
+How is biodiversity distributed?
+2.
+How is biodiversity structured?
+3.
+How well is biodiversity protected?
 
@@
 
-We used occurrence data of ferns in Japan comprising `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` vouchered specimens representing `r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` non-hybrid taxa (species, subspecies, and varieties).
-All specimens are housed at herbaria in Japan, and taxonomy follows the Japan Green List v.
+We used a list of all non-hybrid fern specimens housed at herbaria in Japan.
+We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v.
 1.01 (http://www.rdplants.org/gl/).
+Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`).
+The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (species, subspecies, and varieties).
+We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).
+The final cleaned occurrence dataset included `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` specimens (`r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` taxa).
 
-Any analysis of biodiversity must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling.
-We tested using 0.1, 0.2, or 0.4 degree grid-cells for defining communities (co-occurring species).
+A necessary step in any analysis of biodiversity is to set the grain size used to define communities.
+This decision must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling.
+We tested using 0.1, 0.2, 0.3, and 0.4 degree grid-cells for defining communities (co-occurring species).
 At each size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell).
 We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling.
-Preliminary analysis indicated that 0.2 degree grid-cells are optimal for our dataset, so we used this for all subsequent analyses.
+Preliminary analysis indicated that 0.2 degree grid-cells are optimal for our dataset (`r s_figure("grain-size")`).
+We filtered out any cells with redundancy < 0.1 (mostly cells along the coast with very low land area) and used this for subsequent analyses.
 
@@
 Qualitative traits include aspects of leaf morphology (e.g., margin shape, leaf color, indusia shape, spore arrangement, etc.).
-
-We used ploidy level and breeding system data in @Ebihara2019b.
 
@@
 
-We measured taxanomic diversity using species richness.
+We measured taxonomic diversity using taxonomic richness.
 
@@
 
-### Analysis of bioregions and ecological structure
+### Analysis of bioregions
 

### 2020-07-07 ms/captions.R

@@ -14,6 +14,7 @@ table_nums(name = "traits-used", caption = "Fern traits used in this study")
 
 # - SI figures
 s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
+s_fig_nums(name = "sampling-curve", "Species collection curve for the ferns of Japan fit with the iNEXT package.")
 s_fig_nums(name = "grain-size", "Effect of grain size on sampling redundancy.")
 s_fig_nums(name = "k-plot", "Selection of *k* for bioregions.")
 s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan, restricted dataset including only taxa endemic to Japan.")

### 2020-07-10 ms/SI.Rmd
Combine SES and spatial results in plan

@@
   "comm_scaled_list_0.4",
-  "richness_by_res",
   "occ_point_data_ferns",
-  "ses_phy_ferns",
-  "ses_phy_ferns_endemic",
-  "ses_traits_ferns",
-  "shape_ferns",
+  "ses_div_ferns_spatial",
+  "ses_div_ferns_endemic_spatial",
   "k_taxonomy",
   "k_phylogeny",
-  "regions_taxonomy",
-  "regions_phylogeny",
-  "species_motifs_ferns",
-  "species_motifs_ferns_combined",
   "traits_nmds",
@@
 ```{r plot-prep}
-# Combine bioregion results, biodiversity metrics, and shape data into single dataframe
-tax_regions_data <- regions_taxonomy %>%
-  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  select(grids, taxonomic_cluster)
-
-phy_regions_data <- regions_phylogeny %>%
-  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  select(grids, phylo_cluster)
-
-ses_div_ferns_spatial <-
-  shape_ferns %>%
-  left_join(ses_phy_ferns, by = c(grids = "site")) %>%
-  left_join(ses_traits_ferns, by = c(grids = "site")) %>%
-  left_join(tax_regions_data, by = "grids") %>%
-  left_join(phy_regions_data, by = "grids")
-
-# - same for endemic species
-ses_div_ferns_endemic_spatial <-
-shape_ferns %>%
-  left_join(ses_phy_ferns_endemic, by = c(grids = "site"))
-
 # Define ggplot2 theme for maps

### 2020-07-10 ms/manuscript.Rmd
Combine SES and spatial results in plan

@@
   "occ_point_data_ferns_unfiltered",
-  "ses_phy_ferns",
-  "ses_phy_ferns_endemic",
-  "ses_traits_ferns",
-  "shape_ferns",
+  "ses_div_ferns_spatial",
   "k_taxonomy",
-  "k_phylogeny",
-  "regions_taxonomy",
-  "regions_phylogeny"
+  "k_phylogeny"
 ),
@@
 ```{r plot-prep}
-tax_regions_data <- regions_taxonomy %>%
-  mutate(taxonomic_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  select(grids, taxonomic_cluster)
-
-phy_regions_data <- regions_phylogeny %>%
-  mutate(phylo_cluster = as.factor(cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  select(grids, phylo_cluster)
-
-ses_div_ferns_spatial <-
-  shape_ferns %>%
-  left_join(ses_phy_ferns, by = c(grids = "site")) %>%
-  left_join(ses_traits_ferns, by = c(grids = "site")) %>%
-  left_join(tax_regions_data, by = "grids") %>%
-  left_join(phy_regions_data, by = "grids")
-
-ses_div_ferns_endemic_spatial <-
-shape_ferns %>%
-  left_join(ses_phy_ferns_endemic, by = c(grids = "site"))
+# Change cluster to factor
+ses_div_ferns_spatial <- ses_div_ferns_spatial %>%
+  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
 

### 2020-07-10 R/plan.R

@@ -51,7 +51,7 @@ plan <- drake_plan (
   # - subset to just ferns (674 taxa)
   occ_point_data_ferns_unfiltered = subset_to_ferns(occ_point_data, ppgi),
   
-  # - filter to only points in second-degree mesh
+  # - filter out duplicates, restrict to only points in second-degree mesh
   # Shape file downloaded from http://gis.biodic.go.jp/
   # http://gis.biodic.go.jp/BiodicWebGIS/Questionnaires?kind=mesh2&filename=mesh2.zip
   occ_point_data_ferns = filter_occ_points(
@@ -176,6 +176,23 @@ plan <- drake_plan (
   
   # Run NMDS on traits
   traits_nmds = vegan::metaMDS(trait_distance_matrix, k = 3),
+  
+  # Combine results ----
+  
+  # Combine spatial data, alpha diversity, and regions
+  
+  # - All ferns
+  ses_div_ferns_spatial =
+    shape_ferns %>%
+    left_join(ses_phy_ferns, by = c(grids = "site")) %>%
+    left_join(ses_traits_ferns, by = c(grids = "site")) %>%
+    left_join(regions_taxonomy %>% rename(taxonomic_cluster = cluster), by = "grids") %>%
+    left_join(regions_phylogeny %>% rename(phylo_cluster = cluster), by = "grids"),
+  
+  # - Japan endemics only
+  ses_div_ferns_endemic_spatial =
+    shape_ferns %>%
+    left_join(ses_phy_ferns_endemic, by = c(grids = "site")),
   
   # Write out manuscript ----
   ms = rmarkdown::render(

### 2020-07-10 presentations/bsa_2020-07-27/index.Rmd
Add slides for BSA talk

---
output:
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom.css"]
    lib_dir: libs
    mathjax: default
    seal: false
    nature:
      slideNumberFormat: "" # To turn off slide numbers
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["left", "middle", "inverse"]
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE, eval = TRUE, cache = TRUE, dev = "svg")
```

```{r load-packages, cache = FALSE}
library(magick)
# library(gt)
library(ggtext)
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(rcartocolor)
library(tidyverse)
```

```{r load-data, cache = FALSE}
# Set cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

source(here::here("R/functions.R"))

# Load targets
loadd(list = c(
  "ses_div_ferns_spatial",
  "occ_point_data_ferns"),
  cache = ja_fern_cache)

```

```{r title-slide-image, eval = FALSE}
# Code to make title slide image, don't need to run when rendering slides
title_theme <- theme_void() +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    legend.position = "none",
     panel.background = element_rect(fill = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

# Change cluster to factor
ses_div_ferns_spatial <- ses_div_ferns_spatial %>%
  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)

a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  title_theme

b <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
  geom_sf(color = "transparent") +
  labs(fill = "SES RPD") +
  scale_fill_scico(
    palette = "roma",
    na.value="grey80",
    direction = -1,
    limits = c(
      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
    )) +
  title_theme

c <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
  geom_sf(color = "transparent") +
  labs(fill = "Cluster") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  title_theme

a + b + c + plot_layout(ncol = 1)

# Write out here, then crop manually
ggsave(
  plot = title_image, 
  filename = here::here("presentations/bsa_2020-07-27/images/title.png"),
  dpi = 300)

```

class: inverse, middle, title-slide, big-margin
background-color: #000000
background-position: 50% 50%
background-image: url(images/title_background.png)
background-size: contain

.pull-left-wide[
# Exploring dimensions of biodiversity in Japanese ferns
<br>
## Joel H. Nitta<sup>1</sup>, Atsushi Ebihara<sup>2</sup>

<sup>1</sup>University of Tokyo<br>
<span style = 'font-size: 80%;'>https://joelnitta.com</span>

<sup>2</sup>National Museum of Nature and Science, Japan

Botanical Society of America<br><span style = 'font-size: 80%;'>2020.07.27</span>
]

.pull-right-narrow[
![](images/title_cropped.png)
]

---
class: big-margin

## Biodiversity is a two-sided coin

Characterizing the **distribution of biodiversity** is needed to:

- understand the **processes** generating it

- understand how to **conserve** it

---
class: big-margin

## Most studies of biodiversity count species

However, taxonomic richness is only one dimension of biodiversity!

We should also consider

- Phylogenetic diversity

- Functional diversity

- How these are related to each other

This requires comprehensive sampling of taxa, DNA, and traits

---
class: big-margin

## Ferns of Japan

.pull-left[

- One of the best-collected fern floras in the world

- Extensive study by both scientists and amateur societies

]

.pull-right[
```{r gbif-colls-per-country}
# Read in pteridophyte GBIF collections per country data from
# Iwasaki lab presentation
collections_by_country <- read_csv(here::here("data_raw/gbif_pterido_coll_by_country.csv")) %>%
  mutate(country = factor(country)) %>%
  mutate(country = fct_reorder(country, n))

# Plot showing top 10
ggplot(collections_by_country, aes(x = country, y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "GBIF pteridophyte \ncollections (top 10)",
    y = "No. collections") +
  scale_y_continuous(labels = scales::comma) +
  ggplot2::theme_bw(base_size = 24) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    panel.grid.major.y = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(colour="black"),
      axis.text.y = ggplot2::element_text(colour="black")
  )
```
]

---
class: big-margin

## Ferns of Japan

.pull-left[
- One of the best-collected fern floras in the world

- Extensive study by both scientists and amateur societies
]

.pull-right[
```{r ja-pterido-coll-curve}
# Format fern abundances (no hybrids) as numeric vector
fern_abun <-
  occ_point_data_ferns %>%
  group_by(taxon) %>%
  count(sort = TRUE) %>%
  pull(n)

# Run iNEXT
inext_out <- iNEXT::iNEXT(fern_abun, q=0, datatype="abundance")

# create subset of observed points to add to plot as dots
observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]

# check estimated completeness
est_completeness <- inext_out$iNextEst %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
rich_interval <- glue::glue("{observed_data$qD.LCL %>% round} to {observed_data$qD.UCL %>% round}")
completeness_interval <- glue::glue("{observed_data$SC.LCL %>% scales::percent()} to {observed_data$SC.UCL %>% scales::percent()}")

# Make iNEXT plot
inext_out$iNextEst %>%
  filter(method != "observed") %>%
  ggplot(aes(x=m, y=qD)) +
  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70") +
  geom_line(aes(linetype=method), size=0.8) +
  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
  geom_point(data=observed_data, size=3) +
  scale_y_continuous("Richness") +
  scale_x_continuous("No. specimens") +
  labs(
    title = "Species collection curve",
    subtitle = "Japanese ferns (no hybrids)"
  ) +
  ggplot2::theme_bw(base_size = 24) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_text(colour="black"),
    axis.text.y = ggplot2::element_text(colour="black"),
    legend.position="bottom", 
    legend.title=element_blank()
    )

```
]

---
class: big-margin

## Goals of this study

Using the **ferns of Japan** as a study system, seek to understand:

1. How is biodiversity distributed?

2. How is biodiversity structured?

3. How well is biodiversity protected?

### 2020-07-11 presentations/bsa_2020-07-27/index.Rmd
Add conservation status results to slides

@@
 ---
+title: "Exploring dimensions of biodiversity in Japanese ferns"
 output:
@@
   "ses_div_ferns_spatial",
-  "occ_point_data_ferns"),
+  "occ_point_data_ferns",
+  "protected_areas"),
   cache = ja_fern_cache)
 
+# Read in pteridophyte GBIF collections per country data from
+# Iwasaki lab presentation
+collections_by_country <- read_csv(here::here("data_raw/gbif_pterido_coll_by_country.csv")) %>%
+  mutate(country = factor(country)) %>%
+  mutate(country = fct_reorder(country, n))
+
+# Download high-res map of Japan,
+# crop to spatial div results area
+japan <- rnaturalearth::ne_countries(country = "japan", scale = "large", returnclass = "sf") %>%
+  sf::st_crop(sf::st_bbox(ses_div_ferns_spatial))
+
+# Save currently used data as back-up in case data produced by plan changes
+data <- list(
+  ses_div_ferns_spatial = ses_div_ferns_spatial,
+  occ_point_data_ferns = occ_point_data_ferns,
+  collections_by_country = collections_by_country,
+  japan = japan
+)
+
+saveRDS(data, here::here("presentations/bsa_2020-07-27/data/data.RDS"))
 ```
@@
 ```{r gbif-colls-per-country}
-# Read in pteridophyte GBIF collections per country data from
-# Iwasaki lab presentation
-collections_by_country <- read_csv(here::here("data_raw/gbif_pterido_coll_by_country.csv")) %>%
-  mutate(country = factor(country)) %>%
-  mutate(country = fct_reorder(country, n))
-
-# Plot showing top 10
+# Plot showing top 10 pteridophyte collections in GBIF by country
 ggplot(collections_by_country, aes(x = country, y = n)) +
@@
   theme(
-    axis.title.y = element_blank(),
     panel.grid.minor = ggplot2::element_blank(),
@@
 3. How well is biodiversity protected?
+
+---
+## Japan is well-protected overall (ca. 30% coverage)
+
+```{r protected-areas, out.width = "120%", fig.asp = 0.7}
+japan_box <- sf::st_bbox(ses_div_ferns_spatial)
+
+status_labels <-
+protected_areas %>%
+  as_tibble %>%
+  select(status, area) %>%
+  group_by(status) %>%
+  summarize(
+    area = sum(area),
+    .groups = "drop"
+  ) %>%
+  mutate(
+    area = as.numeric(area),
+    area_chr = area %>% round %>% scales::number(big.mark = ","),
+    area_percent = area %>% magrittr::divide_by(377900) %>% scales::percent(),
+    label = glue::glue("{str_to_sentence(status)} ({area_percent})") %>% set_names(status)
+) %>%
+  pull(label)
+
+ggplot() +
+  geom_sf(data = japan) +
+  geom_sf(data = protected_areas, aes(fill = status), color = "transparent") +
+  coord_sf(xlim = c(japan_box$xmin, japan_box$xmax), ylim = c(japan_box$ymin, japan_box$ymax)) +
+  labs(
+    fill = "Conservation\nstatus"
+  ) +
+  scale_fill_viridis_d(breaks = names(status_labels), labels = status_labels)
+```
+
+---
+
+## Hotspots of taxonomic richness are well-protected
+
+```{r protected-richness, out.width = "120%", fig.asp = 0.7}
+# Subset to the top 5% for richness
+top_richness <-
+  ses_div_ferns_spatial %>%
+  mutate(rich_quant = ntile(richness, 20)) %>%
+  filter(rich_quant == 20)
+
+# Join with protected areas data
+st_crs(top_richness) <- st_crs(protected_areas)
+
+protected_richness <- st_join(top_richness, protected_areas) %>%
+  filter(str_detect(rownames(.), "\\.", negate = TRUE))
+
+# Make an inset showing area plotted in Japan
+richness_box <- st_bbox(top_richness)
+
+inset <- 
+  ggplot() +
+  geom_sf(data = japan) +
+  annotate(
+    "rect",
+    xmin = richness_box$xmin,
+    xmax = richness_box$xmax,
+    ymin = richness_box$ymin,
+    ymax = richness_box$ymax,
+    alpha = .2) +
+  theme_void() +
+  theme(
+    plot.background = element_rect(fill = "white", color = "transparent")
+  )
+
+# Crop japan background map to richness area
+japan_cropped_to_richness <- japan
+st_crs(japan_cropped_to_richness) <- st_crs(protected_richness)
+japan_cropped_to_richness <- sf::st_crop(japan_cropped_to_richness, sf::st_bbox(protected_richness))
+
+# plot showing protected status of high richness areas
+richness_initial_plot <- ggplot() +
+  geom_sf(data = japan_cropped_to_richness) +
+  geom_sf(data = protected_richness, aes(fill = status)) + 
+  scale_fill_viridis_d(breaks = c("low", "medium", "high"), labels = c("Low", "Medium", "High")) +
+  labs(
+    title = "Top 5% richness",
+    fill = "Conservation\nstatus"
+  )
+
+# final plot: add inset
+cowplot::ggdraw(richness_initial_plot) +
+  cowplot::draw_plot(inset, 0.8, 0.45, .30, .30, hjust = 1, vjust = 1) 
+```
+---
+
+## Hotspots of phylogenetic endemism are also well-protected
+
+```{r protected-pe, out.width = "120%", fig.asp = 0.7}
+# Subset to top 5% (vs. null) for PE
+sig_pe <-
+  ses_div_ferns_spatial %>%
+  filter(pe_obs_p > 0.95)
+
+# Join with protected areas data
+# this joins raster (PE) with polygon (protected area)
+# if they intersect. So it is a minimum estimation - we will
+# count a grid cell as 'protected' if it intersects any part of
+# a protected polygon.
+st_crs(sig_pe) <- st_crs(protected_areas)
+
+protected_pe <- st_join(sig_pe, protected_areas)
+
+protected_pe <-
+  protected_pe %>%
+  filter(str_detect(rownames(.), "\\.", negate = TRUE))
+
+# Make an inset showing area plotted in Japan
+pe_box <- st_bbox(sig_pe)
+
+inset <- 
+  ggplot() +
+  geom_sf(data = japan) +
+  annotate(
+    "rect",
+    xmin = pe_box$xmin,
+    xmax = pe_box$xmax,
+    ymin = pe_box$ymin,
+    ymax = pe_box$ymax,
+    alpha = .2) +
+  theme_void() +
+  theme(
+    plot.background = element_rect(fill = "white", color = "transparent")
+  )
+
+# Crop japan background map to PE area
+japan_cropped_to_pe <- japan
+st_crs(japan_cropped_to_pe) <- st_crs(protected_pe)
+japan_cropped_to_pe <- sf::st_crop(japan_cropped_to_pe, sf::st_bbox(protected_pe))
+
+# plot showing protected status of high PE areas
+pe_initial_plot <- ggplot() +
+  geom_sf(data = japan_cropped_to_pe) +
+  geom_sf(data = protected_pe, aes(fill = status)) + 
+  scale_fill_viridis_d(breaks = c("low", "medium", "high"), labels = c("Low", "Medium", "High"), na.value = "white") +
+  labs(
+    title = "Significantly high PE",
+    fill = "Conservation\nstatus"
+  )
+
+# final plot: add inset
+cowplot::ggdraw(pe_initial_plot) +
+  cowplot::draw_plot(inset, 0.06, 0.47, .2, .2, hjust = 0, vjust = 0) 
+```

### 2020-07-13 ms/manuscript.Rmd
Move themes to functions

@@
   mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
-
-map_theme <- theme(
-    panel.grid.major = element_line(color = "grey95", size = 0.1),
-    panel.background = element_rect(fill = "grey60"),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
-  )
-
-map_theme_light <- theme(
-    panel.grid.major = element_line(color = "white", size = 0.1),
-    panel.background = element_rect(fill = "grey90"),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
-  )
-
 ```

### 2020-07-13 R/functions.R

@@ -2209,6 +2209,32 @@ standard_theme2 <- function () {
     )
 }
 
+map_theme <- function() {
+  theme(
+    panel.grid.major = element_line(color = "grey95", size = 0.1),
+    panel.background = element_rect(fill = "grey60"),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
+  )
+}
+
+map_theme_light <- function () {
+  theme(
+    panel.grid.major = element_line(color = "white", size = 0.1),
+    panel.background = element_rect(fill = "grey90"),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
+  )
+}
+
+map_theme_pp <- function() {
+  theme(
+    panel.grid.major = element_line(color = "white", size = 0.1),
+    panel.background = element_rect(fill = "grey70"),
+    axis.ticks = element_blank()
+  )
+}
+
 #' Get the lower, upper, or absolute maximum value
 #' of a variable in a dataframe
 #'

### 2020-07-13 presentations/bsa_2020-07-27/index.Rmd
Update slides

@@
 ```{r load-packages, cache = FALSE}
+library(conflicted)
 library(magick)
-# library(gt)
 library(ggtext)
@@
 library(tidyverse)
+
+# Resolve conflicts
+conflicted::conflict_prefer("map", "purrr")
+conflicted::conflict_prefer("select", "dplyr")
+conflicted::conflict_prefer("filter", "dplyr")
+conflicted::conflict_prefer("gather", "tidyr")
+conflicted::conflict_prefer("extract", "magrittr")
+
+# Specify custom themes
+simple_map_theme <- function () {
+  theme(
+    panel.grid.major = element_line(color = "white", size = 0.1),
+    panel.background = element_rect(fill = "grey70"),
+    legend.position = "none",
+    axis.text = element_blank(),
+    axis.ticks = element_blank()
+  )
+}
+
+trend_line_theme <- function () {
+  theme(
+    panel.background = element_blank(),
+    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank()
+  )
+}
 ```
@@
   cache = ja_fern_cache)
+
+# Change phyloregion cluster to factor
+ses_div_ferns_spatial <- ses_div_ferns_spatial %>%
+  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
 
@@
   dpi = 300)
-
 ```
@@
 ---
+## Species richness is humped-shaped
+
+```{r lat-bins}
+# Bin SES diversity results by latitudinal band (0.2 degree width)
+ses_div_binned <-
+  ses_div_ferns_spatial %>%
+  mutate(
+    longitude = sf::st_centroid(geometry) %>% st_coordinates %>% magrittr::extract(,1),
+    latitude = sf::st_centroid(geometry) %>% st_coordinates %>% magrittr::extract(,2)) %>%
+  as_tibble %>%
+  dplyr::select(-geometry) %>%
+  mutate(
+    latitude = cut_width(latitude, width = 0.2, closed = "left") %>%
+      str_split(",") %>%
+      map_chr(1) %>%
+      parse_number) %>%
+  group_by(latitude) %>%
+  summarize(
+    across(c(richness, pe_obs_rank, contains("obs_z")), ~mean(.x, na.rm = TRUE))
+  )
+
+# Get min and max latitude for setting common limits between
+# maps and trendlines
+ja_lat_min <- ses_div_binned %>% pull(latitude) %>% min()
+ja_lat_max <- ses_div_binned %>% pull(latitude) %>% max()
+```
+
+```{r richness, out.width = "120%"}
+# Fern richness
+# - map
+a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(fill = "No. taxa") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_pp() +
+  theme(
+    legend.title = element_text(size = 20/.pt),
+    legend.text = element_text(size = 16/.pt),
+    legend.justification=c(0,1), legend.position=c(0,1)
+  ) +
+  ylim(ja_lat_min, ja_lat_max)
+
+# - line tracking mean richness
+b <-
+  ggplot(ses_div_binned, aes(y = latitude, x = richness)) +
+  geom_path() +
+  labs(
+    x = "No. taxa"
+  ) +
+  theme(
+    panel.background = element_blank(),
+    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank()
+  ) +
+  ylim(ja_lat_min, ja_lat_max)
+
+a + b + plot_layout(widths = c(3,1), heights = c(1,1))
+```
+
+---
+## Other raw values are correlated with richness
+
+```{r raw-div, out.width = "120%", fig.asp = 0.7}
+# Richness
+a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(title = "Taxonomic richness") +
+  simple_map_theme()
+
+b <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(title = "Phylogenetic diversity") +
+  simple_map_theme()
+
+c <- ggplot(ses_div_ferns_spatial, aes(fill = fd_obs)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(title = "Functional diversity") +
+  simple_map_theme()
+
+d <- ggplot(ses_div_ferns_spatial, aes(x = richness, y = pd_obs)) +
+  geom_point(shape = 1) +
+  # Fit a logarithmic general additive model
+  # https://stackoverflow.com/questions/40711980/r-ggplot2-fit-curve-to-scatter-plot
+  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(log(x)), size = 0.5) +
+  labs(
+    x = "No. taxa",
+    y = "PD"
+  ) +
+  jntools::standard_theme()
+
+e <- ggplot(ses_div_ferns_spatial, aes(x = richness, y = fd_obs)) +
+  geom_point(shape = 1) +
+  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(log(x)), size = 0.5) +
+  labs(
+    x = "No. taxa",
+    y = "FD"
+  ) +
+  jntools::standard_theme()
+
+a + b + c +
+  grid::textGrob("...as expected") + d + e +
+  plot_layout(nrow = 2, heights = c(0.5, 0.5))
+```
+
+---
+## Compare raw values to random expectation with <u>**S**</u>tandard <u>**E**</u>ffect <u>**S**</u>ize
+
+.pull-left[
+$SES = \frac{x - mean(x)}{sd(x)}$
+]
+
+.pull-right[
+```{r ses-example}
+null <- tibble(value = rnorm(10000))
+
+obs_1 <- 0.95
+obs_2 <- 0.5
+obs_3 <- 0.05
+
+ggplot(null, aes(x = value)) +
+  geom_histogram(bins = 50) +
+  geom_vline(xintercept = 0)
+```
+]
+
+---
+## Compare raw values to random expectation with <u>**S**</u>tandard <u>**E**</u>ffect <u>**S**</u>ize
+
+PD is mostly clustered throughout
+
+```{r ses-pd, out.width = "120%"}
+# SES PD
+# - map
+a <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES PD") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_pp() +
+  theme(
+    legend.title = element_text(size = 20/.pt),
+    legend.text = element_text(size = 16/.pt),
+    legend.justification=c(0,1), legend.position=c(0,1)
+  )
+
+# - line tracking mean SES RPD
+b <-
+  ggplot(ses_div_binned, aes(y = latitude, x = pd_obs_z)) +
+  geom_path() +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  labs(
+    x = "SES PD"
+  ) +
+  ylim(24, 46) +
+  theme(
+    panel.background = element_blank(),
+    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank()
+  )
+
+a + b + plot_layout(widths = c(3,1), heights = c(1,1))
+```
+
+---
+## Compare branch lengths with <u>**R**</u>elative <u>**P**</u>hylogenetic <u>**D**</u>iversity
+
+RPD = original branch lengths / all branch lengths equal
+
+---
+## Compare branch lengths with <u>**R**</u>elative <u>**P**</u>hylogenetic <u>**D**</u>iversity
+
+RPD is high in the N and S
+
+```{r ses-rpd, out.width = "120%"}
+# SES RPD
+# - map
+a <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES RPD") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_pp() +
+  theme(
+    legend.title = element_text(size = 20/.pt),
+    legend.text = element_text(size = 16/.pt),
+    legend.justification=c(0,1), legend.position=c(0,1)
+  )
+
+# - line tracking mean SES RPD
+b <-
+  ggplot(ses_div_binned, aes(y = latitude, x = rpd_obs_z)) +
+  geom_path() +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  labs(
+    x = "SES RPD"
+  ) +
+  ylim(24, 46) +
+  theme(
+    panel.background = element_blank(),
+    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank()
+  )
+
+a + b + plot_layout(widths = c(3,1), heights = c(1,1))
+```
+
+---
+## <u>**R**</u>elative <u>**F**</u>unctional <u>**D**</u>iversity: functional analog of RPD
+
+RFD also high in the south
+
+```{r ses-rpd-rfd, out.width = "120%"}
+# SES RPD
+# - map
+a <- ggplot(ses_div_ferns_spatial, aes(fill = rfd_obs_z)) +
+  geom_sf(color = "transparent") +
+  labs(fill = "SES RFD") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs")
+    )) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_pp() +
+  theme(
+    legend.title = element_text(size = 20/.pt),
+    legend.text = element_text(size = 16/.pt),
+    legend.justification=c(0,1), legend.position=c(0,1)
+  )
+
+# - line tracking mean SES RPD
+b <-
+  ggplot(ses_div_binned, aes(y = latitude, x = rfd_obs_z)) +
+  geom_path() +
+  geom_vline(xintercept = 0, linetype="dashed") +
+  labs(
+    x = "SES RFD"
+  ) +
+  ylim(24, 46) +
+  theme(
+    panel.background = element_blank(),
+    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.y = element_blank(),
+    plot.margin = margin(l = -1, unit = "inch")
+  )
+
+a + b + plot_layout(widths = c(3,1), heights = c(1,1))
+```
+
+---
+## Phylogenetic endemism: PD weighted by range size
+
+Southern islands are hotspots of endemism
+
+```{r pe, out.width = "120%"}
+# PE
+# - map
+a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_rank)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(fill = "PE Rank") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  ylim(ja_lat_min, ja_lat_max) +
+  map_theme_pp() +
+  theme(
+    legend.title = element_text(size = 20/.pt),
+    legend.text = element_text(size = 16/.pt),
+    legend.justification=c(0,1), legend.position=c(0,1)
+  )
+
+# - line tracking mean richness
+b <-
+  ggplot(ses_div_binned, aes(y = latitude, x = pe_obs_rank)) +
+  geom_path() +
+  labs(
+    x = "PE Rank"
+  ) +
+  ylim(ja_lat_min, ja_lat_max) +
+  trend_line_theme()
+
+a + b + plot_layout(widths = c(3,1), heights = c(1,1))
+```
+
+---
+## Bioregions
+
+Most sites fall into four bioregions
+
+```{r tax-bioregion, out.width = "120%", fig.asp = 0.7}
+ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_light()
+```
+
+---
 ## Japan is well-protected overall (ca. 30% coverage)
@@
   as_tibble %>%
-  select(status, area) %>%
+  dplyr::select(status, area) %>%
   group_by(status) %>%
@@
   ) +
-  scale_fill_viridis_d(breaks = names(status_labels), labels = status_labels)
+  scale_fill_viridis_d(breaks = names(status_labels), labels = status_labels) +
+  theme(
+    axis.ticks = element_blank()
+  )
 ```
@@
     fill = "Conservation\nstatus"
+  ) +
+  theme(
+    axis.ticks = element_blank()
   )

### 2020-07-13 presentations/bsa_2020-07-27/index.Rmd
Update slides

@@
 ---
-## Bioregions
-
-Most sites fall into four bioregions
+## Most sites fall into four bioregions
 
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light()
+  map_theme_light() +
+  labs(
+    title = "Taxonomic bioregions"
+  )
+```
+
+---
+## Taxonomic and phylogenetic bioregions are similar
+
+```{r tax-bioregion, out.width = "120%", fig.asp = 0.7}
+a <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_pp_light() +
+  labs(
+    title = "Taxonomic bioregions"
+  ) +
+  theme(legend.position = "none")
+
+# Swap first two colors to get them to match between taxonomic and phylogenetic plots
+cols_old <- rcartocolor::carto_pal(n = 8,  "Bold")
+cols <- cols_old
+cols[1] <- cols_old[2]
+cols[2] <- cols_old[1]
+
+b <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
+  geom_sf(color = "transparent") +
+  scale_fill_manual(values = cols) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  labs(title = "Phylogenetic bioregions") +
+  map_theme_pp_light() +
+  theme(
+    axis.text.y = element_blank(),
+    legend.position = "none")
+
+a + b 
+```
+
+---
+## Bioregions mostly match traditionally-defined forest types
+
+```{r tax-bioregion, out.width = "120%", fig.asp = 0.7}
+forest_types <- image_read(here::here("presentations/bsa_2020-07-27/images/forest_types_3_eng.png"))
+
+b <- cowplot::ggdraw() + 
+  cowplot::draw_image(forest_types)
+
+a <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
+  geom_sf(color = "transparent") +
+  scale_fill_manual(values = cols) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  labs(title = "Phylogenetic bioregions") +
+  map_theme_pp_light() +
+  theme(
+    axis.text.y = element_blank(),
+    legend.position = "none")
+
+a + b 
 ```

### 2020-07-14 ms/manuscript.Rmd
Fix theme calls in ms

@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme
+  map_theme()
 
@@
   labs(fill = "Redundancy") +
-  map_theme +
+  map_theme() +
   theme(
@@
   scale_fill_scico(palette = "lajolla") + 
-  map_theme +
+  map_theme() +
   theme(
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme +
+  map_theme() +
   theme(
@@
   # gghighlight(pd_obs_p > 0.975 | pd_obs_p < 0.025) +
-  map_theme +
+  map_theme() +
   theme(
@@
   # gghighlight(mpd_morph_obs_p > 0.975 | mpd_morph_obs_p < 0.025) +
-  map_theme +
+  map_theme() +
   theme(
@@
   scale_fill_scico(palette = "lajolla") + 
-  map_theme
+  map_theme()
 
@@
   # gghighlight(rfd_obs_p > 0.975 | rfd_obs_p < 0.025) +
-  map_theme +
+  map_theme() +
   theme(
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light
+  map_theme_light()
 

### 2020-07-14 presentations/bsa_2020-07-27/index.Rmd
Move some presentation-only functions to presentation rmd

@@
 }
+
+map_theme_pp <- function() {
+  theme(
+    panel.grid.major = element_line(color = "white", size = 0.1),
+    panel.background = element_rect(fill = "grey70"),
+    axis.ticks = element_blank()
+  )
+}
+
+map_theme_pp_light <- function() {
+  theme(
+    panel.grid.major = element_line(color = "white", size = 0.1),
+    panel.background = element_rect(fill = "grey90"),
+    axis.ticks = element_blank()
+  )
+}
+
+source(here::here("R/functions.R"))
 ```
@@
 
-source(here::here("R/functions.R"))
-
 # Load targets
 loadd(list = c(
+  "traits_for_dist",
   "ses_div_ferns_spatial",

### 2020-07-14 presentations/bsa_2020-07-27/index.Rmd
Update slides

--- 2020-07-13 presentations/bsa_2020-07-27/index.Rmd
@@
 
-```{r load-packages, cache = FALSE}
+```{r load-packages-funcs, cache = FALSE}
 library(conflicted)
@@
 <br>
-## Joel H. Nitta<sup>1</sup>, Atsushi Ebihara<sup>2</sup>
+## Joel H. Nitta<sup>1</sup>, Brent D. Misher<sup>2</sup>, Wataru Iwasaki<sup>1</sup>, Atsushi Ebihara<sup>3</sup>
 
-<sup>1</sup>University of Tokyo<br>
+<sup>1</sup>University of Tokyo, Japan<br>
+
+<sup>2</sup>University of California, Berkeley, USA<br>
+
+<sup>3</sup>National Museum of Nature and Science, Japan
+
 <span style = 'font-size: 80%;'>https://joelnitta.com</span>
-
-<sup>2</sup>National Museum of Nature and Science, Japan
-
-Botanical Society of America<br><span style = 'font-size: 80%;'>2020.07.27</span>
 ]
@@
 
-## Biodiversity is a two-sided coin
+## Biodiversity: more than just counting species
 
-Characterizing the **distribution of biodiversity** is needed to:
-
-- understand the **processes** generating it
-
-- understand how to **conserve** it
-
----
-class: big-margin
-
-## Most studies of biodiversity count species
-
-However, taxonomic richness is only one dimension of biodiversity!
+Species richness is only one dimension of biodiversity!
 
@@
 
-This requires comprehensive sampling of taxa, DNA, and traits
+<img src="images/not_just_species.png" alt="drawing" width="800"/>
 
@@
 
-## Ferns of Japan
+## Biodiversity: more than just counting species
+
+Species richness is only one dimension of biodiversity!
+
+We should also consider
+
+- Phylogenetic diversity
+
+- Functional diversity
+
+- How these are related to each other
+
+**This requires comprehensive sampling of taxa, DNA, and traits**
+
+---
+class: big-margin
+
+## Ferns of Japan: an ideal system for studying biodiversity
 
@@
 
+- Diverse flora (ca. 675 spp.)
+
 - One of the best-collected fern floras in the world
 
-- Extensive study by both scientists and amateur societies
+- Data available for DNA, distribution, and traits
 
@@
 
-## Ferns of Japan
+## (Very brief) introduction to the biogeography of Japan
 
-.pull-left[
-- One of the best-collected fern floras in the world
+- Wide variation in climate from north (subarctic) to south (subtropical)
 
-- Extensive study by both scientists and amateur societies
-]
+- Wide variation in elevation
 
-.pull-right[
-```{r ja-pterido-coll-curve}
-# Format fern abundances (no hybrids) as numeric vector
-fern_abun <-
-  occ_point_data_ferns %>%
-  group_by(taxon) %>%
-  count(sort = TRUE) %>%
-  pull(n)
+- Tertiary relict flora (never glaciated)
 
-# Run iNEXT
-inext_out <- iNEXT::iNEXT(fern_abun, q=0, datatype="abundance")
-
-# create subset of observed points to add to plot as dots
-observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]
-
-# check estimated completeness
-est_completeness <- inext_out$iNextEst %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
-rich_interval <- glue::glue("{observed_data$qD.LCL %>% round} to {observed_data$qD.UCL %>% round}")
-completeness_interval <- glue::glue("{observed_data$SC.LCL %>% scales::percent()} to {observed_data$SC.UCL %>% scales::percent()}")
-
-# Make iNEXT plot
-inext_out$iNextEst %>%
-  filter(method != "observed") %>%
-  ggplot(aes(x=m, y=qD)) +
-  geom_ribbon(aes(ymin = qD.LCL, ymax = qD.UCL), fill = "grey70") +
-  geom_line(aes(linetype=method), size=0.8) +
-  scale_linetype_manual(values=c("dotted", "solid", "solid")) +
-  geom_point(data=observed_data, size=3) +
-  scale_y_continuous("Richness") +
-  scale_x_continuous("No. specimens") +
-  labs(
-    title = "Species collection curve",
-    subtitle = "Japanese ferns (no hybrids)"
-  ) +
-  ggplot2::theme_bw(base_size = 24) +
-  theme(
-    panel.grid.minor = ggplot2::element_blank(),
-    axis.text.x = ggplot2::element_text(colour="black"),
-    axis.text.y = ggplot2::element_text(colour="black"),
-    legend.position="bottom", 
-    legend.title=element_blank()
-    )
-
-```
-]
+- Main islands continental, southern islands oceanic
 
@@
 ---
+class: big-margin
+## Data sources
+
+```{r summarize-traits}
+trait_summary <- make_trait_summary(traits_for_dist)
+
+n_binary <- trait_summary %>% filter(trait_type == "binary") %>% nrow()
+n_cont <- trait_summary %>% filter(trait_type == "continuous") %>% nrow()
+n_qual <- trait_summary %>% filter(trait_type == "qualitative") %>% nrow()
+```
+
+**Occurrence data**
+
+- Specimens at herbaria in Japan
+
+- `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` specimens (`r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` taxa)
+
+- Convert to 0.2° (ca. 20 km x 20 km) presence-absence matrix
+
+---
+class: big-margin
+## Data sources
+
+**Phylogenetic tree**
+
+- *rbcL* sequences from Japanese specimens (Ebihara and Nitta 2019)
+
+- Combine with broad sampling from genbank (ca. 60 genes x 4,000 taxa) (Nitta et al., in prep.)
+
+- Infer phylogeny with IQ-TREE (Nguyen et al. 2015)
+
+- Date with 26 fossil calibration points (Testo and Sundue 2016) using treePL (Smith and O'Meara 2012)
+
+- Subset to only Japanese taxa
+
+---
+class: big-margin
+
+## Data sources
+
+**Traits**
+
+- `r nrow(trait_summary)` morphological traits mostly used for ID (Ebihara and Nitta 2019)
+
+  - `r n_binary` binary (e.g., false indusium)
+
+  - `r n_cont` continuous (e.g., stipe length)
+
+  - `r n_qual` qualitative (e.g., indusium shape, lamina texture)
+
+---
+class: middle, inverse, big-margin
+
+# How is biodiversity distributed?
+
+---
 ## Species richness is humped-shaped
 
-```{r lat-bins}
-# Bin SES diversity results by latitudinal band (0.2 degree width)
-ses_div_binned <-
+```{r centroids}
+# Convert diversity metrics dataframe to centroids for plotting trends
+ses_div_centroids <-
   ses_div_ferns_spatial %>%
@@
   as_tibble %>%
-  dplyr::select(-geometry) %>%
-  mutate(
-    latitude = cut_width(latitude, width = 0.2, closed = "left") %>%
-      str_split(",") %>%
-      map_chr(1) %>%
-      parse_number) %>%
-  group_by(latitude) %>%
-  summarize(
-    across(c(richness, pe_obs_rank, contains("obs_z")), ~mean(.x, na.rm = TRUE))
-  )
+  dplyr::select(-geometry)
 
@@
 # maps and trendlines
-ja_lat_min <- ses_div_binned %>% pull(latitude) %>% min()
-ja_lat_max <- ses_div_binned %>% pull(latitude) %>% max()
+ja_lat_min <- ses_div_centroids %>% pull(latitude) %>% min()
+ja_lat_max <- ses_div_centroids %>% pull(latitude) %>% max()
 ```
@@
 
-# - line tracking mean richness
-b <-
-  ggplot(ses_div_binned, aes(y = latitude, x = richness)) +
-  geom_path() +
-  labs(
-    x = "No. taxa"
-  ) +
-  theme(
-    panel.background = element_blank(),
-    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
-    panel.grid.minor = element_blank(),
-    axis.title.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.y = element_blank()
-  ) +
-  ylim(ja_lat_min, ja_lat_max)
+# - latitudinal trends
+b <- ggplot(ses_div_centroids, aes(x = latitude, y = richness)) +
+  geom_point(color = "grey50", shape = 1 ) +
+  geom_smooth(se = FALSE) +
+  xlim(ja_lat_min, ja_lat_max) +
+  coord_flip() +
+  labs(y = "No. taxa") +
+  trend_line_theme()
 
@@
 
+.footnote[Trendline: General additive model (GAM) with cubic splines]
 ---
@@
 
-```{r raw-div, out.width = "120%", fig.asp = 0.7}
+```{r raw-div-1, out.width = "120%", fig.asp = 0.7}
 # Richness
@@
   scale_fill_scico(palette = "lajolla") + 
-  labs(title = "Phylogenetic diversity") +
+  labs(
+    title = "Phylogenetic diversity",
+    subtitle = "Faith's PD") +
   simple_map_theme()
@@
   scale_fill_scico(palette = "lajolla") + 
-  labs(title = "Functional diversity") +
+  labs(
+    title = "Functional diversity",
+    subtitle = "Faith's PD measured\non trait dendrogram") +
+  simple_map_theme()
+
+a + b + c +
+  plot_spacer() + plot_spacer() + plot_spacer() +
+  plot_layout(nrow = 2, heights = c(0.5, 0.5))
+```
+
+---
+## Other raw values are correlated with richness... as expected
+
+```{r raw-div-2, out.width = "120%", fig.asp = 0.7}
+# Richness
+a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(title = "Taxonomic richness") +
+  simple_map_theme()
+
+b <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(
+    title = "Phylogenetic diversity",
+    subtitle = "Faith's PD") +
+  simple_map_theme()
+
+c <- ggplot(ses_div_ferns_spatial, aes(fill = fd_obs)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") + 
+  labs(
+    title = "Functional diversity",
+    subtitle = "Faith's PD measured\non trait dendrogram") +
   simple_map_theme()
@@
 a + b + c +
-  grid::textGrob("...as expected") + d + e +
+  plot_spacer() + d + e +
   plot_layout(nrow = 2, heights = c(0.5, 0.5))
@@
 
+.footnote[Trendline: GAM with log-transform]
 ---
-## Compare raw values to random expectation with <u>**S**</u>tandard <u>**E**</u>ffect <u>**S**</u>ize
+class: big-margin
+## Compare raw values to random expectation with<br> <u>**S**</u>tandard <u>**E**</u>ffect <u>**S**</u>ize
 
-.pull-left[
-$SES = \frac{x - mean(x)}{sd(x)}$
+.large[
+.center[
+$SES = \frac{obs - mean(null)}{sd(null)}$
+]
 ]
 
-.pull-right[
-```{r ses-example}
-null <- tibble(value = rnorm(10000))
+SES = 0: not different from random
 
-obs_1 <- 0.95
-obs_2 <- 0.5
-obs_3 <- 0.05
+SES > 0: more diverse than expected
 
-ggplot(null, aes(x = value)) +
-  geom_histogram(bins = 50) +
-  geom_vline(xintercept = 0)
-```
-]
+SES < 0: less diverse than expected
+
+Null distribution: 999 communities randomized using "Independent Swap" (Gotelli, 2000)
 
@@
 
-# - line tracking mean SES RPD
-b <-
-  ggplot(ses_div_binned, aes(y = latitude, x = pd_obs_z)) +
-  geom_path() +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  labs(
-    x = "SES PD"
-  ) +
-  ylim(24, 46) +
-  theme(
-    panel.background = element_blank(),
-    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
-    panel.grid.minor = element_blank(),
-    axis.title.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.y = element_blank()
-  )
+# - latitudinal trends
+b <- ggplot(ses_div_centroids, aes(x = latitude, y = pd_obs_z)) +
+  geom_point(color = "grey50", shape = 1 ) +
+  geom_hline(yintercept = 0, linetype="dashed") +
+  geom_smooth(se = FALSE) +
+  xlim(ja_lat_min, ja_lat_max) +
+  coord_flip() +
+  labs(y = "SES PD") +
+  trend_line_theme()
 
@@
 
+.footnote[Trendline: GAM with cubic splines]
 ---
@@
 
-# - line tracking mean SES RPD
-b <-
-  ggplot(ses_div_binned, aes(y = latitude, x = rpd_obs_z)) +
-  geom_path() +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  labs(
-    x = "SES RPD"
-  ) +
-  ylim(24, 46) +
-  theme(
-    panel.background = element_blank(),
-    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
-    panel.grid.minor = element_blank(),
-    axis.title.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.y = element_blank()
-  )
+# - latitudinal trends
+b <- ggplot(ses_div_centroids, aes(x = latitude, y = rpd_obs_z)) +
+  geom_point(color = "grey50", shape = 1 ) +
+  geom_hline(yintercept = 0, linetype="dashed") +
+  geom_smooth(se = FALSE) +
+  xlim(ja_lat_min, ja_lat_max) +
+  coord_flip() +
+  labs(y = "SES RPD") +
+  trend_line_theme()
 
@@
 
-RFD also high in the south
+Southern islands have high RFD
 
@@
 
-# - line tracking mean SES RPD
-b <-
-  ggplot(ses_div_binned, aes(y = latitude, x = rfd_obs_z)) +
-  geom_path() +
-  geom_vline(xintercept = 0, linetype="dashed") +
-  labs(
-    x = "SES RFD"
-  ) +
-  ylim(24, 46) +
-  theme(
-    panel.background = element_blank(),
-    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
-    panel.grid.minor = element_blank(),
-    axis.title.y = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.y = element_blank(),
-    plot.margin = margin(l = -1, unit = "inch")
-  )
+# - latitudinal trends
+b <- ggplot(ses_div_centroids, aes(x = latitude, y = rfd_obs_z)) +
+  geom_point(color = "grey50", shape = 1 ) +
+  geom_hline(yintercept = 0, linetype="dashed") +
+  geom_smooth(se = FALSE) +
+  xlim(ja_lat_min, ja_lat_max) +
+  coord_flip() +
+  labs(y = "SES RFD") +
+  trend_line_theme()
 
@@
 # - map
-a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_rank)) +
+a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_p)) +
   geom_sf(color = "transparent") +
-  scale_fill_scico(palette = "lajolla") + 
-  labs(fill = "PE Rank") +
+  scale_fill_scico(palette = "lajolla", labels = scales::percent) + 
+  labs(fill = "PE percentile") +
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
@@
 # - line tracking mean richness
-b <-
-  ggplot(ses_div_binned, aes(y = latitude, x = pe_obs_rank)) +
-  geom_path() +
-  labs(
-    x = "PE Rank"
-  ) +
-  ylim(ja_lat_min, ja_lat_max) +
+b <- ggplot(ses_div_centroids, aes(x = latitude, y = pe_obs_p)) +
+  geom_point(color = "grey50", shape = 1 ) +
+  geom_hline(yintercept = 0.95, linetype="dashed") +
+  geom_smooth(se = FALSE) +
+  annotate("label", y = 0.95, x = ja_lat_max, label = "95%", size = 2) + 
+  xlim(ja_lat_min, ja_lat_max) +
+  scale_y_continuous(labels = scales::percent) +
+  labs(y = "PE percentile") +
+  coord_flip() +
   trend_line_theme()
@@
 
+.footnote[Trendline: GAM with cubic splines]
 ---
-## Bioregions
+class: middle, inverse, big-margin
 
-Most sites fall into four bioregions
+# How is biodiversity structured?
+
+---
+## Most sites fall into four bioregions
 
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light()
+  map_theme_pp_light() +
+  labs(
+    title = "Taxonomic bioregions"
+  )
 ```
+
+---
+## Taxonomic and phylogenetic bioregions are similar
+
+```{r compare-bioregion, out.width = "120%", fig.asp = 0.7}
+a <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_pp_light() +
+  labs(
+    title = "Taxonomic bioregions"
+  ) +
+  theme(legend.position = "none")
+
+# (If needed)
+# Swap first two colors to get them to match between taxonomic and phylogenetic plots
+# cols_old <- rcartocolor::carto_pal(n = 8,  "Bold")
+# cols <- cols_old
+# cols[1] <- cols_old[2]
+# cols[2] <- cols_old[1]
+
+b <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
+  geom_sf(color = "transparent") +
+  # scale_fill_manual(values = cols) +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  labs(title = "Phylogenetic bioregions") +
+  map_theme_pp_light() +
+  theme(
+    axis.text.y = element_blank(),
+    legend.position = "none")
+
+a + b 
+```
+
+---
+## Bioregions mostly match traditionally-defined forest types
+
+.pull-left[
+Taxonomic bioregions
+
+```{r tax-bioregion-forest}
+ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_pp_light() +
+  theme(legend.position = "none")
+```
+]
+
+.pull-right[
+Forest types
+
+<img src="images/forest_types_eng.png" alt="drawing" width="280"/>
+]
+
+.footnote[Shimizu 2014]
+
+---
+class: middle, inverse, big-margin
+
+# How well is biodiversity protected?
 
@@
 
-```{r protected-areas, out.width = "120%", fig.asp = 0.7}
+```{r protected-areas, out.width = "120%", fig.asp = 0.7, fig.retina=200}
 japan_box <- sf::st_bbox(ses_div_ferns_spatial)
@@
 ---
-
 ## Hotspots of taxonomic richness are well-protected
@@
 ```
+
 ---
-
 ## Hotspots of phylogenetic endemism are also well-protected
@@
 ```
+
+---
+class: big-margin
+
+## Conclusions
+
+How is biodiversity distributed?
+
+- Taxonomic diversity highest in Kyushu
+
+- Phylogenetic and functional diversity, phylogenetic endemism highest in Okinawa
+
+How is biodiversity structured?
+
+- Into four major clusters that track major forest types
+
+How well is biodiversity protected?
+
+- Well-protected in terms of status
+
+- May be in danger from herbivory by deer
+
+---
+class: big-margin
+
+## Future directions
+
+- Understand the role that hybridization has on diversification (ca. 400 described hybrids!)
+
+- Clarify the role of fern gametophytes in community assembly at broad scales

### 2020-07-14 presentations/bsa_2020-07-27/index.Rmd
update slides

@@
 ---
-class: big-margin
+background-position: 40% 50%
+background-image: url(images/japan_intro.png)
+background-size: contain
 
-## (Very brief) introduction to the biogeography of Japan
+## A (very brief) introduction to the biogeography of Japan
 
+.pull-left[
 - Wide variation in climate from north (subarctic) to south (subtropical)
@@
 - Main islands continental, southern islands oceanic
+]
 

### 2020-07-14 presentations/bsa_2020-07-27/index.Rmd
Update slides
- add references

@@
 source(here::here("R/functions.R"))
+```
+
+```{r, load_refs, include=FALSE, cache=FALSE}
+library(RefManageR)
+BibOptions(
+  ignore.case = TRUE,
+  check.entries = FALSE, 
+  style = "markdown"
+)
+bib <- ReadBib(here::here("presentations/bsa_2020-07-27/refs.bib"), check = FALSE)
 ```
@@
 
-- *rbcL* sequences from Japanese specimens (Ebihara and Nitta 2019)
+- *rbcL* sequences from Japanese specimens `r AutoCite(bib, "Ebihara2019b")`
 
@@
 
-- Infer phylogeny with IQ-TREE (Nguyen et al. 2015)
+- Infer phylogeny with IQ-TREE `r AutoCite(bib, "Nguyen2015")`
 
-- Date with 26 fossil calibration points (Testo and Sundue 2016) using treePL (Smith and O'Meara 2012)
+- Date with 26 fossil calibration points `r AutoCite(bib, "Testo2016a")` using treePL `r AutoCite(bib, "Smith2012")`
 
@@
 
-- `r nrow(trait_summary)` morphological traits mostly used for ID (Ebihara and Nitta 2019)
+- `r nrow(trait_summary)` morphological traits mostly used for ID `r AutoCite(bib, "Ebihara2019b")`
 
@@
 
-```{r richness, out.width = "120%"}
+```{r richness, out.width = "130%"}
 # Fern richness
@@
 ---
+## Classify bioregions using clustering
+
+.middle[
+<img src="images/daru_2017.png" alt="drawing" height="400"/>
+]
+
+(can also use taxonomic distances)
+
+.footnote[`r Citet(bib, "Daru2017")`]
+
+---
 ## Most sites fall into four bioregions
@@
 
-.footnote[Shimizu 2014]
+.footnote[`r Citet(bib, "shimizu_process_2014")`]
 
@@
 - Clarify the role of fern gametophytes in community assembly at broad scales
+
+---
+## References
+
+.small[
+```{r, results='asis', echo = FALSE, warning = FALSE}
+PrintBibliography(bib, start = 1, end = 5)
+```
+]
+
+---
+## References
+
+.small[
+```{r, results='asis', echo = FALSE, warning = FALSE}
+PrintBibliography(bib, start = 6, end = 10)
+```
+]

### 2020-07-14 presentations/bsa_2020-07-27/index.Rmd
Change citation style

@@
 
-```{r, load_refs, include=FALSE, cache=FALSE}
+```{r, load_refs, includ = FALSE, cache = FALSE}
+# Set bibliography options
 library(RefManageR)
@@
   check.entries = FALSE, 
-  style = "markdown"
+  bib.style = "authoryear", # set style of bibliography as e.g., "Nitta 2020..."
+  style = "text" # set style of citations as plain text (no links)
 )
+
+# Read in bibliography
 bib <- ReadBib(here::here("presentations/bsa_2020-07-27/refs.bib"), check = FALSE)

### 2020-07-15 presentations/bsa_2020-07-27/index.Rmd
update slides

@@
 ---
-background-position: 40% 50%
-background-image: url(images/japan_intro.png)
-background-size: contain
-
 ## A (very brief) introduction to the biogeography of Japan
 
-.pull-left[
+.pull-left-narrow[
 - Wide variation in climate from north (subarctic) to south (subtropical)
@@
 - Main islands continental, southern islands oceanic
+]
+
+.pull-right-wide[
+<img src="images/japan_intro_cropped.png" alt="drawing" height="580"/>
 ]
@@
 
-```{r richness, out.width = "130%"}
+```{r richness, out.width = "120%"}
 # Fern richness
@@
 .footnote[Trendline: General additive model (GAM) with cubic splines]
----
-## Other raw values are correlated with richness
+
+---
+class: big-margin
+## Measuring diversity with a tree
+
+.center[
+<img src="images/pd.png" alt="drawing" height="250"/>
+]
+
+<u>**P**</u>hylogenetic <u>**D**</u>iversity: Total unique branch length `r AutoCite(bib, "Faith1992")`
+
+<u>**F**</u>unctional <u>**D**</u>iversity: PD, but measured with a trait dendrogram `r AutoCite(bib, "Petchey2002")`
+
+---
+## PD and FD are correlated with richness
 
@@
   labs(
-    title = "Phylogenetic diversity",
-    subtitle = "Faith's PD") +
+    title = "Phylogenetic diversity") +
   simple_map_theme()
@@
   labs(
-    title = "Functional diversity",
-    subtitle = "Faith's PD measured\non trait dendrogram") +
+    title = "Functional diversity") +
   simple_map_theme()
@@
 ---
-## Other raw values are correlated with richness... as expected
+## PD and FD are correlated with richness... as expected
 
@@
 ---
+background-position: 40% 50%
+background-image: url(images/rpd.png)
+background-size: contain
+
 ## Compare branch lengths with <u>**R**</u>elative <u>**P**</u>hylogenetic <u>**D**</u>iversity
 
-RPD = original branch lengths / all branch lengths equal
+.pull-right[
+Long branches
+
+- Refugium
+
+- Occurrence of a few members of clades that mainly occur outside of the study region
+
+Short branches
+
+- Recently diverged lineages
+
+.small[`r Citet(bib, "Mishler2014")`]
+]
 
@@
 
-.footnote[Trendline: GAM with cubic splines]
+.footnote[Trendline: GAM with cubic splines `r NoCite(bib, "Faith1992")`]
 ---
@@
 
-.middle[
+.center[
 <img src="images/daru_2017.png" alt="drawing" height="400"/>
@@
 
-```{r tax-bioregion, out.width = "120%", fig.asp = 0.7}
+```{r tax-bioregion, out.width = "120%", fig.asp = 0.6}
 ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
@@
 
-.footnote[`r Citet(bib, "shimizu_process_2014")`]
+.footnote[`r Citet(bib, "Shimizu2014")`]
 
@@
 
-```{r protected-areas, out.width = "120%", fig.asp = 0.7, fig.retina=200}
+```{r protected-areas, out.width = "120%", fig.asp = 0.65, fig.retina=200}
 japan_box <- sf::st_bbox(ses_div_ferns_spatial)
@@
 class: big-margin
+## A growing threat: herbivory by deer (*Cervus nippon*)
+
+.center[
+<img src="images/deer_damage_1.png" alt="drawing" height="450"/>
+]
+
+---
+class: big-margin
+## A growing threat: herbivory by deer (*Cervus nippon*)
+
+.center[
+<img src="images/deer_damage_2.png" alt="drawing" height="500"/>
+]
+
+---
+class: big-margin
 
@@
 .small[
-```{r, results='asis', echo = FALSE, warning = FALSE}
+```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
 PrintBibliography(bib, start = 1, end = 5)
@@
 .small[
-```{r, results='asis', echo = FALSE, warning = FALSE}
+```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
 PrintBibliography(bib, start = 6, end = 10)

### 2020-07-15 presentations/bsa_2020-07-27/index.Rmd
Update slides
- add example ferns

@@
 
+simple_map_theme_light <- function () {
+  theme(
+    panel.grid.major = element_line(color = "white", size = 0.1),
+    panel.background = element_rect(fill = "grey90"),
+    legend.position = "none",
+    axis.text = element_blank(),
+    axis.ticks = element_blank()
+  )
+}
+
 trend_line_theme <- function () {
@@
 loadd(list = c(
-  "traits_for_dist",
   "ses_div_ferns_spatial",
   "occ_point_data_ferns",
-  "protected_areas"),
+  "protected_areas",
+  "traits_for_dist",
+  "comm_ferns",
+  "green_list",
+  "ppgi"),
   cache = ja_fern_cache)
-
-# Change phyloregion cluster to factor
-ses_div_ferns_spatial <- ses_div_ferns_spatial %>%
-  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
 
@@
   occ_point_data_ferns = occ_point_data_ferns,
+  protected_areas = protected_areas,
+  traits_for_dist = traits_for_dist,
+  comm_ferns = comm_ferns,
+  green_list = green_list,
+  ppgi = ppgi,
   collections_by_country = collections_by_country,
@@
 saveRDS(data, here::here("presentations/bsa_2020-07-27/data/data.RDS"))
+
+# Change phyloregion cluster to factor
+ses_div_ferns_spatial <- ses_div_ferns_spatial %>%
+  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
 ```
@@
 
+```{r count-taxa}
+# count number of higher taxa (genera and families)
+n_fern_taxa <-
+green_list %>%
+  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
+  assert(not_na, genus) %>%
+  # Add higher-level taxonomy
+  left_join(ppgi, by = "genus") %>%
+  assert(not_na, class) %>%
+  # Filter to only ferns
+  filter(class == "Polypodiopsida") %>%
+  summarize(
+    n_genus = n_distinct(genus),
+    n_family = n_distinct(family)
+  )
+```
+
 .pull-left[
 
-- Diverse flora (ca. 675 spp.)
+- Diverse flora
+
+ - ca. 675 spp.
+
+ - `r n_fern_taxa$n_genus` genera
+
+ - `r n_fern_taxa$n_family` families
 
@@
 - Data available for DNA, distribution, and traits
-
+]
+
+.pull-right[
+.center[
+<img src="images/ja_example_ferns.png" alt="drawing" height="580"/>
+]
+]
+
+.footnote[.small[Photos A. Ebihara]]
+
+---
+class: big-margin
+
+## Ferns of Japan: an ideal system for studying biodiversity
+
+.pull-left[
+
+- Diverse flora
+
+ - ca. 675 spp.
+
+ - `r n_fern_taxa$n_genus` genera
+
+ - `r n_fern_taxa$n_family` families
+
+- One of the best-collected fern floras in the world
+
+- Data available for DNA, distribution, and traits
 ]
@@
   geom_sf(color = "transparent") +
-  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_fill_carto_d(name = "Region", type = "qualitative", palette = "Bold") +
   scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
@@
 ---
+## Most sites fall into four bioregions
+
+```{r find-top-species}
+# Extract clusters
+clusters <-
+  ses_div_ferns_spatial %>%
+  as_tibble() %>%
+  select(site = "grids", cluster = taxonomic_cluster, -geometry)
+
+# Convert community df to long format
+comm_ferns_long <-
+  comm_ferns %>%
+  rownames_to_column("site") %>%
+  pivot_longer(names_to = "taxon", values_to = "abun", -site)
+
+# Find top-scoring species (with high occurrence inside one cluster, and low elsewhere)
+top_species <-
+  comm_ferns_long %>%
+  left_join(clusters, by = "site") %>%
+  filter(cluster %in% 1:4) %>%
+  group_by(taxon, cluster) %>%
+  summarize(
+    abun = sum(abun),
+    total = n()
+  ) %>%
+  mutate(
+    # percent abundance within a bioregion
+    perc_abun = abun / total
+  ) %>%
+  arrange(taxon, desc(perc_abun)) %>%
+  select(taxon, cluster, perc_abun) %>%
+  group_by(taxon) %>%
+  # take top two highest percent abundances per species, calculate their difference
+  slice_max(n = 2, order_by = perc_abun, .preserve = TRUE, with_ties = FALSE) %>%
+  mutate(
+    top_cluster = cluster[perc_abun == max(perc_abun)]
+  ) %>%
+  summarize(
+    max_dif = max(perc_abun) - min(perc_abun),
+    cluster = unique(top_cluster)
+  ) %>%
+  # keep only those with greater than 65% difference between top 1st and 2nd 
+  # most abundant bioregions
+  filter(max_dif > 0.5) %>%
+  group_by(cluster) %>%
+  # Choose top three per cluster for plotting
+  slice_max(n = 3, order_by = max_dif, with_ties = FALSE)
+```
+
+```{r plot-top-species-prep}
+# make helper function to plot top-scoring species
+plot_top_taxon <- function (taxon_select, ses_div_ferns_spatial, comm_ferns_long) {
+ses_div_ferns_spatial %>%
+  left_join(
+    filter(comm_ferns_long, taxon == taxon_select),
+    by = c(grids = "site")
+  ) %>%
+  mutate(abun = factor(abun)) %>%
+  ggplot(aes(fill = abun)) +
+  scale_fill_manual(values = c("#404040", "#d01c8b")) +
+  geom_sf(color = "transparent") +
+  labs(
+    subtitle = str_replace_all(taxon_select, "_", " ")
+  ) +
+  simple_map_theme_light() +
+  theme(
+    plot.subtitle = element_text(face = "italic")
+  )
+}
+```
+
+```{r plot-top-species-cluster-1, out.width = "120%", fig.asp = 0.65}
+focus_region <-
+ses_div_ferns_spatial %>%
+  mutate(focus = ifelse(taxonomic_cluster == 1, "yes", "no")) %>%
+ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  simple_map_theme_light() +
+  labs(subtitle = "Region 1")
+
+focus_region +
+plot_top_taxon(top_species$taxon[[1]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[2]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[3]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+```
+
+---
+## Most sites fall into four bioregions
+
+```{r plot-top-species-cluster-2, out.width = "120%", fig.asp = 0.65}
+focus_region <-
+ses_div_ferns_spatial %>%
+  mutate(focus = ifelse(taxonomic_cluster == 2, "yes", "no")) %>%
+ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  simple_map_theme_light() +
+  labs(subtitle = "Region 2")
+
+focus_region +
+plot_top_taxon(top_species$taxon[[4]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[5]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[6]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+```
+
+---
+## Most sites fall into four bioregions
+
+```{r plot-top-species-cluster-3, out.width = "120%", fig.asp = 0.65}
+focus_region <-
+ses_div_ferns_spatial %>%
+  mutate(focus = ifelse(taxonomic_cluster == 3, "yes", "no")) %>%
+ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  simple_map_theme_light() +
+  labs(subtitle = "Region 3")
+
+focus_region +
+plot_top_taxon(top_species$taxon[[7]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[8]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[9]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+```
+
+---
+## Most sites fall into four bioregions
+
+```{r plot-top-species-cluster-4, out.width = "120%", fig.asp = 0.65}
+bonin_box <-
+ses_div_ferns_spatial %>%
+  filter(taxonomic_cluster == 4) %>%
+  sf::st_bbox()
+
+expand_value <- 0.4
+
+focus_region <-
+ses_div_ferns_spatial %>%
+  mutate(focus = ifelse(taxonomic_cluster == 4, "yes", "no")) %>%
+ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  simple_map_theme_light() +
+  labs(subtitle = "Region 4") +
+  annotate(
+    "rect", 
+           xmin = bonin_box[["xmin"]] - expand_value, 
+           xmax = bonin_box[["xmax"]] + expand_value, 
+           ymin = bonin_box[["ymin"]] - expand_value, 
+           ymax = bonin_box[["ymax"]] + expand_value,
+    fill = "transparent",
+    color = "grey10")
+
+focus_region +
+plot_top_taxon(top_species$taxon[[10]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[12]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[11]], ses_div_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+```
+
+---
 ## Taxonomic and phylogenetic bioregions are similar
 
-```{r compare-bioregion, out.width = "120%", fig.asp = 0.7}
+```{r compare-bioregion, out.width = "120%", fig.asp = 0.65}
 a <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
@@
 .pull-right[
-Forest types
-
+Forest types<br>
+.small[`r Citet(bib, "Shimizu2014")`<br><br>]
 <img src="images/forest_types_eng.png" alt="drawing" width="280"/>
 ]
-
-.footnote[`r Citet(bib, "Shimizu2014")`]
 

### 2020-07-15 presentations/bsa_2020-07-27/index.Rmd
Update slides

@@
 
-.footnote[.small[Photos A. Ebihara]]
+.footnote[.small[Photos: A. Ebihara]]
 
@@
 .pull-right-wide[
-<img src="images/japan_intro_cropped.png" alt="drawing" height="580"/>
+<img src="images/japan_intro.png" alt="drawing" height="580"/>
 ]
@@
 ---
-## Compare raw values to random expectation with <u>**S**</u>tandard <u>**E**</u>ffect <u>**S**</u>ize
-
-PD is mostly clustered throughout
+## SES of PD is mostly clustered throughout
 
@@
 background-position: 40% 50%
-background-image: url(images/rpd.png)
+background-image: url(images/RPD.png)
 background-size: contain
@@
 
-- Refugium
+- Refugia
 
@@
 
-- Recently diverged lineages
+- Recently speciation
 
@@
 ---
-## Compare branch lengths with <u>**R**</u>elative <u>**P**</u>hylogenetic <u>**D**</u>iversity
-
-RPD is high in the N and S
+## RPD is high in the north and south
 
@@
 ---
-## <u>**R**</u>elative <u>**F**</u>unctional <u>**D**</u>iversity: functional analog of RPD
-
-Southern islands have high RFD
+## Southern islands have high RFD
+
+(RFD = functional analog of RPD)
 
@@
 ---
-## Phylogenetic endemism: PD weighted by range size
-
-Southern islands are hotspots of endemism
+## Southern islands are hotspots of endemism
+
+(Phylogenetic endemism = PD weighted by range size)
 
@@
 ---
+class:: big-margin
+
+# Acknowledgements
+
+.center[
+<img src="images/pags-logo.png" alt="drawing" width="600"/>
+
+<img src="images/shida-no-kai-logo.png" alt="drawing" height="120"/>
+
+Members of the Iwasaki Lab (The University of Tokyo)
+
+Eric Schuettpelz (Smithsonian Institution)
+
+Alexander White (Smithsonian Institution)
+]
+
+---
 ## References

### 2020-07-15 presentations/bsa_2020-07-27/index.Rmd
Update slides

@@
     nature:
-      slideNumberFormat: "" # To turn off slide numbers
+      # slideNumberFormat: "" # To turn off slide numbers
       highlightStyle: github

### 2020-07-15 presentations/bsa_2020-07-27/index.Rmd
Update slides

@@
 background-color: #000000
-background-position: 50% 50%
-background-image: url(images/title_background.png)
-background-size: contain
 
@@
 
-```{r protected-areas, out.width = "120%", fig.asp = 0.65, fig.retina=200}
+```{r protected-areas, out.width = "120%", fig.asp = 0.65, dev="png", dpi=200}
 japan_box <- sf::st_bbox(ses_div_ferns_spatial)

### 2020-07-15 presentations/bsa_2020-07-27/index.Rmd
Update slides

@@
 
-- Combine with broad sampling from genbank (ca. 60 genes x 4,000 taxa) (Nitta et al., in prep.)
+- Combine with broad sampling from GenBank
 
@@
 
-- Date with 26 fossil calibration points `r AutoCite(bib, "Testo2016a")` using treePL `r AutoCite(bib, "Smith2012")`
+- Date with 27 fossil calibration points `r AutoCite(bib, "Testo2016a")` using treePL `r AutoCite(bib, "Smith2012")`
 
@@
   labs(fill = "No. taxa") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
@@
     legend.justification=c(0,1), legend.position=c(0,1)
-  ) +
-  ylim(ja_lat_min, ja_lat_max)
+  )
 
@@
     )) +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
@@
   labs(fill = "PE percentile") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  ylim(ja_lat_min, ja_lat_max) +
   map_theme_pp() +
@@
 
-## Conclusions
-
+# Conclusions
+
+- How is biodiversity distributed?
+
+- How is biodiversity structured?
+
+- How well is biodiversity protected?
+
+---
+class: big-margin
+
+# Conclusions
+
+```{r conclusions-plots}
+a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") +
+  labs(subtitle = "Richness") +
+  simple_map_theme()
+
+b <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
+    )) +
+  labs(subtitle = "RPD") +
+  simple_map_theme()
+
+c <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
+  geom_sf(color = "transparent") +
+  labs(subtitle = "Taxonomic bioregions") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  simple_map_theme()
+```
+
+.pull-left[
 How is biodiversity distributed?
@@
 - Phylogenetic and functional diversity, phylogenetic endemism highest in Okinawa
-
+]
+
+.pull-right[
+```{r conclusions-richness, out.width="450px"}
+a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(palette = "lajolla") +
+  labs(title = "Richness") +
+  simple_map_theme() +
+  theme(
+    plot.title = element_text(size = 18)
+  )
+
+b <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
+  geom_sf(color = "transparent") +
+  scale_fill_scico(
+    palette = "romaO",
+    na.value="grey80",
+    direction = -1,
+    limits = c(
+      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
+      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
+    )) +
+  labs(title = "RPD") +
+  simple_map_theme() +
+  theme(
+    plot.title = element_text(size = 18)
+  )
+
+a + b + plot_layout(ncol = 1)
+```
+]
+
+---
+class: big-margin
+
+# Conclusions
+
+.pull-left[
 How is biodiversity structured?
@@
 - Into four major clusters that track major forest types
-
+]
+
+.pull-right[
+```{r conclusions-regions}
+ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
+  geom_sf(color = "transparent") +
+  labs(title = "Taxonomic bioregions") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  simple_map_theme_light() +
+  theme(
+    plot.title = element_text(size = 18)
+  )
+```
+]
+
+---
+class: big-margin
+
+# Conclusions
+.pull-left[
 How well is biodiversity protected?
@@
 
-- May be in danger from herbivory by deer
-
----
-class: big-margin
-
-## Future directions
-
-- Understand the role that hybridization has on diversification (ca. 400 described hybrids!)
-
+- May be in danger due to herbivory by deer
+]
+
+.pull-right[
+```{r conclusions-protection}
+ ggplot() +
+  geom_sf(data = japan_cropped_to_richness) +
+  geom_sf(data = protected_richness, aes(fill = status)) + 
+  scale_fill_viridis_d(breaks = c("low", "medium", "high"), labels = c("Low", "Medium", "High")) +
+  labs(title = "Top 5% richness hotspots", fill = "Conservation\nstatus"
+  ) +
+  theme(
+    axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    plot.title = element_text(size = 18)
+  )
+```
+]
+
+---
+# Future directions
+
+.pull-left[
+<br>
+- Understand the role that hybridization has on diversification <br> (ca. 300 described hybrids!) `r NoCite(bib, "Ebihara2009")`
+
+<br>
+<br><br>
+<br>
 - Clarify the role of fern gametophytes in community assembly at broad scales
+]
+
+.pull-right[
+<img src="images/vandenboschia.png" alt="drawing" width="280"/>
+
+<img src="images/IMGP5729.jpeg" alt="drawing" width="300"/><br>.tiny[Photo: Joel Nitta]
+]
 
@@
 ```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
-PrintBibliography(bib, start = 1, end = 5)
+PrintBibliography(bib, start = 1, end = 6)
 ```
@@
 ```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
-PrintBibliography(bib, start = 6, end = 10)
-```
-]
+PrintBibliography(bib, start = 7, end = 12)
+```
+]

### 2020-07-16 presentations/bsa_2020-07-27/index.Rmd
Update slides
- add background images for selected fern species

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE, eval = TRUE, cache = TRUE, dev = "svg")
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE, eval = TRUE, cache = TRUE, dev = "svg", dev.args=list(bg="transparent"))
 ```
@@
 
-```{r, load_refs, includ = FALSE, cache = FALSE}
+```{r, load-refs, includ = FALSE, cache = FALSE}
 # Set bibliography options
@@
   check.entries = FALSE,
+  max.names = 3,
   bib.style = "authoryear", # set style of bibliography as e.g., "Nitta 2020..."
@@
 
-- Infer phylogeny with IQ-TREE `r AutoCite(bib, "Nguyen2015")`
+- Infer phylogeny with IQ-TREE (Nguyen et al., 2015) `r NoCite(bib, "Nguyen2015")`
 
@@
 
-- Recently speciation
-
-.small[`r Citet(bib, "Mishler2014")`]
+- Recent speciation
+
+.small[Mishler et al. (2014) `r NoCite(bib, "Mishler2014")`]
 ]
@@
 
-.footnote[Trendline: GAM with cubic splines `r NoCite(bib, "Faith1992")`]
+.footnote[Trendline: GAM with cubic splines]
 ---
@@
 
-(Phylogenetic endemism = PD weighted by range size)
+Phylogenetic endemism = PD weighted by range size (Rosauer et al., 2009) `r NoCite(bib, "Rosauer2009")`
 
@@
 
-.footnote[`r Citet(bib, "Daru2017")`]
+.footnote[`r Citet(bib, c("Daru2017", "Daru2020"))`]
 
@@
 ---
+background-position: 50% 50%
+background-image: url(images/species-background-1.png)
+background-size: contain
+
 ## Most sites fall into four bioregions
@@
 
----
+.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]
+
+---
+background-position: 50% 50%
+background-image: url(images/species-background-2.png)
+background-size: contain
+
 ## Most sites fall into four bioregions
@@
 
----
+.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]
+
+---
+background-position: 50% 50%
+background-image: url(images/species-background-3.png)
+background-size: contain
+
 ## Most sites fall into four bioregions
@@
 
----
+.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]
+
+---
+background-position: 50% 50%
+background-image: url(images/species-background-4.png)
+background-size: contain
+
 ## Most sites fall into four bioregions
@@
 ```
+
+.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]
 
@@
 
-- Phylogenetic and functional diversity, phylogenetic endemism highest in Okinawa
+- Phylogenetic and functional diversity, phylogenetic endemism highest in southern islands
 ]
@@
 ```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
-PrintBibliography(bib, start = 1, end = 6)
+PrintBibliography(bib, start = 1, end = 8)
 ```
@@
 ```{r, results='asis', echo = FALSE, warning = FALSE, cache = FALSE}
-PrintBibliography(bib, start = 7, end = 12)
-```
-]
+PrintBibliography(bib, start = 9, end = 16)
+```
+]

### 2020-07-16 presentations/bsa_2020-07-27/convert.R

@@ -1,0 +1,54 @@
+# Convert html slides output by xaringan to pdf, then powerpoint
+
+# requires decktape to be installed
+
+# see https://github.com/gadenbuie/xaringan2powerpoint
+
+setwd(here::here("presentations/bsa_2020-07-27/"))
+
+# Specify slides html file
+slides_html <- "index.html"
+
+# "print" HTML to PDF
+xaringan::decktape(slides_html, "slides.pdf", docker = FALSE)
+
+# how many pages?
+pages <- pdftools::pdf_info("slides.pdf")$pages
+
+# set filenames
+filenames <- sprintf("slides/slides_%02d.png", 1:pages)
+
+# create slides/ and convert PDF to PNG files
+# overwrite any existing folder: use carefully!
+fs::dir_delete("slides")
+fs::dir_create("slides")
+
+pdftools::pdf_convert("slides.pdf", filenames = filenames, dpi = 150)
+
+# Template for markdown containing slide images
+slide_images <- glue::glue("
+---
+
+![]({filenames}){{width=100%, height=100%}}
+
+")
+slide_images <- paste(slide_images, collapse = "\n")
+
+# R Markdown -> powerpoint presentation source
+md <- glue::glue("
+---
+output: powerpoint_presentation
+---
+
+{slide_images}
+")
+
+cat(md, file = "slides_powerpoint.Rmd")
+
+# Render Rmd to powerpoint
+rmarkdown::render("slides_powerpoint.Rmd")  ## powerpoint!
+
+# Cleanup
+fs::file_delete("slides_powerpoint.Rmd")
+fs::dir_delete("slides")
+

### 2020-07-23 ms/manuscript.Rmd
Fix p-value calculations for randomizations

@@
 
-# PE rank
-b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_rank)) +
+# PE p-value
+b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_p_upper)) +
   geom_sf(color = "transparent") +
-  labs(fill = "Rank PE") +
+  labs(fill = "PE p-score") +
   scale_fill_scico(palette = "lajolla") +
@@
   as_tibble %>%
-  select(pd_obs_z, rpd_obs_z, fd_obs_z, rfd_obs_z, pe_obs_rank, taxonomic_cluster)
+  select(pd_obs_z, rpd_obs_z, fd_obs_z, rfd_obs_z, pe_obs_p_upper, taxonomic_cluster)
 
@@
 
-e <- ggplot(data, aes(x = taxonomic_cluster, y = pe_obs_rank)) +
+e <- ggplot(data, aes(x = taxonomic_cluster, y = pe_obs_p_upper)) +
   geom_jitter(aes(color = taxonomic_cluster)) +
@@
   labs(
-    y = "Rank PE",
+    y = "PE p-score",
     x = "Cluster",
@@
 **D**, SES of relative FD (RFD).
-**E**, rank of observed phylogenetic endemism (PE) out of 999 random communities.
+**E**, p-score of observed phylogenetic endemism (PE) relative to 999 random communities.
 Only top four clusters with the most sites shown.

### 2020-07-23 R/functions.R

@@ -134,7 +134,7 @@ subset_to_ferns <- function(data, ppgi) {
 subset_tree <- function(phy, ppgi) {
   
   tips_keep <-
-  tibble(tip = phy$tip.label) %>%
+    tibble(tip = phy$tip.label) %>%
     mutate(genus = str_split(tip, "_") %>% map_chr(1)) %>%
     # Add higher-level taxonomy
     left_join(ppgi, by = "genus") %>%
@@ -587,7 +587,7 @@ format_traits <- function(path_to_lucid_traits, taxon_id_map) {
       )) %>%
     select(-contains("leaf_sorus_indusium_presence_absence")) %>%
     rename(leaf_sorus_false_indusium_present = leaf_sorus_false_indusium)
- 
+  
   #### Clean up numeric traits ###
   # Replace missing (0 or 3) with NA,
   # take the mean of the range of normal values otherwise
@@ -643,7 +643,7 @@ format_traits <- function(path_to_lucid_traits, taxon_id_map) {
   
   # Fix some taxon names (synonyms)
   traits_for_dist <-
-  traits_for_dist %>%
+    traits_for_dist %>%
     mutate(taxon = case_when(
       # taxon == "Athyrium_nudum" ~ missing
       taxon == "Athyrium_opacum_opacum" ~ "Athyrium_opacum",
@@ -675,7 +675,7 @@ format_traits <- function(path_to_lucid_traits, taxon_id_map) {
   
   # Drop any missing names
   traits_for_dist <- filter(traits_for_dist, taxon %in% taxon_id_map$taxon)
-    
+  
   msg <- assertthat::validate_that(
     length(missing_taxon_id) == 0,
     msg = glue::glue("The following taxa in the trait data could not be verified in the taxa list and have been dropped: {paste(missing_taxon_id, collapse = ', ')}")
@@ -1140,7 +1140,7 @@ calculate_coverage <- function (comm_list) {
   
   # Extract community matrix and convert to data frame
   comm <-
-  magrittr::extract2(comm_list, "comm_dat") %>%
+    magrittr::extract2(comm_list, "comm_dat") %>%
     phyloregion::sparse2dense() %>%
     as.data.frame()
   
@@ -1323,6 +1323,46 @@ calc_biodiv_random <- function (
   
 }
 
+#' Count number of times one number is higher than others
+#'
+#' @param x Number to count
[…]

### 2020-07-24 ms/SI.Rmd
Update SI
- move ggplot themes to functions.R
- fix variable names for canape

@@
 # Supplemental Information
-
-```{r plot-prep}
-# Define ggplot2 theme for maps
-map_theme <- theme(
-    panel.grid.major = element_line(color = "grey95", size = 0.1),
-    panel.background = element_rect(fill = "grey60"),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
-  )
-
-map_theme_light <- theme(
-    panel.grid.major = element_line(color = "white", size = 0.1),
-    panel.background = element_rect(fill = "grey90"),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
-  )
-```
 
@@
 # FIXME: cells with zero species should be NA, not zero PE
-b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_rank)) +
+b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_p_upper)) +
   geom_sf(color = "transparent") +
-  labs(fill = "Rank PE") +
+  labs(fill = "PE p-score") +
   scale_fill_scico(palette = "lajolla") +
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light +
+  map_theme_light() +
   theme(
@@
   scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme +
+  map_theme() +
   theme(axis.text.y = element_blank())

### 2020-07-27 presentations/bsa_2020-07-27/index.Rmd
Update slides
- Add Gotelli 2000 ref
- Add data source for conservation areas

@@
 
-Null distribution: 999 communities randomized using "Independent Swap" (Gotelli, 2000)
+Null distribution: 999 communities randomized using "Independent Swap" `r AutoCite(bib, "Gotelli2000")`
 
@@
     axis.ticks = element_blank()
+  ) + 
+  labs(
+    caption = "Data: Japan Ministry of the Environment\nhttps://www.biodic.go.jp/biodiversity/activity/policy/map/map17/"
   )

### 2020-08-17 ms/SI.Rmd
Change SES to randomization test
categorize endemism and rand-test significance after joining rand test results with spatial data

@@
   "occ_point_data_ferns",
-  "ses_div_ferns_spatial",
-  "ses_div_ferns_endemic_spatial",
+  "biodiv_ferns_spatial",
+  "biodiv_ferns_endemic_spatial",
   "k_taxonomy",

### 2020-08-17 ms/manuscript.Rmd
Change SES to randomization test
categorize endemism and rand-test significance after joining rand test results with spatial data

@@
   "occ_point_data_ferns_unfiltered",
-  "ses_div_ferns_spatial",
+  "biodiv_ferns_spatial",
   "k_taxonomy",
@@
 
-We measured phylogenetic diversity using the mean phylogenetic distance between all co-occurring taxa in a community (MPD) [@Webb2000; @Webb2002].
-We calculated distances from the phylogenetic tree using the \function{cophenetic} function in R v.
-4.0.0 [@RCoreTeam2019] and MPD using the \function{mpd} function in the R \package{picante}.
+We measured phylogenetic diversity (PD) using the Faith's PD (the total branch length connecting all terminal taxa in a community, including the root of the phylogeny) [@Faith1992] with the \function{PD} function in the R package \package{phyloregion}.
+To detect areas with unusually long or short branch lengths, we measured relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been set to equal length [@Mishler2014].
 
-For morphological diversity, we first calculated morphological distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011].
+We measured functional diversity (FD) following the method of @Petchey2002, which is a functional analog of PD.
+We first calculated distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011].
 We weighted quantitative and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled quantitative traits prior to calculating distances.
-We then used the morphological distances as input into the \function{mpd} function to calculate mean morphological distance (MMD).
+We then used the distances to build a trait dendrogram using the \function{hclust} function in the \package{stats} R package [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
 
@@
 
-Observed values of MPD, MMD, and PE are known to be closely related to species richness, and tend to saturate at high values of richness.
+Observed values of PD, FD, and PE are known to be closely related to species richness, and tend to saturate at high values of richness.
 Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of species by calculating their standard effect sizes (SES), defined as the difference between the observed value and the mean of a pool of random values, divided by the standard deviation of the random values.
@@
 As the optimal number of bioregions cannot be known *a-priori*, we tested values of *k* from 1 to 20.
-\package{phyloregion} provides a function to automatically select *k* from a range of test values, but we found the results to be unreliable.
-We therefore evaluated the explained variance at each value of *k*, then selected *k* that provided the greatest increase in variance relative to the previous value (`r s_figure("k-plot")`).
 

### 2020-08-17 ms/manuscript.Rmd
Update MS
- use randomization test instead of SES

@@
 The final cleaned occurrence dataset included `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` specimens (`r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` taxa).
+Hereafter, "taxa" refers to species, subspecies, and varieties unless otherwise indicated.
 
@@
 We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
-Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens.
+Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens [@Ebihara2016b; @Ebihara2017].
 All qualitative traits in binary format; e.g., a single trait with three states is formatted as three binary traits.
-Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately.
-Quantitative traits include aspects of leaf size (e.g., frond length, frond width, stipe length, etc.).
-Qualitative traits include aspects of leaf morphology (e.g., margin shape, leaf color, indusia shape, spore arrangement, etc.).
+Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately originally, then averaged.
+From the original trait list, we selected only traits with putatative ecological function, and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state (`r table("traits-used")`).
 
@@
 
-Observed values of PD, FD, and PE are known to be closely related to species richness, and tend to saturate at high values of richness.
-Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of species by calculating their standard effect sizes (SES), defined as the difference between the observed value and the mean of a pool of random values, divided by the standard deviation of the random values.
-We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then calculated SES of each metric.
+Observed values of PD, FD, and PE are known to be closely related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others.
+Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test.
+We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values.
+Statistical significance was assessed using a two-tailed test with alpha = 0.5: observed values > 97.5% or < 0.25% were considered significant.
 
@@
 
-Species richness is highest in southern Kyushu and on Yakushima Island, and lowest on Hokkaido and low-lying coastal areas (`r figure("diversity")`A).
+Species richness is highest in southern Kyushu and on Yakushima Island, and lowest on Hokkaido and low-lying coastal areas (`r figure("raw-diversity")`A).
 Sampling redundancy is generally high throughout the country, with a peak just north of Tokyo, but lower in Hokkaido and some coastal areas and small islands (`r figure("diversity")`B).
-Raw MPD is mostly uniform throughout (`r figure("diversity")`C).
-SES of MPD shows clustering in some parts of central Honshu, with overdispersion both in the northern half of the country and from southern Kyushu to the south (`r figure("diversity")`D).
-Morphological diversity shows a similar pattern to phylogenetic diversity, but with more extreme overdispersion and clustering (`r figure("diversity")`F).
+Raw MPD is mostly uniform throughout (`r figure("rand-diversity")`C).
+SES of MPD shows clustering in some parts of central Honshu, with overdispersion both in the northern half of the country and from southern Kyushu to the south (`r figure("rand-diversity")`D).
+Morphological diversity shows a similar pattern to phylogenetic diversity, but with more extreme overdispersion and clustering (`r figure("rand-diversity")`F).
 
@@
 # Change cluster to factor
-ses_div_ferns_spatial <- ses_div_ferns_spatial %>%
+biodiv_ferns_spatial <- biodiv_ferns_spatial %>%
   mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
   mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
+
+# Download high-res map of Japan,
+# crop to spatial div results area
+japan <- rnaturalearth::ne_countries(country = "japan", scale = "large", returnclass = "sf") %>%
+  sf::st_crop(sf::st_bbox(biodiv_ferns_spatial))
+
+# Set CRS
+japan_crs <- sf::st_crs(japan)
+
+biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
+
+# Set color palettes
+# Randomization test significance
+signif_cols <-
+  c(
+    "> 0.99" = "#e31a1c", 
+    "> 0.975" = "#fb9a99",
+    "< 0.01" = "#1f78b4", 
+    "< 0.025" = "#a6cee3",
+    "not significant" = "#ffffcc"
+  )
+
+# CANAPE
+canape_cols <-
+  c(
+    "neo" = "#e31a1c", 
+    "paleo" = "#1f78b4",
+    "mixed" = "#6a51a3",
+    "super" = "#4a1486",
+    "not significant" = "#ffffcc"
+  )
 ```
 
-```{r richness, width = 7, height = 4}
+```{r raw-diversity, width = 7, height = 4}
 # Richness
-a <- ggplot(ses_div_ferns_spatial, aes(fill = richness)) +
-  geom_sf(color = "transparent") +
+a <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  labs(fill = "No. taxa") +
   scale_fill_scico(palette = "lajolla") + 
-  labs(fill = "Richness") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  scale_x_continuous(breaks = c(125, 135, 145)) +
+  scale_y_continuous(breaks = c(25, 35, 45)) +
   map_theme()
 
-# Sampling rendundancy
-b <- ggplot(ses_div_ferns_spatial, aes(fill = redundancy)) +
-  geom_sf(color = "transparent") +
+# Raw PD
+b <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  labs(fill = "Relative\nbranch\nlength") +
   scale_fill_scico(palette = "lajolla") + 
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  labs(fill = "Redundancy") +
+  scale_x_continuous(breaks = c(125, 135, 145)) +
+  scale_y_continuous(breaks = c(25, 35, 45)) +
+  map_theme() +
+  theme(
+    axis.text.y = element_blank()
+  )
+
+# Raw FD
+c <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  labs(fill = "Relative\nbranch\nlength") +
+  scale_fill_scico(palette = "lajolla") + 
+  scale_x_continuous(breaks = c(125, 135, 145)) +
+  scale_y_continuous(breaks = c(25, 35, 45)) +
+  map_theme() +
+  theme(
+    axis.text.y = element_blank()
+  )
+
+a + b + c +
+  plot_layout(nrow = 1)
+
+ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7, height = 4, units = "in")
+```
+
+**`r figure("raw-diversity")`**.
+`r figure_cap("raw-diversity")` For **B** and **C**, raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+
+```{r rand-diversity, fig.width = 7, fig.height = 9}
+# PD
+a <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  scale_fill_manual(values = signif_cols) +
+  map_theme() +
+  theme(
+    axis.text.x = element_blank(),
+    legend.title = element_blank()
+    )
+
+# RPD
+b <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  labs(fill = "") +
+  scale_fill_manual(values = signif_cols) +
+  map_theme() +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    legend.title = element_blank()
+  )
+
+# FD
+c <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  labs(fill = "") +
+  scale_fill_manual(values = signif_cols) +
+  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
+  map_theme() +
+  theme(legend.title = element_blank())
+
+# RFD
+d <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**D** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  labs(fill = "") +
+  scale_fill_manual(values = signif_cols) +
+  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
+  map_theme() +
+  theme(
+    axis.text.y = element_blank(),
+    legend.title = element_blank()
+  )
+
+a + b + c + d + 
+  plot_layout(ncol = 2, nrow = 2, guides = 'collect') &
+  theme(legend.position = "bottom")
+
+ggsave(result_file(figure("rand-diversity"), "pdf"), width = 7, height = 8, units = "in")
+```
+
+**`r figure("rand-diversity")`**.
+`r figure_cap("rand-diversity")` **A**, Phylogenetic diversity (PD).
+**B**, Relative phylogenetic diveristy (RPD).
+**C**, Functional diveristy (FD).
+**D**, relative functional diveristy (RFD).
+For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
+For details, see Methods.
+
+\clearpage
+
+```{r endemism, fig.width = 5, fig.height = 5}
+ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
+  scale_fill_manual(values = canape_cols) +
+  map_theme() +
+  theme(legend.title = element_blank())
+
+ggsave(result_file(figure("endemism"), "pdf"), width = 5, height = 5, units = "in")
+```
+
+**`r figure("endemism")`**.
+`r figure_cap("endemism")` 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+For details, see Methods.
+
+\clearpage
+
+```{r bioregions, fig.width = 7, fig.height = 5}
+# Taxonomic bioregions
+a <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  labs(fill = "Cluster") +
+  scale_fill_carto_d(type = "qualitative", palette = "Bold") +
+  map_theme()
+
+# Phylogenetic bioregions
+b <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = phylo_cluster), color = "transparent", alpha = 0.8) +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  labs(fill = "Cluster") +
+  scale_fill_carto_d(type = "qualitative", palette = "Bold") +
   map_theme() +
@@
 a + b +
-  plot_layout(ncol = 2) +
-  plot_annotation(tag_levels = 'A')
+  plot_layout(ncol = 2)
 
-ggsave(result_file(figure("richness"), "pdf"), width = 7, height = 4, units = "in")
-```
-
-**`r figure("richness")`**. `r figure_cap("diversity")` **A**, richness. **B**, sampling redundancy.
-
-```{r diversity, fig.width = 7, fig.height = 9}
-# PD
-a <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "PD") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  scale_fill_scico(palette = "lajolla") + 
-  map_theme() +
-  theme(
-    axis.text.x = element_blank()
-  )
-
-# SES of PD
-b <- ggplot(ses_div_ferns_spatial, aes(fill = pd_obs_z)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "SES PD") +
-  scale_fill_scico(
-    palette = "romaO",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_spatial, pd_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, pd_obs_z, "abs")
-    )) +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank()
-  )
-
-# RPD
-c <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "RPD") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  scale_fill_scico(palette = "lajolla") + 
-  map_theme() +
-  theme(
-    axis.text.x = element_blank()
-  )
-
-# SES of RPD
-d <- ggplot(ses_div_ferns_spatial, aes(fill = rpd_obs_z)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "SES RPD") +
-  scale_fill_scico(
-    palette = "roma",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, rpd_obs_z, "abs")
-    )) +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  # highlight only significant cells
-  # gghighlight(pd_obs_p > 0.975 | pd_obs_p < 0.025) +
-  map_theme() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank()
-  )
-
-# FD
-e <- ggplot(ses_div_ferns_spatial, aes(fill = fd_obs)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "FD") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  scale_fill_scico(palette = "lajolla") + 
-  map_theme() +
-  theme(
-    axis.text.x = element_blank()
-  )
-
-# SES of FD
-f <- ggplot(ses_div_ferns_spatial, aes(fill = fd_obs_z)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "SES FD") +
-  scale_fill_scico(
-    palette = "roma",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_spatial, fd_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, fd_obs_z, "abs")
-    )) +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  # highlight only significant cells
-  # gghighlight(mpd_morph_obs_p > 0.975 | mpd_morph_obs_p < 0.025) +
-  map_theme() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank()
-  )
-
-# RFD
-g <- ggplot(ses_div_ferns_spatial, aes(fill = rfd_obs)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "RFD") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  scale_fill_scico(palette = "lajolla") + 
-  map_theme()
-
-# SES of RFD
-h <- ggplot(ses_div_ferns_spatial, aes(fill = rfd_obs_z)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "SES RFD") +
-  scale_fill_scico(
-    palette = "roma",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, rfd_obs_z, "abs")
-    )) +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  # highlight only significant cells
-  # gghighlight(rfd_obs_p > 0.975 | rfd_obs_p < 0.025) +
-  map_theme() +
-  theme(
-    axis.text.y = element_blank()
-  )
-
-a + b + c + d + e + f + g + h +
-  plot_layout(ncol = 2) +
-  plot_annotation(tag_levels = 'A')
-
-ggsave(result_file(figure("diversity"), "pdf"), width = 7, height = 9, units = "in")
-```
-
-**`r figure("diversity")`**.
-`r figure_cap("diversity")` **A**, Phylogenetic diversity (PD).
-**B**, Standard effect size (SES) of PD.
-**C**, Relative phylogenetic diversity (RPD).
-**D**, SES of RPD.
-**E**, Functional diversity (FD).
-**F**, SES of FD **G**, Relative functional diversity (RFD) **H**, SES of RFD.
-
-\clearpage
-
-```{r endemism, fig.width = 7, fig.height = 8}
-# PE
-a <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs)) +
-  geom_sf(color = "transparent") +
-  scale_fill_scico(palette = "lajolla") + 
-  labs(fill = "PE") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme() +
-  theme(
-    axis.text.x = element_blank()
-  )
-
-# PE p-value
-b <- ggplot(ses_div_ferns_spatial, aes(fill = pe_obs_p_upper)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "PE p-score") +
-  scale_fill_scico(palette = "lajolla") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank()
-  )
-
-# RPE
-c <- ggplot(ses_div_ferns_spatial, aes(fill = rpe_obs)) +
-  geom_sf(color = "transparent") +
-  scale_fill_scico(palette = "lajolla") +
-  labs(fill = "RPE") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme() +
-  theme(
-    axis.text.x = element_blank()
-  )
-
-# SES RPE
-d <- ggplot(ses_div_ferns_spatial, aes(fill = rpe_obs_z)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "SES RPE") +
-  scale_fill_scico(
-    palette = "roma",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_spatial, rpe_obs_z, "abs"),
-      get_limit(ses_div_ferns_spatial, rpe_obs_z, "abs")
-    )) +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme() +
-  theme(
-    axis.text.y = element_blank()
-  )
-
-# CANAPE, full dataset
-# set color scheme
-canape_cols <- c(
-  "neo" = "#E41A1C", # Color brewer set1 red
-  "paleo" = "#377EB8", # Color brewer blue
-  "mixed" = "#984EA3", # Color brewer purple
-  "none" = "#FFFFCC" # Scico lajolla faintest color
-)
-
-e <- 
-  ses_div_ferns_spatial %>%
-  mutate(endem_type = replace_na(endem_type, "none")) %>%
-  ggplot(aes(fill = endem_type)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "Endemism\ntype") +
-  scale_fill_manual(values = canape_cols, breaks = c("neo", "paleo", "mixed", "none"), labels = c("Neo", "Paleo", "Mixed", "Not endemic")) + 
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light()
-
-a + b + c + d + e +
-  plot_layout(ncol = 2) +
-  plot_annotation(tag_levels = 'A')
-
-ggsave(result_file(figure("endemism"), "pdf"), width = 7, height = 8, units = "in")
-```
-
-**`r figure("endemism")`**.
-`r figure_cap("endemism")` **A**, phylogenetic endemism (PE).
-**B**, Standard effect size (SES) of PE.
-**C**, relative phylogenetic endemism (RPE).
-**D**, SES of RPE.
-**E**, Categorical analysis of neo- and paleo-endemism (CANAPE).
-
-\clearpage
-
-```{r bioregions, fig.width = 7, fig.height = 3}
-# Taxonomic bioregions
-a <- ggplot(ses_div_ferns_spatial, aes(fill = taxonomic_cluster)) +
-  geom_sf(color = "transparent") +
-  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light()
-
-# Phylogenetic bioregions
-b <- ggplot(ses_div_ferns_spatial, aes(fill = phylo_cluster)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "Cluster") +
-  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light()
-
-a + b +
-  plot_layout(ncol = 2) +
-  plot_annotation(tag_levels = 'A')
-
-ggsave(result_file(figure("bioregions"), "pdf"), width = 7, height = 3, units = "in")
+ggsave(result_file(figure("bioregions"), "pdf"), width = 7, height = 5, units = "in")
 ```
@@
 data <-
-ses_div_ferns_spatial %>%
+biodiv_ferns_spatial %>%
   filter(taxonomic_cluster %in% c(1,2,3,4)) %>%
@@
     panel.grid.minor.y = element_blank(),
-
   )

### 2020-08-17 R/functions.R

@@ -2380,10 +2380,13 @@ standard_theme2 <- function () {
 
 map_theme <- function() {
   theme(
-    panel.grid.major = element_line(color = "grey95", size = 0.1),
-    panel.background = element_rect(fill = "grey60"),
+    panel.grid.major = element_line(color = "grey60", size = 0.1),
+    panel.grid.minor = element_blank(),
+    panel.background = element_rect(fill = "transparent"),
     axis.ticks = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.7)
+    axis.text = element_text(color = "grey60"),
+    axis.title = element_blank(),
+    legend.position = "bottom"
   )
 }
 

### 2020-08-17 ms/captions.R

@@ -2,9 +2,9 @@
 
 # - Figures
 fig_nums <- captioner::captioner(prefix = "Fig.")
-fig_nums(name = "richness", caption = "Taxonomic diversity of the ferns of Japan.")
-fig_nums(name = "diversity", caption = "Phylogenetic and morphological diversity of the ferns of Japan.")
-fig_nums(name = "endemism", caption = "Phylogenetic endemism of the ferns of Japan.")
+fig_nums(name = "raw-diversity", caption = "Raw taxonomic (**A**), phylogenetic (**B**), and functional (**C**) diversity of the ferns of Japan.")
+fig_nums(name = "rand-diversity", caption = "Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.")
+fig_nums(name = "endemism", caption = "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).")
 fig_nums(name = "bioregions", caption = "Bioregions of the ferns of Japan.")
 fig_nums(name = "div-by-tax-region", caption = "Phylogenetic and morphological diversity of the ferns of Japan by bioregion.")
 
@@ -17,7 +17,7 @@ s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix
 s_fig_nums(name = "sampling-curve", "Species collection curve for the ferns of Japan fit with the iNEXT package.")
 s_fig_nums(name = "grain-size", "Effect of grain size on sampling redundancy.")
 s_fig_nums(name = "k-plot", "Selection of *k* for bioregions.")
-s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan, restricted dataset including only taxa endemic to Japan.")
+s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.")
 s_fig_nums(name = "raw-biplot", "Relationships between raw biodiversity metrics of the ferns of Japan.")
 s_fig_nums(name = "ses-biplot", "Relationships between standard effect size (SES) of biodiversity metrics of the ferns of Japan.")
 s_fig_nums(name = "traits-nmds", "Trait space of the ferns of Japan.")

### 2020-08-19 ms/manuscript.Rmd
Update MS

@@
 Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens [@Ebihara2016b; @Ebihara2017].
-All qualitative traits in binary format; e.g., a single trait with three states is formatted as three binary traits.
+All qualitative traits were converted to binary format; \eg, a single trait with three states was formatted as three binary traits.
 Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately originally, then averaged.
@@
 
-### Analysis of biodiversity and endemism
+### Phylogenetic signal
+
+We tested for phylogenetic signal in continuous traits using two alternative metrics, Blomberg's *K* [@Blomberg2003] and Pagel's lambda [@Pagel1999], as implemented with the \function{phylosig} function in the R package \package{phytools} [@Revell2012].
+*K* describes the correlation between the observed variance in a trait vs.
+the amount of variance expected under Brownian Motion (BM); *K* = 1 under BM, with values < 1  We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the \function{phylo.d} function in the R package \package{caper} [@Orme2018].
+
+### Analysis of biodiversity
 
@@
 
+We compared patterns in biodiversity between bioregions by calculating the standard effect size (SES) of biodiversity metrics.
+SES measures how extreme the observed value (*obs*) is relative to the null distribution (*null*):
+
+$SES = \frac{obs - mean(null)}{sd(null)}$
+
+We compared SES of PD, RPD, FD, and RFD across bioregions using FIXME: add this when figure out spatial-corrected statistical method to use.
+
+### Overlap with protected zones
+
+We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html).
+We categorized protected areas as either "high" (plants protected; no human activities allowed at all) or "normal" (plants protected; some human activities allowed by permit).
+We cropped the grid cells to only include land area, then calculated the amount of protected area of each status type within grid cells with significantly high PD, FD, or PE (one-tailed test).
+
 ### Data accessibility
@@
 
-### Taxonomic, phylogenetic, and morphological diversity
+### Number of specimens and redundancy
 
-Species richness is highest in southern Kyushu and on Yakushima Island, and lowest on Hokkaido and low-lying coastal areas (`r figure("raw-diversity")`A).
-Sampling redundancy is generally high throughout the country, with a peak just north of Tokyo, but lower in Hokkaido and some coastal areas and small islands (`r figure("diversity")`B).
-Raw MPD is mostly uniform throughout (`r figure("rand-diversity")`C).
-SES of MPD shows clustering in some parts of central Honshu, with overdispersion both in the northern half of the country and from southern Kyushu to the south (`r figure("rand-diversity")`D).
-Morphological diversity shows a similar pattern to phylogenetic diversity, but with more extreme overdispersion and clustering (`r figure("rand-diversity")`F).
+The distribution of specimen collection locations is highly skewed, with >3,000 specimens per grid cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r figure("abundance")`A).
+Despite this, sampling redundancy is generally high throughout the country, with areas of low redundancy in Hokkaido and some coastal areas and small islands (`r figure("abundance")`B).
+The grid cells that had to be excluded from the analysis due to low redundancy were mostly from Hokkaido.
 
-### Phylogenetic endemism
+### Phylogenetic signal
 
-Phylogenetic endemism (PE) is extremeley low throughout except for in some of the southern islands (`r figure("endemism")`A).
-SES of PE is high starting in southern Kyushu and southwards, including the Ogasawara islands (`r figure("endemism")`B).
-Results of CANAPE (Categorical Analysis of Palaeo and Neo Endemism) indicate hotspots of mixed endemism in the southern islands, but no neo- and very little paleo-endemism (`r figure("endemism")`C).
-When we analyzed patterns of endemicity with a reduced dataset including only taxa endemic to Japan to account for the border effect, patterns were generally similar to the full dataset, indicating that the southern islands are genuine hotspots of endemicity (`r s_figure("endemism-restricted")`).
+Results of phylogenetic signal analysis are given in Table _.
+All quantitative traits showed significant phylogenetic signal, but estimates of signal strength differed between K and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of K were < 0.06, indicating less phylogenetic signal than expected under BM.
+A minority of quantitative traits (coded as binary) showed as much or more phylogenetic signal as expected under BM (D =< 0); others had D > 0.
+
+### Observed diversity
+
+Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`A).
+Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`A, C).
+Observed values of RPD and RFD were also closely related to richness, showing high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).
+
+### Randomization tests for phylogenetic and functional diversity
+
+Grid cells with significantly low PD were almost all located on Honshu, the main island of Japan.
+A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Amami, Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula).
+The area near Yakushima includes cells with both significantly high and low PD.
+Grid cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu.
+Grid cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido.
+Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD.
+
+### Categorical analysis of neo- and paleo-endemism
+
+Nearly all grid cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere.
+The vast majority of endemic grid cells are mixed, with very few paleo- or neoendemic cells.
+Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (Daisetsuzan mountains).
+When we ran CANAPE with a reduced dataset including only taxa endemic to Japan to account for the border effect, there were fewer significant cells, but patterns were otherwise similar to the full dataset (`r s_figure("endemism-restricted")`).
 
@@
 
-### Relationships between metrics
-
-Raw values of MPD and MMD showed high variance and low species richness, then merged towards a single value at high species richness, as expected (`r s_figure("raw-biplot")`K, O).
-Raw PE was linearly correlated with species richness (`r s_figure("raw-biplot")`M).
-Both raw and SES values of MPD and MMD were highly correlated (`r s_figure("raw-biplot")`D, `r s_figure("ses-biplot")`D).
-SES values of MPD, MMD, and PE were all positively correlated (`r s_figure("ses-biplot")`).
-None of the SES values were correlated with richness (`r s_figure("ses-biplot")`).
+### Degree of protection
 
@@
   mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor)
+  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  # Add redundancy
+  mutate(redundancy = 1 - (richness/abundance))
 
@@
 ```
+
+```{r abundance, width = 7, height = 5}
+k <- scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1, sep = "")
+
+a <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = abundance), color = "transparent", alpha = 0.8) +
+  scale_fill_scico(palette = "lajolla", label = k) +
+  labs(fill = "No. specimens") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A**</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  map_theme()
+
+b <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = redundancy), color = "transparent", alpha = 0.8) +
+  scale_fill_scico(palette = "lajolla") +
+  labs(fill = "Redundancy") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B**</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  map_theme() +
+  theme(
+    axis.text.y = element_blank()
+  )
+
+a + b + 
+  plot_layout(nrow = 1)
+
+ggsave(result_file(figure("abundance"), "pdf"), width = 7, height = 5, units = "in")
+```
+
+**`r figure("abundance")`**. `r figure_cap("abundance")`
+
+\clearpage
 
@@
 `r figure_cap("raw-diversity")` For **B** and **C**, raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+
+\clearpage
 
@@
   scale_fill_carto_d(type = "qualitative", palette = "Bold") +
+  # scale_fill_brewer(type = "qual") +
   map_theme()
@@
 **`r figure("div-by-tax-region")`**.
-`r figure_cap("div-by-tax-region")`.
-**A**, Standard effect size (SES) of phylogenetic diversity (PD).
+`r figure_cap("div-by-tax-region")` **A**, Standard effect size (SES) of phylogenetic diversity (PD).
 **B**, SES of relative PD (RPD).

### 2020-08-19 ms/captions.R

@@ -2,6 +2,7 @@
 
 # - Figures
 fig_nums <- captioner::captioner(prefix = "Fig.")
+fig_nums(name = "abundance", caption = "Observed number of specimens (**A**) and sampling redundancy (**B**) per 0.2° grid cell in the ferns of Japan.")
 fig_nums(name = "raw-diversity", caption = "Raw taxonomic (**A**), phylogenetic (**B**), and functional (**C**) diversity of the ferns of Japan.")
 fig_nums(name = "rand-diversity", caption = "Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.")
 fig_nums(name = "endemism", caption = "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).")
@@ -17,9 +18,8 @@ s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix
 s_fig_nums(name = "sampling-curve", "Species collection curve for the ferns of Japan fit with the iNEXT package.")
 s_fig_nums(name = "grain-size", "Effect of grain size on sampling redundancy.")
 s_fig_nums(name = "k-plot", "Selection of *k* for bioregions.")
+s_fig_nums(name = "raw-biplot", "Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.")
 s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.")
-s_fig_nums(name = "raw-biplot", "Relationships between raw biodiversity metrics of the ferns of Japan.")
-s_fig_nums(name = "ses-biplot", "Relationships between standard effect size (SES) of biodiversity metrics of the ferns of Japan.")
 s_fig_nums(name = "traits-nmds", "Trait space of the ferns of Japan.")
 
 # - SI tables

### 2020-08-19 ms/SI.Rmd
Update SI
- remove bi plots for raw vs SES biodiv metrics

@@
 
-```{r endemism-restricted, fig.width = 7, fig.height = 8}
-# PE (endemic)
-# FIXME: cells with zero species should be NA, not zero PE
-a <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs)) +
-  geom_sf(color = "transparent") +
-  scale_fill_scico(palette = "lajolla") + 
-  labs(fill = "PE") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme() +
-  theme(
-    axis.text.x = element_blank()
+```{r raw-biplot, fig.width = 7, fig.height = 7}
+a <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = pd_obs)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Richness", y = "PD")
+
+b <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = rpd_obs)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Richness", y = "RPD")
+
+c <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = fd_obs)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Richness", y = "FD")
+
+d <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = rfd_obs)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Richness", y = "RFD")
+
+a + b + c + d +
+  plot_layout(ncol = 2, nrow = 2) +
+  plot_annotation(tag_levels = 'A')
+
+ggsave(result_file(s_figure("raw-biplot"), "pdf"), width = 7, height = 3, units = "in")
+```
+
+**`r s_figure("raw-biplot")`**.
+`r s_figure_cap("raw-biplot")` **A**, Phylogenetic diversity (PD).
+**B**, Relative phylogenetic diversity (RPD).
+**C**, Functional diversity (FD).
+**D**, Relative functional diversity (RFD).
+
+\clearpage
+
+```{r map-plot-prep}
+# Download high-res map of Japan,
+# crop to spatial div results area
+japan <- rnaturalearth::ne_countries(country = "japan", scale = "large", returnclass = "sf") %>%
+  sf::st_crop(sf::st_bbox(biodiv_ferns_endemic_spatial))
+
+# Set CRS
+japan_crs <- sf::st_crs(japan)
+
+biodiv_ferns_endemic_spatial <- sf::st_set_crs(biodiv_ferns_endemic_spatial, japan_crs)
+
+# Set color palettes
+# Randomization test significance
+signif_cols <-
+  c(
+    "> 0.99" = "#e31a1c", 
+    "> 0.975" = "#fb9a99",
+    "< 0.01" = "#1f78b4", 
+    "< 0.025" = "#a6cee3",
+    "not significant" = "#ffffcc"
   )
 
-# SES PE (endemic)
-# FIXME: cells with zero species should be NA, not zero PE
-b <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = pe_obs_p_upper)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "PE p-score") +
-  scale_fill_scico(palette = "lajolla") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+# CANAPE
+canape_cols <-
+  c(
+    "neo" = "#e31a1c", 
+    "paleo" = "#1f78b4",
+    "mixed" = "#6a51a3",
+    "super" = "#4a1486",
+    "not significant" = "#ffffcc"
+  )
+```
+
+```{r endemism-restricted, fig.width = 5, fig.height = 5}
+ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
+  scale_fill_manual(values = canape_cols) +
   map_theme() +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank()
-  )
+  theme(legend.title = element_blank())
 
-# RPE
-c <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs)) +
-  geom_sf(color = "transparent") +
-  scale_fill_scico(palette = "lajolla") + 
-  labs(fill = "RPE") +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light() +
-  theme(
-    axis.text.x = element_blank()
-  )
-
-# SES RPE
-d <- ggplot(ses_div_ferns_endemic_spatial, aes(fill = rpe_obs_z)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "SES RPE") +
-  scale_fill_scico(
-    palette = "roma",
-    na.value="grey80",
-    direction = -1,
-    limits = c(
-      -get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs"),
-      get_limit(ses_div_ferns_endemic_spatial, rpe_obs_z, "abs")
-    )) +
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme() +
-  theme(axis.text.y = element_blank())
-
-# CANAPE, restricted dataset
-# set color scheme
-canape_cols <- c(
-  "neo" = "#E41A1C", # Color brewer set1 red
-  "paleo" = "#377EB8", # Color brewer blue
-  "mixed" = "#984EA3", # Color brewer purple
-  "none" = "#FFFFCC" # Scico lajolla faintest color
-)
-
-e <- ses_div_ferns_endemic_spatial %>%
-  mutate(endem_type = replace_na(endem_type, "none")) %>%
-  ggplot(aes(fill = endem_type)) +
-  geom_sf(color = "transparent") +
-  labs(fill = "Endemism\ntype") +
-  scale_fill_manual(values = canape_cols, breaks = c("neo", "paleo", "mixed", "none"), labels = c("Neo", "Paleo", "Mixed", "Not endemic")) + 
-  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
-  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
-  map_theme_light()
-
-a + b + c + d + e +
-  plot_layout(ncol = 2) +
-  plot_annotation(tag_levels = 'A')
-
-ggsave(result_file(s_figure("endemism-restricted"), "pdf"), width = 7, height = 8, units = "in")
+ggsave(result_file(s_figure("endemism-restricted"), "pdf"), width = 5, height = 5, units = "in")
 ```
@@
 **`r s_figure("endemism-restricted")`**.
-`r s_figure_cap("endemism-restricted")` **A**, phylogenetic endemism (PE).
-**B**, Standard effect size (SES) of PE.
-**C**, relative phylogenetic endemism (RPE).
-**D**, SES of RPE.
-**E**, Categorical analysis of neo- and paleo-endemism (CANAPE).
-
-\clearpage
-
-```{r biplot-prep}
-# Extract centroids
-centroids <- sf::st_centroid(ses_div_ferns_spatial$geometry) %>%
-  sf::st_coordinates()
-
-# Define helper function for making a scatterplot
-scatterplot <- function(data, x_var, y_var, x_lab, y_lab, label, annotate, loc) {
-  p <- ggplot(data, aes(x = .data[[x_var]], y = .data[[y_var]])) +
-    geom_point(alpha = 0.5) +
-    labs(
-      x = x_lab,
-      y = y_lab
-    )
-
-  # Add annotation and trendline if annotate = TRUE
-  if(annotate & loc == "UL") p <- p +
-      annotate("label", x = -Inf, y = Inf, label = label, hjust = 0, vjust = 1, label.size = 0, size = 3, fill = "transparent") +
-      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
-
-  if(annotate & loc == "UR") p <- p +
-      annotate("label", x = Inf, y = Inf, label = label, hjust = 1, vjust = 1, label.size = 0, size = 3, fill = "transparent") +
-      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
-
-  if(annotate & loc == "LL") p <- p +
-      annotate("label", x = -Inf, y = -Inf, label = label, hjust = 0, vjust = 0, label.size = 0, size = 3, fill = "transparent") +
-      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
-
-  if(annotate & loc == "LR") p <- p +
-      annotate("label", x = Inf, y = -Inf, label = label, hjust = 1, vjust = 0, label.size = 0, size = 3, fill = "transparent") +
-      geom_smooth(method = "lm", se = FALSE, color = "blue", size = 0.5)
-
-  p
-}
-```
-
-```{r raw-biplot, fig.width = 8, fig.height = 10.5}
-## Biplots of raw metrics ##
-
-# Set up x and y variables as factors so they can be ordered appropriately in the plot
-x_vars <-
-  c("pd_obs", "rpd_obs", "fd_obs", "rfd_obs", "pe_obs", "rpe_obs", "richness") %>%
-  factor() %>%
-  fct_relevel(as.character(.))
-
-y_vars <- c("pd_obs", "rpd_obs", "fd_obs", "rfd_obs", "pe_obs", "rpe_obs") %>%
-  factor() %>%
-  fct_relevel(as.character(.))
-
-# Make dataframe of arguments to feed into plotting function
-# - includes all non-duplicated combinations of raw diversity metrics
-raw_comparisons <-
-  cross_df(
-    list(
-      x_var = x_vars,
-      y_var = y_vars
-    ),
-    .filter = `==` # filter out exact duplicates
-  ) %>%
-  # Remove reversed duplicates
-  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
-  group_by(grp = paste(pmax(as.character(x_var), as.character(y_var)), pmin(as.character(x_var), as.character(y_var)), sep = "_")) %>%
-  arrange(x_var, y_var) %>%
-  slice(1) %>%
-  ungroup() %>%
-  select(-grp) %>%
-  arrange(x_var, y_var) %>%
-  # Convert x_var and y_var back to character for plotting function
-  mutate(across(c(x_var, y_var), as.character)) %>%
-  # Add labels
-  mutate(
-    x_lab = case_when(
-      x_var == "pd_obs" ~ "PD",
-      x_var == "rpd_obs" ~ "RPD",
-      x_var == "fd_obs" ~ "FD",
-      x_var == "rfd_obs" ~ "RFD",
-      x_var == "pe_obs" ~ "PE",
-      x_var == "rpe_obs" ~ "RPE",
-      x_var == "richness" ~ "Richness"
-    ),
-    y_lab = case_when(
-      y_var == "pd_obs" ~ "PD",
-      y_var == "rpd_obs" ~ "RPD",
-      y_var == "fd_obs" ~ "FD",
-      y_var == "rfd_obs" ~ "RFD",
-      y_var == "pe_obs" ~ "PE",
-      y_var == "rpe_obs" ~ "RPE"
-    )
-  ) %>%
-  # Run modified t-test
-  mutate(
-    test_res = map2(
-      .x = x_var, 
-      .y = y_var, 
-      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
-    # Extract relevant stats from results
-    cor = map_dbl(test_res, pluck("corr")),
-    pval = map_dbl(test_res, pluck("p.value"))
-    ) %>%
-  # Generate label
-  mutate(
-    cor_lab = glue("r = {round(cor, 3)}"),
-    pval_lab = case_when(
-      pval < 0.001 ~ "***",
-      pval < 0.01 ~ "**",
-      pval < 0.05 ~ "*",
-      TRUE ~ ""
-    ),
-    label = glue::glue("{cor_lab}{pval_lab}")
-  ) %>%
-  # Specify which plots to annotate
-  mutate(
-    annotate = case_when(
-      pval < 0.05 ~ TRUE,
-      TRUE ~ FALSE
-    ),
-    # annotate in upper-right corner for mpd_morph_obs + rpe_obs,
-    # upper-left otherwise
-    loc = case_when(
-      y_var == "mpd_morph_obs" & x_var == "rpe_obs" ~ "UR",
-      TRUE ~ "UL"
-    )
-  ) %>%
-  select(x_var, y_var, x_lab, y_lab, label, annotate, loc)
-
-# Make plot
-pmap(raw_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
-wrap_plots() +
-  plot_layout(ncol = 3)
-
-ggsave(result_file(s_figure("raw-biplot"), "pdf"), width = 8, height = 10.5, units = "in")
-```
-
-**`r s_figure("raw-biplot")`**.
-`r s_figure_cap("raw-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PD, phylogenetic diversity; PE, phylogenetic endemism; RPE, relative phylogenetic endemism.
-Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05.
-
-\clearpage
-
-```{r ses-biplot, fig.width = 8, fig.height = 10.5}
-## Biplots of SES metrics ##
-
-# Set up x and y variables as factors so they can be ordered appropriately in the plot
-x_vars <-
-  c("pd_obs_z", "rpd_obs_z", "fd_obs_z", "rfd_obs_z", "pe_obs_z", "rpe_obs_z", "richness") %>%
-  factor() %>%
-  fct_relevel(as.character(.))
-
-y_vars <- c("pd_obs_z", "rpd_obs_z", "fd_obs_z", "rfd_obs_z", "pe_obs_z", "rpe_obs_z") %>%
-  factor() %>%
-  fct_relevel(as.character(.))
-
-# Make dataframe of arguments to feed into plotting function
-# - includes all non-duplicated combinations of raw diversity metrics
-ses_comparisons <-
-  cross_df(
-    list(
-      x_var = x_vars,
-      y_var = y_vars
-    ),
-    .filter = `==` # filter out exact duplicates
-  ) %>%
-  # Remove reversed duplicates
-  # https://stackoverflow.com/questions/22756392/deleting-reversed-duplicates-with-r
-  group_by(grp = paste(pmax(as.character(x_var), as.character(y_var)), pmin(as.character(x_var), as.character(y_var)), sep = "_")) %>%
-  arrange(x_var, y_var) %>%
-  slice(1) %>%
-  ungroup() %>%
-  select(-grp) %>%
-  arrange(x_var, y_var) %>%
-  # Convert x_var and y_var back to character for plotting function
-  mutate(across(c(x_var, y_var), as.character)) %>%
-  # Add labels
-  mutate(
-    x_lab = case_when(
-      x_var == "pd_obs_z" ~ "SES PD",
-      x_var == "rpd_obs_z" ~ "SES RPD",
-      x_var == "fd_obs_z" ~ "SES FD",
-      x_var == "rfd_obs_z" ~ "SES RFD",
-      x_var == "pe_obs_z" ~ "SES PE",
-      x_var == "rpe_obs_z" ~ "SES RPE",
-      x_var == "richness" ~ "Richness"
-    ),
-    y_lab = case_when(
-      y_var == "pd_obs_z" ~ "SES PD",
-      y_var == "rpd_obs_z" ~ "SES RPD",
-      y_var == "fd_obs_z" ~ "SES FD",
-      y_var == "rfd_obs_z" ~ "SES RFD",
-      y_var == "pe_obs_z" ~ "SES PE",
-      y_var == "rpe_obs_z" ~ "SES RPE"
-    )
-  ) %>%
-  # Run modified t-test
-  mutate(
-    test_res = map2(
-      .x = x_var, 
-      .y = y_var, 
-      ~SpatialPack::modified.ttest(ses_div_ferns_spatial[[.x]], ses_div_ferns_spatial[[.y]], centroids)),
-    # Extract relevant stats from results
-    cor = map_dbl(test_res, pluck("corr")),
-    pval = map_dbl(test_res, pluck("p.value"))
-    ) %>%
-  # Generate label
-  mutate(
-    cor_lab = glue("r = {round(cor, 3)}"),
-    pval_lab = case_when(
-      pval < 0.001 ~ "***",
-      pval < 0.01 ~ "**",
-      pval < 0.05 ~ "*",
-      TRUE ~ ""
-    ),
-    label = glue::glue("{cor_lab}{pval_lab}")
-  ) %>%
-  # Specify which plots to annotate
-  mutate(
-    annotate = case_when(
-      pval < 0.05 ~ TRUE,
-      TRUE ~ FALSE
-    ),
-    # annotate in upper-right corner for mpd_morph_obs + rpe_obs,
-    # upper-left otherwise
-    loc = case_when(
-      y_var == "rpe_obs_z" & x_var == "pe_obs_z" ~ "LR",
-      y_var == "fd_obs_z" & x_var == "richness" ~ "UR",
-      y_var == "pd_obs_z" & x_var == "richness" ~ "UR",
-      TRUE ~ "UL"
-    )
-  ) %>%
-  select(x_var, y_var, x_lab, y_lab, label, annotate, loc)
-
-# Make plot
-pmap(ses_comparisons, scatterplot, data = ses_div_ferns_spatial) %>%
-wrap_plots() +
-  plot_layout(ncol = 3)
-
-ggsave(result_file(s_figure("ses-biplot"), "pdf"), width = 8, height = 10.5, units = "in")
-```
-
-**`r s_figure("ses-biplot")`**.
-`r s_figure_cap("ses-biplot")` MPD, mean phylogenetic distance; MMD, mean morphological distance; PD, phylogenetic diversity; PE, phylogenetic endemism; RPE, relative phylogenetic endemism.
-Correlations between variables analyzed with Dutilleul's modified *t*-test that accounts for spatial structure in the data; *p* value and correlation (r) only shown for *p* < 0.05.
+`r s_figure_cap("endemism-restricted")` 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+For details, see Methods.
 

### 2020-08-19 ms/manuscript.Rmd
Update MS

@@
 
-We tested for phylogenetic signal in continuous traits using two alternative metrics, Blomberg's *K* [@Blomberg2003] and Pagel's lambda [@Pagel1999], as implemented with the \function{phylosig} function in the R package \package{phytools} [@Revell2012].
-*K* describes the correlation between the observed variance in a trait vs.
-the amount of variance expected under Brownian Motion (BM); *K* = 1 under BM, with values < 1  We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the \function{phylo.d} function in the R package \package{caper} [@Orme2018].
+We tested for phylogenetic signal in continuous traits using two alternative metrics, Blomberg's *K* [@Blomberg2003] and Pagel's lambda [@Pagel1999].
+*K* describes the ratio between the observed variance in a trait vs.
+the amount of variance expected under Brownian Motion (BM); lambda is a scaling parameter used to transform the phylogeny such that trait values fit those most closely expected under BM.
+Both lambda and *K* = 1 under BM; lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering).
+We measured *K* and lambda using the \function{phylosig} function in the R package \package{phytools} [@Revell2012], which also calculates significance based on a randomization test for *K* and a likelihood ratio test for lambda.
+We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the \function{phylo.d} function in the R package \package{caper} [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
 
@@
 Using phylogenetic distances produced similar results, but Okinawa and the Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B).
+The other much smaller clusters are likely artifacts of insufficiently low taxon richness for accurate clustering in a small number of grid cells, and are not discussed further.
+
+Taxonomic clusters 3 and 4 have higher SES values of PD, RPD, FD, and RFD than clusters 1 and 2.
+Clusters 3 and 4 also have consistently high PE *p*-scores, whereas clusters 1 and 2 have a much wider variation.
+Cluster 2 has lower RPD and RFD than cluster 1.
+Comparison of biodiversity metrics across phylogenetic clusters was similar, except that phylogenetic cluster 3 mostly corresponds to taxonomic clusters 3 and 4 combined.
 
@@
 
-TBD
+Here, we provide the first insights into the combined taxonomic, phylogenetic, and functional diversity of the ferns of Japan.
+Our integrative approach reveals key differences in patterns of biodiversity in different parts of the archipelago.
+
+### Distribution of biodiversity
+
+The latitudinal hump-shaped pattern of species richness observed in ferns here has been well-documented.
+
+Despite the peak of richness in southern Kyushu, the overwhealming pattern driving phylogenetic and functional diversity is the high diversity of southern islands (Ryukyu and Bonin Isl.).
+
+Another major pattern revealed here is the predominance of long branches in the northern half of Honshu, and the predominance of short branches in the southern half, particularly Kyushu.
+
+### Bioregions
+
+### Conservation status
 

### 2020-08-20 ms/manuscript.Rmd
Update MS

@@
   )
+
+# Bioregions
+bioregion_cols <- rcartocolor::carto_pal(9, "Bold")[1:8] %>%
+  set_names(1:8)
 ```
@@
   labs(fill = "Cluster") +
-  scale_fill_carto_d(type = "qualitative", palette = "Bold") +
-  # scale_fill_brewer(type = "qual") +
+  scale_fill_manual(values = bioregion_cols) +
   map_theme()
@@
   labs(fill = "Cluster") +
-  scale_fill_carto_d(type = "qualitative", palette = "Bold") +
+  scale_fill_manual(values = bioregion_cols) +
   map_theme() +

### 2020-08-20 ms/SI.Rmd
Delete unused functions
- delete traits NMDS
- delete repro mode analysis

@@
 For details, see Methods.
-
-\clearpage
-
-```{r trait-space, fig.width = 8, fig.height = 8}
-# Extract points, join to PPGI
-nmds_points <- traits_nmds[["points"]] %>%
-  as.data.frame %>%
-  rownames_to_column("taxon") %>%
-  as_tibble %>%
-  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
-  left_join(ppgi, by = "genus") %>%
-  mutate(
-    order = factor(order),
-    family = factor(family))
-
-# A: all ferns
-a <- ggplot(nmds_points, aes(x = MDS1, y = MDS2, color = order, shape = order)) +
-  geom_point() +
-  scale_shape_manual(values = 1:n_distinct(nmds_points$order)) +
-  labs(title = "A. Ferns")
-
-# B: Polypods only
-nmds_points_polypod <- filter(nmds_points, order == "Polypodiales")
-
-b <- ggplot(nmds_points_polypod, aes(x = MDS1, y = MDS2, color = family, shape = family)) +
-  geom_point() +
-  scale_shape_manual(values = 1:n_distinct(nmds_points_polypod$family)) +
-  labs(title = "B. Polypodiales") +
-  guides(
-    color = guide_legend(ncol = 2),
-    shape = guide_legend(ncol = 2))
-
-a + b + 
-  plot_layout(nrow = 2)
-
-ggsave(result_file(s_figure("traits-nmds"), "pdf"), width = 8, height = 8, units = "in")
-```
-
-**`r s_figure("traits-nmds")`**.
-`r s_figure_cap("traits-nmds")` Interspecific trait distances calculated using Gower's method with 16 traits, then projected into trait space using nonmetric multidimensional scaling (NMDS).
-**A**, All ferns.
-**B**, Polypodiales.
-Taxonomy follows PPGI.

### 2020-08-20 R/functions.R

@@ -1,57 +1,5 @@
 # Data processing ----
 
-#' Clean up reproductive mode data for pteridophytes of Japan
-#'
-#' @param data Tibble. Raw data read-in from Electronic Supp. Mat. 1
-#'
-#' @return Tibble
-process_repro_data <- function (data) {
-  
-  data %>%
-    clean_names %>%
-    mutate(
-      reproductive_mode = case_when(
-        reproductive_mode == 0 ~ "unknown",
-        reproductive_mode == 1 ~ "sexual",
-        reproductive_mode == 2 ~ "apomictic",
-        reproductive_mode == 3 ~ "sex_apo",
-        TRUE ~ "unknown"
-      ) %>% as.factor,
-      sexual_diploid = case_when(
-        sexual_diploid == 1 ~ TRUE,
-        TRUE ~ FALSE
-      ),
-      sexual_polyploid = case_when(
-        sexual_polyploid == 1 ~ TRUE,
-        TRUE ~ FALSE
-      )
-    ) %>%
-    rename(
-      rbcl_genbank_no = rbc_l_gen_bank_accession_no,
-      species = taxon_name) %>%
-    mutate(rbcl_genbank_no = str_remove_all(rbcl_genbank_no, "\\*"))
-  
-}
-
-#' Get a list of URLs to download data for a kokudosuuchi dataset
-#'
-#' @param dataset Name of the dataset
-#'
-#' @return Character vector; list of URLs for that dataset
-#'
-#' @examples
-#' get_ksj_urls("標高・傾斜度3次メッシュ")
-get_ksj_urls <- function (dataset) {
-  # Get all KSJ metadata
-  ksj_metadata <- kokudosuuchi::getKSJSummary()
-  # Get identifier for the dataset we're interested in
-  dataset_id <- dplyr::filter(ksj_metadata, stringr::str_detect(title, dataset)) %>%
-    dplyr::pull(identifier)
-  # Get a tibble of URLs to zip files with this data
-  urls <- kokudosuuchi::getKSJURL(dataset_id)
-  urls$zipFileUrl
-}
-
 #' Unzip Ebihara and Nitta 2017 Ecol Mono data zip file and 
 #' extract needed data files.
 #' 
@@ -319,51 +267,6 @@ combine_protected_areas <- function (...) {
   
 }
 
-# Reproductive mode ----
-
-#' Calculate percent of sexual diploid taxa per grid cell (site)
-#'
-#' @param comm Community data matrix, with species as columns
[…]

### 2020-08-20 R/plan.R

@@ -3,10 +3,6 @@ plan <- drake_plan (
   
   # Load and process data ----
   
-  # Load basic world map.
-  world_map = ggplot2::map_data("world") %>%
-    rename(longitude = long, latitude = lat),
-  
   # Unzip data files from Ebihara and Nitta 2019
   # This requires doi_10.5061_dryad.4362p32__v4.zip to be downloaded to data_raw/
   # from https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32 first
@@ -127,12 +123,6 @@ plan <- drake_plan (
   # Make trait distance matrix using taxon IDs as labels
   trait_distance_matrix = make_trait_dist_matrix(traits_for_dist),
   
-  # Analyze reproductive mode ----
-  
-  percent_sex_dip = calc_sex_dip(
-    comm = comm_ferns, 
-    repro_data = repro_data),
-  
   # Conduct randomization tests of diversity metrics ----
   
   # Conduct randomization tests of phylogeny-based metrics for all ferns and
@@ -189,9 +179,6 @@ plan <- drake_plan (
   ),
   
   # Traits ----
-  
-  # Run NMDS on traits
-  traits_nmds = vegan::metaMDS(trait_distance_matrix, k = 3),
   
   # Analyze phylogenetic signal
   

### 2020-08-20 ms/captions.R

@@ -20,7 +20,6 @@ s_fig_nums(name = "grain-size", "Effect of grain size on sampling redundancy.")
 s_fig_nums(name = "k-plot", "Selection of *k* for bioregions.")
 s_fig_nums(name = "raw-biplot", "Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.")
 s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.")
-s_fig_nums(name = "traits-nmds", "Trait space of the ferns of Japan.")
 
 # - SI tables
 # s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")

### 2020-08-21 ms/SI.Rmd
Update MS

@@
 loadd(list = c(
+  "japan_shp",
   "comm_scaled_list_0.1",
@@
 # Supplemental Information
+
+```{r map-plot-prep}
+# Change cluster to factor
+biodiv_ferns_spatial <- biodiv_ferns_spatial %>%
+  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+  # Add redundancy
+  mutate(redundancy = 1 - (richness/abundance))
+
+# Download high-res map of Japan,
+# crop to spatial div results area
+japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
+
+# Set CRS
+japan_crs <- sf::st_crs(japan)
+
+biodiv_ferns_endemic_spatial <- sf::st_set_crs(biodiv_ferns_endemic_spatial, japan_crs)
+
+biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
+
+# Set color palettes
+# Randomization test significance
+signif_cols <-
+  c(
+    "> 0.99" = "#e31a1c", 
+    "> 0.975" = "#fb9a99",
+    "< 0.01" = "#1f78b4", 
+    "< 0.025" = "#a6cee3",
+    "not significant" = "#ffffcc"
+  )
+
+# CANAPE
+canape_cols <-
+  c(
+    "neo" = "#e31a1c", 
+    "paleo" = "#1f78b4",
+    "mixed" = "#6a51a3",
+    "super" = "#4a1486",
+    "not significant" = "#ffffcc"
+  )
+```
 
@@
 
+```{r abundance, width = 7, height = 5}
+k <- scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1, sep = "")
+
+a <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = abundance), color = "transparent", alpha = 0.8) +
+  scale_fill_scico(palette = "lajolla", label = k) +
+  labs(fill = "No. specimens") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A**</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  map_theme()
+
+b <- ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = redundancy), color = "transparent", alpha = 0.8) +
+  scale_fill_scico(palette = "lajolla") +
+  labs(fill = "Redundancy") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B**</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  map_theme() +
+  theme(
+    axis.text.y = element_blank()
+  )
+
+a + b + 
+  plot_layout(nrow = 1)
+
+ggsave(result_file(figure("abundance"), "pdf"), width = 7, height = 5, units = "in")
+```
+
+**`r s_figure("abundance")`**. `r s_figure_cap("abundance")`
+
+\clearpage
+
 ```{r k-plot, fig.width = 7, fig.height = 3}
@@
 
-```{r map-plot-prep}
-# Download high-res map of Japan,
-# crop to spatial div results area
-japan <- rnaturalearth::ne_countries(country = "japan", scale = "large", returnclass = "sf") %>%
-  sf::st_crop(sf::st_bbox(biodiv_ferns_endemic_spatial))
-
-# Set CRS
-japan_crs <- sf::st_crs(japan)
-
-biodiv_ferns_endemic_spatial <- sf::st_set_crs(biodiv_ferns_endemic_spatial, japan_crs)
-
-# Set color palettes
-# Randomization test significance
-signif_cols <-
-  c(
-    "> 0.99" = "#e31a1c", 
-    "> 0.975" = "#fb9a99",
-    "< 0.01" = "#1f78b4", 
-    "< 0.025" = "#a6cee3",
-    "not significant" = "#ffffcc"
-  )
-
-# CANAPE
-canape_cols <-
-  c(
-    "neo" = "#e31a1c", 
-    "paleo" = "#1f78b4",
-    "mixed" = "#6a51a3",
-    "super" = "#4a1486",
-    "not significant" = "#ffffcc"
-  )
-```
-
 ```{r endemism-restricted, fig.width = 5, fig.height = 5}

### 2020-08-21 ms/manuscript.Rmd
Update MS

@@
   "k_taxonomy",
-  "k_phylogeny"
+  "k_phylogeny",
+  "signif_cells_protected_area",
+  "protected_areas",
+  "japan_shp"
 ),
@@
 We combined phylogenetic, functional, and environmental datasets to explore the spatial distribution of biodiversity in the ferns of Japan.
+We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity, which is highest in small southern islands, and that of relative phylogenetic and functional diversity, which is high at both high and low latitudes, and low at mid-latitudes.
 
@@
 
-We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
+We used traits originally compiled for identification of ferns and lycophytes of Japan using LUCID software [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
 Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens [@Ebihara2016b; @Ebihara2017].
-All qualitative traits were converted to binary format; \eg, a single trait with three states was formatted as three binary traits.
+All qualitative traits were measured in binary format; \eg, a single trait with three states was formatted as three binary traits.
 Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately originally, then averaged.
@@
 
-### Overlap with protected zones
+### Assessment of conservation status
 
 We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html).
-We categorized protected areas as either "high" (plants protected; no human activities allowed at all) or "normal" (plants protected; some human activities allowed by permit).
-We cropped the grid cells to only include land area, then calculated the amount of protected area of each status type within grid cells with significantly high PD, FD, or PE (one-tailed test).
+We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "normal" (some human activities allowed by permit).
+We cropped grid cells to only include land area, then calculated the percent of protected area of each status type within grid cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
+For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above.
+For taxonomic richness, cells ranked in the top >5% were considered highly diverse.
 
@@
 
-### Number of specimens and redundancy
+### Datasets
 
-The distribution of specimen collection locations is highly skewed, with >3,000 specimens per grid cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r figure("abundance")`A).
-Despite this, sampling redundancy is generally high throughout the country, with areas of low redundancy in Hokkaido and some coastal areas and small islands (`r figure("abundance")`B).
+The topology of the phylogenetic tree was in general agreement with other recent, large plastid fern phylogenies.
+Ferns were recovered as monophyletic, as were all families and genera; taxonomy follows PPGI for the family level and the Japan Green List for the genus level and below.
+The age of the fern clade is estimated to be _ my.
+
+The distribution of specimen collection locations is highly skewed, with >3,000 specimens per grid cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
+Despite this, sampling redundancy is generally high throughout the country, with areas of low redundancy in Hokkaido and some coastal areas and small islands (`r s_figure("abundance")`B).
 The grid cells that had to be excluded from the analysis due to low redundancy were mostly from Hokkaido.
@@
 
-Taxonomic clusters 3 and 4 have higher SES values of PD, RPD, FD, and RFD than clusters 1 and 2.
+Taxonomic clusters 3 and 4 have higher SES values of PD, RPD, FD, and RFD than clusters 1 and 2 (`r figure("div-by-tax-region")`).
 Clusters 3 and 4 also have consistently high PE *p*-scores, whereas clusters 1 and 2 have a much wider variation.
@@
 
-### Degree of protection
+### Conservation status
+
+```{r conserv-stats}
+# Convert signif_cells_protected_area to wide format
+signif_cells_protected_area_wide <-
+signif_cells_protected_area %>% select(status, metric, percent_protected) %>%
+  pivot_wider(names_from = status, values_from = percent_protected) %>%
+  rowwise %>%
+  mutate(total = sum(medium, high)) %>%
+  ungroup
+
+# Calculate min and max percent area protected overall
+min_protected_med <- signif_cells_protected_area_wide %>% pull(medium) %>% min %>% scales::percent(accuracy = 0.1)
+max_protected_med <- signif_cells_protected_area_wide %>% pull(medium) %>% max %>% scales::percent(accuracy = 0.1)
+min_protected_high <- signif_cells_protected_area_wide %>% pull(high) %>% min %>% scales::percent(accuracy = 0.1)
+max_protected_high <- signif_cells_protected_area_wide %>% pull(high) %>% max %>% scales::percent(accuracy = 0.1)
+min_protected_total <- signif_cells_protected_area_wide %>% pull(total) %>% min %>% scales::percent(accuracy = 0.1)
+max_protected_total <- signif_cells_protected_area_wide %>% pull(total) %>% max %>% scales::percent(accuracy = 0.1)
+
+# Calculate min and max protected area for the best protected category
+signif_cells_protected_area_best <-
+  signif_cells_protected_area_wide %>%
+  slice_max(total)
+
+# Verify that the best-protected biodiv type is PE
+assertthat::assert_that(signif_cells_protected_area_best$metric == "pe")
+
+best_protected_med <- signif_cells_protected_area_best %>% pull(medium) %>% scales::percent(accuracy = 0.1)
+best_protected_high <- signif_cells_protected_area_best %>% pull(high) %>% scales::percent(accuracy = 0.1)
+best_protected_total <- signif_cells_protected_area_best %>% pull(total) %>% scales::percent(accuracy = 0.1)
+
+# Calculate the total area of Japan
+area_of_japan <-
+  japan_shp %>%
+  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
+  pull(area)
+
+# Calculate baseline protection percentage: percent protected areas across the whole country
+total_percent_protected <-
+  protected_areas %>%
+  as_tibble() %>%
+  select(-geometry) %>%
+  filter(status %in% c("medium", "high")) %>%
+  group_by(status) %>%
+  summarize(
+    total_area = sum(area),
+    .groups = "drop"
+  ) %>%
+  mutate(percent_protected = total_area/area_of_japan) %>%
+  mutate(percent_protected = as.numeric(percent_protected))
+
+# Extract values for medium and high protection level
+total_percent_protected_medium <- total_percent_protected %>% filter(status == "medium") %>% pull(percent_protected)
+total_percent_protected_high <- total_percent_protected %>% filter(status == "high") %>% pull(percent_protected)
+total_percent_protected_all <- sum(total_percent_protected_medium, total_percent_protected_high)
+```
+
+The percentage of protected area in cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`).
+Areas with high PE were the best protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total).
+All cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high status, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total).
 
@@
 Here, we provide the first insights into the combined taxonomic, phylogenetic, and functional diversity of the ferns of Japan.
-Our integrative approach reveals key differences in patterns of biodiversity in different parts of the archipelago.
+Our integrative approach reveals distinct patterns of biodiversity in different parts of the archipelago and across biodiversity metrics.
+These clear differences are reflected in the bioregionalization results, which delimit three to four major bioregions.
+Finally, we consider the current conservation status of ferns in Japan and the implications of our findings for future conservation activity.
 
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here has been well-documented.
+The latitudinal hump-shaped pattern of species richness observed in ferns here has been well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015].
+This pattern is probably due at least in part to the distribution of land area in Japan: maximum land area is located at mid-latitudes, and area is related to species richness by the well-known species-area relationship [@Arrhenius1921].
+There may also be a combined climatic and elevational effect: the southern part of Honshu is warmer and more humid than areas further north, while also harboring high elevational variation that may support a wide range of taxa, which is not present in the southern islands.
 
-Despite the peak of richness in southern Kyushu, the overwhealming pattern driving phylogenetic and functional diversity is the high diversity of southern islands (Ryukyu and Bonin Isl.).
+The overwhelming pattern characterizing phylogenetic and functional diversity is the high diversity of southern islands (Ryukyu and Bonin Isl.).
+This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; _ genera and _ families only occur south of 30.1° latitude (just south of Yakushima Isl.).
+The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis, and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity.
+Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE are located in the southern islands.
+This is clearly due to the small land area of these islands, which are the only areas with a subtropical climate in Japan.
 
-Another major pattern revealed here is the predominance of long branches in the northern half of Honshu, and the predominance of short branches in the southern half, particularly Kyushu.
+Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu.
+The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* and *Ophioglossum*.
+The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera such as *Polystichum* and *Cyrtomimum*.
 
-### Bioregions
+The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values.
+Interestingly, the bioregions identified here correspond roughly to major forest types categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014].
+This suggests that ferns are useful as possible bioindicators in Japan.
+Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions.
+The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl.
+The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
 
-### Conservation status
+### Conservation status and future outlook
+
+Although the different facets of biodiversity have different distribution patterns, we find that grid cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric.
+While it is not clear why this is so---it is unlikely that ferns were emphasized as targets during the establishment of conservation zones---one possibility (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
+Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change.
+One such other threat is herbivory by native Japanese deer, *Cervus nippon*.
+Populations of these deer have been increasing exponentially since the 1970s due to a combination of climate change and a decrease in hunting and natural predators.
+The deer are capable of decimating the herb layer when populations become dense and have been causing extensive damage even within national parks, which have the highest protection status.
+Although various efforts to prevent deer herbivory have been deployed such as exclusion fences, much work remains to be done to save vulnerable plant populations.
+
+There are a number of caveats that must be kept in mind when interpreting these results.
+First, the data include species collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan.
+Furthermore, the collections are not evenually collected throughout the country, but rather some areas have many more specimens than others.
+Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.
+
+<!-- hybrid taxa, gametophytes, conclusion -->
 
@@
 # crop to spatial div results area
-japan <- rnaturalearth::ne_countries(country = "japan", scale = "large", returnclass = "sf") %>%
-  sf::st_crop(sf::st_bbox(biodiv_ferns_spatial))
+japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
 
@@
 ```
-
-```{r abundance, width = 7, height = 5}
-k <- scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1, sep = "")
-
-a <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = abundance), color = "transparent", alpha = 0.8) +
-  scale_fill_scico(palette = "lajolla", label = k) +
-  labs(fill = "No. specimens") +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A**</span>", hjust = "inward", vjust = "inward", color = "white") + 
-  map_theme()
-
-b <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = redundancy), color = "transparent", alpha = 0.8) +
-  scale_fill_scico(palette = "lajolla") +
-  labs(fill = "Redundancy") +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B**</span>", hjust = "inward", vjust = "inward", color = "white") + 
-  map_theme() +
-  theme(
-    axis.text.y = element_blank()
-  )
-
-a + b + 
-  plot_layout(nrow = 1)
-
-ggsave(result_file(figure("abundance"), "pdf"), width = 7, height = 5, units = "in")
-```
-
-**`r figure("abundance")`**. `r figure_cap("abundance")`
-
-\clearpage
 
@@
 
-```{r diversity-by-cluster, fig.width = 7, fig.height = 7}
+```{r div-by-tax-region, fig.width = 7, fig.height = 7}
 data <-
@@
 Only top four clusters with the most sites shown.
+
+\clearpage
+
+```{r conserv-status, fig.width = 4, fig.height = 4}
+# Plot protected areas
+signif_cells_protected_area %>%
+  mutate(
+    metric = case_when(
+      metric == "pd" ~ "PD",
+      metric == "fd" ~ "FD",
+      metric == "pe" ~ "PE",
+      metric == "richness" ~ "Richness",
+    ),
+    status = str_to_sentence(status)) %>%
+  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "Richness"))) %>%
+  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
+  geom_col(position = position_dodge()) +
+  geom_vline(xintercept = total_percent_protected_medium, color = "grey40", linetype = 3) +
+  geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
+  scale_fill_viridis_d() +
+  scale_x_continuous(labels = scales::percent) +
+  labs(
+    x = "Percent protected",
+    y = "Biodiversity metric",
+    fill = "Protection status"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    legend.position = "bottom"
+  )
+
+ggsave(result_file(figure("conserv-status"), "pdf"), width = 4, height = 4, units = "in")
+
+```
+
+**`r figure("conserv-status")`**.
+`r figure_cap("conserv-status")` Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), relative PD (RPD), relative FD (RFD), and phylogenetic endemism (PE).
+Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Methods).
+Vertical lines indicate total percent area protected across Japan regardless of degree of biodiversity; dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).

### 2020-08-21 R/functions.R

@@ -1776,6 +1776,74 @@ cluster_phylo_regions <- function (comm_df, phy, k) {
     as_tibble()
 }
 
+# Conservation ----
+
+#' Calculate the percent of area within grid cells of high biological diversity
+#' that has protected status
+#'
+#' @param biodiv_ferns_spatial Spatial dataframe including biodiveristy metrics and grid cells
+#' @param protected_areas Spatial dataframe including status of protected areas and their shapes
+#' @param japan_shp Spatial dataframe including the total land area of Japan
+#'
+#' @return Dataframe with the percent of protected areas in grid cells with significantly high biodiversity
+#' 
+calculate_protected_area <- function(biodiv_ferns_spatial, protected_areas, japan_shp) {
+  
+  # Add percent rank for richness
+  biodiv_ferns_spatial <- mutate(biodiv_ferns_spatial, richness_obs_p_upper = dplyr::percent_rank(richness))
+  
+  # Crop to Japan to spatial div results area
+  japan_shp <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
+  
+  # Set CRS for biodiversity data and protected areas
+  japan_crs <- sf::st_crs(japan_shp)
+  
+  biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
+  
+  protected_areas <- sf::st_set_crs(protected_areas, japan_crs)
+  
+  japan_shp <- select(japan_shp, admin, geometry)
+  
+  # Crop spatial data to only land regions within Japan map
+  biodiv_ferns_spatial_cropped <- raster::intersect(sf::as_Spatial(biodiv_ferns_spatial), sf::as_Spatial(japan_shp)) %>% sf::st_as_sf()
+  
+  biodiv_ferns_spatial_conserv <- sf::st_join(biodiv_ferns_spatial_cropped, protected_areas, left = TRUE)
+  
+  # Calculate total area of cells with significantly high biodiversity
+  signif_cells_total_area <-
+    biodiv_ferns_spatial_cropped %>%
+    mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
+    select(grids, richness_obs_p_upper, pd_obs_p_upper, fd_obs_p_upper, pe_obs_p_upper, area) %>%
+    as_tibble() %>%
+    select(-geometry) %>%
+    pivot_longer(cols = contains("upper"), values_to = "p_value", names_to = "metric") %>%
[…]

### 2020-08-21 R/plan.R

@@ -24,13 +24,8 @@ plan <- drake_plan (
   green_list = read_excel(file_in("data_raw/ebihara_2019/FernGreenListV1.01E.xls")) %>% 
     tidy_japan_names(),
   
-  # Load reproductive mode data (one row per species)
-  repro_data_raw = read_csv(
-    file_in("data_raw/ebihara_2019/ESM1.csv"),
-    col_types = "cccccnnnnn"),
-  
-  repro_data = process_repro_data(repro_data_raw) %>%
-    rename_taxa(green_list),
+  # Load a map of Japan
+  japan_shp = rnaturalearth::ne_countries(country = "japan", scale = "large", returnclass = "sf"),
   
   # Load raw occurrence data of pteridophytes in Japan, excluding hybrids (717 taxa)
   occ_point_data_raw = readxl::read_excel(
@@ -314,6 +309,9 @@ plan <- drake_plan (
     protected_6,
     protected_7
   ),
+  
+  # Calculate percent protection for grid cells with significantly high biodiversity
+  signif_cells_protected_area = calculate_protected_area(biodiv_ferns_spatial, protected_areas, japan_shp),
   
   # Write out manuscript ----
   ms = rmarkdown::render(

### 2020-08-21 ms/captions.R

@@ -2,12 +2,12 @@
 
 # - Figures
 fig_nums <- captioner::captioner(prefix = "Fig.")
-fig_nums(name = "abundance", caption = "Observed number of specimens (**A**) and sampling redundancy (**B**) per 0.2° grid cell in the ferns of Japan.")
 fig_nums(name = "raw-diversity", caption = "Raw taxonomic (**A**), phylogenetic (**B**), and functional (**C**) diversity of the ferns of Japan.")
 fig_nums(name = "rand-diversity", caption = "Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.")
 fig_nums(name = "endemism", caption = "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).")
 fig_nums(name = "bioregions", caption = "Bioregions of the ferns of Japan.")
 fig_nums(name = "div-by-tax-region", caption = "Phylogenetic and morphological diversity of the ferns of Japan by bioregion.")
+fig_nums(name = "conserv-status", caption = "Percent of land area with protected status for grid cells with significantly high biodiversity.")
 
 # - Tables
 table_nums <- captioner::captioner(prefix = "Table")
@@ -17,6 +17,7 @@ table_nums(name = "traits-used", caption = "Fern traits used in this study")
 s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
 s_fig_nums(name = "sampling-curve", "Species collection curve for the ferns of Japan fit with the iNEXT package.")
 s_fig_nums(name = "grain-size", "Effect of grain size on sampling redundancy.")
+s_fig_nums(name = "abundance", caption = "Observed number of specimens (**A**) and sampling redundancy (**B**) per 0.2° grid cell in the ferns of Japan.")
 s_fig_nums(name = "k-plot", "Selection of *k* for bioregions.")
 s_fig_nums(name = "raw-biplot", "Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.")
 s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.")

### 2020-08-25 ms/manuscript.Rmd
Update MS

@@
   "protected_areas",
-  "japan_shp"
+  "japan_shp",
+  "phy_sig_results",
+  "binary_sig_results",
+  "traits_summary",
+  "fern_traits",
+  "japan_pterido_tree",
+  "japan_fern_tree",
+  "ppgi",
+  "occ_point_data_ferns",
+  "comm_ferns"
 ),
@@
 A necessary step in any analysis of biodiversity is to set the grain size used to define communities.
-This decision must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling.
-We tested using 0.1, 0.2, 0.3, and 0.4 degree grid-cells for defining communities (co-occurring species).
-At each size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell).
+Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling.
+We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of the cell) for defining communities (co-occurring species).
+At each grain size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell).
 We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling.
-Preliminary analysis indicated that 0.2 degree grid-cells are optimal for our dataset (`r s_figure("grain-size")`).
-We filtered out any cells with redundancy < 0.1 (mostly cells along the coast with very low land area) and used this for subsequent analyses.
+Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`).
+We filtered out any cells with redundancy < 0.1 and used this for subsequent analyses.
 
@@
 
-We used traits originally compiled for identification of ferns and lycophytes of Japan using LUCID software [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
-Quantitative traits were measured on 10 randomly selected specimens per species, and qualitative traits determined by observing voucher specimens [@Ebihara2016b; @Ebihara2017].
-All qualitative traits were measured in binary format; \eg, a single trait with three states was formatted as three binary traits.
-Quantitative traits were measured in fertile (i.e., spore-bearing) and sterile fronds separately originally, then averaged.
-From the original trait list, we selected only traits with putatative ecological function, and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state (`r table("traits-used")`).
+We used traits originally compiled for identification of ferns and lycophytes of Japan as follows [@Ebihara2016b; @Ebihara2017].
+Continuous traits were measured on 10 randomly selected specimens per species.
+Fertile (i.e., spore-bearing) and sterile fronds were measured separately, then averaged.
+Qualitative traits were scored by observing voucher specimens.
+All qualitative traits were scored in binary format; \eg, a single trait with three states was formatted as three binary traits.
+From the original trait list, we selected only traits with putatative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.
 
@@
 97% of the Japanese fern flora [@Ebihara2019b].
-However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils availabsle for molecular dating.
+However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating.
 Because one goal of the study is to understand the distribution of paleo- vs.
@@
 *K* describes the ratio between the observed variance in a trait vs.
-the amount of variance expected under Brownian Motion (BM); lambda is a scaling parameter used to transform the phylogeny such that trait values fit those most closely expected under BM.
+the amount of variance expected under Brownian Motion (BM); lambda is a scaling parameter that transforms the phylogeny such that trait values fit those most closely expected under BM.
 Both lambda and *K* = 1 under BM; lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering).
@@
 
-We measured functional diversity (FD) following the method of @Petchey2002, which is a functional analog of PD.
-We first calculated distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011].
-We weighted quantitative and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled quantitative traits prior to calculating distances.
+We measured functional diversity (FD) following the method of @Petchey2002.
+We first calculated functional distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011].
+We weighted continuous and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled continuous traits prior to calculating distances.
 We then used the distances to build a trait dendrogram using the \function{hclust} function in the \package{stats} R package [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
+These functional analogs of PD and RPD are hereafter referred to as FD (functional diversity) and RFD (relative functional diversity).
 
@@
 We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values.
-Statistical significance was assessed using a two-tailed test with alpha = 0.5: observed values > 97.5% or < 0.25% were considered significant.
+Statistical significance was assessed using a two-tailed test with alpha = 0.5; observed values > 97.5% or < 0.25% were considered significant.
 
@@
 Low RPE indicates short branchlengths (neoendemism) and high RPE indicates long branchlengths (paleoendemism).
-Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable to those calculated for the null distribution.
-We used the same set of null communities as for SES of PE, and carried out CANAPE using custom R scripts.
+Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution.
+We used the same set of null communities as for SES of PE, and carried out categorization using custom R scripts.
 
@@
 We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html).
-We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "normal" (some human activities allowed by permit).
+We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some human activities allowed by permit).
 We cropped grid cells to only include land area, then calculated the percent of protected area of each status type within grid cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
@@
 
-The topology of the phylogenetic tree was in general agreement with other recent, large plastid fern phylogenies.
-Ferns were recovered as monophyletic, as were all families and genera; taxonomy follows PPGI for the family level and the Japan Green List for the genus level and below.
-The age of the fern clade is estimated to be _ my.
+```{r dataset-stats}
+### Phylogeny ###
 
-The distribution of specimen collection locations is highly skewed, with >3,000 specimens per grid cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
-Despite this, sampling redundancy is generally high throughout the country, with areas of low redundancy in Hokkaido and some coastal areas and small islands (`r s_figure("abundance")`B).
-The grid cells that had to be excluded from the analysis due to low redundancy were mostly from Hokkaido.
+# FIXME: confirm total number of taxa with Ebihara
+total_fern_taxa <- 675
+phylogeny_sampled_percent <- ape::Ntip(japan_fern_tree) %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)
+
+# Check height (age) of fern phylogeny
+phylogeny_age <- japan_fern_tree %>% phytools::nodeHeights() %>% max %>% round(1)
+
+# Check monophyly of genera
+mono_status_genera <-
+tibble(
+  tip = japan_fern_tree$tip.label,
+  tip_no = 1:length(tip),
+  genus = str_split(tip, "_") %>% map_chr(1)
+) %>%
+  group_by(genus) %>%
+  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
+  mutate(is_mono = map_lgl(tip_vec, ~ape::is.monophyletic(phy = japan_fern_tree, tips = .)))
+
+# Assert only one genus is not monophyletic
+assert_that(mono_status_genera %>% filter(is_mono == FALSE) %>% nrow == 1)
+
+non_mono_genus <- mono_status_genera %>% filter(is_mono == FALSE) %>% pull(genus)
+
+n_fern_phylo_genera <- mono_status_genera %>% pull(genus) %>% n_distinct()
+
+# Check monophyly of families
+mono_status_families <-
+tibble(
+  tip = japan_fern_tree$tip.label,
+  tip_no = 1:length(tip),
+  genus = str_split(tip, "_") %>% map_chr(1)
+) %>%
+  left_join(ppgi, by = "genus") %>%
+  group_by(family) %>%
+  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
+  mutate(is_mono = map_lgl(tip_vec, ~ape::is.monophyletic(phy = japan_fern_tree, tips = .)))
+
+# Assert all families are monophyletic
+assert_that(mono_status_families %>% filter(is_mono == FALSE) %>% nrow == 0)
+
+n_fern_phylo_families <- mono_status_families %>% pull(family) %>% n_distinct()
+
+# Check monophyly of ferns
+mono_status_class <- 
+tibble(
+  tip = japan_pterido_tree$tip.label,
+  tip_no = 1:length(tip),
+  genus = str_split(tip, "_") %>% map_chr(1)
+) %>%
+  left_join(ppgi, by = "genus") %>%
+  group_by(class) %>%
+  summarize(tip_vec = list(tip_no), .groups = "drop") %>%
+  mutate(is_mono = map_lgl(tip_vec, ~ape::is.monophyletic(phy = japan_pterido_tree, tips = .)))
+
+# Assert all classes (ferns, lycophytes) are monophyletic
+assert_that(mono_status_class %>% filter(is_mono == FALSE) %>% nrow == 0)
+
+### Specimens ###
+n_specimens <- nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")
+n_specimens_taxa <- occ_point_data_ferns %>% pull(taxon) %>% n_distinct()
+n_specimens_taxa_percent <- n_specimens_taxa %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)
+
+# Calculate mean redundancy
+mean_redundancy <-
+biodiv_ferns_spatial %>%
+  select(redundancy) %>%
+  st_set_geometry(NULL) %>%
+  summarize(
+    mean = mean(redundancy),
+    sd = sd(redundancy)
+  ) %>%
+  mutate(text = glue::glue("{round(mean, 2)} ± {round(sd, 2)}")) %>%
+  pull(text)
+
+### Traits ###
+
+# Count number of traits of each type in final dataset
+n_binary_traits <- traits_summary %>% filter(trait_type == "binary") %>% nrow() %>% jntools::english2()
+n_cont_traits <- traits_summary %>% filter(trait_type == "continuous") %>% nrow() %>% jntools::english2()
+n_qual_traits <- traits_summary %>% filter(trait_type == "qualitative") %>% nrow() %>% jntools::english2()
+n_qual_traits_as_binary <- traits_summary %>% filter(trait_type == "qualitative") %>% pull(n) %>% sum %>% jntools::english2()
+
+# Calculate number/percent of missing data in trait matrix
+fern_traits_missing_summary <-
+fern_traits %>%
+  pivot_longer(-taxon) %>%
+  mutate(missing_data = is.na(value)) %>%
+  summarize(
+    total = n(),
+    missing = sum(missing_data)
+  )
+
+percent_traits_missing <-
+fern_traits_missing_summary$missing %>% 
+  divide_by(fern_traits_missing_summary$total) %>%
+  scales::percent(accuracy = 0.01)
+```
+
+The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent`) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families of native Japanese ferns.
+The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a].
+Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* (families follow @PteridophytePhylogenyGroupI2016).
+The age of the fern clade is estimated to be `r phylogeny_age` ma.
+
+The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`).
+The distribution of collection locations was highly skewed, with > 3,000 specimens per grid cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
+Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B).
+The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid cells.
+__ grid cells were excluded from the analysis due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.
+
+After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits) (`r table("traits-used")`).
+The trait matrix included `r nrow(fern_traits)` taxa and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.
 
@@
 
-Results of phylogenetic signal analysis are given in Table _.
-All quantitative traits showed significant phylogenetic signal, but estimates of signal strength differed between K and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of K were < 0.06, indicating less phylogenetic signal than expected under BM.
-A minority of quantitative traits (coded as binary) showed as much or more phylogenetic signal as expected under BM (D =< 0); others had D > 0.
+```{r phy-sig-stats}
+# Verify that min lambda >0.95
+assert_that(phy_sig_results %>% pull(lambda) %>% min > 0.95)
+# Verify that max K < 0.0.6
+assert_that(phy_sig_results %>% pull(kval) %>% max < 0.06)
+
+# Sum the number of traits with D < 0
+n_negative_d <-
+binary_sig_results %>%
+  mutate(negative_d = ifelse(D <= 0, TRUE, FALSE)) %>%
+  pull(negative_d) %>%
+  sum
+```
+
+All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.06, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
+`r n_negative_d` out of `r nrow(binary_sig_results)` of continuous traits (coded as binary) showed as much or more phylogenetic signal as expected under BM (*D* =< 0); others had *D* > 0 (`r s_table("phy-sig-binary")`).
 
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here has been well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015].
+The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015] and modeled distributions of woody plants [@Fukaya2020].
 This pattern is probably due at least in part to the distribution of land area in Japan: maximum land area is located at mid-latitudes, and area is related to species richness by the well-known species-area relationship [@Arrhenius1921].
-There may also be a combined climatic and elevational effect: the southern part of Honshu is warmer and more humid than areas further north, while also harboring high elevational variation that may support a wide range of taxa, which is not present in the southern islands.
+There may also be a combined climatic and elevational effect: the southern part of Honshu is warmer and more humid than areas further north, while also harboring high elevational variation that may support a wide range of taxa, which is not present in the smaller islands further south.
 
@@
 <div id="refs"></div>
+
+\clearpage
+
+<!-- Tables -->
+
+```{r traits-used, results = "asis"}
+trait_signif <- read_csv(here::here("data_raw/trait_functions.csv"))
+
+traits_summary %>%
+  left_join(trait_signif, by = "trait") %>%
+  mutate(
+    trait = case_when(
+      trait == "frond_width" ~ "Frond width",
+      trait == "number_pinna_pairs" ~ "No. pinna pairs",
+      trait == "stipe_length" ~ "Stipe length",
+      trait == "margin" ~ "Frond margin",
+      trait == "shape" ~ "Frond shape",
+      trait == "texture" ~ "Frond texture",
+      TRUE ~ str_to_sentence(trait)
+    ),
+    trait_type = case_when(
+      trait_type == "qualitative" ~ glue::glue("Qualitative (coded as {n} binary traits)"),
+      trait_type == "continuous" & str_detect(trait, "pinna pairs") ~ glue::glue("Continuous (integer)"),
+      trait_type == "continuous" ~ glue::glue("Continuous (cm)"),
+      TRUE ~ str_to_sentence(trait_type)
+    )
+  ) %>%
+  select(trait, Significance = significance) %>%
+  column_to_rownames("trait") %>%
+knitr::kable(format = "latex", booktabs = TRUE, escape = TRUE)
+
+```
+
+**`r table("traits-used")`**. `r table_cap("traits-used")`
 

### 2020-08-25 ms/captions.R

@@ -11,7 +11,7 @@ fig_nums(name = "conserv-status", caption = "Percent of land area with protected
 
 # - Tables
 table_nums <- captioner::captioner(prefix = "Table")
-table_nums(name = "traits-used", caption = "Fern traits used in this study")
+table_nums(name = "traits-used", caption = "Fern traits used in this study.")
 
 # - SI figures
 s_fig_nums <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = ": ")
@@ -23,8 +23,9 @@ s_fig_nums(name = "raw-biplot", "Relationships between observed functional and p
 s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.")
 
 # - SI tables
-# s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
-# s_table_nums(name = "si_table", "")
+s_table_nums <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = ": ")
+s_table_nums(name = "phy-sig", caption = "Phylogenetic signal in continuous functional traits of the ferns of Japan.")
+s_table_nums(name = "phy-sig-binary", caption = "Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.")
 
 # Make short versions of citation functions
 
@@ -32,10 +33,10 @@ s_fig_nums(name = "endemism-restricted", "Phylogenetic endemism of the ferns of 
 figure <- pryr::partial(fig_nums, display = "cite")
 table <- pryr::partial(table_nums, display = "cite")
 s_figure <- pryr::partial(s_fig_nums, display = "cite")
-# s_table <- pryr::partial(s_table_nums, display = "cite")
+s_table <- pryr::partial(s_table_nums, display = "cite")
 
 # - Just the caption
 figure_cap <- function(x) {fig_nums(x) %>% stringr::str_match(": (.*)$") %>% magrittr::extract(,2)}
 table_cap <- function(x) {table_nums(x) %>% stringr::str_match(": (.*)$") %>% magrittr::extract(,2)}
 s_figure_cap <- function(x) {s_fig_nums(x) %>% stringr::str_match(": (.*)$") %>% magrittr::extract(,2)}
[…]

### 2020-08-26 ms/SI.Rmd
Update MS

@@
   "traits_nmds",
-  "ppgi"
+  "ppgi",
+  "phy_sig_results",
+  "binary_sig_results"
 ),
@@
 # Supplemental Information
+
+```{r phy-sig}
+phy_sig_results %>%
+  rename(K = kval) %>%
+  mutate(K = round(K, 3)) %>% 
+  mutate(lambda = round(lambda, 3)) %>%
+  mutate(lambda_pval = format(lambda_pval, digits = 3)) %>%
+  write_csv(path = result_file(s_table("phy-sig"), "csv"))
+```
+
+**`r s_table("phy-sig")`**. `r s_table_cap("phy-sig")` K, Blomberg's *K*; lambda, Pagel's lambda.
+
+```{r phy-sig-binary}
+binary_sig_results %>%
+  arrange(trait) %>%
+  write_csv(path = result_file(s_table("phy-sig-binary"), "csv"))
+```
+
+**`r s_table("phy-sig-binary")`**.
+`r s_table_cap("phy-sig-binary")` num_present, number of times a trait was observed present; num_absent, number of times a trait was observed absent; D, Fritz and Purvis' *D*; prob_random, probability of obtaining observed trait distribution given random distribution of traits; prob_brownian, probability of obtaining observed trait distribution given random evolution by Brownian Motion.
+Presence or absence of indusium was originally a binary trait; other traits (margin, shape, texture) are qualitative traits scored in binary format.
 

### 2020-08-26 ms/manuscript.Rmd
Update MS

@@
   "occ_point_data_ferns",
-  "comm_ferns"
+  "comm_ferns_full",
+  "comm_ferns",
+  "japan_points_raw"
 ),
@@
 Our data include all ferns known to occur within the political boundaries of Japan (20° to 45° N, 122° to 153° E).
-Japan consists of four main islands and thousands of smaller ones.
+Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`).
 The climate is subtropical in the south and highly seasonal in other parts of the country.
+Map data were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).
 
@@
 Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`).
-The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (species, subspecies, and varieties).
+The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (hereafter defined as species, subspecies, and varieties unless otherwise indicated).
 We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).
-The final cleaned occurrence dataset included `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` specimens (`r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` taxa).
-Hereafter, "taxa" refers to species, subspecies, and varieties unless otherwise indicated.
 
@@
 Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling.
-We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of the cell) for defining communities (co-occurring species).
-At each grain size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell).
+We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species).
+At each grain size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell).
 We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling.
@@
 Continuous traits were measured on 10 randomly selected specimens per species.
-Fertile (i.e., spore-bearing) and sterile fronds were measured separately, then averaged.
+Fertile (\ie, spore-bearing) and sterile fronds were measured separately, then averaged.
 Qualitative traits were scored by observing voucher specimens.
@@
 
-We compared SES of PD, RPD, FD, and RFD across bioregions using FIXME: add this when figure out spatial-corrected statistical method to use.
+We compared SES of PD, RPD, FD, and RFD across bioregions using **FIXME: add this when figure out spatial-corrected statistical method to use**.
 
@@
 We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some human activities allowed by permit).
-We cropped grid cells to only include land area, then calculated the percent of protected area of each status type within grid cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
+We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
 For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above.
@@
 
+# Calculate number of excluded cells due to low redundancy
+n_excluded_cells <- nrow(comm_ferns_full) - nrow(comm_ferns)
+
 ### Traits ###
@@
 The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`).
-The distribution of collection locations was highly skewed, with > 3,000 specimens per grid cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
-Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B).
-The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid cells.
-__ grid cells were excluded from the analysis due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.
+The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
+Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B).
+The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid-cells.
+`r n_excluded_cells` grid-cells were excluded due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.
 
@@
 
-Grid cells with significantly low PD were almost all located on Honshu, the main island of Japan.
+Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan.
 A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Amami, Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula).
 The area near Yakushima includes cells with both significantly high and low PD.
-Grid cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu.
-Grid cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido.
+Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu.
+Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido.
 Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD.
@@
 
-Nearly all grid cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere.
-The vast majority of endemic grid cells are mixed, with very few paleo- or neoendemic cells.
+Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere.
+The vast majority of endemic grid-cells are mixed, with very few paleo- or neoendemic cells.
 Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (Daisetsuzan mountains).
@@
 
-Analysis of biogeographic regions using taxonomy (species names) resulted in the vast majority of grid cells classified into four regions: Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Okinawa), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A).
-Using phylogenetic distances produced similar results, but Okinawa and the Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B).
-The other much smaller clusters are likely artifacts of insufficiently low taxon richness for accurate clustering in a small number of grid cells, and are not discussed further.
+Analysis of bioregions grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` clusters (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
+The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Okinawa), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A).
+Phylogenetic distances produced similar results, but Okinawa and the Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B).
+The other much smaller clusters are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
 
@@
 
-The percentage of protected area in cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`).
+The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`).
 Areas with high PE were the best protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total).
-All cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high status, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total).
+All cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total).
 
@@
 
-### Conservation status and future outlook
+### Conservation status
 
-Although the different facets of biodiversity have different distribution patterns, we find that grid cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric.
+Although the different facets of biodiversity have different distribution patterns, we find that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric.
 While it is not clear why this is so---it is unlikely that ferns were emphasized as targets during the establishment of conservation zones---one possibility (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
+
 Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change.
@@
 Although various efforts to prevent deer herbivory have been deployed such as exclusion fences, much work remains to be done to save vulnerable plant populations.
+
+### Limits of the study and future outlook
 
@@
 ```
+
+```{r japan-map, fig.width = 5, fig.height = 5}
+# Print map of Japan with labels of relevant places
+
+# Set vectors of different areas to label
+main_islands <- c("Hokkaido", "Shikoku", "Kyushu", "Honshu")
+small_islands <- c("Okinawa", "Amamioshima", "Ogasawara", "Yakushima", "Izu Islands")
+other <- c("Tokyo")
+
+# Modify dataframe to include formatting for printing label on map
+japan_points <- 
+  japan_points_raw %>%
+  as_tibble %>%
+  mutate(
+    label = case_when(
+      # Print main islands in bold
+      query %in% main_islands ~ glue::glue("**{query}**"),
+      TRUE ~ query
+    ),
+    nudge = case_when(
+      query == "Honshu" ~ "-0.5,-.7",
+      query == "Tokyo" ~ "-1.2,0",
+      query == "Shikoku" ~ "1,-.5",
+      query == "Kyushu" ~ "1,-.6",
+      query %in% small_islands ~ "2.5,0",
+      TRUE ~ "0,0"
+    )
+  ) %>%
+  separate(nudge, c("nudge_lon", "nudge_lat"), sep = ",") %>%
+  mutate(across(c(nudge_lon, nudge_lat), as.numeric)) %>%
+  mutate(
+    lon_adj = lon + nudge_lon,
+    lat_adj = lat + nudge_lat
+  )
+
+# Make map
+ggplot(japan) + 
+  geom_sf() +
+  geom_richtext(
+    data = japan_points, 
+    aes(x = lon_adj, y = lat_adj, label = label),
+    size = 2.5,
+    fill = NA, label.color = NA, # remove background and outline
+    label.padding = grid::unit(rep(0, 4), "pt") # remove padding)
+  ) +
+  geom_point(
+    data = filter(japan_points, query == "Tokyo"),
+    aes(x = lon, y = lat)
+  ) +
+  map_theme()
+
+ggsave(result_file(figure("japan-map"), "pdf"), width = 5, height = 5, units = "in")
+```
+
+**`r figure("japan-map")`**. `r figure_cap("japan-map")`
+
+\clearpage
 

### 2020-08-26 ms/captions.R

@@ -2,6 +2,7 @@
 
 # - Figures
 fig_nums <- captioner::captioner(prefix = "Fig.")
+fig_nums(name = "japan-map", caption = "Map of Japan showing names of places mentioned in the text. Names of the four main islands in bold.")
 fig_nums(name = "raw-diversity", caption = "Raw taxonomic (**A**), phylogenetic (**B**), and functional (**C**) diversity of the ferns of Japan.")
 fig_nums(name = "rand-diversity", caption = "Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.")
 fig_nums(name = "endemism", caption = "Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).")

### 2020-09-08 presentations/bsj_2020-09-19/index.Rmd
Start on BSJ presentation

---
output:
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom.css"]
    lib_dir: libs
    mathjax: default
    seal: false
    nature:
      slideNumberFormat: "" # To turn off slide numbers
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["left", "middle", "inverse"]
editor_options:
  chunk_output_type: console
---
class: inverse, middle, title-slide, big-margin
background-color: #000000
background-position: 50% 50%
background-image: url(images/title_background.png)
background-size: contain

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE, eval = TRUE, cache = TRUE, dev = "svg", dev.args=list(bg="transparent"))
```

```{r load-packages-funcs, cache = FALSE}
library(conflicted)
library(magick)
library(ggtext)
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(rcartocolor)
library(tidyverse)

# Resolve conflicts
conflicted::conflict_prefer("map", "purrr")
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
conflicted::conflict_prefer("gather", "tidyr")
conflicted::conflict_prefer("extract", "magrittr")

# Specify custom themes
simple_map_theme <- function () {
  theme(
    panel.grid.major = element_line(color = "white", size = 0.1),
    panel.background = element_rect(fill = "grey70"),
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
}

simple_map_theme_light <- function () {
  theme(
    panel.grid.major = element_line(color = "white", size = 0.1),
    panel.background = element_rect(fill = "grey90"),
    legend.position = "none",
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
}

trend_line_theme <- function () {
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )
}

map_theme_pp <- function() {
  theme(
    panel.grid.major = element_line(color = "white", size = 0.1),
    panel.background = element_rect(fill = "grey70"),
    axis.ticks = element_blank()
  )
}

map_theme_pp_light <- function() {
  theme(
    panel.grid.major = element_line(color = "white", size = 0.1),
    panel.background = element_rect(fill = "grey90"),
    axis.ticks = element_blank()
  )
}

# Set color palettes
# Randomization test significance
signif_cols <-
  c(
    "> 0.99" = "#e31a1c",
    "> 0.975" = "#fb9a99",
    "< 0.01" = "#1f78b4",
    "< 0.025" = "#a6cee3",
    "not significant" = "#ffffcc"
  )

# CANAPE
canape_cols <-
  c(
    "neo" = "#e31a1c",
    "paleo" = "#1f78b4",
    "mixed" = "#6a51a3",
    "super" = "#4a1486",
    "not significant" = "#ffffcc"
  )

# Bioregions
bioregion_cols <- rcartocolor::carto_pal(9, "Bold")[1:8] %>%
  set_names(1:8)

source(here::here("R/functions.R"))
```

```{r, load-refs, includ = FALSE, cache = FALSE}
# Set bibliography options
library(RefManageR)
BibOptions(
  ignore.case = TRUE,
  check.entries = FALSE,
  max.names = 3,
  bib.style = "authoryear", # set style of bibliography as e.g., "Nitta 2020..."
  style = "text" # set style of citations as plain text (no links)
)

# Read in bibliography
bib <- ReadBib(here::here("presentations/bsa_2020-07-27/refs.bib"), check = FALSE)
```

```{r load-data, cache = FALSE}
# Set cache
ja_fern_cache <- new_cache(here::here("ja_fern_cache"))

# Load targets
loadd(list = c(
  "biodiv_ferns_spatial",
  "occ_point_data_ferns",
  "protected_areas",
  "traits_for_dist",
  "comm_ferns",
  "green_list",
  "ppgi",
  "japan_shp"),
  cache = ja_fern_cache)

# Read in pteridophyte GBIF collections per country data from
# Iwasaki lab presentation
collections_by_country <- read_csv(here::here("data_raw/gbif_pterido_coll_by_country.csv")) %>%
  mutate(country = factor(country)) %>%
  mutate(country = fct_reorder(country, n))

# Save currently used data as back-up in case data produced by plan changes
data <- list(
  biodiv_ferns_spatial = biodiv_ferns_spatial,
  occ_point_data_ferns = occ_point_data_ferns,
  protected_areas = protected_areas,
  traits_for_dist = traits_for_dist,
  comm_ferns = comm_ferns,
  green_list = green_list,
  ppgi = ppgi,
  japan_shp = japan_shp,
  collections_by_country = collections_by_country
)

saveRDS(data, here::here("presentations/bsj_2020-09-19/data/data.RDS"))

# Change cluster to factor
biodiv_ferns_spatial <- biodiv_ferns_spatial %>%
  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
  # Add redundancy
  mutate(redundancy = 1 - (richness/abundance))

# Download high-res map of Japan,
# crop to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))

# Set CRS
japan_crs <- sf::st_crs(japan)

biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
```

```{r title-slide-image, eval = FALSE}
# Code to make title slide image, don't need to run when rendering slides
title_theme <- theme_void() +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    legend.position = "none",
     panel.background = element_rect(fill = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

a <- ggplot(biodiv_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  title_theme

b <- ggplot(data = biodiv_ferns_spatial, aes(fill = rpd_signif)) +
  geom_sf(color = "transparent", alpha = 1.0) +
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_blank()
  ) +
  title_theme

c <- ggplot(biodiv_ferns_spatial, aes(fill = phylo_cluster)) +
  geom_sf(color = "transparent") +
  labs(fill = "Cluster") +
  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
  title_theme

title_image <- a + b + c + plot_layout(ncol = 1)

# Write out here, then crop manually
ggsave(
  plot = title_image,
  filename = here::here("presentations/bsj_2020-09-19/images/title.png"),
  dpi = 300)
```

.pull-left-wide[
# 日本のシダ植物多様性の<br>多次元解析
<br>
## 新田ジョエル<sup>1</sup>、海老原淳<sup>2</sup>

<sup>1</sup>東京大学理学系研究科・岩崎研究室<br>
<span style = 'font-size: 80%;'>https://joelnitta.com</span>

<sup>2</sup>国立科学博物館・植物研究部

日本植物学会第84回大会<br><span style = 'font-size: 80%;'>2020.03.01</span>
]

.pull-right-narrow[
<img src="images/title_cropped.png" alt="drawing" width="300"/>
]

---
class: big-margin

## 生物多様性は種数だけではない

種数は生物多様性の複数ある次元の一つだけ

他には

- 系統的多様性

- 機能的多様性

- それぞれの関係性

も考慮しなくてはなりません

<img src="images/not_just_species.png" alt="drawing" width="800"/>

---
class: big-margin

## 生物多様性は種数だけではない

種数は生物多様性の複数ある次元の一つだけ

他には

- 系統的多様性

- 機能的多様性

- それぞれの関係性

も考慮しなくてはなりません。

しかし、これは大量のデータを必要とする作業です。

---
class: big-margin

```{r count-taxa}
# count number of higher taxa (genera and families)
n_fern_taxa <-
green_list %>%
  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
  assert(not_na, genus) %>%
  # Add higher-level taxonomy
  left_join(ppgi, by = "genus") %>%
  assert(not_na, class) %>%
  # Filter to only ferns
  filter(class == "Polypodiopsida") %>%
  summarize(
    n_genus = n_distinct(genus),
    n_family = n_distinct(family)
  )
```

.pull-left[

## 日本のシダ植物：生物多様性のモデルシステムとして

- 約675種 ・`r n_fern_taxa$n_genus`属・`r n_fern_taxa$n_family`科

- 全世界でももっとも研究が進んでいるシダ植物フロラの一つ

- DNA、種の分布、種の機能的特徴も網羅的にサンプリング済
]

.pull-right[
.center[
<img src="images/ja_example_ferns.png" alt="drawing" height="580"/>
]
]

.footnote[.small[Photos: A. Ebihara]]

---
class: big-margin

.pull-left[

## 日本のシダ植物：生物多様性のモデルシステムとして

- 約675種 ・`r n_fern_taxa$n_genus`属・`r n_fern_taxa$n_family`科

- 全世界でももっとも研究が進んでいるシダ植物フロラの一つ

- DNA、種の分布、種の機能的特徴も網羅的にサンプリング済
]

.pull-right[
```{r gbif-colls-per-country}
# Plot showing top 10 pteridophyte collections in GBIF by country
ggplot(collections_by_country, aes(x = country, y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "GBIF pteridophyte \ncollections (top 10)",
    y = "No. collections") +
  scale_y_continuous(labels = scales::comma) +
  ggplot2::theme_bw(base_size = 24) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    panel.grid.major.y = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(colour="black"),
      axis.text.y = ggplot2::element_text(colour="black")
  )
```
]

---
## 日本の生物地理学入門（短縮バージョン）

.pull-left-narrow[
- 北（亜寒帯）から南（亜熱帯）において気候変動が広い

- 標高も変動が広い

- 第三紀のレフュジア（氷河に覆われたことがない）

- 本州は大陸島

- 沖縄、小笠原は海洋島
]

.pull-right-wide[
<img src="images/japan_intro.png" alt="drawing" height="580"/>
]

---
class: big-margin

## 本研究の目的

**日本産シダ類** において、以下のことを把握する：

1. 生物多様性がどのように分布しているのか？

2. 生物多様性がどのような構造を持っているのか？

3. 生物多様性がどれくらい守れているのか？

---
class: big-margin
## データ

```{r summarize-traits}
trait_summary <- make_trait_summary(traits_for_dist)

n_binary <- trait_summary %>% filter(trait_type == "binary") %>% nrow()
n_cont <- trait_summary %>% filter(trait_type == "continuous") %>% nrow()
n_qual <- trait_summary %>% filter(trait_type == "qualitative") %>% nrow()
```

**分布データ**

- 日本の植物標本庫に収納されている標本

- `r nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")` 点 (`r occ_point_data_ferns %>% pull(taxon) %>% n_distinct()` 種)

- 上のリストを0.2°（約20 km x 20 km）有無マトリックスに変換する

---
class: big-margin
## データ

**系統樹**

- 日本の標本から葉緑体遺伝子*rbcL* `r AutoCite(bib, "Ebihara2019b")`

- GenBankにある広いサンプリングと合わせる

- IQ-TREEによる系統解析 (Nguyen et al., 2015) `r NoCite(bib, "Nguyen2015")`

- 27点の化石`r AutoCite(bib, "Testo2016a")` を取り入れたtreePL `r AutoCite(bib, "Smith2012")`による分岐年代推定

- 日本の種だけに部分集合

---
class: big-margin

## データ

**機能的特徴**

- 元々種同定に使われた機能的特徴が`r nrow(trait_summary)`個 `r AutoCite(bib, "Ebihara2019b")`

- 例：包膜の有無、葉柄の長さ、葉の厚さ

---
class: middle, inverse, big-margin

# 生物多様性がどのように分布しているのか？

---
## 種数は山型

```{r centroids}
# Convert diversity metrics dataframe to centroids for plotting trends
ses_div_centroids <-
  biodiv_ferns_spatial %>%
  mutate(
    longitude = sf::st_centroid(geometry) %>% st_coordinates %>% magrittr::extract(,1),
    latitude = sf::st_centroid(geometry) %>% st_coordinates %>% magrittr::extract(,2)) %>%
  as_tibble %>%
  dplyr::select(-geometry)

# Get min and max latitude for setting common limits between
# maps and trendlines
ja_lat_min <- ses_div_centroids %>% pull(latitude) %>% min()
ja_lat_max <- ses_div_centroids %>% pull(latitude) %>% max()
```

```{r richness, out.width = "120%"}
# Fern richness
# - map
a <-
  ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
  geom_sf() +
  scale_fill_scico(palette = "lajolla") +
  labs(fill = "No. taxa") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  theme(
    legend.title = element_text(size = 20/.pt),
    legend.text = element_text(size = 16/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )

# - latitudinal trends
b <- ggplot(ses_div_centroids, aes(x = latitude, y = richness)) +
  geom_point(color = "grey50", shape = 1 ) +
  geom_smooth(se = FALSE) +
  xlim(ja_lat_min, ja_lat_max) +
  coord_flip() +
  labs(y = "No. taxa") +
  trend_line_theme()

a + b + plot_layout(widths = c(3,1), heights = c(1,1))
```

.footnote[Trendline: General additive model (GAM) with cubic splines]

---
class: big-margin
## 系統樹を使った生物多様性の測り方

.center[
<img src="images/pd.png" alt="drawing" height="250"/>
]

<u>**P**</u>hylogenetic <u>**D**</u>iversity: 系統樹の枝長の総計 `r AutoCite(bib, "Faith1992")`

<u>**F**</u>unctional <u>**D**</u>iversity: 測り方はPDと同様が、系統樹の代わりに形態的特徴によるdendrogramを使う `r AutoCite(bib, "Petchey2002")`

---
## PDとFDは種数と相関している

```{r raw-div-1, out.width = "120%", fig.asp = 0.7}
# Richness
a <- ggplot(biodiv_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  labs(title = "Taxonomic richness") +
  simple_map_theme()

b <- ggplot(biodiv_ferns_spatial, aes(fill = pd_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  labs(
    title = "Phylogenetic diversity") +
  simple_map_theme()

c <- ggplot(biodiv_ferns_spatial, aes(fill = fd_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  labs(
    title = "Functional diversity") +
  simple_map_theme()

a + b + c +
  plot_spacer() + plot_spacer() + plot_spacer() +
  plot_layout(nrow = 2, heights = c(0.5, 0.5))
```

---
## PDとFDは種数と相関している（予想通りに）

```{r raw-div-2, out.width = "120%", fig.asp = 0.7}
# Richness
a <- ggplot(biodiv_ferns_spatial, aes(fill = richness)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  labs(title = "Taxonomic richness") +
  simple_map_theme()

b <- ggplot(biodiv_ferns_spatial, aes(fill = pd_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  labs(
    title = "Phylogenetic diversity") +
  simple_map_theme()

c <- ggplot(biodiv_ferns_spatial, aes(fill = fd_obs)) +
  geom_sf(color = "transparent") +
  scale_fill_scico(palette = "lajolla") +
  labs(
    title = "Functional diversity") +
  simple_map_theme()

d <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = pd_obs)) +
  geom_point(shape = 1) +
  # Fit a logarithmic general additive model
  # https://stackoverflow.com/questions/40711980/r-ggplot2-fit-curve-to-scatter-plot
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(log(x)), size = 0.5) +
  labs(
    x = "No. taxa",
    y = "PD"
  ) +
  jntools::standard_theme()

e <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = fd_obs)) +
  geom_point(shape = 1) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(log(x)), size = 0.5) +
  labs(
    x = "No. taxa",
    y = "FD"
  ) +
  jntools::standard_theme()

a + b + c +
  plot_spacer() + d + e +
  plot_layout(nrow = 2, heights = c(0.5, 0.5))
```

.footnote[Trendline: GAM with log-transform]

---
class: big-margin

## 観測値とランダムな値との比較

"Independent Swap" `r AutoCite(bib, "Gotelli2000")`によって999個のランダムな群衆を作成し、観察された値と比較する

p-valueによって意義を評価する

---

## PDは本州に低く、琉球に高い

```{r pd-rand, out.width = "120%", fig.asp = 0.7}
ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = signif_cols) +
  labs(fill = "Phylogenetic diversity<br>*p*-value") +
  map_theme() +
  theme(
    legend.title = element_markdown(),
    legend.position = "right"
  )
```

---
background-position: 40% 50%
background-image: url(images/RPD.png)
background-size: contain

## <u>**R**</u>elative <u>**P**</u>hylogenetic <u>**D**</u>iversityで枝長を比較する

.pull-right[
有意義に長い枝

- レフュージア

- 対象地域以外に主な分布をもつ分類群の存在

有意義に短い枝

- 最近の種分化

.small[Mishler et al. (2014) `r NoCite(bib, "Mishler2014")`]
]

---
## 系統樹の枝は九州に短く、南北に長い

```{r rpd-rand, out.width = "120%", fig.asp = 0.7}
ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = signif_cols) +
  labs(fill = "Relative<br>phylogenetic diversity<br>*p*-value") +
  map_theme() +
  theme(
    legend.title = element_markdown(),
    legend.position = "right"
  )
```

---
## 機能的形質は系統的多様性と似ている

```{r rfd-rand, out.width = "120%", fig.asp = 0.7}
ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = signif_cols) +
  labs(fill = "Relative<br>functional diversity<br>*p*-value") +
  map_theme() +
  theme(
    legend.title = element_markdown(),
    legend.position = "right"
  )
```

---
## 琉球と小笠原は固有種のホットスポット

```{r pe, out.width = "120%", fig.asp = 0.7}
ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = canape_cols) +
  labs(fill = "Endemism type") +
  map_theme() +
  theme(
    legend.title = element_markdown(),
    legend.position = "right"
  )
```

---
class: middle, inverse, big-margin

# 生物多様性がどのような構造を持っているのか？

---
## バイオリージョンをクラスター法で分類する

.center[
<img src="images/daru_2017.png" alt="drawing" height="400"/>
]

.footnote[`r Citet(bib, c("Daru2017", "Daru2020"))`]

---
## ほとんどの地点は四つのバイオリージョンのどれかに所属する

```{r tax-bioregion, out.width = "120%", fig.asp = 0.7}
ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent", alpha = 0.8) +
  labs(fill = "Taxonomic<br>bioregion") +
  scale_fill_manual(values = bioregion_cols) +
  map_theme() +
  theme(
    legend.title = element_markdown(),
    legend.position = "right"
  )
```

---
## 系統的距離を使うと琉球と小笠原が一つのバイオリージョンになる

```{r phy-bioregion, out.width = "120%", fig.asp = 0.7}
ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = phylo_cluster), color = "transparent", alpha = 0.8) +
  labs(fill = "Phylogenetic<br>bioregion") +
  scale_fill_manual(values = bioregion_cols) +
  map_theme() +
  theme(
    legend.title = element_markdown(),
    legend.position = "right"
  )
```

### 2020-09-18 ms/SI.Rmd
Render ms word

@@
     fig_caption: yes
-    keep_tex: no
+    keep_tex: yes
     latex_engine: xelatex

### 2020-09-18 R/functions.R

@@ -40,11 +40,14 @@ unzip_ebihara_2019 <- function (dryad_zip_file, exdir, ...) {
 tidy_japan_names <- function (data) {
   data %>%
     select(taxon_id = ID20160331, scientific_name = `GreenList Name`,
-           endemic = Endemism, conservation_status = RL2012) %>%
+           endemic = Endemism, conservation_status = RL2012,
+           hybrid = Hybrid) %>%
     # Simplify taxon names, replace space with underscore
     taxastand::add_parsed_names(scientific_name, taxon) %>%
     mutate(taxon = str_replace_all(taxon, " ", "_")) %>%
     mutate(taxon_id = as.character(taxon_id)) %>%
+    # Change hybrid to TRUE/FALSE
+    mutate(hybrid = ifelse(is.na(hybrid), FALSE, TRUE)) %>%
     assert(not_na, taxon, taxon_id, scientific_name) %>%
     assert(is_uniq, taxon, taxon_id, scientific_name)
 }
@@ -2341,3 +2344,45 @@ result_file <- function (result_num, extension) {
     fs::path_ext_set(extension)
 }
 
+
+# Dummy function to track arbitary output from rmarkdown::render()
+render_tracked <- function (tracked_output, ...) {
+  rmarkdown::render(...)
+}
+
+#' Convert latex file to docx
+#' 
+#' Requires pandoc to be installed and on command line
+#'
+#' @param latex Path to input latex file.
+#' @param docx Path to output docx file.
+#' @param template Path to template docx file.
+#' @param wd Working directory to run conversion. Should be same as
+#' directory containing any files needed to render latex to pdf.
+#'
+#' @return List including STDOUT of pandoc; externally, the
+#' docx file will be rendered in `wd`.
+#' 
+latex2docx <- function (latex, docx, template = NULL, wd = getwd()) {
+  
+  assertthat::assert_that(assertthat::is.readable(latex))
+  
+  assertthat::assert_that(assertthat::is.dir(fs::path_dir(docx)))
+  
+  latex <- fs::path_abs(latex)
+  
+  docx <- fs::path_abs(docx)
+  
+  template <- if (!is.null(template)) {
+    glue::glue("--reference-doc={fs::path_abs(template)}")
+  } else {
+    NULL
+  }
+  
+  processx::run(
+    command = "pandoc",
+    args = c("-s", latex, template, "-o", docx),
+    wd = wd
+  )
+  
+}

### 2020-09-18 R/plan.R

@@ -327,14 +327,37 @@ plan <- drake_plan (
   signif_cells_protected_area = calculate_protected_area(biodiv_ferns_spatial, protected_areas, japan_shp),
   
   # Write out manuscript ----
-  ms = rmarkdown::render(
-    knitr_in("ms/manuscript.Rmd"),
-    output_dir = here::here("results"),
-    quiet = TRUE),
   
-  si = rmarkdown::render(
-    knitr_in("ms/SI.Rmd"),
+  docx_template_path = target("ms/template.docx", format = "file"),
+  
+  ms_pdf = render_tracked(
+    knitr_in("ms/manuscript.Rmd"),
+    quiet = TRUE,
     output_dir = here::here("results"),
-    quiet = TRUE)
+    tracked_output = here::here("results/manuscript.tex") %>% file_out()
+  ),
+  
+  # Next use the latex to convert to docx with pandoc
+  ms_docx = latex2docx(
+    latex = here::here("results/manuscript.tex") %>% file_in(),
+    docx = here::here("results/manuscript.docx") %>% file_out(),
+    template = docx_template_path,
+    wd = here::here("results")
+  ),
+  
+  # Do same with SI
+  si_pdf = render_tracked(
+    knitr_in("ms/SI.Rmd"),
+    quiet = TRUE,
+    output_dir = here::here("results"),
+    tracked_output = here::here("results/SI.tex") %>% file_out()
+  ),
+  
+  si_docx = latex2docx(
+    latex = here::here("results/SI.tex") %>% file_in(),
+    docx = here::here("results/SI.docx") %>% file_out(),
+    template = docx_template_path,
+    wd = here::here("results")
+  )
   
 )

### 2020-09-18 ms/manuscript.Rmd
Update MS

@@
   "comm_ferns",
-  "japan_points_raw"
+  "japan_points_raw",
+  "green_list"
 ),
@@
 
-The ferns of Japan are one of the best-characterized regional floras in the world due to decades of intensive collection and taxonomic revision.
-We combined phylogenetic, functional, and environmental datasets to explore the spatial distribution of biodiversity in the ferns of Japan.
-We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity, which is highest in small southern islands, and that of relative phylogenetic and functional diversity, which is high at both high and low latitudes, and low at mid-latitudes.
+**Aim** To characterize multiple facets of biodiversity and understand how these relate to bioregions and conservation status.
+
+**Location** Japan
+
+**Taxon** Ferns
+
+**Methods** We used a list of >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
+We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and assessed significance using a null distribution of 999 random communities.
+We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism.
+We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions.
+We compared the overlap of significantly diverse grid-cells with existing conservation zones to assess conservation status.
+
+**Results** We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes.
+Grid-cells were grouped in to three (phylogenetic) or four (taxonomic) major bioregions.
+Conservation status (% protected area) is higher than the overall rate for grid-cells with significantly high biodiversity.
+
+**Main conclusions** The integrative approach used here reveals previously undetected patterns of biodiversity in the ferns of Japan.
+The different distributions of each diversity type reflect the major bioregions.
+Although peak areas of different diversity types are largely non-overlapping, each one has a relatively high protection status.
+Despite this, future conservation efforts must be cognizant of dangers not addressed here such as climate change and herbivory by deer.
 
@@
 
-Our data include all ferns known to occur within the political boundaries of Japan (20° to 45° N, 122° to 153° E).
+Our data include all ferns known to occur within the political boundaries of Japan (20°N to 45°N, 122°E to 153°E).
 Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`).
+The main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ryukyu and Ogasawara archipelagos are oceanic islands that have never been connected to the mainland.
 The climate is subtropical in the south and highly seasonal in other parts of the country.
@@
 Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`).
-The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (hereafter defined as species, subspecies, and varieties unless otherwise indicated).
+The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
 We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).
@@
 ```{r dataset-stats}
+### Taxon list ###
+# Make list of all native ferns in Japan (based off Green List), includes hybrids
+fern_list <-
+green_list %>%
+  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
+  left_join(ppgi) %>%
+  filter(class == "Polypodiopsida")
+
 ### Phylogeny ###
-
-# FIXME: confirm total number of taxa with Ebihara
-total_fern_taxa <- 675
+total_fern_taxa <- fern_list %>%
+  filter(hybrid == FALSE) %>%
+  pull(taxon) %>%
+  n_distinct()
+
 phylogeny_sampled_percent <- ape::Ntip(japan_fern_tree) %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)
@@
 
+```{r distribution-stats}
+# Calculate number of genera and families restricted to lat < 30.1
+southern_genera <-
+  occ_point_data_ferns %>%
+  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
+  group_by(genus) %>%
+  summarize(
+    max_lat = max(latitude),
+    .groups = "drop"
+  ) %>%
+  filter(max_lat < 30.1)
+
+n_southern_genera <- southern_genera %>% pull(genus) %>% n_distinct %>% english2
+
+southern_families <-
+  occ_point_data_ferns %>%
+  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
+  left_join(ppgi, by = "genus") %>%
+  group_by(family) %>%
+  summarize(
+    max_lat = max(latitude),
+    .groups = "drop"
+  ) %>%
+  filter(max_lat < 30.1)
+
+n_southern_families <- southern_families %>% pull(family) %>% n_distinct %>% english2
+```
+
 The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015] and modeled distributions of woody plants [@Fukaya2020].
@@
 
-The overwhelming pattern characterizing phylogenetic and functional diversity is the high diversity of southern islands (Ryukyu and Bonin Isl.).
-This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; _ genera and _ families only occur south of 30.1° latitude (just south of Yakushima Isl.).
+The overwhelming pattern characterizing phylogenetic and functional diversity is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C).
+This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
 The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis, and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity.
@@
 
-Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu.
-The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* and *Ophioglossum*.
-The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera such as *Polystichum* and *Cyrtomimum*.
-
-The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values.
+Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`B).
+The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* and *Botrychium*.
+The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera such as *Dryopteris* and *Deparia*.
+
+The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`).
 Interestingly, the bioregions identified here correspond roughly to major forest types categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014].
@@
 Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions.
-The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl.
+The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl (`r figure("bioregions")`).
 The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
@@
 
-Although the different facets of biodiversity have different distribution patterns, we find that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric.
+Although the different facets of biodiversity have different distribution patterns, we find that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric (`r figure("conserv-status")`).
 While it is not clear why this is so---it is unlikely that ferns were emphasized as targets during the establishment of conservation zones---one possibility (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
@@
 
-There are a number of caveats that must be kept in mind when interpreting these results.
-First, the data include species collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan.
-Furthermore, the collections are not evenually collected throughout the country, but rather some areas have many more specimens than others.
+```{r limits-stats}
+# Summarize occurrences by latitude
+occ_data_by_lat <-
+occ_point_data_ferns %>%
+  group_by(taxon) %>%
+  summarize(
+    min_lat = min(latitude),
+    max_lat = max(latitude)
+  ) %>%
+  mutate(
+    lat_span = max_lat - min_lat
+  )
+
+# Find the species with the widest latitudinal distribution
+taxon_with_widest_lat <- occ_data_by_lat %>%
+  slice_max(lat_span) %>%
+  left_join(green_list, by = "taxon") %>%
+  taxastand::add_parsed_names(sci_name = scientific_name, species, author) %>%
+  mutate(text = glue("*{species}* {author}"))
+
+# Find number of species occurring across areas defined by straits
+n_widespread_taxa <-
+occ_data_by_lat %>%
+  # Filter by approx location of Tokara strait (29) and Tsugaru Strait (41.5)
+  filter(min_lat < 29, max_lat > 41.5) %>%
+  pull(taxon) %>%
+  n_distinct %>%
+  english2
+
+# Check number of hybrid fern taxa
+n_hybrid_fern_taxa <- fern_list %>%
+  filter(hybrid == TRUE) %>%
+  pull(taxon) %>%
+  n_distinct()
+```
+
+There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research.
+First, the data include species collected over a wide time period, so they do not represent the current distribution of ferns in Japan.
+Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (`r s_figure("abundance")`).
 Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.
+
+The extreme evolutionary distinctiveness of southern islands is both a major feature of the Japanese fern flora and a challenge for analysis.
+The tropical lineages occurring in these islands are likely responsible for the comparatively low (\ie, clustered) values of PD observed on Honshu (`r figure("rand-diversity")`A).
+If a more restricted null model was used that did not consider all taxa capable of dispersing anywhere within the study area, different results would likely be obtained for the randomization test [@Mishler2020].
+However, ferns are known to have high dispersal abilities [@Tryon1970; @Barrington1993; @Moran2001].
+In our dataset, the taxon with the greatest latitudinal distribution, `r taxon_with_widest_lat$text`, spans nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N).
+Therefore, it seems reasonable to assume that ferns are capable of dispersing across the entire study area.
+
+One piece of biodiversity that we did not consider here is hybrids.
+The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa.
+Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification.
+Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
+
+Finally, a long under-appreciated aspect of fern ecology is the role of the gametophyte.
+Unlike seed plants, ferns have gametophytes that are capable of growing independently from the sporophyte.
+Futhermore, the two halves of the life cycle have vastly different morphology and ecophysiology, and may even occur over partially to completely disjunct ranges [reviewed in @Pinson2016a].
+Here, as in the vast majority of fern ecology studies, we consider only the sporophyte phase.
+However, recent studies in Japan [@Ebihara2013; @Ebihara2019a] and elsewhere [@Nitta2017] have revealed different patterns in the distribution of gametophytes and sporophytes.
+The comprehensive molecular sampling of sporophytes makes this area ideal for conducting high-throughput DNA barcoding analyses to compare patterns of biodiversity between life cycles in ferns across large spatial scales.
 

### 2020-09-18 presentations/bsj_2020-09-19/index.Rmd
Update BSJ presentation

@@
 )
-
-# Read in bibliography
-bib <- ReadBib(here::here("presentations/bsa_2020-07-27/refs.bib"), check = FALSE)
 ```
@@
   "ppgi",
-  "japan_shp"),
+  "japan_shp",
+  "signif_cells_protected_area"),
   cache = ja_fern_cache)
@@
   mutate(country = fct_reorder(country, n))
+
+# Read in bibliography
+bib <- ReadBib(here::here("presentations/bsa_2020-07-27/refs.bib"), check = FALSE)
+
+# Read in Green List in Japanese
+green_list_ja <- readxl::read_excel(here::here("presentations/bsj_2020-09-19/data/FernGreenListV1.01.xls")) %>%
+  transmute(taxon_id = as.character(ID20160331), name_ja = 新リスト和名)
 
@@
   japan_shp = japan_shp,
-  collections_by_country = collections_by_country
+  collections_by_country = collections_by_country,
+  bib = bib,
+  green_list_ja = green_list_ja,
+  signif_cells_protected_area = signif_cells_protected_area
 )
@@
 saveRDS(data, here::here("presentations/bsj_2020-09-19/data/data.RDS"))
+```
 
+```{r prep-data}
 # Change cluster to factor
@@
 # Set CRS
-japan_crs <- sf::st_crs(japan)
-
-biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
+st_crs(japan) <- 4326
+st_crs(biodiv_ferns_spatial) <- 4326
+st_crs(protected_areas) <- 4326
 ```
@@
 <br>
-## 新田ジョエル<sup>1</sup>、海老原淳<sup>2</sup>
+## 新田ジョエル<sup>1</sup>・Brent Mishler<sup>2</sup><br>岩崎渉<sup>1</sup>・海老原淳<sup>3</sup>
 
-<sup>1</sup>東京大学理学系研究科・岩崎研究室<br>
+<sup>1</sup>東京大学理学系研究科<br>
+
+<sup>2</sup>University of California, Berkeley
+
+<sup>3</sup>国立科学博物館
+
+日本植物学会第84回大会<br><span style = 'font-size: 80%;'>2020.03.01</span>
+
 <span style = 'font-size: 80%;'>https://joelnitta.com</span>
 
-<sup>2</sup>国立科学博物館・植物研究部
-
-日本植物学会第84回大会<br><span style = 'font-size: 80%;'>2020.03.01</span>
 ]
@@
 a + b + plot_layout(widths = c(3,1), heights = c(1,1))
+
+ggsave(here::here("presentations/bsj_2020-09-19/richness.png"), height = 6, width = 10, units = "in")
 ```
@@
 ```{r pe, out.width = "120%", fig.asp = 0.7}
+biodiv_ferns_spatial_simple_endem <-
+  biodiv_ferns_spatial %>%
+  mutate(endem_type = case_when(
+    endem_type == "super" ~ "mixed",
+    TRUE ~ as.character(endem_type)
+  )) %>%
+  mutate(endem_type = factor(endem_type, levels = c("neo", "paleo", "mixed", "not significant")))
+
+# CANAPE
+canape_cols_simple <-
+  c(
+    "neo" = "#e31a1c",
+    "paleo" = "#1f78b4",
+    "mixed" = "#6a51a3",
+    "not significant" = "#ffffcc"
+  )
+
 ggplot() +
   geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
+  geom_sf(data = biodiv_ferns_spatial_simple_endem, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
   scale_fill_manual(values = canape_cols) +
@@
   )
+
+ggsave(here::here("presentations/bsj_2020-09-19/canape_simple.png"), height = 6, width = 6, units = "in")
+
 ```
@@
 ---
-## ほとんどの地点は四つのバイオリージョンのどれかに所属する
+## 主なバイオリージョンが四つある
 
@@
 ---
-## 系統的距離を使うと琉球と小笠原が一つのバイオリージョンになる
+系統的距離を使うと琉球と小笠原が一つのバイオリージョンになる
 
@@
 ```
+
+---
+background-position: 50% 50%
+background-image: url(images/species-background-1.png)
+background-size: contain
+
+## 主なバイオリージョンが四つある
+
+```{r find-top-species}
+# Extract clusters
+clusters <-
+  biodiv_ferns_spatial %>%
+  as_tibble() %>%
+  select(site = "grids", cluster = phylo_cluster, -geometry)
+
+# Convert community df to long format
+comm_ferns_long <-
+  comm_ferns %>%
+  rownames_to_column("site") %>%
+  pivot_longer(names_to = "taxon", values_to = "abun", -site)
+
+# Find top-scoring species (with high occurrence inside one cluster, and low elsewhere)
+top_species <-
+  comm_ferns_long %>%
+  left_join(clusters, by = "site") %>%
+  filter(cluster %in% 1:4) %>%
+  group_by(taxon, cluster) %>%
+  summarize(
+    abun = sum(abun),
+    total = n()
+  ) %>%
+  mutate(
+    # percent abundance within a bioregion
+    perc_abun = abun / total
+  ) %>%
+  arrange(taxon, desc(perc_abun)) %>%
+  select(taxon, cluster, perc_abun) %>%
+  group_by(taxon) %>%
+  # take top two highest percent abundances per species, calculate their difference
+  slice_max(n = 2, order_by = perc_abun, .preserve = TRUE, with_ties = FALSE) %>%
+  mutate(
+    top_cluster = cluster[perc_abun == max(perc_abun)]
+  ) %>%
+  summarize(
+    max_dif = max(perc_abun) - min(perc_abun),
+    cluster = unique(top_cluster)
+  ) %>%
+  # keep only those with greater than 65% difference between top 1st and 2nd
+  # most abundant bioregions
+  filter(max_dif > 0.5) %>%
+  group_by(cluster) %>%
+  # Choose top three per cluster for plotting
+  slice_max(n = 3, order_by = max_dif, with_ties = FALSE) %>%
+  # Add name in Japanese
+  left_join(select(green_list, taxon, taxon_id), by = "taxon") %>%
+  left_join(green_list_ja, by = "taxon_id")
+```
+
+```{r plot-top-species-prep}
+# make helper function to plot top-scoring species
+plot_top_taxon <- function (taxon_select, taxon_print, biodiv_ferns_spatial, comm_ferns_long) {
+biodiv_ferns_spatial %>%
+  left_join(
+    filter(comm_ferns_long, taxon == taxon_select),
+    by = c(grids = "site")
+  ) %>%
+  mutate(abun = factor(abun)) %>%
+  ggplot(aes(fill = abun)) +
+  scale_fill_manual(values = c("#404040", "#d01c8b")) +
+  geom_sf(color = "transparent") +
+  labs(subtitle = taxon_print) +
+  simple_map_theme_light() +
+  theme(plot.subtitle = element_text(family = "HiraKakuPro-W3"))
+}
+```
+
+```{r plot-top-species-cluster-1, out.width = "120%", fig.asp = 0.65}
+focus_region <-
+biodiv_ferns_spatial %>%
+  mutate(focus = ifelse(taxonomic_cluster == 1, "yes", "no")) %>%
+ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  simple_map_theme_light() +
+  labs(subtitle = "Region 1")
+
+focus_region +
+plot_top_taxon(top_species$taxon[[1]], taxon_print = top_species$name_ja[[1]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[2]], taxon_print = top_species$name_ja[[2]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[3]], taxon_print = top_species$name_ja[[3]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+
+ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-cluster-1.png"), height = 6, width = 6, units = "in")
+
+# One species per region
+plot_top_taxon(top_species$taxon[[2]], taxon_print = top_species$name_ja[[2]], biodiv_ferns_spatial, comm_ferns_long) + 
+plot_top_taxon(top_species$taxon[[4]], taxon_print = top_species$name_ja[[4]], biodiv_ferns_spatial, comm_ferns_long) + 
+plot_top_taxon(top_species$taxon[[7]], taxon_print = top_species$name_ja[[7]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+
+ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-one-each-phylo.png"), height = 6, width = 6, units = "in")
+```
+
+.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]
+
+---
+background-position: 50% 50%
+background-image: url(images/species-background-2.png)
+background-size: contain
+
+## 主なバイオリージョンが四つある
+
+```{r plot-top-species-cluster-2, out.width = "120%", fig.asp = 0.65}
+focus_region <-
+biodiv_ferns_spatial %>%
+  mutate(focus = ifelse(taxonomic_cluster == 2, "yes", "no")) %>%
+ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  simple_map_theme_light() +
+  labs(subtitle = "Region 2")
+
+focus_region +
+plot_top_taxon(top_species$taxon[[4]], taxon_print = top_species$name_ja[[4]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[5]], taxon_print = top_species$name_ja[[5]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[6]], taxon_print = top_species$name_ja[[6]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+
+ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-cluster-2.png"), height = 6, width = 6, units = "in")
+```
+
+.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]
+
+---
+background-position: 50% 50%
+background-image: url(images/species-background-3.png)
+background-size: contain
+
+## 主なバイオリージョンが四つある
+
+```{r plot-top-species-cluster-3, out.width = "120%", fig.asp = 0.65}
+focus_region <-
+biodiv_ferns_spatial %>%
+  mutate(focus = ifelse(taxonomic_cluster == 3, "yes", "no")) %>%
+ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  simple_map_theme_light() +
+  labs(subtitle = "Region 3")
+
+focus_region +
+plot_top_taxon(top_species$taxon[[7]], taxon_print = top_species$name_ja[[7]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[8]], taxon_print = top_species$name_ja[[8]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[9]], taxon_print = top_species$name_ja[[9]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+
+ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-cluster-3.png"), height = 6, width = 6, units = "in")
+```
+
+.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]
+
+---
+background-position: 50% 50%
+background-image: url(images/species-background-4.png)
+background-size: contain
+
+## 主なバイオリージョンが四つある
+
+```{r plot-top-species-cluster-4, out.width = "120%", fig.asp = 0.65}
+bonin_box <-
+biodiv_ferns_spatial %>%
+  filter(taxonomic_cluster == 4) %>%
+  sf::st_bbox()
+
+expand_value <- 0.4
+
+focus_region <-
+biodiv_ferns_spatial %>%
+  mutate(focus = ifelse(taxonomic_cluster == 4, "yes", "no")) %>%
+ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  simple_map_theme_light() +
+  labs(subtitle = "Region 4") +
+  annotate(
+    "rect",
+           xmin = bonin_box[["xmin"]] - expand_value,
+           xmax = bonin_box[["xmax"]] + expand_value,
+           ymin = bonin_box[["ymin"]] - expand_value,
+           ymax = bonin_box[["ymax"]] + expand_value,
+    fill = "transparent",
+    color = "grey10")
+
+focus_region +
+plot_top_taxon(top_species$taxon[[10]], taxon_print = top_species$name_ja[[10]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[12]], taxon_print = top_species$name_ja[[12]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_top_taxon(top_species$taxon[[11]], taxon_print = top_species$name_ja[[11]], biodiv_ferns_spatial, comm_ferns_long) +
+plot_layout(ncol = 2, nrow = 2)
+
+ggsave(here::here("presentations/bsj_2020-09-19/plot-top-species-cluster-4.png"), height = 6, width = 6, units = "in")
+```
+
+.footnote[`r NoCite(bib, c("Ebihara2016b", "Ebihara2017"))`]
+
+---
+## バイオリージョンは従来の植生帯と概ね一致
+
+.pull-left[
+Taxonomic bioregions
+
+```{r tax-bioregion-forest}
+ggplot(biodiv_ferns_spatial, aes(fill = taxonomic_cluster)) +
+  geom_sf(color = "transparent") +
+  scale_fill_carto_d(name = "Cluster", type = "qualitative", palette = "Bold") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme_pp_light() +
+  theme(legend.position = "none")
+```
+]
+
+.pull-right[
+植生帯<br>
+.small[`r Citet(bib, "Shimizu2014")`<br><br>]
+<img src="images/forest_types_eng.png" alt="drawing" width="280"/>
+]
+
+---
+class: middle, inverse, big-margin
+
+# 生物多様性はどれくらい守られているのか？
+
+---
+class: middle, inverse, big-margin
+
+# How well is biodiversity protected?
+
+---
+## 日本は一般的によく守られている（平均30%）
+
+```{r protected-areas-graph}
+# Calculate the total area of Japan
+area_of_japan <- 377900
+
+# Calculate baseline protection percentage: percent protected areas across the whole country
+total_percent_protected <-
+  protected_areas %>%
+  as_tibble() %>%
+  select(-geometry) %>%
+  filter(status %in% c("medium", "high")) %>%
+  group_by(status) %>%
+  summarize(
+    total_area = sum(area),
+    .groups = "drop"
+  ) %>%
+  mutate(percent_protected = total_area/area_of_japan) %>%
+  mutate(percent_protected = as.numeric(percent_protected))
+
+# Extract values for medium and high protection level
+total_percent_protected_medium <- total_percent_protected %>% filter(status == "medium") %>% pull(percent_protected)
+total_percent_protected_high <- total_percent_protected %>% filter(status == "high") %>% pull(percent_protected)
+total_percent_protected_all <- sum(total_percent_protected_medium, total_percent_protected_high)
+
+show_col(viridis_pal(option = "D")(3))
+
+protection_cols <- c(
+  Medium = "#21908CFF",
+  High = "#FDE725FF"
+)
+
+# Plot protected areas
+signif_cells_protected_area %>%
+  mutate(
+    metric = case_when(
+      metric == "pd" ~ "PD",
+      metric == "fd" ~ "FD",
+      metric == "pe" ~ "PE",
+      metric == "richness" ~ "Richness",
+    ),
+    status = str_to_sentence(status)) %>%
+  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "Richness"))) %>%
+  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
+  geom_col(position = position_dodge()) +
+  geom_vline(xintercept = total_percent_protected_medium, color = "grey40", linetype = 3) +
+  geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
+  scale_fill_manual(values = protection_cols) +
+  scale_x_continuous(labels = scales::percent) +
+  labs(
+    x = "Percent protected",
+    y = "Biodiversity metric",
+    fill = "Conservation\nstatus"
+  ) +
+  theme_bw(base_size = 18) +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    legend.position = "right"
+  )
+
+ggsave(here::here("presentations/bsj_2020-09-19/protected_bar.png"), height = 6, width = 8, units = "in")
+```

### 2021-04-08 presentations/jsps_2020-03-01/index.Rmd
Update presentation

@@
 library(magick)
-library(gt)
 library(ggtext)

### 2021-04-16 ms/manuscript.Rmd
Remove drake files

@@
 # NOTE: This Rmd is meant to be knit as part of the drake plan (R/plan.R).
-# If using interactively, run `source("_drake.R")` first.
-
-loadd(list = c(
+# If using interactively, run `source("_targets.R")` first.
+
+tar_load(c(
   "occ_point_data_ferns",
@@
   "green_list"
-),
-cache = ja_fern_cache)
-
-phylo_cache <- new_cache(here::here("phylo_cache"))
-
-loadd(list = c(
+))
+
+phylo_cache <- drake::new_cache(here::here("phylo_cache"))
+
+drake::loadd(list = c(
   "plastome_calibration_dates",

### 2021-04-16 _drake.R

@@ -1,32 +1,0 @@
-# Specify non-global environment
-# to get around captioner modifying global env
-# (cf https://github.com/ropensci/drake/issues/749)
-envir <- new.env(parent = globalenv())
-
-# Load packages, functions, and plan
-source("R/packages.R")
-source("R/functions.R", local = envir)
-source("R/plan.R")
-
-# Setup cache
-ja_fern_cache <- new_cache("ja_fern_cache")
-options(rstudio_drake_cache = ja_fern_cache)
-
-# Specify parallel back-end
-options(clustermq.scheduler = "multicore")
-
-# Configure settings for making plan
-# (choose either parallel or serial, comment-out the other)
-
-# - parallel
-drake_config(
-  plan,
-  parallelism = "clustermq",
-  jobs = 6, # Change to match number of cores available!
-  cache = ja_fern_cache,
-  seed = 0,
-  envir = envir
-)
-
-# - serial
-# drake_config(plan, verbose = 1, cache = ja_fern_cache, seed = 0, envir = envir)

### 2021-04-16 phylo_drake.R

@@ -1,11 +1,0 @@
-# Load packages, functions, and plan
-source("R/packages.R")
-source("R/functions.R")
-source("R/phylo_plan.R")
-
-# Setup cache
-phylo_cache <- new_cache("phylo_cache")
-options(rstudio_drake_cache = phylo_cache)
-
-# Specify settings for making plan
-drake_config(plan, verbose = 1, cache = phylo_cache, seed = 0)

### 2021-04-16 ms/manuscript.Rmd
Update MS

@@
 
-Ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are very densely sampled.
+The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled.
+One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for 221/721 taxa [70%]), which is likely to effect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
 Japan is situated along a latitudinal gradient spanning highly season temperate areas in the north to subtropical, mostly aseasonal islands in the south.
@@
 
-### Trait data
+### Morphological trait data
 
@@
 From the original trait list, we selected only traits with putatative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.
+
+### Reproductive mode data
+
+We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown.
+We calculated the percentage of apomictic species in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa.
 
@@
 Statistical significance was assessed using a two-tailed test with alpha = 0.5; observed values > 97.5% or < 0.25% were considered significant.
+We also calculated the standard effect size (SES) of each biodiversity metric.
+SES measures how extreme the observed value (*obs*) is relative to the null distribution (*null*):
+
+$SES = \frac{obs - mean(null)}{sd(null)}$
+
+We tested for a relationship between SES of each biodiversity metric and reproductive mode (% apomictic taxa) initially with linear models constructed using the R function 'lm'.
+Next, we checked for evidence of spatial autocorrelation in the linear model residuals with Moran's I using 'moran.mc' function in the spdep package.
+Spatial autocorrelatoin was frequently detected, so to acccount for this we constructed spatial linear mixed models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014].
+We then again checked for any remaining spatial autocorrelation in the mixed models with Moran's I.
 
@@
 As the optimal number of bioregions cannot be known *a-priori*, we tested values of *k* from 1 to 20.
-
-We compared patterns in biodiversity between bioregions by calculating the standard effect size (SES) of biodiversity metrics.
-SES measures how extreme the observed value (*obs*) is relative to the null distribution (*null*):
-
-$SES = \frac{obs - mean(null)}{sd(null)}$
-
-We compared SES of PD, RPD, FD, and RFD across bioregions using **FIXME: add this when figure out spatial-corrected statistical method to use**.
 
@@
 Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD.
+
+### Correlations between biodiversity and reproductive mode
+
+Spatial autocorrelation as measured with Moran's I was detected in linear models, but was not in any of the spatial mixed linear models.
+SES of RPD in particular shows a strong negative correlation with % apomictic taxa (_ estimated effect _, SD).
 
@@
 Here, we provide the first insights into the combined taxonomic, phylogenetic, and functional diversity of the ferns of Japan.
-Our integrative approach reveals distinct patterns of biodiversity in different parts of the archipelago and across biodiversity metrics.
+Our integrative approach reveals distinct patterns of biodiversity in different parts of the archipelago and across biodiversity metrics, and suggest reproductive mode as an important driver in some cases.
 These clear differences are reflected in the bioregionalization results, which delimit three to four major bioregions.

### 2021-04-23 ms/manuscript.Rmd
Update MS

@@
   "japan_points_raw",
+  "plastome_calibration_dates",
+  "plastome_tree",
   "green_list"
 ))
-
-phylo_cache <- drake::new_cache(here::here("phylo_cache"))
-
-drake::loadd(list = c(
-  "plastome_calibration_dates",
-  "plastome_tree"
-),
-cache = phylo_cache)
 
@@
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled.
-One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for 221/721 taxa [70%]), which is likely to effect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
+One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for 221/721 taxa [70%]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
+
 Japan is situated along a latitudinal gradient spanning highly season temperate areas in the north to subtropical, mostly aseasonal islands in the south.
+Furthermore, it is a mountainous country with great variation in elevation on the larger islands.
+Plant distributions often reflect physiological adaptations to climate (e.g., percipiation and temperature).
+Therefore, we expect the spatial distribution of biodviversity to be determined by climatic variation in addition to reproductive mode.
 
@@
 We calculated the percentage of apomictic species in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa.
+
+### Climate data
+
+We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `getData` function in the R package `raster`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1), temperature seasonality (bio4), annual precipitation (bio12), and precipitation seasonality (bio15).
+We checked for correlation between climatic variables while taking into account spatial autocorrelation using a modified *t*-test with the `modified.ttest` in the `SpatialPack` R package; as a result, we excluded temperature seasonality, which was correlated with mean annual temperature.
+We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
+The R package 'sf' was used for GIS analyses.
 
@@
 # Verify that min lambda >0.95
-assert_that(phy_sig_results %>% pull(lambda) %>% min > 0.95)
+# assert_that(phy_sig_results %>% pull(lambda) %>% min > 0.95)
 # Verify that max K < 0.0.6
-assert_that(phy_sig_results %>% pull(kval) %>% max < 0.06)
+# assert_that(phy_sig_results %>% pull(kval) %>% max < 0.06)
 
@@
 
-All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.06, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
+<!-- All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.06, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
 `r n_negative_d` out of `r nrow(binary_sig_results)` of continuous traits (coded as binary) showed as much or more phylogenetic signal as expected under BM (*D* =< 0); others had *D* > 0 (`r s_table("phy-sig-binary")`).
+-->
 

### 2021-04-23 ms/manuscript.Rmd
Add functions to load climate data from WorldClim

@@
 ```{r plot-prep}
-# Change cluster to factor
-biodiv_ferns_spatial <- biodiv_ferns_spatial %>%
-  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  # Add redundancy
-  mutate(redundancy = 1 - (richness/abundance))
-
-# Download high-res map of Japan,
-# crop to spatial div results area
+# crop high-res map of Japan to spatial div results area
 japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
@@
 japan_crs <- sf::st_crs(japan)
-
 biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
 
-# Set color palettes
-# Randomization test significance
+# Set color palettes:
+# - Randomization test significance
 signif_cols <-
@@
   )
-
-# CANAPE
+# - CANAPE
 canape_cols <-
@@
   )
-
-# Bioregions
+# - Bioregions
 bioregion_cols <- rcartocolor::carto_pal(9, "Bold")[1:8] %>%

### 2021-04-23 R/functions.R

@@ -538,6 +538,61 @@ process_repro_data <- function (repro_data_raw, green_list) {
   
 }
 
+#' Load WorldClim data cropped to the area including Japan
+#'
+#' The WorldClim data first needs to be downloaded with this:
+#' raster::getData("worldclim", download = TRUE, var = "bio", res = 2.5, path = path)
+#'
+#' @param path Path to data downloaded with raster::getData()
+#'
+#' @return Simple features dataframe with four climate variables 
+#' 
+load_ja_worldclim_data <- function(path) {
+  
+  # Load the WorldClim dataset at 2.5 minute resolution, crop it to the area around Japan
+  raster::getData("worldclim", download = FALSE, var = "bio", res = 2.5, path = path) %>%
+    raster::crop(raster::extent(122, 154, 20, 46)) %>%
+    # convert to sf
+    spex::polygonize() %>%
+    # select relevant variables https://www.worldclim.org/data/bioclim.html
+    select(
+      temp = bio1, # Annual Mean Temperature
+      temp_season = bio4, # Temperature Seasonality (standard deviation ×100)
+      precip = bio12, # Annual Precipitation
+      precip_season = bio15 # Precipitation Seasonality (Coefficient of Variation)
+    ) %>%
+    # Set CRS to NA (like other spatial dataframes in this analysis)
+    st_set_crs(NA)
+  
+}
+
+#' Join climate data to occurrence data,
+#' calculate mean climate values per grid cell
+#'
+#' @param shape_ferns Simple features data frame; grid of fern occurrence points
+#' generated by shape_from_comm_scaled_list()
+#' @param ja_climate_data Simple features data frame; climate data in Japan at the
+#' 2.5 minute scale
+#'
+#' @return Tibble with mean values of each climate variable per grid cell
+#' 
+calc_mean_climate <- function(shape_ferns, ja_climate_data) {
+  
+  # Check for correctly formatted columns
+  assert_that("grids" %in% colnames(shape_ferns))
+  assert_that("temp" %in% colnames(ja_climate_data))
+  
+  shape_ferns %>%
+    select(grids) %>%
+    st_join(ja_climate_data) %>%
+    group_by(grids) %>%
+    mutate(across(c(temp, temp_season, precip, precip_season), ~mean(., na.rm = TRUE))) %>%
+    ungroup %>%
+    select(grids, temp, temp_season, precip, precip_season) %>%
+    unique() %>%
+    st_drop_geometry()
+}
+
[…]

### 2021-04-23 _targets.R

@@ -256,6 +256,15 @@ tar_plan(
   # Make trait distance matrix using taxon IDs as labels
   trait_distance_matrix = make_trait_dist_matrix(traits_for_dist),
   
+  # Climate data ----
+  
+  # Load climate data downloaded from WorldClim database
+  tar_file(worldclim_dir, "data_raw/world_clim"),
+  ja_climate_data = load_ja_worldclim_data(worldclim_dir),
+  
+  # Calculate mean climate values in each grid cell
+  mean_climate = calc_mean_climate(shape_ferns, ja_climate_data),
+  
   # Randomization tests of diversity metrics ----
   
   # Make list of communities for looping
@@ -371,7 +380,12 @@ tar_plan(
     left_join(regions_taxonomy %>% rename(taxonomic_cluster = cluster), by = "grids") %>%
     left_join(regions_phylogeny %>% rename(phylo_cluster = cluster), by = "grids") %>%
     categorize_endemism() %>%
-    categorize_signif(),
+    categorize_signif() %>%
+    # Format factors
+    mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+    mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
+    # Add redundancy
+    mutate(redundancy = 1 - (richness/abundance)),
   
   # - Only those with reproductive mode data available
   biodiv_ferns_repro_spatial =
@@ -539,6 +553,14 @@ tar_plan(
   ),
   
   # Calculate percent protection for grid cells with significantly high biodiversity
-  signif_cells_protected_area = calculate_protected_area(biodiv_ferns_spatial, protected_areas, japan_shp)
+  signif_cells_protected_area = calculate_protected_area(biodiv_ferns_spatial, protected_areas, japan_shp),
+  
+  # Render manuscript
+  tar_render(
+    ms,
+    path = "ms/manuscript.Rmd",
+    output_dir = "results",
+    
+  )
   
 )

### 2021-04-27 ms/manuscript.Rmd
Update ref processing script

@@
 title: "Exploring dimensions of biodiversity in Japanese ferns"
-bibliography: "main_library.bib"
+bibliography: 
+  - references.bib
+  - references_other.yaml
 mainfont: Roboto-Regular.ttf
@@
 We measured functional diversity (FD) following the method of @Petchey2002.
-We first calculated functional distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R package \package{FD} [@Laliberte2011].
+We first calculated functional distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R package \package{FD} [@Laliberte2010].
 We weighted continuous and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled continuous traits prior to calculating distances.

### 2021-05-18 ms/manuscript.Rmd
Update MS

@@
 **Methods** We used a list of >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
-We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and assessed significance using a null distribution of 999 random communities.
+We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode.
 We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism.
@@
 
-Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences) and analytical tools have only become available relatively recently.
+Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences) and analytic tools have only become available relatively recently.
 These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism.
@@
 
-Japan is situated along a latitudinal gradient spanning highly season temperate areas in the north to subtropical, mostly aseasonal islands in the south.
+Japan is situated along a latitudinal gradient spanning seasonal, temperate areas in the north to subtropical, mostly aseasonal islands in the south.
 Furthermore, it is a mountainous country with great variation in elevation on the larger islands.
-Plant distributions often reflect physiological adaptations to climate (e.g., percipiation and temperature).
-Therefore, we expect the spatial distribution of biodviversity to be determined by climatic variation in addition to reproductive mode.
+Plant distributions often reflect physiological adaptations to climate (e.g., precipitation and temperature).
+Therefore, we expect the spatial distribution of biodiversity to be determined by climatic variation in addition to reproductive mode.
 
@@
 The main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ryukyu and Ogasawara archipelagos are oceanic islands that have never been connected to the mainland.
-The climate is subtropical in the south and highly seasonal in other parts of the country.
+The climate is subtropical in the south and temperate in other parts of the country.
 Map data were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).
@@
 We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species).
-At each grain size, species occurrences were converted into a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell).
-We calculated sampling redundance (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling.
+At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell).
+We calculated sampling redundancy (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling.
 Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`).
@@
 All qualitative traits were scored in binary format; \eg, a single trait with three states was formatted as three binary traits.
-From the original trait list, we selected only traits with putatative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.
+From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.
 
@@
 
-We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `getData` function in the R package `raster`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1), temperature seasonality (bio4), annual precipitation (bio12), and precipitation seasonality (bio15).
-We checked for correlation between climatic variables while taking into account spatial autocorrelation using a modified *t*-test with the `modified.ttest` in the `SpatialPack` R package; as a result, we excluded temperature seasonality, which was correlated with mean annual temperature.
+We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `getData` function in the R package `raster`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15).
 We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
@@
 
-We tested for a relationship between SES of each biodiversity metric and reproductive mode (% apomictic taxa) initially with linear models constructed using the R function 'lm'.
-Next, we checked for evidence of spatial autocorrelation in the linear model residuals with Moran's I using 'moran.mc' function in the spdep package.
-Spatial autocorrelatoin was frequently detected, so to acccount for this we constructed spatial linear mixed models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014].
-We then again checked for any remaining spatial autocorrelation in the mixed models with Moran's I.
-
 We tested whether endemic areas have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
@@
 To account for this, we also calculated all endemism-related metrics using a dataset including only species endemic to Japan.
+
+### Spatial models
+
+We first tested for spatial autocorrelation in biodiversity metrics and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `moran.mc` function in the `spdep` package.
+In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05).
+Therefore, we used spatially explicit methods for modeling.
+
+We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `modified.ttest` in the `SpatialPack` R package.
+This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature.
+Therefore, we excluded temperature seasonality from modeling.
+We suspected that despite the correlation, % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we conducted analyses separately for this variable as described below.
+
+We constructed a set of linear models for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) as dependent on all possible combinations of environmental variables (temperature, precipitation, and precipitation seasonality).
+Initial inspection of the relationship between richness and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature for the richness model only.
+We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014], then identified the best-fitting model for each response variable as the one with the greatest log-likelihood.
+We verified the significance of each term of the best model by conducting likelihood ratio tests (LRTs) between the full model and a model with the term of interest removed, in sequence.
+We also verified the significance of the full model relative to a null model (only including the Matérn correlation matrix) using LRTs.
+
+To compare the model fit between using temperature (environmental model) or % apomictic taxa (reproductive model) in response variables with a genetic component (SES of PD, SES of RPD, PE p-score), we constructed additional models where % apomictic taxa was used in place of temperature in the best-fitting model for each response variable.
+We then conducted LRTs between the environmental model vs.
+the reproductive model, as well as between the full reproductive model and models with each term of interest removed, and between the full reproductive model vs.
+the null model.
 

### 2021-05-18 ms/manuscript.Rmd
Temporarily comment-out plastome tree stuff until phylo analysis is working

@@
   "plastome_calibration_dates",
-  "plastome_tree",
+  # FIXME: remove until phylo analysis is working "plastome_tree",
   "green_list"
@@
 neo-endemic areas (defined in units of time), we require an ultrametric tree.
-To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total).
+To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (
+<!-- FIXME: remove until phylo analysis is working  `r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` -->
+OTUs total).
 We conducted phylogenetic analysis with IQTREE [@Nguyen2015] and molecular dating with treePL [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a.

### 2021-05-19 ms/manuscript.Rmd
Update MS

@@
 The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness.
-Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots and motivating conservation strategies to preserve them.
+Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots [@Myers2000] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006].
 However, richness alone cannot provide a complete picture of biodiversity.
@@
 
-However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data or can be lead astray.
-
+However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale.
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled.

### 2021-05-19 ms/manuscript.Rmd
Update MS

@@
 We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014], then identified the best-fitting model for each response variable as the one with the greatest log-likelihood.
-We verified the significance of each term of the best model by conducting likelihood ratio tests (LRTs) between the full model and a model with the term of interest removed, in sequence.
-We also verified the significance of the full model relative to a null model (only including the Matérn correlation matrix) using LRTs.
-
-To compare the model fit between using temperature (environmental model) or % apomictic taxa (reproductive model) in response variables with a genetic component (SES of PD, SES of RPD, PE p-score), we constructed additional models where % apomictic taxa was used in place of temperature in the best-fitting model for each response variable.
+We verified the significance of each term of the best model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence.
+We also verified the significance of the full model relative to the null model (only including the Matérn correlation matrix) using LRTs.
+
+To compare between the models fit using temperature (environmental model) vs.
+% apomictic taxa (reproductive model) in response variables with a genetic component (SES of PD, SES of RPD, PE p-score), we constructed additional models where % apomictic taxa was used in place of temperature in the best-fitting model for each response variable.
 We then conducted LRTs between the environmental model vs.

### 2021-05-19 ms/manuscript.Rmd
Update MS

@@
 
-We constructed a set of linear models for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) as dependent on all possible combinations of environmental variables (temperature, precipitation, and precipitation seasonality).
+We constructed a linear model for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) dependent on all environmental variables (temperature, precipitation, and precipitation seasonality).
+We did not include any interaction terms as we had no a-priori reason to do so.
 Initial inspection of the relationship between richness and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature for the richness model only.
-We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014], then identified the best-fitting model for each response variable as the one with the greatest log-likelihood.
-We verified the significance of each term of the best model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence.
+We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014].
+We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence.
 We also verified the significance of the full model relative to the null model (only including the Matérn correlation matrix) using LRTs.
@@
 To compare between the models fit using temperature (environmental model) vs.
-% apomictic taxa (reproductive model) in response variables with a genetic component (SES of PD, SES of RPD, PE p-score), we constructed additional models where % apomictic taxa was used in place of temperature in the best-fitting model for each response variable.
-We then conducted LRTs between the environmental model vs.
-the reproductive model, as well as between the full reproductive model and models with each term of interest removed, and between the full reproductive model vs.
-the null model.
+% apomictic taxa (reproductive model) in response variables with a genetic component (SES of PD, SES of RPD, PE p-score), we constructed additional models where % apomictic taxa was used in place of temperature.
+We then conducted LRTs as described above, as well as between the environmental model vs.
+the reproductive model.
 

### 2021-05-19 ms/manuscript.Rmd
WIP: temporarily comment-out phylogeny parts of MS

@@
   "japan_points_raw",
-  "plastome_calibration_dates",
+  # "plastome_calibration_dates",
   # FIXME: remove until phylo analysis is working "plastome_tree",
@@
 To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (
-<!-- FIXME: remove until phylo analysis is working  `r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` -->
+<!-- FIXME: remove until phylo analysis is working  `r #ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` -->
 OTUs total).
-We conducted phylogenetic analysis with IQTREE [@Nguyen2015] and molecular dating with treePL [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a.
+We conducted phylogenetic analysis with IQTREE [@Nguyen2015] and molecular dating with treePL [@Smith2012] using `r #nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a.
 We then trimmed the tree to Japanese taxa only and used this for all further analyses.

### 2021-05-26 ms/manuscript.Rmd
Update MS

@@
 This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature.
-Therefore, we excluded temperature seasonality from modeling.
-We suspected that despite the correlation, % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we conducted analyses separately for this variable as described below.
-
-We constructed a linear model for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) dependent on all environmental variables (temperature, precipitation, and precipitation seasonality).
+The dataset including % apomictic taxa is smaller than the full dataset (_ vs _ taxa, respectively), so for the main modeling analysis we only included the following environmental variables: temperature, precipitation, and precipitation seasonality.
+
+We constructed a linear model for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) dependent on the environmental variables.
 We did not include any interaction terms as we had no a-priori reason to do so.
 Initial inspection of the relationship between richness and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature for the richness model only.
-We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell with the R package spaMM [@Rousset2014].
+We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell as a random effect with the R package spaMM [@Rousset2014].
 We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence.
@@
 
-To compare between the models fit using temperature (environmental model) vs.
-% apomictic taxa (reproductive model) in response variables with a genetic component (SES of PD, SES of RPD, PE p-score), we constructed additional models where % apomictic taxa was used in place of temperature.
-We then conducted LRTs as described above, as well as between the environmental model vs.
-the reproductive model.
+We postulated that % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis.
+These models included response variables with a genetic component (SES of PD, SES of RPD, PE p-score) as dependent on either % apomictic taxa or temperature, precipitation, and precipitation seasonality, with random spatial effects as described above.
+We compared the fit between models including % apomictic taxa or temperature using conditional Akaike Information Criterion (cAIC).
 

### 2021-05-27 ms/manuscript.Rmd
Parameterize markdown to render to doc and pdf separately

@@
 title: "Exploring dimensions of biodiversity in Japanese ferns"
-bibliography: 
-  - references.bib
-  - references_other.yaml
 mainfont: Roboto-Regular.ttf
@@
   - BoldItalicFont=Roboto-BoldItalic.ttf
+bibliography: 
+  - references.yaml
+  - references_other.yaml
 output: 
+  officedown::rdocx_document:
+    base_format: bookdown::word_document2
+    pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
+    toc: no
+    reference_num: false
+    reference_docx: template.docx
   bookdown::pdf_document2:
-    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
+    citation_package: default
     toc: no
@@
     fig_caption: yes
-    keep_tex: yes
     latex_engine: xelatex
     pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
-    includes:
-      in_header: newphyt.sty
-header-includes: 
-  \usepackage{float}
-  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
+params:
+  doc_type: doc
 editor_options: 
   chunk_output_type: console
-# nocite: | 
 ---
@@
 ```{r setup, include=FALSE}
+library(officedown)
+library(officer)
+library(flextable)
+library(kableExtra)
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
-```
-
-```{r load-data, cache = FALSE}
-# NOTE: This Rmd is meant to be knit as part of the drake plan (R/plan.R).
-# If using interactively, run `source("_targets.R")` first.
-
+
+# Run these lines if rendering outside of targets plan
+# library(targets)
+# library(tarchetypes)
+# source(here::here("R/packages.R"))
+# source(here::here("R/functions.R"))
+```
+
+```{r macros}
+# Define formattting for some common custom words that may vary in style between journals
+ie <- "*ie*"
+eg <- "*eg*"
+# R packages
+pack <- function(x) {glue::glue("'{x}'")}
+# R functions
+func <- function(x) {glue::glue("'{x}'")}
+# Pagebreak for MS Word (doc) only
+pagebreak_doc <- function() {ifelse(params$doc_type == "doc", run_pagebreak(), return(""))}
+# Pagebreak for pdf only
+pagebreak_pdf <- function() {ifelse(params$doc_type == "pdf", return("\\clearpage"), return(""))}
+# Pagebreak for both
+pagebreak <- function() {
+  if (params$doc_type == "doc") {
+    run_pagebreak()
+  } else if (params$doc_type == "pdf") {
+    return("\\clearpage")
+  } else {
+    stop("doc_type parameter must be 'doc' or 'pdf'")
+  }
+}
+```
+
+```{r load-data}
 tar_load(c(
@@
 ))
-
-# Load captions
-source(here::here("ms/captions.R"))
 ```
@@
 
-\clearpage
+`r pagebreak()`
 
@@
 Continuous traits were measured on 10 randomly selected specimens per species.
-Fertile (\ie, spore-bearing) and sterile fronds were measured separately, then averaged.
+Fertile (`r ie`, spore-bearing) and sterile fronds were measured separately, then averaged.
 Qualitative traits were scored by observing voucher specimens.
-All qualitative traits were scored in binary format; \eg, a single trait with three states was formatted as three binary traits.
+All qualitative traits were scored in binary format; `r eg`, a single trait with three states was formatted as three binary traits.
 From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.
@@
 
-We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `getData` function in the R package `raster`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15).
+We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15).
 We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
-The R package 'sf' was used for GIS analyses.
+The R package `r pack("sf")` was used for GIS analyses.
 
@@
 Both lambda and *K* = 1 under BM; lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering).
-We measured *K* and lambda using the \function{phylosig} function in the R package \package{phytools} [@Revell2012], which also calculates significance based on a randomization test for *K* and a likelihood ratio test for lambda.
-We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the \function{phylo.d} function in the R package \package{caper} [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
+We measured *K* and lambda using the `r func("phylosig")` function in the R package `r pack("phytools")` [@Revell2012], which also calculates significance based on a randomization test for *K* and a likelihood ratio test for lambda.
+We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the `r func("phylo.d")` function in the R package `r pack("caper")` [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
 
@@
 
-We measured phylogenetic diversity (PD) using the Faith's PD (the total branch length connecting all terminal taxa in a community, including the root of the phylogeny) [@Faith1992] with the \function{PD} function in the R package \package{phyloregion}.
+We measured phylogenetic diversity (PD) using the Faith's PD (the total branch length connecting all terminal taxa in a community, including the root of the phylogeny) [@Faith1992] with the `r func("PD")` function in the R package `r pack("phyloregion")`.
 To detect areas with unusually long or short branch lengths, we measured relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been set to equal length [@Mishler2014].
@@
 We measured functional diversity (FD) following the method of @Petchey2002.
-We first calculated functional distances using Gower's method [@Gower1971] on the trait data with the \function{gowdis} function in the R package \package{FD} [@Laliberte2010].
+We first calculated functional distances using Gower's method [@Gower1971] on the trait data with the `r func("gowdis")` function in the R package `r pack("FD")` [@Laliberte2010].
 We weighted continuous and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled continuous traits prior to calculating distances.
-We then used the distances to build a trait dendrogram using the \function{hclust} function in the \package{stats} R package [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
+We then used the distances to build a trait dendrogram using the `r func("hclust")` function in the `r pack("stats")` R package [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
 These functional analogs of PD and RPD are hereafter referred to as FD (functional diversity) and RFD (relative functional diversity).
 
-We measured phylogenetic endemism (PE) [@Rosauer2009], which is the sum of all branches at a site inversely weighted by their range, using the \function{phylo{\_}endemism} function in the R package \package{phyloregion}.
+We measured phylogenetic endemism (PE) [@Rosauer2009], which is the sum of all branches at a site inversely weighted by their range, using the `r func("phylo_endemism")` function in the R package `r pack("phyloregion")`.
 
@@
 
-We analyzed bioregions using the framework of @Daru2020 and their R package \package{phyloregion}.
+We analyzed bioregions using the framework of @Daru2020 and their R package `r pack("phyloregion")`.
 Briefly, this involves first calculating beta diversity (change in composition between communities) using either species names (taxonomic bioregions) or phylogenetic distances (phylogenetic bioregions).
@@
 
-Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan.
+Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`A).
 A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Amami, Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula).
 The area near Yakushima includes cells with both significantly high and low PD.
-Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu.
+Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu (`r figure("rand-diversity")`B).
 Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido.
-Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD.
+Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD (`r figure("rand-diversity")`C, D).
 
@@
 
-Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere.
+Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`).
 The vast majority of endemic grid-cells are mixed, with very few paleo- or neoendemic cells.
@@
 The extreme evolutionary distinctiveness of southern islands is both a major feature of the Japanese fern flora and a challenge for analysis.
-The tropical lineages occurring in these islands are likely responsible for the comparatively low (\ie, clustered) values of PD observed on Honshu (`r figure("rand-diversity")`A).
+The tropical lineages occurring in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on Honshu (`r figure("rand-diversity")`A).
 If a more restricted null model was used that did not consider all taxa capable of dispersing anywhere within the study area, different results would likely be obtained for the randomization test [@Mishler2020].
@@
 
-\clearpage
-
-<!-- Tables -->
-
-```{r traits-used, results = "asis"}
+`r pagebreak()`
+
+```{r traits-used-prep}
+# Read in trait metadata
 trait_signif <- read_csv(here::here("data_raw/trait_functions.csv"))
 
-traits_summary %>%
+# Format traits to look nice
+traits_table <- traits_summary %>%
   left_join(trait_signif, by = "trait") %>%
@@
   ) %>%
-  select(trait, Significance = significance) %>%
+  select(trait, Significance = significance)
+```
+
+```{r traits-used-pdf, results = "asis", eval = params$doc_type == "pdf"}
+# Print trait table to pdf
+traits_table %>%
   column_to_rownames("trait") %>%
   knitr::kable(format = "latex", booktabs = TRUE, escape = TRUE) 
-
-```
-
-**`r table("traits-used")`**. `r table_cap("traits-used")`
-
-\clearpage
-
-<!-- Figures -->
+```
+
+```{r traits-used-doc, results = "asis", eval = params$doc_type == "doc"}
+# Print trait table to doc
+traits_table %>%
+  flextable() %>%
+  set_header_labels(trait = "") %>%
+  width(1, width = 1.25) %>%
+  width(2, width = 4)
+```
+
+**`r table("traits-used")`**. Fern traits used in this study.
+
+`r pagebreak()`
 
@@
 
-```{r japan-map, fig.width = 5, fig.height = 5}
-# Print map of Japan with labels of relevant places
+```{r japan-map-save}
+# Plot map of Japan with labels of relevant places
 
@@
 # Make map
-ggplot(japan) + 
+japan_map_plot <- ggplot(japan) + 
   geom_sf() +
@@
 
-ggsave(result_file(figure("japan-map"), "pdf"), width = 5, height = 5, units = "in")
-```
-
-**`r figure("japan-map")`**. `r figure_cap("japan-map")`
-
-\clearpage
-
-```{r raw-diversity, width = 7, height = 4}
+ggsave(result_file(figure("japan-map"), "pdf"), japan_map_plot, width = 5, height = 5, units = "in")
+```
+
+```{r japan-map-print, fig.width = 5, fig.height = 5, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+japan_map_plot
+```
+
+**`r figure("japan-map")`**. Map of Japan showing names of places mentioned in the text. Names of the four main islands in bold.
+
+`r pagebreak_pdf()`
+
+```{r raw-diversity-save}
 # Richness
@@
 
-a + b + c +
+raw_diversity_plot <- a + b + c +
   plot_layout(nrow = 1)
 
-ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7, height = 4, units = "in")
+ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7, height = 4, units = "in", raw_diversity_plot)
+```
+
+```{r raw-diversity-print, fig.width = 7, fig.height = 4, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+raw_diversity_plot
 ```
@@
 **`r figure("raw-diversity")`**.
-`r figure_cap("raw-diversity")` For **B** and **C**, raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
-
-\clearpage
-
-```{r rand-diversity, fig.width = 7, fig.height = 9}
+Raw taxonomic (**A**), phylogenetic (**B**), and functional (**C**) diversity of the ferns of Japan.
+For **B** and **C**, raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+
+`r pagebreak_pdf()`
+
+```{r rand-diversity-save}
 # PD
@@
 
-a + b + c + d + 
+rand_diversity_plot <- a + b + c + d + 
   plot_layout(ncol = 2, nrow = 2, guides = 'collect') &
@@
 
-ggsave(result_file(figure("rand-diversity"), "pdf"), width = 7, height = 8, units = "in")
+ggsave(result_file(figure("rand-diversity"), "pdf"), rand_diversity_plot, width = 7, height = 8, units = "in")
+```
+
+```{r rand-diversity-print, fig.width = 7, fig.height = 9, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+rand_diversity_plot
 ```
@@
 **`r figure("rand-diversity")`**.
-`r figure_cap("rand-diversity")` **A**, Phylogenetic diversity (PD).
+Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.
+**A**, Phylogenetic diversity (PD).
 **B**, Relative phylogenetic diveristy (RPD).
@@
 
-\clearpage
-
-```{r endemism, fig.width = 5, fig.height = 5}
-ggplot() +
+`r pagebreak_pdf()`
+
+```{r endemism-save, fig.width = 5, fig.height = 5}
+endemism_plot <- ggplot() +
   geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
@@
 
-ggsave(result_file(figure("endemism"), "pdf"), width = 5, height = 5, units = "in")
+ggsave(result_file(figure("endemism"), "pdf"), endemism_plot, width = 5, height = 5, units = "in")
+```
+
+```{r endemism-print, fig.width = 5, fig.height = 5, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+endemism_plot
 ```
@@
 **`r figure("endemism")`**.
-`r figure_cap("endemism")` 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).
+'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
 For details, see Methods.
 
-\clearpage
-
-```{r bioregions, fig.width = 7, fig.height = 5}
+`r pagebreak_pdf()`
+
+```{r bioregions-save}
 # Taxonomic bioregions
@@
 
-a + b +
+bioregions_plot <- a + b +
   plot_layout(ncol = 2)
 
-ggsave(result_file(figure("bioregions"), "pdf"), width = 7, height = 5, units = "in")
+ggsave(result_file(figure("bioregions"), "pdf"), bioregions_plot, width = 7, height = 5, units = "in")
+```
+
+```{r bioregions-print, fig.width = 7, fig.height = 5, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+bioregions_plot
 ```
@@
 **`r figure("bioregions")`**.
-`r figure_cap("bioregions")` **A**, taxonomic bioregions.
+Bioregions of the ferns of Japan.
+**A**, taxonomic bioregions.
 **B**, phylogenetic bioregions.
-Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances into *k* = 8 or 7 groups, respectively, using the R package \package{phyloregion}.
-
-\clearpage
-
-```{r div-by-tax-region, fig.width = 7, fig.height = 7}
+Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances into *k* = 8 or 7 groups, respectively, using the R package `r pack("phyloregion")`.
+
+`r pagebreak_pdf()`
+
+```{r div-by-tax-region-save}
 data <-
@@
 
-a + b + c + d + e +
+div_by_tax_region_plot <- a + b + c + d + e +
   plot_layout(ncol = 2) +
@@
 
-ggsave(result_file(figure("div-by-tax-region"), "pdf"), width = 7, height = 7, units = "in")
-
+ggsave(result_file(figure("div-by-tax-region"), "pdf"), div_by_tax_region_plot, width = 7, height = 7, units = "in")
+```
+
+```{r div-by-tax-region-print, fig.width = 7, fig.height = 7, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+div_by_tax_region_plot
 ```
@@
 **`r figure("div-by-tax-region")`**.
-`r figure_cap("div-by-tax-region")` **A**, Standard effect size (SES) of phylogenetic diversity (PD).
+Phylogenetic and morphological diversity of the ferns of Japan by bioregion.
+**A**, Standard effect size (SES) of phylogenetic diversity (PD).
 **B**, SES of relative PD (RPD).
@@
 
-\clearpage
-
-```{r conserv-status, fig.width = 4, fig.height = 4}
+`r pagebreak_pdf()`
+
+```{r conserv-status-save, fig.width = 4, fig.height = 4}
 # Plot protected areas
-signif_cells_protected_area %>%
+conserv_status_plot <- signif_cells_protected_area %>%
   mutate(
@@
 
-ggsave(result_file(figure("conserv-status"), "pdf"), width = 4, height = 4, units = "in")
-
+ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 4, height = 4, units = "in")
+```
+
+```{r conserv-status-print, fig.width = 4, fig.height = 4, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+conserv_status_plot
 ```
@@
 **`r figure("conserv-status")`**.
-`r figure_cap("conserv-status")` Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), relative PD (RPD), relative FD (RFD), and phylogenetic endemism (PE).
+Percent of land area with protected status for grid cells with significantly high biodiversity.
+Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), relative PD (RPD), relative FD (RFD), and phylogenetic endemism (PE).
 Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Methods).

### 2021-05-27 R/functions.R

@@ -3099,6 +3099,37 @@ latex2docx <- function (latex, docx, template = NULL, wd = getwd()) {
   
 }
 
+# Captions ----
+
+# Follow Journal of Plant Research (JPR) style:
+# - All figures are to be numbered using Arabic numerals.
+# - Figures should always be cited in text in consecutive numerical order.
+# - Figure parts should be denoted by lowercase letters (a, b, c, etc.).
+# - If an appendix appears in your article and it contains one or more figures, 
+#   continue the consecutive numbering of the main text. 
+# - Do not number the appendix figures, "A1, A2, A3, etc." 
+# - Figures in online appendices (Electronic Supplementary Material) should, 
+#   however, be numbered separately
+
+# - First define the "full" version, which would include a caption
+# (except I never use the caption in the function, and instead replace with 'blank')
+figure_full <- captioner::captioner(prefix = "Fig.", suffix = "")
+table_full <- captioner::captioner(prefix = "Table")
+s_figure_full <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = "")
+s_table_full <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = "")
+
+# - Make a short function that prints only the object type and number, e.g., "Fig. 1"
+figure <- pryr::partial(figure_full, display = "cite", caption = "blank")
+table <- pryr::partial(table_full, display = "cite", caption = "blank")
+s_figure <- pryr::partial(s_figure_full, display = "cite", caption = "blank")
+s_table <- pryr::partial(s_table_full, display = "cite", caption = "blank")
+
+# - Make a short function that prints only the number (e.g., "1")
+figure_num <- function (name) {figure(name) %>% str_remove("Fig. ")}
+table_num <- function (name) {table(name) %>% str_remove("Table ")}
+s_figure_num <- function (name) {s_figure(name) %>% str_remove("Fig. ")}
+s_table_num <- function (name) {s_table(name) %>% str_remove("Table ")}
+
 # Tests ----
 
 # Run tests to make sure custom functions work as expected

### 2021-05-27 _targets.R

@@ -456,7 +456,7 @@ tar_plan(
   ),
   
   ## Analyze Moran's I ----
-
+  
   # Make list of distances for run_moran_mc()
   # - environmental dataset
   dist_list_env = make_dist_list(biodiv_ferns_cent_env),
@@ -484,7 +484,7 @@ tar_plan(
     ),
     pattern = map(data_for_moran)
   ),
-
+  
   ## Spatial models ----
   
   # Prepare datasets for looping
@@ -633,11 +633,22 @@ tar_plan(
   signif_cells_protected_area = calculate_protected_area(biodiv_ferns_spatial, protected_areas, japan_shp),
   
   # Render manuscript
+  # - doc
   tar_render(
-    ms,
+    ms_doc,
     path = "ms/manuscript.Rmd",
-    output_dir = "results",
-    
+    output_format = "officedown::rdocx_document",
+    knit_root_dir = here::here(),
+    output_file = here::here("results/manuscript.docx"),
+    params = list(doc_type = "doc")
+  ),
+  # - pdf
+  tar_render(
+    ms_pdf,
+    path = "ms/manuscript.Rmd",
+    output_format = "bookdown::pdf_document2",
+    knit_root_dir = here::here(),
+    output_file = here::here("results/manuscript.pdf"),
+    params = list(doc_type = "pdf")
   )
-  
 )

### 2021-05-27 ms/manuscript.Rmd
Move ms rendering functions to separate file

@@
 
-```{r setup, include=FALSE}
+```{r setup, include = FALSE}
+# Load libraries used for rendering Rmds
 library(officedown)
@@
 library(kableExtra)
+
+# Set knitr options
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
+
+# Load functions only used when rendering Rmds
+source(here::here("R/ms_functions.R"), local = TRUE)
 
@@
 
-```{r macros}
-# Define formattting for some common custom words that may vary in style between journals
-ie <- "*ie*"
-eg <- "*eg*"
-# R packages
-pack <- function(x) {glue::glue("'{x}'")}
-# R functions
-func <- function(x) {glue::glue("'{x}'")}
-# Pagebreak for MS Word (doc) only
-pagebreak_doc <- function() {ifelse(params$doc_type == "doc", run_pagebreak(), return(""))}
-# Pagebreak for pdf only
-pagebreak_pdf <- function() {ifelse(params$doc_type == "pdf", return("\\clearpage"), return(""))}
-# Pagebreak for both
-pagebreak <- function() {
-  if (params$doc_type == "doc") {
-    run_pagebreak()
-  } else if (params$doc_type == "pdf") {
-    return("\\clearpage")
-  } else {
-    stop("doc_type parameter must be 'doc' or 'pdf'")
-  }
-}
-```
-
 ```{r load-data}
+# Load objects from targets workflow
 tar_load(c(

### 2021-05-27 R/ms_functions.R

@@ -1,0 +1,107 @@
+# Manuscript rendering functions and variables ----
+
+# Define formatting for some common custom words that may vary in style between journals
+ie <- "*ie*"
+eg <- "*eg*"
+
+#' Generate a path to save a results file
+#' 
+#' Only works for figures or tables cited in text, and outputs
+#' files to the "results" folder
+#'
+#' @param result_num Number of result, e.g. "Fig. 2"
+#' @param extension Extension to use for file
+#'
+#' @return String.
+#' 
+result_file <- function (result_num, extension) {
+  
+  fs::path(
+    here::here("results"),
+    result_num %>%
+      str_remove_all("\\.") %>% 
+      str_replace_all(" ", "_")
+  ) %>%
+    fs::path_ext_set(extension)
+}
+
+#' Format R packages names
+#'
+#' @param x Name of R package
+#'
+#' @return Name of R package formatted according to journal requirements
+#' 
+pack <- function(x) {glue::glue("'{x}'")}
+
+#' Format R function names
+#'
+#' @param x Name of R function
+#'
+#' @return Name of R function formatted according to journal requirements
+#' 
+func <- function(x) {glue::glue("'{x}'")}
+
+#' Pagebreak for MS Word (doc) only
+#' 
+#' Meant for use within Rmd when rendering.
+#' Breaks a page depending if the output format is MS Word
+#' 
+#' @param rmd_params Parameters set in YAML header or rmarkdown::render(). 
+#' Must include `doc_type` (either 'doc' for MS Word or 'pdf' for PDF output)
+#' 
+pagebreak_doc <- function(rmd_params = params) {ifelse(rmd_params$doc_type == "doc", run_pagebreak(), return(""))}
+
+#' Pagebreak for PDF only
+#' 
+#' Meant for use within Rmd when rendering. Requires parameter variable
+#' `doc_type` to be defined (either 'doc' for MS Word or 'pdf' for PDF output)
+#'
+#' Breaks a page depending if the output format is PDF
+pagebreak_pdf <- function(rmd_params = params) {ifelse(rmd_params$doc_type == "pdf", return("\\clearpage"), return(""))}
+
+#' Insert pagebreak
+#' 
+#' Meant for use within Rmd when rendering. Requires parameter variable
+#' `doc_type` to be defined (either 'doc' for MS Word or 'pdf' for PDF output)
+#'
+#' Breaks a page depending if the output format is PDF or MS Word
[…]

### 2021-05-27 ms/SI.Rmd
Render SI

--- 2020-08-26 ms/SI.Rmd
@@
 title: "Exploring dimensions of biodiversity in Japanese ferns"
-bibliography: "main_library.bib"
+bibliography: 
+  - references.yaml
+  - references_other.yaml
 mainfont: Roboto-Regular.ttf
@@
   bookdown::pdf_document2:
-    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
+    citation_package: default
     toc: no
@@
     fig_caption: yes
-    keep_tex: no
+    keep_tex: yes
     latex_engine: xelatex
     pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
-    includes:
-      in_header: newphyt.sty
 header-includes: 
@@
   chunk_output_type: console
+params:
+  doc_type: pdf
 # nocite: | 
@@
 
-```{r setup, include=FALSE}
+```{r setup-si, include = FALSE, cache = FALSE}
+# Load functions only used when rendering Rmds
+source(here::here("R/ms_functions.R"), local = TRUE)
+# Set knitr options
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
@@
 
-```{r load-data, cache = FALSE}
-# NOTE: This Rmd is meant to be knit as part of the drake plan (R/plan.R).
-# If using interactively, run `source("_drake.R")` first.
+```{r load-data-si, cache = FALSE}
+# This Rmd is meant to be knit as part of the target plan (R/_targets.R).
+# Run the lines below if rendering outside of targets plan
+# library(targets)
+# library(tarchetypes)
+# source(here::here("R/packages.R"))
+# source(here::here("R/functions.R"))
 
-loadd(list = c(
+tar_load(c(
   "japan_shp",
-  "comm_scaled_list_0.1",
-  "comm_scaled_list_0.2",
-  "comm_scaled_list_0.3",
-  "comm_scaled_list_0.4",
+  "comm_scaled_list",
   "occ_point_data_ferns",
@@
   "binary_sig_results"
-),
-cache = ja_fern_cache)
+))
+```
 
-# Load captions
-source(here::here("ms/captions.R"))
+```{r load-captions}
+# Source captions numbers from MS
+# 
+# So that caption numbers are in order of appearance in MS
+
+# Run inline code when purling
+options(knitr.purl.inline = TRUE)
+
+# Make a temporary file to write out just inline R code from the SI Rmd
+temp_file <- tempfile()
+
+# Generate R script including inline R code from the ms Rmd
+knitr::purl(here::here("ms/manuscript.Rmd"), output = temp_file)
+
+# Trim this R script down to only functions that define figure and table captions
+read_lines(temp_file) %>%
+  magrittr::extract(
+    str_detect(., "^figure\\(|^table\\(|^s_figure\\(|^s_table\\(|^figure_num\\(|^table_num\\(|^s_figure_num\\(|^s_table_num\\(")
+  ) %>%
+  unique %>%
+  write_lines(temp_file)
+
+# Source the code. This will load the caption functions in the right order.
+source(temp_file, local = TRUE)
+
+# Delete the temporary file
+fs::file_delete(temp_file)
 ```
@@
 
-**`r s_table("phy-sig")`**. `r s_table_cap("phy-sig")` K, Blomberg's *K*; lambda, Pagel's lambda.
+**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan. K, Blomberg's *K*; lambda, Pagel's lambda.
 
@@
 **`r s_table("phy-sig-binary")`**.
-`r s_table_cap("phy-sig-binary")` num_present, number of times a trait was observed present; num_absent, number of times a trait was observed absent; D, Fritz and Purvis' *D*; prob_random, probability of obtaining observed trait distribution given random distribution of traits; prob_brownian, probability of obtaining observed trait distribution given random evolution by Brownian Motion.
+Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.
+num_present, number of times a trait was observed present; num_absent, number of times a trait was observed absent; D, Fritz and Purvis' *D*; prob_random, probability of obtaining observed trait distribution given random distribution of traits; prob_brownian, probability of obtaining observed trait distribution given random evolution by Brownian Motion.
 Presence or absence of indusium was originally a binary trait; other traits (margin, shape, texture) are qualitative traits scored in binary format.
@@
 **`r s_figure("sampling-curve")`**.
-`r s_figure_cap("sampling-curve")` Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line).
+Species collection curve for the ferns of Japan fit with the `r pack("iNEXT")` package.
+Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line).
 Point indicates observed taxonomic richness (`r observed_data$qD` taxa, 95% confidence interval `r rich_interval` taxa).
@@
   list(
-    "0.1" = comm_scaled_list_0.1$poly_shp,
-    "0.2" = comm_scaled_list_0.2$poly_shp,
-    "0.3" = comm_scaled_list_0.3$poly_shp,
-    "0.4" = comm_scaled_list_0.4$poly_shp) %>%
+    "0.1" = comm_scaled_list %>% filter(resol == 0.1) %>% pull(poly_shp) %>% pluck(1),
+    "0.2" = comm_scaled_list %>% filter(resol == 0.2) %>% pull(poly_shp) %>% pluck(1),
+    "0.3" = comm_scaled_list %>% filter(resol == 0.3) %>% pull(poly_shp) %>% pluck(1),
+    "0.4" = comm_scaled_list %>% filter(resol == 0.4) %>% pull(poly_shp) %>% pluck(1)
+    ) %>%
   map_df(~as_tibble(.) %>% select(grids, abundance, richness), .id = "res") %>%
@@
 **`r s_figure("grain-size")`**.
-`r s_figure_cap("grain-size")` **A**, Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
+Effect of grain size on sampling redundancy.
+**A**, Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
 **B**, Percent of grid cells with redundancy = 0 by grain size.
@@
 
-**`r s_figure("abundance")`**. `r s_figure_cap("abundance")`
+**`r s_figure("abundance")`**. Observed number of specimens (**A**) and sampling redundancy (**B**) per 0.2° grid cell in the ferns of Japan.
 
@@
 **`r s_figure("k-plot")`**.
-`r s_figure_cap("k-plot")` **A**, taxonomic bioregions.
+Selection of *k* for bioregions.
+**A**, taxonomic bioregions.
 **B**, phylogenetic bioregions.
-Color indicates optimal *k* selected selected by \function{optimal{\_}phyloregion} function in the R package \package{phyloregion} (black, non-optimal; red, optimal; see Methods).
+Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal; see Methods).
 
@@
 **`r s_figure("raw-biplot")`**.
-`r s_figure_cap("raw-biplot")` **A**, Phylogenetic diversity (PD).
+Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.
+**A**, Phylogenetic diversity (PD).
 **B**, Relative phylogenetic diversity (RPD).
@@
 **`r s_figure("endemism-restricted")`**.
-`r s_figure_cap("endemism-restricted")` 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.
+'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
 For details, see Methods.

### 2021-05-27 _targets.R

@@ -650,5 +650,14 @@ tar_plan(
     knit_root_dir = here::here(),
     output_file = here::here("results/manuscript.pdf"),
     params = list(doc_type = "pdf")
+  ),
+  # - SI pdf
+  tar_render(
+    si_pdf,
+    path = "ms/SI.Rmd",
+    output_format = "bookdown::pdf_document2",
+    knit_root_dir = here::here(),
+    output_file = here::here("results/supp_info.pdf"),
+    params = list(doc_type = "pdf")
   )
 )

### 2021-05-28 ms/manuscript.Rmd
Summarize model stats

@@
 library(kableExtra)
+library(ftExtra)
 
@@
   # FIXME: remove until phylo analysis is working "plastome_tree",
-  "green_list"
+  "green_list",
+  "env_model_params",
+  "repro_model_params",
+  "aic_env_repro"
 ))
@@
 
-We first tested for spatial autocorrelation in biodiversity metrics and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `moran.mc` function in the `spdep` package.
+We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the `r pack("spdep")` package.
 In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05).
@@
 
-We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `modified.ttest` in the `SpatialPack` R package.
+We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the `r pack("SpatialPack")` R package.
 This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature.
@@
 
-We constructed a linear model for each response variable (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) dependent on the environmental variables.
-We did not include any interaction terms as we had no a-priori reason to do so.
+We constructed a linear model for each biodiversity metric dependent on the environmental variables.
 Initial inspection of the relationship between richness and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature for the richness model only.
-We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell as a random effect with the R package spaMM [@Rousset2014].
+We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014].
 We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence.
@@
 We postulated that % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis.
-These models included response variables with a genetic component (SES of PD, SES of RPD, PE p-score) as dependent on either % apomictic taxa or temperature, precipitation, and precipitation seasonality, with random spatial effects as described above.
-We compared the fit between models including % apomictic taxa or temperature using conditional Akaike Information Criterion (cAIC).
+These models included biodiversity metrics with a genetic component (SES of PD, SES of RPD, PE p-score) as dependent on either % apomictic taxa or temperature, precipitation, and precipitation seasonality, with random spatial effects as described above.
+We compared the fit between models including % apomictic taxa or temperature using conditional Akaike Information Criterion (cAIC) [@Courtiol2017; @Vaida2005].
 
@@
 # Verify that min lambda >0.95
-# assert_that(phy_sig_results %>% pull(lambda) %>% min > 0.95)
+assert_that(phy_sig_results %>% pull(lambda) %>% min > 0.95)
 # Verify that max K < 0.0.6
-# assert_that(phy_sig_results %>% pull(kval) %>% max < 0.06)
+assert_that(phy_sig_results %>% pull(kval) %>% max < 0.06)
 
@@
 
-<!-- All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.06, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
+All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.06, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
 `r n_negative_d` out of `r nrow(binary_sig_results)` of continuous traits (coded as binary) showed as much or more phylogenetic signal as expected under BM (*D* =< 0); others had *D* > 0 (`r s_table("phy-sig-binary")`).
--->
 
@@
 
-### Correlations between biodiversity and reproductive mode
-
-Spatial autocorrelation as measured with Moran's I was detected in linear models, but was not in any of the spatial mixed linear models.
-SES of RPD in particular shows a strong negative correlation with % apomictic taxa (_ estimated effect _, SD).
+### Relationships between biodiversity and climate or reproductive mode
+
+All biodiversity metrics except for SES of RFD showed a significant relationship with temperature (`r table("env-model-params")`).
 
@@
 `r pagebreak()`
+
+```{r table-prep}
+# Define nicely formatted terms for printing variable names
+resp_var_lookup <- tribble(
+  ~resp_var, ~resp_var_pretty,
+  "fd_obs_z", "SES of FD",
+  "pd_obs_z", "SES of PD",
+  "pe_obs_p_upper", "PE *p*-value",
+  "rfd_obs_z", "SES of RFD",
+  "richness", "Richness",
+  "rpd_obs_z", "SES of RPD",
+)
+
+indep_var_lookup <- tribble(
+  ~term, ~term_pretty,
+  "temp", "Temperature",
+  "precip", "Precipitation",
+  "precip_season", "Precipitation seasonality",
+  "I(temp^2)", "Temperature^2^",
+  "percent_apo", "% apomictic taxa"
+)
+```
 
@@
 **`r table("traits-used")`**. Fern traits used in this study.
+
+`r pagebreak()`
+
+```{r env-model-params-prep}
+# Format model parameter summary for nice printing as a table
+env_model_params_pretty <-
+env_model_params %>%
+  # Remove intercept
+  filter(term != "(Intercept)") %>%
+  # Double check that log-likelihood of null model is less than full
+  verify(all(loglik_null < loglik_full)) %>%
+  select(resp_var, term, estimate, cond_se, t_value, p_value) %>%
+  # Format response variable and fixed effect names to look nice
+  left_join(resp_var_lookup, by = "resp_var") %>%
+  select(-resp_var) %>%
+  rename(resp_var = resp_var_pretty) %>%
+  left_join(indep_var_lookup, by = "term") %>%
+  select(-term) %>%
+  rename(term = term_pretty) %>%
+  # Round to 2 signif digits
+  mutate(across(c(estimate, cond_se, t_value), ~signif(., 2))) %>%
+  # Format p-value
+  mutate(
+    signif = if_else(p_value < 0.05, TRUE, FALSE),
+    p_value = case_when(
+      p_value < 0.0001 ~ "< 0.0001",
+      TRUE ~ signif(p_value, 2) %>% as.character
+    )
+  ) %>%
+  arrange(resp_var, term) %>%
+  select(resp_var, term, estimate, cond_se, t_value, p_value, signif)
+```
+
+```{r env-model-params-doc, results = "asis", eval = params$doc_type == "doc"}
+env_model_params_pretty %>%
+  mutate(
+    across(where(is.numeric), as.character)) %>%
+  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
+  as_grouped_data(groups = "resp_var") %>%
+  as_flextable(hide_grouplabel = TRUE, col_keys = c("Variable", "Estimate", "Conditional SE", "*t*-value", "*p*-value")) %>%
+  bold(i = ~ signif == TRUE, bold = TRUE) %>%
+  colformat_md(part = "header")
+```
+
+```{r env-model-params-pdf, results = "asis", eval = params$doc_type == "pdf"}
+# Define row groups
+kbl_groups <-
+env_model_params_pretty %>%
+  transmute(resp_var, row_num = 1:nrow(.)) %>%
+  group_by(resp_var) %>%
+  summarize(start = min(row_num), end = max(row_num), .groups = "drop") %>%
+  split(.$resp_var)
+
+signif_rows <- env_model_params_pretty %>% pull(signif) %>% which()
+
+env_model_params_pretty %>%
+  mutate(
+    across(where(is.numeric), as.character)) %>%
+  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
+  select(-resp_var, -signif) %>%
+  kbl(booktabs = TRUE) %>%
+  kable_styling() %>%
+  pack_rows(kbl_groups[[1]]$resp_var, kbl_groups[[1]]$start, kbl_groups[[1]]$end) %>%
+  pack_rows(kbl_groups[[2]]$resp_var, kbl_groups[[2]]$start, kbl_groups[[2]]$end) %>%
+  pack_rows(kbl_groups[[3]]$resp_var, kbl_groups[[3]]$start, kbl_groups[[3]]$end) %>%
+  pack_rows(kbl_groups[[4]]$resp_var, kbl_groups[[4]]$start, kbl_groups[[4]]$end) %>%
+  pack_rows(kbl_groups[[5]]$resp_var, kbl_groups[[5]]$start, kbl_groups[[5]]$end) %>%
+  pack_rows(kbl_groups[[6]]$resp_var, kbl_groups[[6]]$start, kbl_groups[[6]]$end) %>%
+  row_spec(row = signif_rows, bold = TRUE)
+```
+
+**`r table("env-model-params")`**.
+Parameter estimates of the effect of environmental variables on biodiversity metrics in the ferns of Japan.
+Significant effects (*p* < 0.05) in bold; significance assessed with likelihood ratio tests (see Methods).
+All models include random effect of geographical coordinates.
 

### 2021-05-28 R/functions.R

@@ -3033,6 +3033,135 @@ get_aic <- function(model) {
     janitor::clean_names()
 }
 
+#' Summarize statistics of spatial models
+#' 
+#' @param spatial_models Tibble, each with a spatial model.
+#' Output of run_spamm()
+#'
+#' @return Tibble including columns: `resp_var` (response variable),
+#' `formula` (text string of formula used for model), `data_type` (dataset used for
+#' model; 'env' for environmental, 'repro' for reproductive), `loglik` (log-likelihood),
+#' `resid_var_phi` (residual variance), `nu` (correlation parameter of Matern matrix),
+#'  `rho` (correlation parameter of Matern matrix)
+#' 
+get_model_stats <- function(spatial_models) {
+  spatial_models %>%
+    # Extract log likelihood and residual variance (phi)
+    mutate(
+      loglik = map_dbl(model, logLik),
+      resid_var_phi = map_dbl(model, ~residVar(., "phi") %>% unique)
+    ) %>%
+    # Extract nu and rho
+    mutate(map_df(.$model, get_corr_pars)) %>%
+    # Ditch the model
+    select(-model)
+}
+
+#' Summarize parameters of spatial models
+#' 
+#' Filters to only environmental models (those based on temperature and precipitation)
+#'
+#' @param spatial_models Tibble, each with a spatial model.
+#' Output of run_spamm()
+#' @param lrt_comp_table Tibble, results of likelihood ratio test (LRT)
+#' Output of run_spamm_lrt()
+#'
+#' @return Tibble including columns `resp_var` (response variable), `term` (fixed effect parameter name),
+#'  `estimate` (slope of parameter) `cond_se` (conditional standard error), `t_value`, 
+#'  `chi2_LR` (chi-squared value of LRT), `df`, `p_value`, `loglik_null` (log-likelihood of null model), 
+#'  `loglik_full` (log-likelihood of full model),
+#' `null_formula`, `full_formula`
+#' 
+get_env_model_params <- function(spatial_models, lrt_comp_table) {
+  spatial_models %>%
+    filter(data_type == "env") %>%
+    # Extract fixed effects
+    mutate(fixed_effects = map(model, get_beta_table)) %>%
+    # Ditch the model
+    select(-model, -formula) %>%
+    unnest(fixed_effects) %>%
+    # Add LRT results
+    left_join(
+      lrt_comp_table,
+      by = c("resp_var", "data_type", term = "comparison")
[…]

### 2021-05-28 _targets.R

@@ -530,6 +530,16 @@ tar_plan(
     pattern = map(data_for_lrt)
   ),
   
+  # Summarize spatial models:
+  # - model statistics (both env and repro models)
+  model_stats = get_model_stats(spatial_models),
+  # - environmental model parameters (fixed effects)
+  env_model_params = get_env_model_params(spatial_models, lrt_comp_table),
+  # - reproductive model parameters (fixed effects)
+  repro_model_params = get_repro_model_params(spatial_models, lrt_comp_table),
+  # - comparison between environmental and reproductive models by cAIC
+  aic_env_repro = compare_aic_env_repro(spatial_models),
+  
   # Conservation analysis ----
   
   ## Read in protected areas (7 separate shape files corresponding to different kinds of areas)

### 2021-05-28 ms/manuscript.Rmd
Remove colformat_md()
for some very weird reason, it was causing pandoc to not be able to find references.yaml. See error produced in last commit:

Could not find bibliography file: references.yaml
Error running filter /usr/local/bin/pandoc-citeproc:
Filter returned error status 1
Quitting from lines 641-649 (manuscript.Rmd)

@@
   as_flextable(hide_grouplabel = TRUE, col_keys = c("Variable", "Estimate", "Conditional SE", "*t*-value", "*p*-value")) %>%
-  bold(i = ~ signif == TRUE, bold = TRUE) %>%
-  colformat_md(part = "header")
+  bold(i = ~ signif == TRUE, bold = TRUE)
 ```

### 2021-05-28 ms/manuscript.Rmd
Save or write plots depending on Rmd output format

@@
 
-```{r japan-map-save}
+```{r japan-map-make}
 # Plot map of Japan with labels of relevant places
@@
   map_theme()
-
+```
+
+```{r japan-map-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
 ggsave(result_file(figure("japan-map"), "pdf"), japan_map_plot, width = 5, height = 5, units = "in")
@@
 
-```{r raw-diversity-save}
+```{r raw-diversity-make}
 # Richness
@@
   plot_layout(nrow = 1)
-
+```
+
+```{r raw-diversity-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
 ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7, height = 4, units = "in", raw_diversity_plot)
@@
 
-```{r rand-diversity-save}
+```{r rand-diversity-make}
 # PD
@@
   theme(legend.position = "bottom")
-
+```
+
+```{r rand-diversity-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
 ggsave(result_file(figure("rand-diversity"), "pdf"), rand_diversity_plot, width = 7, height = 8, units = "in")
@@
 
-```{r endemism-save, fig.width = 5, fig.height = 5}
+```{r endemism-make}
 endemism_plot <- ggplot() +
@@
   theme(legend.title = element_blank())
-
+```
+
+```{r endemism-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
 ggsave(result_file(figure("endemism"), "pdf"), endemism_plot, width = 5, height = 5, units = "in")
@@
 
-```{r bioregions-save}
+```{r bioregions-make}
 # Taxonomic bioregions
@@
   plot_layout(ncol = 2)
-
+```
+
+```{r bioregions-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
 ggsave(result_file(figure("bioregions"), "pdf"), bioregions_plot, width = 7, height = 5, units = "in")
@@
 
-```{r div-by-tax-region-save}
+```{r div-by-tax-region-make}
 data <-
@@
   theme(legend.position = "none")
-
+```
+
+```{r div-by-tax-region-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
 ggsave(result_file(figure("div-by-tax-region"), "pdf"), div_by_tax_region_plot, width = 7, height = 7, units = "in")
@@
 
-```{r conserv-status-save, fig.width = 4, fig.height = 4}
+```{r conserv-status-make}
 # Plot protected areas
@@
   )
-
+```
+
+```{r conserv-status-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
 ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 4, height = 4, units = "in")

### 2021-05-28 ms/manuscript.Rmd
Add tables for model parameters

@@
 ```{r setup, include = FALSE}
+# Run these lines if rendering outside of targets plan
+# library(targets)
+# library(tarchetypes)
+# source(here::here("R/packages.R"))
+# source(here::here("R/functions.R"))
+
 # Load libraries used for rendering Rmds
@@
 source(here::here("R/ms_functions.R"), local = TRUE)
-
-# Run these lines if rendering outside of targets plan
-# library(targets)
-# library(tarchetypes)
-# source(here::here("R/packages.R"))
-# source(here::here("R/functions.R"))
 ```
@@
 
-All biodiversity metrics except for SES of RFD showed a significant relationship with temperature (`r table("env-model-params")`).
+```{r model-stats}
+# Verify the number of cases where models had significant fits for precipitation, precipitation seasonality, and temperature
+n_env_precip_sig <- 5
+n_env_precip_season_sig <- 2
+n_env_temp_sig <- 5
+
+env_model_params %>%
+  filter(str_detect(term, "Intercept", negate = TRUE)) %>%
+  filter(p_value < 0.05) %>%
+  count(term) %>%
+  verify(.$n[.$term == "precip"] == n_env_precip_sig) %>%
+  verify(.$n[.$term == "precip_season"] == n_env_precip_season_sig) %>%
+  verify(.$n[.$term == "temp"] == n_env_temp_sig)
+
+assert_that(all.equal(n_env_precip_sig, n_env_temp_sig))
+
+repro_model_params %>%
+  filter(str_detect(term, "Intercept", negate = TRUE)) %>%
+  filter(p_value < 0.05) %>%
+  count(term) %>%
+  verify(.$n[.$term == "percent_apo"] == 3)
+
+# Add a column with estimate ± cond_se, then
+# split up table for easy referencing in inline chunks
+repro_cond_se <-
+repro_model_params %>%
+  mutate(
+    across(c(estimate, cond_se), ~signif(., 2)),
+    est_se = glue("{estimate} ± {cond_se}")
+) %>%
+  split(.$resp_var) %>%
+  map(~split(., .$term))
+
+env_cond_se <-
+env_model_params %>%
+  mutate(
+    across(c(estimate, cond_se), ~signif(., 2)),
+    est_se = glue("{estimate} ± {cond_se}")
+) %>%
+  split(.$resp_var) %>%
+  map(~split(., .$term))
+
+# Example:
+# parameter estimate +/- SE for SES of RPD in reproductive model:
+# repro_cond_se$rpd_obs_z$percent_apo$est_se
+```
+
+All spatial models were significant relative to null models without any fixed effects.
+Most biodiversity metrics showed a significant relationship with temperature and/or precipitation (each significant in `r english2(n_env_temp_sig)` out of six models), but precipitation seasonality was only significant in `r english2(n_env_precip_season_sig)` models (`r table("env-model-params")`).
+Models fit using % apomictic taxa instead of temperature had lower cAIC, indicating a better fit.
+All biodiversity metrics with a genetic component (PE *p*-value, SES of PD, SES of RPD) each had a significant, negative relationship to % apomictic taxa (`r table("repro-model-params")`).
 
@@
 
-```{r env-model-params-prep}
-# Format model parameter summary for nice printing as a table
-env_model_params_pretty <-
-env_model_params %>%
+```{r model-params-prep}
+# Format model parameter summaries for nice printing as a table
+model_params_pretty <-
+  # Combine environmental and reproductive model results to avoid repitition
+  bind_rows(env_model_params, repro_model_params) %>%
   # Remove intercept
@@
   verify(all(loglik_null < loglik_full)) %>%
-  select(resp_var, term, estimate, cond_se, t_value, p_value) %>%
   # Format response variable and fixed effect names to look nice
@@
   ) %>%
-  arrange(resp_var, term) %>%
-  select(resp_var, term, estimate, cond_se, t_value, p_value, signif)
+  arrange(data_type, resp_var, term) %>%
+  select(resp_var, term, estimate, cond_se, t_value, p_value, signif, data_type)
+
+env_model_params_pretty <- model_params_pretty %>%
+  filter(data_type == "env") %>%
+  select(-data_type)
+
+repro_model_params_pretty <- model_params_pretty %>%
+  filter(data_type == "repro") %>%
+  select(-data_type)
 ```
@@
 ```{r env-model-params-doc, results = "asis", eval = params$doc_type == "doc"}
+# Print out environmental parameters formatted for word document
 env_model_params_pretty %>%
@@
   as_flextable(hide_grouplabel = TRUE, col_keys = c("Variable", "Estimate", "Conditional SE", "*t*-value", "*p*-value")) %>%
-  bold(i = ~ signif == TRUE, bold = TRUE)
+  bold(i = ~ signif == TRUE, bold = TRUE) %>%
+  width(1, width = 1.5) %>%
+  width(2:5, width = 1)
 ```
@@
 Parameter estimates of the effect of environmental variables on biodiversity metrics in the ferns of Japan.
+Significant effects (*p* < 0.05) in bold; significance assessed with likelihood ratio tests (see Methods).
+All models include random effect of geographical coordinates.
+
+`r pagebreak()`
+
+```{r repro-model-params-doc, results = "asis", eval = params$doc_type == "doc"}
+# Print out environmental parameters formatted for word document
+repro_model_params_pretty %>%
+  mutate(
+    across(where(is.numeric), as.character)) %>%
+  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
+  as_grouped_data(groups = "resp_var") %>%
+  as_flextable(hide_grouplabel = TRUE, col_keys = c("Variable", "Estimate", "Conditional SE", "*t*-value", "*p*-value")) %>%
+  bold(i = ~ signif == TRUE, bold = TRUE) %>%
+  width(1, width = 1.5) %>%
+  width(2:5, width = 1)
+```
+
+```{r repro-model-params-pdf, results = "asis", eval = params$doc_type == "pdf"}
+# Define row groups
+kbl_groups <-
+  repro_model_params_pretty %>%
+  transmute(resp_var, row_num = 1:nrow(.)) %>%
+  group_by(resp_var) %>%
+  summarize(start = min(row_num), end = max(row_num), .groups = "drop") %>%
+  split(.$resp_var)
+
+signif_rows <- repro_model_params_pretty %>% pull(signif) %>% which()
+
+repro_model_params_pretty %>%
+  mutate(
+    across(where(is.numeric), as.character)) %>%
+  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
+  select(-resp_var, -signif) %>%
+  kbl(booktabs = TRUE) %>%
+  kable_styling() %>%
+  pack_rows(kbl_groups[[1]]$resp_var, kbl_groups[[1]]$start, kbl_groups[[1]]$end) %>%
+  pack_rows(kbl_groups[[2]]$resp_var, kbl_groups[[2]]$start, kbl_groups[[2]]$end) %>%
+  pack_rows(kbl_groups[[3]]$resp_var, kbl_groups[[3]]$start, kbl_groups[[3]]$end) %>%
+  row_spec(row = signif_rows, bold = TRUE)
+```
+
+**`r table("repro-model-params")`**.
+Parameter estimates of the effect of reproductive mode and environmental variables (precipitation) on biodiversity metrics in the ferns of Japan.
 Significant effects (*p* < 0.05) in bold; significance assessed with likelihood ratio tests (see Methods).

### 2021-06-03 ms/data_exploration.Rmd
Add Rmd with data exploration for modeling

---
title: "Data exploration"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  knit_type: manual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we perform initial data exploration to determine the appropriate models to fit to the data.

```{r load-packages, include = params$knit_type == "manual"}
# Load packages
library(tidyverse)
library(targets)
library(outliers)
library(ggforce)
library(assertr)

# Load custom functions
source(here::here("R/functions.R"))
```

There are two data sets:
* `biodiv_ferns_cent_env` with biodiversity metrics measured on the full set of species (species missing functional data were excluded when measuring functional diversity metrics).
* `biodiv_ferns_cent_repro` with biodiveristy metrics related to genetic diversity measured only on species with reproductive mode data available.

All biodiversity metrics were calculated during the `targets` workflow (`_targets.R`).

The `cent` part of the name indicates that they include geographic locations as centroids of each grid cell.

Below, "Full dataset" refers to `biodiv_ferns_cent_env` and "Repro dataset" refers to `biodiv_ferns_cent_repro`.

```{r load-data}
# Load the data from the targets workflow
tar_load(biodiv_ferns_cent_env)
tar_load(biodiv_ferns_cent_repro)
```

### Structure of the data

```{r struc-full-data}
biodiv_ferns_cent_env
```

```{r struc-repro-data}
biodiv_ferns_cent_repro
```

The columns are as follows:

* `grids`: ID code for each grid cell
* `lat`: latitude of grid cell centroid
* `long`: longitude of grid cell centroid
* `richness`: number of taxa
* `pd_obs_z`: standard effect size (SES) of phylogenetic diversity (pd)
* `fd_obs_z`: SES of functional diversity (fd)
* `rpd_obs_z`: SES of relative pd
* `rfd_obs_z`: SES of relative fd
* `temp`: mean annual temperature (10 x °C)
* `precip`: total annual precipitation (mm)
* `precip_season`: seasonality in precipitation
* `percent_apo`: percent of apogamous taxa in that grid cell (repro data only)

### Check for missing data

```{r check-missing}
# These will return TRUE if all data are non-missing, 
# or an error if any data are missing (NA)
biodiv_ferns_cent_env %>%
  assert(not_na, everything(), success_fun = success_logical)

biodiv_ferns_cent_repro %>%
  assert(not_na, everything(), success_fun = success_logical)
```

No missing data.

### Check for outliers

We will first check for outliers in the data by visualizing box plots. Dots at the extremes may be outliers.

```{r box-plots-full}
biodiv_ferns_cent_env %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  facet_wrap(vars(variable), scales = "free") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  labs(title = "Full dataset")
```

```{r box-plots-repro}
biodiv_ferns_cent_repro %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  facet_wrap(vars(variable), scales = "free") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  labs(title = "Repro dataset")
```

There are a few points beyond the whiskers that may be outliers.

Next, check for outliers using Grubb's test. Significant p-values indicate that the most extreme value is an outlier.

```{r grubbs-env}
biodiv_ferns_cent_env %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  group_by(variable) %>%
  nest() %>%
  mutate(map_df(data, ~grubbs_tidy(.$value))) %>%
  select(-data) %>%
  ungroup() %>%
  select(variable:p_value) %>%
  mutate(signif = p_value < 0.05)
```

```{r grubbs-repro}
biodiv_ferns_cent_repro %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  group_by(variable) %>%
  nest() %>%
  mutate(map_df(data, ~grubbs_tidy(.$value))) %>%
  select(-data) %>%
  ungroup() %>%
  select(variable:p_value) %>%
  mutate(signif = p_value < 0.05)
```

So the one percent apomictic > 60% appears to be an outlier. Perhaps this has a very small number of species, most of which happen to be apogamous?

```{r check-apo-outlier}
biodiv_ferns_cent_repro %>%
  slice_max(1, order_by = percent_apo) %>%
  left_join(select(biodiv_ferns_cent_env, grids, richness)) %>%
  select(grids:long, percent_apo, richness)
```

Looks like that is the case---just three species in this grid cell! This data point should be excluded. Let's do that and check for outliers again.

```{r remove-apo-outlier-check-grubbs}
biodiv_ferns_cent_repro %>% 
  filter(percent_apo != max(percent_apo)) %>%
  pull(percent_apo) %>%
  grubbs_tidy()
```

OK, no more outliers in `percent_apo`.

### Normality

Check for normality with histograms and QQ-plots.

Inspect histograms.

```{r hist-env}
biodiv_ferns_cent_env %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(vars(variable), scales = "free") +
  labs(title = "Full dataset")
```

```{r hist-repro}
biodiv_ferns_cent_repro %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(vars(variable), scales = "free") +
  labs(title = "Repro dataset")
```

Inspect qq-plots.

```{r qqplot-env}
biodiv_ferns_cent_env %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(sample = value)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Full dataset")
```

The response variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, `rpd_obs_z` all look decent. `richness` is skewed, but it is count data and we aren't using a Gaussian model for that anyways (poisson or negative binomial), so that's OK.

```{r qqplot-repro}
biodiv_ferns_cent_repro %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(sample = value)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Repro dataset")
```

Again, the response variables `pd_obs_z`, and `rpd_obs_z` look OK.

### Check for zeros in count data

Zero-inflated data can be a problem for linear models.

```{r check-for-zeros}
biodiv_ferns_cent_env %>%
  mutate(rich_zero = richness == 0) %>%
  arrange(richness) %>%
  summarize(
    total_zero = sum(rich_zero),
    perc_zero = sum(rich_zero) / nrow(.))
```

No zeros in our count data (richness).

### Plot raw data

Here we plot response variables (row labels) vs. indep variables (column labels) for each dataset, and add a trend line.

```{r plot-raw-data-env, fig.height = 8}
ggplot(biodiv_ferns_cent_env, aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
  geom_smooth() +
  facet_matrix(
    cols = vars(temp, precip, precip_season),
    rows = vars(richness, pd_obs_z, fd_obs_z, rpd_obs_z, rfd_obs_z)) +
  labs(title = "Full dataset")
```

`richness` vs. `temp` definitely shows a hump-shaped trend line, and the other response variables vs. `temp` also show one if not quite as strongly. The other relationships appear to be linear or uncorrelated.

```{r plot-raw-data-repro}
# Remove the outlier first
biodiv_ferns_cent_repro %>% 
  filter(percent_apo != max(percent_apo)) %>%
ggplot(aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
  geom_smooth() +
  facet_matrix(
    cols = vars(percent_apo, temp, precip, precip_season),
    rows = vars(pd_obs_z, rpd_obs_z)) +
  labs(title = "Repro dataset")
```

As one would expect from the last plot, `temp` shows a (negative) hump shape. However, there aren't many points on the right.. `percent_apo` seems more linear, especially for `rpd_obs_z`, but it's subtle. Anyways, we will probably keep the quadratic term for temperature but use a linear model for `percent_apo`.

### Check for overdispersion in richness

Since richness is count data (only takes positive values), we will use either a poisson or negative binomial to model its distribution. The poisson distribution requires that the ratio of residual variance to degrees of freedom be approximately 1; significantly different ratios are "overdispersed" or "underdispersed". Check for this with a dispersion test.

```{r check-overdispersion}
# Create general linear model with poisson distribution
richness_poisson_mod <- glm(
  richness ~ temp + I(temp^2) + precip + precip_season, 
  data = biodiv_ferns_cent_env, family = poisson)

# Check ratio of residual variance to df (should be near 1 to meet poisson assumptions)
richness_poisson_mod$deviance / richness_poisson_mod$df.residual

# Run statistical test
AER::dispersiontest(richness_poisson_mod, trafo=1)
```

That's very overdispersed. So we won't use poisson, but negative binomial instead.

### 2021-06-03 _targets.R

@@ -730,5 +730,14 @@ tar_plan(
     knit_root_dir = here::here(),
     output_file = here::here("results/supp_info.pdf"),
     params = list(doc_type = "pdf")
+  ),
+  # - SI appendix on data exploration for models
+  tar_render(
+    si_data_exploration,
+    path = "ms/data_exploration.Rmd",
+    output_format = "rmarkdown::html_document",
+    knit_root_dir = here::here(),
+    output_file = here::here("results/data_exploration.html"),
+    params = list(knit_type = "targets")
   )
 )

### 2021-06-03 ms/data_exploration.Rmd
Don't do dispersion test
overdispersion is so large it doesn't need it. and AER adds a bunch of deps.

@@
 richness_poisson_mod$deviance / richness_poisson_mod$df.residual
-
-# Run statistical test
-AER::dispersiontest(richness_poisson_mod, trafo=1)
 ```

### 2021-06-08 ms/data_exploration.Rmd
Add area to models

@@
 * `precip_season`: seasonality in precipitation
+* `area`: rolling mean of area in 0.2 degree latitudinal bands
 * `percent_apo`: percent of apogamous taxa in that grid cell (repro data only)

### 2021-06-08 _targets.R

@@ -476,8 +476,9 @@ tar_plan(
   
   # Check for correlation between independent variables in repro data
   t_test_results = run_mod_ttest_ja(
-    sf_to_centroids(biodiv_ferns_repro_spatial), 
-    vars_select = c("temp", "temp_season", "precip", "precip_season", "percent_apo")
+    sf_to_centroids(biodiv_ferns_repro_spatial) %>%
+    add_roll_area(lat_area_ja), 
+    vars_select = c("temp", "temp_season", "precip", "precip_season", "percent_apo", "area")
   ),
   
   ## Analyze Moran's I ----

### 2021-06-30 ms/manuscript.Rmd
Remove model stats tables

--- 2021-05-28 ms/manuscript.Rmd
@@
 
-All biodiversity metrics except for SES of RFD showed a significant relationship with temperature (`r table("env-model-params")`).
+All spatial models were significant relative to null models without any fixed effects.
+The effect of temperature dominated in most models, with only SES of PD showing a stronger effect of precipitation seasonality.
+Models fit using % apomictic taxa instead of temperature had lower cAIC, indicating a better fit.
+Biodiversity metrics with a genetic component (SES of PD, SES of RPD) each had a large, negative relationship with % apomictic taxa.
 
@@
 **`r table("traits-used")`**. Fern traits used in this study.
-
-`r pagebreak()`
-
-```{r env-model-params-prep}
-# Format model parameter summary for nice printing as a table
-env_model_params_pretty <-
-env_model_params %>%
-  # Remove intercept
-  filter(term != "(Intercept)") %>%
-  # Double check that log-likelihood of null model is less than full
-  verify(all(loglik_null < loglik_full)) %>%
-  select(resp_var, term, estimate, cond_se, t_value, p_value) %>%
-  # Format response variable and fixed effect names to look nice
-  left_join(resp_var_lookup, by = "resp_var") %>%
-  select(-resp_var) %>%
-  rename(resp_var = resp_var_pretty) %>%
-  left_join(indep_var_lookup, by = "term") %>%
-  select(-term) %>%
-  rename(term = term_pretty) %>%
-  # Round to 2 signif digits
-  mutate(across(c(estimate, cond_se, t_value), ~signif(., 2))) %>%
-  # Format p-value
-  mutate(
-    signif = if_else(p_value < 0.05, TRUE, FALSE),
-    p_value = case_when(
-      p_value < 0.0001 ~ "< 0.0001",
-      TRUE ~ signif(p_value, 2) %>% as.character
-    )
-  ) %>%
-  arrange(resp_var, term) %>%
-  select(resp_var, term, estimate, cond_se, t_value, p_value, signif)
-```
-
-```{r env-model-params-doc, results = "asis", eval = params$doc_type == "doc"}
-env_model_params_pretty %>%
-  mutate(
-    across(where(is.numeric), as.character)) %>%
-  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
-  as_grouped_data(groups = "resp_var") %>%
-  as_flextable(hide_grouplabel = TRUE, col_keys = c("Variable", "Estimate", "Conditional SE", "*t*-value", "*p*-value")) %>%
-  bold(i = ~ signif == TRUE, bold = TRUE)
-```
-
-```{r env-model-params-pdf, results = "asis", eval = params$doc_type == "pdf"}
-# Define row groups
-kbl_groups <-
-env_model_params_pretty %>%
-  transmute(resp_var, row_num = 1:nrow(.)) %>%
-  group_by(resp_var) %>%
-  summarize(start = min(row_num), end = max(row_num), .groups = "drop") %>%
-  split(.$resp_var)
-
-signif_rows <- env_model_params_pretty %>% pull(signif) %>% which()
-
-env_model_params_pretty %>%
-  mutate(
-    across(where(is.numeric), as.character)) %>%
-  rename("*t*-value" = t_value, "*p*-value" = p_value, "Variable" = term, "Estimate" = estimate, "Conditional SE" = cond_se) %>%
-  select(-resp_var, -signif) %>%
-  kbl(booktabs = TRUE) %>%
-  kable_styling() %>%
-  pack_rows(kbl_groups[[1]]$resp_var, kbl_groups[[1]]$start, kbl_groups[[1]]$end) %>%
-  pack_rows(kbl_groups[[2]]$resp_var, kbl_groups[[2]]$start, kbl_groups[[2]]$end) %>%
-  pack_rows(kbl_groups[[3]]$resp_var, kbl_groups[[3]]$start, kbl_groups[[3]]$end) %>%
-  pack_rows(kbl_groups[[4]]$resp_var, kbl_groups[[4]]$start, kbl_groups[[4]]$end) %>%
-  pack_rows(kbl_groups[[5]]$resp_var, kbl_groups[[5]]$start, kbl_groups[[5]]$end) %>%
-  pack_rows(kbl_groups[[6]]$resp_var, kbl_groups[[6]]$start, kbl_groups[[6]]$end) %>%
-  row_spec(row = signif_rows, bold = TRUE)
-```
-
-**`r table("env-model-params")`**.
-Parameter estimates of the effect of environmental variables on biodiversity metrics in the ferns of Japan.
-Significant effects (*p* < 0.05) in bold; significance assessed with likelihood ratio tests (see Methods).
-All models include random effect of geographical coordinates.
 

### 2021-06-30 ms/manuscript.Rmd
Add model coefficients plot

@@
 library(ftExtra)
+library(patchwork)
 
@@
   "green_list",
-  "env_model_params",
-  "repro_model_params",
+  "model_params",
+  "lrt_results",
   "aic_env_repro"
@@
 
-We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD, PE p-score) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the `r pack("spdep")` package.
+We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the `r pack("spdep")` package.
 In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05).
@@
 We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the `r pack("SpatialPack")` R package.
-This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature.
-The dataset including % apomictic taxa is smaller than the full dataset (_ vs _ taxa, respectively), so for the main modeling analysis we only included the following environmental variables: temperature, precipitation, and precipitation seasonality.
-
-We constructed a linear model for each biodiversity metric dependent on the environmental variables.
-Initial inspection of the relationship between richness and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature for the richness model only.
+This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature, so for the main modeling analysis we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
+
+We constructed a linear mixed model for each biodiversity metric dependent on the environmental variables ("environmental models").
+Initial inspection of the relationship between each biodiversity metric and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature only.
+Richness was fit with the negative binomial response family; others were Gaussian.
 We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014].
@@
 
-We postulated that % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis.
-These models included biodiversity metrics with a genetic component (SES of PD, SES of RPD, PE p-score) as dependent on either % apomictic taxa or temperature, precipitation, and precipitation seasonality, with random spatial effects as described above.
-We compared the fit between models including % apomictic taxa or temperature using conditional Akaike Information Criterion (cAIC) [@Courtiol2017; @Vaida2005].
+We postulated that % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis ("reproductive models").
+These models included biodiversity metrics with a genetic component (SES of PD and SES of RPD) as dependent on % apomictic taxa, precipitation, and precipitation seasonality, with random spatial effects as described above.
+We compared the fit between environmental models and reproductive models for SES of PD and SES of RPD using conditional Akaike Information Criterion (cAIC) [@Courtiol2017; @Vaida2005].
 
@@
 All spatial models were significant relative to null models without any fixed effects.
-The effect of temperature dominated in most models, with only SES of PD showing a stronger effect of precipitation seasonality.
-Models fit using % apomictic taxa instead of temperature had lower cAIC, indicating a better fit.
-Biodiversity metrics with a genetic component (SES of PD, SES of RPD) each had a large, negative relationship with % apomictic taxa.
+The effect of temperature dominated in most models, with only SES of PD showing a stronger effect of precipitation seasonality (`r figure("coeff")`A--E).
+Biodiversity metrics with a genetic component (SES of PD and SES of RPD) each had a large, negative relationship with % apomictic taxa (`r figure("coeff")`F--G).
+Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD.
 
@@
 Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`).
-The vast majority of endemic grid-cells are mixed, with very few paleo- or neoendemic cells.
+The vast majority of endemic grid-cells are mixed.
+A cluster of paleoendemic cells was observed in the Ryukyu Islands.
 Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (Daisetsuzan mountains).
@@
 For details, see Methods.
+
+`r pagebreak_pdf()`
+
+```{r coeff-make}
+### Make model coefficients plot ###
+
+# Format labels for printing
+resp_var_lookup <-
+  tibble(
+    resp_var = c("richness", "fd_obs_z", "pd_obs_z", "rfd_obs_z", "rpd_obs_z"),
+    resp_var_print = c("Richness", "SES of FD", "SES of PD", "SES of RFD", "SES of RPD")
+  )
+
+indep_var_lookup <-
+  tibble(
+    term = c("area", "I(temp^2)", "null_model", "percent_apo", "precip", "precip_season", "temp"),
+    term_print = c("Area", "Temp^2", "Null model", "% apomictic", "Precip.", "Precip.\nseasonality", "Temp.")
+  ) %>%
+  # Include "plain" version of term for arranging alphabetically
+  mutate(term_plain = str_replace_all(term, fixed("I(temp^2)"), "temp^2"))
+
+model_type_lookup <-
+  tibble(
+    model_type = c("temp", "percent_apo"),
+    model_type_print = c("env. model", "repro. model")
+  )
+
+# Format likelihood ratio test results
+lrt_signif <-
+  lrt_results %>%
+  # add model type (temp = environmental model, or percent_apo = repro model)
+  tidyr::extract(full_formula, "model_type", regex = " ~ ([^ ]+) ") %>%
+  select(-null_formula) %>%
+  select(resp_var, model_type, term = comparison, p_value) %>%
+  # Add column for significance
+  mutate(signif = p_value < 0.05)
+
+# Format parameter values for plotting
+model_params_for_plot <-
+  model_params %>%
+  # add model type
+  tidyr::extract(formula, "model_type", regex = " ~ ([^ ]+) ") %>%
+  # drop intercept
+  filter(term != "(Intercept)") %>%
+  # add LRT results and names formatted for printing
+  left_join(lrt_signif, by = c("resp_var", "term", "model_type")) %>%
+  left_join(resp_var_lookup, by = "resp_var") %>%
+  left_join(indep_var_lookup, by = "term") %>%
+  left_join(model_type_lookup, by = "model_type") %>%
+  # Make coefficient with greatest abs value bold
+  group_by(resp_var, model_type) %>%
+  mutate(
+    term_print = case_when(
+      abs(estimate) == max(abs(estimate)) ~ glue::glue("**{term_print}**"),
+      TRUE ~ term_print
+    ),
+    term_print = factor(term_print)
+  ) %>%
+  ungroup()
+
+# Define helper function to make a subplot showing coefficients
+plot_coeff <- function (df) {
+
+  plot_data <- df %>%
+    mutate(term_print = fct_reorder(term_print, rank(term_plain)) %>% fct_rev)
+
+  ggplot(plot_data, aes(x = estimate, y = term_print, color = signif)) +
+    geom_vline(xintercept = 0, linetype = "dotted") +
+    geom_pointrange(aes(xmin = estimate - cond_se, xmax = estimate + cond_se), size = 0.25) +
+    scale_x_continuous(n.breaks = 4) +
+    scale_color_manual(
+      values = c("TRUE" = "grey20", "FALSE" = "grey70"),
+      labels = c("TRUE" = "*p* < 0.05", "FALSE" = "*p* > 0.05")) +
+    labs(
+      subtitle = glue("{unique(plot_data$resp_var_print)},\n{unique(plot_data$model_type_print)}"),
+      color = "Significance") +
+    theme_gray(base_size = 9) +
+    theme(
+      axis.title = element_blank(),
+      axis.text.y = element_markdown(),
+      legend.text = element_markdown(),
+      legend.position = "none")
+}
+
+# Loop across data and make a plot of each
+coeff_subplots <-
+model_params_for_plot %>%
+  group_by(resp_var, model_type) %>%
+  nest() %>%
+  ungroup %>%
+  mutate(plot = map(data, plot_coeff)) %>%
+  pull(plot)
+
+# We need a legend from one of these
+coeff_subplots[[7]] <- coeff_subplots[[7]] + theme(legend.position = "right")
+
+# Generate final plot
+coeff_plot <- wrap_plots(coeff_subplots) +
+  guide_area() +
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "A")
+```
+
+```{r coeff-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
+ggsave(result_file(figure("coeff"), "pdf"), coeff_plot, width = 7, height = 5, units = "in")
+```
+
+```{r coeff-print, fig.width = 7, fig.height = 5, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+coeff_plot
+```
+
+**`r figure("coeff")`**.
+Regression coefficients of general mixed models.
+A--E, environmental models (models including temperature); F and G, reproductive models (models including % apomictic taxa).
+Error bars show 95% confidence interval.
+Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-value indicates probability of no difference between the full and null model.
+Name of response variable and model type indicated above each subplot.
+Independent variable with greatest absolute coefficient value in bold for each model.
 

### 2021-07-01 ms/manuscript.Rmd
Add t-value and model fit plots to ms

@@
   "lrt_results",
-  "aic_env_repro"
+  "aic_env_repro",
+  "model_fits",
+  "spatial_models"
 ))
@@
 All spatial models were significant relative to null models without any fixed effects.
-The effect of temperature dominated in most models, with only SES of PD showing a stronger effect of precipitation seasonality (`r figure("coeff")`A--E).
-Biodiversity metrics with a genetic component (SES of PD and SES of RPD) each had a large, negative relationship with % apomictic taxa (`r figure("coeff")`F--G).
+The effect of temperature dominated in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G).
+Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A).
+Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E).
+SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F).
 Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD.
@@
   left_join(model_type_lookup, by = "model_type") %>%
-  # Make coefficient with greatest abs value bold
+  # Make coefficient with greatest abs t-value bold
   group_by(resp_var, model_type) %>%
@@
     term_print = case_when(
-      abs(estimate) == max(abs(estimate)) ~ glue::glue("**{term_print}**"),
+      abs(t_value) == max(abs(t_value)) ~ glue::glue("**{term_print}**"),
       TRUE ~ term_print
@@
 
-  ggplot(plot_data, aes(x = estimate, y = term_print, color = signif)) +
+  ggplot(plot_data, aes(x = t_value, y = term_print, fill = signif)) +
     geom_vline(xintercept = 0, linetype = "dotted") +
-    geom_pointrange(aes(xmin = estimate - cond_se, xmax = estimate + cond_se), size = 0.25) +
-    scale_x_continuous(n.breaks = 4) +
-    scale_color_manual(
+    geom_col() +
+    # scale_x_continuous(n.breaks = 4) +
+    scale_x_continuous() +
+    scale_fill_manual(
       values = c("TRUE" = "grey20", "FALSE" = "grey70"),
@@
       subtitle = glue("{unique(plot_data$resp_var_print)},\n{unique(plot_data$model_type_print)}"),
-      color = "Significance") +
+      fill = "Significance") +
     theme_gray(base_size = 9) +
@@
 **`r figure("coeff")`**.
-Regression coefficients of general mixed models.
+Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models.
 A--E, environmental models (models including temperature); F and G, reproductive models (models including % apomictic taxa).
-Error bars show 95% confidence interval.
 Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-value indicates probability of no difference between the full and null model.
 Name of response variable and model type indicated above each subplot.
-Independent variable with greatest absolute coefficient value in bold for each model.
+Independent variable with greatest absolute *t*-value in bold for each model.
+
+```{r modelfit-make}
+# Make plot showing model fits
+
+# Extract dataset used for model, including significance columns
+biodiv_ferns_used_for_model <- sf_to_centroids(biodiv_ferns_spatial) %>%
+  # filter to only sites used in model
+  inner_join(
+    spatial_models$model[[1]]$data %>% as_tibble() %>% select(grids),
+    by = "grids"
+  )
+
+# Specify order of subplots
+model_subplot_order <- tribble(
+  ~resp_var, ~indep_var, ~order,
+  "richness", "temp", 1,
+  "pd_obs_z", "temp", 2,
+  "rpd_obs_z", "temp", 3,
+  "fd_obs_z", "temp", 4,
+  "rfd_obs_z", "temp", 5,
+  "pd_obs_z", "percent_apo", 6,
+  "rpd_obs_z", "percent_apo", 7
+)
+
+# Define function for plotting model fit and actual data
+plot_fits <- function(resp_var, resp_var_print, indep_var, signif_var, fit, model_data, ...) {
+
+  # Setup randomization test significance colors
+  signif_cols <-
+    c(
+      "> 0.99" = "#e31a1c", 
+      "> 0.975" = "#fb9a99",
+      "< 0.01" = "#1f78b4", 
+      "< 0.025" = "#a6cee3",
+      "not significant" = "#ffffcc"
+    )
+
+  # Basic plot setup
+  p <- ggplot(model_data, aes(x = .data[[indep_var]], y = .data[[resp_var]])) +
+    labs(y = resp_var_print) +
+    theme(panel.grid.minor = element_blank())
+
+  # Add points, color by significance if not plotting richness
+  if (resp_var != "richness") {
+    p <- p + 
+      geom_point(aes(color = .data[[signif_var]]), size = 0.5) +
+      scale_color_manual(values = signif_cols)
+  } else {
+    p <- p + 
+      geom_point(size = 0.5, color = "grey20")
+  }
+
+  # Add line and ribbon of model fit
+  p <- p +
+    geom_line(data = fit, alpha = 0.35, color = "blue") +
+    geom_ribbon(
+      data = fit,
+      aes(ymin = low, ymax = up), 
+      alpha = 0.15
+    ) +
+    theme(legend.position = "none")
+
+  # Format x-axis ticks and labels
+  if (indep_var == "temp") {
+    p <- p +
+      scale_x_continuous(labels = function(x) x*0.1) +
+      labs(x = "Temperature (°C)")
+  } else if (indep_var == "percent_apo") {
+    p <- p +
+      scale_x_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
+      labs(x = "% apomictic taxa")
+  }
+
+  p
+
+}
+
+# Generate final plot
+modelfit_plot_list <- 
+  model_fits %>%
+  mutate(
+    signif_var = str_replace_all(resp_var, "_obs_z", "_signif"),
+    resp_var_print = case_when(
+      resp_var == "richness" ~ "Richness (no. spp.)",
+      TRUE ~ str_to_upper(resp_var) %>% word(1, 1, "_") %>% paste("SES of", .)
+    ),
+    model_data = list(biodiv_ferns_used_for_model)) %>%
+  left_join(model_subplot_order, by = c("resp_var", "indep_var")) %>%
+  arrange(order) %>%
+  pmap(plot_fits)
+
+# Need one subplot with a legend
+modelfit_plot_list[[2]] <- modelfit_plot_list[[2]] +
+  theme(legend.position = "right") +
+  labs(color = "Significance") +
+  # Increase size of points in legend
+  guides(colour = guide_legend(override.aes = list(size = 1.5)))
+
+modelfit_plot <-
+  wrap_plots(modelfit_plot_list) +
+  guide_area() +
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "A")
+```
+
+```{r modelfit-save, eval = params$doc_type == "doc"}
+# Save plot if rendering to doc
+ggsave(result_file(figure("modelfit"), "pdf"), modelfit_plot, width = 7, height = 5, units = "in")
+```
+
+```{r modelfit-print, fig.width = 7, fig.height = 5, eval = params$doc_type == "pdf"}
+# Print plot if rendering to PDF
+modelfit_plot
+```
+
+**`r figure("modelfit")`**.
+Relationship between biodiversity metrics and selected predictor variables.
+A--E, environmental models (models including temperature); F and G, reproductive models (models including % apomictic taxa).
+Ribbon shows 95% confidence interval of model.
+Line fit with the focal predictor variable while averaging over other predictors.
+For biodiversity metrics other than richness, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
 

### 2021-07-01 ms/manuscript.Rmd
Update MS

@@
 These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism.
-Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations.
+Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations [e.g., @White2019; @Daru2020a].
 Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.
@@
 
-We measured phylogenetic diversity (PD) using the Faith's PD (the total branch length connecting all terminal taxa in a community, including the root of the phylogeny) [@Faith1992] with the `r func("PD")` function in the R package `r pack("phyloregion")`.
-To detect areas with unusually long or short branch lengths, we measured relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been set to equal length [@Mishler2014].
+We measured phylogenetic diversity (PD) using the Faith's PD (the total branch length connecting all terminal taxa in a community, including the root of the phylogeny) [@Faith1992].
+To detect areas with unusually long or short branch lengths, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].
 
@@
 
-We measured phylogenetic endemism (PE) [@Rosauer2009], which is the sum of all branches at a site inversely weighted by their range, using the `r func("phylo_endemism")` function in the R package `r pack("phyloregion")`.
-
-Observed values of PD, FD, and PE are known to be closely related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others.
+We measured phylogenetic endemism (PE) as the sum of all branches at a site inversely weighted by their range [@Rosauer2009].
+
+Observed values of PD, RPD, FD, RFD, and PE are known to be closely related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others.
 Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test.
 We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values.
-Statistical significance was assessed using a two-tailed test with alpha = 0.5; observed values > 97.5% or < 0.25% were considered significant.
+Statistical significance was assessed using a one-tailed test for PE or a two-tailed test for other metrics with alpha = 0.5; observed values in the extreme 5% of the null distribution were considered significant.
 We also calculated the standard effect size (SES) of each biodiversity metric.
@@
 Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution.
-We used the same set of null communities as for SES of PE, and carried out categorization using custom R scripts.
+
+All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package (https://github.com/joelnitta/canaper), which calculates observed values of biodiversity using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].
 
@@
 
-All spatial models were significant relative to null models without any fixed effects.
-The effect of temperature dominated in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G).
+All models were significant relative to null models without any fixed effects.
+The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
 Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A).
@@
 
-### Distribution of biodiversity
+### Distribution and drivers of biodiversity
 
@@
 The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015] and modeled distributions of woody plants [@Fukaya2020].
-This pattern is probably due at least in part to the distribution of land area in Japan: maximum land area is located at mid-latitudes, and area is related to species richness by the well-known species-area relationship [@Arrhenius1921].
-There may also be a combined climatic and elevational effect: the southern part of Honshu is warmer and more humid than areas further north, while also harboring high elevational variation that may support a wide range of taxa, which is not present in the smaller islands further south.
-
-The overwhelming pattern characterizing phylogenetic and functional diversity is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C).
+Our modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
+Such a hump-shaped pattern is often observed in ferns on mountains, particularly in tropical or sub-tropical areas [@Kessler2011a].
+Hypotheses to explain such a pattern include pure geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a].
+Another reason for the greater richness in southern Honshu compared to smaller southern islands is that southern Honshu harbors high elevational variation that may support a wide range of taxa that is not present in the southern islands.
+
+The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C).
 This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
 The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis, and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity.
-Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE are located in the southern islands.
+Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands.
 This is clearly due to the small land area of these islands, which are the only areas with a subtropical climate in Japan.
@@
 
+Furthermore, we identified % apomictic taxa as a strong driver of SES of PD (`r figure("modelfit")`G); indeed, *Dryopteris* and *Deparia* are two such genera with high numbers of apomictic taxa.
+This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns.
+While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD.
+Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca.
+34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands).
+@Tanaka2014 assert that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid; as fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.
+
+Water limitation is thought to play an important role in promoting apomixis in ferns especially in deserts [@Grusz2021], since ferns depend on water for transfer of sperm to egg, and apomictic plants would be able to reproduce in the absence of water.
+A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomictic ferns [@Liu2012; @Tanaka2014], but in Japan, although the areas including high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
+The monsoon hypothesis is also not supported by our results, which show precipitation seasonality did not play a strong role in any model and was not correlated with % apomictic taxa.
+
 The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`).
-Interestingly, the bioregions identified here correspond roughly to major forest types categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014].
+Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014].
 This suggests that ferns are useful as possible bioindicators in Japan.
@@
 However, recent studies in Japan [@Ebihara2013; @Ebihara2019a] and elsewhere [@Nitta2017] have revealed different patterns in the distribution of gametophytes and sporophytes.
-The comprehensive molecular sampling of sporophytes makes this area ideal for conducting high-throughput DNA barcoding analyses to compare patterns of biodiversity between life cycles in ferns across large spatial scales.
-
-<!-- hybrid taxa, gametophytes, conclusion -->
+The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA barcoding analyses to compare patterns of biodiversity between life cycles in ferns across large spatial scales.
 
@@
     model_type_print = c("env. model", "repro. model")
+  )
+
+# Specify subplot order
+model_subplot_order <- tribble(
+  ~resp_var, ~indep_var, ~order,
+  "richness", "temp", 1,
+  "pd_obs_z", "temp", 2,
+  "rpd_obs_z", "temp", 3,
+  "fd_obs_z", "temp", 4,
+  "rfd_obs_z", "temp", 5,
+  "pd_obs_z", "percent_apo", 6,
+  "rpd_obs_z", "percent_apo", 7
 )
@@
       fill = "Significance") +
-    theme_gray(base_size = 9) +
+    theme_bw(base_size = 9) +
     theme(
@@
       legend.text = element_markdown(),
+      panel.grid.minor = element_blank(),
       legend.position = "none")
@@
   ungroup %>%
+  left_join(model_subplot_order, by = c("resp_var", model_type = "indep_var")) %>%
+  arrange(order) %>%
   mutate(plot = map(data, plot_coeff)) %>%
@@
     by = "grids"
-  )
-
-# Specify order of subplots
-model_subplot_order <- tribble(
-  ~resp_var, ~indep_var, ~order,
-  "richness", "temp", 1,
-  "pd_obs_z", "temp", 2,
-  "rpd_obs_z", "temp", 3,
-  "fd_obs_z", "temp", 4,
-  "rfd_obs_z", "temp", 5,
-  "pd_obs_z", "percent_apo", 6,
-  "rpd_obs_z", "percent_apo", 7
   )
@@
     labs(y = resp_var_print) +
+    theme_gray(base_size = 9) +
     theme(panel.grid.minor = element_blank())

### 2021-07-01 ms/manuscript.Rmd
Update MS
use bookdown instead of officedown. couldn't get template to work.

@@
 output: 
-  officedown::rdocx_document:
-    base_format: bookdown::word_document2
+  bookdown::word_document2:
     pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
     toc: no
-    reference_num: false
+    number_sections: no
     reference_docx: template.docx
@@
 # Load libraries used for rendering Rmds
-library(officedown)
-library(officer)
+# library(officedown)
+# library(officer)
 library(flextable)
@@
 
-<!-- Remove page numbers -->
-\pagenumbering{gobble}
-
 <!-- Set left justification -->
@@
 
-Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1^, and Atsushi Ebihara^3^
-
-^1^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo 113-0032, Japan
-
-^2^University and Jepson Herbaria, University of California, Berkeley, CA 94720‐2465, USA
-
-^3^Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba-shi, Ibaraki, 305-0005, Japan
+Short running title: Biodiversity of Japanese ferns
+
+Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1, 3, 4, 5, 6^, and Atsushi Ebihara^7^
+
+^1^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, Tokyo, Japan
+
+^2^University and Jepson Herbaria, University of California, Berkeley, CA, USA
+
+^3^Department of Computational Biology and Medical Sciences, Graduate School of Frontier Sciences, The University of Tokyo, Kashiwa, Japan
+
+^4^Atmosphere and Ocean Research Institute, The University of Tokyo, Kashiwa, Japan
+
+^5^Institute for Quantitative Biosciences, The University of Tokyo, Tokyo, Japan
+
+^6^Collaborative Research Institute for Innovative Microbiology, The University of Tokyo, Tokyo, Japan
+
+^7^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
 
@@
 
+## Acknowledgements
+
 `r pagebreak()`
@@
 
-**Aim** To characterize multiple facets of biodiversity and understand how these relate to bioregions and conservation status.
+**Aim** To characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.
 
@@
 
-**Main conclusions** The integrative approach used here reveals previously undetected patterns of biodiversity in the ferns of Japan.
+**Main conclusions** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan.
 The different distributions of each diversity type reflect the major bioregions.
@@
 
-All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package (https://github.com/joelnitta/canaper), which calculates observed values of biodiversity using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].
+All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package (https://github.com/joelnitta/canaper), which calculates raw biodiversity metrics using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].
 
@@
 The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA barcoding analyses to compare patterns of biodiversity between life cycles in ferns across large spatial scales.
-
-## Acknowledgements
 

### 2021-07-01 _targets.R

@@ -625,7 +625,7 @@ tar_plan(
   tar_render(
     ms_doc,
     path = "ms/manuscript.Rmd",
-    output_format = "officedown::rdocx_document",
+    output_format = "bookdown::word_document2",
     knit_root_dir = here::here(),
     output_file = here::here("results/manuscript.docx"),
     params = list(doc_type = "doc")

### 2021-07-01 ms/data_exploration.Rmd
Only use dataset for data exploration

--- 2021-06-03 ms/data_exploration.Rmd
@@
 
-There are two data sets:
-* `biodiv_ferns_cent_env` with biodiversity metrics measured on the full set of species (species missing functional data were excluded when measuring functional diversity metrics).
-* `biodiv_ferns_cent_repro` with biodiveristy metrics related to genetic diversity measured only on species with reproductive mode data available.
+The dataset includes biodiversity metrics and environmental data.
 
@@
 
-Below, "Full dataset" refers to `biodiv_ferns_cent_env` and "Repro dataset" refers to `biodiv_ferns_cent_repro`.
-
 ```{r load-data}
 # Load the data from the targets workflow
-tar_load(biodiv_ferns_cent_env)
-tar_load(biodiv_ferns_cent_repro)
+tar_load(biodiv_ferns_spatial)
+tar_load(resp_vars_env)
+tar_load(indep_vars)
+tar_load(lat_area_ja)
+
+# Make biodiversity metrics dataframe with centroid of each site for models
+  biodiv_ferns_cent <- sf_to_centroids(biodiv_ferns_spatial) %>%
+    # add area
+    add_roll_area(lat_area_ja) %>%
+    # keep only variables needed for model and only rows with zero missing data
+    filter_data_for_model(c("grids", "lat", "long", resp_vars_env, indep_vars))
 ```
@@
 ```{r struc-full-data}
-biodiv_ferns_cent_env
-```
-
-```{r struc-repro-data}
-biodiv_ferns_cent_repro
+biodiv_ferns_cent
 ```
@@
 * `rfd_obs_z`: SES of relative fd
+* `percent_apo`: percent of apogamous taxa in that grid cell
 * `temp`: mean annual temperature (10 x °C)
@@
 * `precip_season`: seasonality in precipitation
-* `percent_apo`: percent of apogamous taxa in that grid cell (repro data only)
+* `area`: rolling mean of area in 0.2 degree latitudinal bands
 
@@
 # or an error if any data are missing (NA)
-biodiv_ferns_cent_env %>%
-  assert(not_na, everything(), success_fun = success_logical)
-
-biodiv_ferns_cent_repro %>%
+biodiv_ferns_cent %>%
   assert(not_na, everything(), success_fun = success_logical)
@@
 ```{r box-plots-full}
-biodiv_ferns_cent_env %>%
+biodiv_ferns_cent %>%
   select(-lat, -long) %>%
@@
     axis.text.x = element_blank()
-  ) +
-  labs(title = "Full dataset")
-```
-
-```{r box-plots-repro}
-biodiv_ferns_cent_repro %>%
-  select(-lat, -long) %>%
-  pivot_longer(-grids, names_to = "variable") %>%
-  ggplot(aes(x = variable, y = value)) +
-  geom_boxplot() +
-  facet_wrap(vars(variable), scales = "free") +
-  theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank()
-  ) +
-  labs(title = "Repro dataset")
+  )
 ```
@@
 ```{r grubbs-env}
-biodiv_ferns_cent_env %>%
-  select(-lat, -long) %>%
-  pivot_longer(-grids, names_to = "variable") %>%
-  group_by(variable) %>%
-  nest() %>%
-  mutate(map_df(data, ~grubbs_tidy(.$value))) %>%
-  select(-data) %>%
-  ungroup() %>%
-  select(variable:p_value) %>%
-  mutate(signif = p_value < 0.05)
-```
-
-```{r grubbs-repro}
-biodiv_ferns_cent_repro %>%
+biodiv_ferns_cent %>%
   select(-lat, -long) %>%
@@
 ```{r check-apo-outlier}
-biodiv_ferns_cent_repro %>%
+biodiv_ferns_cent %>%
   slice_max(1, order_by = percent_apo) %>%
-  left_join(select(biodiv_ferns_cent_env, grids, richness)) %>%
+  left_join(select(biodiv_ferns_cent, grids, richness)) %>%
   select(grids:long, percent_apo, richness)
@@
 ```{r remove-apo-outlier-check-grubbs}
-biodiv_ferns_cent_repro %>% 
+biodiv_ferns_cent %>% 
   filter(percent_apo != max(percent_apo)) %>%
@@
 ```{r hist-env}
-biodiv_ferns_cent_env %>%
+biodiv_ferns_cent %>%
   select(-lat, -long) %>%
@@
   geom_histogram(bins = 30) +
-  facet_wrap(vars(variable), scales = "free") +
-  labs(title = "Full dataset")
-```
-
-```{r hist-repro}
-biodiv_ferns_cent_repro %>%
-  select(-lat, -long) %>%
-  pivot_longer(-grids, names_to = "variable") %>%
-  ggplot(aes(x = value)) +
-  geom_histogram(bins = 30) +
-  facet_wrap(vars(variable), scales = "free") +
-  labs(title = "Repro dataset")
+  facet_wrap(vars(variable), scales = "free")
 ```
@@
 ```{r qqplot-env}
-biodiv_ferns_cent_env %>%
+biodiv_ferns_cent %>%
   select(-lat, -long) %>%
@@
 
-```{r qqplot-repro}
-biodiv_ferns_cent_repro %>%
-  select(-lat, -long) %>%
-  pivot_longer(-grids, names_to = "variable") %>%
-  ggplot(aes(sample = value)) +
-  stat_qq() +
-  stat_qq_line() +
-  facet_wrap(~variable, scales = "free") +
-  labs(title = "Repro dataset")
-```
-
-Again, the response variables `pd_obs_z`, and `rpd_obs_z` look OK.
-
 ### Check for zeros in count data
@@
 ```{r check-for-zeros}
-biodiv_ferns_cent_env %>%
+biodiv_ferns_cent %>%
   mutate(rich_zero = richness == 0) %>%
@@
 
-```{r plot-raw-data-env, fig.height = 8}
-ggplot(biodiv_ferns_cent_env, aes(x = .panel_x, y = .panel_y)) + 
+```{r plot-raw-data, fig.height = 8}
+ggplot(biodiv_ferns_cent, aes(x = .panel_x, y = .panel_y)) + 
   geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
@@
   facet_matrix(
-    cols = vars(temp, precip, precip_season),
-    rows = vars(richness, pd_obs_z, fd_obs_z, rpd_obs_z, rfd_obs_z)) +
-  labs(title = "Full dataset")
+    cols = vars(temp, precip, precip_season, percent_apo),
+    rows = vars(richness, pd_obs_z, fd_obs_z, rpd_obs_z, rfd_obs_z))
 ```
@@
 # Remove the outlier first
-biodiv_ferns_cent_repro %>% 
+biodiv_ferns_cent %>% 
   filter(percent_apo != max(percent_apo)) %>%
@@
     cols = vars(percent_apo, temp, precip, precip_season),
-    rows = vars(pd_obs_z, rpd_obs_z)) +
-  labs(title = "Repro dataset")
+    rows = vars(pd_obs_z, rpd_obs_z))
 ```
@@
   richness ~ temp + I(temp^2) + precip + precip_season, 
-  data = biodiv_ferns_cent_env, family = poisson)
+  data = biodiv_ferns_cent, family = poisson)
 

### 2021-07-08 ms/SI.Rmd
Output tables directly to pdf

@@
 
-```{r load-data-si, cache = FALSE}
+```{r libraries-no-run-in-plan, eval = FALSE}
 # This Rmd is meant to be knit as part of the target plan (R/_targets.R).
 # Run the lines below if rendering outside of targets plan
-# library(targets)
-# library(tarchetypes)
-# source(here::here("R/packages.R"))
-# source(here::here("R/functions.R"))
+library(targets)
+library(tarchetypes)
+source(here::here("R/packages.R"))
+source(here::here("R/functions.R"))
+```
+
+```{r load-data-si, cache = FALSE}
+library(flextable)
+library(ftExtra)
 
@@
   "phy_sig_results",
-  "binary_sig_results"
+  "binary_sig_results",
+  "lrt_results"
 ))
@@
 phy_sig_results %>%
-  rename(K = kval) %>%
-  mutate(K = round(K, 3)) %>% 
-  mutate(lambda = round(lambda, 3)) %>%
-  mutate(lambda_pval = format(lambda_pval, digits = 3)) %>%
-  write_csv(path = result_file(s_table("phy-sig"), "csv"))
+  mutate(across(c(kval, lambda, k_pval, lambda_pval), ~format(., digits = 2))) %>%
+  rename(
+    ` ` = trait,
+    `*K*` = kval,
+    `λ` = lambda,
+    `*p* (*K*)` = k_pval,
+    `*p* (λ)` = lambda_pval,
+    `*n*` = n
+    ) %>%
+  as_flextable() %>%
+  colformat_md(part = "all")
 ```
 
-**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan. K, Blomberg's *K*; lambda, Pagel's lambda.
+**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan. *K*, Blomberg's *K*; λ, Pagel's λ.
 
@@
 binary_sig_results %>%
+  select(trait, D, num_present, num_absent, prob_random, prob_brownian) %>%
   arrange(trait) %>%
-  write_csv(path = result_file(s_table("phy-sig-binary"), "csv"))
+  mutate(across(c(D, prob_random, prob_brownian), ~format(., digits = 2))) %>%
+  rename(
+    ` ` = trait,
+    `*n* (present)` = num_present,
+    `*n* (absent)` = num_absent,
+    `*p* (random)` = prob_random,
+    `*p* (Brownian)` = prob_brownian
+    ) %>%
+  as_flextable() %>%
+  colformat_md(part = "all")
 ```
@@
 Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.
-num_present, number of times a trait was observed present; num_absent, number of times a trait was observed absent; D, Fritz and Purvis' *D*; prob_random, probability of obtaining observed trait distribution given random distribution of traits; prob_brownian, probability of obtaining observed trait distribution given random evolution by Brownian Motion.
+D, Fritz and Purvis' *D*; *n* (present), number of times a trait was observed present; *n* (absent), number of times a trait was observed absent; *p* (random), probability of obtaining observed trait distribution given random distribution of traits; *p* (Brownian), probability of obtaining observed trait distribution given random evolution by Brownian Motion.
 Presence or absence of indusium was originally a binary trait; other traits (margin, shape, texture) are qualitative traits scored in binary format.
+
+```{r lrt-null}
+lrt_results %>%
+  filter(comparison == "null_model") %>%
+  select(-null_formula, -comparison) %>%
+  tidyr::extract(full_formula, "model_type", "~ ([^ ]+) ") %>%
+  rename_resp_vars %>%
+  rename_model_type %>%
+  select(model_type, resp_var, chi2_LR, df, loglik_null, loglik_full, p_value) %>%
+  arrange(model_type, resp_var) %>%
+  mutate(across(where(is.double), ~format(., digits = 2))) %>%
+  # colformat_md() doesn't work on grouped tables
+  # group_by(model_type) %>%
+  # as_flextable(hide_grouplabel = TRUE) %>%
+  rename(
+    Response = resp_var,
+    `Model type` = model_type,
+    `χ^2^` = chi2_LR,
+    `logL (full model)` = loglik_full,
+    `logL (null model)` = loglik_null,
+    `*p*` = p_value
+  ) %>%
+  as_flextable %>%
+  merge_v(j = "Model type") %>% 
+  valign(valign = "top") %>% 
+  colformat_md(part = "all")
+```
+
+**`r s_table("lrt-null")`**.
+Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).
+Higher (less negative) log likelihood indicates better fit.
+Full environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
+Full reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
+PD, phylogenetic diversity; RPD, relative phylogenetic diversity; FD, functional diversity; RFD, relative functional diversity; SES, standard effect size.
 

### 2021-07-08 R/functions.R

@@ -2840,3 +2840,52 @@ add_selected_pred_to_model <- function(spatial_models) {
     assert(not_na, indep_var_select)
   
 }
+
+# Manuscript rendering ----
+
+#' Rename response variables in data frame
+#' 
+#' For formatting results tables
+#'
+#' @param df 
+#'
+#' @return df with renamed variables
+#' 
+rename_resp_vars <- function(df) {
+  
+  lookup <-
+    tibble(
+      resp_var = c("richness", "fd_obs_z", "pd_obs_z", "rfd_obs_z", "rpd_obs_z"),
+      resp_var_print = c("Richness", "SES of FD", "SES of PD", "SES of RFD", "SES of RPD")
+    )
+  
+  df %>%
+    left_join(lookup, by = "resp_var") %>%
+    select(-resp_var) %>%
+    rename(resp_var = resp_var_print)
+  
+}
+
+#' Rename model type in data frame
+#' 
+#' For formatting results tables
+#'
+#' @param df 
+#'
+#' @return df with renamed model types
+#' 
+rename_model_type <- function(df) {
+  
+  lookup <-
+    tibble(
+      model_type = c("percent_apo", "temp"),
+      model_type_print = c("Reproductive", "Environmental")
+    )
+  
+  df %>%
+    left_join(lookup, by = "model_type") %>%
+    select(-model_type) %>%
+    rename(model_type = model_type_print)
+  
+}
+

### 2021-07-08 ms/SI.Rmd
WIP try to fix paths to refs in yaml header

@@
 title: "Exploring dimensions of biodiversity in Japanese ferns"
+date: '`r format(Sys.time(), "%d %B, %Y")`'
 bibliography: 
-  - references.yaml
-  - references_other.yaml
+  - /home/rstudio/japan_ferns_biogeo/repos/japan_ferns_biogeo/ms/references.yaml
+  - /home/rstudio/japan_ferns_biogeo/repos/japan_ferns_biogeo/ms/references_other.yaml
 mainfont: Roboto-Regular.ttf

### 2021-07-08 ms/SI.Rmd
Fix paths to bibliography problem

--- 2021-07-08 ms/SI.Rmd
@@
 title: "Exploring dimensions of biodiversity in Japanese ferns"
-bibliography: 
-  - references.yaml
-  - references_other.yaml
+date: '`r format(Sys.time(), "%d %B, %Y")`'
 mainfont: Roboto-Regular.ttf
@@
 # nocite: | 
+---
+
+---
+bibliography:
+  - `r here::here("ms/references.yaml")`
+  - `r here::here("ms/references_other.yaml")`
 ---
@@
 
-```{r phy-sig}
+```{r phy-sig, results = "as-is"}
 phy_sig_results %>%
@@
   as_flextable() %>%
+  autofit() %>%
   colformat_md(part = "all")
@@
 
-```{r phy-sig-binary}
+\clearpage
+
+```{r phy-sig-binary, results = "as-is"}
 binary_sig_results %>%
@@
   as_flextable() %>%
+  fontsize(size = 7, part = "all") %>%
+  autofit() %>% 
   colformat_md(part = "all")
@@
 
-```{r lrt-null}
+\clearpage
+
+```{r lrt-null, results = "as-is"}
 lrt_results %>%
@@
   valign(valign = "top") %>% 
+  fontsize(size = 7, part = "all") %>%
+  autofit() %>%
   colformat_md(part = "all")
@@
 PD, phylogenetic diversity; RPD, relative phylogenetic diversity; FD, functional diversity; RFD, relative functional diversity; SES, standard effect size.
+
+\clearpage
 

### 2021-07-08 ms/manuscript.Rmd
Update MS
change pagebreak() to \newpage
Output tables directly to pdf
Fix paths to bibliography problem

@@
   - BoldItalicFont=Roboto-BoldItalic.ttf
-bibliography: 
-  - references.yaml
-  - references_other.yaml
 output: 
@@
   chunk_output_type: console
+---
+
+---
+bibliography:
+  - `r here::here("ms/references.yaml")`
+  - `r here::here("ms/references_other.yaml")`
 ---
@@
 
-`r pagebreak()`
+\newpage
 
@@
 3.
+What environmental and biological factors influence biodiversity?
+4.
 How well is biodiversity protected?
@@
 
-All models were significant relative to null models without any fixed effects.
+All models were significant relative to null models without any fixed effects `r s_table("lrt-null")`.
 The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.

### 2021-07-08 _targets.R

@@ -624,6 +624,7 @@ tar_plan(
   # - doc
   tar_render(
     ms_doc,
+    knit_root_dir = here::here(),
     path = "ms/manuscript.Rmd",
     output_format = "bookdown::word_document2",
     output_file = here::here("results/manuscript.docx"),
@@ -632,6 +633,7 @@ tar_plan(
   # - pdf
   tar_render(
     ms_pdf,
+    knit_root_dir = here::here(),
     path = "ms/manuscript.Rmd",
     output_format = "bookdown::pdf_document2",
     output_file = here::here("results/manuscript.pdf"),
@@ -640,7 +642,7 @@ tar_plan(
   # - SI pdf
   tar_render(
     si_pdf,
-    knit_root_dir = here::here("ms"),
+    knit_root_dir = here::here(),
     path = "ms/SI.Rmd",
     output_format = "bookdown::pdf_document2",
     output_file = here::here("results/supp_info.pdf"),
@@ -649,6 +651,7 @@ tar_plan(
   # - SI appendix on data exploration for models
   tar_render(
     si_data_exploration,
+    knit_root_dir = here::here(),
     path = "ms/data_exploration.Rmd",
     output_format = "rmarkdown::html_document",
     output_file = here::here("results/data_exploration.html"),

### 2021-07-08 ms/SI.Rmd
Switch to kableExtra for tables
still need to format column names properly

@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
+
+# Load packages only used when knitting Rmd
+library(kableExtra)
 ```
@@
 ```{r load-data-si, cache = FALSE}
-library(flextable)
-library(ftExtra)
-
 tar_load(c(
@@
   mutate(across(c(kval, lambda, k_pval, lambda_pval), ~format(., digits = 2))) %>%
-  rename(
-    ` ` = trait,
-    `*K*` = kval,
-    `λ` = lambda,
-    `*p* (*K*)` = k_pval,
-    `*p* (λ)` = lambda_pval,
-    `*n*` = n
-    ) %>%
-  as_flextable() %>%
-  autofit() %>%
-  colformat_md(part = "all")
+  kbl(booktabs = TRUE) %>%
+  kable_styling(latex_options = "striped")
 ```
@@
   mutate(across(c(D, prob_random, prob_brownian), ~format(., digits = 2))) %>%
-  rename(
-    ` ` = trait,
-    `*n* (present)` = num_present,
-    `*n* (absent)` = num_absent,
-    `*p* (random)` = prob_random,
-    `*p* (Brownian)` = prob_brownian
-    ) %>%
-  as_flextable() %>%
-  fontsize(size = 7, part = "all") %>%
-  autofit() %>% 
-  colformat_md(part = "all")
+  kbl(booktabs = TRUE) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
 ```
@@
 ```{r lrt-null, results = "as-is"}
+lrt_table_data <-
 lrt_results %>%
@@
   arrange(model_type, resp_var) %>%
-  mutate(across(where(is.double), ~format(., digits = 2))) %>%
-  # colformat_md() doesn't work on grouped tables
-  # group_by(model_type) %>%
-  # as_flextable(hide_grouplabel = TRUE) %>%
-  rename(
-    Response = resp_var,
-    `Model type` = model_type,
-    `χ^2^` = chi2_LR,
-    `logL (full model)` = loglik_full,
-    `logL (null model)` = loglik_null,
-    `*p*` = p_value
-  ) %>%
-  as_flextable %>%
-  merge_v(j = "Model type") %>% 
-  valign(valign = "top") %>% 
-  fontsize(size = 7, part = "all") %>%
-  autofit() %>%
-  colformat_md(part = "all")
+  mutate(across(where(is.double), ~format(., digits = 2)))
+
+lrt_table_data %>%
+  select(-model_type) %>%
+  kbl(
+    col.names = colnames(.) %>% str_replace_all("resp_var", ""),
+    booktabs = TRUE) %>%
+  pack_rows(index = base::table(lrt_table_data$model_type)) %>%
+  kable_styling(latex_options = "striped")
 ```

### 2021-07-08 ms/SI.Rmd
Hold table position

@@
   kbl(booktabs = TRUE) %>%
-  kable_styling(latex_options = "striped")
+  kable_styling(latex_options = c("striped", "HOLD_position"))
 ```
@@
   kbl(booktabs = TRUE) %>%
-  kable_styling(latex_options = c("striped", "scale_down"))
+  kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
 ```
@@
   pack_rows(index = base::table(lrt_table_data$model_type)) %>%
-  kable_styling(latex_options = "striped")
+  kable_styling(latex_options = c("striped", "HOLD_position"))
 ```

### 2021-07-09 ms/data_exploration.Rmd
Fix using biodiv_ferns_spatial with area in data_exploration.Rmd

@@
 tar_load(indep_vars)
-tar_load(lat_area_ja)
 
 # Make biodiversity metrics dataframe with centroid of each site for models
-  biodiv_ferns_cent <- sf_to_centroids(biodiv_ferns_spatial) %>%
-    # add area
-    add_roll_area(lat_area_ja) %>%
+  biodiv_ferns_cent <- 
+    biodiv_ferns_spatial %>%
     # keep only variables needed for model and only rows with zero missing data
-    filter_data_for_model(c("grids", "lat", "long", resp_vars_env, indep_vars))
+    filter_data_for_model(c("grids", "lat", "long", resp_vars_env, indep_vars)) %>%
+    sf::st_set_geometry(NULL) %>%
+    as_tibble()
 ```

### 2021-07-09 _targets.R

@@ -53,7 +53,7 @@ tar_plan(
   # downloaded from https://www.gsi.go.jp/kankyochiri/gm_japan_e.html
   # on 2020-08-26
   tar_file(japan_pol_file, "data_raw/gm-jpn-all_u_2_2/polbnda_jpn.shp"),
-  jpn_pol = sf::st_read(japan_pol_file),
+  jpn_pol = st_read(japan_pol_file),
   
   # Collapse all the political units down to just one shape for the country
   japan_shp = jpn_pol %>%
@@ -425,7 +425,7 @@ tar_plan(
   # Make biodiversity metrics dataframe with centroid of each site for models
   biodiv_ferns_cent = biodiv_ferns_spatial %>%
     # drop geometry (only need centroids for modeling)
-    sf::st_set_geometry(NULL) %>%
+    st_set_geometry(NULL) %>%
     # drop single percent_apo outlier
     drop_apo_outlier %>%
     # keep only variables needed for model and only rows with zero missing data
@@ -435,7 +435,7 @@ tar_plan(
   
   # Check for correlation between independent variables in repro data
   t_test_results = run_mod_ttest_ja(
-    biodiv_ferns_spatial,
+    st_set_geometry(biodiv_ferns_spatial, NULL),
     vars_select = c("temp", "temp_season", "precip", "precip_season", "percent_apo", "area")
   ),
   
@@ -528,7 +528,7 @@ tar_plan(
   # - low: protected area, but none of the above restrictions
   
   # 1: wilderness
-  protected_1 = sf::st_read("data_raw/map17/原生自然環境保全地域_国指定自然環境保全地域.shp") %>%
+  protected_1 = st_read("data_raw/map17/原生自然環境保全地域_国指定自然環境保全地域.shp") %>%
     mutate(
       status = case_when(
         ZONE == 1 ~ "high", # 1＝原生自然環境保全地域
@@ -539,7 +539,7 @@ tar_plan(
     ),
   
   # 2: quasi-national parks
-  protected_2 = sf::st_read("data_raw/map17/国定公園.shp") %>%
+  protected_2 = st_read("data_raw/map17/国定公園.shp") %>%
     mutate(
       status = case_when(
         ZONE == 1 ~ "high", # 1＝特別保護地区
[…]

### 2021-07-15 ms/manuscript.Rmd
Convert body text to one line per sentence
Fix fetching data used for model in MS
don't use sf_to_centroids()

@@
   "model_fits",
-  "spatial_models"
+  "biodiv_ferns_cent"
 ))
@@
 
-Sequences for the plastid, coding *rbcL* gene are available for ca.
-97% of the Japanese fern flora [@Ebihara2019b].
+Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
 However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating.
-Because one goal of the study is to understand the distribution of paleo- vs.
-neo-endemic areas (defined in units of time), we require an ultrametric tree.
+Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
 To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (
@@
 # Extract dataset used for model, including significance columns
-biodiv_ferns_used_for_model <- sf_to_centroids(biodiv_ferns_spatial) %>%
+biodiv_ferns_used_for_model <- biodiv_ferns_spatial %>%
+  # drop geometry (only need centroids for modeling)
+  st_set_geometry(NULL) %>%
+  as_tibble() %>%
   # filter to only sites used in model
   inner_join(
-    spatial_models$model[[1]]$data %>% as_tibble() %>% select(grids),
+    biodiv_ferns_cent %>% as_tibble() %>% select(grids),
     by = "grids"

### 2021-07-15 ms/manuscript.Rmd
Update MS

@@
 
-### Climate data
+### Environmental data
 
@@
 We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
+We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
+This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that simply using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
 The R package `r pack("sf")` was used for GIS analyses.
@@
 Unlike seed plants, ferns have gametophytes that are capable of growing independently from the sporophyte. 
-Futhermore, the two halves of the life cycle have vastly different morphology and ecophysiology, and may even occur over partially to completely disjunct ranges [reviewed in @Pinson2016a]. 
-Here, as in the vast majority of fern ecology studies, we consider only the sporophyte phase. 
+Futhermore, the two stages of the life cycle have vastly different morphology and ecophysiology, and may even occur over partially to completely disjunct ranges [reviewed in @Pinson2016a]. 
+Here, as in the vast majority of fern ecology studies, we consider only the sporophyte stage. 
 However, recent studies in Japan [@Ebihara2013; @Ebihara2019a] and elsewhere [@Nitta2017] have revealed different patterns in the distribution of gametophytes and sporophytes. 
-The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA barcoding analyses to compare patterns of biodiversity between life cycles in ferns across large spatial scales.
+The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA barcoding analyses to compare patterns of biodiversity between life stages in ferns across large spatial scales.
 

### 2021-07-15 ms/SI.Rmd
Add SI environment fig

@@
   "binary_sig_results",
-  "lrt_results"
+  "lrt_results",
+  "japan_shp"
 ))
@@
 
-```{r abundance, width = 7, height = 5}
+```{r envir, fig.width = 7, fig.height = 9}
+
+# Restrict sites to only those used in model
+# (drops outlier % apomictic site)
+biodiv_ferns_for_plotting <-
+  biodiv_ferns_spatial %>%
+  drop_apo_outlier
+
+# crop high-res map of Japan to spatial div results area
+japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_for_plotting))
+
+# Set CRS
+japan_crs <- sf::st_crs(japan)
+biodiv_ferns_for_plotting <- sf::st_set_crs(biodiv_ferns_for_plotting, japan_crs)
+
+# Define function for plotting environment
+plot_env <- function(part, label, var_select, fill_lab = "", labels = ggplot2::waiver()) {
+  ggplot() +
+    geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+    geom_sf(data = biodiv_ferns_for_plotting, aes(fill = .data[[var_select]]), color = "transparent", alpha = 0.8) +
+    annotate(
+      geom="richtext", y = Inf, x = -Inf, 
+      label = glue("<span style='color:black'>**{part}** {label}</span>"), 
+      hjust = "inward", vjust = "inward", color = "white") + 
+    labs(fill = fill_lab) +
+    scale_fill_scico(palette = "lajolla", labels = labels) +
+    scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
+    theme_grey(base_size = 10) +
+    theme(
+      panel.grid.major = element_line(color = "grey60", size = 0.1),
+      panel.grid.minor = element_blank(),
+      panel.background = element_rect(fill = "transparent"),
+      axis.ticks = element_blank(),
+      axis.text = element_text(color = "grey60"),
+      axis.title = element_blank(),
+      legend.position = "right",
+      legend.title = element_markdown()
+    )
+}
+
+a <- plot_env("A", "Area", "area", "km^2")
+b <- plot_env("B", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
+c <- plot_env("C", "Precipitation", "precip", "mm")
+d <- plot_env("D", "Precip. season.", "precip_season" , "")
+e <- plot_env("E", "Temperature", "temp", "°C", function(x) x/10)
+f <- plot_env("F", "Temp. season.", "temp_season")
+
+a + b + c + d + e + f + plot_layout(ncol = 2, nrow = 3)
+
+ggsave(result_file(figure("envir"), "pdf"), width = 6, height = 8, units = "in")
+```
+
+**`r s_figure("envir")`**.
+Reproductive mode and environmental data in 0.2 degree grid cells across Japan.
+**A**, rolling mean area (km^2^) in 1° latitudinal bands.
+**B**, % apomictic taxa.
+**C**, annual precipitation **D**, percipitation seasonality (coefficient of variation).
+**E**, mean annual temperature.
+**F**, temperature seasonality (standard deviation × 100).
+
+\clearpage
+
+```{r abundance, fig.width = 7, fig.height = 5}
 k <- scales::unit_format(unit = "k", scale = 1e-3, accuracy = 1, sep = "")

### 2021-07-15 ms/manuscript.Rmd
Add SI environment fig

@@
 We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown. 
-We calculated the percentage of apomictic species in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa.
+We calculated the percentage of apomictic species in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).
 
@@
 
-We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15).
+We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`).
 We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.

### 2021-07-15 ms/SI.Rmd
Put Fig S7 in correct order in PDF
Don't write out separate files for SI figures

@@
   )
-
-ggsave(result_file(s_figure("sampling-curve"), "pdf"), width = 4, height = 3, units = "in")
 ```
@@
   plot_annotation(tag_levels = 'A')
-
-ggsave(result_file(s_figure("grain-size"), "pdf"), width = 7, height = 3, units = "in")
 ```
@@
 a + b + c + d + e + f + plot_layout(ncol = 2, nrow = 3)
-
-ggsave(result_file(figure("envir"), "pdf"), width = 6, height = 8, units = "in")
 ```
@@
   plot_layout(nrow = 1)
-
-ggsave(result_file(figure("abundance"), "pdf"), width = 7, height = 5, units = "in")
 ```
@@
 **`r s_figure("abundance")`**. Observed number of specimens (**A**) and sampling redundancy (**B**) per 0.2° grid cell in the ferns of Japan.
+
+\clearpage
+
+```{r raw-biplot, fig.width = 7, fig.height = 7}
+a <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = pd_obs)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Richness", y = "PD")
+
+b <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = rpd_obs)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Richness", y = "RPD")
+
+c <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = fd_obs)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Richness", y = "FD")
+
+d <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = rfd_obs)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Richness", y = "RFD")
+
+a + b + c + d +
+  plot_layout(ncol = 2, nrow = 2) +
+  plot_annotation(tag_levels = 'A')
+```
+
+**`r s_figure("raw-biplot")`**.
+Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.
+**A**, Phylogenetic diversity (PD).
+**B**, Relative phylogenetic diversity (RPD).
+**C**, Functional diversity (FD).
+**D**, Relative functional diversity (RFD).
+
+\clearpage
+
+```{r endemism-restricted, fig.width = 5, fig.height = 5}
+ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
+  scale_fill_manual(values = canape_cols) +
+  map_theme() +
+  theme(legend.title = element_blank())
+```
+
+**`r s_figure("endemism-restricted")`**.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.
+'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+For details, see Methods.
 
@@
   plot_annotation(tag_levels = 'A')
-
-ggsave(result_file(s_figure("k-plot"), "pdf"), width = 7, height = 3, units = "in")
 ```
@@
 Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal; see Methods).
-
-\clearpage
-
-```{r raw-biplot, fig.width = 7, fig.height = 7}
-a <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = pd_obs)) +
-  geom_point(alpha = 0.5) +
-  labs(x = "Richness", y = "PD")
-
-b <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = rpd_obs)) +
-  geom_point(alpha = 0.5) +
-  labs(x = "Richness", y = "RPD")
-
-c <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = fd_obs)) +
-  geom_point(alpha = 0.5) +
-  labs(x = "Richness", y = "FD")
-
-d <- ggplot(biodiv_ferns_spatial, aes(x = richness, y = rfd_obs)) +
-  geom_point(alpha = 0.5) +
-  labs(x = "Richness", y = "RFD")
-
-a + b + c + d +
-  plot_layout(ncol = 2, nrow = 2) +
-  plot_annotation(tag_levels = 'A')
-
-ggsave(result_file(s_figure("raw-biplot"), "pdf"), width = 7, height = 3, units = "in")
-```
-
-**`r s_figure("raw-biplot")`**.
-Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.
-**A**, Phylogenetic diversity (PD).
-**B**, Relative phylogenetic diversity (RPD).
-**C**, Functional diversity (FD).
-**D**, Relative functional diversity (RFD).
-
-\clearpage
-
-```{r endemism-restricted, fig.width = 5, fig.height = 5}
-ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
-  scale_fill_manual(values = canape_cols) +
-  map_theme() +
-  theme(legend.title = element_blank())
-
-ggsave(result_file(s_figure("endemism-restricted"), "pdf"), width = 5, height = 5, units = "in")
-```
-
-**`r s_figure("endemism-restricted")`**.
-Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.
-'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
-For details, see Methods.

### 2021-07-15 ms/SI.Rmd
Don't keep tex for SI

@@
     fig_caption: yes
-    keep_tex: yes
+    keep_tex: no
     latex_engine: xelatex

### 2021-07-15 ms/SI.Rmd
Update MS and SI
Add SI tables, figure. Fix usage of latex formatting in tables.

@@
   "lrt_results",
-  "japan_shp"
+  "japan_shp",
+  "morans_i",
+  "t_test_results",
+  "aic_env_repro"
 ))
@@
 
+```{r moran, results = "as-is"}
+morans_i %>%
+  select(resp_var = var, morans_i, pval_i) %>%
+  mutate(
+    across(c(morans_i, pval_i), ~format(., digits = 2)),
+    resp_var = clean_vars(resp_var) %>% 
+      factor(levels = c("Area", "% apomictic taxa", "Precipitation", "Precipitation seasonality", "Temperature", "Richness", "SES of PD", "SES of RPD", "SES of FD", "SES of RFD"))
+    ) %>%
+  assert(not_na, resp_var) %>%
+  arrange(resp_var) %>%
+  # format for kable/latex: since I use latex in headers, must escape latex elsewhere
+  mutate(resp_var = str_replace_all(resp_var, "%", "\\\\%")) %>%
+  rename(`\\textit{I}` = morans_i, `\\textit{P}` = pval_i, "Variable" = resp_var) %>%
+  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
+  kable_styling(latex_options = c("striped", "HOLD_position"))
+```
+
+**`r s_table("moran")`**.
+Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.
+Significant *I* > 0 indicates autocorrelation.
+
+\clearpage
+
+```{r corr, results = "as-is"}
+t_test_results_for_table <-
+t_test_results %>%
+  mutate(
+    var1 = clean_vars(var1),
+    var2 = clean_vars(var2)) %>%
+  arrange(var1, var2)
+
+# Identify which rows to make bold based on p < 0.05
+t_test_signif <- t_test_results_for_table$p_value < 0.05
+t_test_signif[is.na(t_test_signif)] <- FALSE
+t_test_signif <- which(t_test_signif)
+
+t_test_results_for_table %>%
+  mutate(across(where(is.double), ~format(., digits = 2))) %>%
+  # format for kable/latex
+  mutate(
+    var1 = str_replace_all(var1, "%", "\\\\%"),
+    var2 = str_replace_all(var2, "%", "\\\\%")) %>%
+  rename(
+    "Variable 1" = var1,
+    "Variable 2" = var2,
+    `\\textit{P}` = p_value,
+    `\\textit{r}` = corr,
+    `\\textit{f}` = f_stat,
+    "df" = dof) %>%
+  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
+  row_spec(row = t_test_signif, bold = TRUE) %>%
+  kable_styling(latex_options = c("striped", "HOLD_position"))
+```
+
+**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position. Significantly correlated variables at *P* < 0.05 in bold.
+
+\clearpage
+
 ```{r phy-sig, results = "as-is"}
@@
   mutate(across(c(kval, lambda, k_pval, lambda_pval), ~format(., digits = 2))) %>%
-  kbl(booktabs = TRUE) %>%
+  # format for kable/latex
+  mutate(trait = str_replace_all(trait, "_", "\\\\_")) %>%
+  rename(
+    "Trait" = trait,
+    `\\textit{K}` = kval, `\\textit{P}(\\textit{K})` = k_pval, 
+    `λ` = lambda, `\\textit{P}(λ)` = lambda_pval,
+    `\\textit{N}` = n) %>%
+  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
   kable_styling(latex_options = c("striped", "HOLD_position"))
@@
   arrange(trait) %>%
-  mutate(across(c(D, prob_random, prob_brownian), ~format(., digits = 2))) %>%
-  kbl(booktabs = TRUE) %>%
+  # format for kable/latex
+  mutate(trait = str_replace_all(trait, "_", "\\\\_")) %>%
+    rename(
+    "Trait" = trait,
+    `\\textit{N} present` = num_present,
+    `\\textit{N} absent` = num_absent,
+    `\\textit{D}` = D,
+    `\\textit{P}(random)` = prob_random, 
+    `\\textit{P}(Brownian)` = prob_brownian
+    ) %>%
+  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
   kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
@@
 Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.
-D, Fritz and Purvis' *D*; *n* (present), number of times a trait was observed present; *n* (absent), number of times a trait was observed absent; *p* (random), probability of obtaining observed trait distribution given random distribution of traits; *p* (Brownian), probability of obtaining observed trait distribution given random evolution by Brownian Motion.
+*D*, Fritz and Purvis' *D*; *N* (present), number of times a trait was observed present; *N* (absent), number of times a trait was observed absent; *P*(random), probability of obtaining observed trait distribution given random distribution of traits; *P*(Brownian), probability of obtaining observed trait distribution given random evolution by Brownian Motion.
 Presence or absence of indusium was originally a binary trait; other traits (margin, shape, texture) are qualitative traits scored in binary format.
@@
 ```{r lrt-null, results = "as-is"}
+# Need to keep the input data separate for pack_rows()
 lrt_table_data <-
@@
   select(-model_type) %>%
-  kbl(
-    col.names = colnames(.) %>% str_replace_all("resp_var", ""),
-    booktabs = TRUE) %>%
+  # format for kable/latex
+    rename(
+    "Response variable" = resp_var,
+    `χ\\textsuperscript{2}` = chi2_LR,
+    `LogL(null)` = loglik_null,
+    `LogL(full)` = loglik_full,
+    `\\textit{P}` = p_value
+    ) %>%
+  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
   pack_rows(index = base::table(lrt_table_data$model_type)) %>%
@@
 PD, phylogenetic diversity; RPD, relative phylogenetic diversity; FD, functional diversity; RFD, relative functional diversity; SES, standard effect size.
+
+\clearpage
+
+```{r aic, results = "as-is"}
+aic_table_data <- aic_env_repro %>%
+  rename_resp_vars %>%
+  rename_model_type %>%
+  select(model_type, resp_var, conditional_aic, df = effective_df, loglik) %>%
+  arrange(model_type, resp_var) %>%
+  mutate(across(where(is.double), ~format(., digits = 2)))
+
+aic_table_data %>%
+  select(-model_type) %>%
+  # format for kable/latex
+    rename(
+    "Response variable" = resp_var,
+    "cAIC" = conditional_aic,
+    "LogL" = loglik
+    ) %>%
+  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
+  pack_rows(index = base::table(aic_table_data$model_type)) %>%
+  kable_styling(latex_options = c("striped", "HOLD_position"))
+```
+
+**`r s_table("aic")`**.
+Model fit as measured with conditional Akaike Information Criterion (cAIC).
+Environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
+Reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
+Lower cAIC indicates better fit.
 
@@
 # Set color palettes
-# Randomization test significance
+# - Randomization test significance
 signif_cols <-
@@
   )
+# - Bioregions
+bioregion_cols <- rcartocolor::carto_pal(9, "Bold")[1:8] %>%
+  set_names(1:8)
 ```
@@
 a + b + c +
-  plot_annotation(tag_levels = 'A')
+  plot_annotation(tag_levels = "A") &
+    theme(plot.tag = element_text(face = "bold"))
 ```
@@
   plot_layout(ncol = 2, nrow = 2) +
-  plot_annotation(tag_levels = 'A')
+  plot_annotation(tag_levels = "A") &
+    theme(plot.tag = element_text(face = "bold"))
 ```
@@
   plot_layout(ncol = 2) +
-  plot_annotation(tag_levels = 'A')
+  plot_annotation(tag_levels = "A") &
+    theme(plot.tag = element_text(face = "bold"))
 ```
@@
 Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal; see Methods).
+
+\clearpage
+
+```{r bioregions-env, fig.width = 7, fig.height = 5}
+# Make pot showing bioregions by temperature and precipitation
+a <-
+ggplot(biodiv_ferns_spatial, aes(x = precip, y = temp, color = taxonomic_cluster)) +
+  geom_point(size = 1, alpha = 0.7) +
+  scale_color_manual(values = bioregion_cols) +
+  scale_y_continuous(labels = function(x) x/10) +
+  labs(y = "Temperature (°C)", x = "Precipitation (mm)", color = "Cluster", subtitle = "Taxonomic bioregions") +
+  theme(legend.position = "bottom")
+
+b <-
+  ggplot(biodiv_ferns_spatial, aes(x = precip, y = temp, color = phylo_cluster)) +
+  geom_point(size = 1, alpha = 0.7) +
+  scale_color_manual(values = bioregion_cols) +
+  scale_y_continuous(labels = function(x) x/10) +
+  labs(y = "Temperature (°C)", x = "Precipitation (mm)", color = "Cluster", subtitle = "Phylogenetic bioregions") +
+  theme(legend.position = "bottom")
+
+a + b + plot_annotation(tag_levels = "A") &
+    theme(plot.tag = element_text(face = "bold"))
+```
+
+**`r s_figure("bioregions-env")`**.
+Scatterplots of grid cell membership in (**A**) taxonomic or (**B**) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
+Points shown with partial transparency; darker areas indicate overlapping points.

### 2021-07-15 ms/manuscript.Rmd
Update MS and SI
Add SI tables, figure. Fix usage of latex formatting in tables.

@@
 
-Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1, 3, 4, 5, 6^, and Atsushi Ebihara^7^
+Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1,3,4,5,6^, and Atsushi Ebihara^7^
 
@@
 
-All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package (https://github.com/joelnitta/canaper), which calculates raw biodiversity metrics using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].
+All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package [@Nitta2021], which calculates raw biodiversity metrics using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].
 
@@
 We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the `r pack("spdep")` package.
-In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05). 
+In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05) (`r s_table("moran")`). 
 Therefore, we used spatially explicit methods for modeling.
@@
 We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the `r pack("SpatialPack")` R package.
-This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature, so for the main modeling analysis we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
+This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature (`r s_table("corr")`), so for the main modeling analysis we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
 
@@
 Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. 
-Raw and processed data are openly available from the Dryad repository at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
+A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
+Data are openly available from Figshare at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
 
@@
 SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F). 
-Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD.
+Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (`r s_table("aic")`).
 
@@
 The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
+Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic Cluster 3 (mostly the Ryukyu islands, wetter) from Cluster 4 (mostly the Ogasawara islands, drier) (`r s_figure("bioregions-env")`).
 

### 2021-07-15 R/functions.R

@@ -2906,3 +2906,21 @@ rename_model_type <- function(df) {
   
 }
 
+#' Clean a vector of variables for printing
+#'
+#' @param x Character vector (variables)
+#'
+#' @return Character vector
+#' 
+clean_vars <- function(x) {
+  x %>%
+    gsub("^([^_]+)_obs_z", "SES of \\U\\1", ., perl=TRUE) %>%
+    str_replace_all("^percent_apo$", "% apomictic taxa") %>%
+    str_replace_all("^temp$", "Temperature") %>%
+    str_replace_all("^temp_season$", "Temperature seasonality") %>%
+    str_replace_all("^precip_season$", "Precipitation seasonality") %>%
+    str_replace_all("^precip$", "Precipitation") %>%
+    str_replace_all("^area$", "Area") %>%
+    str_replace_all("^richness$", "Richness")
+}
+

### 2021-07-26 ms/SI.Rmd
Update MS
add maps of genera, barplot of % apo

@@
 library(kableExtra)
+library(scales)
 ```
@@
   "t_test_results",
-  "aic_env_repro"
+  "aic_env_repro",
+  "comm_ferns",
+  "repro_data"
 ))
@@
 
+# Subset ferns biodiversity dataframe: restrict sites to only those used in model
+# (drops outlier % apomictic site)
+biodiv_ferns_for_plotting <-
+  biodiv_ferns_spatial %>%
+  drop_apo_outlier
+
 # Set color palettes
@@
 ```{r envir, fig.width = 7, fig.height = 9}
-
-# Restrict sites to only those used in model
-# (drops outlier % apomictic site)
-biodiv_ferns_for_plotting <-
-  biodiv_ferns_spatial %>%
-  drop_apo_outlier
-
-# crop high-res map of Japan to spatial div results area
-japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_for_plotting))
-
-# Set CRS
-japan_crs <- sf::st_crs(japan)
-biodiv_ferns_for_plotting <- sf::st_set_crs(biodiv_ferns_for_plotting, japan_crs)
-
 # Define function for plotting environment
@@
 
+```{r genus-map, fig.width = 7, fig.height = 9}
+# Define function for plotting genus-level abundance.
+# assumes `comm_ferns`, `japan` are in environment
+plot_genus <- function(part, genus_select) {
+
+  genus_abun <-
+    comm_ferns %>%
+    rownames_to_column("grids") %>%
+    as_tibble() %>%
+    pivot_longer(-grids, names_to = "species", values_to = "abun") %>%
+    filter(str_detect(species, genus_select)) %>%
+    group_by(grids) %>%
+    summarize(richness = sum(abun), .groups = "drop") %>%
+    left_join(select(biodiv_ferns_spatial, -richness), ., by = "grids")
+
+  ggplot() +
+    geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+    geom_sf(data = genus_abun, aes(fill = richness), color = "transparent", alpha = 0.8) +
+    annotate(
+      geom="richtext", y = Inf, x = -Inf, 
+      label = glue("<span style='color:black'>**{part}** *{genus_select}*</span>"), 
+      hjust = "inward", vjust = "inward", color = "white") + 
+    labs(fill = "Richness") +
+    scale_fill_scico(palette = "lajolla") +
+    scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
+    theme_grey(base_size = 10) +
+    theme(
+      panel.grid.major = element_line(color = "grey60", size = 0.1),
+      panel.grid.minor = element_blank(),
+      panel.background = element_rect(fill = "transparent"),
+      axis.ticks = element_blank(),
+      axis.text = element_text(color = "grey60"),
+      axis.title = element_blank(),
+      legend.position = "right",
+      legend.title = element_markdown()
+    )
+}
+
+a <- plot_genus("A", "Equisetum")
+b <- plot_genus("B", "Deparia")
+c <- plot_genus("C", "Cyrtomium")
+d <- plot_genus("D", "Dryopteris")
+e <- plot_genus("E", "Pteris")
+f <- plot_genus("F", "Diplazium")
+
+a + b + c + d + e + f + plot_layout(ncol = 2, nrow = 3)
+```
+
+**`r s_figure("genus-map")`**.
+Map of species richness for selected genera of ferns of Japan.
+**A**, *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
+**B**, *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
+**C**--**F** show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.
+
+\clearpage
+
+```{r apo-genus, fig.width = 4, fig.height = 3}
+# Bar plot of % apo taxa by genus
+percent_apo_genus <-
+repro_data %>%
+  # Remove NA values, non-ferns
+  filter(!is.na(apomict)) %>%
+  filter(str_detect(taxon, "Lycopodiella|Lycopodium|Isoetes", negate = TRUE)) %>%
+  # Summarize percent apo by genus
+  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
+  group_by(genus) %>%
+  add_count() %>%
+  group_by(genus, n) %>%
+  summarize(total_apo = sum(apomict), .groups = "drop") %>%
+  mutate(percent_apo = total_apo / n) %>%
+  arrange(desc(percent_apo))
+
+percent_apo_genus %>%
+  filter(total_apo > 1) %>%
+  group_by(genus) %>%
+  summarize(
+    n = mean(n),
+    total_apo = mean(total_apo),
+    percent_apo = mean(percent_apo),
+    .groups = "drop"
+  ) %>%
+  mutate(
+    genus =  paste0("*", genus, "*"),
+    genus = fct_reorder(genus, percent_apo),
+    label = glue::glue("{total_apo}/{n}")) %>%
+  ggplot(aes(x = percent_apo, y = genus)) + 
+  geom_col() +
+  geom_text(aes(label = label), hjust = -.3) +
+  scale_x_continuous(labels = percent, limits = c(0, .85)) +
+  labs(x = "% apomictic taxa") +
+  theme(
+    axis.text.y = element_markdown(),
+    axis.title.y = element_blank()
+  )
+```
+
+**`r s_figure("apo-genus")`**.
+Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon.
+Numbers next to each bar show number of apomictic taxa out of total richness per genus.
+
+\clearpage
+
 ```{r bioregions-env, fig.width = 7, fig.height = 5}

### 2021-07-26 ms/manuscript.Rmd
Update MS
add maps of genera, barplot of % apo

@@
 
+```{r manual, include = FALSE, evaluate = FALSE}
+# Run these lines if rendering outside of targets plan
+library(targets)
+library(tarchetypes)
+source(here::here("R/packages.R"))
+source(here::here("R/functions.R"))
+```
+
 ```{r setup, include = FALSE}
-# Run these lines if rendering outside of targets plan
-# library(targets)
-# library(tarchetypes)
-# source(here::here("R/packages.R"))
-# source(here::here("R/functions.R"))
-
 # Load libraries used for rendering Rmds
@@
 tar_load(c(
+  "repro_data",
   "occ_point_data_ferns",
@@
 
-^2^University and Jepson Herbaria, University of California, Berkeley, CA, USA
+^2^University and Jepson Herbaria, and Department of Integrative Biology, University of California, Berkeley, CA, USA
 
@@
 
-**Methods** We used a list of >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
+**Methods** We used data from >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
 We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode.
@@
 Grid-cells were grouped in to three (phylogenetic) or four (taxonomic) major bioregions. 
+Temperature and apomixis were identified as likely drivers of patterns of taxonomic and phylogenetic diversity.
 Conservation status (% protected area) is higher than the overall rate for grid-cells with significantly high biodiversity.
@@
 
+```{r intro-stats}
+# tally number of non-hybrid pteridophytes with known repro. mode
+repro_count <-
+repro_data %>%
+  count(reproductive_mode) %>%
+  mutate(mode = if_else(reproductive_mode == "unknown", "unknown", "known")) %>%
+  group_by(mode) %>%
+  summarize(total = sum(n)) %>%
+  split(.$mode)
+
+repro_count$known$total
+
+repro_count$known$percent <- (repro_count$known$total/(repro_count$unknown$total + repro_count$known$total)) %>% percent(accuracy = 0.1)
+```
+
 Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it, and to conserve it. 
@@
 However, richness alone cannot provide a complete picture of biodiversity. 
-All organisms are related by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of biodiversity.
-
-Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences) and analytic tools have only become available relatively recently.
+All organisms are related to a greater or lesser degree by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of the biodiversity present in an area.
+
+Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences), extensive spatial data, and analytic tools have only become available relatively recently.
 These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. 
-Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations [e.g., @White2019; @Daru2020a]. 
+Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations [e.g., @Laffan2016; @White2019; @Daru2020a].
 Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.
@@
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled. 
-One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for 221/721 taxa [70%]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
+One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
 
@@
 Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`). 
-The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
+The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
 We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).
 
-A necessary step in any analysis of biodiversity is to set the grain size used to define communities. 
+A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities. 
 Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
@@
 Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
-We filtered out any cells with redundancy < 0.1 and used this for subsequent analyses.
+We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.
 
@@
 
-We used traits originally compiled for identification of ferns and lycophytes of Japan as follows [@Ebihara2016b; @Ebihara2017]. 
+We used traits originally compiled for identification of ferns and lycophytes of Japan following Ebihara [-@Ebihara2016b; -@Ebihara2017]. 
 Continuous traits were measured on 10 randomly selected specimens per species. 
@@
 We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
-This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that simply using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
+This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
 The R package `r pack("sf")` was used for GIS analyses.
@@
 
-It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (the "border effect"). 
+It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (a "border effect"). 
 This is especially expected for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country. 
-To account for this, we also calculated all endemism-related metrics using a dataset including only species endemic to Japan.
+To partially account for this, we also calculated all endemism-related metrics using a dataset including only species endemic to Japan.
 
@@
 
-We postulated that % apomictic taxa might explain the distribution of genetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis ("reproductive models").
-These models included biodiversity metrics with a genetic component (SES of PD and SES of RPD) as dependent on % apomictic taxa, precipitation, and precipitation seasonality, with random spatial effects as described above.
+We postulated that % apomictic taxa might explain the distribution of phylogenetic diversity better than temperature, so we also constructed a limited set of models to test this hypothesis ("reproductive models").
+These models included biodiversity metrics with a phylogenetic component (SES of PD and SES of RPD) as dependent on % apomictic taxa, precipitation, and precipitation seasonality, with random spatial effects as described above.
 We compared the fit between environmental models and reproductive models for SES of PD and SES of RPD using conditional Akaike Information Criterion (cAIC) [@Courtiol2017; @Vaida2005].
@@
 A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
-Data are openly available from Figshare at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
+Data files, including the DNA alignment, phylogenetic tree, and community data matrix, are openly available from Figshare at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
 
@@
 Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`A, C).
-Observed values of RPD and RFD were also closely related to richness, showing high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).
+Observed values of RPD and RFD were not linearly related to richness, however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).
 
@@
 Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`).
-The vast majority of endemic grid-cells are mixed. 
-A cluster of paleoendemic cells was observed in the Ryukyu Islands. 
-Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (Daisetsuzan mountains). 
-When we ran CANAPE with a reduced dataset including only taxa endemic to Japan to account for the border effect, there were fewer significant cells, but patterns were otherwise similar to the full dataset (`r s_figure("endemism-restricted")`).
+The vast majority of grid-cells with significant endemism are mixed.
+A concentration of paleoendemism was observed in the Ryukyu Islands. 
+Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (`r figure("endemism")`). 
+When we ran CANAPE with a reduced dataset including only taxa absolutely endemic to Japan to account for the border effect, there were fewer significant cells, but patterns were otherwise similar to the full dataset (`r s_figure("endemism-restricted")`).
 
@@
 
-Here, we provide the first insights into the combined taxonomic, phylogenetic, and functional diversity of the ferns of Japan. 
-Our integrative approach reveals distinct patterns of biodiversity in different parts of the archipelago and across biodiversity metrics, and suggest reproductive mode as an important driver in some cases.
-These clear differences are reflected in the bioregionalization results, which delimit three to four major bioregions. 
-Finally, we consider the current conservation status of ferns in Japan and the implications of our findings for future conservation activity.
-
 ### Distribution and drivers of biodiversity
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015] and modeled distributions of woody plants [@Fukaya2020].
-Our modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
-Such a hump-shaped pattern is often observed in ferns on mountains, particularly in tropical or sub-tropical areas [@Kessler2011a]. 
-Hypotheses to explain such a pattern include pure geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
-Another reason for the greater richness in southern Honshu compared to smaller southern islands is that southern Honshu harbors high elevational variation that may support a wide range of taxa that is not present in the southern islands.
+The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants [@Fukaya2020].
+Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
+Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018a; @Qian2021] to be important drivers of spatial fern diversity.
+A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a]. 
+Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
+Another reason for the greater richness in southern Honshu compared to smaller southern islands is that southern Honshu harbors high elevational variation that may support a wider range of taxa than is present in the southern islands.
 
@@
 Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`B).
-The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* and *Botrychium*.
-The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera such as *Dryopteris* and *Deparia*.
-
-Furthermore, we identified % apomictic taxa as a strong driver of SES of PD (`r figure("modelfit")`G); indeed, *Dryopteris* and *Deparia* are two such genera with high numbers of apomictic taxa. 
+The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`A).
+The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).
+
+Furthermore, we identified % apomictic taxa as a strong driver of SES of PD (`r figure("modelfit")`G).
+Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
 This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns.
-While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD.
+While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`).
 Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca.
 34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands).
-@Tanaka2014 assert that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid; as fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.
+@Tanaka2014 assert that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
+As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.
 
@@
 Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions.
-The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl (`r figure("bioregions")`).
+The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl.
+(`r figure("bioregions")`).
 The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
-Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic Cluster 3 (mostly the Ryukyu islands, wetter) from Cluster 4 (mostly the Ogasawara islands, drier) (`r s_figure("bioregions-env")`).
+Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic Cluster 3 (mostly the Ryukyu islands; wetter) from Cluster 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).
 

### 2021-07-28 ms/manuscript.Rmd
Update MS
conservation discussion

@@
 # Load libraries used for rendering Rmds
-# library(officedown)
-# library(officer)
 library(flextable)
@@
 library(ftExtra)
+library(scales)
 library(patchwork)
@@
 34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands).
-@Tanaka2014 assert that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
+@Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
 As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.
@@
 Although the different facets of biodiversity have different distribution patterns, we find that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric (`r figure("conserv-status")`).
-While it is not clear why this is so---it is unlikely that ferns were emphasized as targets during the establishment of conservation zones---one possibility (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
+One possible reason is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas (A. Ebihara, pers. comm.). 
+Another (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
 
@@
 One such other threat is herbivory by native Japanese deer, *Cervus nippon*. 
-Populations of these deer have been increasing exponentially since the 1970s due to a combination of climate change and a decrease in hunting and natural predators. 
-The deer are capable of decimating the herb layer when populations become dense and have been causing extensive damage even within national parks, which have the highest protection status. 
-Although various efforts to prevent deer herbivory have been deployed such as exclusion fences, much work remains to be done to save vulnerable plant populations.
+Populations of these deer have increased rapidly since the 1970s due to a combination of climate change and a decrease in human hunters and natural predators [@Takatsuki2009]. 
+At dense populations, the deer can prevent forest regeneration and result in plant communities dominated by a limited number of species unpalatable to deer [@Takatsuki2009], although lower population sizes may promote plant diversity as per the intermediate disturbance hypothesis [@Nishizawa2016].
+They have caused extensive damage even within national parks [@Tokita2006], which have the highest protection status. 
+Although various efforts to prevent deer herbivory have been deployed such as culling and exclusion fences [@Takatsuki2009], much work remains to be done to save vulnerable plant populations.
 

### 2021-07-29 ms/manuscript.Rmd
Update MS
Use markdown table for traits, to include refs

@@
 There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research. 
-First, the data include species collected over a wide time period, so they do not represent the current distribution of ferns in Japan. 
+First, the data include species collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
 Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (`r s_figure("abundance")`). 
@@
 Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
+Methods for calculating PD would need to be modified to take into account multiple lineages contributing to a terminal taxon.
 
@@
 However, recent studies in Japan [@Ebihara2013; @Ebihara2019a] and elsewhere [@Nitta2017] have revealed different patterns in the distribution of gametophytes and sporophytes. 
-The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA barcoding analyses to compare patterns of biodiversity between life stages in ferns across large spatial scales.
+The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA sequencing analyses to compare patterns of biodiversity between life stages in ferns across large spatial scales.
 
@@
 
-```{r table-prep}
+| Trait               | Significance                                                | Reference         | 
+|---------------------|-------------------------------------------------------------|-------------------|
+| Indusium (b)        | Protects developing spores                                  | [@Poppinga2015]   |
+| Frond width (c)     | Scales with overall plant size                              | [@Creese2011]     |
+| No. pinna pairs (c) | Less divided fronds have lower rates of evapotranspiration  | [@Kluge2007]      |
+| Stipe length (c)    | Shorter lengths minimize resistance                         | [@WatkinsJr.2010] |
+| Frond margin (q)    | Thermoregulation                                            | [@Nicotra2011]    |
+| Frond shape (q)     | Less divided fronds have lower rates of evapotranspiration  | [@Kluge2007]      |
+| Frond texture (q)   | Thicker leaves have lower rates of evapotranspiration       | [@Givnish1987]    |
+
+**`r table("traits-used")`**.
+Fern traits used in this study.
+Letters in parentheses indicate trait type: b, binary; c, continuous; q, qualitative.
+Qualitative traits were coded as binary for analysis (see Methods).
+
+\newpage
+
+```{r plot-prep}
 # Define nicely formatted terms for printing variable names
@@
 )
-```
-
-```{r traits-used-prep}
-# Read in trait metadata
-trait_signif <- read_csv(here::here("data_raw/trait_functions.csv"))
-
-# Format traits to look nice
-traits_table <- traits_summary %>%
-  left_join(trait_signif, by = "trait") %>%
-  mutate(
-    trait = case_when(
-      trait == "frond_width" ~ "Frond width",
-      trait == "number_pinna_pairs" ~ "No. pinna pairs",
-      trait == "stipe_length" ~ "Stipe length",
-      trait == "margin" ~ "Frond margin",
-      trait == "shape" ~ "Frond shape",
-      trait == "texture" ~ "Frond texture",
-      TRUE ~ str_to_sentence(trait)
-    ),
-    trait_type = case_when(
-      trait_type == "qualitative" ~ glue::glue("Qualitative (coded as {n} binary traits)"),
-      trait_type == "continuous" & str_detect(trait, "pinna pairs") ~ glue::glue("Continuous (integer)"),
-      trait_type == "continuous" ~ glue::glue("Continuous (cm)"),
-      TRUE ~ str_to_sentence(trait_type)
-    )
-  ) %>%
-  select(trait, Significance = significance)
-```
-
-```{r traits-used-pdf, results = "asis", eval = params$doc_type == "pdf"}
-# Print trait table to pdf
-traits_table %>%
-  column_to_rownames("trait") %>%
-  knitr::kable(format = "latex", booktabs = TRUE, escape = TRUE) 
-```
-
-```{r traits-used-doc, results = "asis", eval = params$doc_type == "doc"}
-# Print trait table to doc
-traits_table %>%
-  flextable() %>%
-  set_header_labels(trait = "") %>%
-  width(1, width = 1.25) %>%
-  width(2, width = 4)
-```
-
-**`r table("traits-used")`**. Fern traits used in this study.
-
-\newpage
-
-```{r plot-prep}
-# crop high-res map of Japan to spatial div results area
+
+# Crop high-res map of Japan to spatial div results area
 japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))

### 2021-07-30 ms/manuscript.Rmd
Update MS
Fix some captions

@@
 | Trait               | Significance                                               | Reference       | 
-|---------------------|-------------------------------------------------------------|-------------------|
-| Indusium (b)        | Protects developing spores                                  | [@Poppinga2015]   |
-| Frond width (c)     | Scales with overall plant size                              | [@Creese2011]     |
-| No. pinna pairs (c) | Less divided fronds have lower rates of evapotranspiration  | [@Kluge2007]      |
-| Stipe length (c)    | Shorter lengths minimize resistance                         | [@WatkinsJr.2010] |
-| Frond margin (q)    | Thermoregulation                                            | [@Nicotra2011]    |
-| Frond shape (q)     | Less divided fronds have lower rates of evapotranspiration  | [@Kluge2007]      |
-| Frond texture (q)   | Thicker leaves have lower rates of evapotranspiration       | [@Givnish1987]    |
+|---------------------|------------------------------------------------------------|-----------------|
+| Indusium (b)        | Protects developing spores                                 | @Poppinga2015   |
+| Frond width (c)     | Scales with overall plant size                             | @Creese2011     |
+| No. pinna pairs (c) | Less divided fronds have lower rates of evapotranspiration | @Kluge2007      |
+| Stipe length (c)    | Shorter lengths minimize resistance                        | @WatkinsJr.2010 |
+| Frond margin (q)    | Thermoregulation                                           | @Nicotra2011    |
+| Frond shape (q)     | Less divided fronds have lower rates of evapotranspiration | @Kluge2007      |
+| Frond texture (q)   | Thicker leaves have lower rates of evapotranspiration      | @Givnish1987    |
 
@@
 Letters in parentheses indicate trait type: b, binary; c, continuous; q, qualitative.
-Qualitative traits were coded as binary for analysis (see Methods).
+Qualitative traits were coded as binary for analysis (see Materials and Methods).
 
@@
 For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
-For details, see Methods.
+For details, see Materials and Methods.
 
@@
 Line fit with the focal predictor variable while averaging over other predictors.
-For biodiversity metrics other than richness, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
+For phylogenetic diversity (PD), relative PD (RPD), functional diversity (FD), and relative FD (RFD), the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*P*-value) calculated using a two-tailed test (see Materials and Methods).
 
@@
 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
-For details, see Methods.
+For details, see Materials and Methods.
 

### 2021-07-30 ms/manuscript.Rmd
Update MS
fix map of Japan, place names

@@
 
-**Methods** We used data from >300,000 fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
+**Methods** We used data from >`r nrow(occ_point_data_ferns) %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
 We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode.
@@
 Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`A). 
-A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Amami, Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula). 
+A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula). 
 The area near Yakushima includes cells with both significantly high and low PD. 
@@
 The vast majority of grid-cells with significant endemism are mixed.
-A concentration of paleoendemism was observed in the Ryukyu Islands. 
-Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in central Hokkaido (`r figure("endemism")`). 
-When we ran CANAPE with a reduced dataset including only taxa absolutely endemic to Japan to account for the border effect, there were fewer significant cells, but patterns were otherwise similar to the full dataset (`r s_figure("endemism-restricted")`).
+Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands. 
+Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`). 
+When we ran CANAPE with a reduced dataset including only taxa absolutely endemic to Japan to account for the border effect, there were fewer significant cells, but a similar pattern of significantly endemic cells mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).
 
@@
 Analysis of bioregions grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` clusters (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Okinawa), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A).
-Phylogenetic distances produced similar results, but Okinawa and the Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B). 
+The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Ryukyu Islands, excluding Yakushima), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A).
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B). 
 The other much smaller clusters are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
@@
 main_islands <- c("Hokkaido", "Shikoku", "Kyushu", "Honshu")
-small_islands <- c("Okinawa", "Amamioshima", "Ogasawara", "Yakushima", "Izu Islands")
+small_islands <- c("Okinawa", "Amami", "Ogasawara Islands", "Yakushima", "Izu Islands", "Iriomote", "Miyako")
 other <- c("Tokyo")
@@
   japan_points_raw %>%
-  as_tibble %>%
+  filter(is.na(lat_end)) %>%
+  select(-lat_end, -lon_end) %>%
   mutate(
@@
       query == "Tokyo" ~ "-1.2,0",
-      query == "Shikoku" ~ "1,-.5",
-      query == "Kyushu" ~ "1,-.6",
-      query %in% small_islands ~ "2.5,0",
+      query == "Shikoku" ~ "1.3,-.6",
+      query == "Kyushu" ~ "1.6,-.4",
+      query == "Miyako" ~ "1.5, 0",
+      query == "Iriomote" ~ "1.8, 0",
+      query == "Izu Islands" ~ "2.0, 0",
+      query == "Yakushima" ~ "2.2, 0",
+      query == "Ogasawara Islands" ~ "3.0, 0",
+      query == "Izu Peninsula" ~ "2.0, -0.3",
+      query == "Kii Peninsula" ~ "2.0, -0.5",
+      query == "Ryukyu Islands" ~ "8.0, 1.0",
+      query %in% small_islands ~ "2.0, 0",
       TRUE ~ "0,0"
@@
 
+japan_lines <- japan_points_raw %>%
+  filter(!is.na(lat_end))
+
 # Make map
 japan_map_plot <- ggplot(japan) + 
-  geom_sf() +
+  geom_sf(size = 0.25) +
+  geom_segment(data = japan_lines, aes(x = lon, y = lat, xend = lon_end, yend = lat_end, linetype = query)) +
+  scale_linetype_manual(values = c("Tsugaru Strait" = "solid", "Tokara Strait" = "longdash")) +
   geom_richtext(
@@
   ) +
-  map_theme()
+  map_theme() +
+  theme(legend.position = "none")
+
+ggsave("plot.pdf", japan_map_plot, width = 5, height = 5, units = "in")
 ```
@@
 
-**`r figure("japan-map")`**. Map of Japan showing names of places mentioned in the text. Names of the four main islands in bold.
+**`r figure("japan-map")`**.
+Map of Japan showing names of places mentioned in the text.
+Names of the four main islands in bold.
+The Ryukyu Islands include Yakushima, Amami, Okinawa, Miyako, and Iriomote.
+Solid line shows location of Tsugaru Strait; dashed line shows Tokara Strait.
 

### 2021-08-02 ms/SI.Rmd
Update MS
add bioregions dendrogram, phylo. tree

@@
 library(scales)
+library(ggtree)
 ```
@@
   "comm_ferns",
-  "repro_data"
+  "repro_data",
+  "japan_pterido_tree"
 ))
@@
 
+```{r tree, fig.width = 8, fig.height = 8}
+phy <- japan_pterido_tree
+
+phy$tip.label <- str_replace_all(phy$tip.label, "_", " ")
+
+# Make a dataframe (tibble) with node IDs (integers) and their corresponding bootstrap support values
+# The tibble has two columns: one called "node", the other can be named as we like (here, "bootstrap")
+bs_tibble <- tibble(
+  # hard-code node ID: internal nodes start after tip nodes,
+  # and phy$node.label is in the same order as internal nodes
+  node=1:Nnode(phy) + Ntip(phy), 
+  # Don't print BS < 50%
+  # (or to show all BS, just do `bootstrap = phy$node.label`)
+  bootstrap_raw = phy$node.label,
+  bootstrap = parse_number(bootstrap_raw))
+
+p <- ggtree(phy, layout = "fan", size = 0.25, open.angle = 3) %<+% 
+  bs_tibble +
+  geom_nodepoint(aes(color = bootstrap), size = 0.2) +
+  geom_tiplab(size = 0.6)  +
+  theme(legend.justification = c(1, 0), legend.position = c(1, 0)) +
+  scale_color_scico(palette = "turku", direction = -1, na.value = "grey", name = "BS") +
+  geom_treescale(x = 220, y = 708, width = 100, fontsize = 0)
+
+rotate_tree(p, 3)
+```
+
+**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding hybrids. Points at nodes indicate bootstrap support. Scalebar 100 my.
+
+\clearpage
+
 ```{r abundance, fig.width = 7, fig.height = 5}
@@
 
+```{r dendrogram-a, fig.width = 7, fig.height = 9}
+# Extract dendrogram from phyloregions analysis, convert to phylogeny
+dendro_tax <- attributes(k_taxonomy$df)$meta$hclust.obj %>% as.phylo() %>% ladderize()
+
+# Map taxonomic clusters to grid cells
+clusters_tax <-
+  biodiv_ferns_spatial %>%
+  as_tibble() %>%
+  select(grids, cluster = taxonomic_cluster) %>%
+  # Convert cluster to numeric to use max() later
+  mutate(cluster = levels(cluster)[cluster] %>% as.numeric)
+
+cluster_nodes_tax_tbl <- map_df(1:max(clusters_tax$cluster), ~map_cluster_to_nodes(dendro_tax, clusters_tax, .)) %>%
+  mutate(cluster = as.factor(cluster))
+
+# Use the ggtree::`%<+%` operator to map the bootstrap values onto the ggtree object
+ggtree(dendro_tax) %<+% 
+  cluster_nodes_tax_tbl +
+  aes(color = cluster) +
+  scale_color_manual(
+    values = bioregion_cols, na.value = "grey", name = "Cluster") +
+  theme(legend.position = "right") +
+  labs(subtitle = "**A** Taxonomic") +
+  theme(plot.subtitle = element_markdown())
+```
+
+**`r s_figure("dendrogram")`**.
+Bioregion dendrogram.
+Tips are sites (names not shown); color shows cluster; internal branches between clusters 'NA'.
+**A**, taxonomic bioregions.
+**B**, phylogenetic bioregions.
+\clearpage
+
+```{r dendrogram-b, fig.width = 7, fig.height = 9}
+# Extract dendrogram from phyloregions analysis, convert to phylogeny
+dendro_phy <- attributes(k_phylogeny$df)$meta$hclust.obj %>% as.phylo() %>% ladderize()
+
+# Map taxonomic clusters to grid cells
+clusters_phy <-
+  biodiv_ferns_spatial %>%
+  as_tibble() %>%
+  select(grids, cluster = phylo_cluster) %>%
+  # Convert cluster to numeric to use max() later
+  mutate(cluster = levels(cluster)[cluster] %>% as.numeric)
+
+cluster_nodes_phy_tbl <- map_df(1:max(clusters_phy$cluster), ~map_cluster_to_nodes(dendro_phy, clusters_phy, .)) %>%
+  mutate(cluster = as.factor(cluster))
+
+# Use the ggtree::`%<+%` operator to map the bootstrap values onto the ggtree object
+ggtree(dendro_phy) %<+% 
+  cluster_nodes_phy_tbl +
+  aes(color = cluster) +
+  scale_color_manual(values = bioregion_cols, na.value = "grey", name = "Cluster") +
+  theme(legend.position = "right") +
+  labs(subtitle = "**B** Phylogenetic") +
+  theme(plot.subtitle = element_markdown())
+```
+
+**`r s_figure("dendrogram")`**, continued.
+
+\clearpage
+
 ```{r genus-map, fig.width = 7, fig.height = 9}

### 2021-08-02 ms/manuscript.Rmd
Update MS
add bioregions dendrogram, phylo. tree

@@
 The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent`) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families of native Japanese ferns. 
-The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a]. 
+The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
 Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* (families follow @PteridophytePhylogenyGroupI2016). 
@@
 Analysis of bioregions grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` clusters (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Ryukyu Islands, excluding Yakushima), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B). 
+The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Ryukyu Islands, excluding Yakushima), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B; `r s_figure("dendrogram")`B). 
 The other much smaller clusters are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.

### 2021-08-02 R/functions.R

@@ -2895,3 +2895,30 @@ clean_vars <- function(x) {
     str_replace_all("^richness$", "Richness")
 }
 
+#' Make a tibble matching the nodes of a dendrogram to the cluster they belong to
+#'
+#' @param dendro Dendrogram; list of class "phylo". Dendrogram clustering sites. Part of output of phyloregion::phyloregion()
+#' @param tax_clusters Tibble with two columns, "grids" (site names) and "cluster" (ID of cluster for each site)
+#' @param cluster_select Value indicating the cluster to select for making the tibble
+#'
+#' @return Tibble
+#' 
+map_cluster_to_nodes <- function(dendro, tax_clusters, cluster_select) {
+  
+  # Get vector of tips matching the selected cluster
+  tips <- tax_clusters %>% filter(cluster == cluster_select) %>% pull(grids)
+  
+  if(length(tips) == 1) {
+    # If only one tip, just return that node (tip) number
+    nodes <- which(dendro$tip.label == tips)
+  } else if (length(tips) > 1) {
+    # If multiple tips, get MRCA, then all descendents
+    node_mrca <- ape::getMRCA(dendro, tips)
+    nodes <- phytools::getDescendants(dendro, node_mrca)
+  } else {
+    stop("Can't find that cluster")
+  }
+  
+  # Return a tibble mapping nodes to the cluster
+  tibble(node = nodes, cluster = cluster_select)
+}

### 2021-08-03 ms/SI.Rmd
Adjust line width of dendrogram
so all sites are distinct

@@
 # Use the ggtree::`%<+%` operator to map the bootstrap values onto the ggtree object
-ggtree(dendro_tax) %<+% 
+ggtree(dendro_tax, size = 0.15) %<+% 
   cluster_nodes_tax_tbl +
@@
 # Use the ggtree::`%<+%` operator to map the bootstrap values onto the ggtree object
-ggtree(dendro_phy) %<+% 
+ggtree(dendro_phy, size = 0.15) %<+% 
   cluster_nodes_phy_tbl +

### 2021-08-06 ms/manuscript.Rmd
Add tree stats back in to MS
Update k-val for most recent phylogeny

@@
   "japan_points_raw",
-  # "plastome_calibration_dates",
-  # FIXME: remove until phylo analysis is working "plastome_tree",
+  "plastome_calibration_dates",
+  "plastome_tree",
   "green_list",
@@
 Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
-To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (
-<!-- FIXME: remove until phylo analysis is working  `r #ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` -->
-OTUs total). 
-We conducted phylogenetic analysis with IQTREE [@Nguyen2015] and molecular dating with treePL [@Smith2012] using `r #nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a. 
+To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total).
+We conducted phylogenetic analysis with IQTREE [@Nguyen2015] and molecular dating with treePL [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a. 
 We then trimmed the tree to Japanese taxa only and used this for all further analyses.
@@
 # Verify that max K < 0.0.6
-assert_that(phy_sig_results %>% pull(kval) %>% max < 0.06)
+assert_that(phy_sig_results %>% pull(kval) %>% max < 0.17)
 
@@
 
-All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.06, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
+All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.17, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
 `r n_negative_d` out of `r nrow(binary_sig_results)` of continuous traits (coded as binary) showed as much or more phylogenetic signal as expected under BM (*D* =< 0); others had *D* > 0 (`r s_table("phy-sig-binary")`).

### 2021-08-18 ms/SI.Rmd
Add non-ultrametric model

@@
   "repro_data",
-  "japan_pterido_tree"
+  "japan_pterido_tree",
+  "model_fits_not_ult",
+  "biodiv_ferns_cent"
 ))
@@
   set_names(1:8)
+
+# Specify subplot order
+model_subplot_order <- tribble(
+  ~resp_var, ~indep_var, ~order,
+  "richness", "temp", 1,
+  "pd_obs_z", "temp", 2,
+  "rpd_obs_z", "temp", 3,
+  "fd_obs_z", "temp", 4,
+  "rfd_obs_z", "temp", 5,
+  "pd_obs_z", "percent_apo", 6,
+  "rpd_obs_z", "percent_apo", 7
+)
 ```
@@
 
+```{r modelfit-not-ult, fig.width = 7, fig.height = 3}
+# Extract dataset used for model, including significance columns
+biodiv_ferns_used_for_model <- biodiv_ferns_spatial %>%
+  # drop geometry (only need centroids for modeling)
+  st_set_geometry(NULL) %>%
+  as_tibble() %>%
+  # filter to only sites used in model
+  inner_join(
+    biodiv_ferns_cent %>% as_tibble() %>% select(grids),
+    by = "grids"
+  )
+
+# Generate final plot
+modelfit_plot_list <- 
+  model_fits_not_ult %>%
+  mutate(
+    signif_var = str_replace_all(resp_var, "_obs_z", "_signif"),
+    resp_var_print = case_when(
+      resp_var == "richness" ~ "Richness (no. spp.)",
+      TRUE ~ str_to_upper(resp_var) %>% word(1, 1, "_") %>% paste("SES of", .)
+    ),
+    model_data = list(biodiv_ferns_used_for_model)) %>%
+  left_join(model_subplot_order, by = c("resp_var", "indep_var")) %>%
+  arrange(order) %>%
+  pmap(plot_fits)
+
+# Need one subplot with a legend
+modelfit_plot_list[[2]] <- modelfit_plot_list[[2]] +
+  theme(legend.position = "right") +
+  labs(color = "Significance") +
+  # Increase size of points in legend
+  guides(colour = guide_legend(override.aes = list(size = 1.5)))
+
+wrap_plots(modelfit_plot_list) +
+  guide_area() +
+  plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = "A") 
+```
+
+**`r figure("modelfit-not-ult")`**.
+Relationship between (**A**) phylogenetic diversity (PD) and (**B**) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
+Ribbon shows 95% confidence interval of model.
+Line fit with the focal predictor variable while averaging over other predictors.
+For RPD, the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*P*-value) calculated using a two-tailed test (see Materials and Methods).
+
+\clearpage
+
 ```{r bioregions-env, fig.width = 7, fig.height = 5}

### 2021-08-18 ms/manuscript.Rmd
Add non-ultrametric model

@@
 
-Furthermore, we identified % apomictic taxa as a strong driver of SES of PD (`r figure("modelfit")`G).
+Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`G).
 Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
 This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns.
+We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), consistent with an effect on genetic diversity (`r s_figure("modelfit-not-ult")`B).
 While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`).
@@
   )
-
-# Define function for plotting model fit and actual data
-plot_fits <- function(resp_var, resp_var_print, indep_var, signif_var, fit, model_data, ...) {
-
-  # Setup randomization test significance colors
-  signif_cols <-
-    c(
-      "> 0.99" = "#e31a1c", 
-      "> 0.975" = "#fb9a99",
-      "< 0.01" = "#1f78b4", 
-      "< 0.025" = "#a6cee3",
-      "not significant" = "#ffffcc"
-    )
-
-  # Basic plot setup
-  p <- ggplot(model_data, aes(x = .data[[indep_var]], y = .data[[resp_var]])) +
-    labs(y = resp_var_print) +
-    theme_gray(base_size = 9) +
-    theme(panel.grid.minor = element_blank())
-
-  # Add points, color by significance if not plotting richness
-  if (resp_var != "richness") {
-    p <- p + 
-      geom_point(aes(color = .data[[signif_var]]), size = 0.5) +
-      scale_color_manual(values = signif_cols)
-  } else {
-    p <- p + 
-      geom_point(size = 0.5, color = "grey20")
-  }
-
-  # Add line and ribbon of model fit
-  p <- p +
-    geom_line(data = fit, alpha = 0.35, color = "blue") +
-    geom_ribbon(
-      data = fit,
-      aes(ymin = low, ymax = up), 
-      alpha = 0.15
-    ) +
-    theme(legend.position = "none")
-
-  # Format x-axis ticks and labels
-  if (indep_var == "temp") {
-    p <- p +
-      scale_x_continuous(labels = function(x) x*0.1) +
-      labs(x = "Temperature (°C)")
-  } else if (indep_var == "percent_apo") {
-    p <- p +
-      scale_x_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
-      labs(x = "% apomictic taxa")
-  }
-
-  p
-
-}
 

### 2021-08-18 R/functions.R

@@ -2848,3 +2848,57 @@ map_cluster_to_nodes <- function(dendro, tax_clusters, cluster_select) {
   # Return a tibble mapping nodes to the cluster
   tibble(node = nodes, cluster = cluster_select)
 }
+
+# Define function for plotting model fit and actual data
+plot_fits <- function(resp_var, resp_var_print, indep_var, signif_var, fit, model_data, ...) {
+  
+  # Setup randomization test significance colors
+  signif_cols <-
+    c(
+      "> 0.99" = "#e31a1c", 
+      "> 0.975" = "#fb9a99",
+      "< 0.01" = "#1f78b4", 
+      "< 0.025" = "#a6cee3",
+      "not significant" = "#ffffcc"
+    )
+  
+  # Basic plot setup
+  p <- ggplot(model_data, aes(x = .data[[indep_var]], y = .data[[resp_var]])) +
+    labs(y = resp_var_print) +
+    theme_gray(base_size = 9) +
+    theme(panel.grid.minor = element_blank())
+  
+  # Add points, color by significance if not plotting richness
+  if (resp_var != "richness") {
+    p <- p + 
+      geom_point(aes(color = .data[[signif_var]]), size = 0.5) +
+      scale_color_manual(values = signif_cols)
+  } else {
+    p <- p + 
+      geom_point(size = 0.5, color = "grey20")
+  }
+  
+  # Add line and ribbon of model fit
+  p <- p +
+    geom_line(data = fit, alpha = 0.35, color = "blue") +
+    geom_ribbon(
+      data = fit,
+      aes(ymin = low, ymax = up), 
+      alpha = 0.15
+    ) +
+    theme(legend.position = "none")
+  
+  # Format x-axis ticks and labels
+  if (indep_var == "temp") {
+    p <- p +
+      scale_x_continuous(labels = function(x) x*0.1) +
+      labs(x = "Temperature (°C)")
+  } else if (indep_var == "percent_apo") {
+    p <- p +
+      scale_x_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
+      labs(x = "% apomictic taxa")
+  }
+  
+  p
+  
+}

### 2021-08-18 _targets.R

@@ -488,13 +488,21 @@ tar_plan(
   ## Build spatial models ----
   
   # Prepare datasets for looping
+  # - ultrametric tree (full  analysis)
   data_for_spamm = prepare_data_for_spamm(
     resp_var_env = resp_vars_env,
     resp_var_repro = resp_vars_repro,
     biodiv_ferns_cent = biodiv_ferns_cent
   ),
+  # - non ultrametric tree (% apo only)
+  data_for_spamm_not_ult = prepare_data_for_spamm(
+    resp_var_env = resp_vars_env,
+    resp_var_repro = resp_vars_repro,
+    biodiv_ferns_cent = biodiv_ferns_not_ult
+  ) %>% filter(str_detect(formula, "percent_apo")),
   
   # Loop across each formula and build a spatial model
+  # - ultrametric tree (full  analysis)
   tar_target(
     spatial_models,
     run_spamm(
@@ -504,10 +512,20 @@ tar_plan(
     ),
     pattern = map(data_for_spamm)
   ),
+  # - non ultrametric tree (% apo only)
+  tar_target(
+    spatial_models_not_ult,
+    run_spamm(
+      formula = data_for_spamm_not_ult$formula[[1]], 
+      data = data_for_spamm_not_ult$data[[1]], 
+      resp_var = data_for_spamm_not_ult$resp_var[[1]]
+    ),
+    pattern = map(data_for_spamm_not_ult)
+  ),
   
   ## Conduct LRTs ----
   
-  # Prepare data for looping
+  # Prepare data for looping (full analysis only)
   data_for_lrt = prepare_data_for_lrt(spatial_models, biodiv_ferns_cent),
   
   # Conduct LRTs in loop
@@ -535,10 +553,17 @@ tar_plan(
   ## Predict model fits ---
 
   # Extract model fits in loop: use temperature for environmental models, and percent_apo for repro models
+  # - ultrametric tree (full  analysis)
   tar_target(
     model_fits,
     predict_fit(spatial_models),
     pattern = map(spatial_models)
+  ),
+  # - non ultrametric tree (% apo only)
+  tar_target(
+    model_fits_not_ult,
+    predict_fit(spatial_models_not_ult),
+    pattern = map(spatial_models_not_ult)
   ),
   
   # Conservation analysis ----

### 2021-08-19 ms/SI.Rmd
Update MS
- Make figure part letters bold, small-caps, in parenthesis per J Biogeo style
- Add dependency on files used to build MS

@@
 ---
-title: "Exploring dimensions of biodiversity in Japanese ferns"
-date: '`r format(Sys.time(), "%d %B, %Y")`'
+title: "Appendix S1: Supporting Information"
 mainfont: Roboto-Regular.ttf
@@
 \setlength{\parindent}{0ex}
-
-# Supplemental Information
 
@@
 a + b + c +
-  plot_annotation(tag_levels = "A") &
+  plot_annotation(tag_levels = "a",
+    tag_prefix = "(",
+    tag_suffix = ")") &
     theme(plot.tag = element_text(face = "bold"))
@@
 Effect of grain size on sampling redundancy.
-**A**, Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
-**B**, Percent of grid cells with redundancy = 0 by grain size.
-**C**, Distribution of redundancy values by grain size.
+(a), Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
+(b), Percent of grid cells with redundancy = 0 by grain size.
+(c), Distribution of redundancy values by grain size.
 There is a sudden improvement in redundancy values from 0.1 to 0.2 degree grid-cells, but less improvement as grain size is increased beyond 0.2 degrees.
@@
       geom="richtext", y = Inf, x = -Inf, 
-      label = glue("<span style='color:black'>**{part}** {label}</span>"), 
+      label = glue("<span style='color:black'>**({part})** {label}</span>"), 
       hjust = "inward", vjust = "inward", color = "white") + 
@@
 
-a <- plot_env("A", "Area", "area", "km^2")
-b <- plot_env("B", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
-c <- plot_env("C", "Precipitation", "precip", "mm")
-d <- plot_env("D", "Precip. season.", "precip_season" , "")
-e <- plot_env("E", "Temperature", "temp", "°C", function(x) x/10)
-f <- plot_env("F", "Temp. season.", "temp_season")
+a <- plot_env("a", "Area", "area", "km^2")
+b <- plot_env("b", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
+c <- plot_env("c", "Precipitation", "precip", "mm")
+d <- plot_env("d", "Precip. season.", "precip_season" , "")
+e <- plot_env("e", "Temperature", "temp", "°C", function(x) x/10)
+f <- plot_env("f", "Temp. season.", "temp_season")
 
@@
 Reproductive mode and environmental data in 0.2 degree grid cells across Japan.
-**A**, rolling mean area (km^2^) in 1° latitudinal bands.
-**B**, % apomictic taxa.
-**C**, annual precipitation **D**, percipitation seasonality (coefficient of variation).
-**E**, mean annual temperature.
-**F**, temperature seasonality (standard deviation × 100).
+(a), rolling mean area (km^2^) in 1° latitudinal bands.
+(b), % apomictic taxa.
+(c), annual precipitation.
+(d), percipitation seasonality (coefficient of variation).
+(e), mean annual temperature.
+(f), temperature seasonality (standard deviation × 100).
 
@@
   labs(fill = "No. specimens") +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A**</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
   map_theme()
@@
   labs(fill = "Redundancy") +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B**</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
   map_theme() +
@@
 
-**`r s_figure("abundance")`**. Observed number of specimens (**A**) and sampling redundancy (**B**) per 0.2° grid cell in the ferns of Japan.
+**`r s_figure("abundance")`**. Observed number of specimens (a) and sampling redundancy (b) per 0.2° grid cell in the ferns of Japan.
 
@@
   plot_layout(ncol = 2, nrow = 2) +
-  plot_annotation(tag_levels = "A") &
+  plot_annotation(tag_levels = "a",
+    tag_prefix = "(",
+    tag_suffix = ")") &
     theme(plot.tag = element_text(face = "bold"))
@@
 Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.
-**A**, Phylogenetic diversity (PD).
-**B**, Relative phylogenetic diversity (RPD).
-**C**, Functional diversity (FD).
-**D**, Relative functional diversity (RFD).
+(a), Phylogenetic diversity (PD).
+(b), Relative phylogenetic diversity (RPD).
+(c), Functional diversity (FD).
+(d), Relative functional diversity (RFD).
 
@@
   plot_layout(ncol = 2) +
-  plot_annotation(tag_levels = "A") &
+  plot_annotation(tag_levels = "a",
+    tag_prefix = "(",
+    tag_suffix = ")") &
     theme(plot.tag = element_text(face = "bold"))
@@
 Selection of *k* for bioregions.
-**A**, taxonomic bioregions.
-**B**, phylogenetic bioregions.
+(a), taxonomic bioregions.
+(b), phylogenetic bioregions.
 Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal; see Methods).
@@
   theme(legend.position = "right") +
-  labs(subtitle = "**A** Taxonomic") +
+  labs(subtitle = "**(a)** Taxonomic") +
   theme(plot.subtitle = element_markdown())
@@
 Tips are sites (names not shown); color shows cluster; internal branches between clusters 'NA'.
-**A**, taxonomic bioregions.
-**B**, phylogenetic bioregions.
+(a), taxonomic bioregions.
+(b), phylogenetic bioregions.
 \clearpage
@@
   theme(legend.position = "right") +
-  labs(subtitle = "**B** Phylogenetic") +
+  labs(subtitle = "**(b)** Phylogenetic") +
   theme(plot.subtitle = element_markdown())
@@
       geom="richtext", y = Inf, x = -Inf, 
-      label = glue("<span style='color:black'>**{part}** *{genus_select}*</span>"), 
+      label = glue("<span style='color:black'>**({part})** *{genus_select}*</span>"), 
       hjust = "inward", vjust = "inward", color = "white") + 
@@
 
-a <- plot_genus("A", "Equisetum")
-b <- plot_genus("B", "Deparia")
-c <- plot_genus("C", "Cyrtomium")
-d <- plot_genus("D", "Dryopteris")
-e <- plot_genus("E", "Pteris")
-f <- plot_genus("F", "Diplazium")
+a <- plot_genus("a", "Equisetum")
+b <- plot_genus("b", "Deparia")
+c <- plot_genus("c", "Cyrtomium")
+d <- plot_genus("d", "Dryopteris")
+e <- plot_genus("e", "Pteris")
+f <- plot_genus("f", "Diplazium")
 
@@
 Map of species richness for selected genera of ferns of Japan.
-**A**, *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
-**B**, *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
-**C**--**F** show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.
+(a), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
+(b), *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
+(c)--(f) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.
 
@@
   plot_layout(guides = "collect") +
-  plot_annotation(tag_levels = "A") 
+  plot_annotation(tag_levels = "a",
+                  tag_prefix = "(",
+                  tag_suffix = ")") &
+  theme(plot.tag = element_text(face = "bold"))
 ```
 
-**`r figure("modelfit-not-ult")`**.
-Relationship between (**A**) phylogenetic diversity (PD) and (**B**) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
+**`r s_figure("modelfit-not-ult")`**.
+Relationship between (a) phylogenetic diversity (PD) and (b) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
 Ribbon shows 95% confidence interval of model.
@@
 
-a + b + plot_annotation(tag_levels = "A") &
+a + b + plot_annotation(tag_levels = "a",
+    tag_prefix = "(",
+    tag_suffix = ")") &
     theme(plot.tag = element_text(face = "bold"))
@@
 **`r s_figure("bioregions-env")`**.
-Scatterplots of grid cell membership in (**A**) taxonomic or (**B**) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
+Scatterplots of grid cell membership in (a) taxonomic or (b) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
 Points shown with partial transparency; darker areas indicate overlapping points.

### 2021-08-19 ms/manuscript.Rmd
Update MS
- Make figure part letters bold, small-caps, in parenthesis per J Biogeo style
- Add dependency on files used to build MS

@@
   "model_fits",
-  "biodiv_ferns_cent"
+  "biodiv_ferns_cent",
+  "refs_yaml",
+  "refs_other_yaml",
+  "template_file",
+  "csl_file"
 ))
@@
 
-Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1,3,4,5,6^, and Atsushi Ebihara^7^
+Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^
 
@@
 
-^3^Department of Computational Biology and Medical Sciences, Graduate School of Frontier Sciences, The University of Tokyo, Kashiwa, Japan
-
-^4^Atmosphere and Ocean Research Institute, The University of Tokyo, Kashiwa, Japan
-
-^5^Institute for Quantitative Biosciences, The University of Tokyo, Tokyo, Japan
-
-^6^Collaborative Research Institute for Innovative Microbiology, The University of Tokyo, Tokyo, Japan
-
-^7^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
+^3^Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan
+
+^4^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
 
@@
 ## Acknowledgements
+
+FIXME: add acknowledgements (not including reviewers)
 
@@
 
-**Key words:** biodiversity, biogeography, phylogenetic diversity.
+<!-- may include title words -->
+**Key words:** apomixis, biodiversity, biogeography, ferns, Japan, phylogenetic diversity.
 
@@
 
-Our data include all ferns known to occur within the political boundaries of Japan (20°N to 45°N, 122°E to 153°E). 
+Our data include all ferns known to occur within the boundaries of Japan (20°N to 45°N, 122°E to 153°E). 
 Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
@@
 1.01 (http://www.rdplants.org/gl/).
-Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")`). 
+Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
 The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
@@
 We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown. 
-We calculated the percentage of apomictic species in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).
+We calculated the percentage of apomictic species in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`b).
 
@@
 To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total).
-We conducted phylogenetic analysis with IQTREE [@Nguyen2015] and molecular dating with treePL [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a. 
+We conducted phylogenetic analysis with `r software("IQTREE")` [@Nguyen2015] and molecular dating with `r software("treePL")` [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a.
 We then trimmed the tree to Japanese taxa only and used this for all further analyses.
+Tree visualization was done with the `r pack("ggtree")` R package [@Yu2017].
 
@@
 For taxonomic richness, cells ranked in the top >5% were considered highly diverse.
-
-### Data accessibility
-
-Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. 
-A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
-Data files, including the DNA alignment, phylogenetic tree, and community data matrix, are openly available from Figshare at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
 
@@
 The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`). 
-The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A). 
-Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B).
+The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`a). 
+Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`b).
 The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid-cells.
@@
 assert_that(phy_sig_results %>% pull(lambda) %>% min > 0.95)
-# Verify that max K < 0.0.6
-assert_that(phy_sig_results %>% pull(kval) %>% max < 0.17)
+# Verify that max K < 0.6
+assert_that(phy_sig_results %>% pull(kval) %>% max < 0.10)
 
@@
 
-All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.17, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
+All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.1, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
 `r n_negative_d` out of `r nrow(binary_sig_results)` of continuous traits (coded as binary) showed as much or more phylogenetic signal as expected under BM (*D* =< 0); others had *D* > 0 (`r s_table("phy-sig-binary")`).
@@
 
-Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`A).
-Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`A, C).
-Observed values of RPD and RFD were not linearly related to richness, however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).
+Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`a).
+Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`b, c), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`a, c).
+Observed values of RPD and RFD were not linearly related to richness, however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`b, d).
 
@@
 
-Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`A). 
+Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`a). 
 A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula). 
 The area near Yakushima includes cells with both significantly high and low PD. 
-Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu (`r figure("rand-diversity")`B). 
+Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
 Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido. 
-Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD (`r figure("rand-diversity")`C, D).
+Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD (`r figure("rand-diversity")`c, d).
 
@@
 All models were significant relative to null models without any fixed effects `r s_table("lrt-null")`. 
-The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
-Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A). 
-Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E). 
-SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F). 
+The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`a--e), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`f--g); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
+Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`a). 
+Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`b--e). 
+SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`g); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`f). 
 Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (`r s_table("aic")`).
@@
 Analysis of bioregions grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` clusters (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Ryukyu Islands, excluding Yakushima), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single cluster (`r figure("bioregions")`B; `r s_figure("dendrogram")`B). 
+The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Ryukyu Islands, excluding Yakushima), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a).
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single cluster (`r figure("bioregions")`b; `r s_figure("dendrogram")`b). 
 The other much smaller clusters are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants [@Fukaya2020].
-Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
+The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants [@Fukaya2020].
+Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`a).
 Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018a; @Qian2021] to be important drivers of spatial fern diversity.
@@
 
-The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C). 
+The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`a, c). 
 This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
@@
 
-Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`B).
-The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`A).
-The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).
-
-Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`G).
+Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`)b.
+The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a).
+The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).
+
+Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`g).
 Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
 This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns.
-We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), consistent with an effect on genetic diversity (`r s_figure("modelfit-not-ult")`B).
+We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), consistent with an effect on genetic diversity (`r s_figure("modelfit-not-ult")`b).
 While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`).
@@
 The extreme evolutionary distinctiveness of southern islands is both a major feature of the Japanese fern flora and a challenge for analysis. 
-The tropical lineages occurring in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on Honshu (`r figure("rand-diversity")`A). 
+The tropical lineages occurring in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on Honshu (`r figure("rand-diversity")`a). 
 If a more restricted null model was used that did not consider all taxa capable of dispersing anywhere within the study area, different results would likely be obtained for the randomization test [@Mishler2020].
@@
 ## Author Contribution
-
-## References
-<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
-<div id="refs"></div>
 
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "No. taxa") +
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "Relative\nbranch\nlength") +
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Relative\nbranch\nlength") +
@@
 **`r figure("raw-diversity")`**.
-Raw taxonomic (**A**), phylogenetic (**B**), and functional (**C**) diversity of the ferns of Japan.
-For **B** and **C**, raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+Raw taxonomic (a), phylogenetic (b), and functional (c) diversity of the ferns of Japan.
+For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
 
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   scale_fill_manual(values = signif_cols) +
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**D** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(d)** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
 Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.
-**A**, Phylogenetic diversity (PD).
-**B**, Relative phylogenetic diveristy (RPD).
-**C**, Functional diveristy (FD).
-**D**, relative functional diveristy (RFD).
+(a), Phylogenetic diversity (PD).
+(b), Relative phylogenetic diveristy (RPD).
+(c), Functional diveristy (FD).
+(d), relative functional diveristy (RFD).
 For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
@@
   plot_layout(guides = "collect") +
-  plot_annotation(tag_levels = "A")
+  plot_annotation(
+    tag_levels = "a",
+    tag_prefix = "(",
+    tag_suffix = ")") &
+  theme(plot.tag = element_text(face = "bold"))
 ```
@@
 Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models.
-A--E, environmental models (models including temperature); F and G, reproductive models (models including % apomictic taxa).
+(a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa).
 Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-value indicates probability of no difference between the full and null model.
@@
   plot_layout(guides = "collect") +
-  plot_annotation(tag_levels = "A")
+  plot_annotation(
+    tag_levels = "a",
+    tag_prefix = "(",
+    tag_suffix = ")")  &
+  theme(plot.tag = element_text(face = "bold"))
 ```
@@
 Relationship between biodiversity metrics and selected predictor variables.
-A--E, environmental models (models including temperature); F and G, reproductive models (models including % apomictic taxa).
+(a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa).
 Ribbon shows 95% confidence interval of model.
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Cluster") +
@@
   geom_sf(data = biodiv_ferns_spatial, aes(fill = phylo_cluster), color = "transparent", alpha = 0.8) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Cluster") +
@@
 Bioregions of the ferns of Japan.
-**A**, taxonomic bioregions.
-**B**, phylogenetic bioregions.
-Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances into *k* = 8 or 7 groups, respectively, using the R package `r pack("phyloregion")`.
+(a), taxonomic bioregions.
+(b), phylogenetic bioregions.
+Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances into *k* = 8 or 7 groups, respectively.
 
@@
   labs(
-    y = "PE p-score",
+    y = "PE *p*-score",
     x = "Cluster",
@@
   theme(
+    axis.title.y = element_markdown(),
     panel.grid.minor.x = element_blank(),
@@
   plot_annotation(
-    tag_levels = 'A'
+    tag_levels = "a",
+    tag_prefix = "(",
+    tag_suffix = ")"
   ) &
-  theme(legend.position = "none")
+  theme(
+    legend.position = "none",
+    plot.tag = element_text(face = "bold"))
 ```
@@
 Phylogenetic and morphological diversity of the ferns of Japan by bioregion.
-**A**, Standard effect size (SES) of phylogenetic diversity (PD).
-**B**, SES of relative PD (RPD).
-**C**, SES of functional diversity (FD).
-**D**, SES of relative FD (RFD).
-**E**, p-score of observed phylogenetic endemism (PE) relative to 999 random communities.
+(a), Standard effect size (SES) of phylogenetic diversity (PD).
+(b), SES of relative PD (RPD).
+(c), SES of functional diversity (FD).
+(d), SES of relative FD (RFD).
+(e), *p*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
 Only top four clusters with the most sites shown.
@@
 Vertical lines indicate total percent area protected across Japan regardless of degree of biodiversity; dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).
+
+## Data availability statement
+
+Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. 
+A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
+Data files, including the DNA alignment, phylogenetic tree, and community data matrix, are openly available from Dryad at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
+
+## References
+<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
+<div id="refs"></div>
+
+## Biosketch
+
+Joel H. Nitta's research focuses on patterns and processes of diversification in ferns. For more information, see https://joelnitta.com.
+
+Author contributions: J.H.N., B.M., and A.E.
+conceived the ideas; A.E.
+provided the data; J.H.N.
+analyzed the data; I.W.
+provided computational resources; J.H.N.
+wrote the manuscript with input from all co-authors.

### 2021-08-19 _targets.R

@@ -669,8 +669,14 @@ tar_plan(
   # Calculate percent protection for grid cells with significantly high biodiversity
   signif_cells_protected_area = calculate_protected_area(biodiv_ferns_spatial, protected_areas, japan_shp),
   
-  # Render manuscript
-  # - doc
+  # Render manuscript ----
+  # Track ms files
+  tar_file(refs_yaml, here::here("ms/references.yaml")),
+  tar_file(refs_other_yaml, here::here("ms/references_other.yaml")),
+  tar_file(template_file, here::here("ms/template.docx")),
+  tar_file(csl_file, here::here("ms/journal-of-biogeography.csl")),
+  
+  # MS, docx format
   tar_render(
     ms_doc,
     knit_root_dir = here::here(),
@@ -679,7 +685,7 @@ tar_plan(
     output_file = here::here("results/manuscript.docx"),
     params = list(doc_type = "doc")
   ),
-  # - pdf
+  # MS, pdf format
   tar_render(
     ms_pdf,
     knit_root_dir = here::here(),
@@ -688,7 +694,7 @@ tar_plan(
     output_file = here::here("results/manuscript.pdf"),
     params = list(doc_type = "pdf")
   ),
-  # - SI pdf
+  # SI
   tar_render(
     si_pdf,
     knit_root_dir = here::here(),
@@ -697,7 +703,7 @@ tar_plan(
     output_file = here::here("results/supp_info.pdf"),
     params = list(doc_type = "pdf")
   ),
-  # - SI appendix on data exploration for models
+  # SI appendix on data exploration for models
   tar_render(
     si_data_exploration,
     knit_root_dir = here::here(),

### 2021-08-19 ms/manuscript.Rmd
Use APA (6th ed) for citations
as per J. Biogeog. guidance

@@
   bookdown::word_document2:
-    pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
+    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
     toc: no
@@
     latex_engine: xelatex
-    pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
+    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
 params:

### 2021-08-19 ms/manuscript.Rmd
Update MS
Fix description of some results

@@
 Temperature and apomixis were identified as likely drivers of patterns of taxonomic and phylogenetic diversity.
-Conservation status (% protected area) is higher than the overall rate for grid-cells with significantly high biodiversity.
+Conservation status (% protected area) is generally higher than the overall rate for grid-cells with significantly high biodiversity.
 
@@
 However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale. 
-The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled. 
+The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b]. 
 One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
@@
 
-We used traits originally compiled for identification of ferns and lycophytes of Japan following Ebihara [-@Ebihara2016b; -@Ebihara2017]. 
+We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017]. 
 Continuous traits were measured on 10 randomly selected specimens per species. 
@@
 This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
-The R package `r pack("sf")` was used for GIS analyses.
+The R package `r pack("sf")` was used for GIS analyses [@Pebesma2018].
 
@@
 
-All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package [@Nitta2021], which calculates raw biodiversity metrics using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].
+All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package [@Nitta2021a], which calculates raw biodiversity metrics using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].
 
@@
 Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`a). 
-A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Izu Peninsula, Kii Peninsula). 
-The area near Yakushima includes cells with both significantly high and low PD. 
+A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
 Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
 Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido. 
-Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity, but with more significant cells for FD than PD and fewer significant cells for RFD than RPD (`r figure("rand-diversity")`c, d).
+Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`c, d).
 
@@
 
-All models were significant relative to null models without any fixed effects `r s_table("lrt-null")`. 
+All models were significant relative to null models without any fixed effects (`r s_table("lrt-null")`). 
 The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`a--e), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`f--g); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
@@
 SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`g); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`f). 
-Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (`r s_table("aic")`).
+Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD and SES of PD (`r s_table("aic")`).
 
@@
 Areas with high PE were the best protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total). 
-All cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total).
+Cells with significantly high biodiversity generally had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total); only FD was lower than the overall rate for highly protected areas.
 
@@
 
-Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`)b.
+Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`b).
 The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a).
@@
 The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA sequencing analyses to compare patterns of biodiversity between life stages in ferns across large spatial scales.
-
-## Author Contribution
 
@@
   theme(legend.position = "none")
-
-ggsave("plot.pdf", japan_map_plot, width = 5, height = 5, units = "in")
 ```
@@
 **`r figure("div-by-tax-region")`**.
-Phylogenetic and morphological diversity of the ferns of Japan by bioregion.
+Phylogenetic and morphological diversity of the ferns of Japan by taxonomic bioregion.
 (a), Standard effect size (SES) of phylogenetic diversity (PD).
@@
 Percent of land area with protected status for grid cells with significantly high biodiversity.
-Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), relative PD (RPD), relative FD (RFD), and phylogenetic endemism (PE).
+Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
 Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Methods).

### 2021-08-20 ms/SI.Rmd
Fix CSL file in SI

@@
     latex_engine: xelatex
-    pandoc_args: [ "--csl", "journal-of-biogeography.csl"]
+    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
 header-includes: 

### 2021-08-20 R/process_refs.R
Update reference processing
Combine other references and main_library before filtering

@@ -7,37 +7,88 @@ library(tidyverse)
 library(yaml)
 library(assertr)
 
-# Set path to RMD file
-rmd_file <- "ms/manuscript.Rmd"
+#' Filter a list of references in YAML format to those occurring in an Rmd file
+#'
+#' @param rmd_file Character vector; Path to Rmd file(s)
+#' @param yaml_in String or list; if string, the path to the YAML file to filter.
+#' If list, should be result of reading in a YAML file with yaml::read_yaml()
+#' @param yaml_out Path to write filtered YAML reference file
+#'
+#' @return NULL; externally, the filtered YAML will be written to `yaml_out`
+#' 
+filter_refs_yaml <- function(rmd_file, yaml_in = "ms/main_library.yaml", yaml_out = "ms/references.yaml") {
+  
+  # Helper function to extract citations (formatted like `[@key]`) from an Rmd file
+  extract_citations <- function(rmd_file) {
+      read_lines(rmd_file) %>%
+      stringr::str_split(" |;") %>% 
+      unlist %>% 
+      magrittr::extract(., stringr::str_detect(., "@")) %>% 
+      stringr::str_remove_all("\\[|\\]|\\)|\\(|\\.$|,|\\{|\\}") %>% 
+      magrittr::extract(., stringr::str_detect(., "^@|^-@")) %>% 
+      stringr::str_remove_all("^@|^-@") %>% 
+      unique %>% 
+      sort %>%
+      tibble(key = .)
+  }
+  
+  # Parse RMD file and extract citation keys
+  citations <- map_df(rmd_file, extract_citations)
+    
+  # Read in YAML including all references exported from Zotero
+  if(inherits(yaml_in, "character")) {
+    ref_yaml <- yaml::read_yaml(yaml_in)
+  } else if(inherits(yaml_in, "list")) {
+    ref_yaml <- yaml_in
+  } else {
+    stop("`yaml_in` must be a path to a YAML file (string) or a list read in with yaml::read_yaml()")
+  }
+  
+  # Extract all citation keys from full YAML
+  cite_keys_all <- map_chr(ref_yaml$references, "id") %>%
+    tibble(
+      key = .,
+      order = 1:length(.)
+    )
+  
+  # Check that all keys in the yaml are present in the input YAML
+  missing <- citations %>% anti_join(cite_keys_all, by = "key")
+  
+  assertthat::assert_that(
[…]

### 2021-08-20 ms/SI.Rmd
Update reference processing
Combine other references and main_library before filtering
Add references to SI

@@
   - `r here::here("ms/references.yaml")`
-  - `r here::here("ms/references_other.yaml")`
 ---
@@
 **`r s_figure("sampling-curve")`**.
-Species collection curve for the ferns of Japan fit with the `r pack("iNEXT")` package.
+Species collection curve for the ferns of Japan fit with the `r pack("iNEXT")` package [@Hsieh2016].
 Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line).
@@
 Points shown with partial transparency; darker areas indicate overlapping points.
+
+## References

### 2021-08-20 ms/manuscript.Rmd
Update MS
according to comments from WI
Update reference processing
Combine other references and main_library before filtering
Count plastome sampling

@@
   "template_file",
-  "csl_file"
+  "csl_file",
+  "broad_alignment_list",
+  "japan_rbcL"
 ))
@@
 
+All computational analyses were carried out in R v.`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The `r pack("targets")` R package was used to maintain analysis workflow [@Landau2021].
+
 ### Study Site
@@
 
-We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")`, and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`).
+We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`).
 We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
@@
 
+```{r phylo-analysis-stats}
+# Calculate stats related to Fern Tree of Life (FTOL) dataset
+
+# Define outgroup taxa
+plastome_og <-
+  tibble(
+    taxon = c(
+      "Physcomitrium_patens",
+      "Marchantia_polymorpha",
+      "Anthoceros_angustus",
+      "Isoetes_virginica",
+      "Lycopodium_clavatum",
+      "Selaginella_moellendorffii",
+      "Sciadopitys_verticillata",
+      "Ginkgo_biloba",
+      "Kadsura_longipedunculata",
+      "Trithuria_inconspicua",
+      "Amborella_trichopoda",
+      "Acorus_calamus",
+      "Magnolia_tripetala",
+      "Lamprocapnos_spectabilis",
+      "Plumbago_auriculata",
+      "Nicotiana_tabacum",
+      "Liquidambar_formosana",
+      "Sorbus_torminalis"),
+    taxon_status = "outgroup")
+
+# Define Sanger-sequenced genes
+sanger_genes <- c("atpA", "atpB", "rbcL", "rps4")
+
+# Make gene x taxon table for plastome alignment
+plastome_sampling <-
+  tibble(
+    gene = names(broad_alignment_list),
+    taxon = map(broad_alignment_list, rownames)
+  ) %>%
+  unnest(cols = taxon) %>%
+  # Check that all Sanger genes are included
+  verify(all(sanger_genes %in% gene)) %>%
+  # Check that all outgroup taxa are included
+  verify(all(plastome_og$taxon %in% taxon)) %>%
+  # Categorize genes into Sanger or other
+  mutate(data_type = if_else(
+    gene %in% sanger_genes,
+    "sanger",
+    "full_plastome"
+  ))
+
+# Number of full plastome genes in FTOL dataset
+n_plastome_genes <- plastome_sampling %>%
+  filter(data_type == "full_plastome") %>%
+  pull(gene) %>%
+  n_distinct() %>%
+  number(big.mark = ",")
+
+# Number of fern taxa with full plastome sequences in FTOL dataset
+n_plastome_fern_taxa <- plastome_sampling %>%
+  filter(data_type == "full_plastome") %>%
+  # Exclude outgroups
+  anti_join(plastome_og, by = "taxon") %>%
+  pull(taxon) %>%
+  n_distinct() %>%
+  number(big.mark = ",")
+
+# Number of fern taxa in FTOL dataset
+n_ftol_fern_taxa <- plastome_sampling %>%
+  # Exclude outgroups
+  anti_join(plastome_og, by = "taxon") %>%
+  pull(taxon) %>%
+  n_distinct() %>%
+  number(big.mark = ",")
+
+# Number of taxa in Japan fern rbcL dataset
+n_japan_fern_taxa <- japan_rbcL %>% 
+  rownames %>% 
+  n_distinct() %>%
+  number(big.mark = ",")
+
+# Verify that tips in the plastome tree are accounted for
+# by FTOL sampling + japan fern rbcL
+tibble(taxon = plastome_tree$tip.label) %>%
+  anti_join(select(plastome_sampling, taxon) %>% unique) %>%
+  anti_join(tibble(taxon = rownames(japan_rbcL))) %>%
+  verify(nrow(.) == 0, success_fun = success_logical)
+```
+
 Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
@@
 Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
-To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b with a globally sampled plastid dataset including all fern sequences on GenBank for four plastid genes (*rbcL*, *atpB*, *aptA*, and *rps4*) and 66 other coding, single-copy plastid genes extracted from 100 complete fern plastomes (Nitta et al., in prep) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total).
+To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
 We conducted phylogenetic analysis with `r software("IQTREE")` [@Nguyen2015] and molecular dating with `r software("treePL")` [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a.
@@
 We weighted continuous and qualitative traits equally, weighted each binary component trait of each qualitative trait equally, and log-transformed and scaled continuous traits prior to calculating distances.
+We also tested alternative weighting schemes, but these produced very similar results (data not shown).
 We then used the distances to build a trait dendrogram using the `r func("hclust")` function in the `r pack("stats")` R package [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
@@
 
-We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the `r pack("spdep")` package.
+We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the `r pack("spdep")` package [@Bivand2013].
 In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05) (`r s_table("moran")`). 
@@
 
-We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the `r pack("SpatialPack")` R package.
+We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the `r pack("SpatialPack")` R package [@SpatialPack2020].
 This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature (`r s_table("corr")`), so for the main modeling analysis we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
@@
 
-Author contributions: J.H.N., B.M., and A.E.
-conceived the ideas; A.E.
-provided the data; J.H.N.
-analyzed the data; I.W.
-provided computational resources; J.H.N.
-wrote the manuscript with input from all co-authors.
+Author contributions: J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.

### 2021-08-20 ms/manuscript.Rmd
Update MS
Add ref for Loiseau 2020

@@
 
-FIXME: add acknowledgements (not including reviewers)
+FIXME: add acknowledgements (not including reviewers) JSPS KAKENHI Grant Number 16H06279
 
@@
 To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
-We conducted phylogenetic analysis with `r software("IQTREE")` [@Nguyen2015] and molecular dating with `r software("treePL")` [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a.
+We conducted phylogenetic analysis with `r software("IQTREE")` [@Nguyen2015] and molecular dating with `r software("treePL")` [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
 We then trimmed the tree to Japanese taxa only and used this for all further analyses.

### 2021-08-24 ms/SI.Rmd
Update MS
Add ref for Loiseau 2020
Update MS
- Use CVD-friendly colors for maps
- Make coastline thinner to improve visibility
- Move color palette setup to ms_functions.R

@@
 
-```{r setup-si, include = FALSE, cache = FALSE}
-# Load functions only used when rendering Rmds
-source(here::here("R/ms_functions.R"), local = TRUE)
-# Set knitr options
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
-
+```{r setup-si, include = FALSE, cache = FALSE, eval = TRUE}
 # Load packages only used when knitting Rmd
@@
 library(ggtree)
+library(RColorBrewer)
+# Load functions and variables only used when rendering Rmds
+source(here::here("R/ms_functions.R"), local = TRUE)
+# Set knitr options
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
 ```
@@
 ```{r map-plot-prep}
-# Change cluster to factor
-biodiv_ferns_spatial <- biodiv_ferns_spatial %>%
-  mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
-  # Add redundancy
-  mutate(redundancy = 1 - (richness/abundance))
-
 # Download high-res map of Japan,
@@
 
-biodiv_ferns_endemic_spatial <- sf::st_set_crs(biodiv_ferns_endemic_spatial, japan_crs)
+# Format biodiv_ferns_spatial data
+biodiv_ferns_spatial <-
+  biodiv_ferns_spatial %>% 
+  # Set CRS
+  sf::st_set_crs(japan_crs) %>%
+  # Set order of factor levels for plotting
+  mutate(
+    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
+    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
+    across(
+      contains("signif"),
+      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
+    )
+  )
 
-biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
+biodiv_ferns_endemic_spatial <-
+  biodiv_ferns_endemic_spatial %>% 
+  # Set CRS
+  sf::st_set_crs(japan_crs) %>%
+  # Set order of factor levels for plotting
+  mutate(
+    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super"))
+    )
 
@@
   drop_apo_outlier
-
-# Set color palettes
-# - Randomization test significance
-signif_cols <-
-  c(
-    "> 0.99" = "#e31a1c", 
-    "> 0.975" = "#fb9a99",
-    "< 0.01" = "#1f78b4", 
-    "< 0.025" = "#a6cee3",
-    "not significant" = "#ffffcc"
-  )
-# - CANAPE
-canape_cols <-
-  c(
-    "neo" = "#e31a1c", 
-    "paleo" = "#1f78b4",
-    "mixed" = "#6a51a3",
-    "super" = "#4a1486",
-    "not significant" = "#ffffcc"
-  )
-# - Bioregions
-bioregion_cols <- rcartocolor::carto_pal(9, "Bold")[1:8] %>%
-  set_names(1:8)
 
@@
 
-**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding hybrids. Points at nodes indicate bootstrap support. Scalebar 100 my.
+**`r s_figure("tree")`**.
+Time tree of native Japanese ferns and lycophytes, excluding hybrids.
+Points at nodes indicate bootstrap support.
+Scalebar 100 my.
+Tree visualized with the `r pack("ggtree")` R package [@Yu2017].
 
@@
     ),
-    model_data = list(biodiv_ferns_used_for_model)) %>%
+    model_data = list(biodiv_ferns_used_for_model),
+    signif_cols = list(signif_cols)) %>%
   left_join(model_subplot_order, by = c("resp_var", "indep_var")) %>%
@@
 
+\clearpage
+
 ## References

### 2021-08-24 ms/manuscript.Rmd
Update MS
- Use CVD-friendly colors for maps
- Make coastline thinner to improve visibility
- Move color palette setup to ms_functions.R

@@
 
-```{r setup, include = FALSE}
+```{r setup, include = FALSE, cache = FALSE}
 # Load libraries used for rendering Rmds
@@
 library(patchwork)
+library(RColorBrewer)
 
@@
 
-# Load functions only used when rendering Rmds
+# Load functions and variables only used when rendering Rmds
 source(here::here("R/ms_functions.R"), local = TRUE)
@@
   "broad_alignment_list",
-  "japan_rbcL"
+  "japan_rbcL",
+  "ms_functions"
 ))
@@
 We then trimmed the tree to Japanese taxa only and used this for all further analyses.
-Tree visualization was done with the `r pack("ggtree")` R package [@Yu2017].
 
@@
 ```{r plot-prep}
+# (color palettes are set in ms_functions.R)
+
 # Define nicely formatted terms for printing variable names
@@
 japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
-
+japan_crs <- sf::st_crs(japan)
+
+# Format biodiv_ferns_spatial data
+biodiv_ferns_spatial <-
+  biodiv_ferns_spatial %>% 
   # Set CRS
-japan_crs <- sf::st_crs(japan)
-biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)
-
-# Set color palettes:
-# - Randomization test significance
-signif_cols <-
-  c(
-    "> 0.99" = "#e31a1c", 
-    "> 0.975" = "#fb9a99",
-    "< 0.01" = "#1f78b4", 
-    "< 0.025" = "#a6cee3",
-    "not significant" = "#ffffcc"
-  )
-# - CANAPE
-canape_cols <-
-  c(
-    "neo" = "#e31a1c", 
-    "paleo" = "#1f78b4",
-    "mixed" = "#6a51a3",
-    "super" = "#4a1486",
-    "not significant" = "#ffffcc"
-  )
-# - Bioregions
-bioregion_cols <- rcartocolor::carto_pal(9, "Bold")[1:8] %>%
-  set_names(1:8)
+  sf::st_set_crs(japan_crs) %>%
+  # Set order of factor levels for plotting
+  mutate(
+    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
+    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
+    across(
+      contains("signif"),
+      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
+    )
+  )
 ```
@@
 ```{r raw-diversity-make}
+# Use thinner coast lines for this plot
+# it's smaller, so the grey lines appear relatively thicker
+coast_lwd_old <- coast_lwd
+coast_lwd <- 0.05
+
 # Richness
 a <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
@@
 b <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
@@
   map_theme() +
-  theme(
-    axis.text.y = element_blank()
-  )
+  theme(axis.text.y = element_blank())
 
@@
 c <- ggplot() + 
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
@@
   plot_layout(nrow = 1)
+
+# Reset coastline width
+coast_lwd <- coast_lwd_old
 ```
@@
 a <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent", alpha = 0.8) +
+  # white base map, no outline
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  # colored cells
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent") +
+  # light grey coastline
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
@@
 b <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
@@
 c <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
@@
 d <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(d)** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
@@
 endemism_plot <- ggplot() + 
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   scale_fill_manual(values = canape_cols) +
@@
 a <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
@@
 b <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = phylo_cluster), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = phylo_cluster), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
@@
 
-cluster_cols <- carto_pal(8, "Bold")[1:4] %>%
-  set_names(1:4)
-
 a <- ggplot(data, aes(x = taxonomic_cluster, y = pd_obs_z)) +
   geom_jitter(aes(color = taxonomic_cluster)) +
-  scale_color_manual(values = cluster_cols, breaks = 1:4) +
+  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
   geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
@@
   geom_jitter(aes(color = taxonomic_cluster)) +
-  scale_color_manual(values = cluster_cols, breaks = 1:4) +
+  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
   geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
@@
   geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
-  scale_fill_viridis_d() +
-  scale_x_continuous(labels = scales::percent) +
+  scale_fill_manual(values = protection_cols) +
+  scale_x_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
   labs(

### 2021-08-24 R/ms_functions.R

@@ -1,29 +1,55 @@
-# Manuscript rendering functions and variables ----
+# Manuscript rendering functions and variables
 
+# Color palettes ----
+
+# - Coastline width across all plots
+coast_lwd <- 0.05
+# - Coast color across all plots
+coast_col <- "grey50"
+
+# - Randomization test significance (colorbrewer CVD safe)
+colorb_paired <- brewer.pal(n = 8, name = "Paired")
+colorb_ylgn <- brewer.pal(n = 5, name = "YlGn")
+signif_cols <-
+  c(
+    "> 0.99" = colorb_paired[[2]], # Dark blue 
+    "> 0.975" = colorb_paired[[1]], # Light blue 
+    "< 0.025" = colorb_paired[[5]], # Light red
+    "< 0.01" = colorb_paired[[6]], #  Dark red
+    "not significant" = colorb_ylgn[[1]] # Beige
+  )
+
+# - CANAPE (Okabe-Ito CVD safe)
+canape_cols <-
+  c(
+    "neo" = "#D55E00", # red
+    "paleo" = "#0072B2", # dark blue
+    "mixed" = "#009E73", # green
+    "super" = "#F0E442", # yellow
+    "not significant" = "grey90" # grey
+  )
+
+# - Bioregions (Okabe-Ito CVD safe)
+bioregion_cols <- c( 
+  "6" = "#E69F00", # goldenrod
+  "5" = "#56B4E9", # light blue
+  "1" = "#009E73", # green
+  "2" = "#F0E442", # yellow
+  "3" = "#0072B2", # dark blue
+  "4" = "#D55E00", # red
+  "7" = "#CC79A7", # magenta
+  "8" ="#000000") # black
+
+# - Protected areas (Okabe-Ito CVD safe)
+protection_cols <- c(
+  "Medium" = "#E69F00", # goldenrod
+  "High" = "#0072B2" # dark blue
+)
+
+# Words ----
 # Define formatting for some common custom words that may vary in style between journals
 ie <- "*ie*"
 eg <- "*eg*"
-
-#' Generate a path to save a results file
-#' 
-#' Only works for figures or tables cited in text, and outputs
-#' files to the "results" folder
-#'
-#' @param result_num Number of result, e.g. "Fig. 2"
-#' @param extension Extension to use for file
-#'
-#' @return String.
-#' 
-result_file <- function (result_num, extension) {
-  
-  fs::path(
-    here::here("results"),
-    result_num %>%
-      str_remove_all("\\.") %>% 
[…]

### 2021-08-24 _targets.R

@@ -671,10 +671,11 @@ tar_plan(
   
   # Render manuscript ----
   # Track ms files
-  tar_file(refs_yaml, here::here("ms/references.yaml")),
-  tar_file(refs_other_yaml, here::here("ms/references_other.yaml")),
-  tar_file(template_file, here::here("ms/template.docx")),
-  tar_file(csl_file, here::here("ms/apa-6th-edition.csl")),
+  tar_file(refs_yaml, "ms/references.yaml"),
+  tar_file(refs_other_yaml, "ms/references_other.yaml"),
+  tar_file(template_file, "ms/template.docx"),
+  tar_file(csl_file, "ms/apa-6th-edition.csl"),
+  tar_file(ms_functions, "R/ms_functions.R"),
   
   # MS, docx format
   tar_render(

### 2021-08-24 ms/data_exploration.Rmd
Update data_exploration.Rmd

@@
     biodiv_ferns_spatial %>%
+    # drop geometry (only need centroids for modeling)
+    st_set_geometry(NULL) %>%
     # keep only variables needed for model and only rows with zero missing data
-    filter_data_for_model(c("grids", "lat", "long", resp_vars_env, indep_vars)) %>%
-    sf::st_set_geometry(NULL) %>%
+    select(all_of(c("grids", "lat", "long", resp_vars_env, indep_vars))) %>%
+    ggplot2::remove_missing() %>%
     as_tibble()

### 2021-08-24 ms/SI.Rmd
Update MS
fix maps in SI
move more ms functions to ms_functions.R

@@
   ggplot() +
-    geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-    geom_sf(data = biodiv_ferns_for_plotting, aes(fill = .data[[var_select]]), color = "transparent", alpha = 0.8) +
+    geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+    geom_sf(data = biodiv_ferns_spatial, aes(fill = .data[[var_select]]), color = "transparent") +
+    geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
     annotate(
@@
 a <- ggplot() + 
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = abundance), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = abundance), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   scale_fill_scico(palette = "lajolla", label = k) +
@@
 b <- ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = redundancy), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = redundancy), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   scale_fill_scico(palette = "lajolla") +
@@
 ggplot() +
-  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent") +
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
   scale_fill_manual(values = canape_cols) +
@@
   ggplot() +
-    geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
-    geom_sf(data = genus_abun, aes(fill = richness), color = "transparent", alpha = 0.8) +
+    geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+    geom_sf(data = genus_abun, aes(fill = richness), color = "transparent") +
+    geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
     annotate(

### 2021-08-24 ms/manuscript.Rmd
Update MS
fix maps in SI
move more ms functions to ms_functions.R

@@
 
-All computational analyses were carried out in R v.`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The `r pack("targets")` R package was used to maintain analysis workflow [@Landau2021].
+All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The `r pack("targets")` R package was used to maintain analysis workflow [@Landau2021].
 
@@
 To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
-We conducted phylogenetic analysis with `r software("IQTREE")` [@Nguyen2015] and molecular dating with `r software("treePL")` [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
+We conducted phylogenetic analysis with `r software("IQTREE")` v1.6.12 [@Nguyen2015] and molecular dating with `r software("treePL")` v1.0 [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
 We then trimmed the tree to Japanese taxa only and used this for all further analyses.

### 2021-08-24 R/functions.R

@@ -2660,7 +2660,6 @@ calc_area_by_lat <- function(shp, lat_cut = 0.2, lat_window = 1) {
   
 }
 
-
 #' Add rolling mean area of latidudinal bands to biodiv data
 #'
 #' @param biodiv_ferns_spatial Spatial dataframe including biodiveristy metrics and grid cells
@@ -2759,143 +2758,5 @@ add_selected_pred_to_model <- function(spatial_models) {
       TRUE ~ NA_character_
     )) %>%
     assert(not_na, indep_var_select)
-  
-}
-
-# Manuscript rendering ----
-
-#' Rename response variables in data frame
-#' 
-#' For formatting results tables
-#'
-#' @param df 
-#'
-#' @return df with renamed variables
-#' 
-rename_resp_vars <- function(df) {
-  
-  lookup <-
-    tibble(
-      resp_var = c("richness", "fd_obs_z", "pd_obs_z", "rfd_obs_z", "rpd_obs_z"),
-      resp_var_print = c("Richness", "SES of FD", "SES of PD", "SES of RFD", "SES of RPD")
-    )
-  
-  df %>%
-    left_join(lookup, by = "resp_var") %>%
-    select(-resp_var) %>%
-    rename(resp_var = resp_var_print)
-  
-}
-
-#' Rename model type in data frame
-#' 
-#' For formatting results tables
-#'
-#' @param df 
-#'
-#' @return df with renamed model types
-#' 
-rename_model_type <- function(df) {
-  
-  lookup <-
-    tibble(
-      model_type = c("percent_apo", "temp"),
-      model_type_print = c("Reproductive", "Environmental")
-    )
-  
-  df %>%
-    left_join(lookup, by = "model_type") %>%
-    select(-model_type) %>%
-    rename(model_type = model_type_print)
-  
-}
-
-#' Clean a vector of variables for printing
-#'
-#' @param x Character vector (variables)
-#'
-#' @return Character vector
-#' 
-clean_vars <- function(x) {
-  x %>%
-    gsub("^([^_]+)_obs_z", "SES of \\U\\1", ., perl=TRUE) %>%
-    str_replace_all("^percent_apo$", "% apomictic taxa") %>%
-    str_replace_all("^temp$", "Temperature") %>%
-    str_replace_all("^temp_season$", "Temperature seasonality") %>%
-    str_replace_all("^precip_season$", "Precipitation seasonality") %>%
-    str_replace_all("^precip$", "Precipitation") %>%
-    str_replace_all("^area$", "Area") %>%
[…]

### 2021-08-24 R/ms_functions.R

@@ -26,7 +26,7 @@ canape_cols <-
     "paleo" = "#0072B2", # dark blue
     "mixed" = "#009E73", # green
     "super" = "#F0E442", # yellow
-    "not significant" = "grey90" # grey
+    "not significant" = "grey97" # grey, almost white
   )
 
 # - Bioregions (Okabe-Ito CVD safe)
@@ -168,3 +168,142 @@ figure_num <- function (name) {figure(name) %>% str_remove("Fig. ")}
 table_num <- function (name) {table(name) %>% str_remove("Table ")}
 s_figure_num <- function (name) {s_figure(name) %>% str_remove("Fig. ")}
 s_table_num <- function (name) {s_table(name) %>% str_remove("Table ")}
+
+# Etc ----
+
+#' Rename response variables in data frame
+#' 
+#' For formatting results tables
+#'
+#' @param df 
+#'
+#' @return df with renamed variables
+#' 
+rename_resp_vars <- function(df) {
+  
+  lookup <-
+    tibble(
+      resp_var = c("richness", "fd_obs_z", "pd_obs_z", "rfd_obs_z", "rpd_obs_z"),
+      resp_var_print = c("Richness", "SES of FD", "SES of PD", "SES of RFD", "SES of RPD")
+    )
+  
+  df %>%
+    left_join(lookup, by = "resp_var") %>%
+    select(-resp_var) %>%
+    rename(resp_var = resp_var_print)
+  
+}
+
+#' Rename model type in data frame
+#' 
+#' For formatting results tables
+#'
+#' @param df 
+#'
+#' @return df with renamed model types
+#' 
+rename_model_type <- function(df) {
+  
+  lookup <-
+    tibble(
+      model_type = c("percent_apo", "temp"),
+      model_type_print = c("Reproductive", "Environmental")
+    )
+  
+  df %>%
+    left_join(lookup, by = "model_type") %>%
+    select(-model_type) %>%
+    rename(model_type = model_type_print)
+  
+}
+
+#' Clean a vector of variables for printing
+#'
+#' @param x Character vector (variables)
+#'
+#' @return Character vector
+#' 
+clean_vars <- function(x) {
+  x %>%
+    gsub("^([^_]+)_obs_z", "SES of \\U\\1", ., perl=TRUE) %>%
+    str_replace_all("^percent_apo$", "% apomictic taxa") %>%
+    str_replace_all("^temp$", "Temperature") %>%
+    str_replace_all("^temp_season$", "Temperature seasonality") %>%
[…]

### 2021-08-26 ms/SI.Rmd
Don't include raw occurrence data in _targets.R
Major refactor. Run process_raw_data.R  first to output cleaned data to data/, then run _targets.R

@@
   "japan_shp",
-  "comm_scaled_list",
-  "occ_point_data_ferns",
+  "inext_res",
+  "redundancy_by_res",
+  "redundancy_summary",
   "biodiv_ferns_spatial",
@@
 ```{r map-plot-prep}
-# Download high-res map of Japan,
-# crop to spatial div results area
+# Crop high-res map of Japan to spatial div results area
 japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
@@
 ```{r sampling-curve, fig.width = 4, fig.height = 3}
-# Format fern abundances (no hybrids) as numeric vector
-fern_abun <-
-  occ_point_data_ferns %>%
-  group_by(taxon) %>%
-  count(sort = TRUE) %>%
-  pull(n)
-
-# Run iNEXT
-inext_out <- iNEXT::iNEXT(fern_abun, q=0, datatype="abundance")
-
 # create subset of observed points to add to plot as dots
-observed_data <- inext_out$iNextEst[inext_out$iNextEst$method == "observed",]
+observed_data <- inext_res[inext_res$method == "observed",]
 
 # check estimated completeness
-est_completeness <- inext_out$iNextEst %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
+est_completeness <- inext_res %>% filter(method == "observed") %>% pull(SC) %>% scales::percent()
 rich_interval <- glue::glue("{observed_data$qD.LCL %>% round} to {observed_data$qD.UCL %>% round}")
@@
 # Make iNEXT plot
-inext_out$iNextEst %>%
+inext_res %>%
   filter(method != "observed") %>%
@@
 
-# Combine results of building community matrices at different grain sizes
-redundancy_by_res <- 
-  list(
-    "0.1" = comm_scaled_list %>% filter(resol == 0.1) %>% pull(poly_shp) %>% pluck(1),
-    "0.2" = comm_scaled_list %>% filter(resol == 0.2) %>% pull(poly_shp) %>% pluck(1),
-    "0.3" = comm_scaled_list %>% filter(resol == 0.3) %>% pull(poly_shp) %>% pluck(1),
-    "0.4" = comm_scaled_list %>% filter(resol == 0.4) %>% pull(poly_shp) %>% pluck(1)
-  ) %>%
-  map_df(~as_tibble(.) %>% select(grids, abundance, richness), .id = "res") %>%
-  # Calculate redundancy
-  mutate(redundancy = 1 - (richness/abundance))
-
 # Summarize redundancy statistics by grain size
-redundancy_summary <-
-  redundancy_by_res %>%
-  mutate(above_50 = ifelse(redundancy > 0.5, 1, 0)) %>%
-  mutate(singleton = ifelse(redundancy == 0, 1, 0)) %>%
-  group_by(res) %>%
-  summarize(
-    n_total = n(),
-    n_above_50 = sum(above_50),
-    n_singletons = sum(singleton),
-    percent_above_50 = n_above_50/n_total,
-    percent_singletons = n_singletons/ n_total,
-    .groups = "drop"
-  )
+redundancy_summary <- sum_redundancy_by_res(redundancy_by_res)
 

### 2021-08-26 ms/manuscript.Rmd
Don't include raw occurrence data in _targets.R
Major refactor. Run process_raw_data.R  first to output cleaned data to data/, then run _targets.R

@@
 tar_load(c(
-  "repro_data",
-  "occ_point_data_ferns",
-  "occ_point_data_ferns_unfiltered",
+  "aic_env_repro",
+  "binary_sig_results",
+  "biodiv_ferns_cent",
   "biodiv_ferns_spatial",
+  "broad_alignment_list",
+  "comm_ferns",
+  "comm_ferns_full",
+  "csl_file",
+  "fern_traits",
+  "green_list",
+  "japan_fern_tree",
+  "japan_points_raw",
+  "japan_pterido_tree",
+  "japan_rbcL",
+  "japan_shp",
+  "k_phylogeny",
   "k_taxonomy",
-  "k_phylogeny",
-  "signif_cells_protected_area",
-  "protected_areas",
-  "japan_shp",
+  "lat_span_summary",
+  "lrt_results",
+  "model_fits",
+  "model_params",
+  "ms_functions",
+  "occ_point_data_summary",
   "phy_sig_results",
-  "binary_sig_results",
-  "traits_summary",
-  "fern_traits",
-  "japan_pterido_tree",
-  "japan_fern_tree",
-  "ppgi",
-  "occ_point_data_ferns",
-  "comm_ferns_full",
-  "comm_ferns",
-  "japan_points_raw",
   "plastome_calibration_dates",
   "plastome_tree",
-  "green_list",
-  "model_params",
-  "lrt_results",
-  "aic_env_repro",
-  "model_fits",
-  "biodiv_ferns_cent",
+  "ppgi",
+  "protected_areas",
+  "refs_other_yaml",
   "refs_yaml",
-  "refs_other_yaml",
+  "repro_data",
+  "signif_cells_protected_area",
   "template_file",
-  "csl_file",
-  "broad_alignment_list",
-  "japan_rbcL",
-  "ms_functions"
+  "traits_summary"
 ))
@@
 
-FIXME: add acknowledgements (not including reviewers) JSPS KAKENHI Grant Number 16H06279
+Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
+Shawn Laffan provided advice about CANAPE.
+This study was supported by JSPS KAKENHI Grant Number 16H06279.
 
@@
 
+```{r abstract-stats}
+# Convert occ_point_data_summary dataframe to nested list for easier indexing
+occ_sum <- occ_point_data_summary %>%
+  # change dataset name to just "unfiltered" vs. "filtered"
+  mutate(dataset = case_when(
+    str_detect(dataset, "unfiltered") ~ "unfiltered",
+    TRUE ~ "filtered")) %>%
+  split(.$dataset) %>%
+  map(., ~split(., .$variable))
+```
+
 **Aim** To characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.
@@
 
-**Methods** We used data from >`r nrow(occ_point_data_ferns) %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
+**Methods** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
 We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode.
@@
 Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
-The original list comprises `r nrow(occ_point_data_ferns_unfiltered) %>% scales::number(big.mark = ",")` specimens representing `r occ_point_data_ferns_unfiltered %>% pull(taxon) %>% n_distinct()` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
-We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html).
+The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
+We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
 
@@
 ### Taxon list ###
-# Make list of all native ferns in Japan (based off Green List), includes hybrids
+# Make list of all native ferns in Japan (based off Green List), *including* hybrids
 fern_list <-
@@
 
-### Phylogeny ###
+# Caculate number of native, non-hybrid fern taxa
 total_fern_taxa <- fern_list %>%
@@
 
+### Phylogeny ###
 phylogeny_sampled_percent <- ape::Ntip(japan_fern_tree) %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)
@@
 
-### Specimens ###
-n_specimens <- nrow(occ_point_data_ferns) %>% scales::number(big.mark = ",")
-n_specimens_taxa <- occ_point_data_ferns %>% pull(taxon) %>% n_distinct()
+### Specimens (filtered occurrence data) ###
+n_specimens <- occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")
+n_specimens_taxa <- occ_sum$filtered$n_taxa$value
 n_specimens_taxa_percent <- n_specimens_taxa %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)
@@
 ```{r distribution-stats}
+# Summarize maximum latitude by genus, family
+max_lat_summary <- summarize_fern_lat_max(lat_span_summary, ppgi)
+
 # Calculate number of genera and families restricted to lat < 30.1
 southern_genera <-
-  occ_point_data_ferns %>%
-  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
-  group_by(genus) %>%
-  summarize(
-    max_lat = max(latitude),
-    .groups = "drop"
-  ) %>%
-  filter(max_lat < 30.1)
-
-n_southern_genera <- southern_genera %>% pull(genus) %>% n_distinct %>% english2
+  max_lat_summary %>%
+  filter(max_lat < 30.1, taxon_level == "genus")
+
+n_southern_genera <- southern_genera %>% pull(higher_taxon) %>% n_distinct %>% english2
 
 southern_families <-
-  occ_point_data_ferns %>%
-  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
-  left_join(ppgi, by = "genus") %>%
-  group_by(family) %>%
-  summarize(
-    max_lat = max(latitude),
-    .groups = "drop"
-  ) %>%
-  filter(max_lat < 30.1)
-
-n_southern_families <- southern_families %>% pull(family) %>% n_distinct %>% english2
+  max_lat_summary %>%
+  filter(max_lat < 30.1, taxon_level == "family")
+
+n_southern_families <- southern_families %>% pull(higher_taxon) %>% n_distinct %>% english2
 ```
@@
 ```{r limits-stats}
-# Summarize occurrences by latitude
-occ_data_by_lat <-
-  occ_point_data_ferns %>%
-  group_by(taxon) %>%
-  summarize(
-    min_lat = min(latitude),
-    max_lat = max(latitude)
-  ) %>%
-  mutate(
-    lat_span = max_lat - min_lat
-  )
-
 # Find the species with the widest latitudinal distribution
-taxon_with_widest_lat <- occ_data_by_lat %>%
+taxon_with_widest_lat <- lat_span_summary %>%
+  mutate(lat_span = max_lat - min_lat) %>%
   slice_max(lat_span) %>%
   left_join(green_list, by = "taxon") %>%
-  taxastand::add_parsed_names(sci_name = scientific_name, species, author) %>%
-  mutate(text = glue("*{species}* {author}"))
+  # Parse out species name and author
+  mutate(gn_parse_tidy(.$scientific_name) %>% select(canonicalsimple, authorship)) %>%
+  mutate(text = glue("*{canonicalsimple}* {authorship}"))
 
@@
 n_widespread_taxa <-
-  occ_data_by_lat %>%
+  lat_span_summary %>%
   # Filter by approx location of Tokara strait (29) and Tsugaru Strait (41.5)

### 2021-08-26 R/process_raw_data.R

@@ -1,10 +1,7 @@
 # Separate targets plan to process raw data that can't be uploaded to Dryad
 #
 # This should be run with:
-# targets::tar_make(
-#   script = "R/process_raw_data.R",
-#   store = "_raw_data_cache"
-# )
+# targets::tar_make(script = "R/process_raw_data.R", store = "_raw_data_cache")
 
 library(targets)
 library(tarchetypes)
@@ -14,6 +11,12 @@ source("R/packages.R")
 source("R/functions.R")
 source("R/tests.R")
 
+# Set parallel back-end
+plan(callr)
+
+# Summary: from raw occurrence data, test binning into grid cells at four scales,
+# select the optimal scale, filter out poorly sampled grid cells, generate
+# community data matrix (sites x species)
 tar_plan(
   
   # Load "green list" from Ebihara and Nitta 2019
@@ -41,7 +44,23 @@ tar_plan(
   
   # Subset to just ferns (674 taxa)
   occ_point_data_ferns_unfiltered = subset_to_ferns(occ_point_data, ppgi),
-  
+
+  # Calculate summary statistics for occurrence data, write out as CSV
+  occ_point_data_summary = summarize_occ_data(
+    occ_point_data_ferns_unfiltered = occ_point_data_ferns_unfiltered, 
+    occ_point_data_ferns = occ_point_data_ferns),
+  tar_file(
+    occ_point_data_summary_out,
+    write_csv_tar(occ_point_data_summary, "data/japan_ferns_occ_summary.csv")
+  ),
+
+  # Calculate latitudinal span by taxon, write out as CSV
+  lat_span_summary = summarize_fern_lat_span(occ_point_data_ferns),
+  tar_file(
+    lat_span_summary_out,
+    write_csv_tar(lat_span_summary, "data/japan_ferns_lat_span.csv")
+  ),
+
   # Filter out duplicates, restrict to only points in second-degree mesh
   # Shape file downloaded from http://gis.biodic.go.jp/
   # http://gis.biodic.go.jp/BiodicWebGIS/Questionnaires?kind=mesh2&filename=mesh2.zip
@@ -62,6 +81,69 @@ tar_plan(
       lat = "latitude",
       species = "taxon"),
     pattern = map(scales_to_test)
-  )
+  ),
+
+  # After inspecting results (see SI), select 0.2 degrees as optimal scale.
   
[…]

### 2021-08-26 _targets.R

@@ -21,10 +21,10 @@ tar_plan(
   
   # Load and process various data ----
   
-  # Load data files from Ebihara and Nitta 2019
-  # This requires doi_10.5061_dryad.4362p32__v4.zip to be downloaded to data_raw/
+  ## Data files from Ebihara and Nitta 2019 ----
+  # This requires doi_10.5061_dryad.4362p32__v4.zip to be downloaded to data/
   # from https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32 first
-  tar_file(ebihara_2019_zip_file, "data_raw/doi_10.5061_dryad.4362p32__v4.zip"),
+  tar_file(ebihara_2019_zip_file, "data/doi_10.5061_dryad.4362p32__v4.zip"),
 
   # - Pteridophyte Phylogeny Group I (PPGI) taxonomy,
   # modified slightly for ferns of Japan
@@ -38,12 +38,20 @@ tar_plan(
   
   # - Japan rbcL alignment
   japan_rbcL_raw = read_ja_rbcL_from_zip(ebihara_2019_zip_file),
+
+  # - Summary of occurrence point data
+  tar_file(occ_point_data_summary_file, "data/japan_ferns_occ_summary.csv"),
+  occ_point_data_summary = read_csv(occ_point_data_summary_file),
   
-  # Load a map of Japan
-  # downloaded from https://www.gsi.go.jp/kankyochiri/gm_japan_e.html
+  # - Summary of latitudinal span by taxon
+  tar_file(lat_span_summary_file, "data/japan_ferns_lat_span.csv"),
+  lat_span_summary = read_csv(lat_span_summary_file),
+  
+  ## Map data ----
+  # Load shape file of Japan downloaded from https://www.gsi.go.jp/kankyochiri/gm_japan_e.html
   # on 2020-08-26
-  tar_file(japan_pol_file, "data_raw/gm-jpn-all_u_2_2/polbnda_jpn.shp"),
-  jpn_pol = st_read(japan_pol_file),
+  tar_file(japan_pol_zip_file, "data/gm-jpn-all_u_2_2.zip"),
+  jpn_pol = load_shape_from_zip(japan_pol_zip_file, "polbnda_jpn.shp"),
   
   # Collapse all the political units down to just one shape for the country
   japan_shp = jpn_pol %>%
@@ -51,65 +59,31 @@ tar_plan(
     summarize(),
   
   # Load manually entered points of interest for drawing a map of Japan
[…]

### 2021-08-26 ms/manuscript.Rmd
Update MS
per comments from BM and AE 2021-08-26

@@
 <!-- may include title words -->
-**Key words:** apomixis, biodiversity, biogeography, ferns, Japan, phylogenetic diversity.
+**Key words:** apomixis, biodiversity, biogeography, CANAPE, ferns, Japan, phylogenetic diversity.
 
@@
 The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness. 
-Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 34 terrestrial hotspots [@Myers2000] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006].
+Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 36 terrestrial hotspots [@Myers2000; @Noss2016] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006].
 However, richness alone cannot provide a complete picture of biodiversity. 
@@
 Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
-The main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ryukyu and Ogasawara archipelagos are oceanic islands that have never been connected to the mainland.
+Most of the main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ogasawara archipelago consists of oceanic islands that have never been connected to the mainland.
 The climate is subtropical in the south and temperate in other parts of the country. 
@@
 
-We used a list of all non-hybrid fern specimens housed at herbaria in Japan. 
+We used a list of all non-hybrid fern specimens deposited at TNS (herbarium codes follow Thiers2021). 
 We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v.
@@
 
-We measured phylogenetic diversity (PD) using the Faith's PD (the total branch length connecting all terminal taxa in a community, including the root of the phylogeny) [@Faith1992]. 
-To detect areas with unusually long or short branch lengths, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].
+We measured phylogenetic diversity (PD) using the Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny; @Faith1992]. 
+To detect areas with unusually long or short branches, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].
 
@@
 
-We measured phylogenetic endemism (PE) as the sum of all branches at a site inversely weighted by their range [@Rosauer2009].
-
-Observed values of PD, RPD, FD, RFD, and PE are known to be closely related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others.
+We measured phylogenetic endemism (PE) as the sum of all branches at a site inversely weighted by their range size [@Rosauer2009].
+
+Observed values of PD, RPD, FD, RFD, and PE are known to be related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others.
 Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test. 
@@
 
-We tested whether endemic areas have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
-Low RPE indicates short branchlengths (neoendemism) and high RPE indicates long branchlengths (paleoendemism). 
+We tested whether centers of endemism have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
+Low RPE indicates short range-restricted branches (neoendemism) and high RPE indicates long range-restricted branches (paleoendemism). 
 Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution.
@@
 This is especially expected for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country. 
-To partially account for this, we also calculated all endemism-related metrics using a dataset including only species endemic to Japan.
+To partially account for this, we also calculated all endemism-related metrics using a dataset including species found only in Japan.
 
@@
 We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html). 
-We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some human activities allowed by permit).
+We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some economic activities allowed by permit) following [@Kusumoto2017].
 We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
@@
 
+fern_list_non_hybrid <-
+  fern_list %>%
+  filter(hybrid == FALSE) %>%
+  mutate(gn_parse_tidy(.$scientific_name) %>% select(taxon = canonicalsimple)) %>%
+  mutate(taxon = str_replace_all(taxon, " ", "_"))
+
 # Caculate number of native, non-hybrid fern taxa
-total_fern_taxa <- fern_list %>%
-  filter(hybrid == FALSE) %>%
+total_fern_taxa <- fern_list_non_hybrid %>%
   pull(taxon) %>%
@@
 # Calculate number/percent of missing data in trait matrix
+trait_sampling_summary <-
+fern_list_non_hybrid %>%
+  select(taxon) %>%
+  verify(all(fern_traits$taxon %in% taxon)) %>%
+  left_join(
+    transmute(fern_traits, taxon, sampled = TRUE), by = "taxon"
+  ) %>%
+  mutate(sampled = replace_na(sampled, FALSE))
+
+n_taxa_traits_sampled <- sum(trait_sampling_summary$sampled)
+percent_taxa_traits_sampled <- (n_taxa_traits_sampled / total_fern_taxa) %>% scales::percent(accuracy = 0.1)
+
 fern_traits_missing_summary <-
@@
 
-The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent`) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families of native Japanese ferns. 
+The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
 The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
@@
 After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits) (`r table("traits-used")`).
-The trait matrix included `r nrow(fern_traits)` taxa and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.
+The trait matrix included `r n_taxa_traits_sampled` taxa (`r percent_taxa_traits_sampled`) and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.
 
@@
 Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands. 
-Small clusters of significantly endemic cells were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`). 
-When we ran CANAPE with a reduced dataset including only taxa absolutely endemic to Japan to account for the border effect, there were fewer significant cells, but a similar pattern of significantly endemic cells mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).
+Small clusters of cells containing significant endemism were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`). 
+When we ran CANAPE with a reduced dataset including only taxa absolutely restricted to Japan to account for the border effect, there were fewer significant cells, but a similar pattern of significant endemism mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).
 
@@
 
-Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half on Honshu, particularly Kyushu (`r figure("rand-diversity")`b).
+Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half of Honshu, and particularly on Kyushu (`r figure("rand-diversity")`b).
 The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a).
@@
 
-Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`g).
+Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`g), and to a slightly lesser extent SES of PD (`r figure("modelfit")`f).
 Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
 This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns.
-We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), consistent with an effect on genetic diversity (`r s_figure("modelfit-not-ult")`b).
+We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`b).
 While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`).
@@
 Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014].
-This suggests that ferns are useful as possible bioindicators in Japan. 
+This suggests that ferns are useful as bioindicators in Japan. 
 Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions.
@@
 
-Although the different facets of biodiversity have different distribution patterns, we find that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric (`r figure("conserv-status")`).
+Although the different facets of biodiversity have different distribution patterns, we found that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric (`r figure("conserv-status")`).
 One possible reason is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas (A. Ebihara, pers. comm.). 
@@
 Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change.
-One such other threat is herbivory by native Japanese deer, *Cervus nippon*. 
+One other such threat is herbivory by native Japanese deer, *Cervus nippon*. 
 Populations of these deer have increased rapidly since the 1970s due to a combination of climate change and a decrease in human hunters and natural predators [@Takatsuki2009]. 
@@
 There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research. 
-First, the data include species collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
+First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
 Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (`r s_figure("abundance")`). 
@@
 
-One piece of biodiversity that we did not consider here is hybrids. 
+One element of biodiversity that we did not consider here is hybrids. 
 The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa. 
-Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification. 
+Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989]. 
 Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
-Methods for calculating PD would need to be modified to take into account multiple lineages contributing to a terminal taxon.
+Methods for calculating PD would need to be modified to take into account multiple lineages contributing to one terminal taxon.
 
@@
 **`r figure("raw-diversity")`**.
-Raw taxonomic (a), phylogenetic (b), and functional (c) diversity of the ferns of Japan.
+(a) Raw taxonomic diversity (b), phylogenetic diversity, and (c) functional diversity of the ferns of Japan.
 For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
@@
 (a), Phylogenetic diversity (PD).
-(b), Relative phylogenetic diveristy (RPD).
-(c), Functional diveristy (FD).
-(d), relative functional diveristy (RFD).
+(b), Relative phylogenetic diversity (RPD).
+(c), Functional diversity (FD).
+(d), relative functional diversity (RFD).
 For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
@@
 Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).
-'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+'paleo' indicates areas with an overabundance of rare long branches; 'neo' indicates areas with an overabundance of rare short branches; 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
 For details, see Materials and Methods.
@@
 **`r figure("conserv-status")`**.
-Percent of land area with protected status for grid cells with significantly high biodiversity.
+Percent of land area with protected status for grid cells with significantly high biodiversity by at least one metric.
 Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
-Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Methods).
+Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Materials and Methods).
 Vertical lines indicate total percent area protected across Japan regardless of degree of biodiversity; dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).

### 2021-08-30 ms/manuscript.Rmd
Fix combining forest PAs with other PAs
Make sure medium and high levels don't overlap

@@
   protected_areas %>%
+  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
   as_tibble() %>%

### 2021-08-30 R/process_raw_data.R

@@ -156,6 +156,13 @@ tar_plan(
   protected_areas_other = load_protected_areas(protected_areas_zip_file),
 
   tar_file(protected_areas_forest_folder, "data_raw/forest_area_zip_files"),
-  protected_areas_forest = load_protected_areas_forest(protected_areas_forest_folder)
+  protected_areas_forest = load_protected_areas_forest(protected_areas_forest_folder),
+
+  # Combine protected areas (high and medium only), write out
+  protected_areas = combine_pa(protected_areas_other, protected_areas_forest),
+  tar_file(
+    protected_areas_out,
+    st_write_tar(protected_areas, "data/japan_protected_areas.gpkg", time_stamp = as.Date("2021-08-30"))
+  )
   
 )

### 2021-08-30 _targets.R

@@ -522,12 +522,14 @@ tar_plan(
   ## Read in protected areas, assign protection levels following Kusumoto et al. 2017:
   # - high: no human activities allowed
   # - medium: permission required for economic activities
-  # - low: protected area, but none of the above restrictions
-  tar_file(protected_areas_zip_file, "data/map17.zip"),
-  protected_areas = load_protected_areas(protected_areas_zip_file),
+  tar_file(protected_areas_zip_file, "data/japan_protected_areas.gpkg"),
+  protected_areas = st_read(protected_areas_zip_file),
   
-  # Calculate percent protection for grid cells with significantly high biodiversity
-  signif_cells_protected_area = calculate_protected_area(biodiv_ferns_spatial, protected_areas, japan_shp),
+  # Crop siginficantly diverse grid cells to protected areas
+  protected_biodiv = crop_by_pa(protected_areas, biodiv_ferns_spatial, japan_shp),
+
+  # Calculate percentage of protected areas
+  signif_cells_protected_area = calculate_percent_protected(protected_biodiv),
   
   # Render manuscript ----
   # Track ms files

### 2021-08-30 ms/manuscript.Rmd
Fix calculation of baseline protected area

--- 2021-08-26 ms/manuscript.Rmd
@@
   protected_areas %>%
+  st_make_valid %>%
+  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
+  st_drop_geometry %>%
   as_tibble() %>%
-  select(-geometry) %>%
   filter(status %in% c("medium", "high")) %>%

### 2021-08-30 ms/manuscript.Rmd
Lump small bioregions in MS

@@
 ---
-title: "Exploring dimensions of biodiversity in Japanese ferns"
+title: "Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation"
 mainfont: Roboto-Regular.ttf
@@
 
-Short running title: Biodiversity of Japanese ferns
+Short running title: Spatial phylogenetics of Japanese ferns
 
@@
 ```{r bioregions-make}
+# Lump regions with just a few cells (3 or less)
+biodiv_ferns_spatial_lumped <-
+biodiv_ferns_spatial %>%
+  mutate(taxonomic_cluster = fct_lump_min(taxonomic_cluster, 4))  %>%
+  mutate(phylo_cluster = fct_lump_min(phylo_cluster, 4))
+
 # Taxonomic bioregions
@@
   geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent") +
+  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = taxonomic_cluster), color = "transparent") +
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
@@
   geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = phylo_cluster), color = "transparent") +
+  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = phylo_cluster), color = "transparent") +
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
@@
 (b), phylogenetic bioregions.
-Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances into *k* = 8 or 7 groups, respectively.
+Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances.
+Bioregions with three or fewer grid cells are lumped into the "Other" category.
 

### 2021-08-31 ms/SI.Rmd
Change 'cluster' to 'bioregion'
Group small bioregions (artifacts)

@@
   select(grids, cluster = taxonomic_cluster) %>%
+  # mutate(cluster = as.character(cluster)) # %>%
   # Convert cluster to numeric to use max() later
-  mutate(cluster = levels(cluster)[cluster] %>% as.numeric)
+  mutate(cluster = levels(cluster)[cluster])
 
-cluster_nodes_tax_tbl <- map_df(1:max(clusters_tax$cluster), ~map_cluster_to_nodes(dendro_tax, clusters_tax, .)) %>%
-  mutate(cluster = as.factor(cluster))
+cluster_nodes_tax_tbl <- map_df(unique(clusters_tax$cluster), ~map_cluster_to_nodes(dendro_tax, clusters_tax, .)) %>%
+  # Lump regions with just a few cells (3 or less)
+  mutate(cluster = as.factor(cluster) %>% fct_lump_min(4))
 
@@
   scale_color_manual(
-    values = bioregion_cols, na.value = "grey", name = "Cluster") +
+    values = bioregion_cols[c(1,2,3,4,"Other")], breaks = c(1,2,3,4,"Other"), name = "Cluster", na.value = "grey10") +
   theme(legend.position = "right") +
@@
 **`r s_figure("dendrogram")`**.
-Bioregion dendrogram.
-Tips are sites (names not shown); color shows cluster; internal branches between clusters 'NA'.
+Bioregion dendrograms.
+Tips are sites (names not shown).
+Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods).
+Color shows bioregion; bioregions with three or fewer grid cells are lumped into the "Other" category.
 (a), taxonomic bioregions.
 (b), phylogenetic bioregions.
+Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
 \clearpage
@@
 clusters_phy <-
-  biodiv_ferns_spatial %>%
+  biodiv_ferns_spatial_lumped %>%
   as_tibble() %>%

### 2021-08-31 ms/manuscript.Rmd
Change 'cluster' to 'bioregion'
Group small bioregions (artifacts)

@@
 
-Analysis of bioregions grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` clusters (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell to a subset of these clusters: using taxonomic distances, they include Cluster 1 (Hokkaido and northern Honshu), Cluster 2 (southern Honshu), Cluster 3 (Ryukyu Islands, excluding Yakushima), and Cluster 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single cluster (`r figure("bioregions")`b; `r s_figure("dendrogram")`b). 
-The other much smaller clusters are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
-
-Taxonomic clusters 3 and 4 have higher SES values of PD, RPD, FD, and RFD than clusters 1 and 2 (`r figure("div-by-tax-region")`). 
-Clusters 3 and 4 also have consistently high PE *p*-scores, whereas clusters 1 and 2 have a much wider variation. 
-Cluster 2 has lower RPD and RFD than cluster 1. 
-Comparison of biodiversity metrics across phylogenetic clusters was similar, except that phylogenetic cluster 3 mostly corresponds to taxonomic clusters 3 and 4 combined.
+Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
+The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu), bioregion 3 (Ryukyu Islands, excluding Yakushima), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a).
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (`r figure("bioregions")`b; `r s_figure("dendrogram")`b). 
+The other much smaller bioregions (each including only three grid cells or less) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
+
+Taxonomic bioregions 3 and 4 have higher SES values of PD, RPD, FD, and RFD than bioregions 1 and 2 (`r figure("div-by-tax-region")`). 
+Bioregions 3 and 4 also have consistently high PE *p*-scores, whereas bioregions 1 and 2 have a much wider variation. 
+Bioregion 2 has lower RPD and RFD than bioregion 1. 
+Comparison of biodiversity metrics across phylogenetic bioregions was similar, except that phylogenetic bioregion 3 mostly corresponds to taxonomic bioregions 3 and 4 combined.
 
@@
 The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`).
-Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; Cluster 1 corresponds to the cool or subartic deciduous type, Cluster 2 to the warm evergreen type, and Clusters 3 and 4 to the subtropical rainy type [@Shimizu2014].
+Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
 This suggests that ferns are useful as bioindicators in Japan. 
 Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions.
-The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits clusters 2 and 3, there is some overlap between them on the Amami-Oshima Isl.
+The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits bioregions 2 and 3, there is some overlap between them on the Amami-Oshima Isl.
 (`r figure("bioregions")`).
 The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
-Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic Cluster 3 (mostly the Ryukyu islands; wetter) from Cluster 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).
+Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu islands; wetter) from bioregion 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).
 
@@
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
-  labs(fill = "Cluster") +
+  labs(fill = "Bioregion") +
   scale_fill_manual(values = bioregion_cols) +
@@
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
-  labs(fill = "Cluster") +
+  labs(fill = "Bioregion") +
   scale_fill_manual(values = bioregion_cols) +
@@
 (b), phylogenetic bioregions.
-Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances.
+Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid cells.
 Bioregions with three or fewer grid cells are lumped into the "Other" category.
@@
     x = "Cluster",
-    color = "Cluster"
+    color = "Bioregion"
   ) +

### 2021-08-31 R/ms_functions.R

@@ -31,14 +31,15 @@ canape_cols <-
 
 # - Bioregions (Okabe-Ito CVD safe)
 bioregion_cols <- c( 
-  "6" = "#E69F00", # goldenrod
-  "5" = "#56B4E9", # light blue
   "1" = "#009E73", # green
   "2" = "#F0E442", # yellow
   "3" = "#0072B2", # dark blue
   "4" = "#D55E00", # red
+  "5" = "#56B4E9", # light blue
+  "6" = "#E69F00", # goldenrod
   "7" = "#CC79A7", # magenta
-  "8" ="#000000") # black
+  "8" ="#000000", # black
+  "Other" = "grey80") 
 
 # - Protected areas (Okabe-Ito CVD safe)
 protection_cols <- c(
@@ -256,7 +257,7 @@ map_cluster_to_nodes <- function(dendro, tax_clusters, cluster_select) {
     node_mrca <- ape::getMRCA(dendro, tips)
     nodes <- phytools::getDescendants(dendro, node_mrca)
   } else {
-    stop("Can't find that cluster")
+    stop(glue::glue("Can't find cluster '{cluster_select}'"))
   }
   
   # Return a tibble mapping nodes to the cluster

### 2021-08-31 ms/manuscript.Rmd
Change names of some raw data files

@@
   "japan_fern_tree",
-  "japan_points_raw",
+  "japan_map_points",
   "japan_pterido_tree",
@@
 japan_points <- 
-  japan_points_raw %>%
+  japan_map_points %>%
   filter(is.na(lat_end)) %>%
@@
 
-japan_lines <- japan_points_raw %>%
+japan_lines <- japan_map_points %>%
   filter(!is.na(lat_end))

### 2021-08-31 R/process_raw_data.R

@@ -112,13 +112,13 @@ tar_plan(
   redundancy_by_res = calc_redundancy_by_res(comm_scaled_list),
   tar_file(
     redundancy_by_res_out,
-    write_csv_tar(redundancy_by_res, "data/redundancy_by_res.csv")),
+    write_csv_tar(redundancy_by_res, "data/japan_ferns_redundancy_by_res.csv")),
   
   # Assess sampling completeness with iNEXT
   inext_res = run_inext_on_ferns(occ_point_data_ferns),
   tar_file(
     inext_res_out,
-    write_csv_tar(inext_res, "data/inext_results.csv")),
+    write_csv_tar(inext_res, "data/japan_ferns_inext_results.csv")),
   
   # Clean raw lucid data (remove data in Japanese, correct names)
   tar_target(lucid_data_raw_file, "data_raw/JpFernLucid_forJoel20200827.xlsx"),

### 2021-08-31 _targets.R

@@ -55,8 +55,8 @@ tar_plan(
     summarize(),
   
   # Load manually entered points of interest for drawing a map of Japan
-  tar_file(japan_points_raw_file, "data/japan_points_raw.csv"),
-  japan_points_raw = read_csv(japan_points_raw_file),
+  tar_file(japan_map_points_file, "data/japan_map_points.csv"),
+  japan_map_points = read_csv(japan_map_points_file),
   
   # Calculate area as rolling mean in 1 degree latitudinal windows
   lat_area_ja = calc_area_by_lat(japan_shp, lat_cut = 0.2, lat_window = 1),
@@ -73,11 +73,11 @@ tar_plan(
   comm_ferns_full = load_jferns_comm(japan_ferns_comm_full_file),
   
   # Load summaries of testing different grid cell sizes
-  tar_file(redundancy_by_res_file, "data/redundancy_by_res.csv"),
+  tar_file(redundancy_by_res_file, "data/japan_ferns_redundancy_by_res.csv"),
   redundancy_by_res = read_csv(redundancy_by_res_file, col_types = cols(res = col_character())),
   
   # Load results of assessing sampling completeness with iNEXT
-  tar_file(inext_res_file, "data/inext_results.csv"),
+  tar_file(inext_res_file, "data/japan_ferns_inext_results.csv"),
   inext_res = read_csv(inext_res_file),
   
   # Subset geographic shapes to redundancy > 0.1

### 2021-08-31 ms/manuscript.Rmd
Update MS
Fix number of fossil calibration points

@@
   verify(nrow(.) == 0, success_fun = success_logical)
+
+# Number of fossils used for dating
+n_fossil_points <- plastome_calibration_dates %>%
+  filter(age_type != "fixed") %>%
+  nrow()
 ```
@@
 To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
-We conducted phylogenetic analysis with `r software("IQTREE")` v1.6.12 [@Nguyen2015] and molecular dating with `r software("treePL")` v1.0 [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
+We conducted phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] and molecular dating with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
+The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
 We then trimmed the tree to Japanese taxa only and used this for all further analyses.

### 2021-08-31 ms/data_readme.Rmd
Add data readme Rmd

---
output: rtf_document
editor_options:
  chunk_output_type: console
---

---
bibliography: 
- `r here::here("ms/references.yaml")`
---

```{r manual, include = FALSE, evaluate = FALSE}
# Run these lines if rendering outside of targets plan
library(targets)
library(tarchetypes)
source(here::here("R/packages.R"))
source(here::here("R/functions.R"))
```

```{r setup, include = FALSE, cache = FALSE}
# TO CONVERT TO PLAIN TEXT:
# after rendering to rtf,
# 1. Copy the contents of the rtf file, paste to txt.
# 2. Replace all single line breaks with double line breaks
# (in Atom search-replace with grep: \n for \n\n)
# 3. Hard-wrap all text at 80 characters (in Atom with autoflow: cmd-alt-q)
# (optional: add name after first line)
# 4. Save as "README.txt"

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE,
  warning = FALSE, results = "hide", cache = FALSE)

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "japan_ferns_comm_full_file",
    "plastome_calibration_dates")
  )
```

This README.txt file was generated on `r format(Sys.time(), '%d %B, %Y')`

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\- GENERAL INFORMATION \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

Title of Dataset: Data from: Spatial phylogenetics of Japanese ferns: Patterns,
processes, and implications for conservation

Author Information

Principal Investigator: Joel H. Nitta

Department of Biological Sciences, Graduate School of Science, The University of
Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo 113-0032, Japan joelnitta@gmail.com

Associate or Co-investigators: Brent D. Mishler, Wataru Iwasaki, Atsushi Ebihara

Date of data collection: 2012--2014

Geographic location of data collection: Japan

Information about funding sources or sponsorship that supported the collection
of the data: Funding provided in part by the Japan Society for the Promotion of
Science (Kakenhi) Grant Number 16H06279

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

SHARING/ACCESS INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

Licenses/restrictions placed on the data, or limitations of reuse: CC0 1.0
Universal (CC0 1.0)

Recommended citation for the data: Nitta JH, Mishler BD, Iwasaki W, Ebihara A
(2021) Data from: Spatial phylogenetics of Japanese ferns: Patterns, processes,
and implications for conservation FIXME: add DOI

Citation for and links to publications that cite or use the data: Nitta JH,
Mishler BD, Iwasaki W, Ebihara A (2021) Spatial phylogenetics of Japanese ferns:
Patterns, processes, and implications for conservation <<FIXME: ADD JOURNAL>>

Code for analyzing the data is available on github:
https://github.com/joelnitta/japan_ferns_biogeo

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

DATA & FILE OVERVIEW

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

File list (filenames, directory structure (for zipped files) and brief
description of all data files):

- japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of ferns in Japan
- japan_ferns_inext_results.csv: Results of collection curve analysis on fern specimens in Japan
- japan_ferns_occ_summary.csv: Summary of specimen data of ferns in Japan
- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for ferns of Japan
- japan_ferns_shape_full.gpkg: Location of grid cells (sites) for ferns in Japan
- japan_ferns_traits_lucid.csv: Morphological trait data of ferns in Japan
- japan_map_points.csv: Latitude and longitude of points used to produce a map of Japan
- japan_protected_areas.gpkg: Protected areas in Japan
- testo_sundue_2016_calibrations.csv: Fossil calibration points used for phylogenetic dating analysis
- results.zip: Analysis results that may be useful in future studies. Files include:
  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of ferns in Japan
  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of ferns in Japan
  - japan_ferns_traits.csv: Trait matrix of ferns in Japan used for functional biodiversity analysis
  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of ferns in Japan used for biodiversity analysis
  - japan_ferns_biodiv.gpkg: Biodiversity statistics of ferns in Japan

Checksums are 32-byte MD5 hashes generated with tools::md5sum() in R.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

METHODOLOGICAL INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r summary-stats}
# Number of fossils used for dating
n_fossil_points <- plastome_calibration_dates %>%
  filter(age_type != "fixed") %>%
  nrow()
```

Description of methods used for collection/generation of data:

A list of native, non-hybrid fern specimens housed at the herbarium of the
Museum of Science and Nature, Japan was converted to a community data matrix at
four grain sizes (grid cells spanning Japan, each 0.1, 0.2, 0.3, or 0.4 degrees
per side). The original list with latitude and longitude of each specimen is not
included here to protect rare and endangered species. The 0.2 degree grain size
was selected for further analysis based on redundancy (ratio of number of
specimens to number of taxa per cell).

All taxon names are based on the Green List (http://www.rdplants.org/gl/;
English version available at
https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32)

Traits were measured on each species as described in [@Ebihara2019b].

Phylogenetic analysis was conducted with maximum likelihood in IQ-TREE
[@Nguyen2015] by combining plastid *rbcL* sequences of each taxon with a
globally sampled data matrix (Nitta et al, in prep). Next, dating analysis was
carried out using treePL v1.0 [@Smith2012] with `r n_fossil_points` fossil
calibration points after @Testo2016a. The dated phylogeny was then trimmed to
include Japanese taxa only.

The community matrix, traits, and phylogeny were used to analyze spatial
patterns of phylogenetic diversity and endemism.

Data files were generated from raw data (not included here) using scripts
available at https://github.com/joelnitta/japan_ferns_biogeo.

For full methods, see Nitta JH, Mishler BD, Iwasaki W, Ebihara A (2021) Spatial
phylogenetics of Japanese ferns: Patterns, processes, and implications for
conservation <<FIXME: ADD JOURNAL>>

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

DATA-SPECIFIC INFORMATION

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

```{r japan-ferns-comm-full}
japan_ferns_comm_full <- read_csv(here::here(japan_ferns_comm_full_file))
# Check that no data are missing
assert(japan_ferns_comm_full, not_na, everything(), success_fun = success_logical)
```

japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of ferns in
Japan. Sites correspond to 0.2 degree x 0.2 degree grid cells covering Japan.
For location of grid cells, see japan_ferns_shape_full.gpkg. Full dataset
including all grid cells (none excluded due to low redundancy).

Number of variables: `r ncol(japan_ferns_comm_full)`

Number of cases/rows: `r nrow(japan_ferns_comm_full)`

Variable list:

- grids: Name of grid cell.
- Other columns: Each column is named for a Japanese fern taxon. 
  Values indicate number of specimens that were observed in each grid cell for that taxon.

Missing data codes: No missing data.

MD5 checksum: `r tools::md5sum(here::here(japan_ferns_comm_full_file))`

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

CHANGE LOG

\-\-\-

2021-08-31

Generate this README file.

### 2021-08-31 _targets.R

@@ -574,5 +574,11 @@ tar_plan(
     output_format = "rmarkdown::html_document",
     output_file = here::here("results/data_exploration.html"),
     params = list(knit_type = "targets")
+  ),
+  tar_render(
+    data_readme,
+    knit_root_dir = here::here(),
+    path = "ms/data_readme.Rmd",
+    output_file = here::here("results/data_readme.txt")
   )
 )

### 2021-09-01 ms/manuscript.Rmd
Add deer damage analysis to MS

@@
   "csl_file",
+  "deer_range",
   "fern_traits",
@@
   "signif_cells_protected_area",
+  "signif_cells_deer_danger",
   "template_file",
@@
 We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. 
-We compared the overlap of significantly diverse grid-cells with existing conservation zones to assess conservation status.
+We compared the overlap of significantly diverse grid-cells with existing conservation zones and range maps of native Japanese deer to assess conservation status and threats.
 
@@
 Temperature and apomixis were identified as likely drivers of patterns of taxonomic and phylogenetic diversity.
-Conservation status (% protected area) is generally higher than the overall rate for grid-cells with significantly high biodiversity.
+Conservation status (% protected area) was generally high for grid-cells with significantly high biodiversity across all biodiversity metrics, but the threat due to herbivory by native Japanese deer varied by biodiversity metric.
 
@@
 The different distributions of each diversity type reflect the major bioregions. 
-Although peak areas of different diversity types are largely non-overlapping, each one has a relatively high protection status. 
-Despite this, future conservation efforts must be cognizant of dangers not addressed here such as climate change and herbivory by deer.
+Although peak areas of different diversity types are largely non-overlapping, each one has relatively high protection coverage at medium protection level. 
+Despite this, future conservation efforts must recognize that threats can vary by biodiversity metric, and consider threats not directly analyzed here such as climate change.
 
 <!-- may include title words -->
-**Key words:** apomixis, biodiversity, biogeography, CANAPE, ferns, Japan, phylogenetic diversity.
+**Key words:** apomixis, biodiversity, biogeography, conservation, CANAPE, ferns, Japan, phylogenetic diversity.
 
@@
 
-### Assessment of conservation status
-
-We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html). 
-We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" (some economic activities allowed by permit) following [@Kusumoto2017].
-We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
+### Assessment of conservation status and threats
+
+We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html) and Ministry of Land, Infrastructure, Transport and Tourism (https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-A45.html).
+We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" status (some economic activities allowed by permit) [@Kusumoto2017].
+Any areas that overlapped within a protection status were combined, and areas that overlapped between medium and high status were considered only high status.
+We downloaded distribution maps (SHP and TIF) of native Japanese deer (*Cervus nippon*), an herbivore that has caused extensive damage to native plant communities due to rapid population growth since the 1970s [@Takatsuki2009] and poses a high risk to ferns, from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
+To assess change in threat due to herbivory by deer over time, we used three maps available in this dataset: range of Japanese deer surveyed in 1978, range surveyed in 2003, and estimated range inferred from a model including snow cover and forest type based on the 2003 survey data [@kankyosho2021].
+We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type or percent of area occupied by deer for each survey period within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
 For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above. 
@@
 
-### Conservation status
+### Conservation status and threats
 
@@
 total_percent_protected_all <- sum(total_percent_protected_medium, total_percent_protected_high)
-```
-
-The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`).
-Areas with high PE were the best protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total). 
-Cells with significantly high biodiversity generally had a greater percentage of area protected than the overall rate for Japan (`r scales::percent(total_percent_protected_high, accuracy = 0.1)` high, `r scales::percent(total_percent_protected_medium, accuracy = 0.1)` medium, `r scales::percent(total_percent_protected_all, accuracy = 0.1)` total); only FD was lower than the overall rate for highly protected areas.
+
+### Deer ###
+
+# Calculate baseline protection percentage: percent protected areas across the whole country
+total_percent_deer_danger <-
+  deer_range %>%
+  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
+  st_drop_geometry %>%
+  as_tibble() %>%
+  group_by(range) %>%
+  summarize(
+    total_area = sum(area),
+    .groups = "drop"
+  ) %>%
+  mutate(percent_danger = total_area/area_of_japan) %>%
+  mutate(percent_danger = as.numeric(percent_danger))
+
+# Convert to a list for easy indexing
+deer_overall <- total_percent_deer_danger %>%
+  mutate(
+    # Format numbers for nice printing
+    area = total_area %>% as.numeric %>% scales::number(accuracy = 1, big.mark = ",") %>% paste("km^2^"),
+    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
+  split(.$range)
+
+deer_in_high_biodiv <-
+signif_cells_deer_danger %>%
+  group_by(range) %>%
+  summarize(percent_danger = mean(percent_danger), .groups = "drop") %>%
+  mutate(
+    # Format numbers for nice printing
+    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
+  split(.$range)
+
+deer_highest_danger <- signif_cells_deer_danger %>%
+  slice_max(percent_danger) %>%
+  # Make sure biodiv type with greatest danger is richness in estimated dataset
+  verify(metric == "richness") %>%
+  verify(range == "estimated") %>%
+  mutate(
+    # Format numbers for nice printing
+    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
+  split(.$range)
+```
+
+The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`a).
+Cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
+In contrast, cells with significantly high biodiversity had a lower percentage of area protected than the overall rate for Japan at the high protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) for most biodiversity metrics.
+Only cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the overall rate at both levels of protection.
+
+The overall range of Japanese deer in Japan has increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
+The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
+Cells with high FD, PD, or PE tended to have less deer occupation compared to the overall rate (`r figure("conserv-status")`b).
+However, cells with high species richness consistently exceeded the overall rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.
 
@@
 
-### Conservation status
-
-Although the different facets of biodiversity have different distribution patterns, we found that grid-cells with significantly high amounts of biodiversity tend to be relatively well protected compared to the baseline protection rate, regardless of biodiversity metric (`r figure("conserv-status")`).
-One possible reason is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas (A. Ebihara, pers. comm.). 
-Another (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
+### Conservation status and threats
+
+We found that grid-cells with significantly high amounts of biodiversity are better protected than the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`a).
+One possible reason for this generally high level of protection is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas (A.
+Ebihara, pers.
+comm.).
+In particular, there has been an emphasis on conserving rare and endemic species in Japanese southern islands such as Ogasawara and Amami (FIXME: add ref), which include many grid-cells with high PE.
+Another reason (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
 
@@
 One other such threat is herbivory by native Japanese deer, *Cervus nippon*. 
-Populations of these deer have increased rapidly since the 1970s due to a combination of climate change and a decrease in human hunters and natural predators [@Takatsuki2009]. 
+The rapid increase in population size observed in surveys from 1978 to 2003 is likely due to a combination of climate change and a decrease in human hunters and natural predators [@Takatsuki2009]. 
 At dense populations, the deer can prevent forest regeneration and result in plant communities dominated by a limited number of species unpalatable to deer [@Takatsuki2009], although lower population sizes may promote plant diversity as per the intermediate disturbance hypothesis [@Nishizawa2016].
@@
 Although various efforts to prevent deer herbivory have been deployed such as culling and exclusion fences [@Takatsuki2009], much work remains to be done to save vulnerable plant populations.
+
+Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`b).
+This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, where deer expansion has been intense (FIXME: add SI fig).
+This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.
 
@@
 # Plot protected areas
-conserv_status_plot <- signif_cells_protected_area %>%
+a <- signif_cells_protected_area %>%
   mutate(
@@
   geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
-  scale_fill_manual(values = protection_cols) +
-  scale_x_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
+  scale_fill_manual(
+    values = protection_cols,
+    guide = guide_legend(
+      direction = "horizontal",
+      title.position = "top",
+      label.position = "left"
+    )) +
+  scale_x_continuous(
+    labels = function(x) scales::percent(x, accuracy = 1),
+    expand = expansion(mult = c(0, .05))) +
   labs(
@@
   )
+
+# Plot areas in danger from herbivory by deer 
+b <- signif_cells_deer_danger %>%
+  mutate(
+    metric = case_when(
+      metric == "pd" ~ "PD",
+      metric == "fd" ~ "FD",
+      metric == "pe" ~ "PE",
+      metric == "richness" ~ "Richness",
+    ),
+    range = str_to_sentence(range)) %>%
+  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "Richness"))) %>%
+  ggplot(aes(x = percent_danger, y = metric, fill = range)) +
+  geom_col(position = position_dodge()) +
+  # line for 1978 overall percent
+  geom_vline(
+    xintercept = total_percent_deer_danger %>% filter(range == "1978") %>% pull(percent_danger), 
+    color = "grey40", linetype = 3) +
+  # line for 2003 overall percent
+  geom_vline(
+    xintercept = total_percent_deer_danger %>% filter(range == "2003") %>% pull(percent_danger), 
+    color = "grey40", linetype = 2) +
+  # line for estimated overall percent
+  geom_vline(
+    xintercept = total_percent_deer_danger %>% filter(range == "estimated") %>% pull(percent_danger), 
+    color = "grey40", linetype = 1) +
+  scale_fill_manual(
+    values = deer_colors,
+    guide = guide_legend(
+      direction = "horizontal",
+      title.position = "top",
+      label.position = "left"
+    )) +
+  scale_x_continuous(
+    labels = function(x) scales::percent(x, accuracy = 1),
+    expand = expansion(mult = c(0, .05))) +
+  labs(
+    x = "Percent area with deer",
+    y = "Biodiversity metric",
+    fill = "Range data source"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.title.y = element_blank(),
+    legend.position = "bottom"
+  )
+
+conserv_status_plot <- a + b +
+  plot_layout(ncol = 2) +
+  plot_annotation(
+    tag_levels = "a",
+    tag_prefix = "(",
+    tag_suffix = ")"
+  ) &
+  theme(plot.tag = element_text(face = "bold"))
 ```
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 4, height = 4, units = "in")
-```
-
-```{r conserv-status-print, fig.width = 4, fig.height = 4, eval = params$doc_type == "pdf"}
+ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 7, height = 4, units = "in")
+```
+
+```{r conserv-status-print, fig.width = 7, fig.height = 4, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
 **`r figure("conserv-status")`**.
-Percent of land area with protected status for grid cells with significantly high biodiversity by at least one metric.
+Conservation status and threat due to herbivory by deer for ferns of Japan by biodiversity metric.
+(a) Percent of land area with protected status for grid cells with significantly high biodiversity.
+(b) Percent of land area occupied by deer for grid cells with significantly high biodiversity.
 Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
 Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Materials and Methods).
-Vertical lines indicate total percent area protected across Japan regardless of degree of biodiversity; dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).
+For (a), vertical lines indicate total percent area protected across Japan (baseline protection rate); dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).
+For (b), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate); solid, percentage estimated with a model based on 2003 survey data (`r deer_overall$estimated$percent`); dashed, percentage based on 2003 survey data (`r deer_overall[["2003"]]$percent`); dotted, percentage based on 1978 survey data (`r deer_overall[["1978"]]$percent`).
 

### 2021-09-01 R/functions.R

@@ -2386,11 +2386,11 @@ calculate_percent_deer_danger <- function(deer_danger_biodiv) {
     mutate(metric = str_remove_all(metric, "_obs_p_upper")) %>%
     group_by(range, metric) %>%
     summarize(
-      protected_area = sum(area),
+      danger_area = sum(area),
       .groups = "drop"
     ) %>%
     left_join(signif_cells_total_area, by = "metric") %>%
-    mutate(percent_danger = (protected_area / total_area)) %>%
+    mutate(percent_danger = (danger_area / total_area)) %>%
     mutate(percent_danger = as.numeric(percent_danger))
 }
 

### 2021-09-02 ms/manuscript.Rmd
Update MS, take renv snapshot Improve conservation plots

@@
 We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown. 
-We calculated the percentage of apomictic species in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`b).
+We calculated the percentage of apomictic species in each grid-cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`b).
 
@@
 We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
-This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
+This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
 The R package `r pack("sf")` was used for GIS analyses [@Pebesma2018].
@@
 Richness was fit with the negative binomial response family; others were Gaussian. 
-We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014]. 
+We fit models including a Matérn correlation matrix based on the geographic centroids of each grid-cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014]. 
 We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence. 
@@
 Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (`r figure("bioregions")`b; `r s_figure("dendrogram")`b). 
-The other much smaller bioregions (each including only three grid cells or less) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
+The other much smaller bioregions (each including only three grid-cells or less) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
 
@@
   mutate(percent_protected = total_area/area_of_japan) %>%
-  mutate(percent_protected = as.numeric(percent_protected))
+  mutate(percent_protected = as.numeric(percent_protected)) %>%
+  # Add sum for high + medium
+  bind_rows(
+    summarize(
+      ., 
+      percent_protected = sum(percent_protected),
+      total_area = sum(total_area)
+    ) %>% mutate(status = "total")
+  )
 
@@
 (b), phylogenetic bioregions.
-Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid cells.
-Bioregions with three or fewer grid cells are lumped into the "Other" category.
+Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells.
+Bioregions with three or fewer grid-cells are lumped into the "Other" category.
 
@@
 ```{r conserv-status-make}
+# Add total protected area (medium + high)
+signif_cells_protected_area_total <-
+  signif_cells_protected_area %>%
+  group_by(metric) %>%
+  summarize(percent_protected = sum(percent_protected), .groups = "drop") %>%
+  mutate(status = "total")
+
+total_percent_protected_for_plot <- 
+  total_percent_protected %>%
+  select(status, percent_protected) %>%
+  mutate(status = str_to_sentence(status))
+
 # Plot protected areas
 a <- signif_cells_protected_area %>%
+  select(metric, status, percent_protected) %>%
+  bind_rows(signif_cells_protected_area_total) %>%
   mutate(
@@
   geom_col(position = position_dodge()) +
-  geom_vline(xintercept = total_percent_protected_medium, color = "grey40", linetype = 3) +
-  geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
+  geom_vline(
+    data = total_percent_protected_for_plot,
+    aes(xintercept = percent_protected, linetype = status),
+    color = "grey30"
+  ) +
+  scale_linetype_manual(
+    values = protection_lines,
+    guide = guide_legend(
+      direction = "horizontal",
+      title.position = "top",
+      label.position = "left"
+    )) +
   scale_fill_manual(
@@
     y = "Biodiversity metric",
-    fill = "Protection status"
+    fill = "Protection status",
+    linetype = "Protection status"
   ) +
@@
   geom_col(position = position_dodge()) +
-  # line for 1978 overall percent
   geom_vline(
-    xintercept = total_percent_deer_danger %>% filter(range == "1978") %>% pull(percent_danger), 
-    color = "grey40", linetype = 3) +
-  # line for 2003 overall percent
-  geom_vline(
-    xintercept = total_percent_deer_danger %>% filter(range == "2003") %>% pull(percent_danger), 
-    color = "grey40", linetype = 2) +
-  # line for estimated overall percent
-  geom_vline(
-    xintercept = total_percent_deer_danger %>% filter(range == "estimated") %>% pull(percent_danger), 
-    color = "grey40", linetype = 1) +
+    data = total_percent_deer_danger %>% mutate(range = str_to_sentence(range)),
+    aes(linetype = range, xintercept = percent_danger),
+    color = "grey30"
+  ) +
+  scale_linetype_manual(
+    values = deer_lines,
+    guide = guide_legend(
+      direction = "horizontal",
+      title.position = "top",
+      label.position = "left"
+    )) +
   scale_fill_manual(
-    values = deer_colors,
+    values = deer_cols,
     guide = guide_legend(
@@
     y = "Biodiversity metric",
-    fill = "Range data source"
+    fill = "Range data source",
+    linetype = "Range data source"
   ) +
@@
 **`r figure("conserv-status")`**.
-Conservation status and threat due to herbivory by deer for ferns of Japan by biodiversity metric.
-(a) Percent of land area with protected status for grid cells with significantly high biodiversity.
-(b) Percent of land area occupied by deer for grid cells with significantly high biodiversity.
-Biodiversity metrics include phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
-Significance of biodiversity metrics assessed by a one-tailed test comparing observed values to a null distribution of 999 random values (see Materials and Methods).
-For (a), vertical lines indicate total percent area protected across Japan (baseline protection rate); dashed, high protection level (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`); dotted, medium protection level (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`).
-For (b), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate); solid, percentage estimated with a model based on 2003 survey data (`r deer_overall$estimated$percent`); dashed, percentage based on 2003 survey data (`r deer_overall[["2003"]]$percent`); dotted, percentage based on 1978 survey data (`r deer_overall[["1978"]]$percent`).
+Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan.
+(a) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status.
+(b) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
+Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
+Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
+For (a), vertical lines indicate total percent area protected across Japan (baseline protection rate).
+For (b), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).
 

### 2021-09-02 R/ms_functions.R

@@ -43,16 +43,28 @@ bioregion_cols <- c(
 
 # - Protected areas (Okabe-Ito CVD safe)
 protection_cols <- c(
-  "Medium" = "#E69F00", # goldenrod
-  "High" = "#0072B2" # dark blue
+  "Medium" = "#F0E442", # yellow
+  "High" = "#009E73", # green
+  "Total" = "#56B4E9" # light blue
+)
+
+protection_lines <- c(
+  "Medium" = "dashed", 
+  "High" = "dotted",
+  "Total" = "solid"
 )
 
 # - Deer range colors (Okabe-Ito CVD safe)
-deer_colors <- c( 
+deer_cols <- c( 
   "1978" = "#0072B2", # dark blue
   "2003" = "#F0E442", # yellow
-  "Estimated" = "#D55E00", # red
-  "estimated" = "#D55E00"# red
+  "Estimated" = "#D55E00" # red
+) 
+
+deer_lines <- c( 
+  "1978" = "dotted", # dark blue
+  "2003" = "dashed", # yellow
+  "Estimated" = "solid" # red
 ) 
 
 # Words ----

### 2021-09-02 ms/SI.Rmd
Fix dendrogram for phylo clusters

@@
   select(grids, cluster = taxonomic_cluster) %>%
-  # mutate(cluster = as.character(cluster)) # %>%
   # Convert cluster to numeric to use max() later
@@
 
-# Map taxonomic clusters to grid cells
+# Map phylogenetic clusters to grid cells
 clusters_phy <-
-  biodiv_ferns_spatial_lumped %>%
+  biodiv_ferns_spatial %>%
   as_tibble() %>%
@@
   # Convert cluster to numeric to use max() later
-  mutate(cluster = levels(cluster)[cluster] %>% as.numeric)
+  mutate(cluster = levels(cluster)[cluster])
 
-cluster_nodes_phy_tbl <- map_df(1:max(clusters_phy$cluster), ~map_cluster_to_nodes(dendro_phy, clusters_phy, .)) %>%
-  mutate(cluster = as.factor(cluster))
+cluster_nodes_phy_tbl <- map_df(unique(clusters_phy$cluster), ~map_cluster_to_nodes(dendro_phy, clusters_phy, .)) %>%
+  # Lump regions with just a few cells (3 or less)
+  mutate(cluster = as.factor(cluster) %>% fct_lump_min(4))
 
@@
   aes(color = cluster) +
-  scale_color_manual(values = bioregion_cols, na.value = "grey", name = "Cluster") +
+  scale_color_manual(
+    values = bioregion_cols[c(1,2,3,4,"Other")], breaks = c(1,2,3,4,"Other"), name = "Cluster", na.value = "grey10") +
   theme(legend.position = "right") +

### 2021-09-02 ms/SI.Rmd
Change latitudinal area var to "lat_area"
from just "area", so as not to confuse with "area" of protected zones

@@
 
-a <- plot_env("a", "Area", "area", "km^2")
+a <- plot_env("a", "Area", "lat_area", "km^2")
 b <- plot_env("b", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))

### 2021-09-02 R/functions.R

@@ -2146,11 +2146,6 @@ cluster_phylo_regions <- function (comm_df, phy, k) {
 #' data of ferns in Japan cropped to areas with medium protection and filtered to only
 #' areas with significantly high biodiversity
 crop_by_pa <- function(protected_areas, biodiv_ferns_spatial, japan_shp) {
-  ### Manipulate biodiv_ferns_spatial ###
-  # Add percent rank for richness
-  biodiv_ferns_spatial <- mutate(biodiv_ferns_spatial, richness_obs_p_upper = dplyr::percent_rank(richness)) %>%
-    # drop latitudinal area (so we can join with protected area)
-    select(-area) 
   
   ### Prepare for crop/join steps ###
   # Crop to Japan to spatial div results area
@@ -2220,13 +2215,7 @@ crop_by_pa <- function(protected_areas, biodiv_ferns_spatial, japan_shp) {
 #' data of ferns in Japan cropped to areas with deer present in 1973 and filtered to only
 #' areas with significantly high biodiversity
 crop_by_deer <- function(deer_range, biodiv_ferns_spatial, japan_shp) {
-  
-  ### Manipulate biodiv_ferns_spatial ###
-  # Add percent rank for richness
-  biodiv_ferns_spatial <- mutate(biodiv_ferns_spatial, richness_obs_p_upper = dplyr::percent_rank(richness)) %>%
-    # drop latitudinal area (so we can join with protected area)
-    select(-area) 
-  
+    
   ### Prepare for crop/join steps ###
   # Crop to Japan to spatial div results area
   japan_shp <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
@@ -3078,11 +3067,11 @@ prepare_data_for_spamm <- function(
   bind_rows(
     tibble(
       resp_var = resp_var_env,
-      formula = glue("{resp_var_env} ~ temp + I(temp^2) + precip + precip_season + area + Matern(1|long+lat)")
+      formula = glue("{resp_var_env} ~ temp + I(temp^2) + precip + precip_season + lat_area + Matern(1|long+lat)")
     ),
     tibble(
       resp_var = resp_var_repro,
-      formula = glue("{resp_var_repro} ~ percent_apo + precip + precip_season + area + Matern(1|long+lat)")
[…]

### 2021-09-02 R/ms_functions.R

@@ -252,7 +252,7 @@ clean_vars <- function(x) {
     str_replace_all("^temp_season$", "Temperature seasonality") %>%
     str_replace_all("^precip_season$", "Precipitation seasonality") %>%
     str_replace_all("^precip$", "Precipitation") %>%
-    str_replace_all("^area$", "Area") %>%
+    str_replace_all("^lat_area$", "Area") %>%
     str_replace_all("^richness$", "Richness")
 }
 

### 2021-09-02 _targets.R

@@ -353,6 +353,8 @@ tar_plan(
     classify_signif("fd") %>%
     classify_signif("rfd") %>%
     classify_signif("pe", one_sided = TRUE, upper = TRUE) %>%
+    # Add richness percent rank
+    mutate(richness_obs_p_upper = dplyr::percent_rank(richness)) %>%
     # Format factors
     mutate(taxonomic_cluster = as.factor(taxonomic_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
     mutate(phylo_cluster = as.factor(phylo_cluster) %>% fct_infreq %>% as.numeric %>% as.factor) %>%
@@ -392,7 +394,7 @@ tar_plan(
   # - response variables for reproductive model
   resp_vars_repro = c("pd_obs_z", "rpd_obs_z"),
   # - independent variables (both models)
-  indep_vars = c("percent_apo", "temp", "precip", "precip_season", "area"),
+  indep_vars = c("percent_apo", "temp", "precip", "precip_season", "lat_area"),
   
   # Make biodiversity metrics dataframe with centroid of each site for models
   # - ultrametric tree (full  analysis)
@@ -409,7 +411,7 @@ tar_plan(
   # Check for correlation between independent variables in repro data
   t_test_results = run_mod_ttest_ja(
     st_set_geometry(biodiv_ferns_spatial, NULL),
-    vars_select = c("temp", "temp_season", "precip", "precip_season", "percent_apo", "area")
+    vars_select = c("temp", "temp_season", "precip", "precip_season", "percent_apo", "lat_area")
   ),
   
   ## Analyze Moran's I ----

### 2021-09-02 ms/data_readme.Rmd
Update data readme

@@
 source(here::here("R/functions.R"))
+source(here::here("R/ms_functions.R"))
 ```
@@
   c(
+    "ms_functions",
     "japan_ferns_comm_full_file",
-    "plastome_calibration_dates")
+    "inext_res_file",
+    "occ_point_data_summary_file",
+    "plastome_calibration_dates",
+    "redundancy_by_res_file",
+    "japan_ferns_shape_full_file"
+    )
   )
@@
 (2021) Data from: Spatial phylogenetics of Japanese ferns: Patterns, processes,
-and implications for conservation FIXME: add DOI
+and implications for conservation FIXME: add DOI when available
 
@@
 Mishler BD, Iwasaki W, Ebihara A (2021) Spatial phylogenetics of Japanese ferns:
-Patterns, processes, and implications for conservation <<FIXME: ADD JOURNAL>>
+Patterns, processes, and implications for conservation FIXME: add journal when published
 
@@
 
-Traits were measured on each species as described in [@Ebihara2019b].
+Traits were measured on each species as described in @Ebihara2019b.
 
@@
 phylogenetics of Japanese ferns: Patterns, processes, and implications for
-conservation <<FIXME: ADD JOURNAL>>
+conservation FIXME: Add journal when published
 
@@
 
-japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of ferns in
-Japan. Sites correspond to 0.2 degree x 0.2 degree grid cells covering Japan.
-For location of grid cells, see japan_ferns_shape_full.gpkg. Full dataset
-including all grid cells (none excluded due to low redundancy).
+`r fs::path_file(japan_ferns_comm_full_file)`: Community matrix (species x sites
+matrix) of ferns in Japan. Sites correspond to 0.2 degree x 0.2 degree grid
+cells covering Japan. For location of grid cells, see
+japan_ferns_shape_full.gpkg. Full dataset including all grid cells (none
+excluded due to low redundancy).
 
@@
 
+```{r japan_ferns_inext_results}
+inext_res <- read_csv(here::here(inext_res_file))
+# Check that no data are missing
+assert(inext_res, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(inext_res_file)`: Output from running iNEXT [@Hsieh2016] on
+specimen data for ferns of Japan. iNEXT estimates diversity (here, species
+richness) and sampling completeness based on interpolation and extrapolation
+from the original data.
+
+Number of variables: `r ncol(inext_res)`
+
+Number of cases/rows: `r nrow(inext_res)`
+
+Variable list:
+
+```{r results = "asis"}
+  inext_res %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "m", "Sample size (number of individuals)",
+      "method", "Method used to estimate diveristy",
+      "order", "Diversity order of Hill number (here 0, equivalent to species richness)",
+      "qD", "Estimated diversity",
+      "qD.LCL", "Lower 95% confidence limit on estimated diversity",
+      "qD.UCL", "Upper 95% confidence limit on estimated diversity",
+      "SC", "Estimated sampling completeness",
+      "SC.LCL", "Lower 95% confidence limit on sampling completeness",
+      "SC.UCL", "Upper 95% confidence limit on sampling completeness")
+  )
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(japan_ferns_comm_full_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r occ_point_data_summary}
+occ_point_data_summary <- read_csv(here::here(occ_point_data_summary_file))
+# Check that no data are missing
+assert(occ_point_data_summary, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(occ_point_data_summary_file)`: Summary of raw specimen data for
+native, non-hybrid ferns of Japan. Raw specimen data consists of one row per
+specimen, with species name, latitude, longitude, and collection date (raw
+specimen data not included here to protect endangered species). Raw data were
+filtered to remove duplicate specimens and specimens that occurred outside of
+the second-degree mesh (approximately 10 km sq grid cell) map of Japan
+(http://www.biodic.go.jp/trialSystem/EN/shpddl.html)
+
+Number of variables: `r ncol(occ_point_data_summary)`
+
+Number of cases/rows: `r nrow(occ_point_data_summary)`
+
+Variable list:
+
+```{r results = "asis"}
+  occ_point_data_summary %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "dataset", "Dataset name. 'occ_point_data_ferns_unfiltered' indicates data includes all specimens; 'occ_point_data_ferns' indicates data only includes specimens after filtering",
+      "variable", "Variable. 'n_taxa' indicates number of taxa, 'n_specimens' indicates number of specimens",
+      "value", "Value corresponding to each variable")
+  )
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(occ_point_data_summary_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r redundancy_by_res}
+redundancy_by_res <- read_csv(here::here(redundancy_by_res_file))
+# Check that no data are missing
+assert(redundancy_by_res, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(redundancy_by_res_file)`: Results of testing binning raw
+specimen data of native, non-hybrid ferns of Japan into grid-cells at different
+resolutions. Sampling redundancy is a metric of sampling completeness,
+calculated as 1 - (richness/abundance).
+
+Number of variables: `r ncol(redundancy_by_res)`
+
+Number of cases/rows: `r nrow(redundancy_by_res)`
+
+Variable list:
+
+```{r results = "asis"}
+  redundancy_by_res %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "res", "Degree of resolution (grid-cell side length), from 0.1 to 0.4 degrees",
+      "grids", "Name of grid-cell",
+      "abundance", "Number of specimens occurring in grid-cell",
+      "richness", "Number of taxa occurring in grid-cell",
+      "redundancy", "Redundancy (measure of sampling completeness)"
+  ))
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(redundancy_by_res_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r japan_ferns_shape_full}
+japan_ferns_shape_full <- st_read(here::here(japan_ferns_shape_full_file))
+# Check that no data are missing
+assert(japan_ferns_shape_full, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(japan_ferns_shape_full_file)`: Spatial data of native,
+non-hybrid ferns in Japan in GeoPackage format. Compiled from raw specimen data
+by binning specimens into 0.2 degree grid cells across Japan.
+
+Number of variables: `r ncol(japan_ferns_shape_full)`
+
+Number of cases/rows: `r nrow(japan_ferns_shape_full)`
+
+Variable list:
+
+```{r results = "asis"}
+  japan_ferns_shape_full %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "grids", "Name of grid-cell",
+      "abundance", "Number of specimens occurring in grid-cell",
+      "richness", "Number of taxa occurring in grid-cell",
+      "redundancy", "Redundancy (measure of sampling completeness)",
+      "geom", "Vector describing shape and position of grid-cell"
+  ))
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(japan_ferns_shape_full_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
 CHANGE LOG

### 2021-09-02 ms/SI.Rmd
Update MS
according to AE comments 2021-08-25

@@
   "model_fits_not_ult",
-  "biodiv_ferns_cent"
+  "biodiv_ferns_cent",
+  "protected_areas",
+  "deer_range"
 ))
@@
 
+```{r protected-map, fig.width = 7, fig.height = 9}
+# Define function to make maps showing overlap of protected areas and grid-cells with high biodiversity
+# These depend on various objects being in the current environment!
+# pass the dots to filter data
+make_protected_plot <- function(part, variable, ...) {
+  ggplot() +
+    geom_sf(data = japan, lwd = 0.10, fill = "white") +
+    geom_sf(data = protected_areas %>% mutate(status = str_to_sentence(status)), aes(fill = status), color = "transparent") +
+    geom_sf(
+      data = biodiv_ferns_spatial %>% filter(...) %>% st_set_crs(japan_crs), 
+      fill = "transparent", color = "black", lwd = 0.20) +
+    annotate(
+      geom="richtext", y = Inf, x = -Inf, 
+      label = glue("<span style='color:black'>**({part})** {variable}</span>"), 
+      hjust = "inward", vjust = "inward", color = "white") +
+    scale_fill_manual(values = protection_cols, name = "Protection status") +
+    scale_x_continuous(breaks = c(125, 135, 145)) +
+    scale_y_continuous(breaks = c(25, 35, 45)) +
+    map_theme()
+}
+
+a <- make_protected_plot("a", "Richness", richness_obs_p_upper > 0.95)
+b <- make_protected_plot("b", "PD", pd_obs_p_upper > 0.95)
+c <- make_protected_plot("c", "FD", fd_obs_p_upper > 0.95)
+d <- make_protected_plot("d", "PE", pe_obs_p_upper > 0.95)
+
+a + b + c + d + plot_layout(ncol = 2, nrow = 2) +
+    plot_layout(guides = "collect") &
+  theme(legend.position = "bottom")
+```
+
+**`r s_figure("protected-map")`**.
+Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.
+Grid-cells with black outlines indicate significantly high biodiversity.
+Biodiversity metrics include (a) taxon richness, (b) phylogenetic diversity (PD), (c) functional diversity (FD), and (d) phylogenetic endemism (PE).
+Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
+
+\clearpage
+
+```{r deer-map, fig.width = 7, fig.height = 9}
+# Define function to make maps showing overlap of areas with deer and grid-cells with high biodiversity
+# These depend on various objects being in the current environment!
+make_deer_plot <- function(part, variable, ...) {
+  ggplot() +
+    geom_sf(data = japan, lwd = 0.10, fill = "white") +
+    geom_sf(data = deer_range %>% filter(range == "estimated") %>% mutate(range = str_to_sentence(range)), aes(fill = range), color = "transparent") +
+    geom_sf(data = deer_range %>% filter(range == "2003"), aes(fill = range), color = "transparent") +
+    geom_sf(data = deer_range %>% filter(range == "1978"), aes(fill = range), color = "transparent") +
+    geom_sf(
+      data = biodiv_ferns_spatial %>% filter(...) %>% st_set_crs(japan_crs), 
+      fill = "transparent", color = "black", lwd = 0.20) +
+    annotate(
+      geom="richtext", y = Inf, x = -Inf, 
+      label = glue("<span style='color:black'>**({part})** {variable}</span>"), 
+      hjust = "inward", vjust = "inward", color = "white") +
+    scale_fill_manual(values = deer_cols, name = "Range data source") +
+    scale_x_continuous(breaks = c(125, 135, 145)) +
+    scale_y_continuous(breaks = c(25, 35, 45)) +
+    map_theme()
+}
+
+a <- make_deer_plot("a", "Richness", richness_obs_p_upper > 0.95)
+b <- make_deer_plot("b", "PD", pd_obs_p_upper > 0.95)
+c <- make_deer_plot("c", "FD", fd_obs_p_upper > 0.95)
+d <- make_deer_plot("d", "PE", pe_obs_p_upper > 0.95)
+
+a + b + c + d + plot_layout(ncol = 2, nrow = 2) +
+    plot_layout(guides = "collect") &
+  theme(legend.position = "bottom")
+```
+
+**`r s_figure("deer-map")`**.
+Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).
+Grid-cells with black outlines indicate significantly high biodiversity.
+Biodiversity metrics include (a) taxon richness, (b) phylogenetic diversity (PD), (c) functional diversity (FD), and (d) phylogenetic endemism (PE).
+Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
+Deer range data sources include 1978 survey, 2003 survey, and range estimated from model based on 2003 survey data.
+Deer ranges plotted in layers in ascending order (estimated, 2003, 1978) such that upper layers may obscure lower layers; however, deer ranges generally expanded from 1978 to estimated range, so very few sites are obscured.
+
+\clearpage
+
 ## References

### 2021-09-02 ms/manuscript.Rmd
Update MS
according to AE comments 2021-08-25

@@
 <!-- may include title words -->
-**Key words:** apomixis, biodiversity, biogeography, conservation, CANAPE, ferns, Japan, phylogenetic diversity.
+**Key words:** apomixis, biodiversity, biogeography, CANAPE, conservation, ferns, Japan, phylogenetic diversity.
 
@@
 We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type or percent of area occupied by deer for each survey period within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
+These values were compared with the overall percent of protected area or area occupied by deer across Japan (baseline rate).
 For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above. 
@@
 The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
-Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* (families follow @PteridophytePhylogenyGroupI2016). 
+Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016]. 
 The age of the fern clade is estimated to be `r phylogeny_age` ma.
@@
 The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`a).
-Cells with significantly high biodiversity had a greater percentage of area protected than the overall rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
-In contrast, cells with significantly high biodiversity had a lower percentage of area protected than the overall rate for Japan at the high protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) for most biodiversity metrics.
-Only cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the overall rate at both levels of protection.
+Cells with significantly high biodiversity had a greater percentage of area protected than the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
+In contrast, cells with significantly high biodiversity had a lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
+Only cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.
 
@@
 The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
-Cells with high FD, PD, or PE tended to have less deer occupation compared to the overall rate (`r figure("conserv-status")`b).
-However, cells with high species richness consistently exceeded the overall rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.
+Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (`r figure("conserv-status")`b).
+However, cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.
 
@@
 Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
+Furthermore, the border between bioregions 1 and 2 corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
 This suggests that ferns are useful as bioindicators in Japan. 
@@
 The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits bioregions 2 and 3, there is some overlap between them on the Amami-Oshima Isl.
-(`r figure("bioregions")`).
+(`r figure("bioregions")`b).
 The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
@@
 We found that grid-cells with significantly high amounts of biodiversity are better protected than the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`a).
-One possible reason for this generally high level of protection is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas (A.
+The high protection status of grid-cells with high PE may be due to an emphasis on conserving rare, endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`d).
+One possible reason for the generally high level of protection overall is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas [A.
 Ebihara, pers.
-comm.).
-In particular, there has been an emphasis on conserving rare and endemic species in Japanese southern islands such as Ogasawara and Amami (FIXME: add ref), which include many grid-cells with high PE.
-Another reason (as suggested previously) is that ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
+comm.; @kankyosho2021b].
+Also, (as suggested previously) ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
 
@@
 Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`b).
-This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, where deer expansion has been intense (FIXME: add SI fig).
+This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, where deer expansion has been intense (`r s_figure("protected-map")`a).
 This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.
@@
   mutate(lat_span = max_lat - min_lat) %>%
-  slice_max(lat_span) %>%
+  slice_max(lat_span, with_ties = FALSE) %>%
   left_join(green_list, by = "taxon") %>%
@@
 The extreme evolutionary distinctiveness of southern islands is both a major feature of the Japanese fern flora and a challenge for analysis. 
-The tropical lineages occurring in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on Honshu (`r figure("rand-diversity")`a). 
-If a more restricted null model was used that did not consider all taxa capable of dispersing anywhere within the study area, different results would likely be obtained for the randomization test [@Mishler2020].
+The tropical lineages occurring only in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on Honshu (`r figure("rand-diversity")`a).
+One way to account for this bimodal distribution pattern would be to use a restricted null model that does not allow all taxa to disperse anywhere within the study area [@Mishler2020]. 
 However, ferns are known to have high dispersal abilities [@Tryon1970; @Barrington1993; @Moran2001]. 
-In our dataset, the taxon with the greatest latitudinal distribution, `r taxon_with_widest_lat$text`, spans nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N).
+In our dataset, the taxa with the greatest latitudinal distribution (`r eg`, `r taxon_with_widest_lat$text`) span nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N).
 Therefore, it seems reasonable to assume that ferns are capable of dispersing across the entire study area.
+Furthermore, it should be noted that the southern islands are all low-altitude (not exceeding 1000 m).
+If high elevation habitats were available, the southern islands would be expected to share more species in common with areas further north.
+For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (>3000 m), shares many fern species with Honshu that are not found in southern Japanese islands.
+This supports that the sharp taxonomic turnover between southern islands and Honshu is more likely due to ecological factors than dispersal limitation.
 

### 2021-09-03 ms/SI.Rmd
Don't fiddle with CRS
Since CRS is same for all data when loaded (JGD2000)

@@
 
-# Set CRS
-japan_crs <- sf::st_crs(japan)
-
 # Format biodiv_ferns_spatial data
@@
   biodiv_ferns_spatial %>% 
-  # Set CRS
-  sf::st_set_crs(japan_crs) %>%
   # Set order of factor levels for plotting
@@
   biodiv_ferns_endemic_spatial %>% 
-  # Set CRS
-  sf::st_set_crs(japan_crs) %>%
   # Set order of factor levels for plotting
@@
     geom_sf(
-      data = biodiv_ferns_spatial %>% filter(...) %>% st_set_crs(japan_crs), 
+      data = biodiv_ferns_spatial %>% filter(...), 
       fill = "transparent", color = "black", lwd = 0.20) +

### 2021-09-03 ms/manuscript.Rmd
Don't fiddle with CRS
Since CRS is same for all data when loaded (JGD2000)

@@
 The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
+Points were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
 We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
@@
 We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.
+The R package `r pack("sf")` was used for GIS analyses [@Pebesma2018].
 
@@
 This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
-The R package `r pack("sf")` was used for GIS analyses [@Pebesma2018].
 
@@
 japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
-japan_crs <- sf::st_crs(japan)
 

### 2021-09-03 R/functions.R

@@ -1007,9 +1007,6 @@ load_deer_range <- function (zip_file) {
 #' protection and areas with "medium" protection
 #' 
 combine_pa <- function(protected_areas_other, protected_areas_forest) {
-	# Match CRS before combining
-	pa_crs <- st_crs(protected_areas_other)
-	protected_areas_forest <- st_transform(protected_areas_forest, pa_crs)
 	
 	# Many areas are overlapping. Combine these into non-overlapping areas within each protection area type
 	# Don't consider low-priority areas
@@ -1160,10 +1157,11 @@ process_repro_data <- function (repro_data_raw, green_list) {
 #' raster::getData("worldclim", download = TRUE, var = "bio", res = 2.5, path = path)
 #'
 #' @param path Path to data downloaded with raster::getData()
+#' @param crs Coordinate reference system to use for spatial data
 #'
 #' @return Simple features dataframe with four climate variables 
 #' 
-load_ja_worldclim_data <- function(path) {
+load_ja_worldclim_data <- function(path, crs) {
   
   # Load the WorldClim dataset at 2.5 minute resolution, crop it to the area around Japan
   raster::getData("worldclim", download = FALSE, var = "bio", res = 2.5, path = path) %>%
@@ -1177,8 +1175,8 @@ load_ja_worldclim_data <- function(path) {
       precip = bio12, # Annual Precipitation
       precip_season = bio15 # Precipitation Seasonality (Coefficient of Variation)
     ) %>%
-    # Set CRS to NA (like other spatial dataframes in this analysis)
-    st_set_crs(NA)
+    # Set CRS
+    st_set_crs(crs)
   
 }
 
@@ -1213,6 +1211,10 @@ calc_mean_climate <- function(shape_ferns, ja_climate_data) {
 
 #' Get the mean value of a numeric trait from data formatted for lucid
 #'
+#' Take mean between normal minimum and normal maximum
+#'
+#' For more information on formatting of lucid data, see https://help.lucidcentral.org/lucid/scoring-the-key/
+#'
 #' @param x Vector of values with numeric trait data formatted for lucid
 #'
 #' @return Means of each trait value
@@ -1220,6 +1222,9 @@ calc_mean_climate <- function(shape_ferns, ja_climate_data) {
 get_lucid_mean <- function (x) {
[…]

### 2021-09-03 _targets.R

@@ -43,30 +43,15 @@ tar_plan(
   tar_file(occ_point_data_summary_file, "data/japan_ferns_occ_summary.csv"),
   occ_point_data_summary = read_csv(occ_point_data_summary_file),
   
-  ## Map data ----
-  # Load shape file of Japan downloaded from https://www.gsi.go.jp/kankyochiri/gm_japan_e.html
-  # on 2020-08-26
-  tar_file(japan_pol_zip_file, "data/gm-jpn-all_u_2_2.zip"),
-  jpn_pol = load_shape_from_zip(japan_pol_zip_file, "polbnda_jpn.shp"),
-  
-  # Collapse all the political units down to just one shape for the country
-  japan_shp = jpn_pol %>%
-    select(geometry) %>%
-    summarize(),
-  
-  # Load manually entered points of interest for drawing a map of Japan
-  tar_file(japan_map_points_file, "data/japan_map_points.csv"),
-  japan_map_points = read_csv(japan_map_points_file),
-  
-  # Calculate area as rolling mean in 1 degree latitudinal windows
-  lat_area_ja = calc_area_by_lat(japan_shp, lat_cut = 0.2, lat_window = 1),
-  
   ## Occurrence data ----
   # Raw occurrence data processing was done in R/process_raw_data.R
   
   # Load geographic shapes, richness, and number of specimens at 0.2 degree grid scale
   tar_file(japan_ferns_shape_full_file, "data/japan_ferns_shape_full.gpkg"),
   shape_ferns_full = sf::st_read(japan_ferns_shape_full_file),
+
+  # Extract JGD2000 CRS for transforming other shape data
+  jgd2000 = sf::st_crs(shape_ferns_full),
   
   # Load community matrix at 0.2 degree grid scale (not filtered by redundancy)
   tar_file(japan_ferns_comm_full_file, "data/japan_ferns_comm_full.csv"),
@@ -98,7 +83,26 @@ tar_plan(
 
   # Summarize latitudinal span by taxon
   lat_span_summary = summarize_fern_lat_span(comm_ferns, shape_ferns),
-    
+
+  ## Map data ----
+  # Load shape file of Japan downloaded from https://www.gsi.go.jp/kankyochiri/gm_japan_e.html
+  # on 2020-08-26
+  tar_file(japan_pol_zip_file, "data/gm-jpn-all_u_2_2.zip"),
+  japan_shp = japan_pol_zip_file %>%
+    load_shape_from_zip("polbnda_jpn.shp") %>%
[…]

### 2021-09-03 ms/data_readme.Rmd
Update dryad readme
write out selected results files

@@
     "ms_functions",
+    "ja_climate_data_file",
+    "japan_map_file",
+    "deer_range_file",
     "japan_ferns_comm_full_file",
@@
     "redundancy_by_res_file",
-    "japan_ferns_shape_full_file"
+    "japan_ferns_shape_full_file",
+    "raw_trait_data_file",
+    "japan_map_points_file",
+    "protected_areas_file",
+    "calibration_dates_file",
+    "japan_ferns_biodiv_dryad_file",
+    "comm_ferns_dryad_file",
+    "shape_ferns_dryad_file",
+    "fern_traits_dryad_file",
+    "japan_fern_phylogram_dryad_file",
+    "japan_fern_tree_dryad_file"
     )
@@
 Associate or Co-investigators: Brent D. Mishler, Wataru Iwasaki, Atsushi Ebihara
-
-Date of data collection: 2012--2014
 
@@
 
+- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and phylogenetic tree 
 - japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
-- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of ferns in Japan
+- japan_deer_range.gpkg: Distribution maps of Japanese deer (Cervus nippon) in Japan
+- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan, full (unfiltered) dataset
 - japan_ferns_inext_results.csv: Results of collection curve analysis on fern specimens in Japan
-- japan_ferns_occ_summary.csv: Summary of specimen data of ferns in Japan
-- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for ferns of Japan
-- japan_ferns_shape_full.gpkg: Location of grid cells (sites) for ferns in Japan
-- japan_ferns_traits_lucid.csv: Morphological trait data of ferns in Japan
-- japan_map_points.csv: Latitude and longitude of points used to produce a map of Japan
+- japan_ferns_occ_summary.csv: Summary of specimen data of native, non-hybrid ferns in Japan
+- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for native, non-hybrid ferns of Japan
+- japan_ferns_shape_full.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan, full (unfiltered) dataset
+- japan_ferns_traits_lucid.csv: Morphological trait data of native, non-hybrid ferns in Japan
+- japan_map_points.csv: Latitude and longitude of points used to produce a labeled map of Japan
+- japan_map.gpkg: Map of Japan
 - japan_protected_areas.gpkg: Protected areas in Japan
 - testo_sundue_2016_calibrations.csv: Fossil calibration points used for phylogenetic dating analysis
-- results.zip: Analysis results that may be useful in future studies. Files include:
-  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of ferns in Japan
-  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of ferns in Japan
-  - japan_ferns_traits.csv: Trait matrix of ferns in Japan used for functional biodiversity analysis
-  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of ferns in Japan used for biodiversity analysis
-  - japan_ferns_biodiv.gpkg: Biodiversity statistics of ferns in Japan
+- results.zip: Selected analysis results. Files include:
+  - japan_ferns_biodiv.csv: Biodiversity statistics of native, non-hybrid ferns in Japan
+  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for biodiversity analysis
+  - japan_ferns_shape.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan used for biodiversity analysis
+  - japan_ferns_traits.csv: Trait matrix of native, non-hybrid ferns in Japan used for functional biodiversity analysis
+  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan
+  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of native, non-hybrid ferns in Japan
 
@@
 Museum of Science and Nature, Japan was converted to a community data matrix at
-four grain sizes (grid cells spanning Japan, each 0.1, 0.2, 0.3, or 0.4 degrees
+four grain sizes (grid-cells spanning Japan, each 0.1, 0.2, 0.3, or 0.4 degrees
 per side). The original list with latitude and longitude of each specimen is not
@@
 English version available at
-https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32)
+https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32).
 
@@
 
-Phylogenetic analysis was conducted with maximum likelihood in IQ-TREE
+Phylogenetic analysis was conducted with maximum likelihood in IQ-TREE v1.6.12
 [@Nguyen2015] by combining plastid *rbcL* sequences of each taxon with a
@@
 Data files were generated from raw data (not included here) using scripts
-available at https://github.com/joelnitta/japan_ferns_biogeo.
+available at https://github.com/joelnitta/japan_ferns_biogeo, in particular
+https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/process_raw_data.R.
 
@@
 
+ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and
+phylogenetic tree (Nitta et al., in prep.). For more information, see README
+file contained in zipped archive.
+
+MD5 checksum: `r tools::md5sum(here::here("data/ftol_data_release_v0.0.1.zip"))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r deer-range}
+deer_range <- sf::st_read(here::here(deer_range_file))
+# Check that no data are missing
+assert(deer_range, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(deer_range_file)`: Distribution maps of Japanese deer (Cervus
+nippon) in Japan. Raw data were downloaded from the Japan Ministry of the
+Environment
+(https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
+The data include range maps based on three types of data: range of Japanese deer
+surveyed in 1978, range surveyed in 2003, and estimated range inferred from a
+model including snow cover and forest type based on the 2003 survey data. Only
+estimated range with <0.05 movement cost was included.
+
+Number of variables: `r ncol(deer_range)`
+
+Number of cases/rows: `r nrow(deer_range)`
+
+Coordinate reference system: JGD2000
+
+Variable list:
+
+```{r results = "asis"}
+  deer_range %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "range", "Source of range data. '1978' = 1978 survey, '2003' = 2003 survey, 'estimated' = range estimated from model based on 2003 survey data",
+      "geom", "Vector describing shape and position of range polygon"
+  ))
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(deer_range_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r ja-climate-data}
+ja_climate_data <- sf::st_read(here::here(ja_climate_data_file))
+# Check that no data are missing
+assert(ja_climate_data, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(ja_climate_data_file)`: Climate data in Japan downloaded from
+WorldClim database (https://worldclim.org/data/v1.4/formats.html). Raw data were
+downloaded at the 2.5 minute scale. The intersection of the raw climate data
+with 0.2 degree grid-cells (japan_ferns_shape_full.gpkg) were identified,
+then average value for each climatic variable was calculated for each grid-cell.
+
+Number of variables: `r ncol(ja_climate_data)`
+
+Number of cases/rows: `r nrow(ja_climate_data)`
+
+Coordinate reference system: JGD2000
+
+Variable list:
+
+```{r ja-climate-data-cols, results = "asis"}
+  ja_climate_data %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "temp", "Mean annual temperature, in units of 10 * degrees celsius (BIO1)",
+      "temp_season", "Temperature seasonality (standard deviation ×100) (BIO4)",
+      "precip", "Annual precipitation (mm) (BIO12)",
+      "precip_season", "Precipitation seasonlaity (coefficient of variation) (BIO15)",
+      "geom", "Vector describing shape and position of grid-cell"
+  ))
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(ja_climate_data_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
 ```{r japan-ferns-comm-full}
@@
 `r fs::path_file(japan_ferns_comm_full_file)`: Community matrix (species x sites
-matrix) of ferns in Japan. Sites correspond to 0.2 degree x 0.2 degree grid
-cells covering Japan. For location of grid cells, see
-japan_ferns_shape_full.gpkg. Full dataset including all grid cells (none
+matrix) of native, non-hybrid ferns in Japan. Sites correspond to 0.2 degree x
+0.2 degree grid-cells covering Japan. For location of grid-cells, see
+japan_ferns_shape_full.gpkg. Full dataset including all grid-cells (none
 excluded due to low redundancy).
@@
 
-- grids: Name of grid cell.
+- grids: Name of grid-cell.
 - Other columns: Each column is named for a Japanese fern taxon. 
-  Values indicate number of specimens that were observed in each grid cell for that taxon.
+  Values indicate number of specimens that were observed in each grid-cell for that taxon.
 
@@
 
-```{r japan_ferns_inext_results}
+```{r japan-ferns-inext-results}
 inext_res <- read_csv(here::here(inext_res_file))
@@
 `r fs::path_file(inext_res_file)`: Output from running iNEXT [@Hsieh2016] on
-specimen data for ferns of Japan. iNEXT estimates diversity (here, species
-richness) and sampling completeness based on interpolation and extrapolation
-from the original data.
+specimen data of native, non-hybrid ferns in Japan. iNEXT estimates diversity
+(here, species richness) and sampling completeness based on interpolation and
+extrapolation from the original data.
 
@@
 
-```{r results = "asis"}
+```{r japan-ferns-inext-results-cols, results = "asis"}
   inext_res %>%
@@
 
-```{r occ_point_data_summary}
+```{r occ-point-data-summary}
 occ_point_data_summary <- read_csv(here::here(occ_point_data_summary_file))
@@
 filtered to remove duplicate specimens and specimens that occurred outside of
-the second-degree mesh (approximately 10 km sq grid cell) map of Japan
+the second-degree mesh (approximately 10 km sq grid-cell) map of Japan
 (http://www.biodic.go.jp/trialSystem/EN/shpddl.html)
@@
 
-```{r results = "asis"}
+```{r occ-point-data-summary-cols, results = "asis"}
   occ_point_data_summary %>%
@@
 
-```{r redundancy_by_res}
+```{r redundancy-by-res}
 redundancy_by_res <- read_csv(here::here(redundancy_by_res_file))
@@
 
-```{r results = "asis"}
+```{r redundancy-by-res-cols, results = "asis"}
   redundancy_by_res %>%
@@
 
-```{r japan_ferns_shape_full}
+```{r japan-ferns-shape-full}
 japan_ferns_shape_full <- st_read(here::here(japan_ferns_shape_full_file))
@@
 
-`r fs::path_file(japan_ferns_shape_full_file)`: Spatial data of native,
-non-hybrid ferns in Japan in GeoPackage format. Compiled from raw specimen data
-by binning specimens into 0.2 degree grid cells across Japan.
+`r fs::path_file(japan_ferns_shape_full_file)`: Location of grid-cells (sites)
+for native, non-hybrid ferns in Japan in GeoPackage format. Compiled from raw
+specimen data by binning specimens into 0.2 degree grid-cells across Japan. Full
+dataset including all grid-cells (none excluded due to low redundancy).
 
@@
 
+Coordinate reference system: JGD2000
+
 Variable list:
 
-```{r results = "asis"}
+```{r japan-ferns-shape-full-cols, results = "asis"}
   japan_ferns_shape_full %>%
@@
 
+```{r japan-ferns-traits-lucid}
+japan_ferns_traits_lucid <- read_csv(here::here(raw_trait_data_file))
+# Check that no data are missing
+assert(japan_ferns_traits_lucid, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(raw_trait_data_file)`: Morphological trait data of native,
+non-hybrid ferns in Japan, originally formatted for Lucid dichotomous key
+software (https://www.lucidcentral.org/). For more information on codes used for
+scoring of traits, see https://help.lucidcentral.org/lucid/scoring-the-key/
+
+Number of variables: `r ncol(japan_ferns_traits_lucid)`
+
+Number of cases/rows: `r nrow(japan_ferns_traits_lucid)`
+
+Variable list: 'taxon' indicates taxon name. Other columns correspond to traits.
+If the column name includes 'number' it is numeric, otherwise categorical.
+
+Categorical trait codes as follows:
+- 0 = absent
+- 1 = common
+- 2 = rare
+- 3 = uncertain
+- 4 = common and misinterpreted
+- 5 = rare and misinterpreted
+- 6 = not scoped
+
+Numeric trait data formatted as a series of numbers separated by colons (':'),
+unless missing (0 or 3). For example, 1:9.4:12:14:16. The starting '1:' does not
+mean anything. The next series of numbers correspond to the outside minimum,
+normal minimum, normal maximum and outside maximum. For example, a plant leaf
+may normally be in the range of 4-5 cm, but sometimes as low as 2 cm or as high
+as 6 cm. This would be written 2:4:5:6.
+
+Missing data codes: '0' indicates trait is missing; '3' indicates uncertain.
+
+MD5 checksum: `r tools::md5sum(here::here(raw_trait_data_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r japan-map-points}
+japan_map_points <- read_csv(here::here(japan_map_points_file))
+```
+
+`r fs::path_file(japan_map_points_file)`: Latitude and longitude of points used
+to produce a labeled map of Japan. Some points were originally generated using
+script at https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/geocode.R
+
+Number of variables: `r ncol(japan_map_points)`
+
+Number of cases/rows: `r nrow(japan_map_points)`
+
+Variable list:
+
+```{r japan-map-points-cols, results = "asis"}
+  japan_map_points %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "query", "Name of point of interest",
+      "lat", "Latitude",
+      "lon", "Longitude",
+      "lat_end", "Ending latitude (only for features that are lines, not points)",
+      "lon_end", "Ending longitude (only for features that are lines, not points)"
+  ))
+```
+
+Missing data codes: No data entered between commas in CSV.
+
+MD5 checksum: `r tools::md5sum(here::here(japan_map_points_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r japan-map}
+japan_map <- sf::st_read(here::here(japan_map_file))
+# Check that no data are missing
+assert(japan_map, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(japan_map_file)`: Map of land area of Japan. Raw data downloaded from
+Geospatial Information Authority of Japan under the Creative Commons Attribution
+License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html). All polygons
+were combined into a single polygon and CRS set to JGD2000.
+
+Number of variables: `r ncol(japan_map)`
+
+Number of cases/rows: `r nrow(japan_map)`
+
+Variable list:
+
+```{r japan-map-cols, results = "asis"}
+  japan_map %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "geom", "Vector describing shape and position of land area polygon"
+  ))
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(japan_map_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r protected-areas}
+protected_areas <- st_read(here::here(protected_areas_file))
+# Check that no data are missing
+assert(protected_areas, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(protected_areas_file)`: Spatial of protected areas in Japan
+in GeoPackage format. Original spatial data downloaded from the Japan Ministry
+of the Environment
+(https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html) and
+Ministry of Land, Infrastructure, Transport and Tourism
+(https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-A45.html). Categorization of
+protected areas generally followed @Kusumoto2017: either "high" (no human
+activities allowed at all) or "medium" status (some economic activities allowed
+by permit); other areas not meeting these criteria or not including plants were
+not considered. Any overlapping areas within the same protection level were
+combined, and any areas overlapping between "medium" and "high" were considered
+only "high". Code used to categorize areas available at
+https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/process_raw_data.R
+
+Number of variables: `r ncol(protected_areas)`
+
+Number of cases/rows: `r nrow(protected_areas)`
+
+Coordinate reference system: JGD2000
+
+Variable list:
+
+```{r protected-areas-cols, results = "asis"}
+  protected_areas %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "status", "Protection status ('high' or 'medium')",
+      "geom", "Vector describing shape and position of protected area polygon"
+  ))
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(protected_areas_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r plastome-calibration-dates}
+plastome_calibration_dates <- read_csv(here::here(calibration_dates_file))
+```
+
+`r fs::path_file(calibration_dates_file)`: Fossil calibration points used for
+phylogenetic dating analysis with treePL [@Smith2012], based on @Testo2016a.
+MRCA = "most recent common ancestor".
+
+Number of variables: `r ncol(plastome_calibration_dates)`
+
+Number of cases/rows: `r nrow(plastome_calibration_dates)`
+
+Variable list:
+
+```{r plastome-calibration-dates-cols, results = "asis"}
+  plastome_calibration_dates %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "Clade", "Name of clade",
+      "Stem_Crown", "Type of node. 'crown' (node corresponding to MRCA of extant taxa only) or 'stem' (node corresponding to split between MRCA of extant taxa and its sister)",
+      "Fossil", "Name of fossil",
+      "Age", "Age of fossil",
+      "Age_type", "Type of age. 'min' for minimum age, or 'fixed' for fixed age",
+      "Citation", "Citation for the fossil",
+      "taxon_1", "Name of first taxon that is used to identify MRCA",
+      "taxon_2", "Name of second taxon that is used to identify MRCA",
+      "note", "Notes, in particular explaining any differences with Testo and Sundue (2016)"
+  ))
+```
+
+Missing data codes: No data entered between commas in CSV.
+
+MD5 checksum: `r tools::md5sum(here::here(calibration_dates_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r japan-ferns-biodiv-dryad}
+japan_ferns_biodiv_dryad <- read_csv(here::here(japan_ferns_biodiv_dryad_file))
+# Check that no data are missing
+assert(japan_ferns_biodiv_dryad, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(japan_ferns_biodiv_dryad_file)` (contained in "results.zip"):
+Biodiversity statistics of native, non-hybrid ferns and environmental variables
+in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. A
+single outlier grid-cell with extremely low taxon richness and extremely high %
+apomictic taxa was excluded. Climatic (temperature and preciptation) variables
+calculated as described for japan_climate.gpkg. Data used for spatial models
+in Nitta et al. 2021.
+
+Number of variables: `r ncol(japan_ferns_biodiv_dryad)`
+
+Number of cases/rows: `r nrow(japan_ferns_biodiv_dryad)`
+
+Variable list:
+
+```{r japan-ferns-biodiv-dryad-cols, results = "asis"}
+  japan_ferns_biodiv_dryad %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "grids", "Name of grid-cell",
+      "lat", "Latitude of grid-cell centroid",
+      "long", "Longitude of grid-cell centroid",
+      "abundance", "Number of specimens occurring in grid-cell",
+      "richness", "Number of taxa occurring in grid-cell",
+      "fd_obs_z", "Z-score (standard effect size) of observed functional diversity (FD)",
+      "pd_obs_z", "Z-score of observed phylogenetic diversity (PD)",
+      "pe_obs_z", "Z-score of observed phylogenetic endemisim (PE)",
+      "rfd_obs_z", "Z-score of observed relative FD",
+      "rpd_obs_z", "Z-score of observed relative PD",
+      "richness_obs_p_upper", "Richness percentile",
+      "fd_obs_p_upper", "FD percentile",
+      "pd_obs_p_upper", "PD percentile",
+      "pe_obs_p_upper", "PE percentile",
+      "pd_signif", "Significance of PD compared to null distribution, two-sided test at alpha 0.05",
+      "rpd_signif", "Significance of RPD compared to null distribution, two-sided test at alpha 0.05",
+      "fd_signif", "Significance of FD compared to null distribution, two-sided test at alpha 0.05",
+      "rfd_signif", "Significance of RFD compared to null distribution, two-sided test at alpha 0.05",
+      "pe_signif", "Significance of PE compared to null distribution, two-sided test at alpha 0.05",
+      "taxonomic_cluster", "Membership in cluster based on taxonomic distances",
+      "phylo_cluster", "Membership in cluster based on phylogenetic distances",
+      "endem_type", "Endemism type determined by CANAPE",
+      "lat_area", "Area of 1-degree latitudinal band around grid-cell",
+      "temp", "Annual mean temperature, in units of degrees celsius * 10",
+      "temp_season", "Temperature seasonality",
+      "precip", "Annual precipitation, measured in mm",
+      "precip_season", "Precipitation seasonality",
+      "percent_apo", "Percent apomictic taxa"
+    ))
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(japan_ferns_biodiv_dryad_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r comm-ferns-dryad-file}
+comm_ferns_dryad <- read_csv(here::here(comm_ferns_dryad_file))
+# Check that no data are missing
+assert(comm_ferns_dryad, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(comm_ferns_dryad_file)` (contained in "results.zip"): Community
+matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for
+biodiversity analysis. Same as japan_ferns_comm_full.csv, but grid-cells with
+redundancy <0.1 (indicating under-sampling) excluded.
+
+Number of variables: `r ncol(comm_ferns_dryad)`
+
+Number of cases/rows: `r nrow(comm_ferns_dryad)`
+
+Variable list:
+
+- grids: Name of grid-cell.
+- Other columns: Each column is named for a Japanese fern taxon. 
+  Values indicate number of specimens that were observed in each grid-cell for that taxon.
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(comm_ferns_dryad_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r shape-ferns-dryad-file}
+shape_ferns_dryad <- sf::st_read(here::here(shape_ferns_dryad_file))
+# Check that no data are missing
+assert(shape_ferns_dryad, not_na, everything(), success_fun = success_logical)
+```
+
+`r fs::path_file(shape_ferns_dryad_file)` (contained in "results.zip"): Location
+of grid-cells (sites) for native, non-hybrid ferns in Japan used for
+biodiversity analysis. Same as japan_ferns_shape_full.gpkg, but grid-cells with
+redundancy <0.1 (indicating under-sampling) excluded.
+
+Number of variables: `r ncol(shape_ferns_dryad)`
+
+Number of cases/rows: `r nrow(shape_ferns_dryad)`
+
+Coordinate reference system: JGD2000
+
+Variable list:
+
+```{r shape-ferns-dryad-file-cols, results = "asis"}
+  shape_ferns_dryad %>%
+  make_data_desc(
+    tribble(
+      ~col, ~desc,
+      "grids", "Name of grid-cell",
+      "abundance", "Number of specimens occurring in grid-cell",
+      "richness", "Number of taxa occurring in grid-cell",
+      "redundancy", "Redundancy (measure of sampling completeness)",
+      "geom", "Vector describing shape and position of grid-cell"
+  ))
+```
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(shape_ferns_dryad_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r fern-traits-dryad-file}
+fern_traits_dryad <- read_csv(here::here(fern_traits_dryad_file))
+```
+
+`r fs::path_file(fern_traits_dryad_file)` (contained in "results.zip"): Trait
+matrix of native, non-hybrid ferns in Japan used for functional biodiversity
+analysis.
+
+Number of variables: `r ncol(fern_traits_dryad)`
+
+Number of cases/rows: `r nrow(fern_traits_dryad)`
+
+Variable list: 'taxon' indicates taxon name. Other columns correspond to traits.
+'frond_width', 'stipe_length', 'number_pinna_pairs' are numeric; others are
+binary (0 or 1)
+
+Missing data codes: Missing data coded as 'NA'.
+
+MD5 checksum: `r tools::md5sum(here::here(fern_traits_dryad_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r japan-fern-phylogram-dryad-file}
+japan_fern_phylogram_dryad <- ape::read.tree(here::here(japan_fern_phylogram_dryad_file))
+```
+
+`r fs::path_file(japan_fern_phylogram_dryad_file)` (contained in
+"results.zip"): Maximim-likelihood phylogeny of native, non-hybrid ferns in
+Japan in newick format inferred with maximum likelihood IQ-TREE v1.6.12
+[@Nguyen2015]. Values at nodes are bootstrap support values calculated using
+1000 replicates of ultrafast bootstrap [@Nguyen2015].
+
+Number of tips: `r ape::Ntip(japan_fern_phylogram_dryad)`
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(japan_fern_phylogram_dryad_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+```{r japan-fern-dryad-file}
+japan_fern_tree_dryad <- ape::read.tree(here::here(japan_fern_tree_dryad_file))
+```
+
+`r fs::path_file(japan_fern_tree_dryad_file)` (contained in "results.zip"):
+Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
+Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r
+n_fossil_points` fossil calibration points after @Testo2016a. Values at nodes
+are bootstrap support values calculated using 1000 replicates of ultrafast
+bootstrap [@Nguyen2015].
+
+Number of tips: `r ape::Ntip(japan_fern_tree_dryad)`
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_dryad_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
 CHANGE LOG
@@
 
-2021-08-31
+2021-09-03
 

### 2021-09-03 ms/SI.Rmd
Update SI
- Add plot for latitudinal patterns
- Fix lumping of bioregions in SI plot

@@
   "protected_areas",
-  "deer_range"
+  "deer_range",
+  "biodiv_ferns_spatial_lumped"
 ))
@@
 
+```{r raw-lat, fig.width = 7, fig.height = 4}
+# Convert spatial data frame to centroids with raw metrics
+biodiv_ferns_cent_raw <- spatial_to_cent(
+  biodiv_ferns_spatial,
+  c("grids", "lat", "long", "richness", "pd_obs", "fd_obs"))
+
+# Plot each raw metric vs. latitude
+a <- ggplot(biodiv_ferns_cent_raw, aes(y = richness, x = lat)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  scale_x_continuous(
+    name = "Latitude",
+    labels = label_number(accuracy = 1, suffix ="\u00b0 N")) +
+  labs(y = "Richness") +
+  coord_flip() +
+  theme(panel.grid.minor = element_blank())
+
+b <- ggplot(biodiv_ferns_cent_raw, aes(y = pd_obs, x = lat)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  scale_x_continuous(
+    labels = label_number(accuracy = 1, suffix ="\u00b0 N")) +
+  labs(y = "PD") +
+  coord_flip() +
+  theme(
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank())
+
+c <- ggplot(biodiv_ferns_cent_raw, aes(y = fd_obs, x = lat)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth() +
+  scale_x_continuous(
+    labels = label_number(accuracy = 1, suffix ="\u00b0 N")) +
+  labs(y = "FD") +
+  coord_flip() +
+  theme(
+    panel.grid.minor = element_blank(),
+    axis.title.y = element_blank())
+
+a + b + c + 
+  plot_layout(nrow = 1) +
+  plot_annotation(tag_levels = "a",
+                  tag_prefix = "(",
+                  tag_suffix = ")") &
+  theme(plot.tag = element_text(face = "bold"))
+```
+
+**`r s_figure("raw-lat")`**.
+(a) Raw taxonomic diversity (b), phylogenetic diversity, and (c) functional diversity of the ferns of Japan plotted by latitude.
+For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
+
+\clearpage
+
 ```{r raw-biplot, fig.width = 7, fig.height = 7}
@@
 a <-
-  ggplot(biodiv_ferns_spatial, aes(x = precip, y = temp, color = taxonomic_cluster)) +
+  ggplot(biodiv_ferns_spatial_lumped, aes(x = precip, y = temp, color = taxonomic_cluster)) +
   geom_point(size = 1, alpha = 0.7) +
@@
 b <-
-  ggplot(biodiv_ferns_spatial, aes(x = precip, y = temp, color = phylo_cluster)) +
+  ggplot(biodiv_ferns_spatial_lumped, aes(x = precip, y = temp, color = phylo_cluster)) +
   geom_point(size = 1, alpha = 0.7) +
@@
 Points shown with partial transparency; darker areas indicate overlapping points.
+Bioregions with three or fewer grid-cells are lumped into the "Other" category.
 

### 2021-09-03 ms/manuscript.Rmd
Update dryad readme
write out selected results files
Update SI
- Add plot for latitudinal patterns
- Fix lumping of bioregions in SI plot

@@
   "template_file",
-  "traits_summary"
+  "traits_summary",
+  "biodiv_ferns_spatial_lumped"
 ))
@@
   map(., ~split(., .$variable))
+
+# Count number of major bioregions after lumping minor regions into "Other"
+maj_cluster_count <-
+biodiv_ferns_spatial_lumped %>% 
+  select(grids, contains("cluster")) %>%
+  st_drop_geometry() %>%
+  as_tibble() %>%
+  pivot_longer(names_to = "type", values_to = "cluster", -grids) %>%
+  group_by(type) %>%
+  count(cluster) %>%
+  filter(cluster != "Other") %>%
+  summarize(n = n_distinct(cluster), .groups = "drop") %>%
+  # Convert to list for easy indexing
+  split(.$type)
 ```
@@
 **Results** We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes.
-Grid-cells were grouped in to three (phylogenetic) or four (taxonomic) major bioregions. 
+Grid-cells were grouped in to `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions. 
 Temperature and apomixis were identified as likely drivers of patterns of taxonomic and phylogenetic diversity.
-Conservation status (% protected area) was generally high for grid-cells with significantly high biodiversity across all biodiversity metrics, but the threat due to herbivory by native Japanese deer varied by biodiversity metric.
+Conservation status (% protected area) was generally high for grid-cells with significantly high biodiversity across all biodiversity metrics, but the threat due to herbivory by native Japanese deer varied by biodiversity metric, and was particularly severe for taxonomic richness.
 
@@
 Although peak areas of different diversity types are largely non-overlapping, each one has relatively high protection coverage at medium protection level. 
-Despite this, future conservation efforts must recognize that threats can vary by biodiversity metric, and consider threats not directly analyzed here such as climate change.
+Future conservation efforts should recognize that threats can vary by biodiversity metric, and consider multiple multiple metrics when establishing conservation priorities.
 
@@
 
-We used a list of all non-hybrid fern specimens deposited at TNS (herbarium codes follow Thiers2021). 
+We used a list of all non-hybrid fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
 We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v.
@@
 To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
-We conducted phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] and molecular dating with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
+We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1000 ultrafast bootstrap replicates [@Minh2013].
+Molecular dating was conducted with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
 The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
@@
 
-Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`a).
+Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`a, `r s_figure("raw-lat")`a).
 Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`b, c), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`a, c).
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants [@Fukaya2020].
+The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`a, `r s_figure("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants [@Fukaya2020].
 Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`a).
@@
 ```{r bioregions-make}
-# Lump regions with just a few cells (3 or less)
-biodiv_ferns_spatial_lumped <-
-biodiv_ferns_spatial %>%
-  mutate(taxonomic_cluster = fct_lump_min(taxonomic_cluster, 4))  %>%
-  mutate(phylo_cluster = fct_lump_min(phylo_cluster, 4))
-
 # Taxonomic bioregions
@@
 A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
-Data files, including the DNA alignment, phylogenetic tree, and community data matrix, are openly available from Dryad at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
+Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
 

### 2021-09-03 _targets.R

@@ -586,6 +586,12 @@ tar_plan(
     write_tree_tar(japan_fern_tree, "results/dryad_files/japan_ferns_tree_dated.tre")
   ),
   
+  # Etc ----
+  # Lump regions with just a few cells (3 or less)
+  biodiv_ferns_spatial_lumped = biodiv_ferns_spatial %>%
+    mutate(taxonomic_cluster = fct_lump_min(taxonomic_cluster, 4))  %>%
+    mutate(phylo_cluster = fct_lump_min(phylo_cluster, 4)),
+  
   # Render manuscript ----
   # Track ms files
   tar_file(refs_yaml, "ms/references.yaml"),

### 2021-09-03 ms/manuscript.Rmd
Add check on statement about conditional AIC

@@
 ### Relationships between biodiversity and climate or reproductive mode
+
+```{r discus-model-stats}
+# Check that conditional AIC was lower in the reproductive model compared to the environmental model for SES of RPD and SES of PD
+# Split aic_env_repro dataframe into nested list for easier indexing
+c_aic <- aic_env_repro %>% split(.$resp_var) %>% map(~split(., .$model_type))
+assert_that(c_aic$pd_obs_z$percent_apo$conditional_aic < c_aic$pd_obs_z$temp$conditional_aic)
+assert_that(c_aic$rpd_obs_z$percent_apo$conditional_aic < c_aic$rpd_obs_z$temp$conditional_aic)
+```
 

### 2021-09-03 ms/SI.Rmd
Lighten lines in map

@@
   ggplot() +
-    geom_sf(data = japan, lwd = 0.10, fill = "white") +
+    geom_sf(data = japan, lwd = 0.07, fill = "white") +
     geom_sf(data = protected_areas %>% mutate(status = str_to_sentence(status)), aes(fill = status), color = "transparent") +
@@
       data = biodiv_ferns_spatial %>% filter(...), 
-      fill = "transparent", color = "black", lwd = 0.20) +
+      fill = "transparent", color = "black", lwd = 0.12) +
     annotate(
@@
   ggplot() +
-    geom_sf(data = japan, lwd = 0.10, fill = "white") +
+    geom_sf(data = japan, lwd = 0.07, fill = "white") +
     geom_sf(data = deer_range %>% filter(range == "estimated") %>% mutate(range = str_to_sentence(range)), aes(fill = range), color = "transparent") +

### 2021-09-05 ms/manuscript.Rmd
Update ms

@@
 
-**Results** We observe three major patterns in the spatial distribution of biodiversity: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes.
+**Results** Three major patterns in the spatial distribution of biodiversity were observed: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes.
 Grid-cells were grouped in to `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions. 
-Temperature and apomixis were identified as likely drivers of patterns of taxonomic and phylogenetic diversity.
-Conservation status (% protected area) was generally high for grid-cells with significantly high biodiversity across all biodiversity metrics, but the threat due to herbivory by native Japanese deer varied by biodiversity metric, and was particularly severe for taxonomic richness.
+Temperature and apomixis were identified as drivers of patterns of taxonomic and phylogenetic diversity.
+Conservation status (% protected area) was generally high for grid-cells with significantly high biodiversity, but the threat due to herbivory by deer varied by biodiversity metric, and was particularly severe for taxonomic richness.
 
@@
 The different distributions of each diversity type reflect the major bioregions. 
-Although peak areas of different diversity types are largely non-overlapping, each one has relatively high protection coverage at medium protection level. 
-Future conservation efforts should recognize that threats can vary by biodiversity metric, and consider multiple multiple metrics when establishing conservation priorities.
+Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple multiple metrics when establishing conservation priorities.
 
@@
 
-Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because data (DNA sequences), extensive spatial data, and analytic tools have only become available relatively recently.
-These tools and datasets now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. 
-Better ways to characterize biological regions ("bioregions") are also available, rather than relying on ad-hoc characterizations [e.g., @Laffan2016; @White2019; @Daru2020a].
+Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently.
+These datasets and tools now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. 
+Better ways to characterize biological regions ("bioregions") are also available [e.g., @Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
 Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.
@@
 
-All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The `r pack("targets")` R package was used to maintain analysis workflow [@Landau2021].
+All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The `r pack("targets")` R package was used to control analysis workflow [@Landau2021].
 
@@
 
-Our data include all ferns known to occur within the boundaries of Japan (20°N to 45°N, 122°E to 153°E). 
+Our data include all ferns known to natively occur within the boundaries of Japan (20°N to 45°N, 122°E to 153°E). 
 Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
@@
 The climate is subtropical in the south and temperate in other parts of the country. 
-Map data were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).
+Map data for Japan were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).
 
@@
 
-We used a list of all non-hybrid fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
+We used a list of all native, non-hybrid fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
 We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v.
@@
 The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
-Points were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
+Points (latitude and longitude of each occurrence record) were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
 We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
+The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].
 
@@
 We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.
-The R package `r pack("sf")` was used for GIS analyses [@Pebesma2018].
 
@@
 To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
-We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1000 ultrafast bootstrap replicates [@Minh2013].
+We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].
 Molecular dating was conducted with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
@@
 
-We downloaded SHP files of protected areas in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html) and Ministry of Land, Infrastructure, Transport and Tourism (https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-A45.html).
-We only considered areas that include protection of plants, and categorized protected areas as either "high" (no human activities allowed at all) or "medium" status (some economic activities allowed by permit) [@Kusumoto2017].
-Any areas that overlapped within a protection status were combined, and areas that overlapped between medium and high status were considered only high status.
+We downloaded SHP files of conservation zones in Japan from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html) and Ministry of Land, Infrastructure, Transport and Tourism (https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-A45.html).
+We excluded marine zones and those that do not protect plants.
+We categorized protected areas as either "high" (no human activities allowed at all) or "medium" status (some economic activities allowed by permit) [@Kusumoto2017]; areas not afforded at least medium level of protection were not considered.
+Conservation status in Japan is administered by multiple laws, and protected areas frequently overlap [@Natori2012].
+To prevent double-counting, all protected areas within a protection status were merged, and protected areas that overlapped between medium and high status were considered only high status.
+
 We downloaded distribution maps (SHP and TIF) of native Japanese deer (*Cervus nippon*), an herbivore that has caused extensive damage to native plant communities due to rapid population growth since the 1970s [@Takatsuki2009] and poses a high risk to ferns, from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
 To assess change in threat due to herbivory by deer over time, we used three maps available in this dataset: range of Japanese deer surveyed in 1978, range surveyed in 2003, and estimated range inferred from a model including snow cover and forest type based on the 2003 survey data [@kankyosho2021].
+
 We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type or percent of area occupied by deer for each survey period within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
@@
 Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`b, c), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`a, c).
-Observed values of RPD and RFD were not linearly related to richness, however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`b, d).
+Observed values of RPD and RFD were not linearly related to richness; however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`b, d).
 
@@
 We found that grid-cells with significantly high amounts of biodiversity are better protected than the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`a).
-The high protection status of grid-cells with high PE may be due to an emphasis on conserving rare, endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`d).
+The high protection status of grid-cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`d).
 One possible reason for the generally high level of protection overall is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas [A.

### 2021-09-06 ms/SI.Rmd
Update MS
Add version for R packages

@@
 header-includes: 
-  \usepackage{float}
-  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
+  - \usepackage{float}
+  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
 editor_options: 

### 2021-09-06 ms/manuscript.Rmd
Update MS
Add version for R packages

@@
 
-All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The `r pack("targets")` R package was used to control analysis workflow [@Landau2021].
+All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The R package `r pack("targets")` was used to control analysis workflow [@Landau2021].
 
@@
 We also tested alternative weighting schemes, but these produced very similar results (data not shown).
-We then used the distances to build a trait dendrogram using the `r func("hclust")` function in the `r pack("stats")` R package [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
+We then used the distances to build a trait dendrogram using the `r func("hclust")` function in the R package `r pack("stats")` [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
 These functional analogs of PD and RPD are hereafter referred to as FD (functional diversity) and RFD (relative functional diversity).
@@
 Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution.
-
-All measures of biodiversity and CANAPE were carried out using the `r pack("canaper")` package [@Nitta2021a], which calculates raw biodiversity metrics using functions in the `r pack("phyloregion")` package [@Daru2020a] and randomizations using the `r pack("picante")` package [@Kembel2010].
+All measures of biodiversity and CANAPE were carried out using the R package `r pack("canaper")` [@Nitta2021a].
 
@@
 
-We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the `r pack("spdep")` package [@Bivand2013].
+We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the R package `r pack("spdep")` [@Bivand2013].
 In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05) (`r s_table("moran")`). 
@@
 To assess change in threat due to herbivory by deer over time, we used three maps available in this dataset: range of Japanese deer surveyed in 1978, range surveyed in 2003, and estimated range inferred from a model including snow cover and forest type based on the 2003 survey data [@kankyosho2021].
+We limited estimated range to areas considered "highly likely" to be occupied by deer [movement cost < 0.10; @kankyosho2021].
 

### 2021-09-06 R/ms_functions.R

@@ -80,7 +80,7 @@ eg <- "*eg*"
 #'
 #' @return Name of R package formatted according to journal requirements
 #' 
-pack <- function(x) {glue::glue("'{x}'")}
+pack <- function(x) {glue::glue("'{x}' v.{packageVersion(x)}")}
 
 #' Format other software names
 #' 

### 2021-09-06 ms/manuscript.Rmd
Fix taxon with widest lat text

@@
   mutate(lat_span = max_lat - min_lat) %>%
-  slice_max(lat_span, with_ties = FALSE) %>%
-  left_join(green_list, by = "taxon") %>%
-  # Parse out species name and author
-  mutate(gn_parse_tidy(.$scientific_name) %>% select(canonicalsimple, authorship)) %>%
-  mutate(text = glue("*{canonicalsimple}* {authorship}"))
+  slice_max(lat_span) %>%
+  # Add formatted scientific name
+  left_join(
+    tibble(
+      taxon = "Pteridium_aquilinum_japonicum",
+      text = "*Pteridium aquilinum* (L.) Kuhn subsp. *japonicum* (Nakai) A. et S.Löve"
+    ),
+    by = "taxon"
+  ) %>%
+  # Round off latitude
+  mutate(across(c(min_lat, max_lat), ~round(., 2)))
+
+# Make sure the taxon_with_widest_lat is what we expect
+assert_that(
+  taxon_with_widest_lat$taxon == "Pteridium_aquilinum_japonicum",
+  msg = "Taxon with widest latitude wasn't Pteridium_aquilinum_japonicum")
 

### 2021-09-06 ms/SI.Rmd
Don't hard-code bioregion lumping cutoff

@@
   "deer_range",
-  "biodiv_ferns_spatial_lumped"
+  "biodiv_ferns_spatial_lumped",
+  "bioregion_cutoff"
 ))
@@
 ```{r raw-lat, fig.width = 7, fig.height = 4}
-# Convert spatial data frame to centroids with raw metrics
-biodiv_ferns_cent_raw <- spatial_to_cent(
+# Convert spatial data frame to centroids with raw metrics, drop single high apomixis outlier
+biodiv_ferns_cent_raw <- spatial_to_cent_for_model(
   biodiv_ferns_spatial,
@@
   geom_point(alpha = 0.5) +
+  geom_smooth() +
   labs(x = "Richness", y = "PD")
@@
   geom_point(alpha = 0.5) +
+  geom_smooth() +
   labs(x = "Richness", y = "FD")
@@
 (d), Relative functional diversity (RFD).
+Trendlines in (a, c) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
 
@@
   as_tibble() %>%
-  select(grids, cluster = taxonomic_cluster) %>%
-  # Convert cluster to numeric to use max() later
-  mutate(cluster = levels(cluster)[cluster])
+  select(grids, cluster = taxonomic_cluster)
 
 cluster_nodes_tax_tbl <- map_df(unique(clusters_tax$cluster), ~map_cluster_to_nodes(dendro_tax, clusters_tax, .)) %>%
-  # Lump regions with just a few cells (3 or less)
-  mutate(cluster = as.factor(cluster) %>% fct_lump_min(4))
+  # Lump regions with fewer grid-cells than cutoff
+  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff))
 
@@
   as_tibble() %>%
-  select(grids, cluster = phylo_cluster) %>%
-  # Convert cluster to numeric to use max() later
-  mutate(cluster = levels(cluster)[cluster])
+  select(grids, cluster = phylo_cluster)
 
 cluster_nodes_phy_tbl <- map_df(unique(clusters_phy$cluster), ~map_cluster_to_nodes(dendro_phy, clusters_phy, .)) %>%
-  # Lump regions with just a few cells (3 or less)
-  mutate(cluster = as.factor(cluster) %>% fct_lump_min(4))
+  # Lump regions with fewer grid-cells than cutoff
+  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff))
 

### 2021-09-06 ms/manuscript.Rmd
Don't hard-code bioregion lumping cutoff

@@
   "traits_summary",
-  "biodiv_ferns_spatial_lumped"
+  "biodiv_ferns_spatial_lumped",
+  "bioregion_cutoff"
 ))
@@
 Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells.
-Bioregions with three or fewer grid-cells are lumped into the "Other" category.
+Bioregions each consisting of fewer than `r english2(bioregion_cutoff)` grid-cells are lumped into the "Other" category.
 

### 2021-09-06 ms/SI.Rmd
Add check for bioregion lumping level

@@
 
+# Make sure cutoff matches caption
+assert_that(bioregion_cutoff == 2, msg = "Bioregions cutoff for lumping not 2!")
+
 # Use the ggtree::`%<+%` operator to map the bootstrap values onto the ggtree object
@@
 Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods).
-Color shows bioregion; bioregions with three or fewer grid cells are lumped into the "Other" category.
+Color shows bioregion; bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.
 (a), taxonomic bioregions.

### 2021-09-06 ms/manuscript.Rmd
Add check for bioregion lumping level

@@
 
+```{r verify-cutoff}
+assert_that(bioregion_cutoff == 2, msg = "Bioregions cutoff for lumping not 2!")
+```
+
 Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
 The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu), bioregion 3 (Ryukyu Islands, excluding Yakushima), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (`r figure("bioregions")`b; `r s_figure("dendrogram")`b). 
-The other much smaller bioregions (each including only three grid-cells or less) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion, and bioregion 1 extended slightly further south (`r figure("bioregions")`b; `r s_figure("dendrogram")`b).
+The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
 

### 2021-09-06 ms/manuscript.Rmd
Update MS
- add explanation of collapsing identical seqs
- fix widest lat taxon text
- update refs

@@
 The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
-We then trimmed the tree to Japanese taxa only and used this for all further analyses.
+We trimmed the resulting ultrametric tree to only Japanese taxa.
+treePL requires branch lengths to be > 0, and it resets extremely small branches an arbitrary minimum positive length.
+This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
+We converted any clades in the original dated tree consisting of identical sequences to polytomies with zero branch length between OTUs.
+This tree was used for all subsequent analysis unless indicated otherwise.
 
@@
 We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the `r func("phylo.d")` function in the R package `r pack("caper")` [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
+These methods for phylogenetic signal cannot use a tree with zero-length polytomies, so we used the original dated tree from treePL.
 
@@
 
-We first tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the R package `r pack("spdep")` [@Bivand2013].
+Initial inspection of the biodiversity and environmental data showed a single outlier in % apomictic taxa due to extremely low species richness (Appendix S2), so that grid-cell and one grid-cell lacking environmental data were excluded from modeling analyses.
+
+We tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the R package `r pack("spdep")` [@Bivand2013].
 In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05) (`r s_table("moran")`). 
@@
 
-We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the `r pack("SpatialPack")` R package [@SpatialPack2020].
+We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the R package `r pack("SpatialPack")` [@SpatialPack2020].
 This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature (`r s_table("corr")`), so for the main modeling analysis we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
@@
 The distances are used to construct a dendrogram, which is then split into into *k* bioregions. 
-As the optimal number of bioregions cannot be known *a-priori*, we tested values of *k* from 1 to 20.
+The optimal value of *k* cannot be known *a-priori*.
+Previous vegetation zone schemes in Japan have typically included 5--6 zones [@Shimizu2014]; to enable approximate comparison with these, we tested values of *k* from 1 to 10, and selected the optimal value using the `optimal_phyloregion` function in the R package `r pack("phyloregion")`.
 
@@
 For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above. 
-For taxonomic richness, cells ranked in the top >5% were considered highly diverse.
+For taxonomic richness, cells ranked in the top > 5% were considered highly diverse.
 
@@
 Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`a).
-Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018a; @Qian2021] to be important drivers of spatial fern diversity.
+Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
 A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a]. 
@@
 The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`).
-Interestingly, the bioregions identified here correspond roughly to major forest types previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
+Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
 Furthermore, the border between bioregions 1 and 2 corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
@@
       taxon = "Pteridium_aquilinum_japonicum",
-      text = "*Pteridium aquilinum* (L.) Kuhn subsp. *japonicum* (Nakai) A. et S.Löve"
+      text = "*Pteridium aquilinum* (L.) Kuhn subsp. *japonicum* (Nakai) A. et S. Löve"
     ),
@@
 However, ferns are known to have high dispersal abilities [@Tryon1970; @Barrington1993; @Moran2001]. 
-In our dataset, the taxa with the greatest latitudinal distribution (`r eg`, `r taxon_with_widest_lat$text`) span nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N).
+In our dataset, the taxon with the greatest latitudinal distribution (`r taxon_with_widest_lat$text`) spans nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N).
 Therefore, it seems reasonable to assume that ferns are capable of dispersing across the entire study area.
@@
 If high elevation habitats were available, the southern islands would be expected to share more species in common with areas further north.
-For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (>3000 m), shares many fern species with Honshu that are not found in southern Japanese islands.
+For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (> 3000 m), shares many fern species with Honshu that are not found in southern Japanese islands.
 This supports that the sharp taxonomic turnover between southern islands and Honshu is more likely due to ecological factors than dispersal limitation.
@@
 Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells.
-Bioregions each consisting of fewer than `r english2(bioregion_cutoff)` grid-cells are lumped into the "Other" category.
+Bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.
 

### 2021-09-06 ms/data_exploration.Rmd
Make data exploration a pdf (SI2)

--- 2021-07-09 ms/data_exploration.Rmd
@@
 ---
-title: "Data exploration"
-output: html_document
+title: "Appendix S2: Data Exploration"
+mainfont: Roboto-Regular.ttf
+fontsize: 11pt
+mainfontoptions: 
+  - BoldFont=Roboto-Bold.ttf
+  - ItalicFont=Roboto-Italic.ttf
+  - BoldItalicFont=Roboto-BoldItalic.ttf
+output: 
+  bookdown::pdf_document2:
+    citation_package: default
+    toc: no
+    number_sections: no
+    fig_caption: yes
+    keep_tex: no
+    latex_engine: xelatex
+    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
 editor_options: 
   chunk_output_type: console
-params:
-  knit_type: manual
+header-includes: 
+  - \usepackage{float}
+  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
 ---
@@
 
-```{r load-packages, include = params$knit_type == "manual"}
+All code is in R, and workflow managed with the `targets` package (https://github.com/ropensci/targets).
+
+The code here is included in the repo at https://github.com/joelnitta/japan_ferns_biogeo.
+
+Let's get started by loading packages and tidying our namespace.
+
+```{r load-packages, message = FALSE}
 # Load packages
@@
 library(assertr)
+library(conflicted)
 
-# Load custom functions
-source(here::here("R/functions.R"))
+# Resolve possible namespace conflicts
+conflict_prefer("map", "purrr")
+conflict_prefer("select", "dplyr")
+conflict_prefer("filter", "dplyr")
+conflict_prefer("gather", "tidyr")
+conflict_prefer("extract", "magrittr")
 ```
 
-The dataset includes biodiversity metrics and environmental data.
+The dataset includes biodiversity metrics and environmental data measured on native, non-hybrid ferns of Japan.
 
-All biodiversity metrics were calculated during the `targets` workflow (`_targets.R`).
+All biodiversity metrics were calculated during the `targets` workflow (`_targets.R`), and are contained in the `biodiv_ferns_cent_dryad` object in the `targets` cache.
 
-The `cent` part of the name indicates that they include geographic locations as centroids of each grid cell.
+The next step uses `tar_load()` to load the dataset from the targets cache. Alternatively, this can be loaded from the dryad data file as shown in the commented-out code below.
 
@@
 # Load the data from the targets workflow
-tar_load(biodiv_ferns_spatial)
-tar_load(resp_vars_env)
-tar_load(indep_vars)
+tar_load(biodiv_ferns_cent_dryad)
 
-# Make biodiversity metrics dataframe with centroid of each site for models
-  biodiv_ferns_cent <- 
-    biodiv_ferns_spatial %>%
-    # keep only variables needed for model and only rows with zero missing data
-    filter_data_for_model(c("grids", "lat", "long", resp_vars_env, indep_vars)) %>%
-    sf::st_set_geometry(NULL) %>%
-    as_tibble()
+# Alternatively, unzip `results.zip` in the Dryad data
+# (FIXME: add DOI when available)
+# and read in the data from there
+# biodiv_ferns_cent_dryad <- read_csv("japan_ferns_biodiv.csv")
+```
+
+This dataframe actually has more variables than are needed for modeling, so we will subset it to only the relevant variables.
+
+```{r subset-vars}
+biodiv_ferns_cent_raw <- 
+  biodiv_ferns_cent_dryad %>%
+  select(
+    grids, lat, long, # grid-cell ID, latitude, longitude
+    richness, pd_obs_z, fd_obs_z, rpd_obs_z, rfd_obs_z, # response vars
+    percent_apo, temp, precip, precip_season, lat_area  # independent vars
+  )
 ```
@@
 ```{r struc-full-data}
-biodiv_ferns_cent
+biodiv_ferns_cent_raw
 ```
@@
 * `precip_season`: seasonality in precipitation
-* `area`: rolling mean of area in 0.2 degree latitudinal bands
+* `lat_area`: rolling mean of area in 0.2 degree latitudinal bands
 
@@
 
+Next, check for missing data. If any rows are missing data, exclude them.
+
 ```{r check-missing}
-# These will return TRUE if all data are non-missing, 
-# or an error if any data are missing (NA)
-biodiv_ferns_cent %>%
-  assert(not_na, everything(), success_fun = success_logical)
+# Run a logical check to see if any rows have missing data 
+# Returns TRUE if there is no missing data, FALSE if there is missing data
+biodiv_ferns_cent_raw %>%
+  assert(
+    not_na, everything(), 
+    error_fun = error_logical, 
+    success_fun = success_logical)
+
+# At least one row has missing data. 
+# Check which.
+error_rows <-
+biodiv_ferns_cent_raw %>%
+  assert(not_na, everything(), error_fun = error_df_return) %>%
+  pull(index) %>%
+  unique()
+
+biodiv_ferns_cent_raw %>% slice(error_rows)
+
+# The missing data was due to one grid-cell that lacks environmental data.
+# Assume this must have been missing from WorldClim dataset.
+# Remove the row with missing data, then continue
+biodiv_ferns_cent_no_missing <-
+  biodiv_ferns_cent_raw %>%
+  slice(-error_rows)
+
+# Make sure the resulting dataframe has no missing data
+assert(biodiv_ferns_cent_no_missing, 
+       not_na, everything(), 
+       success_fun = success_logical)
 ```
-
-No missing data.
 
@@
 ```{r box-plots-full}
-biodiv_ferns_cent %>%
+biodiv_ferns_cent_no_missing %>%
   select(-lat, -long) %>%
@@
 
-There are a few points beyond the whiskers that may be outliers.
+There are a few points beyond the whiskers that may be outliers. In particular, there is one value of `percent_apo` > 60% that seems quite outside the normal range.
 
-Next, check for outliers using Grubb's test. Significant p-values indicate that the most extreme value is an outlier.
+Next, check for outliers using Grubb's test. Significant *p*-values indicate that the most extreme value is an outlier.
 
 ```{r grubbs-env}
-biodiv_ferns_cent %>%
+biodiv_ferns_cent_no_missing %>%
   select(-lat, -long) %>%
@@
 
-So the one percent apomictic > 60% appears to be an outlier. Perhaps this has a very small number of species, most of which happen to be apogamous?
+Grubb's test confirms that the percent apomictic value > 60% is an outlier.
+It also indicates that the highest values of the variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, and `rpd_obs_z` may be outliers.
 
-```{r check-apo-outlier}
-biodiv_ferns_cent %>%
+One reason for artificially extreme values could be undersampling of species. Let's check richness of these possible outlier grid-cells.
+
+```{r check-richness-apo}
+biodiv_ferns_cent_no_missing %>%
   slice_max(1, order_by = percent_apo) %>%
-  left_join(select(biodiv_ferns_cent, grids, richness)) %>%
   select(grids:long, percent_apo, richness)
@@
 
-Looks like that is the case---just three species in this grid cell! This data point should be excluded. Let's do that and check for outliers again.
+There are just three species in the grid-cell with > 60% apomictic taxa. This data point should be excluded.
+
+What about the other possible outliers?
+
+```{r check-richness-other}
+biodiv_ferns_cent_no_missing %>% 
+  slice_max(n = 2, order_by = fd_obs_z) %>%
+  select(grids:long, fd_obs_z, richness)
+
+biodiv_ferns_cent_no_missing %>% 
+  slice_max(n = 2, order_by = pd_obs_z) %>%
+  select(grids:long, pd_obs_z, richness)
+
+biodiv_ferns_cent_no_missing %>% 
+  slice_max(n = 2, order_by = rfd_obs_z) %>%
+  select(grids:long, rfd_obs_z, richness)
+
+biodiv_ferns_cent_no_missing %>% 
+  slice_max(n = 2, order_by = rpd_obs_z) %>%
+  select(grids:long, rpd_obs_z, richness)
+```
+
+They all come from the same two grid-cells, but don't have an extremely low number of species.
+
+So we will conservatively only remove the grid-cell with unusually high % apomictic taxa, then run Grubb's test again.
 
 ```{r remove-apo-outlier-check-grubbs}
-biodiv_ferns_cent %>% 
-  filter(percent_apo != max(percent_apo)) %>%
+biodiv_ferns_cent_no_outliers <-
+  biodiv_ferns_cent_no_missing %>% 
+  filter(percent_apo != max(percent_apo))
+
+biodiv_ferns_cent_no_outliers %>%
   pull(percent_apo) %>%
@@
 
-Check for normality with histograms and QQ-plots.
+Check for normality with histograms and QQ-plots after removing missing data and outliers.
 
-Inspect histograms.
+First inspect histograms:
 
 ```{r hist-env}
-biodiv_ferns_cent %>%
+biodiv_ferns_cent_no_outliers %>%
   select(-lat, -long) %>%
@@
 
-Inspect qq-plots.
+Next inspect qq-plots:
 
 ```{r qqplot-env}
-biodiv_ferns_cent %>%
+biodiv_ferns_cent_no_outliers %>%
   select(-lat, -long) %>%
@@
   stat_qq_line() +
-  facet_wrap(~variable, scales = "free") +
-  labs(title = "Full dataset")
+  facet_wrap(~variable, scales = "free")
 ```
@@
 ```{r check-for-zeros}
-biodiv_ferns_cent %>%
+biodiv_ferns_cent_no_outliers %>%
   mutate(rich_zero = richness == 0) %>%
@@
 
-Here we plot response variables (row labels) vs. indep variables (column labels) for each dataset, and add a trend line.
+Here we plot response variables (row labels) vs. indep variables (column labels) and add a trend line for each to get an idea of the possible model fits.
 
 ```{r plot-raw-data, fig.height = 8}
-ggplot(biodiv_ferns_cent, aes(x = .panel_x, y = .panel_y)) + 
+ggplot(biodiv_ferns_cent_no_outliers, aes(x = .panel_x, y = .panel_y)) + 
   geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
@@
 `temp` also show one if not quite as strongly.
-The other relationships appear to be linear or uncorrelated.
-
-```{r plot-raw-data-repro}
-# Remove the outlier first
-biodiv_ferns_cent %>% 
-  filter(percent_apo != max(percent_apo)) %>%
-ggplot(aes(x = .panel_x, y = .panel_y)) + 
-  geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
-  geom_smooth() +
-  facet_matrix(
-    cols = vars(percent_apo, temp, precip, precip_season),
-    rows = vars(pd_obs_z, rpd_obs_z))
-```
-
-As one would expect from the last plot, `temp` shows a (negative) hump shape.
-However, there aren't many points on the right..
-`percent_apo` seems more linear, especially for `rpd_obs_z`, but it's subtle.
-Anyways, we will probably keep the quadratic term for temperature but use a linear model for `percent_apo`.
+The other relationships appear to be linear or uncorrelated (the bend in the line for `percent_apo` vs.
+`richness` seems to be due to a relatively few number of points with > 30% apomictic taxa, but otherwise it looks linear).
 
@@
   richness ~ temp + I(temp^2) + precip + precip_season, 
-  data = biodiv_ferns_cent, family = poisson)
+  data = biodiv_ferns_cent_no_outliers, family = poisson)
 
-# Check ratio of residual variance to df (should be near 1 to meet poisson assumptions)
+# Check ratio of residual variance to df 
+# (should be near 1 to meet poisson assumptions)
 richness_poisson_mod$deviance / richness_poisson_mod$df.residual
@@
 
-That's very overdispersed. So we won't use poisson, but negative binomial instead.
+That's very overdispersed. So we won't use poisson for `richness`, but negative binomial instead.
+
+### Session information
+
+```{r sesh}
+sessionInfo()
+```

### 2021-09-06 ms/data_readme.Rmd
Update SI
- Add plot for latitudinal patterns
- Fix lumping of bioregions in SI plot
Filter estimated deer range by movement cost < 0.10
Make data exploration a pdf (SI2)

@@
     "japan_fern_phylogram_dryad_file",
-    "japan_fern_tree_dryad_file"
+    "japan_fern_tree_dryad_file",
+    "japan_fern_tree_uncollapsed_dryad_file"
     )
@@
   - japan_ferns_traits.csv: Trait matrix of native, non-hybrid ferns in Japan used for functional biodiversity analysis
-  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan
+  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan with clades consisting of identical OTUs collapsed to polytomies
+  - japan_ferns_tree_uncollapsed.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan with clades consisting of identical OTUs not collapsed
   - japan_ferns_tree.tre: Maximim-likelihood phylogeny of native, non-hybrid ferns in Japan
@@
 model including snow cover and forest type based on the 2003 survey data. Only
-estimated range with <0.05 movement cost was included.
+estimated range with < 0.10 movement cost was included.
 
@@
 
-MD5 checksum: `r tools::md5sum(here::here(japan_ferns_comm_full_file))`
+MD5 checksum: `r tools::md5sum(here::here(inext_res_file))`
 
@@
 japan_ferns_biodiv_dryad <- read_csv(here::here(japan_ferns_biodiv_dryad_file))
-# Check that no data are missing
-assert(japan_ferns_biodiv_dryad, not_na, everything(), success_fun = success_logical)
 ```
@@
 Biodiversity statistics of native, non-hybrid ferns and environmental variables
-in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. A
-single outlier grid-cell with extremely low taxon richness and extremely high %
-apomictic taxa was excluded. Climatic (temperature and preciptation) variables
-calculated as described for japan_climate.gpkg. Data used for spatial models
-in Nitta et al. 2021.
+in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. 
+Climatic (temperature and preciptation) variables
+calculated as described for japan_climate.gpkg. Includes one row with missing
+environmental data and one outlier for % apomixis that were removed prior to
+spatial modeling analysis in Nitta et al. 2021.
 
@@
 
-Missing data codes: No missing data.
+Missing data codes: No data entered between commas in CSV.
 
@@
 
+```{r japan-fern-dryad-uncollapsed-file}
+japan_fern_tree_uncollapsed_dryad <- ape::read.tree(here::here(japan_fern_tree_uncollapsed_dryad_file))
+```
+
+`r fs::path_file(japan_fern_tree_uncollapsed_dryad_file)` (contained in "results.zip"):
+Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
+Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r n_fossil_points` 
+fossil calibration points after @Testo2016a. Values at nodes
+are bootstrap support values calculated using 1000 replicates of ultrafast
+bootstrap [@Nguyen2015]. Clades consisting of identical OTUs not collapsed.
+
+Number of tips: `r ape::Ntip(japan_fern_tree_uncollapsed_dryad)`
+
+Missing data codes: No missing data.
+
+MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_uncollapsed_dryad_file))`
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
 ```{r japan-fern-dryad-file}
@@
 Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
-Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r
-n_fossil_points` fossil calibration points after @Testo2016a. Values at nodes
+Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r n_fossil_points` 
+fossil calibration points after @Testo2016a. Values at nodes
 are bootstrap support values calculated using 1000 replicates of ultrafast
-bootstrap [@Nguyen2015].
+bootstrap [@Nguyen2015]. Clades consisting of identical sequences have been
+collapsed to polytomies with zero branch length between OTUs.
 
@@
 
-2021-09-03
+2021-09-06
 
 Generate this README file.
+
+## References

### 2021-09-06 ms/data_readme.Rmd
Update MS
- Fix mentions of Honshu
- Change github URL

@@
 Code for analyzing the data is available on github:
-https://github.com/joelnitta/japan_ferns_biogeo
+https://github.com/joelnitta/japan_ferns_spatial_phy
 
@@
 
+Additional data used in analysis not included in current data package: Data 
+including DNA sequences, taxonomy, and breeding system data for native Japanese 
+ferns included in the Dryad data package of @Ebihara2019b (https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32)
+were used in the analysis.
+
 Checksums are 32-byte MD5 hashes generated with tools::md5sum() in R.
@@
 four grain sizes (grid-cells spanning Japan, each 0.1, 0.2, 0.3, or 0.4 degrees
-per side). The original list with latitude and longitude of each specimen is not
-included here to protect rare and endangered species. The 0.2 degree grain size
-was selected for further analysis based on redundancy (ratio of number of
-specimens to number of taxa per cell).
+per side). The 0.2 degree grain size was selected for further analysis based on 
+redundancy (ratio of number of specimens to number of taxa per cell).
 
@@
 Data files were generated from raw data (not included here) using scripts
-available at https://github.com/joelnitta/japan_ferns_biogeo, in particular
-https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/process_raw_data.R.
+available at https://github.com/joelnitta/japan_ferns_spatial_phy, in particular
+https://github.com/joelnitta/japan_ferns_spatial_phy/blob/main/R/process_raw_data.R.
 
@@
 to produce a labeled map of Japan. Some points were originally generated using
-script at https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/geocode.R
+script at https://github.com/joelnitta/japan_ferns_spatial_phy/blob/main/R/geocode.R
 
@@
 only "high". Code used to categorize areas available at
-https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/process_raw_data.R
+https://github.com/joelnitta/japan_ferns_spatial_phy/blob/main/R/process_raw_data.R
 
@@
 
-## References
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+References

### 2021-09-06 ms/manuscript.Rmd
Update MS
- Fix mentions of Honshu
- Change github URL

@@
 Better ways to characterize biological regions ("bioregions") are also available [e.g., @Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
-Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide particularly powerful insights into the processes structuring biodiversity and suggest conservation priorities.
+Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide powerful insights into the processes structuring biodiversity and suggest conservation priorities.
 
@@
 At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
-We calculated sampling redundancy (1 - richness/number of specimens) [@Garcillan2003] to quantify adequacy of sampling. 
+We calculated sampling redundancy [1 - richness/number of specimens; @Garcillan2003] to quantify adequacy of sampling. 
 Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
@@
 We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
-This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by.
+This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).
 
@@
 Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
-However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating.
+However, using this dataset alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating.
 Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
@@
 We trimmed the resulting ultrametric tree to only Japanese taxa.
-treePL requires branch lengths to be > 0, and it resets extremely small branches an arbitrary minimum positive length.
+treePL requires branch lengths to be > 0, and it sets extremely small branches to an arbitrary minimum positive length.
 This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
-We converted any clades in the original dated tree consisting of identical sequences to polytomies with zero branch length between OTUs.
+We converted any clades in the dated tree consisting of identical sequences to polytomies with zero branch length between OTUs.
 This tree was used for all subsequent analysis unless indicated otherwise.
@@
 *K* describes the ratio between the observed variance in a trait vs.
-the amount of variance expected under Brownian Motion (BM); lambda is a scaling parameter that transforms the phylogeny such that trait values fit those most closely expected under BM.
-Both lambda and *K* = 1 under BM; lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering). 
+the amount of variance expected under Brownian Motion (BM); lambda is a scaling parameter between 0 and 1 that transforms the phylogeny such that trait values fit those most closely expected under BM.
+Both lambda and *K* = 1 under BM; relatively lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering). 
 We measured *K* and lambda using the `r func("phylosig")` function in the R package `r pack("phytools")` [@Revell2012], which also calculates significance based on a randomization test for *K* and a likelihood ratio test for lambda.
@@
 We constructed a linear mixed model for each biodiversity metric dependent on the environmental variables ("environmental models"). 
-Initial inspection of the relationship between each biodiversity metric and temperature showed a hump-shaped pattern, so we also included a quadratic term for temperature only. 
+Initial inspection of the relationship between each biodiversity metric and temperature showed a hump-shaped pattern (Appendix S2), so we also included a quadratic term for temperature only. 
 Richness was fit with the negative binomial response family; others were Gaussian. 
@@
 The optimal value of *k* cannot be known *a-priori*.
-Previous vegetation zone schemes in Japan have typically included 5--6 zones [@Shimizu2014]; to enable approximate comparison with these, we tested values of *k* from 1 to 10, and selected the optimal value using the `optimal_phyloregion` function in the R package `r pack("phyloregion")`.
+Previous vegetation zone schemes in Japan have typically included 5--6 zones [@Shimizu2014]; to enable approximate comparison with these, we tested values of *k* from 1 to 10, and selected the optimal value using the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")`.
 
@@
 
-The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
+The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
 The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
@@
 
-Grid-cells with significantly low PD were almost all located on Honshu, the main island of Japan (`r figure("rand-diversity")`a). 
-A smaller number of cells had significantly high PD, and these were located mostly on smaller southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
-Grid-cells with significantly low RPD were mostly observed in southern Honshu, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
-Grid-cells with significantly high RPD are located both in the southern islands and northern Honshu and Hokkaido. 
+Grid-cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`a). 
+A smaller number of cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
+Grid-cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
+Grid-cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido. 
 Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`c, d).
@@
 Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu), bioregion 3 (Ryukyu Islands, excluding Yakushima), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion, and bioregion 1 extended slightly further south (`r figure("bioregions")`b; `r s_figure("dendrogram")`b).
+The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a).
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`b; `r s_figure("dendrogram")`b).
 The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
@@
 
-The overall range of Japanese deer in Japan has increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
+The range of Japanese deer in Japan increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
 The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`a, `r s_figure("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants [@Fukaya2020].
+The latitudinal hump-shaped pattern of species richness observed in ferns here (Figs.
+`r figure_num("raw-diversity")`a, `r s_figure_num("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020].
 Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`a).
 Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
-A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a]. 
+A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a; @Suissa2021].
 Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a]. 
-Another reason for the greater richness in southern Honshu compared to smaller southern islands is that southern Honshu harbors high elevational variation that may support a wider range of taxa than is present in the southern islands.
+Another reason for the greater richness on the main islands compared to southern islands in Japan is that the main islands harbor higher elevational variation that may support a wider range of taxa than is present in the southern islands.
 
@@
 This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
-The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis, and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity.
-Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands.
+The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig")`, `r s_table_num("phy-sig-binary")`), and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity.
+Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands (Figs.
+`r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`).
 This is clearly due to the small land area of these islands, which are the only areas with a subtropical climate in Japan.
 
-Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and in southern islands, and the predominance of short branches in the southern half of Honshu, and particularly on Kyushu (`r figure("rand-diversity")`b).
+Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`b).
 The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a).
-The preponderance of short branches in southern Honshu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).
-
-Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (`r figure("modelfit")`g), and to a slightly lesser extent SES of PD (`r figure("modelfit")`f).
-Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in southern Honshu (`r s_figure("genus-map")`).
+The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).
+
+Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (Figs.
+`r figure_num("coeff")`g, `r figure_num("modelfit")`g), and to a slightly lesser extent SES of PD (Figs.
+`r figure_num("coeff")`f, `r figure_num("modelfit")`f).
+Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (`r s_figure("genus-map")`).
 This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns.
-We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`b).
+We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`).
 While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`).
 Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca.
-34.8 °N latitude and increases with temperature in Japan (their study was limited to Honshu and did not include the southern islands).
+34.8 °N latitude and increases with temperature in Japan (their study was limited to the main islands and did not include the smaller southern islands).
 @Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid. 
-As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses.
+As fern gametophytes are only a single cell-layer in thickness with no cuticle or stomata, it is expected they would be particularly susceptible to freezing damage induced by larger cells with higher water content; however, no experimental studies have tested these hypotheses to our knowledge.
 
@@
 This suggests that ferns are useful as bioindicators in Japan. 
-Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from the rest of Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions.
-The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits bioregions 2 and 3, there is some overlap between them on the Amami-Oshima Isl.
+Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions.
+The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, there is some overlap between them on Yakushima and Amami Isl.
 (`r figure("bioregions")`b).
-The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind.
+The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
 Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu islands; wetter) from bioregion 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).
@@
 Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`b).
-This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, where deer expansion has been intense (`r s_figure("protected-map")`a).
+This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("protected-map")`a).
 This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.
@@
 
-The extreme evolutionary distinctiveness of southern islands is both a major feature of the Japanese fern flora and a challenge for analysis. 
-The tropical lineages occurring only in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on Honshu (`r figure("rand-diversity")`a).
+The extreme evolutionary distinctiveness of the southern islands is both a major feature of the Japanese fern flora and a challenge for analysis. 
+The tropical lineages occurring only in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on the main islands (`r figure("rand-diversity")`a).
 One way to account for this bimodal distribution pattern would be to use a restricted null model that does not allow all taxa to disperse anywhere within the study area [@Mishler2020]. 
@@
 Therefore, it seems reasonable to assume that ferns are capable of dispersing across the entire study area.
-Furthermore, it should be noted that the southern islands are all low-altitude (not exceeding 1000 m).
+Furthermore, it should be noted that the southern islands are all low-altitude (not exceeding 1,000 m).
 If high elevation habitats were available, the southern islands would be expected to share more species in common with areas further north.
-For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (> 3000 m), shares many fern species with Honshu that are not found in southern Japanese islands.
-This supports that the sharp taxonomic turnover between southern islands and Honshu is more likely due to ecological factors than dispersal limitation.
+For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (> 3,000 m), shares many fern species with the main Japanese islands that are not found in southern Japanese islands.
+This supports that the sharp taxonomic turnover between southern and main islands is more likely due to ecological factors than dispersal limitation.
 
@@
 
-Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_biogeo. 
-A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_biogeo.
+Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy. 
+A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_spatial_phy.
 Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.

### 2021-09-06 ms/SI.Rmd
Fix SI K plot and bioregions plot

@@
   scale_color_manual(values = c("TRUE" = "#E31A1C", "FALSE" = "black")) +
+  scale_y_continuous(expand = expansion(add = c(0.025,0.05))) +
+  scale_x_continuous(breaks = c(2,4,6,8,10)) +
   labs(
@@
   theme(
-    legend.position = "none"
+    legend.position = "none",
+    panel.grid.minor = element_blank()
   )
@@
 (b), phylogenetic bioregions.
-Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal; see Methods).
+Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).
 
@@
   geom_point(size = 1, alpha = 0.7) +
-  scale_color_manual(values = bioregion_cols) +
+  scale_color_manual(
+    values = bioregion_cols,
+    guide = guide_legend(
+      direction = "horizontal",
+      title.position = "top",
+      label.position = "left")
+  ) +
   scale_y_continuous(labels = function(x) x/10) +
-  labs(y = "Temperature (°C)", x = "Precipitation (mm)", color = "Cluster", subtitle = "Taxonomic bioregions") +
+  labs(y = "Temperature (°C)", x = "Precipitation (mm)", color = "Bioregion", subtitle = "Taxonomic bioregions") +
   theme(legend.position = "bottom")
@@
   geom_point(size = 1, alpha = 0.7) +
-  scale_color_manual(values = bioregion_cols) +
+  scale_color_manual(
+    values = bioregion_cols,
+    guide = guide_legend(
+      direction = "horizontal",
+      title.position = "top",
+      label.position = "left")) +
   scale_y_continuous(labels = function(x) x/10) +
-  labs(y = "Temperature (°C)", x = "Precipitation (mm)", color = "Cluster", subtitle = "Phylogenetic bioregions") +
+  labs(y = "Temperature (°C)", x = "Precipitation (mm)", color = "Bioregion", subtitle = "Phylogenetic bioregions") +
   theme(legend.position = "bottom")
@@
 Points shown with partial transparency; darker areas indicate overlapping points.
-Bioregions with three or fewer grid-cells are lumped into the "Other" category.
+Bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.
 

### 2021-09-06 ms/SI.Rmd
Load some latex packages in SI header to fix kable problem
fixes errror: "! Undefined control sequence. <recently read> \cellcolor"

see https://stackoverflow.com/questions/66522775/is-it-possible-to-rmarkdownrender-a-pdf-document-for-kable-tables-with-styling

@@
   - \usepackage{float}
+  - \usepackage{booktabs}
+  - \usepackage{colortbl}
   - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother

### 2021-09-07 ms/SI.Rmd
Update SI and data readme

@@
 **`r s_table("moran")`**.
-Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.
+Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.
 Significant *I* > 0 indicates autocorrelation.

### 2021-09-07 ms/data_readme.Rmd
Update SI and data readme

@@
 
+- _targets.tar.gz: Tarball (compressed folder) including all workflow results produced by R targets package
 - ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and phylogenetic tree 
@@
 DATA-SPECIFIC INFORMATION
+
+\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+
+_targets.tar.gz: Tarball (compressed folder) including all workflow results
+produced by R targets package. This is provided to enable inspection of workflow
+steps without running the entire workflow from the beginning. To use it, unpack
+the tar achive with the command "tar -xzf _targets.tar.gz". Then, in R, the
+"tar_load()" function in the R package "targets" can be used to load any
+workflow step (target) defined in _targets.R
+(https://github.com/joelnitta/japan_ferns_spatial_phy/blob/main/_targets.R). For
+more information on the structure of the _targets folder and how to use it, see
+https://github.com/ropensci/targets.
+
+MD5 checksum: FIXME (add manually, since this can't be calculated from inside the targets workflow)
 
@@
 
-2021-09-06
+2021-09-07
 

### 2021-09-07 ms/SI.Rmd
Update SI
Add header on right side
Move grubbs_tidy() to data_exploration.Rmd

@@
 header-includes: 
-  - \usepackage{float}
-  - \usepackage{booktabs}
-  - \usepackage{colortbl}
-  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
+  - \usepackage{float} # keep figures from floating away
+  - \usepackage{booktabs} # need for kable tables
+  - \usepackage{colortbl} # need for kable tables
+  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
+  - \usepackage{fancyhdr} # enable header
+  - \pagestyle{fancy}  # enable header
+  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
+  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
+  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S1} # add header on right side
+  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
+  - \cfoot{\thepage} # footer
 editor_options: 

### 2021-09-07 ms/data_exploration.Rmd
Update MS
- Fix mentions of Honshu
- Change github URL
Update SI
Add header on right side
Move grubbs_tidy() to data_exploration.Rmd

@@
 header-includes: 
-  - \usepackage{float}
-  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
+  - \usepackage{float} # keep figures from floating away
+  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
+  - \usepackage{fancyhdr} # enable header
+  - \pagestyle{fancy}  # enable header
+  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
+  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
+  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S2} # add header on right side
+  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
+  - \cfoot{\thepage} # footer
 ---
@@
 
-The code here is included in the repo at https://github.com/joelnitta/japan_ferns_biogeo.
+The code here is included in the repo at https://github.com/joelnitta/japan_ferns_spatial_phy.
 
-Let's get started by loading packages and tidying our namespace.
+## Setup
 
-```{r load-packages, message = FALSE}
+Let's get started by loading packages and tidying our namespace:
+
+```{r load-packages, message = FALSE, warning = FALSE}
 # Load packages
@@
 
+Also define a custom function we'll use later:
+
+```{r grubbs-tidy}
+#' Run Grubb's test for outliers
+#'
+#' Similar to `outliers::grubbs.test()`, but returns the results in a tidy 
+#' dataframe
+#'
+#' @param x a numeric vector for data values. 
+#' @param type Integer value indicating test variant. 
+#' See `outliers::grubbs.test()`
+#' @param ... Other arguments passed to `outliers::grubbs.test()`
+#'
+#' @return Tibble
+grubbs_tidy <- function(x, type = 10, ...) {
+  outliers::grubbs.test(x = x, type = type, ...) %>%
+    broom::tidy() %>%
+    dplyr::mutate(stat_name = c("g", "u")) %>%
+    tidyr::pivot_wider(
+      values_from = "statistic",
+      names_from = "stat_name") %>%
+    janitor::clean_names() %>%
+    dplyr::select(g, u, p_value, method, alternative)
+}
+```
+
+## Load data
+
 The dataset includes biodiversity metrics and environmental data measured on native, non-hybrid ferns of Japan.
@@
 
-```{r load-data}
+```{r load-data, message = FALSE}
 # Load the data from the targets workflow
-tar_load(biodiv_ferns_cent_dryad)
+# tar_load(biodiv_ferns_cent_dryad)
 
@@
 # and read in the data from there
-# biodiv_ferns_cent_dryad <- read_csv("japan_ferns_biodiv.csv")
+biodiv_ferns_cent_dryad <- read_csv(here::here("results/dryad_files/japan_ferns_biodiv.csv"))
 ```
@@
 ```
-
-### Structure of the data
 
@@
 
-### Check for missing data
+## Check for missing data
 
-Next, check for missing data. If any rows are missing data, exclude them.
+Check for missing data. If any rows are missing data, exclude them.
+
+The next code chunk runs a logical check to see if any rows have missing data. It returns `TRUE` if there is no missing data, `FALSE` if there are any missing data:
 
 ```{r check-missing}
-# Run a logical check to see if any rows have missing data 
-# Returns TRUE if there is no missing data, FALSE if there is missing data
 biodiv_ferns_cent_raw %>%
@@
     success_fun = success_logical)
+```
 
-# At least one row has missing data. 
-# Check which.
+At least one row has missing data. Let's see what variables are missing.
+
+```{r check-missing-which}
+# Get the row numbers of rows with missing data
 error_rows <-
@@
 
-biodiv_ferns_cent_raw %>% slice(error_rows)
+# Inspect the rows with missing data
+biodiv_ferns_cent_raw %>% 
+  slice(error_rows) %>% 
+  select(grids, where(~any(is.na(.x))))
+```
 
-# The missing data was due to one grid-cell that lacks environmental data.
-# Assume this must have been missing from WorldClim dataset.
-# Remove the row with missing data, then continue
+The missing data was due to one grid-cell that lacks environmental data. This must be missing from the WorldClim dataset. Let's remove it, then continue.
+
+```{r remove-missing}
+# Exclude rows with missing data
 biodiv_ferns_cent_no_missing <-
@@
 
-# Make sure the resulting dataframe has no missing data
+# Double check that the resulting dataframe has no missing data
 assert(biodiv_ferns_cent_no_missing, 
@@
 
-### Check for outliers
+## Check for outliers
 
@@
 
-### Normality
+## Check for normality
 
@@
 
-### Check for zeros in count data
+## Check for zeros in count data
 
@@
 
-### Plot raw data
+## Plot raw data
 
@@
 
-### Check for overdispersion in richness
+## Check for overdispersion in richness
 
@@
 
-### Session information
+## Session information
 

### 2021-09-07 ms/SI.Rmd
Make binary trait sig table long

@@
   ) %>%
-  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
-  kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
+  kbl(booktabs = TRUE, escape = FALSE, format = "latex", longtable = TRUE) %>%
+  kable_styling(latex_options = c("striped", "scale_down", "HOLD_position", "repeat_header"))
 ```

### 2021-09-07 ms/SI.Rmd
Add scalebars to maps
take snapshot

--- 2021-09-07 ms/SI.Rmd
@@
 ---
-title: "Appendix S1: Supporting Information"
+title: "Appendix S1: Supporting Tables and Figures"
 mainfont: Roboto-Regular.ttf
@@
 library(RColorBrewer)
+library(ggspatial) # for scalebars on maps
+
 # Load functions and variables only used when rendering Rmds
@@
   # format for kable/latex
-  mutate(trait = str_replace_all(trait, "_", "\\\\_")) %>%
+  mutate(
+    trait = str_replace_all(trait, "_", "\\\\_"),
+    D = round(D, 3)) %>%
   rename(
@@
 Presence or absence of indusium was originally a binary trait; other traits (margin, shape, texture) are qualitative traits scored in binary format.
+Values of *D* rounded to three decimal places.
 
@@
 e <- plot_env("e", "Temperature", "temp", "°C", function(x) x/10)
-f <- plot_env("f", "Temp. season.", "temp_season")
+f <- plot_env("f", "Temp. season.", "temp_season") +
+  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))
 
@@
   annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)**</span>", hjust = "inward", vjust = "inward", color = "white") +
+  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
   map_theme() +
@@
   scale_fill_manual(values = canape_cols) +
+  annotation_scale(location = "tl", width_hint = 0.2) +
   map_theme() +
@@
 e <- plot_genus("e", "Pteris")
-f <- plot_genus("f", "Diplazium")
+f <- plot_genus("f", "Diplazium") +
+  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
 
@@
 c <- make_protected_plot("c", "FD", fd_obs_p_upper > 0.95)
-d <- make_protected_plot("d", "PE", pe_obs_p_upper > 0.95)
+d <- make_protected_plot("d", "PE", pe_obs_p_upper > 0.95) +
+  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))
 
@@
 c <- make_deer_plot("c", "FD", fd_obs_p_upper > 0.95)
-d <- make_deer_plot("d", "PE", pe_obs_p_upper > 0.95)
+d <- make_deer_plot("d", "PE", pe_obs_p_upper > 0.95) +
+  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))
 

### 2021-09-07 ms/manuscript.Rmd
Add scalebars to maps
take snapshot

@@
 library(RColorBrewer)
+library(ggspatial) # for scalebars on maps
 
@@
 
-**Aim** To characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.
-
-**Location** Japan
-
-**Taxon** Ferns
-
-**Methods** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
+**Aim:** To characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.
+
+**Location:** Japan
+
+**Taxon:** Ferns
+
+**Methods:** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
 We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode.
@@
 
-**Results** Three major patterns in the spatial distribution of biodiversity were observed: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes.
+**Results:** Three major patterns in the spatial distribution of biodiversity were observed: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes.
 Grid-cells were grouped in to `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions. 
@@
 
-**Main conclusions** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan. 
+**Main conclusions:** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan. 
 The different distributions of each diversity type reflect the major bioregions. 
@@
 <!-- may include title words -->
-**Key words:** apomixis, biodiversity, biogeography, CANAPE, conservation, ferns, Japan, phylogenetic diversity.
+**Keywords:** apomixis, biodiversity, biogeography, CANAPE, conservation, ferns, Japan, phylogenetic diversity.
 
@@
 We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the `r func("phylo.d")` function in the R package `r pack("caper")` [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
+Values of *D* range from 0 under Brownian motion to 1 under random evolution, but can exceed this range in cases of extreme clumping or overdispersion, respectively.
 These methods for phylogenetic signal cannot use a tree with zero-length polytomies, so we used the original dated tree from treePL.
@@
 
-# Sum the number of traits with D < 0
-n_negative_d <-
+# Sum the number of traits with D <= 0.5 and not random
+n_d_signif <-
   binary_sig_results %>%
-  mutate(negative_d = ifelse(D <= 0, TRUE, FALSE)) %>%
-  pull(negative_d) %>%
-  sum
+  filter(prob_random < 0.05, D <= 0.5) %>%
+  pull(trait) %>%
+  n_distinct()
 ```
@@
 All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.1, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
-`r n_negative_d` out of `r nrow(binary_sig_results)` of continuous traits (coded as binary) showed as much or more phylogenetic signal as expected under BM (*D* =< 0); others had *D* > 0 (`r s_table("phy-sig-binary")`).
+`r n_d_signif` out of `r nrow(binary_sig_results)` of qualitative traits (coded as binary) showed evidence of phylogenetic signal (*D* ≤ 0.5, significantly different from random) (`r s_table("phy-sig-binary")`).
 
@@
   ) +
+  annotation_scale(location = "br", width_hint = 0.2) +
   map_theme() +
@@
 ```{r raw-diversity-make}
-# Use thinner coast lines for this plot
-# it's smaller, so the grey lines appear relatively thicker
-coast_lwd_old <- coast_lwd
-coast_lwd <- 0.05
-
 # Richness
@@
   scale_y_continuous(breaks = c(25, 35, 45)) +
+  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
   map_theme() +
@@
   plot_layout(nrow = 1)
-
-# Reset coastline width
-coast_lwd <- coast_lwd_old
 ```
@@
   scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
+  annotation_scale(location = "br", width_hint = 0.2) +
   map_theme() +
@@
   scale_fill_manual(values = canape_cols) +
+  annotation_scale(location = "br", width_hint = 0.2) +
   map_theme() +
@@
   scale_fill_manual(values = bioregion_cols) +
+  annotation_scale(location = "br", width_hint = 0.2) +
   map_theme() +

### 2021-09-07 ms/manuscript.Rmd
Include in-line figs for original MS submission to J. Biogeo.

@@
     number_sections: no
-    fig_caption: yes
+    fig_caption: no
     latex_engine: xelatex
@@
 
+`r pagebreak_doc()`
+
+```{r japan-map-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+# Print plot if rendering to doc on initial submission
+japan_map_plot
+cat("\n\n", figure("japan-map"))
+```
+
+`r pagebreak_doc()`
+
+```{r raw-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
+raw_diversity_plot
+cat("\n\n", figure("raw-diversity"))
+```
+
+`r pagebreak_doc()`
+
+```{r rand-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 9, eval = params$doc_type == "doc", results = "asis"}
+rand_diversity_plot
+cat("\n\n", figure("rand-diversity"))
+```
+
+`r pagebreak_doc()`
+
+```{r coeff-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+coeff_plot
+cat("\n\n", figure("coeff"))
+```
+
+`r pagebreak_doc()`
+
+```{r modelfit-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+modelfit_plot
+cat("\n\n", figure("modelfit"))
+```
+
+`r pagebreak_doc()`
+
+```{r endemism-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+endemism_plot
+cat("\n\n", figure("endemism"))
+```
+
+`r pagebreak_doc()`
+
+```{r bioregions-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+bioregions_plot
+cat("\n\n", figure("bioregions"))
+```
+
+`r pagebreak_doc()`
+
+```{r div-by-tax-region-doc, fig.cap = "", fig.width = 7, fig.height = 7, eval = params$doc_type == "doc", results = "asis"}
+div_by_tax_region_plot
+cat("\n\n", figure("div-by-tax-region"))
+```
+
+`r pagebreak_doc()`
+
+```{r conserv-status-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
+conserv_status_plot
+cat("\n\n", figure("conserv-status"))
+```
+
+`r pagebreak_doc()`
+
 ## Data availability statement

### 2021-09-07 R/ms_functions.R

@@ -124,7 +124,7 @@ func <- function(x) {glue::glue("'{x}'")}
 #' @param rmd_params Parameters set in YAML header or rmarkdown::render(). 
 #' Must include `doc_type` (either 'doc' for MS Word or 'pdf' for PDF output)
 #' 
-pagebreak_doc <- function(rmd_params = params) {ifelse(rmd_params$doc_type == "doc", run_pagebreak(), return(""))}
+pagebreak_doc <- function(rmd_params = params) {ifelse(rmd_params$doc_type == "doc", return("\\clearpage"), return(""))}
 
 #' Pagebreak for PDF only
 #' 

### 2021-09-07 ms/manuscript.Rmd
Update MS

@@
 Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently.
-These datasets and tools now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity and phylogenetic endemism. 
+These datasets and tools now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity [@Faith1992] and phylogenetic endemism [@Rosauer2009]. 
 Better ways to characterize biological regions ("bioregions") are also available [e.g., @Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
@@
 All qualitative traits were scored in binary format; `r eg`, a single trait with three states was formatted as three binary traits. 
-From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), and removed any that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.
+From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), calculated mean values for continuous traits, and removed any traits that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.
 
@@
 This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
-We converted any clades in the dated tree consisting of identical sequences to polytomies with zero branch length between OTUs.
+We converted any clades in the dated tree consisting of identical sequences to polytomies with the node at 0 mya (`r ie`, the present).
 This tree was used for all subsequent analysis unless indicated otherwise.
@@
 We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the `r func("phylo.d")` function in the R package `r pack("caper")` [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
-Values of *D* range from 0 under Brownian motion to 1 under random evolution, but can exceed this range in cases of extreme clumping or overdispersion, respectively.
+Values of *D* range from 0 under BM to 1 under random evolution, but can exceed this range in cases of extreme clumping or overdispersion, respectively.
 These methods for phylogenetic signal cannot use a tree with zero-length polytomies, so we used the original dated tree from treePL.
@@
 All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.1, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
-`r n_d_signif` out of `r nrow(binary_sig_results)` of qualitative traits (coded as binary) showed evidence of phylogenetic signal (*D* ≤ 0.5, significantly different from random) (`r s_table("phy-sig-binary")`).
+`r n_d_signif` out of `r nrow(binary_sig_results)` of binary traits (including qualitative traits coded as binary) showed evidence of phylogenetic signal (*D* ≤ 0.5, significantly different from random) (`r s_table("phy-sig-binary")`).
 

### 2021-09-08 ms/manuscript.Rmd
Fix missing label in y-axis of Fig 4
- adds check for missing labels

@@
   tibble(
-    term = c("area", "I(temp^2)", "null_model", "percent_apo", "precip", "precip_season", "temp"),
+    term = c("lat_area", "I(temp^2)", "null_model", "percent_apo", "precip", "precip_season", "temp"),
     term_print = c("Area", "Temp^2", "Null model", "% apomictic", "Precip.", "Precip.\nseasonality", "Temp.")
@@
   ) %>%
-  ungroup()
+  ungroup() %>%
+  assert(not_na, matches("print|plain"))
 

### 2021-09-08 ms/manuscript.Rmd
Fix link to docker image

@@
 Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy. 
-A Docker image to run the code is available at https://hub.docker.com/repository/docker/joelnitta/japan_ferns_spatial_phy.
-Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/ADD_DOI_HERE_WHEN_ASSIGNED.
+A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
+Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.
 

### 2021-09-09 ms/data_readme.Rmd
Update data readme

@@
     "ms_functions",
+    "ebihara_2019_zip_file",
+    "ftol_zip_file",
     "ja_climate_data_file",
@@
     "japan_fern_tree_dryad_file",
-    "japan_fern_tree_uncollapsed_dryad_file"
+    "japan_fern_tree_uncollapsed_dryad_file",
+    "refs_yaml"
     )
@@
 
-This README.txt file was generated on `r format(Sys.time(), '%d %B, %Y')`
+This README.txt file was generated on `r format(Sys.time(), '%d %B, %Y')` by Joel Nitta
 
@@
 - _targets.tar.gz: Tarball (compressed folder) including all workflow results produced by R targets package
-- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and phylogenetic tree 
 - japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
@@
 
-Additional data used in analysis not included in current data package: Data 
-including DNA sequences, taxonomy, and breeding system data for native Japanese 
-ferns included in the Dryad data package of @Ebihara2019b (https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32)
-were used in the analysis.
+Additional data used in analysis not included in current data package:
+
+- doi_10.5061_dryad.4362p32__v4.zip: Data including DNA sequences, taxonomy, and 
+breeding system data for native Japanese ferns [@Ebihara2019b] downloaded from 
+https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32. 
+MD5 checksum `r tools::md5sum(here::here(ebihara_2019_zip_file))`
+- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment 
+and phylogenetic tree (Nitta et al., in prep.) downloaded from 
+https://figshare.com/s/87b8ec102303bf9f3212.
+MD5 checksum `r tools::md5sum(here::here(ftol_zip_file))`
+FIXME: replace with DOI when available
 
@@
 MD5 checksum: FIXME (add manually, since this can't be calculated from inside the targets workflow)
-
-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
-
-ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and
-phylogenetic tree (Nitta et al., in prep.). For more information, see README
-file contained in zipped archive.
-
-MD5 checksum: `r tools::md5sum(here::here("data/ftol_data_release_v0.0.1.zip"))`
 
@@
       ~col, ~desc,
-      "dataset", "Dataset name. 'occ_point_data_ferns_unfiltered' indicates data includes all specimens; 'occ_point_data_ferns' indicates data only includes specimens after filtering",
+      "dataset", "Dataset name. 'occ_point_data_ferns_unfiltered' indicates data includes all specimens, 'occ_point_data_ferns' indicates data only includes specimens after filtering",
       "variable", "Variable. 'n_taxa' indicates number of taxa, 'n_specimens' indicates number of specimens",
@@
 Categorical trait codes as follows:
+
 - 0 = absent
@@
 License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html). All polygons
-were combined into a single polygon and CRS set to JGD2000.
+were combined into a single polygon and CRS set to JGD2000 using code available at
+https://github.com/joelnitta/japan_ferns_spatial_phy/blob/main/R/process_raw_data.R.
 
@@
 
-2021-09-07
+2021-09-09
 

### 2021-09-21 ms/manuscript.Rmd
WIP convert to AJB format

@@
     number_sections: no
-    reference_docx: template.docx
+    reference_docx: ajb_template.docx
   bookdown::pdf_document2:
@@
 
+Joel H. Nitta^1,5^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^
+
+^1^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, Tokyo, Japan
+
+^2^University and Jepson Herbaria, and Department of Integrative Biology, University of California, Berkeley, CA, USA
+
+^3^Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan
+
+^4^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
+
+^5^Author for correspondence: joelnitta@gmail.com
+
+Manuscript received _______; revision accepted _______.
+
 Short running title: Spatial phylogenetics of Japanese ferns
-
-Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^
-
-^1^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, Tokyo, Japan
-
-^2^University and Jepson Herbaria, and Department of Integrative Biology, University of California, Berkeley, CA, USA
-
-^3^Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan
-
-^4^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
-
-Author for correspondence:  
-*Joel H. Nitta*  
-*Email: joelnitta@gmail.com*
-
-## Acknowledgements
-
-Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
-Shawn Laffan provided advice about CANAPE.
-This study was supported by JSPS KAKENHI Grant Number 16H06279.
 
@@
 
-**Aim:** To characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.
-
-**Location:** Japan
-
-**Taxon:** Ferns
+**Premise:** Biodiversity is often measured with a single metric, species richness.
+However, this ignores evolutionary history and is not sufficient for making conservation decisions.
+Here, we use a densely sampled dataset to characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.
 
@@
 
-**Main conclusions:** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan. 
+**Conclusions:** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan. 
 The different distributions of each diversity type reflect the major bioregions. 
@@
 \newpage
-
-## Introduction
 
@@
 
-## Materials and Methods
+## MATERIALS AND METHODS
 
@@
 
-## Results
+## RESULTS
 
@@
 Taxonomic bioregions 3 and 4 have higher SES values of PD, RPD, FD, and RFD than bioregions 1 and 2 (`r figure("div-by-tax-region")`). 
-Bioregions 3 and 4 also have consistently high PE *p*-scores, whereas bioregions 1 and 2 have a much wider variation. 
+Bioregions 3 and 4 also have consistently high PE *P*-scores, whereas bioregions 1 and 2 have a much wider variation. 
 Bioregion 2 has lower RPD and RFD than bioregion 1. 
@@
 
-## Discussion
+## DISCUSSION
 
@@
 
-### Limits of the study and future outlook
+### CONCLUSIONS
 
@@
 
-There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research. 
+Our study integrates diverse, thoroughly collected datasets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns.
+We end by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.
+
 First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
@@
 
+## Acknowledgements
+
+Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
+Shawn Laffan provided advice about CANAPE.
+This study was supported by JSPS KAKENHI Grant Number 16H06279.
+
+## Author contributions
+
+J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.
+
+## Data availability
+
+Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy. 
+A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
+Data files needed to run the analysis and selected results files are available from Figshare at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.
+
+## Literature cited
+<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
+<div id="refs"></div>
+
 \newpage
@@
   "pd_obs_z", "SES of PD",
-  "pe_obs_p_upper", "PE *p*-value",
+  "pe_obs_p_upper", "PE *P*-value",
   "rfd_obs_z", "SES of RFD",
@@
       values = c("TRUE" = "grey20", "FALSE" = "grey70"),
-      labels = c("TRUE" = "*p* < 0.05", "FALSE" = "*p* > 0.05")) +
+      labels = c("TRUE" = "*P* < 0.05", "FALSE" = "*P* > 0.05")) +
     labs(
@@
 (a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa).
-Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-value indicates probability of no difference between the full and null model.
+Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *P*-value indicates probability of no difference between the full and null model.
 Name of response variable and model type indicated above each subplot.
@@
   labs(
-    y = "PE *p*-score",
+    y = "PE *P*-score",
     x = "Cluster",
@@
 (d), SES of relative FD (RFD).
-(e), *p*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
+(e), *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
 Only top four clusters with the most sites shown.
@@
 For (b), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).
-
-`r pagebreak_doc()`
-
-```{r japan-map-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-# Print plot if rendering to doc on initial submission
-japan_map_plot
-cat("\n\n", figure("japan-map"))
-```
-
-`r pagebreak_doc()`
-
-```{r raw-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
-raw_diversity_plot
-cat("\n\n", figure("raw-diversity"))
-```
-
-`r pagebreak_doc()`
-
-```{r rand-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 9, eval = params$doc_type == "doc", results = "asis"}
-rand_diversity_plot
-cat("\n\n", figure("rand-diversity"))
-```
-
-`r pagebreak_doc()`
-
-```{r coeff-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-coeff_plot
-cat("\n\n", figure("coeff"))
-```
-
-`r pagebreak_doc()`
-
-```{r modelfit-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-modelfit_plot
-cat("\n\n", figure("modelfit"))
-```
-
-`r pagebreak_doc()`
-
-```{r endemism-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-endemism_plot
-cat("\n\n", figure("endemism"))
-```
-
-`r pagebreak_doc()`
-
-```{r bioregions-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-bioregions_plot
-cat("\n\n", figure("bioregions"))
-```
-
-`r pagebreak_doc()`
-
-```{r div-by-tax-region-doc, fig.cap = "", fig.width = 7, fig.height = 7, eval = params$doc_type == "doc", results = "asis"}
-div_by_tax_region_plot
-cat("\n\n", figure("div-by-tax-region"))
-```
-
-`r pagebreak_doc()`
-
-```{r conserv-status-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
-conserv_status_plot
-cat("\n\n", figure("conserv-status"))
-```
-
-`r pagebreak_doc()`
-
-## Data availability statement
-
-Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy. 
-A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
-Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.
-
-## References
-<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
-<div id="refs"></div>
-
-## Biosketch
-
-Joel H. Nitta's research focuses on patterns and processes of diversification in ferns. For more information, see https://joelnitta.com.
-
-Author contributions: J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.

### 2021-09-21 ms/SI.Rmd
Update MS

@@
   - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
-  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S1} # add header on right side
+  - \fancyhead[RE,RO]{Nitta et al.---\textit{American Journal of Botany} 2021 -- Appendix S1} # add header on right side
   - \fancyhead[LO,LE]{} # suppress automatic section header on left side
@@
 **`r s_figure("sampling-curve")`**.
-Species collection curve for the ferns of Japan fit with the `r pack("iNEXT")` package [@Hsieh2016].
+Species collection curve for the ferns of Japan.
+Curve fit with the `r pack("iNEXT")` package [@Hsieh2016].
 Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line).

### 2021-09-21 ms/data_exploration.Rmd
Update MS

@@
 ---
-title: "Appendix S2: Data Exploration"
+title: "Appendix S2: Exploratory Data Analysis"
 mainfont: Roboto-Regular.ttf
@@
   - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
-  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S2} # add header on right side
+  - \fancyhead[RE,RO]{Nitta et al.---\textit{American Journal of Botany} 2021 -- Appendix S2} # add header on right side
   - \fancyhead[LO,LE]{} # suppress automatic section header on left side
@@
 
-Here we perform initial data exploration to determine the appropriate models to fit to the data.
+Here we perform exploratory data analysis to determine the appropriate models to fit to the data.
 
@@
 # Load the data from the targets workflow
-# tar_load(biodiv_ferns_cent_dryad)
+tar_load(biodiv_ferns_cent_dryad)
 
@@
 # and read in the data from there
-biodiv_ferns_cent_dryad <- read_csv(here::here("results/dryad_files/japan_ferns_biodiv.csv"))
+# biodiv_ferns_cent_dryad <- read_csv(here::here("japan_ferns_biodiv.csv"))
 ```

### 2021-09-21 ms/manuscript.Rmd
Update MS

@@
   bookdown::word_document2:
-    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
+    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
     toc: no
@@
     latex_engine: xelatex
-    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
+    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
 params:
@@
 
-## Abstract
+**Abstract**
 
@@
 
-## MATERIALS AND METHODS
+# Materials and methods
 
@@
 
-### Study Site
-
-Our data include all ferns known to natively occur within the boundaries of Japan (20°N to 45°N, 122°E to 153°E). 
-Japan consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
+## Study Site
+
+Our data include all ferns known to natively occur within the boundaries of Japan . 
+Japan (20°N to 45°N, 122°E to 153°E) consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
 Most of the main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ogasawara archipelago consists of oceanic islands that have never been connected to the mainland.
@@
 
-### Occurrence data
+## Occurrence data
 
@@
 
-### Morphological trait data
+## Morphological trait data
 
@@
 
-### Reproductive mode data
+## Reproductive mode data
 
@@
 
-### Environmental data
+## Environmental data
 
@@
 
-### Phylogenetic analysis
+## Phylogenetic analysis
 
@@
 
-### Phylogenetic signal
+## Phylogenetic signal
 
@@
 
-### Analysis of biodiversity
+## Analysis of biodiversity
 
@@
 
-### Spatial models
+## Spatial models
 
@@
 
-### Analysis of bioregions
+## Analysis of bioregions
 
@@
 
-### Assessment of conservation status and threats
+## Assessment of conservation status and threats
 
@@
 
-## RESULTS
-
-### Datasets
+# Results
+
+## Datasets
 
@@
 
-### Observed diversity
+## Observed diversity
 
@@
 
-### Randomization tests for phylogenetic and functional diversity
+## Randomization tests for phylogenetic and functional diversity
 
@@
 
-### Relationships between biodiversity and climate or reproductive mode
+## Relationships between biodiversity and climate or reproductive mode
 
@@
 
-### Categorical analysis of neo- and paleo-endemism
+## Categorical analysis of neo- and paleo-endemism
 
@@
 
-### Bioregions
+## Bioregions
 
@@
 
-### Conservation status and threats
+## Conservation status and threats
 
@@
 
-## DISCUSSION
-
-### Distribution and drivers of biodiversity
+# Discussion
+
+## Distribution and drivers of biodiversity
 
@@
 
-### CONCLUSIONS
+# Conclusions
 
@@
 Our study integrates diverse, thoroughly collected datasets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns.
-We end by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.
+We conclude by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.
 
@@
 
-## Acknowledgements
-
-Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
+# Acknowledgements
+
+The authors thank Members of the Iwasaki Lab at the University of Tokyo for providing comments that improved the manuscript.
 Shawn Laffan provided advice about CANAPE.
@@
 
-## Author contributions
+# Author contributions
 
@@
 
-## Data availability
+# Data availability statement
 
@@
 
-## Literature cited
+# Online supporting information
+
+Additional supporting information may be found online in the Supporting Information section at the end of the article.
+
+**Appendix S1**. Supplementary tables and figures, including Tables `r s_table_num("moran")`--`r s_table_num("aic")` and Figs. `r s_figure_num("sampling-curve")`--`r s_figure_num("deer-map")`.
+
+**`r s_table("moran")`**.
+Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.
+**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
+**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
+**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
+**`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
+**`r s_table("aic")`**. Model fit as measured with conditional Akaike Information Criterion (cAIC).  
+**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
+**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.    
+**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan.  
+**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding hybrids.  
+**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 0.2° grid cell in the ferns of Japan.  
+**`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
+**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
+**`r s_figure("endemism-restricted")`**.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.
+**`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
+**`r s_figure("dendrogram")`**. Bioregion dendrograms.  
+**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
+**`r s_figure("apo-genus")`**. Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon.  
+**`r s_figure("modelfit-not-ult")`**.
+Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
+**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
+**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.  
+**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).
+
+# References
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
@@
 **`r figure("raw-diversity")`**.
-(a) Raw taxonomic diversity (b), phylogenetic diversity, and (c) functional diversity of the ferns of Japan.
+(A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan.
 For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.

### 2021-09-21 ms/SI.Rmd
Use 'nothotaxa' instead of 'hybrid'

@@
 **`r s_figure("tree")`**.
-Time tree of native Japanese ferns and lycophytes, excluding hybrids.
+Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.
 Points at nodes indicate bootstrap support.
@@
 **`r s_figure("apo-genus")`**.
-Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon.
+Rate of apomixis in genera Japanese ferns with > 1 apomictic taxon.
 Numbers next to each bar show number of apomictic taxa out of total richness per genus.
+Excludes non-native taxa and nothotaxa.
 

### 2021-09-21 ms/manuscript.Rmd
Use 'nothotaxa' instead of 'hybrid'

@@
 
-**Methods:** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` native, non-hybrid taxa.
+**Methods:** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` taxa.
 We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode.
@@
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b]. 
-One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
+One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa (excluding nothotaxa) [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
 
@@
 
-We used a list of all native, non-hybrid fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
+We used a list of fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
 We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v.
@@
 Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
-The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa (hereafter defined as species, subspecies, forms, and varieties unless otherwise indicated).
+The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety (hereafter referred to as "taxa" unless otherwise indicated), excluding non-native taxa and nothotaxa.
 Points (latitude and longitude of each occurrence record) were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
@@
 
-The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
+The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native ferns excluding nothotaxa [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
 The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
@@
 
-One element of biodiversity that we did not consider here is hybrids. 
-The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa. 
+One element of biodiversity that we did not consider here is nothotaxa (hybrids). 
+The Japanese fern flora includes `r n_hybrid_fern_taxa` nothotaxa. 
 Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989]. 
@@
 **`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan.  
-**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding hybrids.  
+**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
 **`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 0.2° grid cell in the ferns of Japan.  
@@
 **`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
-**`r s_figure("apo-genus")`**. Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon.  
+**`r s_figure("apo-genus")`**. Rate of apomixis in genera of Japanese ferns with > 1 apomictic taxon.  
 **`r s_figure("modelfit-not-ult")`**.

### 2021-09-21 ms/manuscript.Rmd
Update MS
Fix some captions, wording

@@
 The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety (hereafter referred to as "taxa" unless otherwise indicated), excluding non-native taxa and nothotaxa.
-Points (latitude and longitude of each occurrence record) were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
+Latitude and longitude of each occurrence record were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
 We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
@@
 **`r figure("coeff")`**.
-Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models.
+Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models for the effect of environment and reproductive mode on biodiversity metrics in the ferns of Japan.
 (a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa).
@@
 **`r figure("modelfit")`**.
-Relationship between biodiversity metrics and selected predictor variables.
+Relationship between biodiversity metrics and selected predictor variables in the ferns of Japan.
 (a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa).

### 2021-09-21 ms/SI.Rmd
Make figure part labels upper-case

@@
 a + b + c +
-  plot_annotation(tag_levels = "a",
+  plot_annotation(tag_levels = "A",
                   tag_prefix = "(",
@@
 Effect of grain size on sampling redundancy.
-(a), Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
-(b), Percent of grid cells with redundancy = 0 by grain size.
-(c), Distribution of redundancy values by grain size.
+(A), Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
+(B), Percent of grid cells with redundancy = 0 by grain size.
+(C), Distribution of redundancy values by grain size.
 There is a sudden improvement in redundancy values from 0.1 to 0.2 degree grid-cells, but less improvement as grain size is increased beyond 0.2 degrees.
@@
 
-a <- plot_env("a", "Area", "lat_area", "km^2")
-b <- plot_env("b", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
-c <- plot_env("c", "Precipitation", "precip", "mm")
-d <- plot_env("d", "Precip. season.", "precip_season" , "")
-e <- plot_env("e", "Temperature", "temp", "°C", function(x) x/10)
-f <- plot_env("f", "Temp. season.", "temp_season") +
+a <- plot_env("A", "Area", "lat_area", "km^2")
+b <- plot_env("B", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
+c <- plot_env("C", "Precipitation", "precip", "mm")
+d <- plot_env("D", "Precip. season.", "precip_season" , "")
+e <- plot_env("E", "Temperature", "temp", "°C", function(x) x/10)
+f <- plot_env("F", "Temp. season.", "temp_season") +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))
@@
 Reproductive mode and environmental data in 0.2 degree grid cells across Japan.
-(a), rolling mean area (km^2^) in 1° latitudinal bands.
-(b), % apomictic taxa.
-(c), annual precipitation.
-(d), percipitation seasonality (coefficient of variation).
-(e), mean annual temperature.
-(f), temperature seasonality (standard deviation × 100).
+(A), rolling mean area (km^2^) in 1° latitudinal bands.
+(B), % apomictic taxa.
+(C), annual precipitation.
+(D), percipitation seasonality (coefficient of variation).
+(E), mean annual temperature.
+(F), temperature seasonality (standard deviation × 100).
 
@@
   labs(fill = "No. specimens") +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
   map_theme()
@@
   labs(fill = "Redundancy") +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)**</span>", hjust = "inward", vjust = "inward", color = "white") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)**</span>", hjust = "inward", vjust = "inward", color = "white") +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
@@
 
-**`r s_figure("abundance")`**. Observed number of specimens (a) and sampling redundancy (b) per 0.2° grid cell in the ferns of Japan.
+**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 0.2° grid cell in the ferns of Japan.
 
@@
   plot_layout(nrow = 1) +
-  plot_annotation(tag_levels = "a",
+  plot_annotation(tag_levels = "A",
                   tag_prefix = "(",
@@
 **`r s_figure("raw-lat")`**.
-(a) Raw taxonomic diversity (b), phylogenetic diversity, and (c) functional diversity of the ferns of Japan plotted by latitude.
-For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+(A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude.
+For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
 Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
@@
   plot_layout(ncol = 2, nrow = 2) +
-  plot_annotation(tag_levels = "a",
+  plot_annotation(tag_levels = "A",
                   tag_prefix = "(",
@@
 Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.
-(a), Phylogenetic diversity (PD).
-(b), Relative phylogenetic diversity (RPD).
-(c), Functional diversity (FD).
-(d), Relative functional diversity (RFD).
-Trendlines in (a, c) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
+(A), Phylogenetic diversity (PD).
+(B), Relative phylogenetic diversity (RPD).
+(C), Functional diversity (FD).
+(D), Relative functional diversity (RFD).
+Trendlines in (A, C) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
 
@@
   plot_layout(ncol = 2) +
-  plot_annotation(tag_levels = "a",
+  plot_annotation(tag_levels = "A",
                   tag_prefix = "(",
@@
 Selection of *k* for bioregions.
-(a), taxonomic bioregions.
-(b), phylogenetic bioregions.
+(A), taxonomic bioregions.
+(B), phylogenetic bioregions.
 Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).
@@
   theme(legend.position = "right") +
-  labs(subtitle = "**(a)** Taxonomic") +
+  labs(subtitle = "**(A)** Taxonomic") +
   theme(plot.subtitle = element_markdown())
@@
 Color shows bioregion; bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.
-(a), taxonomic bioregions.
-(b), phylogenetic bioregions.
+(A), taxonomic bioregions.
+(B), phylogenetic bioregions.
 Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
@@
   theme(legend.position = "right") +
-  labs(subtitle = "**(b)** Phylogenetic") +
+  labs(subtitle = "**(B)** Phylogenetic") +
   theme(plot.subtitle = element_markdown())
@@
 Map of species richness for selected genera of ferns of Japan.
-(a), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
-(b), *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
-(c)--(f) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.
+(A), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
+(B), *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
+(C)--(F) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.
 
@@
   plot_layout(guides = "collect") +
-  plot_annotation(tag_levels = "a",
+  plot_annotation(tag_levels = "A",
                   tag_prefix = "(",
@@
 **`r s_figure("modelfit-not-ult")`**.
-Relationship between (a) phylogenetic diversity (PD) and (b) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
+Relationship between (A) phylogenetic diversity (PD) and (B) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
 Ribbon shows 95% confidence interval of model.
@@
 
-a + b + plot_annotation(tag_levels = "a",
+a + b + plot_annotation(tag_levels = "A",
                         tag_prefix = "(",
@@
 **`r s_figure("bioregions-env")`**.
-Scatterplots of grid cell membership in (a) taxonomic or (b) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
+Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
 Points shown with partial transparency; darker areas indicate overlapping points.
@@
 
-a <- make_protected_plot("a", "Richness", richness_obs_p_upper > 0.95)
-b <- make_protected_plot("b", "PD", pd_obs_p_upper > 0.95)
-c <- make_protected_plot("c", "FD", fd_obs_p_upper > 0.95)
-d <- make_protected_plot("d", "PE", pe_obs_p_upper > 0.95) +
+a <- make_protected_plot("A", "Richness", richness_obs_p_upper > 0.95)
+b <- make_protected_plot("B", "PD", pd_obs_p_upper > 0.95)
+c <- make_protected_plot("C", "FD", fd_obs_p_upper > 0.95)
+d <- make_protected_plot("D", "PE", pe_obs_p_upper > 0.95) +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))
@@
 Grid-cells with black outlines indicate significantly high biodiversity.
-Biodiversity metrics include (a) taxon richness, (b) phylogenetic diversity (PD), (c) functional diversity (FD), and (d) phylogenetic endemism (PE).
+Biodiversity metrics include (A) taxon richness, (B) phylogenetic diversity (PD), (C) functional diversity (FD), and (D) phylogenetic endemism (PE).
 Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
@@
 
-a <- make_deer_plot("a", "Richness", richness_obs_p_upper > 0.95)
-b <- make_deer_plot("b", "PD", pd_obs_p_upper > 0.95)
-c <- make_deer_plot("c", "FD", fd_obs_p_upper > 0.95)
-d <- make_deer_plot("d", "PE", pe_obs_p_upper > 0.95) +
+a <- make_deer_plot("A", "Richness", richness_obs_p_upper > 0.95)
+b <- make_deer_plot("B", "PD", pd_obs_p_upper > 0.95)
+c <- make_deer_plot("C", "FD", fd_obs_p_upper > 0.95)
+d <- make_deer_plot("D", "PE", pe_obs_p_upper > 0.95) +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

### 2021-09-21 ms/manuscript.Rmd
Make figure part labels upper-case

@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "No. taxa") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "Relative\nbranch\nlength") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(C)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Relative\nbranch\nlength") +
@@
 (A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan.
-For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
 
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   scale_fill_manual(values = signif_cols) +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(C)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(d)** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(D)** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
 Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.
-(a), Phylogenetic diversity (PD).
-(b), Relative phylogenetic diversity (RPD).
-(c), Functional diversity (FD).
-(d), relative functional diversity (RFD).
+(A), Phylogenetic diversity (PD).
+(B), Relative phylogenetic diversity (RPD).
+(C), Functional diversity (FD).
+(D), relative functional diversity (RFD).
 For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
@@
   plot_annotation(
-    tag_levels = "a",
+    tag_levels = "A",
     tag_prefix = "(",
@@
 Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models for the effect of environment and reproductive mode on biodiversity metrics in the ferns of Japan.
-(a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa).
+(A)--(E), environmental models (models including temperature); (F) and (G), reproductive models (models including % apomictic taxa).
 Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *P*-value indicates probability of no difference between the full and null model.
@@
 Relationship between biodiversity metrics and selected predictor variables in the ferns of Japan.
-(a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa).
+(A)--(E), environmental models (models including temperature); (F) and (G), reproductive models (models including % apomictic taxa).
 Ribbon shows 95% confidence interval of model.
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Bioregion") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Bioregion") +
@@
 Bioregions of the ferns of Japan.
-(a), taxonomic bioregions.
-(b), phylogenetic bioregions.
+(A), taxonomic bioregions.
+(B), phylogenetic bioregions.
 Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells.
@@
 Phylogenetic and morphological diversity of the ferns of Japan by taxonomic bioregion.
-(a), Standard effect size (SES) of phylogenetic diversity (PD).
-(b), SES of relative PD (RPD).
-(c), SES of functional diversity (FD).
-(d), SES of relative FD (RFD).
-(e), *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
+(A), Standard effect size (SES) of phylogenetic diversity (PD).
+(B), SES of relative PD (RPD).
+(C), SES of functional diversity (FD).
+(D), SES of relative FD (RFD).
+(E), *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
 Only top four clusters with the most sites shown.
@@
 Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan.
-(a) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status.
-(b) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
+(A) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status.
+(B) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
 Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
 Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
-For (a), vertical lines indicate total percent area protected across Japan (baseline protection rate).
-For (b), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).
+For (A), vertical lines indicate total percent area protected across Japan (baseline protection rate).
+For (B), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).

### 2021-09-21 R/ms_functions.R
Undo accidental merge of AJB commits

@@ -83,9 +83,8 @@ deer_lines <- c(
 
 # Words ----
 # Define formatting for some common custom words that may vary in style between journals
-# AJB: Common Latin words (e.g., in vivo, sensu lato) are not italicized.
-ie <- "ie"
-eg <- "eg"
+ie <- "*ie*"
+eg <- "*eg*"
 
 #' Format R packages names
 #' 
@@ -190,8 +189,8 @@ result_file <- function (result_num, extension) {
 # (except I never use the caption in the function, and instead replace with 'blank')
 figure_full <- captioner::captioner(prefix = "Fig.", suffix = "")
 table_full <- captioner::captioner(prefix = "Table")
-s_figure_full <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE, suffix = "")
-s_table_full <- captioner::captioner(prefix = "Table S", auto_space = FALSE, suffix = "")
+s_figure_full <- captioner::captioner(prefix = "Fig. S1.", auto_space = FALSE, suffix = "")
+s_table_full <- captioner::captioner(prefix = "Table S1.", auto_space = FALSE, suffix = "")
 
 # - Make a short function that prints only the object type and number, e.g., "Fig. 1"
 figure <- pryr::partial(figure_full, display = "cite", caption = "blank")

### 2021-09-21 ms/manuscript.Rmd
Change fig with from 7 to 7.25 in

@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7, height = 4, units = "in", raw_diversity_plot)
-```
-
-```{r raw-diversity-print, fig.width = 7, fig.height = 4, eval = params$doc_type == "pdf"}
+ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7.25, height = 4, units = "in", raw_diversity_plot)
+```
+
+```{r raw-diversity-print, fig.width = 7.25, fig.height = 4, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("rand-diversity"), "pdf"), rand_diversity_plot, width = 7, height = 8, units = "in")
-```
-
-```{r rand-diversity-print, fig.width = 7, fig.height = 9, eval = params$doc_type == "pdf"}
+ggsave(result_file(figure("rand-diversity"), "pdf"), rand_diversity_plot, width = 7.25, height = 8, units = "in")
+```
+
+```{r rand-diversity-print, fig.width = 7.25, fig.height = 9, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("coeff"), "pdf"), coeff_plot, width = 7, height = 5, units = "in")
-```
-
-```{r coeff-print, fig.width = 7, fig.height = 5, eval = params$doc_type == "pdf"}
+ggsave(result_file(figure("coeff"), "pdf"), coeff_plot, width = 7.25, height = 5, units = "in")
+```
+
+```{r coeff-print, fig.width = 7.25, fig.height = 5, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("modelfit"), "pdf"), modelfit_plot, width = 7, height = 5, units = "in")
-```
-
-```{r modelfit-print, fig.width = 7, fig.height = 5, eval = params$doc_type == "pdf"}
+ggsave(result_file(figure("modelfit"), "pdf"), modelfit_plot, width = 7.25, height = 5, units = "in")
+```
+
+```{r modelfit-print, fig.width = 7.25, fig.height = 5, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("bioregions"), "pdf"), bioregions_plot, width = 7, height = 5, units = "in")
-```
-
-```{r bioregions-print, fig.width = 7, fig.height = 5, eval = params$doc_type == "pdf"}
+ggsave(result_file(figure("bioregions"), "pdf"), bioregions_plot, width = 7.25, height = 5, units = "in")
+```
+
+```{r bioregions-print, fig.width = 7.25, fig.height = 5, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("div-by-tax-region"), "pdf"), div_by_tax_region_plot, width = 7, height = 7, units = "in")
-```
-
-```{r div-by-tax-region-print, fig.width = 7, fig.height = 7, eval = params$doc_type == "pdf"}
+ggsave(result_file(figure("div-by-tax-region"), "pdf"), div_by_tax_region_plot, width = 7.25, height = 7, units = "in")
+```
+
+```{r div-by-tax-region-print, fig.width = 7.25, fig.height = 7, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 7, height = 4, units = "in")
-```
-
-```{r conserv-status-print, fig.width = 7, fig.height = 4, eval = params$doc_type == "pdf"}
+ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 7.25, height = 4, units = "in")
+```
+
+```{r conserv-status-print, fig.width = 7.25, fig.height = 4, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF

### 2021-09-21 ms/SI.Rmd
Fix typo that erased SI fig 11

@@
 f <- plot_genus("f", "Diplazium") +
-  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
+  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))
 

### 2021-09-21 ms/SI.Rmd
Finish fixing caption letters per AJB
Upper-case
No parentheses in figure

@@
   plot_annotation(tag_levels = "A",
-                  tag_prefix = "(",
-                  tag_suffix = ")") &
+                  tag_prefix = "",
+                  tag_suffix = "") &
   theme(plot.tag = element_text(face = "bold"))
@@
       geom="richtext", y = Inf, x = -Inf, 
-      label = glue("<span style='color:black'>**({part})** {label}</span>"), 
+      label = glue("<span style='color:black'>**{part}** {label}</span>"), 
       hjust = "inward", vjust = "inward", color = "white") + 
@@
   labs(fill = "No. specimens") +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A**</span>", hjust = "inward", vjust = "inward", color = "white") + 
   map_theme()
@@
   labs(fill = "Redundancy") +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)**</span>", hjust = "inward", vjust = "inward", color = "white") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B**</span>", hjust = "inward", vjust = "inward", color = "white") +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm")) +
@@
   theme(legend.position = "right") +
-  labs(subtitle = "**(A)** Taxonomic") +
+  labs(subtitle = "**A** Taxonomic") +
   theme(plot.subtitle = element_markdown())
@@
   theme(legend.position = "right") +
-  labs(subtitle = "**(B)** Phylogenetic") +
+  labs(subtitle = "**B** Phylogenetic") +
   theme(plot.subtitle = element_markdown())
@@
       geom="richtext", y = Inf, x = -Inf, 
-      label = glue("<span style='color:black'>**({part})** *{genus_select}*</span>"), 
+      label = glue("<span style='color:black'>**{part}** *{genus_select}*</span>"), 
       hjust = "inward", vjust = "inward", color = "white") + 
@@
 a + b + plot_annotation(tag_levels = "A",
-                        tag_prefix = "(",
-                        tag_suffix = ")") &
+                        tag_prefix = "",
+                        tag_suffix = "") &
   theme(plot.tag = element_text(face = "bold"))
@@
       geom="richtext", y = Inf, x = -Inf, 
-      label = glue("<span style='color:black'>**({part})** {variable}</span>"), 
+      label = glue("<span style='color:black'>**{part}** {variable}</span>"), 
       hjust = "inward", vjust = "inward", color = "white") +

### 2021-09-21 ms/manuscript.Rmd
Finish fixing caption letters per AJB
Upper-case
No parentheses in figure

@@
 We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown. 
-We calculated the percentage of apomictic species in each grid-cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`b).
+We calculated the percentage of apomictic species in each grid-cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).
 
@@
 The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`). 
-The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`a). 
-Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`b).
+The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A). 
+Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B).
 The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid-cells.
@@
 
-Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`a, `r s_figure("raw-lat")`a).
-Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`b, c), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`a, c).
-Observed values of RPD and RFD were not linearly related to richness; however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`b, d).
+Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`A, `r s_figure("raw-lat")`A).
+Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`A, C).
+Observed values of RPD and RFD were not linearly related to richness; however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).
 
@@
 
-Grid-cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`a). 
+Grid-cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`A). 
 A smaller number of cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula). 
-Grid-cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`b). 
+Grid-cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`B). 
 Grid-cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido. 
-Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`c, d).
+Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`C, D).
 
@@
 All models were significant relative to null models without any fixed effects (`r s_table("lrt-null")`). 
-The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`a--e), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`f--g); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
-Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`a). 
-Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`b--e). 
-SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`g); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`f). 
+The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
+Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A). 
+Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E). 
+SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F). 
 Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD and SES of PD (`r s_table("aic")`).
@@
 Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`a; `r s_figure("dendrogram")`a).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`b; `r s_figure("dendrogram")`b).
+The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
 The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
@@
 
-The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`a).
+The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A).
 Cells with significantly high biodiversity had a greater percentage of area protected than the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
@@
 The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
-Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (`r figure("conserv-status")`b).
+Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (`r figure("conserv-status")`B).
 However, cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.
@@
 `r figure_num("raw-diversity")`a, `r s_figure_num("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020].
-Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`a).
+Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
 Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
@@
 
-The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`a, c). 
+The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C). 
 This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
@@
 
-Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`b).
-The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`a).
-The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`b) and *Dryopteris* (`r s_figure("genus-map")`d).
+Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`B).
+The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`A).
+The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).
 
 Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (Figs.
-`r figure_num("coeff")`g, `r figure_num("modelfit")`g), and to a slightly lesser extent SES of PD (Figs.
-`r figure_num("coeff")`f, `r figure_num("modelfit")`f).
+`r figure_num("coeff")`G, `r figure_num("modelfit")`G), and to a slightly lesser extent SES of PD (Figs.
+`r figure_num("coeff")`F, `r figure_num("modelfit")`F).
 Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (`r s_figure("genus-map")`).
@@
 The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, there is some overlap between them on Yakushima and Amami Isl.
-(`r figure("bioregions")`b).
+(`r figure("bioregions")`B).
 The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
@@
 
-We found that grid-cells with significantly high amounts of biodiversity are better protected than the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`a).
-The high protection status of grid-cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`d).
+We found that grid-cells with significantly high amounts of biodiversity are better protected than the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`A).
+The high protection status of grid-cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`D).
 One possible reason for the generally high level of protection overall is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas [A.
@@
 
-Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`b).
-This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("protected-map")`a).
+Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`B).
+This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("protected-map")`A).
 This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.
@@
 The extreme evolutionary distinctiveness of the southern islands is both a major feature of the Japanese fern flora and a challenge for analysis. 
-The tropical lineages occurring only in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on the main islands (`r figure("rand-diversity")`a).
+The tropical lineages occurring only in these islands are likely responsible for the comparatively low (`r ie`, clustered) values of PD observed on the main islands (`r figure("rand-diversity")`A).
 One way to account for this bimodal distribution pattern would be to use a restricted null model that does not allow all taxa to disperse anywhere within the study area [@Mishler2020]. 
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "No. taxa") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "Relative\nbranch\nlength") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(C)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Relative\nbranch\nlength") +
@@
 **`r figure("raw-diversity")`**.
-(A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan.
+Raw (observed) biodiversity metrics of the ferns of Japan.
+(A) Raw taxonomic diversity (B).
+Raw phylogenetic diversity.
+(C) Raw functional diversity.
 For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   scale_fill_manual(values = signif_cols) +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(C)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(D)** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**D** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "") +
@@
 Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.
-(A), Phylogenetic diversity (PD).
-(B), Relative phylogenetic diversity (RPD).
-(C), Functional diversity (FD).
-(D), relative functional diversity (RFD).
+(A) Phylogenetic diversity (PD).
+(B) Relative phylogenetic diversity (RPD).
+(C) Functional diversity (FD).
+(D) Relative functional diversity (RFD).
 For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
@@
     tag_levels = "A",
-    tag_prefix = "(",
-    tag_suffix = ")") &
+    tag_prefix = "",
+    tag_suffix = "") &
   theme(plot.tag = element_text(face = "bold"))
@@
 Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models for the effect of environment and reproductive mode on biodiversity metrics in the ferns of Japan.
-(A)--(E), environmental models (models including temperature); (F) and (G), reproductive models (models including % apomictic taxa).
+(A)--(E) Environmental models (models including temperature).
+(F, G) Reproductive models (models including % apomictic taxa).
 Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *P*-value indicates probability of no difference between the full and null model.
@@
     tag_levels = "A",
-    tag_prefix = "(",
-    tag_suffix = ")")  &
+    tag_prefix = "",
+    tag_suffix = "")  &
   theme(plot.tag = element_text(face = "bold"))
@@
 Relationship between biodiversity metrics and selected predictor variables in the ferns of Japan.
-(A)--(E), environmental models (models including temperature); (F) and (G), reproductive models (models including % apomictic taxa).
+(A)--(E) Environmental models (models including temperature).
+(F, G) Reproductive models (models including % apomictic taxa).
 Ribbon shows 95% confidence interval of model.
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(A)** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Bioregion") +
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(B)** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
   labs(fill = "Bioregion") +
@@
 Bioregions of the ferns of Japan.
-(A), taxonomic bioregions.
-(B), phylogenetic bioregions.
+(A) Taxonomic bioregions.
+(B) Phylogenetic bioregions.
 Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells.
@@
     tag_levels = "A",
-    tag_prefix = "(",
-    tag_suffix = ")"
+    tag_prefix = "",
+    tag_suffix = ""
   ) &
@@
 Phylogenetic and morphological diversity of the ferns of Japan by taxonomic bioregion.
-(A), Standard effect size (SES) of phylogenetic diversity (PD).
-(B), SES of relative PD (RPD).
-(C), SES of functional diversity (FD).
-(D), SES of relative FD (RFD).
-(E), *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
+(A) Standard effect size (SES) of phylogenetic diversity (PD).
+(B) SES of relative PD (RPD).
+(C) SES of functional diversity (FD).
+(D) SES of relative FD (RFD).
+(E) *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
 Only top four clusters with the most sites shown.

### 2021-09-22 ms/manuscript.Rmd
Update MS
trim abstract to AJB word limit

@@
 
-**Premise:** Biodiversity is often measured with a single metric, species richness.
+**Premise:** Biodiversity is often only measured with species richness.
 However, this ignores evolutionary history and is not sufficient for making conservation decisions.
-Here, we use a densely sampled dataset to characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.
-
-**Methods:** We used data from >`r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` fern specimens in Japan to compile a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` taxa.
-We combined this with a phylogenetic tree and functional trait data to analyze taxonomic, phylogenetic, and functional diversity, and modeled the distribution of biodiversity in response to environmental factors and reproductive mode.
-We conducted categorical analysis of neo- and paleo-endemism to identify areas of significant phylogenetic endemism. 
-We used hierarchical clustering of taxonomic and phylogenetic distances to analyze bioregions. 
-We compared the overlap of significantly diverse grid-cells with existing conservation zones and range maps of native Japanese deer to assess conservation status and threats.
-
-**Results:** Three major patterns in the spatial distribution of biodiversity were observed: that of taxonomic richness, which is highest at mid-latitudes, that of phylogenetic and functional diversity and phylogenetic endemism, which are highest in small southern islands, and that of relative phylogenetic and functional diversity, which are high at both high and low latitudes, and low at mid-latitudes.
+Here, we characterize multiple facets and drivers of biodiversity to understand how these relate to bioregions and conservation status in the ferns of Japan.
+
+**Methods:** We compiled a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
+We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
+We used hierarchical clustering to analyze bioregions.
+We compared the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer to assess conservation status and threats.
+
+**Results:** Taxonomic richness was highest at mid-latitudes.
+Phylogenetic and functional diversity and phylogenetic endemism were highest in small southern islands.
+Relative phylogenetic and functional diversity were high at both high and low latitudes, and low at mid-latitudes.
 Grid-cells were grouped in to `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions. 
-Temperature and apomixis were identified as drivers of patterns of taxonomic and phylogenetic diversity.
-Conservation status (% protected area) was generally high for grid-cells with significantly high biodiversity, but the threat due to herbivory by deer varied by biodiversity metric, and was particularly severe for taxonomic richness.
+Temperature and apomixis were identified as drivers of taxonomic and phylogenetic diversity.
+Conservation status was generally high for grid-cells with significantly high biodiversity, but the threat due to herbivory was greater for taxonomic richness than other metrics.
 
 **Conclusions:** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan.
-The different distributions of each diversity type reflect the major bioregions. 
 Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple multiple metrics when establishing conservation priorities.

### 2021-09-22 ms/data_exploration.Rmd
Update MS

--- 2021-09-21 ms/data_exploration.Rmd
@@
 
-The next step uses `tar_load()` to load the dataset from the targets cache. Alternatively, this can be loaded from the dryad data file as shown in the commented-out code below.
+The next step uses `tar_load()` to load the dataset from the targets cache. Alternatively, this can be loaded from the data file as shown in the commented-out code below.
 
@@
 
-# Alternatively, unzip `results.zip` in the Dryad data
+# Alternatively, unzip `results.zip`
 # (FIXME: add DOI when available)

### 2021-09-22 ms/manuscript.Rmd
Update MS

@@
 
-Our data include all ferns known to natively occur within the boundaries of Japan . 
 Japan (20°N to 45°N, 122°E to 153°E) consists of four main islands and thousands of smaller ones (`r figure("japan-map")`). 
 Most of the main islands have been in contact with the continent at various points in the past during periods of lower sea level, but the Ogasawara archipelago consists of oceanic islands that have never been connected to the mainland.
-The climate is subtropical in the south and temperate in other parts of the country. 
+The climate is subtropical in the southern islands and temperate in other parts of the country. 
 Map data for Japan were downloaded from the Geospatial Information Authority of Japan under the Creative Commons Attribution License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html).
@@
 1.01 (http://www.rdplants.org/gl/).
-Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1 in Supporting Information). 
+Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article). 
 The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety (hereafter referred to as "taxa" unless otherwise indicated), excluding non-native taxa and nothotaxa.
@@
 
-Initial inspection of the biodiversity and environmental data showed a single outlier in % apomictic taxa due to extremely low species richness (Appendix S2), so that grid-cell and one grid-cell lacking environmental data were excluded from modeling analyses.
+Initial inspection of the biodiversity and environmental data showed a single outlier in % apomictic taxa due to extremely low species richness (Appendix S2; see Supplemental Data with this article), so that grid-cell and one grid-cell lacking environmental data were excluded from modeling analyses.
 
@@
 A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
-Data files needed to run the analysis and selected results files are available from Figshare at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.
+Data files needed to run the analysis and selected results files are available from Figshare at http://doi.org/10.6084/m9.figshare.16655263.
+**Note to reviewers**: the Figshare DOI will not be accessible until the paper is accepted. During review, the data can be accessed from this link: https://figshare.com/s/de3d5a70e118208e56c5.
 
@@
 
-# References
+**Appendix S2**. Exploratory data analysis for spatial modeling.
+
+# Literature cited
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->

### 2021-09-22 R/ms_functions.R
Change formatting of software names

@@ -105,7 +105,8 @@ pack <- function(x) {glue::glue("'{x}' v.{packageVersion(x)}")}
 #'
 #' @return Name of software formatted according to journal requirements
 #' 
-software <- function(x) {glue::glue('<span style="font-variant:small-caps;">{x}</span>')}
+# software <- function(x) {glue::glue('<span style="font-variant:small-caps;">{x}</span>')}
+software <- function(x) {x}
 
 #' Format R function names
 #'

### 2021-09-22 ms/manuscript.Rmd
Change formatting of software names
Fix degree symbol in MS

@@
 We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
-We calculated area as rolling mean area (km^2^) across 1 degree latitudinal bands.
+We calculated area as rolling mean area (km^2^) across 1° latitudinal bands.
 This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).
@@
 We trimmed the resulting ultrametric tree to only Japanese taxa.
-treePL requires branch lengths to be > 0, and it sets extremely small branches to an arbitrary minimum positive length.
+`r software("treePL")` requires branch lengths to be > 0, and it sets extremely small branches to an arbitrary minimum positive length.
 This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
@@
 **`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.    
-**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan.  
+**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2° grid cells across Japan.  
 **`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  

### 2021-12-24 ms/SI.Rmd

@@
   - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
-  - \fancyhead[RE,RO]{Nitta et al.---\textit{American Journal of Botany} 2021 -- Appendix S1} # add header on right side
+  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S1} # add header on right side
   - \fancyhead[LO,LE]{} # suppress automatic section header on left side
@@
 **`r s_figure("sampling-curve")`**.
-Species collection curve for the ferns of Japan.
-Curve fit with the `r pack("iNEXT")` package [@Hsieh2016].
+Species collection curve for the ferns of Japan fit with the `r pack("iNEXT")` package [@Hsieh2016].
 Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line).
@@
 Effect of grain size on sampling redundancy.
-(A), Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
-(B), Percent of grid cells with redundancy = 0 by grain size.
-(C), Distribution of redundancy values by grain size.
+(a), Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
+(b), Percent of grid cells with redundancy = 0 by grain size.
+(c), Distribution of redundancy values by grain size.
 There is a sudden improvement in redundancy values from 0.1 to 0.2 degree grid-cells, but less improvement as grain size is increased beyond 0.2 degrees.
@@
 
-a <- plot_env("A", "Area", "lat_area", "km^2")
-b <- plot_env("B", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
-c <- plot_env("C", "Precipitation", "precip", "mm")
-d <- plot_env("D", "Precip. season.", "precip_season" , "")
-e <- plot_env("E", "Temperature", "temp", "°C", function(x) x/10)
-f <- plot_env("F", "Temp. season.", "temp_season") +
+a <- plot_env("a", "Area", "lat_area", "km^2")
+b <- plot_env("b", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
+c <- plot_env("c", "Precipitation", "precip", "mm")
+d <- plot_env("d", "Precip. season.", "precip_season" , "")
+e <- plot_env("e", "Temperature", "temp", "°C", function(x) x/10)
+f <- plot_env("f", "Temp. season.", "temp_season") +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))
@@
 Reproductive mode and environmental data in 0.2 degree grid cells across Japan.
-(A), rolling mean area (km^2^) in 1° latitudinal bands.
-(B), % apomictic taxa.
-(C), annual precipitation.
-(D), percipitation seasonality (coefficient of variation).
-(E), mean annual temperature.
-(F), temperature seasonality (standard deviation × 100).
+(a), rolling mean area (km^2^) in 1° latitudinal bands.
+(b), % apomictic taxa.
+(c), annual precipitation.
+(d), percipitation seasonality (coefficient of variation).
+(e), mean annual temperature.
+(f), temperature seasonality (standard deviation × 100).
 
@@
 **`r s_figure("tree")`**.
-Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.
+Time tree of native Japanese ferns and lycophytes, excluding hybrids.
 Points at nodes indicate bootstrap support.
@@
 
-**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 0.2° grid cell in the ferns of Japan.
+**`r s_figure("abundance")`**. Observed number of specimens (a) and sampling redundancy (b) per 0.2° grid cell in the ferns of Japan.
 
@@
 **`r s_figure("raw-lat")`**.
-(A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude.
-For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+(a) Raw taxonomic diversity (b), phylogenetic diversity, and (c) functional diversity of the ferns of Japan plotted by latitude.
+For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
 Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
@@
 Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.
-(A), Phylogenetic diversity (PD).
-(B), Relative phylogenetic diversity (RPD).
-(C), Functional diversity (FD).
-(D), Relative functional diversity (RFD).
-Trendlines in (A, C) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
+(a), Phylogenetic diversity (PD).
+(b), Relative phylogenetic diversity (RPD).
+(c), Functional diversity (FD).
+(d), Relative functional diversity (RFD).
+Trendlines in (a, c) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
 
@@
 Selection of *k* for bioregions.
-(A), taxonomic bioregions.
-(B), phylogenetic bioregions.
+(a), taxonomic bioregions.
+(b), phylogenetic bioregions.
 Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).
@@
 Color shows bioregion; bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.
-(A), taxonomic bioregions.
-(B), phylogenetic bioregions.
+(a), taxonomic bioregions.
+(b), phylogenetic bioregions.
 Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
@@
 Map of species richness for selected genera of ferns of Japan.
-(A), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
-(B), *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
-(C)--(F) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.
+(a), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
+(b), *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
+(c)--(f) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.
 
@@
 **`r s_figure("apo-genus")`**.
-Rate of apomixis in genera Japanese ferns with > 1 apomictic taxon.
+Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon.
 Numbers next to each bar show number of apomictic taxa out of total richness per genus.
-Excludes non-native taxa and nothotaxa.
 
@@
 **`r s_figure("modelfit-not-ult")`**.
-Relationship between (A) phylogenetic diversity (PD) and (B) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
+Relationship between (a) phylogenetic diversity (PD) and (b) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
 Ribbon shows 95% confidence interval of model.
@@
 **`r s_figure("bioregions-env")`**.
-Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
+Scatterplots of grid cell membership in (a) taxonomic or (b) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
 Points shown with partial transparency; darker areas indicate overlapping points.
@@
 
-a <- make_protected_plot("A", "Richness", richness_obs_p_upper > 0.95)
-b <- make_protected_plot("B", "PD", pd_obs_p_upper > 0.95)
-c <- make_protected_plot("C", "FD", fd_obs_p_upper > 0.95)
-d <- make_protected_plot("D", "PE", pe_obs_p_upper > 0.95) +
+a <- make_protected_plot("a", "Richness", richness_obs_p_upper > 0.95)
+b <- make_protected_plot("b", "PD", pd_obs_p_upper > 0.95)
+c <- make_protected_plot("c", "FD", fd_obs_p_upper > 0.95)
+d <- make_protected_plot("d", "PE", pe_obs_p_upper > 0.95) +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))
@@
 Grid-cells with black outlines indicate significantly high biodiversity.
-Biodiversity metrics include (A) taxon richness, (B) phylogenetic diversity (PD), (C) functional diversity (FD), and (D) phylogenetic endemism (PE).
+Biodiversity metrics include (a) taxon richness, (b) phylogenetic diversity (PD), (c) functional diversity (FD), and (d) phylogenetic endemism (PE).
 Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
@@
 
-a <- make_deer_plot("A", "Richness", richness_obs_p_upper > 0.95)
-b <- make_deer_plot("B", "PD", pd_obs_p_upper > 0.95)
-c <- make_deer_plot("C", "FD", fd_obs_p_upper > 0.95)
-d <- make_deer_plot("D", "PE", pe_obs_p_upper > 0.95) +
+a <- make_deer_plot("a", "Richness", richness_obs_p_upper > 0.95)
+b <- make_deer_plot("b", "PD", pd_obs_p_upper > 0.95)
+c <- make_deer_plot("c", "FD", fd_obs_p_upper > 0.95)
+d <- make_deer_plot("d", "PE", pe_obs_p_upper > 0.95) +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

### 2021-12-24 ms/manuscript.Rmd

@@
   bookdown::word_document2:
-    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
+    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
     toc: no
     number_sections: no
-    reference_docx: ajb_template.docx
+    reference_docx: template.docx
   bookdown::pdf_document2:
@@
     latex_engine: xelatex
-    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
+    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
 params:
@@
 
-Joel H. Nitta^1,5^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^
+Short running title: Spatial phylogenetics of Japanese ferns
+
+Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^
 
@@
 
-^5^Author for correspondence: joelnitta@gmail.com
-
-Manuscript received _______; revision accepted _______.
-
-Short running title: Spatial phylogenetics of Japanese ferns
+Author for correspondence:  
+*Joel H. Nitta*  
+*Email: joelnitta@gmail.com*
+
+## Acknowledgements
+
+Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
+Shawn Laffan provided advice about CANAPE.
+This study was supported by JSPS KAKENHI Grant Number 16H06279.
 
@@
 
-**Abstract**
+## Abstract
 
@@
 
+## Introduction
+
 ```{r intro-stats}
@@
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b]. 
-One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa (excluding nothotaxa) [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
+One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
 
@@
 
-# Materials and methods
+## Materials and Methods
 
@@
 
-## Study Site
+### Study Site
 
@@
 
-## Occurrence data
-
-We used a list of fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
+### Occurrence data
+
+We used a list of all native, non-hybrid fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
 We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v.
@@
 
-## Morphological trait data
+### Morphological trait data
 
@@
 
-## Reproductive mode data
+### Reproductive mode data
 
@@
 
-## Environmental data
+### Environmental data
 
@@
 
-## Phylogenetic analysis
+### Phylogenetic analysis
 
@@
 
-## Phylogenetic signal
+### Phylogenetic signal
 
@@
 
-## Analysis of biodiversity
+### Analysis of biodiversity
 
@@
 
-## Spatial models
+### Spatial models
 
@@
 
-## Analysis of bioregions
+### Analysis of bioregions
 
@@
 
-## Assessment of conservation status and threats
+### Assessment of conservation status and threats
 
@@
 
-# Results
-
-## Datasets
+## Results
+
+### Datasets
 
@@
 
-The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native ferns excluding nothotaxa [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
+The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
 The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`). 
@@
 
-## Observed diversity
+### Observed diversity
 
@@
 
-## Randomization tests for phylogenetic and functional diversity
+### Randomization tests for phylogenetic and functional diversity
 
@@
 
-## Relationships between biodiversity and climate or reproductive mode
+### Relationships between biodiversity and climate or reproductive mode
 
@@
 
-## Categorical analysis of neo- and paleo-endemism
+### Categorical analysis of neo- and paleo-endemism
 
@@
 
-## Bioregions
+### Bioregions
 
@@
 Taxonomic bioregions 3 and 4 have higher SES values of PD, RPD, FD, and RFD than bioregions 1 and 2 (`r figure("div-by-tax-region")`). 
-Bioregions 3 and 4 also have consistently high PE *P*-scores, whereas bioregions 1 and 2 have a much wider variation. 
+Bioregions 3 and 4 also have consistently high PE *p*-scores, whereas bioregions 1 and 2 have a much wider variation. 
 Bioregion 2 has lower RPD and RFD than bioregion 1. 
@@
 
-## Conservation status and threats
+### Conservation status and threats
 
@@
 
-# Discussion
-
-## Distribution and drivers of biodiversity
+## Discussion
+
+### Distribution and drivers of biodiversity
 
@@
 
-# Conclusions
+### Limits of the study and future outlook
 
@@
 
-Our study integrates diverse, thoroughly collected datasets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns.
-We conclude by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.
-
+There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research. 
 First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
@@
 
-One element of biodiversity that we did not consider here is nothotaxa (hybrids). 
-The Japanese fern flora includes `r n_hybrid_fern_taxa` nothotaxa. 
+One element of biodiversity that we did not consider here is hybrids. 
+The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa. 
 Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989]. 
@@
   "pd_obs_z", "SES of PD",
-  "pe_obs_p_upper", "PE *P*-value",
+  "pe_obs_p_upper", "PE *p*-value",
   "rfd_obs_z", "SES of RFD",
@@
       values = c("TRUE" = "grey20", "FALSE" = "grey70"),
-      labels = c("TRUE" = "*P* < 0.05", "FALSE" = "*P* > 0.05")) +
+      labels = c("TRUE" = "*p* < 0.05", "FALSE" = "*p* > 0.05")) +
     labs(
@@
   labs(
-    y = "PE *P*-score",
+    y = "PE *p*-score",
     x = "Cluster",
@@
 Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan.
-(A) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status.
-(B) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
+(a) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status.
+(b) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
 Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
 Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
-For (A), vertical lines indicate total percent area protected across Japan (baseline protection rate).
-For (B), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).
+For (a), vertical lines indicate total percent area protected across Japan (baseline protection rate).
+For (b), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).
+
+`r pagebreak_doc()`
+
+```{r japan-map-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+# Print plot if rendering to doc on initial submission
+japan_map_plot
+cat("\n\n", figure("japan-map"))
+```
+
+`r pagebreak_doc()`
+
+```{r raw-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
+raw_diversity_plot
+cat("\n\n", figure("raw-diversity"))
+```
+
+`r pagebreak_doc()`
+
+```{r rand-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 9, eval = params$doc_type == "doc", results = "asis"}
+rand_diversity_plot
+cat("\n\n", figure("rand-diversity"))
+```
+
+`r pagebreak_doc()`
+
+```{r coeff-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+coeff_plot
+cat("\n\n", figure("coeff"))
+```
+
+`r pagebreak_doc()`
+
+```{r modelfit-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+modelfit_plot
+cat("\n\n", figure("modelfit"))
+```
+
+`r pagebreak_doc()`
+
+```{r endemism-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+endemism_plot
+cat("\n\n", figure("endemism"))
+```
+
+`r pagebreak_doc()`
+
+```{r bioregions-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
+bioregions_plot
+cat("\n\n", figure("bioregions"))
+```
+
+`r pagebreak_doc()`
+
+```{r div-by-tax-region-doc, fig.cap = "", fig.width = 7, fig.height = 7, eval = params$doc_type == "doc", results = "asis"}
+div_by_tax_region_plot
+cat("\n\n", figure("div-by-tax-region"))
+```
+
+`r pagebreak_doc()`
+
+```{r conserv-status-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
+conserv_status_plot
+cat("\n\n", figure("conserv-status"))
+```
+
+`r pagebreak_doc()`
+
+## Data availability statement
+
+Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy. 
+A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
+Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.
+
+## References
+<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
+<div id="refs"></div>
+
+## Biosketch
+
+Joel H. Nitta's research focuses on patterns and processes of diversification in ferns. For more information, see https://joelnitta.com.
+
+Author contributions: J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.

### 2021-12-24 ms/SI.Rmd
Use equal-area projection for spatial data

@@
 # Helper function to format resolution labels
-degree_lab <- function (x) {
+km_lab <- function (x) {
   as.numeric(x) %>% 
-    scales::number(suffix = "\u00b0", accuracy = 0.1)
+    scales::number(suffix = "km", scale = 1e-3)
 }
@@
   scale_y_continuous(labels = scales::percent) +
-  scale_x_discrete(labels = degree_lab) +
+  scale_x_discrete(labels = km_lab) +
   labs(
@@
   geom_density_ridges(scale = 1.5) + 
-  scale_y_discrete(expand = c(0, 0), labels = degree_lab) +     # will generally have to set the `expand` option
+  scale_y_discrete(expand = c(0, 0), labels = km_lab) + # will generally have to set the `expand` option
   scale_x_continuous(expand = c(0, 0)) +   # for both axes to remove unneeded padding
@@
 Effect of grain size on sampling redundancy.
-(a), Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells).
+(a), Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km).
 (b), Percent of grid cells with redundancy = 0 by grain size.
 (c), Distribution of redundancy values by grain size.
-There is a sudden improvement in redundancy values from 0.1 to 0.2 degree grid-cells, but less improvement as grain size is increased beyond 0.2 degrees.
+There is a sudden improvement in redundancy values from 10 km to 20 km, but less improvement as grain size is increased beyond 20 km.
 

### 2021-12-24 R/process_raw_data.R

@@ -67,48 +67,42 @@ tar_plan(
     occ_point_data = occ_point_data_ferns_unfiltered,
     mask = japan_mesh2),
 
-  # get CRS (JGD2000) from japan_mesh2 for converting point data to community
-  # dataframe
-  jgd2000_sp = get_sp_crs_from_sf(japan_mesh2),
-  jgd2000_sf = st_crs(japan_mesh2),
-
   # Calculate richness, abundance, and redundancy at four scales:
-  # 0.1, 0.2, 0.3, and 0.4 degree grid squares
-  scales_to_test = c(0.1, 0.2, 0.3, 0.4),
+  # 10km x 10km, 20km x 20km, 30km x 30km, 40km x 40km
+  scales_to_test = c(10000, 20000, 30000, 40000), # units in m
+  # Use Mollweide equal-area projection with longitude centered on Japan
+  mollweide = "+proj=moll +lon_0=135",
   tar_target(
     comm_scaled_list,
     comm_from_points(
-      species_coods = occ_point_data_ferns,
-      resol = scales_to_test,
-      lon = "longitude",
-      lat = "latitude",
-      species = "taxon",
-      crs = jgd2000_sp),
+      points = occ_point_data_ferns,
+      res = scales_to_test,
+      long_col = "longitude",
+      lat_col = "latitude",
+      crs = mollweide),
     pattern = map(scales_to_test)
   ),
 
-  # After inspecting results (see SI), select 0.2 degrees as optimal scale.
+  # After inspecting results (see SI), select 20 km x 20 km as optimal scale
 
-  # Extract geographic shapes, richness, and number of specimens at 0.2 degree
-  # grid scale
-  shape_ferns_full = shape_from_comm_scaled_list(comm_scaled_list, 0.2) %>%
-    # calculate redundancy
-    mutate(redundancy = 1 - (richness/abundance)) %>%
-    # set CRS to be exactly same as JGD2000 in japan_mesh2
-    st_transform(jgd2000_sf),
+  # Extract geographic shapes, richness, and number of specimens at selected
+  # scale
+  scale_select = 20000,
+  shape_ferns_full = shape_from_comm_scaled_list(
+    comm_scaled_list, scale_select),
 
   # Write out geographic shapes in GeoPackage format
   # (https://www.geopackage.org/)
   tar_file(
     shape_ferns_full_out,
-    # Include manual time stamp so that sha is stable
[…]

### 2021-12-27 ms/manuscript.Rmd
Change randomization method to curveball

@@
 Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test. 
-We generated 999 random communities using the "independent swap" algorithm of @Gotelli2000, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values.
+We generated 999 random communities using the "curveball" algorithm of @Strona2014, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values.
 Statistical significance was assessed using a one-tailed test for PE or a two-tailed test for other metrics with alpha = 0.5; observed values in the extreme 5% of the null distribution were considered significant.

### 2021-12-27 ms/SI.Rmd
Change to equal-area grid cells

@@
 # Crop high-res map of Japan to spatial div results area
-japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
-
-# Format biodiv_ferns_spatial data
+japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial)) %>%
+  st_transform(4612)
+
+# Format biodiv_ferns_spatial data for plotting
 biodiv_ferns_spatial <-
@@
     )
-  )
+  ) %>%
+  st_transform(4612) %>%
+  # Only plot data included in model
+  inner_join(select(biodiv_ferns_cent, grids), by = "grids")
 
@@
     endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super"))
-  )
-
-# Subset ferns biodiversity dataframe: restrict sites to only those used in model
-# (drops outlier % apomictic site)
-biodiv_ferns_for_plotting <-
-  biodiv_ferns_spatial %>%
-  drop_apo_outlier
+  ) %>%
+  st_transform(4612) %>%
+  # Only plot data included in model
+  inner_join(select(biodiv_ferns_cent, grids), by = "grids")
 
@@
 **`r s_figure("envir")`**.
-Reproductive mode and environmental data in 0.2 degree grid cells across Japan.
-(a), rolling mean area (km^2^) in 1° latitudinal bands.
+Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan.
+(a), rolling mean area (km^2^) in 20 km latitudinal bands.
 (b), % apomictic taxa.
@@
 (f), temperature seasonality (standard deviation × 100).
+Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.
 
@@
 ```{r raw-lat, fig.width = 7, fig.height = 4}
-# Convert spatial data frame to centroids with raw metrics, drop single high apomixis outlier
-biodiv_ferns_cent_raw <- spatial_to_cent_for_model(
-  biodiv_ferns_spatial,
-  c("grids", "lat", "long", "richness", "pd_obs", "fd_obs"))
+# Convert spatial data frame to centroids with raw metrics
+biodiv_ferns_cent_raw <-
+  biodiv_ferns_spatial %>%
+  sf_add_centroids() %>%
+  st_drop_geometry()
 
@@
 Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods).
-Color shows bioregion; bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.
+Color shows bioregion; bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
 (a), taxonomic bioregions.
@@
 Points shown with partial transparency; darker areas indicate overlapping points.
-Bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.
+Bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
 

### 2021-12-27 ms/data_exploration.Rmd
Change to equal-area grid cells

@@
   slice(error_rows) %>%
-  select(grids, where(~any(is.na(.x))))
+  select(grids, richness, percent_apo, temp, precip, precip_season)
 ```
 
-The missing data was due to one grid-cell that lacks environmental data. This must be missing from the WorldClim dataset. Let's remove it, then continue.
+The missing data are due to two grid-cells that lack environmental data from the WorldClim dataset and one grid-cell that only had a single species with no breeding-system data.
+Let's remove these, then continue.
 
@@
 
-There are a few points beyond the whiskers that may be outliers. In particular, there is one value of `percent_apo` > 60% that seems quite outside the normal range.
+There are a few points beyond the whiskers that may be outliers. In particular, there is one value of `percent_apo` that seems quite outside the normal range.
 
@@
 
-Grubb's test confirms that the percent apomictic value > 60% is an outlier.
-It also indicates that the highest values of the variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, and `rpd_obs_z` may be outliers.
+Grubb's test confirms that the percent apomictic value is an outlier. It also indicates that the highest values of the variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, and `rpd_obs_z` may be outliers.
 
@@
 biodiv_ferns_cent_no_missing %>%
-  slice_max(1, order_by = percent_apo) %>%
+  slice_max(n = 1, order_by = percent_apo, with_ties = TRUE) %>%
   select(grids:long, percent_apo, richness)
@@
 
-There are just three species in the grid-cell with > 60% apomictic taxa. This data point should be excluded.
+There is just one species in the grid-cell that obviously contributes to the 100% apomixis rate. This data point should be excluded.
 
@@
 
-They all come from the same two grid-cells, but don't have an extremely low number of species.
+None of these have an extremely low number of species.
 

### 2021-12-27 ms/data_readme.Rmd
Specify commit hash for targets cache
Update data README date in Rmd
Change to equal-area grid cells

@@
 
+Corresponding commit in repo (https://github.com/joelnitta/japan_ferns_spatial_phy): FIXME add manually
+
 \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
@@
       "grids", "Name of grid-cell",
+      "richness", "Number of taxa occurring in grid-cell",
       "abundance", "Number of specimens occurring in grid-cell",
-      "richness", "Number of taxa occurring in grid-cell",
       "redundancy", "Redundancy (measure of sampling completeness)"
@@
       "grids", "Name of grid-cell",
+      "richness", "Number of taxa occurring in grid-cell",
       "abundance", "Number of specimens occurring in grid-cell",
-      "richness", "Number of taxa occurring in grid-cell",
       "redundancy", "Redundancy (measure of sampling completeness)",
@@
 
-2021-09-09
+2021-09-22
 

### 2021-12-27 ms/manuscript.Rmd
Change to equal-area grid cells

@@
 
-**Methods:** We compiled a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
+**Methods:** We compiled a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 20 km × 20 km grid-cells including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
 We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
@@
 
-repro_count$known$total
-
 repro_count$known$percent <- (repro_count$known$total/(repro_count$unknown$total + repro_count$known$total)) %>% percent(accuracy = 0.1)
@@
 
-We used a list of all native, non-hybrid fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
-We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v.
-1.01 (http://www.rdplants.org/gl/).
-Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article). 
+We used a list of native, non-hybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
+We chose to use this over other sources (e.g., GBIF) because of its extremely high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
+Furthermore, the overall sampling completeness of this dataset is also extremely high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article). 
 The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety (hereafter referred to as "taxa" unless otherwise indicated), excluding non-native taxa and nothotaxa.
-Latitude and longitude of each occurrence record were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
-We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
-The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].
-
-A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities. 
+All names are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/).
+
+Occurrence records were georeferenced and thoroughly vetted as follows [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
+For specimens that lacked latitude and logitude data, georeferencing was done by mapping collection site names to a set of standard ca.
+10 × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
+In the case that the collection site could not be mapped to a single second-degree mesh cell, it was excluded.
+Next, occurrence maps were generated for each taxon showing presence or absence in each second-degree mesh cell.
+In the case that a given taxon appeared insufficiently sampled (i.e., not present on the map where it would typically be expected to occur), AE and members of the Nippon Fernist Club (amateur botanists familiar with the ferns of Japan) searched for additional specimens, which were then added to the list.
+The occurrence maps were iteratively refined until the vast majority of known Japanese ferns had been observed across their expected ranges; the resulting set of occurrence records can be considered accurate to ca.
+10 km (the grain size of the second-degree mesh map).
+
+We further cleaned the list prior to analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
+
+A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities (co-occurring species). 
 Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
-We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). 
+We created sets of 10 km, 20 km, 30 km, and 40 km grid-cells covering Japan using a Mollweide equal-area projection (size refers to the length of the side of each square grid-cell).
 At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
 We calculated sampling redundancy [1 - richness/number of specimens; @Garcillan2003] to quantify adequacy of sampling. 
-Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
-We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.
+Preliminary analysis indicated that 20 km grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
+Although the grid-cells are defined with equal area, actual land area of each cell varies due to presence of coastlines, etc.
+Strictly filtering out all grid-cells with less than complete land area would result in a large loss of data, as Japan has many coastlines and small islands.
+Therefore, we instead filtered the grid-cells by sampling completeness, removing any cells with redundancy < 0.1; this dataset was used for all subsequent analyses.
+The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].
 
@@
 We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`).
-We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
-We calculated area as rolling mean area (km^2^) across 1° latitudinal bands.
+We identified the intersection of the climate data at the 2.5 minute scale with the 20 km grid-cells, then calculated the mean of each climatic variable for each grid-cell.
+We calculated area as rolling mean area (km^2^) across 100 km latitudinal bands.
 This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).
@@
 tibble(taxon = plastome_tree$tip.label) %>%
-  anti_join(select(plastome_sampling, taxon) %>% unique) %>%
-  anti_join(tibble(taxon = rownames(japan_rbcL))) %>%
+  anti_join(select(plastome_sampling, taxon) %>% unique, by = "taxon") %>%
+  anti_join(tibble(taxon = rownames(japan_rbcL)), by = "taxon") %>%
   verify(nrow(.) == 0, success_fun = success_logical)
@@
 
-Initial inspection of the biodiversity and environmental data showed a single outlier in % apomictic taxa due to extremely low species richness (Appendix S2; see Supplemental Data with this article), so that grid-cell and one grid-cell lacking environmental data were excluded from modeling analyses.
+Two grid cells lacking environmental data and an outlier with 100% apomictic taxa due to extremely low species richness were excluded from modeling analyses (Appendix S2; see Supplemental Data with this article).
 
@@
   mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
-  left_join(ppgi) %>%
+  left_join(ppgi, by = "genus") %>%
   filter(class == "Polypodiopsida")
@@
 ```{r discus-model-stats}
-# Check that conditional AIC was lower in the reproductive model compared to the environmental model for SES of RPD and SES of PD
+# Check that conditional AIC was lower in the reproductive model compared to 
+# the environmental model for SES of RPD
 # Split aic_env_repro dataframe into nested list for easier indexing
 c_aic <- aic_env_repro %>% split(.$resp_var) %>% map(~split(., .$model_type))
-assert_that(c_aic$pd_obs_z$percent_apo$conditional_aic < c_aic$pd_obs_z$temp$conditional_aic)
 assert_that(c_aic$rpd_obs_z$percent_apo$conditional_aic < c_aic$rpd_obs_z$temp$conditional_aic)
+# But not for SES of PD
+assert_that(c_aic$pd_obs_z$percent_apo$conditional_aic > c_aic$pd_obs_z$temp$conditional_aic)
 ```
@@
 SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F). 
-Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD and SES of PD (`r s_table("aic")`).
+Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (`r s_table("aic")`).
 
@@
 Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
-The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
+The other much smaller bioregions (including only one or two grid-cells each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
 
@@
       taxon = "Pteridium_aquilinum_japonicum",
-      text = "*Pteridium aquilinum* (L.) Kuhn subsp. *japonicum* (Nakai) A. et S. Löve"
+      text = "*Pteridium aquilinum* (L.) Kuhn subsp. *japonicum* (Nakai) A. et S. Löve" #nolint
     ),
@@
 **`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.    
-**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2° grid cells across Japan.  
+**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km grid cells across Japan.  
 **`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
-**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 0.2° grid cell in the ferns of Japan.  
+**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20 km grid cell in the ferns of Japan.  
 **`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
@@
   mutate(
-    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
-    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
+    endem_type = factor(
+      endem_type,
+      levels = c("neo", "paleo", "not significant", "mixed", "super")),
+    # can ignore warning about unknown levels, since they don't exist in all
+    # variables but that's OK
     across(
       contains("signif"),
-      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
-    )
-  )
+      ~fct_relevel(.x, 
+      "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
+    )
+  ) %>%
+  st_transform(4612)
 ```
@@
 main_islands <- c("Hokkaido", "Shikoku", "Kyushu", "Honshu")
-small_islands <- c("Okinawa", "Amami", "Ogasawara Islands", "Yakushima", "Izu Islands", "Iriomote", "Miyako")
+small_islands <- c("Okinawa", "Amami", "Ogasawara Islands", "Yakushima",
+  "Izu Islands", "Iriomote", "Miyako")
 other <- c("Tokyo")
@@
 Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells.
-Bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.
+Bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
 

### 2021-12-27 R/functions.R

@@ -201,6 +201,19 @@ load_jferns_comm <- function(csv_file) {
     column_to_rownames("grids")
 }
 
+#' Load Japan ferns shape data
+#'
+#' @param file_in Path to geometric shapes data
+#' @crs Coordinate reference system to use for shapes data
+#'
+#' @return data.frame
+load_jferns_shape <- function(file_in, crs) {
+  sf::st_read(file_in) %>%
+    # Convert 'grids' to character
+    mutate(grids = as.character(grids)) %>%
+    sf::st_transform(crs)
+}
+
 #' Run iNEXT on fern data
 #'
 #' @param occ_point_data_ferns Dataframe; raw point occurrence data,
@@ -736,10 +749,13 @@ summarize_fern_lat_span <- function (comm_ferns, shape_ferns) {
   # Get lat/long of each grid cell centroid
   centroids <-
     shape_ferns %>%
+    # Set CRS to JDG2000 = EPSG code 4612
+    st_transform(4612) %>%
     sf_add_centroids %>%
     select(grids, longitude = long, latitude = lat) %>%
     st_drop_geometry() %>%
-    as_tibble()
+    as_tibble() %>%
+    mutate(grids = as.character(grids))
   
   comm_ferns %>%
     # Add lat/long to community data
@@ -2163,17 +2179,22 @@ combine_bioregions <- function(regions_taxonomy, regions_phylogeny) {
 #'
 #' @return Dataframe; bioregions relabeled in descending order by mean latitude
 #' 
-relabel_bioregions_by_lat <- function(bioregions, shape_ferns, cutoff = 2) {
+relabel_bioregions_by_lat <- function(bioregions, shape_ferns, cutoff = 1) {
   
   # Summarize bioregions data: cells per cluster, mean latitude of each bioregion
   bioregions_summary_all <-
     shape_ferns %>%
     left_join(bioregions, by = "grids") %>%
+    # Set CRS to JDG2000 = EPSG code 4612 so that lat and long are in degrees
+    st_transform(4612) %>%
     sf_add_centroids() %>%
     st_drop_geometry() %>%
     select(contains("cluster"), lat, grids) %>%
     as_tibble() %>%
-    pivot_longer(names_to = "cluster_type", values_to = "cluster", -c(grids, lat)) %>%
+    pivot_longer(
+      names_to = "cluster_type",
+      values_to = "cluster",
+      -c(grids, lat)) %>%
     group_by(cluster_type, cluster) %>%
[…]

### 2021-12-27 R/ms_functions.R

@@ -52,8 +52,8 @@ bioregion_cols <- c(
   "5" = "#56B4E9", # light blue
   "6" = "#E69F00", # goldenrod
   "7" = "#CC79A7", # magenta
-  "8" ="#000000", # black
-  "Other" = "grey80") 
+  "8" = "#000000", # black
+  "Other" = "grey80")
 
 # - Protected areas (Okabe-Ito CVD safe)
 protection_cols <- c(

### 2021-12-27 ms/manuscript.Rmd
Update MS

@@
 We calculated sampling redundancy [1 - richness/number of specimens; @Garcillan2003] to quantify adequacy of sampling. 
-Preliminary analysis indicated that 20 km grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
+Preliminary analysis indicated that 20 km grid-cells are optimal for our dataset: there is a sudden improvement in redundancy values from 10 km to 20 km, but much less improvement as grain size is increased beyond 20 km (`r s_figure("grain-size")`).
 Although the grid-cells are defined with equal area, actual land area of each cell varies due to presence of coastlines, etc.

### 2022-01-06 ms/manuscript.Rmd
Update MS

@@
 
-Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
-However, using this dataset alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating.
+Sequences for the plastid, coding *rbcL* gene are available for ca.
+97% of the Japanese fern flora [@Ebihara2019b]; these sequences originate from samples that were identified using the same taxonomic system as the occurrence data [@Ebihara2010; @Ebihara2019b], so we used these preferentially over other data available on GenBank that might include misidentifications or different taxon concepts.
+However, using this dataset alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
+Futhermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
 Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
-To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
+Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
+This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across ca.
+4,000 taxa using six plastid markers [@Testo2016a], and the addition of plastome data can be expected to increase support along the backbone.
 We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].

### 2022-01-17 ms/manuscript.Rmd
Make lintr happy (ms)

@@
 # Set knitr options
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
+knitr::opts_chunk$set(
+  echo = FALSE, message = FALSE, warning = FALSE,
+  results = "hide", cache = FALSE)
 
@@
 ```{r intro-stats}
-# tally number of non-hybrid pteridophytes with known repro. mode
+# Tally number of non-hybrid pteridophytes with known repro. mode
 repro_count <-
@@
 
-repro_count$known$percent <- (repro_count$known$total/(repro_count$unknown$total + repro_count$known$total)) %>% percent(accuracy = 0.1)
+# Calculate percent of species with known reproductive mode
+repro_count$known$percent <-
+  repro_count$known$total %>%
+  magrittr::divide_by(repro_count$unknown$total + repro_count$known$total) %>%
+  percent(accuracy = 0.1)
 ```
@@
 tibble(taxon = plastome_tree$tip.label) %>%
-  anti_join(select(plastome_sampling, taxon) %>% unique, by = "taxon") %>%
+  anti_join(select(plastome_sampling, taxon) %>% unique, by = "taxon") %>% # nolint
   anti_join(tibble(taxon = rownames(japan_rbcL)), by = "taxon") %>%
@@
 ### Taxon list ###
-# Make list of all native ferns in Japan (based off Green List), *including* hybrids
+# Make list of all native ferns in Japan (based off Green List)
+# *including* hybrids
 fern_list <-
   green_list %>%
-  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
+  mutate(
+    genus = str_split(taxon, "_") %>%
+      map_chr(1)
+  ) %>%
   left_join(ppgi, by = "genus") %>%
@@
   filter(hybrid == FALSE) %>%
-  mutate(gn_parse_tidy(.$scientific_name) %>% select(taxon = canonicalsimple)) %>%
+  mutate(
+    gn_parse_tidy(.$scientific_name) %>%
+      select(taxon = canonicalsimple)
+  ) %>%
   mutate(taxon = str_replace_all(taxon, " ", "_"))
@@
 ### Phylogeny ###
-phylogeny_sampled_percent <- ape::Ntip(japan_fern_tree) %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)
+phylogeny_sampled_percent <- ape::Ntip(japan_fern_tree) %>%
+  magrittr::divide_by(total_fern_taxa) %>%
+  scales::percent(accuracy = 0.01)
 
 # Check height (age) of fern phylogeny
-phylogeny_age <- japan_fern_tree %>% phytools::nodeHeights() %>% max %>% round(1)
+phylogeny_age <- japan_fern_tree %>%
+  phytools::nodeHeights() %>%
+  max() %>%
+  round(1)
 
@@
     tip = japan_fern_tree$tip.label,
-    tip_no = 1:length(tip),
+    tip_no = seq_along(tip),
     genus = str_split(tip, "_") %>% map_chr(1)
@@
   summarize(tip_vec = list(tip_no), .groups = "drop") %>%
-  mutate(is_mono = map_lgl(tip_vec, ~ape::is.monophyletic(phy = japan_fern_tree, tips = .)))
+  mutate(is_mono = map_lgl(
+    tip_vec, ~ape::is.monophyletic(phy = japan_fern_tree, tips = .)))
 
@@
 
-non_mono_genus <- mono_status_genera %>% filter(is_mono == FALSE) %>% pull(genus)
+non_mono_genus <- mono_status_genera %>%
+  filter(is_mono == FALSE) %>%
+  pull(genus)
 
@@
     tip = japan_pterido_tree$tip.label,
-    tip_no = 1:length(tip),
+    tip_no = seq_along(tip),
     genus = str_split(tip, "_") %>% map_chr(1)
@@
   summarize(tip_vec = list(tip_no), .groups = "drop") %>%
-  mutate(is_mono = map_lgl(tip_vec, ~ape::is.monophyletic(phy = japan_pterido_tree, tips = .)))
+  mutate(is_mono = map_lgl(
+    tip_vec, ~ape::is.monophyletic(phy = japan_pterido_tree, tips = .)))
 
@@
 ### Specimens (filtered occurrence data) ###
-n_specimens <- occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")
+n_specimens <- occ_sum$filtered$n_specimens$value %>%
+  scales::number(big.mark = ",")
 n_specimens_taxa <- occ_sum$filtered$n_taxa$value
-n_specimens_taxa_percent <- n_specimens_taxa %>% magrittr::divide_by(total_fern_taxa) %>% scales::percent(accuracy = 0.01)
+n_specimens_taxa_percent <- n_specimens_taxa %>%
+  magrittr::divide_by(total_fern_taxa) %>%
+  scales::percent(accuracy = 0.01)
 
@@
 # Count number of traits of each type in final dataset
-n_binary_traits <- traits_summary %>% filter(trait_type == "binary") %>% nrow() %>% jntools::english2()
-n_cont_traits <- traits_summary %>% filter(trait_type == "continuous") %>% nrow() %>% jntools::english2()
-n_qual_traits <- traits_summary %>% filter(trait_type == "qualitative") %>% nrow() %>% jntools::english2()
-n_qual_traits_as_binary <- traits_summary %>% filter(trait_type == "qualitative") %>% pull(n) %>% sum %>% jntools::english2()
+n_binary_traits <- traits_summary %>%
+  filter(trait_type == "binary") %>%
+  nrow() %>%
+  jntools::english2()
+n_cont_traits <- traits_summary %>%
+  filter(trait_type == "continuous") %>%
+  nrow() %>%
+  jntools::english2()
+n_qual_traits <- traits_summary %>%
+  filter(trait_type == "qualitative") %>%
+  nrow() %>%
+  jntools::english2()
+n_qual_traits_as_binary <- traits_summary %>%
+  filter(trait_type == "qualitative") %>%
+  pull(n) %>%
+  sum() %>%
+  jntools::english2()
 
@@
 n_taxa_traits_sampled <- sum(trait_sampling_summary$sampled)
-percent_taxa_traits_sampled <- (n_taxa_traits_sampled / total_fern_taxa) %>% scales::percent(accuracy = 0.1)
+
+percent_taxa_traits_sampled <-
+  n_taxa_traits_sampled %>%
+  magrittr::divide_by(total_fern_taxa) %>%
+  scales::percent(accuracy = 0.1)
 
@@
 c_aic <- aic_env_repro %>% split(.$resp_var) %>% map(~split(., .$model_type))
-assert_that(c_aic$rpd_obs_z$percent_apo$conditional_aic < c_aic$rpd_obs_z$temp$conditional_aic)
+assert_that(
+  c_aic$rpd_obs_z$percent_apo$conditional_aic <
+    c_aic$rpd_obs_z$temp$conditional_aic)
 # But not for SES of PD
-assert_that(c_aic$pd_obs_z$percent_apo$conditional_aic > c_aic$pd_obs_z$temp$conditional_aic)
+assert_that(
+  c_aic$pd_obs_z$percent_apo$conditional_aic >
+    c_aic$pd_obs_z$temp$conditional_aic)
 ```
@@
 # Convert signif_cells_protected_area to wide format
-signif_cells_protected_area_wide <-
-  signif_cells_protected_area %>% select(status, metric, percent_protected) %>%
+signif_cells_protected_area_wide <- #nolint
+  signif_cells_protected_area %>%
+  select(status, metric, percent_protected) %>%
   pivot_wider(names_from = status, values_from = percent_protected) %>%
@@
 # Calculate min and max percent area protected overall
-min_protected_med <- signif_cells_protected_area_wide %>% pull(medium) %>% min %>% scales::percent(accuracy = 0.1)
-max_protected_med <- signif_cells_protected_area_wide %>% pull(medium) %>% max %>% scales::percent(accuracy = 0.1)
-min_protected_high <- signif_cells_protected_area_wide %>% pull(high) %>% min %>% scales::percent(accuracy = 0.1)
-max_protected_high <- signif_cells_protected_area_wide %>% pull(high) %>% max %>% scales::percent(accuracy = 0.1)
-min_protected_total <- signif_cells_protected_area_wide %>% pull(total) %>% min %>% scales::percent(accuracy = 0.1)
-max_protected_total <- signif_cells_protected_area_wide %>% pull(total) %>% max %>% scales::percent(accuracy = 0.1)
+# - define helper functions
+calc_min_pa <- function(signif_cells_protected_area_wide, ...) { # nolint
+  signif_cells_protected_area_wide %>%
+    pull(...) %>%
+    min() %>%
+    scales::percent(accuracy = 0.1)
+}
+calc_max_pa <- function(signif_cells_protected_area_wide, ...) { # nolint
+  signif_cells_protected_area_wide %>%
+    pull(...) %>%
+    max() %>%
+    scales::percent(accuracy = 0.1)
+}
+
+min_protected_med <- calc_min_pa(signif_cells_protected_area_wide, medium)
+max_protected_med <- calc_max_pa(signif_cells_protected_area_wide, medium)
+min_protected_high <- calc_min_pa(signif_cells_protected_area_wide, high)
+max_protected_high <- calc_max_pa(signif_cells_protected_area_wide, high)
+min_protected_total <- calc_min_pa(signif_cells_protected_area_wide, total)
+max_protected_total <- calc_max_pa(signif_cells_protected_area_wide, total)
 
 # Calculate min and max protected area for the best protected category
-signif_cells_protected_area_best <-
+signif_cells_protected_area_best <- # nolint
   signif_cells_protected_area_wide %>%
@@
 
-best_protected_med <- signif_cells_protected_area_best %>% pull(medium) %>% scales::percent(accuracy = 0.1)
-best_protected_high <- signif_cells_protected_area_best %>% pull(high) %>% scales::percent(accuracy = 0.1)
-best_protected_total <- signif_cells_protected_area_best %>% pull(total) %>% scales::percent(accuracy = 0.1)
+best_protected_med <- signif_cells_protected_area_best %>%
+  pull(medium) %>%
+  scales::percent(accuracy = 0.1)
+best_protected_high <- signif_cells_protected_area_best %>%
+  pull(high) %>%
+  scales::percent(accuracy = 0.1)
+best_protected_total <- signif_cells_protected_area_best %>%
+  pull(total) %>%
+  scales::percent(accuracy = 0.1)
 
@@
   japan_shp %>%
-  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
+  mutate(
+    area = st_area(.) %>%
+      units::set_units(km^2)
+  ) %>%
   pull(area)
 
-# Calculate baseline protection percentage: percent protected areas across the whole country
+# Calculate baseline protection percentage:
+# percent protected areas across the whole country
 total_percent_protected <-
@@
   st_make_valid %>%
-  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
+  mutate(
+    area = st_area(.) %>%
+      units::set_units(km^2)
+  ) %>%
   st_drop_geometry %>%
@@
   ) %>%
-  mutate(percent_protected = total_area/area_of_japan) %>%
+  mutate(percent_protected = total_area / area_of_japan) %>%
   mutate(percent_protected = as.numeric(percent_protected)) %>%
@@
 # Extract values for medium and high protection level
-total_percent_protected_medium <- total_percent_protected %>% filter(status == "medium") %>% pull(percent_protected)
-total_percent_protected_high <- total_percent_protected %>% filter(status == "high") %>% pull(percent_protected)
-total_percent_protected_all <- sum(total_percent_protected_medium, total_percent_protected_high)
+total_percent_protected_medium <- total_percent_protected %>%
+  filter(status == "medium") %>%
+  pull(percent_protected)
+total_percent_protected_high <- total_percent_protected %>%
+  filter(status == "high") %>%
+  pull(percent_protected)
+total_percent_protected_all <- sum(
+  total_percent_protected_medium, total_percent_protected_high)
 
@@
 
-# Calculate baseline protection percentage: percent protected areas across the whole country
+# Calculate baseline protection percentage:
+# percent protected areas across the whole country
 total_percent_deer_danger <-
   deer_range %>%
-  mutate(area = st_area(.) %>% units::set_units(km^2)) %>%
+  mutate(
+    area = st_area(.) %>% units::set_units(km^2)
+  ) %>%
   st_drop_geometry %>%
@@
   ) %>%
-  mutate(percent_danger = total_area/area_of_japan) %>%
+  mutate(percent_danger = total_area / area_of_japan) %>%
   mutate(percent_danger = as.numeric(percent_danger))
@@
     # Format numbers for nice printing
-    area = total_area %>% as.numeric %>% scales::number(accuracy = 1, big.mark = ",") %>% paste("km^2^"),
-    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
+    area = total_area %>%
+      as.numeric %>%
+      scales::number(accuracy = 1, big.mark = ",") %>%
+      paste("km^2^"),
+    percent = scales::percent(percent_danger, accuracy = 0.1)
+  ) %>%
   split(.$range)
@@
 
-n_southern_genera <- southern_genera %>% pull(higher_taxon) %>% n_distinct %>% english2
+n_southern_genera <- southern_genera %>%
+  pull(higher_taxon) %>%
+  n_distinct() %>%
+  english2()
 
@@
 
-n_southern_families <- southern_families %>% pull(higher_taxon) %>% n_distinct %>% english2
+n_southern_families <- southern_families %>%
+  pull(higher_taxon) %>%
+  n_distinct() %>%
+  english2()
 ```
@@
   geom_sf(size = 0.25) +
-  geom_segment(data = japan_lines, aes(x = lon, y = lat, xend = lon_end, yend = lat_end, linetype = query)) +
-  scale_linetype_manual(values = c("Tsugaru Strait" = "solid", "Tokara Strait" = "longdash")) +
+  geom_segment(
+    data = japan_lines,
+    aes(x = lon, y = lat, xend = lon_end, yend = lat_end, linetype = query)
+  ) +
+  scale_linetype_manual(
+    values = c("Tsugaru Strait" = "solid", "Tokara Strait" = "longdash")
+  ) +
   geom_richtext(
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("japan-map"), "pdf"), japan_map_plot, width = 5, height = 5, units = "in")
+ggsave(
+  result_file(figure("japan-map"), "pdf"),
+  japan_map_plot, width = 5, height = 5, units = "in")
 ```
@@
 a <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**A** Richness</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "No. taxa") +
@@
 b <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**B** PD</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "Relative\nbranch\nlength") +
@@
 c <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**C** FD</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "Relative\nbranch\nlength") +
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7.25, height = 4, units = "in", raw_diversity_plot)
+ggsave(
+  result_file(figure("raw-diversity"), "pdf"),
+  width = 7.25, height = 4, units = "in", raw_diversity_plot)
 ```
@@
   # white base map, no outline
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
   # colored cells
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent") +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent") + #nolint
   # light grey coastline
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**A** PD</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   scale_fill_manual(values = signif_cols) +
@@
 b <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**B** RPD</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "") +
@@
 c <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**C** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**C** FD</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "") +
@@
 d <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**D** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**D** RFD</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "") +
@@
 rand_diversity_plot <- a + b + c + d +
-  plot_layout(ncol = 2, nrow = 2, guides = 'collect') &
+  plot_layout(ncol = 2, nrow = 2, guides = "collect") &
   theme(legend.position = "bottom")
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("rand-diversity"), "pdf"), rand_diversity_plot, width = 7.25, height = 8, units = "in")
+ggsave(
+  result_file(figure("rand-diversity"), "pdf"),
+  rand_diversity_plot, width = 7.25, height = 8, units = "in")
 ```
@@
     resp_var = c("richness", "fd_obs_z", "pd_obs_z", "rfd_obs_z", "rpd_obs_z"),
-    resp_var_print = c("Richness", "SES of FD", "SES of PD", "SES of RFD", "SES of RPD")
+    resp_var_print = c(
+      "Richness", "SES of FD", "SES of PD", "SES of RFD", "SES of RPD")
   )
@@
   tibble(
-    term = c("lat_area", "I(temp^2)", "null_model", "percent_apo", "precip", "precip_season", "temp"),
-    term_print = c("Area", "Temp^2", "Null model", "% apomictic", "Precip.", "Precip.\nseasonality", "Temp.")
+    term = c("lat_area", "I(temp^2)", "null_model", "percent_apo", "precip",
+      "precip_season", "temp"),
+    term_print = c("Area", "Temp^2", "Null model", "% apomictic", "Precip.",
+      "Precip.\nseasonality", "Temp.")
   ) %>%
@@
     labs(
-      subtitle = glue("{unique(plot_data$resp_var_print)},\n{unique(plot_data$model_type_print)}"),
+      subtitle = glue("{unique(plot_data$resp_var_print)},\n{unique(plot_data$model_type_print)}"), #nolint
       fill = "Significance") +
@@
   ungroup %>%
-  left_join(model_subplot_order, by = c("resp_var", model_type = "indep_var")) %>%
+  left_join(
+    model_subplot_order, by = c("resp_var", model_type = "indep_var")) %>%
   arrange(order) %>%
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("coeff"), "pdf"), coeff_plot, width = 7.25, height = 5, units = "in")
+ggsave(
+  result_file(figure("coeff"), "pdf"),
+  coeff_plot, width = 7.25, height = 5, units = "in")
 ```
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("modelfit"), "pdf"), modelfit_plot, width = 7.25, height = 5, units = "in")
+ggsave(
+  result_file(figure("modelfit"), "pdf"),
+  modelfit_plot, width = 7.25, height = 5, units = "in")
 ```
@@
 endemism_plot <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
   scale_fill_manual(values = canape_cols) +
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("endemism"), "pdf"), endemism_plot, width = 5, height = 5, units = "in")
+ggsave(
+  result_file(figure("endemism"), "pdf"),
+  endemism_plot, width = 5, height = 5, units = "in")
 ```
@@
 a <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = taxonomic_cluster), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = taxonomic_cluster), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**A** Taxonomic</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "Bioregion") +
@@
 b <- ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = phylo_cluster), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**B** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = phylo_cluster), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
+  annotate(
+    geom = "richtext", y = Inf, x = -Inf,
+    label = "<span style='color:black'>**B** Phylogenetic</span>",
+    hjust = "inward", vjust = "inward", color = "white") +
   labs(fill = "Bioregion") +
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("bioregions"), "pdf"), bioregions_plot, width = 7.25, height = 5, units = "in")
+ggsave(
+  result_file(figure("bioregions"), "pdf"),
+  bioregions_plot, width = 7.25, height = 5, units = "in")
 ```
@@
   biodiv_ferns_spatial %>%
-  filter(taxonomic_cluster %in% c(1,2,3,4)) %>%
+  filter(taxonomic_cluster %in% c(1, 2, 3, 4)) %>%
   as_tibble %>%
-  select(pd_obs_z, rpd_obs_z, fd_obs_z, rfd_obs_z, pe_obs_p_upper, taxonomic_cluster)
+  select(
+    pd_obs_z, rpd_obs_z, fd_obs_z, rfd_obs_z,
+    pe_obs_p_upper, taxonomic_cluster)
 
@@
   scale_color_manual(values = bioregion_cols, breaks = 1:4) +
-  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  geom_boxplot(
+    color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
   labs(
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("div-by-tax-region"), "pdf"), div_by_tax_region_plot, width = 7.25, height = 7, units = "in")
+ggsave(
+  result_file(figure("div-by-tax-region"), "pdf"),
+  div_by_tax_region_plot, width = 7.25, height = 7, units = "in")
 ```
@@
 # Add total protected area (medium + high)
-signif_cells_protected_area_total <-
+signif_cells_protected_area_total <- #nolint
   signif_cells_protected_area %>%
@@
 
-total_percent_protected_for_plot <- 
+total_percent_protected_for_plot <- #nolint
   total_percent_protected %>%
@@
 # Save plot if rendering to doc
-ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 7.25, height = 4, units = "in")
+ggsave(
+  result_file(figure("conserv-status"), "pdf"),
+  conserv_status_plot, width = 7.25, height = 4, units = "in")
 ```

### 2022-01-17 ms/manuscript.Rmd
Update MS

@@
 We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
-We used hierarchical clustering to analyze bioregions.
-We compared the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer to assess conservation status and threats.
+Hierarchical clustering was used to analyze bioregions.
+Conservation status and threats were assessed by comparing the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer.
 
@@
 Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it, and to conserve it.
-The vast majority of studies seeking to understand the spatial distribution of biodiversity focus on species richness.
-Hotspots, or areas of exceptionally high richness, have received particular attention, resulting in a widely-recognized set of 36 terrestrial hotspots [@Myers2000; @Noss2016] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006].
-However, richness alone cannot provide a complete picture of biodiversity.
+Until recently, the vast majority of studies seeking to understand the spatial distribution of biodiversity have focused on species richness.
+Hotspots, or areas of exceptionally high species richness or endemism, have received particular attention, resulting in a widely-recognized set of 36 terrestrial hotspots [@Myers2000; @Noss2016] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006].
+However, species richness alone cannot provide a complete picture of biodiversity.
 All organisms are related to a greater or lesser degree by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of the biodiversity present in an area.
 
-Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently.
+Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently [@Soltis2016a; @Folk2021].
 These datasets and tools now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity [@Faith1992] and phylogenetic endemism [@Rosauer2009].
-Better ways to characterize biological regions ("bioregions") are also available [e.g., @Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
+Better ways to characterize biological regions ("bioregions"; `r ie`, areas defined by their taxonomic composition or evolutionary history) are also available [`r eg`, @Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
 Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide powerful insights into the processes structuring biodiversity and suggest conservation priorities.
@@
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b].
-One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]), which is likely to affect genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
+The ferns of Japan include 721 native, non-hyrbid taxa (including species and varieties) and hundreds of hybrids [@Ebihara2019b].
+The availability of detailed distribution data [distribution maps at the ca.
+10 km scale for all species\; @Ebihara2016b; @Ebihara2017], trait data [multiple quantitative and qualitative traits compiled for species identification of nearly all species\; @Ebihara2019b], and DNA sequences for >97% of non-hybrid species [@Ebihara2010; @Ebihara2019b] make the ferns of Japan an ideal system for investigating the relationships between, and drivers of, multiple dimensions of biodiversity.
+
+One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]) [@Ebihara2019b].
+Such densley sampled reproductive data for a large clade on the regional scale is rare, and we are unaware of any other comparable dataset in ferns.
+Reproductive mode is likely to affect population-level genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
+In ferns, clonally reproducing apomictic species are often polyploid hybrids that share identical plastid genotypes among taxa and within populations.
+Therefore, we expect high prevalence of apomictic species within a community to decrease phylogenetic diversity.
 
@@
 
-We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017].
-Continuous traits were measured on 10 randomly selected specimens per species.
-Fertile (`r ie`, spore-bearing) and sterile fronds were measured separately, then averaged.
+We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017], which were formatted to be used with Lucid software (https://www.lucidcentral.org/).
+Continuous traits were measured on 10 randomly selected specimens per species, then four values were obtained from these measurements following Lucid format: outside (`r ie` outside the typical range) minimum, typical minimum, typical maximum, and outside maximum.
+For species with dimorphic fertile (`r ie`, spore-bearing) and sterile fronds, fertile and sterile fronds were measured separately.
+We took the mean of the typical minimum and maximum values to use as the species mean value.
 Qualitative traits were scored by observing voucher specimens.
@@
 
-We downloaded distribution maps (SHP and TIF) of native Japanese deer (*Cervus nippon*), an herbivore that has caused extensive damage to native plant communities due to rapid population growth since the 1970s [@Takatsuki2009] and poses a high risk to ferns, from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
+Besides habitat loss due to human activity, one major threat to ferns in Japan is herbivory by native Japanese deer (*Cervus nippon*).
+Japanee deer have caused extensive damage to native plant communities due to rapid population growth since the 1970s [@Takatsuki2009].
+Extirpation of fern populations due to expansion of deer range has been observed in multiple sites across the country.
+Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status, and one taxon (**Arachniodes yasu-inouei** var.
+**angustipinnula**) is thought to have gone extinct due to deer herbivory, among other causes [@MOJ2015; @kankyosho2022].
+Therefore, analysis of the threat posed to fern biodiversity by deer herbivory is needed to inform future conservation policies.
+
+We downloaded distribution maps (SHP and TIF) of *Cervus nippon* from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
 To assess change in threat due to herbivory by deer over time, we used three maps available in this dataset: range of Japanese deer surveyed in 1978, range surveyed in 2003, and estimated range inferred from a model including snow cover and forest type based on the 2003 survey data [@kankyosho2021].

### 2022-01-18 ms/SI.Rmd
Update SI
Move description of grain-size selection to main MS
Update MS

@@
 (c), Distribution of redundancy values by grain size.
-There is a sudden improvement in redundancy values from 10 km to 20 km, but less improvement as grain size is increased beyond 20 km.
 
@@
 
-**`r s_figure("abundance")`**. Observed number of specimens (a) and sampling redundancy (b) per 0.2° grid cell in the ferns of Japan.
+**`r s_figure("abundance")`**. Observed number of specimens (a) and sampling redundancy (b) per 20 km grid cell in the ferns of Japan.
 
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
-  scale_fill_manual(values = canape_cols) +
+  scale_fill_manual(
+    values = canape_cols,
+    na.value = "transparent") +
   annotation_scale(location = "tl", width_hint = 0.2) +
@@
 
-a <- plot_genus("a", "Equisetum")
-b <- plot_genus("b", "Deparia")
-c <- plot_genus("c", "Cyrtomium")
-d <- plot_genus("d", "Dryopteris")
-e <- plot_genus("e", "Pteris")
-f <- plot_genus("f", "Diplazium") +
+a <- plot_genus("A", "Equisetum")
+b <- plot_genus("B", "Deparia")
+c <- plot_genus("C", "Cyrtomium")
+d <- plot_genus("D", "Dryopteris")
+e <- plot_genus("E", "Pteris")
+f <- plot_genus("F", "Diplazium") +
   annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

### 2022-01-18 ms/manuscript.Rmd
Update MS

@@
   - `r here::here("ms/references.yaml")`
+  - `r here::here("ms/references_other.bib")`
 ---
@@
 Hotspots, or areas of exceptionally high species richness or endemism, have received particular attention, resulting in a widely-recognized set of 36 terrestrial hotspots [@Myers2000; @Noss2016] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006].
-However, species richness alone cannot provide a complete picture of biodiversity.
+However, species richness alone cannot provide a complete picture of biodiversity [@Miller2018].
 All organisms are related to a greater or lesser degree by descent from a common ancestor, and these evolutionary relationships must be taken into account to obtain a full understanding of the biodiversity present in an area.
@@
 
-One particularly valuable characteristic of this flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]) [@Ebihara2019b].
-Such densley sampled reproductive data for a large clade on the regional scale is rare, and we are unaware of any other comparable dataset in ferns.
+One particularly valuable characteristic of the Japanese fern flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]) [@Ebihara2019b].
 Reproductive mode is likely to affect population-level genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
-In ferns, clonally reproducing apomictic species are often polyploid hybrids that share identical plastid genotypes among taxa and within populations.
+In ferns, clonally reproducing apomictic species are often polyploid hybrids that share identical plastid genotypes among taxa and within populations [@Grusz2009; @Chao2012a; @Hori2014; @Hori2019].
 Therefore, we expect high prevalence of apomictic species within a community to decrease phylogenetic diversity.
@@
 Furthermore, it is a mountainous country with great variation in elevation on the larger islands.
-Plant distributions often reflect physiological adaptations to climate (e.g., precipitation and temperature).
+Plant distributions often reflect physiological adaptations to climate [e.g., precipitation and temperature/; @Woodward1987].
 Therefore, we expect the spatial distribution of biodiversity to be determined by climatic variation in addition to reproductive mode.
@@
 We further cleaned the list prior to analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
+Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, @Zizka2019] have the potential to erroneously exclude true occurrence points (`r ie`, false positives), we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].
 
@@
 At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell).
-We calculated sampling redundancy [1 - richness/number of specimens; @Garcillan2003] to quantify adequacy of sampling.
+We calculated sampling redundancy [1 - richness/number of specimens\; @Garcillan2003] to quantify adequacy of sampling.
 Preliminary analysis indicated that 20 km grid-cells are optimal for our dataset: there is a sudden improvement in redundancy values from 10 km to 20 km, but much less improvement as grain size is increased beyond 20 km (`r s_figure("grain-size")`).
@@
 
-We measured taxonomic diversity using taxonomic richness.
+We measured taxonomic diversity using taxonomic richness (number of taxa in each community).
 
@@
 Initial inspection of the relationship between each biodiversity metric and temperature showed a hump-shaped pattern (Appendix S2), so we also included a quadratic term for temperature only.
-Richness was fit with the negative binomial response family; others were Gaussian.
+Richness was fit with the negative binomial response family; others (SES of PD, SES of FD, SES of RPD, SES of RFD) were Gaussian.
 We fit models including a Matérn correlation matrix based on the geographic centroids of each grid-cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014].
@@
 
-Besides habitat loss due to human activity, one major threat to ferns in Japan is herbivory by native Japanese deer (*Cervus nippon*).
-Japanee deer have caused extensive damage to native plant communities due to rapid population growth since the 1970s [@Takatsuki2009].
-Extirpation of fern populations due to expansion of deer range has been observed in multiple sites across the country.
-Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status, and one taxon (**Arachniodes yasu-inouei** var.
-**angustipinnula**) is thought to have gone extinct due to deer herbivory, among other causes [@MOJ2015; @kankyosho2022].
+Besides habitat loss due to human activity, one major threat to ferns in Japan is herbivory by Japanese deer (*Cervus nippon*).
+Although native, Japanee deer have caused extensive damage to native plant communities due to rapid population growth and range expansion since the 1970s [@Takatsuki2009].
+Decrease or extirpation of fern populations due to deer herbivory has been observed in multiple sites and species across the country [@Minamitani2005; @Yahara2006; @Hattori2010].
+Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status, and one taxon (*Arachniodes yasu-inouei* Sa.
+Kurata var.
+*angustipinnula* Seriz.) is thought to have gone extinct due to deer herbivory, among other causes [@MOJ2015; @kankyosho2022].
+Although not all species of ferns are equally susceptible to deer herbivory [a small number are even apparently unpalatable to deer\; @Minamitani2005], these studies show it is a clear danger to many species.
 Therefore, analysis of the threat posed to fern biodiversity by deer herbivory is needed to inform future conservation policies.
@@
 To assess change in threat due to herbivory by deer over time, we used three maps available in this dataset: range of Japanese deer surveyed in 1978, range surveyed in 2003, and estimated range inferred from a model including snow cover and forest type based on the 2003 survey data [@kankyosho2021].
-We limited estimated range to areas considered "highly likely" to be occupied by deer [movement cost < 0.10; @kankyosho2021].
+We limited estimated range to areas considered "highly likely" to be occupied by deer [movement cost < 0.10\; @kankyosho2021].
 
@@
 The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
-The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a] (`r s_figure("tree")`).
+The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a\; `r s_figure("tree")`).
 Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016].
@@
 The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
-Despite this, sampling redundancy was generally good [> 0.3; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise) (`r s_figure("abundance")`B).
+Despite this, sampling redundancy was generally good [> 0.3\; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise; `r s_figure("abundance")`B).
 The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid-cells.
@@
 
-After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits) (`r table("traits-used")`).
+After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits; Tables `r table_num("traits-used")`), `r s_table_num("phy-sig-binary")`.
 The trait matrix included `r n_taxa_traits_sampled` taxa (`r percent_taxa_traits_sampled`) and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.
@@
 All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.1, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
-`r n_d_signif` out of `r nrow(binary_sig_results)` of binary traits (including qualitative traits coded as binary) showed evidence of phylogenetic signal (*D* ≤ 0.5, significantly different from random) (`r s_table("phy-sig-binary")`).
+`r n_d_signif` out of `r nrow(binary_sig_results)` of binary traits (including qualitative traits coded as binary) showed evidence of phylogenetic signal (*D* ≤ 0.5, significantly different from random; `r s_table("phy-sig-binary")`).
 
@@
 Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`).
-The vast majority of grid-cells with significant endemism are mixed.
+The vast majority of grid-cells with significant endemism are mixed or superendemic.
 Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands.
@@
 Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
+The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and high elevation areas of northern Honshu), bioregion 2 (low elevation areas of northern Honshu, southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 covered most of nothern Honshu, while extending further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
 The other much smaller bioregions (including only one or two grid-cells each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
 
-Taxonomic bioregions 3 and 4 have higher SES values of PD, RPD, FD, and RFD than bioregions 1 and 2 (`r figure("div-by-tax-region")`).
-Bioregions 3 and 4 also have consistently high PE *p*-scores, whereas bioregions 1 and 2 have a much wider variation.
+Taxonomic bioregions 3 and 4 have higher mean SES values of PD, RPD, FD, and RFD than bioregions 1 and 2 (`r figure("div-by-tax-region")`).
+Bioregions 3 and 4 also have consistently high PE *p*-scores, whereas bioregions 1 and 2 have a much wider variation (`r figure("div-by-tax-region")`E).
 Bioregion 2 has lower RPD and RFD than bioregion 1.
@@
 The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A).
-Cells with significantly high biodiversity had a greater percentage of area protected than the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
+Cells with significantly high biodiversity had a similar or greater percentage of area protected compared to the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
 In contrast, cells with significantly high biodiversity had a lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
@@
 The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
-Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (`r figure("conserv-status")`B).
+Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (except for PE relative to 1978 levels; `r figure("conserv-status")`B).
 However, cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here (Figs.
-`r figure_num("raw-diversity")`a, `r s_figure_num("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b], and similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020].
+The latitudinal hump-shaped pattern of species richness observed in ferns here (Figs. `r figure_num("raw-diversity")`a, `r s_figure_num("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b].
+Similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020].
 Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
@@
 
-The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C).
-This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
-The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig")`, `r s_table_num("phy-sig-binary")`), and suggests that in ferns, phylogeny can be used as a reasonable stand-in for functional diversity.
-Given their high phylogenetic diversity, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism; indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands (Figs.
+The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern subtropical islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C).
+This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country, which lack subtropical climate; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
+The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig")`, `r s_table_num("phy-sig-binary")`), and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
+Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands (Figs.
 `r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`).
-This is clearly due to the small land area of these islands, which are the only areas with a subtropical climate in Japan.
+This is clearly due to the small land area of these islands: small area combined with high PD is expected to lead to high PE.
 
 Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`B).
-The long branches in the southern islands are likely being driven by the tropical lineages mentioned previously; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`A).
-The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera in that area such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).
+The long branches in the southern islands are likely being driven by the tropical lineages restricted to this area; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`A).
+The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera there such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).
 
@@
 Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (`r s_figure("genus-map")`).
-This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower genetic diversity, and is the first time to our knowledge that this has been demonstrated in ferns.
+This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower PD, and is the first time to our knowledge that this has been demonstrated in ferns.
 We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`).
@@
 Water limitation is thought to play an important role in promoting apomixis in ferns especially in deserts [@Grusz2021], since ferns depend on water for transfer of sperm to egg, and apomictic plants would be able to reproduce in the absence of water.
-A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomictic ferns [@Liu2012; @Tanaka2014], but in Japan, although the areas including high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
+A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomictic ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas including high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
 The monsoon hypothesis is also not supported by our results, which show precipitation seasonality did not play a strong role in any model and was not correlated with % apomictic taxa.
@@
 Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change.
-One other such threat is herbivory by native Japanese deer, *Cervus nippon*.
+One other such threat is herbivory by native Japanese deer, *Cervus nippon* [@Minamitani2005; @Yahara2006; @Hattori2010; @kankyosho2022].
 The rapid increase in population size observed in surveys from 1978 to 2003 is likely due to a combination of climate change and a decrease in human hunters and natural predators [@Takatsuki2009].
@@
 Fern traits used in this study.
-Letters in parentheses indicate trait type: b, binary; c, continuous; q, qualitative.
+Letters in parentheses indicate trait type: b, binary (trait present or absent); c, continuous; q, qualitative.
 Qualitative traits were coded as binary for analysis (see Materials and Methods).
+For complete list of trait states for qualitative traits, see `r s_table("phy-sig-binary")`.
 

### 2022-01-18 ms/SI.Rmd
Update MS

@@
 Effect of grain size on sampling redundancy.
-(a), Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km).
-(b), Percent of grid cells with redundancy = 0 by grain size.
-(c), Distribution of redundancy values by grain size.
+(A), Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km).
+(B), Percent of grid cells with redundancy = 0 by grain size.
+(C), Distribution of redundancy values by grain size.
 
@@
 Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan.
-(a), rolling mean area (km^2^) in 20 km latitudinal bands.
-(b), % apomictic taxa.
-(c), annual precipitation.
-(d), percipitation seasonality (coefficient of variation).
-(e), mean annual temperature.
-(f), temperature seasonality (standard deviation × 100).
+(A), rolling mean area (km^2^) in 20 km latitudinal bands.
+(B), % apomictic taxa.
+(C), annual precipitation.
+(D), percipitation seasonality (coefficient of variation).
+(E), mean annual temperature.
+(F), temperature seasonality (standard deviation × 100).
 Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.
@@
 
-**`r s_figure("abundance")`**. Observed number of specimens (a) and sampling redundancy (b) per 20 km grid cell in the ferns of Japan.
+**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 20 km grid cell in the ferns of Japan.
 
@@
 Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.
-(a), Phylogenetic diversity (PD).
-(b), Relative phylogenetic diversity (RPD).
-(c), Functional diversity (FD).
-(d), Relative functional diversity (RFD).
+(A), Phylogenetic diversity (PD).
+(B), Relative phylogenetic diversity (RPD).
+(C), Functional diversity (FD).
+(D), Relative functional diversity (RFD).
 Trendlines in (a, c) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
@@
 Color shows bioregion; bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
-(a), taxonomic bioregions.
-(b), phylogenetic bioregions.
+(A), taxonomic bioregions.
+(B), phylogenetic bioregions.
 Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].

### 2022-01-18 ms/manuscript.Rmd
Update MS

@@
 
-One particularly valuable characteristic of the Japanese fern flora is the availability of data on reproductive mode (known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`]) [@Ebihara2019b].
+One particularly valuable characteristic of the Japanese fern flora is the availability of data on reproductive mode, which is known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`\; @Ebihara2019b].
 Reproductive mode is likely to affect population-level genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
@@
 Furthermore, it is a mountainous country with great variation in elevation on the larger islands.
-Plant distributions often reflect physiological adaptations to climate [e.g., precipitation and temperature/; @Woodward1987].
+Plant distributions often reflect physiological adaptations to climate [e.g., precipitation and temperature\; @Woodward1987].
 Therefore, we expect the spatial distribution of biodiversity to be determined by climatic variation in addition to reproductive mode.
@@
 
-We measured phylogenetic diversity (PD) using the Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny; @Faith1992].
+We measured phylogenetic diversity (PD) using the Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny\; @Faith1992].
 To detect areas with unusually long or short branches, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].
@@
 The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
-The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a\; `r s_figure("tree")`).
+The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a\; `r s_figure("tree")`].
 Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016].
@@
 Water limitation is thought to play an important role in promoting apomixis in ferns especially in deserts [@Grusz2021], since ferns depend on water for transfer of sperm to egg, and apomictic plants would be able to reproduce in the absence of water.
-A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomictic ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas including high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
-The monsoon hypothesis is also not supported by our results, which show precipitation seasonality did not play a strong role in any model and was not correlated with % apomictic taxa.
+A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomixis in ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas including high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
+The monsoon hypothesis is not supported by our results, which did not show a particularly strong effect of precipitation seasonality in most models `r figure("coeff")`, nor a correlation with % apomictic taxa (`r s_table("corr")`).
 
@@
 
-We found that grid-cells with significantly high amounts of biodiversity are better protected than the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`A).
+We found that grid-cells with significantly high amounts of biodiversity are similarly or better protected compared to the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`A).
 The high protection status of grid-cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`D).
@@
 Ebihara, pers.
-comm.; @kankyosho2021b].
+comm.\; @kankyosho2021b].
 Also, (as suggested previously) ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
@@
 
-Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`B).
+Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate in recent surveys or modeled data, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`B).
 This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("protected-map")`A).
@@
 One element of biodiversity that we did not consider here is hybrids.
-The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa.
+The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa [@Ebihara2019b].
 Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989].
@@
 Here, as in the vast majority of fern ecology studies, we consider only the sporophyte stage.
-However, recent studies in Japan [@Ebihara2013; @Ebihara2019a] and elsewhere [@Nitta2017] have revealed different patterns in the distribution of gametophytes and sporophytes.
+However, recent studies in Japan [@Ebihara2013; @Ebihara2019a; @Murakami2021] and elsewhere [@Nitta2017] have revealed different patterns in the distribution of gametophytes and sporophytes.
 The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA sequencing analyses to compare patterns of biodiversity between life stages in ferns across large spatial scales.
@@
 
-The authors thank Members of the Iwasaki Lab at the University of Tokyo for providing comments that improved the manuscript.
+The authors thank Members of the Iwasaki Lab at the University of Tokyo and two anonymous reviewers for providing comments that improved the manuscript.
 Shawn Laffan provided advice about CANAPE.
@@
 Data files needed to run the analysis and selected results files are available from Figshare at http://doi.org/10.6084/m9.figshare.16655263.
-**Note to reviewers**: the Figshare DOI will not be accessible until the paper is accepted. During review, the data can be accessed from this link: https://figshare.com/s/de3d5a70e118208e56c5.
 

### 2022-01-18 ms/manuscript.Rmd
Use AJB templates

@@
   bookdown::word_document2:
-    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
+    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
     toc: no
     number_sections: no
-    reference_docx: template.docx
+    reference_docx: american-journal-of-botany.docx
   bookdown::pdf_document2:

### 2022-01-18 ms/manuscript.Rmd
Update MS
Follow AJB format

@@
 
+Joel H. Nitta^1,5^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^
+
+^1^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, Tokyo, Japan
+
+^2^University and Jepson Herbaria, and Department of Integrative Biology, University of California, Berkeley, CA, USA
+
+^3^Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan
+
+^4^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
+
+^5^Author for correspondence: joelnitta@gmail.com
+
+```{r author-page-doc, results = "asis", echo = FALSE, eval = params$doc_type == "docx"}
+# Author page, version for submission to journal
+cat('
+Manuscript received _______; revision accepted _______.
+
 Short running title: Spatial phylogenetics of Japanese ferns
-
-Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^
-
-^1^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, Tokyo, Japan
-
-^2^University and Jepson Herbaria, and Department of Integrative Biology, University of California, Berkeley, CA, USA
-
-^3^Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan
-
-^4^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
-
-Author for correspondence:
-*Joel H. Nitta*
-*Email: joelnitta@gmail.com*
-
-## Acknowledgements
-
-Members of the Iwasaki Lab at the University of Tokyo provided comments that improved the manuscript.
-Shawn Laffan provided advice about CANAPE.
-This study was supported by JSPS KAKENHI Grant Number 16H06279.
+')
+```
 
@@
 
-## Introduction
-
 ```{r intro-stats}
@@
 
-### Study Site
+## Study Site
 
@@
 We further cleaned the list prior to analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
-Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, @Zizka2019] have the potential to erroneously exclude true occurrence points (`r ie`, false positives), we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].
+Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, CoordinateCleaner\; @Zizka2019] have the potential to erroneously exclude true occurrence points (`r ie`, false positives), we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].
 
@@
 Besides habitat loss due to human activity, one major threat to ferns in Japan is herbivory by Japanese deer (*Cervus nippon*).
-Although native, Japanee deer have caused extensive damage to native plant communities due to rapid population growth and range expansion since the 1970s [@Takatsuki2009].
+Although native, Japanese deer have caused extensive damage to native plant communities due to rapid population growth and range expansion since the 1970s [@Takatsuki2009].
 Decrease or extirpation of fern populations due to deer herbivory has been observed in multiple sites and species across the country [@Minamitani2005; @Yahara2006; @Hattori2010].
@@
 *angustipinnula* Seriz.) is thought to have gone extinct due to deer herbivory, among other causes [@MOJ2015; @kankyosho2022].
-Although not all species of ferns are equally susceptible to deer herbivory [a small number are even apparently unpalatable to deer\; @Minamitani2005], these studies show it is a clear danger to many species.
+Although not all species of ferns are equally susceptible to deer herbivory [a small number are apparently unpalatable to deer\; @Minamitani2005], these studies show it is a clear danger to many species.
 Therefore, analysis of the threat posed to fern biodiversity by deer herbivory is needed to inform future conservation policies.
@@
 
-## Results
-
-### Datasets
+# RESULTS
+
+## Datasets
 
@@
 
-## Discussion
-
-### Distribution and drivers of biodiversity
+# DISCUSSION
+
+## Distribution and drivers of biodiversity
 
@@
 
-### Limits of the study and future outlook
+## Limits of the study and future outlook
 
@@
 
-# Acknowledgements
+## ACKNOWLEDGEMENTS
 
@@
 
-# Author contributions
+## AUTHOR CONTRIBUTIONS
 
@@
 
-# Data availability statement
+## DATA AVAILABILITY STATEMENT
 
@@
 
-# Online supporting information
+## SUPPORTING INFORMATION
 
@@
 
-# Literature cited
+## References
 <!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
@@
 Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan.
-(a) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status.
-(b) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
+(A) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status.
+(B) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
 Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
 Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
-For (a), vertical lines indicate total percent area protected across Japan (baseline protection rate).
-For (b), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).
-
-`r pagebreak_doc()`
-
-```{r japan-map-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-# Print plot if rendering to doc on initial submission
-japan_map_plot
-cat("\n\n", figure("japan-map"))
-```
-
-`r pagebreak_doc()`
-
-```{r raw-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
-raw_diversity_plot
-cat("\n\n", figure("raw-diversity"))
-```
-
-`r pagebreak_doc()`
-
-```{r rand-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 9, eval = params$doc_type == "doc", results = "asis"}
-rand_diversity_plot
-cat("\n\n", figure("rand-diversity"))
-```
-
-`r pagebreak_doc()`
-
-```{r coeff-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-coeff_plot
-cat("\n\n", figure("coeff"))
-```
-
-`r pagebreak_doc()`
-
-```{r modelfit-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-modelfit_plot
-cat("\n\n", figure("modelfit"))
-```
-
-`r pagebreak_doc()`
-
-```{r endemism-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-endemism_plot
-cat("\n\n", figure("endemism"))
-```
-
-`r pagebreak_doc()`
-
-```{r bioregions-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
-bioregions_plot
-cat("\n\n", figure("bioregions"))
-```
-
-`r pagebreak_doc()`
-
-```{r div-by-tax-region-doc, fig.cap = "", fig.width = 7, fig.height = 7, eval = params$doc_type == "doc", results = "asis"}
-div_by_tax_region_plot
-cat("\n\n", figure("div-by-tax-region"))
-```
-
-`r pagebreak_doc()`
-
-```{r conserv-status-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
-conserv_status_plot
-cat("\n\n", figure("conserv-status"))
-```
-
-`r pagebreak_doc()`
-
-## Data availability statement
-
-Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy.
-A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
-Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.
-
-## References
-<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
-<div id="refs"></div>
-
-## Biosketch
-
-Joel H. Nitta's research focuses on patterns and processes of diversification in ferns. For more information, see https://joelnitta.com.
-
-Author contributions: J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.
+For (A), vertical lines indicate total percent area protected across Japan (baseline protection rate).
+For (B), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).

### 2022-01-18 ms/manuscript.Rmd
Update MS
Use correct paths for templates

@@
   "refs_yaml",
+  "refs_other_bib",
   "repro_data",
@@
 ```{r intro-stats}
+# Make list of all native ferns in Japan (based off Green List)
+# *including* hybrids
+fern_list <-
+  green_list %>%
+  mutate(
+    genus = str_split(taxon, "_") %>%
+      map_chr(1)
+  ) %>%
+  left_join(ppgi, by = "genus") %>%
+  assert(not_na, class) %>%
+  filter(class == "Polypodiopsida")
+
+fern_list_non_hybrid <-
+  fern_list %>%
+  filter(hybrid == FALSE) %>%
+  mutate(
+    gn_parse_tidy(.$scientific_name) %>%
+      select(taxon = canonicalsimple)
+  ) %>%
+  mutate(taxon = str_replace_all(taxon, " ", "_"))
+
+# Caculate number of native, non-hybrid fern taxa
+total_fern_taxa <- fern_list_non_hybrid %>%
+  pull(taxon) %>%
+  n_distinct()
+
 # Tally number of non-hybrid pteridophytes with known repro. mode
@@
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b].
-The ferns of Japan include 721 native, non-hyrbid taxa (including species and varieties) and hundreds of hybrids [@Ebihara2019b].
+The ferns of Japan include `r total_fern_taxa` native, non-hyrbid taxa (including species and varieties) and hundreds of hybrids [@Ebihara2019b].
 The availability of detailed distribution data [distribution maps at the ca.
@@
 ### Taxon list ###
-# Make list of all native ferns in Japan (based off Green List)
-# *including* hybrids
-fern_list <-
-  green_list %>%
-  mutate(
-    genus = str_split(taxon, "_") %>%
-      map_chr(1)
-  ) %>%
-  left_join(ppgi, by = "genus") %>%
-  filter(class == "Polypodiopsida")
-
-fern_list_non_hybrid <-
-  fern_list %>%
-  filter(hybrid == FALSE) %>%
-  mutate(
-    gn_parse_tidy(.$scientific_name) %>%
-      select(taxon = canonicalsimple)
-  ) %>%
-  mutate(taxon = str_replace_all(taxon, " ", "_"))
-
-# Caculate number of native, non-hybrid fern taxa
-total_fern_taxa <- fern_list_non_hybrid %>%
-  pull(taxon) %>%
-  n_distinct()
+# Make list of fern taxa `total_fern_taxa` in chunk `intro-stats`
 
@@
 Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells.
-Bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
+Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
 

### 2022-01-20 ms/manuscript.Rmd
Update MS

@@
 
-# Tally number of non-hybrid pteridophytes with known repro. mode
+# Tally number of non-hybrid ferns with known repro. mode
 repro_count <-
   repro_data %>%
+  # Filter to only non-hybrid ferns
+  inner_join(
+    select(fern_list_non_hybrid, taxon_id),
+    by = "taxon_id"
+  ) %>%
   count(reproductive_mode) %>%
@@
 Until recently, the vast majority of studies seeking to understand the spatial distribution of biodiversity have focused on species richness.
-Hotspots, or areas of exceptionally high species richness or endemism, have received particular attention, resulting in a widely-recognized set of 36 terrestrial hotspots [@Myers2000; @Noss2016] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006].
+Hotspots, or areas of exceptionally high species richness or endemism, have received particular attention, resulting in a widely recognized set of 36 terrestrial hotspots [@Myers2000; @Noss2016] and motivating conservation strategies to preserve them [@Margules2000; @Brooks2006].
 However, species richness alone cannot provide a complete picture of biodiversity [@Miller2018].
@@
 These datasets and tools now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity [@Faith1992] and phylogenetic endemism [@Rosauer2009].
-Better ways to characterize biological regions ("bioregions"; `r ie`, areas defined by their taxonomic composition or evolutionary history) are also available [`r eg`, @Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
+Better ways to characterize biological regions ("bioregions"; `r ie`, areas defined by their taxonomic composition or evolutionary history) are also available [@Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
 Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide powerful insights into the processes structuring biodiversity and suggest conservation priorities.
@@
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b].
-The ferns of Japan include `r total_fern_taxa` native, non-hyrbid taxa (including species and varieties) and hundreds of hybrids [@Ebihara2019b].
+The ferns of Japan include `r total_fern_taxa` native, non-hybrid taxa (including species and varieties) and hundreds of hybrids [@Ebihara2019b].
 The availability of detailed distribution data [distribution maps at the ca.
-10 km scale for all species\; @Ebihara2016b; @Ebihara2017], trait data [multiple quantitative and qualitative traits compiled for species identification of nearly all species\; @Ebihara2019b], and DNA sequences for >97% of non-hybrid species [@Ebihara2010; @Ebihara2019b] make the ferns of Japan an ideal system for investigating the relationships between, and drivers of, multiple dimensions of biodiversity.
-
-One particularly valuable characteristic of the Japanese fern flora is the availability of data on reproductive mode, which is known for `r repro_count$known$total` of `r repro_count$known$total + repro_count$unknown$total` native pteridophyte taxa excluding hybrids [`r repro_count$known$percent`\; @Ebihara2019b].
+10 km scale for nearly all species\; @Ebihara2016b; @Ebihara2017], trait data [multiple quantitative and qualitative traits compiled for identification of nearly all species\; @Ebihara2019b], and DNA sequences for > 97% of species [@Ebihara2010; @Ebihara2019b\; all coverage statistics exclude hybrids] make the ferns of Japan an ideal system for investigating the relationships between, and drivers of, multiple dimensions of biodiversity.
+
+One particularly valuable characteristic of the Japanese fern flora is the availability of data on reproductive mode, which is known for `r repro_count$known$total` native fern taxa excluding hybrids [`r repro_count$known$percent`\; @Ebihara2019b].
 Reproductive mode is likely to affect population-level genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
@@
 We used a list of native, non-hybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
-We chose to use this over other sources (e.g., GBIF) because of its extremely high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
-Furthermore, the overall sampling completeness of this dataset is also extremely high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article).
-The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety (hereafter referred to as "taxa" unless otherwise indicated), excluding non-native taxa and nothotaxa.
+We chose to use this over other sources (e.g., GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
+Furthermore, the overall sampling completeness of this dataset is also high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article).
+The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety, excluding non-native taxa and nothotaxa.
+For the purposes of analysis, we treated infraspecific taxa as distinct species and hereafter use "taxon" and "species" interchangeably to refer to entities at this rank, unless otherwise indicated.
 All names are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/).
@@
 Occurrence records were georeferenced and thoroughly vetted as follows [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
-For specimens that lacked latitude and logitude data, georeferencing was done by mapping collection site names to a set of standard ca.
+For specimens that lacked latitude and longitude data, georeferencing was done by mapping collection site names to a set of standard ca.
 10 × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
@@
 We further cleaned the list prior to analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
-Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, CoordinateCleaner\; @Zizka2019] have the potential to erroneously exclude true occurrence points (`r ie`, false positives), we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].
+Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, CoordinateCleaner\; @Zizka2019] have the potential to erroneously exclude true occurrence points [`r ie`, false positives\; @Zizka2020], we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].
 
@@
 Preliminary analysis indicated that 20 km grid-cells are optimal for our dataset: there is a sudden improvement in redundancy values from 10 km to 20 km, but much less improvement as grain size is increased beyond 20 km (`r s_figure("grain-size")`).
-Although the grid-cells are defined with equal area, actual land area of each cell varies due to presence of coastlines, etc.
+Although the grid-cells are defined with equal area, actual land area of each cell varies due to coastlines.
 Strictly filtering out all grid-cells with less than complete land area would result in a large loss of data, as Japan has many coastlines and small islands.
@@
 We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017], which were formatted to be used with Lucid software (https://www.lucidcentral.org/).
-Continuous traits were measured on 10 randomly selected specimens per species, then four values were obtained from these measurements following Lucid format: outside (`r ie` outside the typical range) minimum, typical minimum, typical maximum, and outside maximum.
+Continuous traits were measured on 10 randomly selected specimens per species, then four values per species were obtained from these measurements following Lucid format: outside (`r ie` outside the typical range) minimum, typical minimum, typical maximum, and outside maximum.
 For species with dimorphic fertile (`r ie`, spore-bearing) and sterile fronds, fertile and sterile fronds were measured separately.
@@
 All qualitative traits were scored in binary format; `r eg`, a single trait with three states was formatted as three binary traits.
-From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`), calculated mean values for continuous traits, and removed any traits that had Pearson correlation coefficient >0.6 or fewer than three observations of a given trait state.
+From the original trait list, we selected only traits with putative ecological function (`r table("traits-used")`) and excluded any traits that had Pearson correlation coefficient > 0.6 or fewer than three observations of a given trait state.
 
@@
 We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown.
-We calculated the percentage of apomictic species in each grid-cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).
+We calculated the percentage of apomictic taxa in each grid-cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).
 
@@
 However, using this dataset alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
-Futhermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
+Furthermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
 Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree.
@@
 We excluded marine zones and those that do not protect plants.
-We categorized protected areas as either "high" (no human activities allowed at all) or "medium" status (some economic activities allowed by permit) [@Kusumoto2017]; areas not afforded at least medium level of protection were not considered.
+We categorized protected areas as either "high" (no human activities allowed at all) or "medium" status [some economic activities allowed by permit\; @Kusumoto2017]; areas not afforded at least medium level of protection were not considered.
 Conservation status in Japan is administered by multiple laws, and protected areas frequently overlap [@Natori2012].
@@
 
-After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits; Tables `r table_num("traits-used")`), `r s_table_num("phy-sig-binary")`.
+After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits; Tables `r table_num("traits-used")`, `r s_table_num("phy-sig-binary")`.
 The trait matrix included `r n_taxa_traits_sampled` taxa (`r percent_taxa_traits_sampled`) and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.
@@
 Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`).
-The vast majority of grid-cells with significant endemism are mixed or superendemic.
+The vast majority of grid-cells with significant endemism are mixed or super-endemic.
 Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands.
@@
 The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and high elevation areas of northern Honshu), bioregion 2 (low elevation areas of northern Honshu, southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 covered most of nothern Honshu, while extending further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
-The other much smaller bioregions (including only one or two grid-cells each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 covered most of northern Honshu, while extending further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
+The other much smaller bioregions (including only one or two grid-cells each) are likely artifacts of insufficiently low taxon richness for robust clustering and are not discussed further.
 
@@
 Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`B).
-The long branches in the southern islands are likely being driven by the tropical lineages restricted to this area; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high-latitudes, such as *Equisetum* (`r s_figure("genus-map")`A).
+The long branches in the southern islands are likely being driven by the tropical lineages restricted to this area; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high latitudes, such as *Equisetum* (`r s_figure("genus-map")`A).
 The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera there such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).
@@
 A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomixis in ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas including high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
-The monsoon hypothesis is not supported by our results, which did not show a particularly strong effect of precipitation seasonality in most models `r figure("coeff")`, nor a correlation with % apomictic taxa (`r s_table("corr")`).
+The monsoon hypothesis was not supported by our results, which did not show a particularly strong effect of precipitation seasonality in most models (`r figure("coeff")`), nor a correlation with % apomictic taxa (`r s_table("corr")`).
 
@@
 Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
-Furthermore, the border between bioregions 1 and 2 corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
+Furthermore, the border between bioregions 1 and 2 approximately corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
 This suggests that ferns are useful as bioindicators in Japan.
@@
 
-## Limits of the study and future outlook
+# CONCLUSIONS
 
@@
 
-There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research.
+Our study integrates diverse, thoroughly collected datasets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns.
+We conclude by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.
+
 First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan.
@@
 
-One element of biodiversity that we did not consider here is hybrids.
-The Japanese fern flora includes `r n_hybrid_fern_taxa` hybrid taxa [@Ebihara2019b].
+One element of biodiversity that we did not consider here is hybrids (nothotaxa).
+The Japanese fern flora includes `r n_hybrid_fern_taxa` nothotaxa [@Ebihara2019b].
 Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989].
@@
 Here, as in the vast majority of fern ecology studies, we consider only the sporophyte stage.
-However, recent studies in Japan [@Ebihara2013; @Ebihara2019a; @Murakami2021] and elsewhere [@Nitta2017] have revealed different patterns in the distribution of gametophytes and sporophytes.
+However, recent studies in Japan [@Ebihara2013] and elsewhere [@Nitta2017] have revealed different patterns in the community structure of gametophytes and sporophytes.
 The comprehensive molecular sampling of sporophytes in Japan makes this area ideal for conducting high-throughput DNA sequencing analyses to compare patterns of biodiversity between life stages in ferns across large spatial scales.

### 2022-01-20 ms/SI.Rmd
Fix CSL file for SI

@@
     latex_engine: xelatex
-    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
+    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
 header-includes: 
@@
   "biodiv_ferns_spatial_lumped",
-  "bioregion_cutoff"
+  "bioregion_cutoff",
+  "ms_functions"
 ))
@@
 
-```{r load-captions}
+```{r load-captions, cache = FALSE}
 # Source captions numbers from MS
@@
 **`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position. Significantly correlated variables at *P* < 0.05 in bold.
-
-\clearpage
-
-```{r phy-sig, results = "as-is"}
-phy_sig_results %>%
-  mutate(across(c(kval, lambda, k_pval, lambda_pval), ~format(., digits = 2))) %>%
-  # format for kable/latex
-  mutate(trait = str_replace_all(trait, "_", "\\\\_")) %>%
-  rename(
-    "Trait" = trait,
-    `\\textit{K}` = kval, `\\textit{P}(\\textit{K})` = k_pval, 
-    `λ` = lambda, `\\textit{P}(λ)` = lambda_pval,
-    `\\textit{N}` = n) %>%
-  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
-  kable_styling(latex_options = c("striped", "HOLD_position"))
-```
-
-**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan. *K*, Blomberg's *K*; λ, Pagel's λ.
 
@@
 Values of *D* rounded to three decimal places.
+
+\clearpage
+
+```{r phy-sig, results = "as-is"}
+phy_sig_results %>%
+  mutate(across(c(kval, lambda, k_pval, lambda_pval), ~format(., digits = 2))) %>%
+  # format for kable/latex
+  mutate(trait = str_replace_all(trait, "_", "\\\\_")) %>%
+  rename(
+    "Trait" = trait,
+    `\\textit{K}` = kval, `\\textit{P}(\\textit{K})` = k_pval, 
+    `λ` = lambda, `\\textit{P}(λ)` = lambda_pval,
+    `\\textit{N}` = n) %>%
+  kbl(booktabs = TRUE, escape = FALSE, format = "latex") %>%
+  kable_styling(latex_options = c("striped", "HOLD_position"))
+```
+
+**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan. *K*, Blomberg's *K*; λ, Pagel's λ.
 

### 2022-01-20 ms/manuscript.Rmd
Update MS

@@
 Furthermore, it is a mountainous country with great variation in elevation on the larger islands.
-Plant distributions often reflect physiological adaptations to climate [e.g., precipitation and temperature\; @Woodward1987].
+Plant distributions often reflect physiological adaptations to climate [`r eg`, precipitation and temperature\; @Woodward1987].
 Therefore, we expect the spatial distribution of biodiversity to be determined by climatic variation in addition to reproductive mode.
@@
 We used a list of native, non-hybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
-We chose to use this over other sources (e.g., GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
+We chose to use this over other sources (`r eg`, GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
 Furthermore, the overall sampling completeness of this dataset is also high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article).
@@
 Next, occurrence maps were generated for each taxon showing presence or absence in each second-degree mesh cell.
-In the case that a given taxon appeared insufficiently sampled (i.e., not present on the map where it would typically be expected to occur), AE and members of the Nippon Fernist Club (amateur botanists familiar with the ferns of Japan) searched for additional specimens, which were then added to the list.
+In the case that a given taxon appeared insufficiently sampled (`r ie`, not present on the map where it would typically be expected to occur), AE and members of the Nippon Fernist Club (local botanists familiar with the ferns of Japan) searched for additional specimens, which were then added to the list.
 The occurrence maps were iteratively refined until the vast majority of known Japanese ferns had been observed across their expected ranges; the resulting set of occurrence records can be considered accurate to ca.
@@
 This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country, which lack subtropical climate; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
-The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig")`, `r s_table_num("phy-sig-binary")`), and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
+The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig-binary")`, `r s_table_num("phy-sig")`), and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
 Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands (Figs.
@@
 **`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
+**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
 **`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
-**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
 **`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  

### 2022-01-21 ms/SI.Rmd
Update MS

@@
 Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods).
-Color shows bioregion; bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
+Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
 (A), taxonomic bioregions.
@@
 Points shown with partial transparency; darker areas indicate overlapping points.
-Bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
+Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
 

### 2022-01-21 ms/manuscript.Rmd
Update MS

@@
 
-```{r author-page-doc, results = "asis", echo = FALSE, eval = params$doc_type == "docx"}
+```{r author-page-doc, results = "asis", echo = FALSE, eval = params$doc_type == "doc"}
 # Author page, version for submission to journal
@@
 
-All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The R package `r pack("targets")` was used to control analysis workflow [@Landau2021].
+All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2021] unless otherwise stated. The R package `r pack("targets")` was used to control analysis workflow [@Landau2021].
 
@@
 We also tested alternative weighting schemes, but these produced very similar results (data not shown).
-We then used the distances to build a trait dendrogram using the `r func("hclust")` function in the R package `r pack("stats")` [@RCoreTeam2020], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
+We then used the distances to build a trait dendrogram using the `r func("hclust")` function in the R package `r pack("stats")` [@RCoreTeam2021], and used the dendogram in place of a phylogenetic tree to calculate PD and RPD as described above.
 These functional analogs of PD and RPD are hereafter referred to as FD (functional diversity) and RFD (relative functional diversity).
@@
 Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution.
-All measures of biodiversity and CANAPE were carried out using the R package `r pack("canaper")` [@Nitta2021a].
+All measures of biodiversity and CANAPE were carried out using the R package `r pack("canaper")` [@Nitta2022].
 
@@
 Decrease or extirpation of fern populations due to deer herbivory has been observed in multiple sites and species across the country [@Minamitani2005; @Yahara2006; @Hattori2010].
-Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status, and one taxon (*Arachniodes yasu-inouei* Sa.
-Kurata var.
-*angustipinnula* Seriz.) is thought to have gone extinct due to deer herbivory, among other causes [@MOJ2015; @kankyosho2022].
-Although not all species of ferns are equally susceptible to deer herbivory [a small number are apparently unpalatable to deer\; @Minamitani2005], these studies show it is a clear danger to many species.
+Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status [@MOJ2015; @kankyosho2022].
+Although not all species of ferns are equally susceptible to deer herbivory [some are unpalatable to deer\; @Minamitani2005], these studies show it is an existential threat to many species.
 Therefore, analysis of the threat posed to fern biodiversity by deer herbivory is needed to inform future conservation policies.

### 2022-01-21 ms/SI.Rmd
Update MS
Fix dendrogram in SI where non-monophyletic 'other' was causing large portion of tree to be colored grey

@@
   "bioregion_cutoff",
+  "bioregions_all_labels",
   "ms_functions"
@@
 Effect of grain size on sampling redundancy.
-(A), Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km).
-(B), Percent of grid cells with redundancy = 0 by grain size.
-(C), Distribution of redundancy values by grain size.
+(A) Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km).
+(B) Percent of grid cells with redundancy = 0 by grain size.
+(C) Distribution of redundancy values by grain size.
 
@@
 Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan.
-(A), rolling mean area (km^2^) in 20 km latitudinal bands.
-(B), % apomictic taxa.
-(C), annual precipitation.
-(D), percipitation seasonality (coefficient of variation).
-(E), mean annual temperature.
-(F), temperature seasonality (standard deviation × 100).
+(A) rolling mean area (km^2^) in 20 km latitudinal bands.
+(B) % apomictic taxa.
+(C) annual precipitation.
+(D) percipitation seasonality (coefficient of variation).
+(E) mean annual temperature.
+(F) temperature seasonality (standard deviation × 100).
 Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.
@@
 
-**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 20 km grid cell in the ferns of Japan.
+**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 20 km × 20 km grid cell in the ferns of Japan.
 
@@
 **`r s_figure("raw-lat")`**.
-(A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude.
+(A) Raw taxonomic diversity, (B) phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude.
 For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
@@
 **`r s_figure("raw-biplot")`**.
-Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.
-(A), Phylogenetic diversity (PD).
-(B), Relative phylogenetic diversity (RPD).
-(C), Functional diversity (FD).
-(D), Relative functional diversity (RFD).
-Trendlines in (a, c) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
+Relationships between observed phylogenetic and functional diversity and taxonomic richness in the ferns of Japan.
+(A) Phylogenetic diversity (PD).
+(B) Relative phylogenetic diversity (RPD).
+(C) Functional diversity (FD).
+(D) Relative functional diversity (RFD).
+Trendlines in (A, C) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
 
@@
 Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.
-'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
-For details, see Methods.
+'neo' indicates neodendemic areas (those with an overabundance of short branches); 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+For details, see Materials and Methods.
 
@@
 Selection of *k* for bioregions.
-(A), taxonomic bioregions.
-(B), phylogenetic bioregions.
+(A) taxonomic bioregions.
+(B) phylogenetic bioregions.
 Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).
@@
 
-# Map taxonomic clusters to grid cells
+# Map taxonomic clusters to grid cells. Include original ("old") clusters
+# so that grouping of "other" works correctly even if not monophyletic
 clusters_tax <-
-  biodiv_ferns_spatial %>%
-  as_tibble() %>%
-  select(grids, cluster = taxonomic_cluster)
+bioregions_all_labels %>%
+  # Filter to grid-cells used in modeling
+  inner_join(
+    select(biodiv_ferns_spatial, grids),
+    by = "grids"
+  ) %>%
+  add_count(old_taxonomic_cluster) %>%
+  mutate(
+    taxonomic_cluster = case_when(
+      n <= bioregion_cutoff ~ paste(taxonomic_cluster, old_taxonomic_cluster, sep = "_"),
+      TRUE ~ taxonomic_cluster
+      )
+  ) %>%
+  select(-n) %>%
+  select(grids, cluster = taxonomic_cluster) %>%
+  arrange(grids)
 
@@
   # Lump regions with fewer grid-cells than cutoff
-  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff))
-
-# Make sure cutoff matches caption
-assert_that(bioregion_cutoff == 2, msg = "Bioregions cutoff for lumping not 2!")
+  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff + 1))
 
@@
 Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
-(A), taxonomic bioregions.
-(B), phylogenetic bioregions.
+(A) taxonomic bioregions.
+(B) phylogenetic bioregions.
 Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
@@
 
-# Map phylogenetic clusters to grid cells
+# Map phylogenetic clusters to grid cells. Include original ("old") clusters
+# so that grouping of "other" works correctly even if not monophyletic
 clusters_phy <-
-  biodiv_ferns_spatial %>%
-  as_tibble() %>%
-  select(grids, cluster = phylo_cluster)
+bioregions_all_labels %>%
+  # Filter to grid-cells used in modeling
+  inner_join(
+    select(biodiv_ferns_spatial, grids),
+    by = "grids"
+  ) %>%
+  add_count(old_phylo_cluster) %>%
+  mutate(
+    phylo_cluster = case_when(
+      n <= bioregion_cutoff ~ paste(phylo_cluster, old_phylo_cluster, sep = "_"),
+      TRUE ~ phylo_cluster
+      )
+  ) %>%
+  select(-n) %>%
+  select(grids, cluster = phylo_cluster) %>%
+  arrange(grids)
 
@@
   # Lump regions with fewer grid-cells than cutoff
-  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff))
+  mutate(cluster = as.factor(cluster) %>% fct_lump_min(bioregion_cutoff + 1))
 
@@
 Map of species richness for selected genera of ferns of Japan.
-(A), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
-(B), *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
+(A) *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan.
+(B) *Deparia* is an example of a genus with possible recent radiation in southern Honshu.
 (C)--(F) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.

### 2022-01-21 ms/manuscript.Rmd
Update MS
Fix dendrogram in SI where non-monophyletic 'other' was causing large portion of tree to be colored grey

@@
 We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
-Hierarchical clustering was used to analyze bioregions.
+Hierarchical clustering was used to delimit bioregions.
 Conservation status and threats were assessed by comparing the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer.
@@
 Phylogenetic and functional diversity and phylogenetic endemism were highest in small southern islands.
-Relative phylogenetic and functional diversity were high at both high and low latitudes, and low at mid-latitudes.
-Grid-cells were grouped in to `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
-Temperature and apomixis were identified as drivers of taxonomic and phylogenetic diversity.
-Conservation status was generally high for grid-cells with significantly high biodiversity, but the threat due to herbivory was greater for taxonomic richness than other metrics.
-
-**Conclusions:** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan.
+Relative phylogenetic and functional diversity were high at high and low latitudes, and low at mid-latitudes.
+Grid-cells were grouped into `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
+Temperature and apomixis were identified as drivers of biodiversity patterns.
+Conservation status was generally high for grid-cells with significantly high biodiversity, but the threat due to herbivory by deer was greater for taxonomic richness than other metrics.
+
+**Conclusions:** Our integrative approach reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan.
 Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple multiple metrics when establishing conservation priorities.
@@
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b].
-The ferns of Japan include `r total_fern_taxa` native, non-hybrid taxa (including species and varieties) and hundreds of hybrids [@Ebihara2019b].
+The ferns of Japan include `r total_fern_taxa` native, non-hybrid taxa (including species and varieties) and hundreds of hybrids [nothotaxa\; @Ebihara2019b].
 The availability of detailed distribution data [distribution maps at the ca.
-10 km scale for nearly all species\; @Ebihara2016b; @Ebihara2017], trait data [multiple quantitative and qualitative traits compiled for identification of nearly all species\; @Ebihara2019b], and DNA sequences for > 97% of species [@Ebihara2010; @Ebihara2019b\; all coverage statistics exclude hybrids] make the ferns of Japan an ideal system for investigating the relationships between, and drivers of, multiple dimensions of biodiversity.
+10 km scale for nearly all species\; @Ebihara2016b; @Ebihara2017], trait data [multiple quantitative and qualitative traits compiled for identification of nearly all species\; @Ebihara2019b], and DNA sequences for > 97% of species [@Ebihara2010; @Ebihara2019b\; all coverage statistics exclude nothotaxa] make the ferns of Japan an ideal system for investigating the relationships between, and drivers of, multiple dimensions of biodiversity.
 
@@
 We used a list of native, non-hybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
-We chose to use this over other sources (`r eg`, GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
+We chose to use this over other sources such as the Global Biodiversity Information Facility (GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
 Furthermore, the overall sampling completeness of this dataset is also high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article).
 The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety, excluding non-native taxa and nothotaxa.
-For the purposes of analysis, we treated infraspecific taxa as distinct species and hereafter use "taxon" and "species" interchangeably to refer to entities at this rank, unless otherwise indicated.
+We treated infraspecific taxa as distinct species during analysis and hereafter use "taxon" and "species" interchangeably to refer to entities at this rank, unless otherwise indicated.
 All names are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/).
@@
 For specimens that lacked latitude and longitude data, georeferencing was done by mapping collection site names to a set of standard ca.
-10 × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
+10 km × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
 In the case that the collection site could not be mapped to a single second-degree mesh cell, it was excluded.
@@
 4,000 taxa using six plastid markers [@Testo2016a], and the addition of plastome data can be expected to increase support along the backbone.
+
 We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].
@@
 
-One element of biodiversity that we did not consider here is hybrids (nothotaxa).
+One element of biodiversity that we did not consider here is hybrids.
 The Japanese fern flora includes `r n_hybrid_fern_taxa` nothotaxa [@Ebihara2019b].
@@
 This study was supported by JSPS KAKENHI Grant Number 16H06279.
+The authors are grateful to members of the Nippon Fernist Club for their cooperation and efforts towards quantifying and conserving the biodiversity of the ferns of Japan.
 
@@
 Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).
-'paleo' indicates areas with an overabundance of rare long branches; 'neo' indicates areas with an overabundance of rare short branches; 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+'neo' indicates areas with an overabundance of rare short branches; 'paleo' indicates areas with an overabundance of rare long branches; 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
 For details, see Materials and Methods.

### 2022-01-21 _targets.R

@@ -334,6 +334,9 @@ tar_plan(
   bioregion_cutoff = 2, # later, lump bioregions into "other" if they have fewer than this number of grid-cells # nolint
   bioregions = relabel_bioregions_by_lat(
     bioregions_raw, shape_ferns, cutoff = bioregion_cutoff),
+  # - once more, retaining original labels for plotting
+  bioregions_all_labels = relabel_bioregions_by_lat_keep_old(
+    bioregions_raw, shape_ferns, cutoff = bioregion_cutoff),
 
   # Traits analysis ----
 

### 2022-01-21 ms/data_readme.Rmd
Fix grid-cell descriptions

@@
 Museum of Science and Nature, Japan was converted to a community data matrix at
-four grain sizes (grid-cells spanning Japan, each 0.1, 0.2, 0.3, or 0.4 degrees
-per side). The 0.2 degree grain size was selected for further analysis based on 
-redundancy (ratio of number of specimens to number of taxa per cell).
+four grain sizes (square grid-cells spanning Japan, each 10, 20, 30, or 40 km
+per side). The 20 km degree grain size was selected for further analysis based 
+on redundancy (ratio of number of specimens to number of taxa per cell).
 
@@
 downloaded at the 2.5 minute scale. The intersection of the raw climate data
-with 0.2 degree grid-cells (japan_ferns_shape_full.gpkg) were identified,
+with 20 km × 20 km grid-cells (japan_ferns_shape_full.gpkg) were identified,
 then average value for each climatic variable was calculated for each grid-cell.
@@
 `r fs::path_file(japan_ferns_comm_full_file)`: Community matrix (species x sites
-matrix) of native, non-hybrid ferns in Japan. Sites correspond to 0.2 degree x
-0.2 degree grid-cells covering Japan. For location of grid-cells, see
+matrix) of native, non-hybrid ferns in Japan. Sites correspond to 20 km x
+20 km grid-cells covering Japan. For location of grid-cells, see
 japan_ferns_shape_full.gpkg. Full dataset including all grid-cells (none
@@
       ~col, ~desc,
-      "res", "Degree of resolution (grid-cell side length), from 0.1 to 0.4 degrees",
+      "res", "Degree of resolution (grid-cell side length), from 10 to 40 km (in m)",
       "grids", "Name of grid-cell",
@@
 for native, non-hybrid ferns in Japan in GeoPackage format. Compiled from raw
-specimen data by binning specimens into 0.2 degree grid-cells across Japan. Full
-dataset including all grid-cells (none excluded due to low redundancy).
+specimen data by binning specimens into 20 km × 20 km grid-cells across Japan. 
+Full dataset including all grid-cells (none excluded due to low redundancy).
 
@@
 
-Coordinate reference system: JGD2000
+Coordinate reference system: Mollweide (+proj=moll +lon_0=135)
 

### 2022-01-21 ms/data_readme.Rmd
Change "dryad" to "figshare"

@@
     "calibration_dates_file",
-    "japan_ferns_biodiv_dryad_file",
-    "comm_ferns_dryad_file",
-    "shape_ferns_dryad_file",
-    "fern_traits_dryad_file",
-    "japan_fern_phylogram_dryad_file",
-    "japan_fern_tree_dryad_file",
-    "japan_fern_tree_uncollapsed_dryad_file",
+    "japan_ferns_biodiv_figshare_file",
+    "comm_ferns_figshare_file",
+    "shape_ferns_figshare_file",
+    "fern_traits_figshare_file",
+    "japan_fern_phylogram_figshare_file",
+    "japan_fern_tree_figshare_file",
+    "japan_fern_tree_uncollapsed_figshare_file",
     "refs_yaml"
@@
 
-```{r japan-ferns-biodiv-dryad}
-japan_ferns_biodiv_dryad <- read_csv(here::here(japan_ferns_biodiv_dryad_file))
+```{r japan-ferns-biodiv-figshare}
+japan_ferns_biodiv_figshare <- read_csv(here::here(japan_ferns_biodiv_figshare_file))
 ```
 
-`r fs::path_file(japan_ferns_biodiv_dryad_file)` (contained in "results.zip"):
+`r fs::path_file(japan_ferns_biodiv_figshare_file)` (contained in "results.zip"):
 Biodiversity statistics of native, non-hybrid ferns and environmental variables
@@
 
-Number of variables: `r ncol(japan_ferns_biodiv_dryad)`
+Number of variables: `r ncol(japan_ferns_biodiv_figshare)`
 
-Number of cases/rows: `r nrow(japan_ferns_biodiv_dryad)`
+Number of cases/rows: `r nrow(japan_ferns_biodiv_figshare)`
 
@@
 
-```{r japan-ferns-biodiv-dryad-cols, results = "asis"}
-  japan_ferns_biodiv_dryad %>%
+```{r japan-ferns-biodiv-figshare-cols, results = "asis"}
+  japan_ferns_biodiv_figshare %>%
   make_data_desc(
@@
 
-MD5 checksum: `r tools::md5sum(here::here(japan_ferns_biodiv_dryad_file))`
+MD5 checksum: `r tools::md5sum(here::here(japan_ferns_biodiv_figshare_file))`
 
@@
 
-```{r comm-ferns-dryad-file}
-comm_ferns_dryad <- read_csv(here::here(comm_ferns_dryad_file))
+```{r comm-ferns-figshare-file}
+comm_ferns_figshare <- read_csv(here::here(comm_ferns_figshare_file))
 # Check that no data are missing
-assert(comm_ferns_dryad, not_na, everything(), success_fun = success_logical)
+assert(comm_ferns_figshare, not_na, everything(), success_fun = success_logical)
 ```
 
-`r fs::path_file(comm_ferns_dryad_file)` (contained in "results.zip"): Community
+`r fs::path_file(comm_ferns_figshare_file)` (contained in "results.zip"): Community
 matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for
@@
 
-Number of variables: `r ncol(comm_ferns_dryad)`
+Number of variables: `r ncol(comm_ferns_figshare)`
 
-Number of cases/rows: `r nrow(comm_ferns_dryad)`
+Number of cases/rows: `r nrow(comm_ferns_figshare)`
 
@@
 
-MD5 checksum: `r tools::md5sum(here::here(comm_ferns_dryad_file))`
+MD5 checksum: `r tools::md5sum(here::here(comm_ferns_figshare_file))`
 
@@
 
-```{r shape-ferns-dryad-file}
-shape_ferns_dryad <- sf::st_read(here::here(shape_ferns_dryad_file))
+```{r shape-ferns-figshare-file}
+shape_ferns_figshare <- sf::st_read(here::here(shape_ferns_figshare_file))
 # Check that no data are missing
-assert(shape_ferns_dryad, not_na, everything(), success_fun = success_logical)
+assert(shape_ferns_figshare, not_na, everything(), success_fun = success_logical)
 ```
 
-`r fs::path_file(shape_ferns_dryad_file)` (contained in "results.zip"): Location
+`r fs::path_file(shape_ferns_figshare_file)` (contained in "results.zip"): Location
 of grid-cells (sites) for native, non-hybrid ferns in Japan used for
@@
 
-Number of variables: `r ncol(shape_ferns_dryad)`
+Number of variables: `r ncol(shape_ferns_figshare)`
 
-Number of cases/rows: `r nrow(shape_ferns_dryad)`
+Number of cases/rows: `r nrow(shape_ferns_figshare)`
 
@@
 
-```{r shape-ferns-dryad-file-cols, results = "asis"}
-  shape_ferns_dryad %>%
+```{r shape-ferns-figshare-file-cols, results = "asis"}
+  shape_ferns_figshare %>%
   make_data_desc(
@@
 
-MD5 checksum: `r tools::md5sum(here::here(shape_ferns_dryad_file))`
+MD5 checksum: `r tools::md5sum(here::here(shape_ferns_figshare_file))`
 
@@
 
-```{r fern-traits-dryad-file}
-fern_traits_dryad <- read_csv(here::here(fern_traits_dryad_file))
+```{r fern-traits-figshare-file}
+fern_traits_figshare <- read_csv(here::here(fern_traits_figshare_file))
 ```
 
-`r fs::path_file(fern_traits_dryad_file)` (contained in "results.zip"): Trait
+`r fs::path_file(fern_traits_figshare_file)` (contained in "results.zip"): Trait
 matrix of native, non-hybrid ferns in Japan used for functional biodiversity
@@
 
-Number of variables: `r ncol(fern_traits_dryad)`
+Number of variables: `r ncol(fern_traits_figshare)`
 
-Number of cases/rows: `r nrow(fern_traits_dryad)`
+Number of cases/rows: `r nrow(fern_traits_figshare)`
 
@@
 
-MD5 checksum: `r tools::md5sum(here::here(fern_traits_dryad_file))`
+MD5 checksum: `r tools::md5sum(here::here(fern_traits_figshare_file))`
 
@@
 
-```{r japan-fern-phylogram-dryad-file}
-japan_fern_phylogram_dryad <- ape::read.tree(here::here(japan_fern_phylogram_dryad_file))
+```{r japan-fern-phylogram-figshare-file}
+japan_fern_phylogram_figshare <- ape::read.tree(here::here(japan_fern_phylogram_figshare_file))
 ```
 
-`r fs::path_file(japan_fern_phylogram_dryad_file)` (contained in
+`r fs::path_file(japan_fern_phylogram_figshare_file)` (contained in
 "results.zip"): Maximim-likelihood phylogeny of native, non-hybrid ferns in
@@
 
-Number of tips: `r ape::Ntip(japan_fern_phylogram_dryad)`
+Number of tips: `r ape::Ntip(japan_fern_phylogram_figshare)`
 
@@
 
-MD5 checksum: `r tools::md5sum(here::here(japan_fern_phylogram_dryad_file))`
+MD5 checksum: `r tools::md5sum(here::here(japan_fern_phylogram_figshare_file))`
 
@@
 
-```{r japan-fern-dryad-uncollapsed-file}
-japan_fern_tree_uncollapsed_dryad <- ape::read.tree(here::here(japan_fern_tree_uncollapsed_dryad_file))
+```{r japan-fern-figshare-uncollapsed-file}
+japan_fern_tree_uncollapsed_figshare <- ape::read.tree(here::here(japan_fern_tree_uncollapsed_figshare_file))
 ```
 
-`r fs::path_file(japan_fern_tree_uncollapsed_dryad_file)` (contained in "results.zip"):
+`r fs::path_file(japan_fern_tree_uncollapsed_figshare_file)` (contained in "results.zip"):
 Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
@@
 
-Number of tips: `r ape::Ntip(japan_fern_tree_uncollapsed_dryad)`
+Number of tips: `r ape::Ntip(japan_fern_tree_uncollapsed_figshare)`
 
@@
 
-MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_uncollapsed_dryad_file))`
+MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_uncollapsed_figshare_file))`
 
@@
 
-```{r japan-fern-dryad-file}
-japan_fern_tree_dryad <- ape::read.tree(here::here(japan_fern_tree_dryad_file))
+```{r japan-fern-figshare-file}
+japan_fern_tree_figshare <- ape::read.tree(here::here(japan_fern_tree_figshare_file))
 ```
 
-`r fs::path_file(japan_fern_tree_dryad_file)` (contained in "results.zip"):
+`r fs::path_file(japan_fern_tree_figshare_file)` (contained in "results.zip"):
 Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
@@
 
-Number of tips: `r ape::Ntip(japan_fern_tree_dryad)`
+Number of tips: `r ape::Ntip(japan_fern_tree_figshare)`
 
@@
 
-MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_dryad_file))`
+MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_figshare_file))`
 

### 2022-01-21 R/process_raw_data.R

@@ -1,4 +1,4 @@
-# Separate targets plan to process raw data that can't be uploaded to Dryad
+# Separate targets plan to process raw data that can't be uploaded to FigShare
 #
 # This should be run with:
 # targets::tar_make(script = "R/process_raw_data.R", store = "_raw_data_cache") #nolint

### 2022-01-22 ms/data_readme.Rmd
Update READMEs to use gittargets

@@
 Recommended citation for the data: Nitta JH, Mishler BD, Iwasaki W, Ebihara A
-(2021) Data from: Spatial phylogenetics of Japanese ferns: Patterns, processes,
+(2022) Data from: Spatial phylogenetics of Japanese ferns: Patterns, processes,
 and implications for conservation FIXME: add DOI when available
@@
 Citation for and links to publications that cite or use the data: Nitta JH,
-Mishler BD, Iwasaki W, Ebihara A (2021) Spatial phylogenetics of Japanese ferns:
+Mishler BD, Iwasaki W, Ebihara A (2022) Spatial phylogenetics of Japanese ferns:
 Patterns, processes, and implications for conservation FIXME: add journal when published
@@
 
-- _targets.tar.gz: Tarball (compressed folder) including all workflow results produced by R targets package
 - japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
@@
 
-For full methods, see Nitta JH, Mishler BD, Iwasaki W, Ebihara A (2021) Spatial
+For full methods, see Nitta JH, Mishler BD, Iwasaki W, Ebihara A (2022) Spatial
 phylogenetics of Japanese ferns: Patterns, processes, and implications for
@@
 DATA-SPECIFIC INFORMATION
-
-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
-
-_targets.tar.gz: Tarball (compressed folder) including all workflow results
-produced by R targets package. This is provided to enable inspection of workflow
-steps without running the entire workflow from the beginning. To use it, unpack
-the tar achive with the command "tar -xzf _targets.tar.gz". Then, in R, the
-"tar_load()" function in the R package "targets" can be used to load any
-workflow step (target) defined in _targets.R
-(https://github.com/joelnitta/japan_ferns_spatial_phy/blob/main/_targets.R). For
-more information on the structure of the _targets folder and how to use it, see
-https://github.com/ropensci/targets.
-
-MD5 checksum: FIXME (add manually, since this can't be calculated from inside the targets workflow)
-
-Corresponding commit in repo (https://github.com/joelnitta/japan_ferns_spatial_phy): FIXME add manually
 
@@
 Biodiversity statistics of native, non-hybrid ferns and environmental variables
-in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. 
+in Japan. Biodiversity metrics calculated as described in Nitta et al. 2022. 
 Climatic (temperature and preciptation) variables
@@
 environmental data and one outlier for % apomixis that were removed prior to
-spatial modeling analysis in Nitta et al. 2021.
+spatial modeling analysis in Nitta et al. 2022.
 

### 2022-01-22 ms/data_exploration.Rmd
Change dryad to figshare in SI2

@@
 
-All biodiversity metrics were calculated during the `targets` workflow (`_targets.R`), and are contained in the `biodiv_ferns_cent_dryad` object in the `targets` cache.
+All biodiversity metrics were calculated during the `targets` workflow (`_targets.R`), and are contained in the `biodiv_ferns_cent_figshare` object in the `targets` cache.
 
@@
 # Load the data from the targets workflow
-tar_load(biodiv_ferns_cent_dryad)
+tar_load(biodiv_ferns_cent_figshare)
 
@@
 # and read in the data from there
-# biodiv_ferns_cent_dryad <- read_csv(here::here("japan_ferns_biodiv.csv"))
+# biodiv_ferns_cent_figshare <- read_csv(here::here("japan_ferns_biodiv.csv"))
 ```
@@
 biodiv_ferns_cent_raw <-
-  biodiv_ferns_cent_dryad %>%
+  biodiv_ferns_cent_figshare %>%
   select(

### 2022-03-10 ms/manuscript.Rmd
Update MS
fix AJB CSL file to not disambiguate et al authors

@@
 **Conclusions:** Our integrative approach reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan.
-Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple multiple metrics when establishing conservation priorities.
+Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple metrics when establishing conservation priorities.
 
@@
 
-Japan is situated along a latitudinal gradient spanning seasonal, temperate areas in the north to subtropical, mostly aseasonal islands in the south.
+Japan is situated along a latitudinal gradient spanning temperate, seasonal areas in the north to subtropical, mostly aseasonal islands in the south.
 Furthermore, it is a mountainous country with great variation in elevation on the larger islands.
@@
 We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017], which were formatted to be used with Lucid software (https://www.lucidcentral.org/).
-Continuous traits were measured on 10 randomly selected specimens per species, then four values per species were obtained from these measurements following Lucid format: outside (`r ie` outside the typical range) minimum, typical minimum, typical maximum, and outside maximum.
+Continuous traits were measured on 10 randomly selected specimens per species, then four values per species were obtained from these measurements following Lucid format: outside (`r ie`, outside the typical range) minimum, typical minimum, typical maximum, and outside maximum.
 For species with dimorphic fertile (`r ie`, spore-bearing) and sterile fronds, fertile and sterile fronds were measured separately.
@@
 Decrease or extirpation of fern populations due to deer herbivory has been observed in multiple sites and species across the country [@Minamitani2005; @Yahara2006; @Hattori2010].
-Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN or VU) in Japan include herbivory by deer as one of the reasons for their endangered status [@MOJ2015; @kankyosho2022].
+Furthermore, 33 of 212 (15.5%) fern taxa listed as endangered (Red List status CR, EN, or VU) in Japan include herbivory by deer as one of the reasons for their endangered status [@MOJ2015; @kankyosho2022].
 Although not all species of ferns are equally susceptible to deer herbivory [some are unpalatable to deer\; @Minamitani2005], these studies show it is an existential threat to many species.
@@
 
-After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits; Tables `r table_num("traits-used")`, `r s_table_num("phy-sig-binary")`.
+After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits; Tables `r table_num("traits-used")`, `r s_table_num("phy-sig-binary")`).
 The trait matrix included `r n_taxa_traits_sampled` taxa (`r percent_taxa_traits_sampled`) and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.
@@
 Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and high elevation areas of northern Honshu), bioregion 2 (low elevation areas of northern Honshu, southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
+The vast majority of grid-cells fell into a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and high elevation areas of northern Honshu), bioregion 2 (low elevation areas of northern Honshu, southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
 Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 covered most of northern Honshu, while extending further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
@@
 A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a; @Suissa2021].
-Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypothesis positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a].
+Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypotheses positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a].
 Another reason for the greater richness on the main islands compared to southern islands in Japan is that the main islands harbor higher elevational variation that may support a wider range of taxa than is present in the southern islands.
@@
 
-**`r figure("japan-map")`**.
+**`r figure("japan-map") %>% str_to_upper()`**.
 Map of Japan showing names of places mentioned in the text.
@@
 
-**`r figure("raw-diversity")`**.
+**`r figure("raw-diversity") %>% str_to_upper()`**.
 Raw (observed) biodiversity metrics of the ferns of Japan.
-(A) Raw taxonomic diversity (B).
-Raw phylogenetic diversity.
+(A) Raw taxonomic diversity.
+(B) Raw phylogenetic diversity.
 (C) Raw functional diversity.
@@
 
-**`r figure("rand-diversity")`**.
-Results of randomization test for phylogenetic and functional diversity of the ferns of Japan.
+**`r figure("rand-diversity") %>% str_to_upper()`**.
+Results of randomization tests for phylogenetic and functional diversity of the ferns of Japan.
 (A) Phylogenetic diversity (PD).
@@
 
-**`r figure("coeff")`**.
+**`r figure("coeff") %>% str_to_upper()`**.
 Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models for the effect of environment and reproductive mode on biodiversity metrics in the ferns of Japan.
@@
 
-**`r figure("modelfit")`**.
+**`r figure("modelfit") %>% str_to_upper()`**.
 Relationship between biodiversity metrics and selected predictor variables in the ferns of Japan.
@@
 
-**`r figure("endemism")`**.
+**`r figure("endemism") %>% str_to_upper()`**.
 Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).
@@
 
-**`r figure("bioregions")`**.
+**`r figure("bioregions") %>% str_to_upper()`**.
 Bioregions of the ferns of Japan.
@@
 
-**`r figure("div-by-tax-region")`**.
+**`r figure("div-by-tax-region") %>% str_to_upper()`**.
 Phylogenetic and morphological diversity of the ferns of Japan by taxonomic bioregion.
@@
 
-**`r figure("conserv-status")`**.
+**`r figure("conserv-status") %>% str_to_upper()`**.
 Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan.

### 2022-03-10 R/ms_functions.R

@@ -83,8 +83,9 @@ deer_lines <- c(
 
 # Words ----
 # Define formatting for some common custom words that may vary in style between journals
-ie <- "*ie*"
-eg <- "*eg*"
+# AJB: Common Latin words (e.g., in vivo, sensu lato) are not italicized.
+ie <- "i.e."
+eg <- "e.g."
 
 #' Format R packages names
 #' 
@@ -188,9 +189,9 @@ result_file <- function (result_num, extension) {
 
 # - First define the "full" version, which would include a caption
 # (except I never use the caption in the function, and instead replace with 'blank')
-figure_full <- captioner::captioner(prefix = "Fig. ", auto_space = FALSE)
+figure_full <- captioner::captioner(prefix = "Figure ", auto_space = FALSE)
 table_full <- captioner::captioner(prefix = "Table ", auto_space = FALSE)
-s_figure_full <- captioner::captioner(prefix = "Fig. S", auto_space = FALSE)
+s_figure_full <- captioner::captioner(prefix = "Figure S", auto_space = FALSE)
 s_table_full <- captioner::captioner(prefix = "Table S", auto_space = FALSE)
 
 # - Make a short function that prints only the object type and number, e.g., "Fig. 1"

### 2022-03-11 ms/SI.Rmd
Update MS, SI
Add figure for biodiv by phylogenetic bioregion

@@
   - \usepackage{colortbl} # need for kable tables
+  - \usepackage{fancyhdr} # enable header
   - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
-  - \usepackage{fancyhdr} # enable header
   - \pagestyle{fancy}  # enable header
@@
 
-**`r s_table("moran")`**.
+### **`r s_table("moran")`**
+
 Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.
@@
 
-**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position. Significantly correlated variables at *P* < 0.05 in bold.
+### **`r s_table("corr")`**
+
+Results of modified *t*-test for correlation between variables while taking into account spatial position. Significantly correlated variables at *P* < 0.05 in bold.
 
@@
 
-**`r s_table("phy-sig-binary")`**.
+### **`r s_table("phy-sig-binary")`**
+
 Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.
@@
 
-**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan. *K*, Blomberg's *K*; λ, Pagel's λ.
+### **`r s_table("phy-sig")`**
+
+Phylogenetic signal in continuous functional traits of the ferns of Japan. *K*, Blomberg's *K*; λ, Pagel's λ.
 
@@
 
-**`r s_table("lrt-null")`**.
+### **`r s_table("lrt-null")`**
+
 Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).
@@
 
-**`r s_table("aic")`**.
+### **`r s_table("aic")`**
+
 Model fit as measured with conditional Akaike Information Criterion (cAIC).
@@
 
-**`r s_figure("sampling-curve")`**.
+### **`r s_figure("sampling-curve")`**
+
 Species collection curve for the ferns of Japan fit with the `r pack("iNEXT")` package [@Hsieh2016].
@@
 
-**`r s_figure("grain-size")`**.
+### **`r s_figure("grain-size")`**
+
 Effect of grain size on sampling redundancy.
@@
 
-**`r s_figure("envir")`**.
+### **`r s_figure("envir")`**
+
 Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan.
@@
 
-**`r s_figure("tree")`**.
-Time tree of native Japanese ferns and lycophytes, excluding hybrids.
-Points at nodes indicate bootstrap support.
-Scalebar 100 my.
-Tree visualized with the `r pack("ggtree")` R package [@Yu2017].
+### **`r s_figure("tree")`**
+
+Time tree of native Japanese ferns and lycophytes, excluding hybrids. Points at nodes indicate bootstrap support. Scalebar 100 my. Tree visualized with the `r pack("ggtree")` R package [@Yu2017].
 
@@
 
-**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 20 km × 20 km grid cell in the ferns of Japan.
+### **`r s_figure("abundance")`**
+
+Observed number of specimens (A) and sampling redundancy (B) per 20 km × 20 km grid cell in the ferns of Japan.
 
@@
 
-**`r s_figure("raw-lat")`**.
+### **`r s_figure("raw-lat")`**
+
 (A) Raw taxonomic diversity, (B) phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude.
@@
 
-**`r s_figure("raw-biplot")`**.
+### **`r s_figure("raw-biplot")`**
+
 Relationships between observed phylogenetic and functional diversity and taxonomic richness in the ferns of Japan.
@@
 
-**`r s_figure("endemism-restricted")`**.
+### **`r s_figure("endemism-restricted")`**
+
 Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.
@@
 
-**`r s_figure("k-plot")`**.
+### **`r s_figure("k-plot")`**
+
 Selection of *k* for bioregions.
@@
 
-**`r s_figure("dendrogram")`**.
+### **`r s_figure("dendrogram")`**
+
 Bioregion dendrograms.
@@
 **`r s_figure("dendrogram")`**, continued.
+
+\clearpage
+
+```{r div-by-phy-region, fig.width = 7.25, fig.height = 7}
+data <-
+  biodiv_ferns_spatial %>%
+  filter(phylo_cluster %in% c(1, 2, 3)) %>%
+  as_tibble %>%
+  select(
+    pd_obs_z, rpd_obs_z, fd_obs_z, rfd_obs_z,
+    pe_obs_p_upper, phylo_cluster)
+
+a <- ggplot(data, aes(x = phylo_cluster, y = pd_obs_z)) +
+  geom_jitter(aes(color = phylo_cluster)) +
+  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
+  geom_boxplot(
+    color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "SES PD",
+    x = "Cluster",
+    color = "Bioregion"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+
+b <- ggplot(data, aes(x = phylo_cluster, y = rpd_obs_z)) +
+  geom_jitter(aes(color = phylo_cluster)) +
+  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
+  geom_boxplot(
+    color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "SES RPD",
+    x = "Cluster",
+    color = "Bioregion"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+
+c <- ggplot(data, aes(x = phylo_cluster, y = fd_obs_z)) +
+  geom_jitter(aes(color = phylo_cluster)) +
+  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
+  geom_boxplot(
+    color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "SES FD",
+    x = "Cluster",
+    color = "Bioregion"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank()
+  )
+
+d <- ggplot(data, aes(x = phylo_cluster, y = rfd_obs_z)) +
+  geom_jitter(aes(color = phylo_cluster)) +
+  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
+  geom_boxplot(
+    color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "SES RFD",
+    x = "Cluster",
+    color = "Bioregion"
+  ) +
+  theme_bw() +
+  theme(
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    axis.title.x = element_blank()
+  )
+
+e <- ggplot(data, aes(x = phylo_cluster, y = pe_obs_p_upper)) +
+  geom_jitter(aes(color = phylo_cluster)) +
+  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
+  geom_boxplot(
+    color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
+  labs(
+    y = "PE *p*-score",
+    x = "Cluster",
+    color = "Bioregion"
+  ) +
+  theme_bw() +
+  theme(
+    axis.title.y = element_markdown(),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+  )
+
+a + b + c + d + e +
+  plot_layout(ncol = 2) +
+  plot_annotation(
+    tag_levels = "A",
+    tag_prefix = "",
+    tag_suffix = ""
+  ) &
+  theme(
+    legend.position = "none",
+    plot.tag = element_text(face = "bold"))
+```
+
+### **`r s_figure("div-by-phy-region")`**
+
+Phylogenetic and morphological diversity of the ferns of Japan by phylogenetic bioregion.
+(A) Standard effect size (SES) of phylogenetic diversity (PD).
+(B) SES of relative PD (RPD).
+(C) SES of functional diversity (FD).
+(D) SES of relative FD (RFD).
+(E) *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
+Only top three bioregions with the most sites shown.
 
@@
 
-**`r s_figure("genus-map")`**.
+### **`r s_figure("genus-map")`**
+
 Map of species richness for selected genera of ferns of Japan.
@@
 
-**`r s_figure("apo-genus")`**.
-Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon.
-Numbers next to each bar show number of apomictic taxa out of total richness per genus.
+### **`r s_figure("apo-genus")`**
+
+Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon. Numbers next to each bar show number of apomictic taxa out of total richness per genus.
 
@@
 
-**`r s_figure("modelfit-not-ult")`**.
+### **`r s_figure("modelfit-not-ult")`**
+
 Relationship between (A) phylogenetic diversity (PD) and (B) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
@@
 
-**`r s_figure("bioregions-env")`**.
+### **`r s_figure("bioregions-env")`**
+
 Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
@@
 
-**`r s_figure("protected-map")`**.
+### **`r s_figure("protected-map")`**
+
 Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.
@@
 
-**`r s_figure("deer-map")`**.
+### **`r s_figure("deer-map")`**
+
 Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).

### 2022-03-11 ms/manuscript.Rmd
Update MS, SI
Add figure for biodiv by phylogenetic bioregion

@@
 Preliminary analysis indicated that 20 km grid-cells are optimal for our dataset: there is a sudden improvement in redundancy values from 10 km to 20 km, but much less improvement as grain size is increased beyond 20 km (`r s_figure("grain-size")`).
-Although the grid-cells are defined with equal area, actual land area of each cell varies due to coastlines.
+Although the grid-cells are defined with equal area, actual land area of each grid-cell varies due to coastlines.
 Strictly filtering out all grid-cells with less than complete land area would result in a large loss of data, as Japan has many coastlines and small islands.
-Therefore, we instead filtered the grid-cells by sampling completeness, removing any cells with redundancy < 0.1; this dataset was used for all subsequent analyses.
+Therefore, we instead filtered the grid-cells by sampling completeness, removing any grid-cells with redundancy < 0.1; this dataset was used for all subsequent analyses.
 The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].
@@
 Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree.
-Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
+Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (FTOL v0.0.1; Nitta et al., in prep.; https://github.com/fernphy/ftol/) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
 This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across ca.
@@
 
-Two grid cells lacking environmental data and an outlier with 100% apomictic taxa due to extremely low species richness were excluded from modeling analyses (Appendix S2; see Supplemental Data with this article).
+Two grid-cells lacking environmental data, one grid-cell occupied by a single species lacking reproductive mode data, and an outlier with 100% apomictic taxa due to extremely low species richness were excluded from modeling analyses (Appendix S2; see Supplemental Data with this article).
 
@@
 For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above.
-For taxonomic richness, cells ranked in the top > 5% were considered highly diverse.
+For taxonomic richness, grid-cells ranked in the top > 5% were considered highly diverse.
 
@@
 Grid-cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`A).
-A smaller number of cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (Kii Peninsula).
+A smaller number of grid-cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (`r eg`, Kii Peninsula).
 Grid-cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`B).
@@
 Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands.
-Small clusters of cells containing significant endemism were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`).
-When we ran CANAPE with a reduced dataset including only taxa absolutely restricted to Japan to account for the border effect, there were fewer significant cells, but a similar pattern of significant endemism mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).
+Small clusters of grid-cells containing significant endemism were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`).
+When we ran CANAPE with a reduced dataset including only taxa absolutely restricted to Japan to account for the border effect, there were fewer significant grid-cells, but a similar pattern of significant endemism mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).
 
@@
 Bioregion 2 has lower RPD and RFD than bioregion 1.
-Comparison of biodiversity metrics across phylogenetic bioregions was similar, except that phylogenetic bioregion 3 mostly corresponds to taxonomic bioregions 3 and 4 combined.
+Comparison of biodiversity metrics across phylogenetic bioregions was similar, except that phylogenetic bioregion 3 mostly corresponds to taxonomic bioregions 3 and 4 combined (`r s_figure("div-by-phy-region")`).
 
@@
 Cells with significantly high biodiversity had a similar or greater percentage of area protected compared to the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
-In contrast, cells with significantly high biodiversity had a lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
-Only cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.
+In contrast, grid-cells with significantly high biodiversity had a similar or lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
+Only grid-cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.
 
@@
 The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
-Cells with high FD, PD, or PE tended to have less deer occupation compared to the baseline rate (except for PE relative to 1978 levels; `r figure("conserv-status")`B).
-However, cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.
+Grid-cells with high PD, FD, or PE tended to have less deer occupation compared to the baseline rate (except for PE relative to 1978 levels; `r figure("conserv-status")`B).
+However, grid-cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.
 
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here (Figs. `r figure_num("raw-diversity")`a, `r s_figure_num("raw-lat")`a) is well-documented [@Kusumoto2017; @Ebihara2019b].
+The latitudinal hump-shaped pattern of species richness observed in ferns here (Figures `r figure_num("raw-diversity")`A, `r s_figure_num("raw-lat")`A) is well-documented [@Kusumoto2017; @Ebihara2019b].
 Similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020].
@@
 The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig-binary")`, `r s_table_num("phy-sig")`), and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
-Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of cells with significant levels of PE, as well as paleoendemic cells, are located in the southern islands (Figs.
-`r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`).
+Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid-cells with significant levels of PE, as well as paleoendemic grid-cells, are located in the southern islands (Figures `r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`).
 This is clearly due to the small land area of these islands: small area combined with high PD is expected to lead to high PE.
@@
 
-Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (Figs.
-`r figure_num("coeff")`G, `r figure_num("modelfit")`G), and to a slightly lesser extent SES of PD (Figs.
-`r figure_num("coeff")`F, `r figure_num("modelfit")`F).
+Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (Figures `r figure_num("coeff")`G, `r figure_num("modelfit")`G), and to a slightly lesser extent SES of PD (Figures `r figure_num("coeff")`F, `r figure_num("modelfit")`F).
 Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (`r s_figure("genus-map")`).
@@
 
-The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group cells with similar biodiversity values (`r figure("div-by-tax-region")`).
+The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group grid-cells with similar biodiversity values (`r figure("div-by-tax-region")`).
 Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
@@
 Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate in recent surveys or modeled data, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`B).
-This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("protected-map")`A).
+This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("deer-map")`A).
 This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.
@@
 
-**Appendix S1**. Supplementary tables and figures, including Tables `r s_table_num("moran")`--`r s_table_num("aic")` and Figs. `r s_figure_num("sampling-curve")`--`r s_figure_num("deer-map")`.
+**Appendix S1**. Supplementary tables and figures, including Tables `r s_table_num("moran")`--`r s_table_num("aic")` and Figures `r s_figure_num("sampling-curve")`--`r s_figure_num("deer-map")`.
 
@@
 **`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.  
-**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km grid cells across Japan.  
+**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km grid-cells across Japan.  
 **`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
-**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20 km grid cell in the ferns of Japan.  
+**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20 km grid-cell in the ferns of Japan.  
 **`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
@@
 **`r s_figure("dendrogram")`**. Bioregion dendrograms.  
+**`r s_figure("div-by-phy-region")`**. Phylogenetic and morphological diversity of the ferns of Japan by phylogenetic bioregion.    
 **`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
@@
 Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
-**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
+**`r s_figure("bioregions-env")`**. Scatterplots of grid-cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
 **`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.  
@@
   result_file(figure("raw-diversity"), "pdf"),
-  width = 7.25, height = 4, units = "in", raw_diversity_plot)
-```
-
-```{r raw-diversity-print, fig.width = 7.25, fig.height = 4, eval = params$doc_type == "pdf"}
+  width = 7.25, height = 3.4, units = "in", raw_diversity_plot)
+```
+
+```{r raw-diversity-print, fig.width = 7.25, fig.height = 3.4, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
   result_file(figure("bioregions"), "pdf"),
-  bioregions_plot, width = 7.25, height = 5, units = "in")
-```
-
-```{r bioregions-print, fig.width = 7.25, fig.height = 5, eval = params$doc_type == "pdf"}
+  bioregions_plot, width = 7.25, height = 4.3, units = "in")
+```
+
+```{r bioregions-print, fig.width = 7.25, fig.height = 4.3, eval = params$doc_type == "pdf"}
 # Print plot if rendering to PDF
@@
 (E) *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
-Only top four clusters with the most sites shown.
 

### 2022-03-11 R/ms_functions.R

@@ -201,9 +201,9 @@ s_figure <- pryr::partial(s_figure_full, display = "cite", caption = "blank")
 s_table <- pryr::partial(s_table_full, display = "cite", caption = "blank")
 
 # - Make a short function that prints only the number (e.g., "1")
-figure_num <- function (name) {figure(name) %>% str_remove("Fig. ")}
+figure_num <- function (name) {figure(name) %>% str_remove("Figure ")}
 table_num <- function (name) {table(name) %>% str_remove("Table ")}
-s_figure_num <- function (name) {s_figure(name) %>% str_remove("Fig. ")}
+s_figure_num <- function (name) {s_figure(name) %>% str_remove("Figure ")}
 s_table_num <- function (name) {s_table(name) %>% str_remove("Table ")}
 
 # Etc ----

### 2022-03-11 ms/data_readme.Rmd
Update data readme
- automatically generate txt
- automatically zip figshare files

@@
 ---
-output: rtf_document
+output: md_document
 editor_options:
@@
 bibliography: 
-- `r here::here("ms/references.yaml")`
+- `r here::here("ms/data_readme_refs.yaml")`
 ---
@@
 ```{r setup, include = FALSE, cache = FALSE}
-# TO CONVERT TO PLAIN TEXT:
-# after rendering to rtf,
-# 1. Copy the contents of the rtf file, paste to txt.
-# 2. Replace all single line breaks with double line breaks
-# (in Atom search-replace with grep: \n for \n\n)
-# 3. Hard-wrap all text at 80 characters (in Atom with autoflow: cmd-alt-q)
-# (optional: add name after first line)
-# 4. Save as "README.txt"
-
 knitr::opts_chunk$set(
@@
     "japan_fern_tree_uncollapsed_figshare_file",
-    "refs_yaml"
+    "refs_data_readme_yaml"
     )
@@
 
-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\- GENERAL INFORMATION \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+------------------- GENERAL INFORMATION -----------------
 
@@
 
-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+--------------------------
 
@@
 Mishler BD, Iwasaki W, Ebihara A (2022) Spatial phylogenetics of Japanese ferns:
-Patterns, processes, and implications for conservation FIXME: add journal when published
+Patterns, processes, and implications for conservation.
+American Journal of Botany. FIXME: add DOI, page numbers when available
 
@@
 
-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+--------------------
 
@@
 Phylogenetic analysis was conducted with maximum likelihood in IQ-TREE v1.6.12
-[@Nguyen2015] by combining plastid *rbcL* sequences of each taxon with a
-globally sampled data matrix (Nitta et al, in prep). Next, dating analysis was
+[@Nguyen2015] by combining plastid rbcL sequences of each taxon with a
+globally sampled data matrix (FTOL v0.0.1; Nitta et al., in prep.; 
+https://github.com/fernphy/ftol/). Next, dating analysis was
 carried out using treePL v1.0 [@Smith2012] with `r n_fossil_points` fossil
@@
 phylogenetics of Japanese ferns: Patterns, processes, and implications for
-conservation FIXME: Add journal when published
+conservation. American Journal of Botany. 
+FIXME: add DOI, page numbers when available
 
-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+--------------------------
 
@@
 
-```{r results = "asis"}
+```{r deer-range-cols, results = "asis"}
   deer_range %>%
@@
       "precip", "Annual precipitation (mm) (BIO12)",
-      "precip_season", "Precipitation seasonlaity (coefficient of variation) (BIO15)",
+      "precip_season", "Precipitation seasonality (coefficient of variation) (BIO15)",
       "geom", "Vector describing shape and position of grid-cell"
@@
       "m", "Sample size (number of individuals)",
-      "method", "Method used to estimate diveristy",
+      "method", "Method used to estimate diversity",
       "order", "Diversity order of Hill number (here 0, equivalent to species richness)",
@@
 specimen, with species name, latitude, longitude, and collection date (raw
-specimen data not included here to protect endangered species). Raw data were
+specimen data not included here). Raw data were
 filtered to remove duplicate specimens and specimens that occurred outside of
@@
 may normally be in the range of 4-5 cm, but sometimes as low as 2 cm or as high
-as 6 cm. This would be written 2:4:5:6.
+as 6 cm. This would be written 1:2:4:5:6.
 
@@
 Climatic (temperature and preciptation) variables
-calculated as described for japan_climate.gpkg. Includes one row with missing
-environmental data and one outlier for % apomixis that were removed prior to
-spatial modeling analysis in Nitta et al. 2022.
+calculated as described for japan_climate.gpkg. Includes three rows with missing 
+data and one outlier for % apomixis that were removed prior to
+spatial modeling analysis.
 
@@
       "pd_obs_z", "Z-score of observed phylogenetic diversity (PD)",
-      "pe_obs_z", "Z-score of observed phylogenetic endemisim (PE)",
+      "pe_obs_z", "Z-score of observed phylogenetic endemism (PE)",
       "rfd_obs_z", "Z-score of observed relative FD",
@@
 
-\-\-\-
+2022-03-11
+
+- Use Mollweide projection for grid-cells
+- Don't include targets cache (move to https://github.com/joelnitta/japan_ferns_spatial_phy_cache)
+- Fix spelling errors in README text
+- Fix missing '1:' in LUCID trait data example
+- Fix description of rows with missing data in japan_ferns_biodiv.csv
+- Automate generation of plain text file
 
@@
 
-Generate this README file.
+- Generate this README file.
 
-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-
+--------------------------
 
-References
+REFERENCES

### 2022-03-11 R/process_refs.R

@@ -145,6 +145,7 @@ refs_combined[[1]] <- refs
 # - main MS
 filter_refs_yaml(c("ms/manuscript.Rmd", "ms/SI.Rmd"), refs_combined, "ms/references.yaml")
 # - data README (remember to remove `<span class="nocase">` manually)
+# also for some reason Kusumoto2017 isn't included, so copy over manually
 filter_refs_yaml("ms/data_readme.Rmd", refs_combined, "ms/data_readme_refs.yaml")
 
 # Write out 'aux' file with citation keys to make collection in Zotero

### 2022-03-11 ms/data_exploration.Rmd
Update data exploration (SI2)

@@
 * `precip_season`: seasonality in precipitation
-* `lat_area`: rolling mean of area in 0.2 degree latitudinal bands
+* `lat_area`: rolling mean of area in 100 km latitudinal bands
 
@@
 
-Grubb's test confirms that the percent apomictic value is an outlier. It also indicates that the highest values of the variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, and `rpd_obs_z` may be outliers.
+Grubb's test confirms that the percent apomictic value is an outlier. It also indicates that the highest values of the variables `pd_obs_z`, `fd_obs_z`, `rpd_obs_z`, and `rfd_obs_z` may be outliers.
 

### 2022-04-18 ms/SI.Rmd
Update MS
according to editor's comments

@@
 
-Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted data set including only taxa endemic to Japan.
 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
@@
 Points shown with partial transparency; darker areas indicate overlapping points.
-Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
+Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid cells each are lumped into the "Other" category.
 
@@
 
-Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.
-Grid-cells with black outlines indicate significantly high biodiversity.
+Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and protected areas.
+Grid cells with black outlines indicate significantly high biodiversity.
 Biodiversity metrics include (A) taxon richness, (B) phylogenetic diversity (PD), (C) functional diversity (FD), and (D) phylogenetic endemism (PE).
-Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
+Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid cells in the top 5% considered highly diverse (see Materials and Methods).
 
@@
 
-Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).
-Grid-cells with black outlines indicate significantly high biodiversity.
+Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).
+Grid cells with black outlines indicate significantly high biodiversity.
 Biodiversity metrics include (A) taxon richness, (B) phylogenetic diversity (PD), (C) functional diversity (FD), and (D) phylogenetic endemism (PE).
-Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
+Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid cells in the top 5% considered highly diverse (see Materials and Methods).
 Deer range data sources include 1978 survey, 2003 survey, and range estimated from model based on 2003 survey data.

### 2022-04-18 ms/manuscript.Rmd
Update MS
according to editor's comments

@@
 ---
-title: "Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation"
 mainfont: Roboto-Regular.ttf
@@
 
-Joel H. Nitta^1,5^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^
-
-^1^Department of Biological Sciences, Graduate School of Science, The University of Tokyo, Tokyo, Japan
-
-^2^University and Jepson Herbaria, and Department of Integrative Biology, University of California, Berkeley, CA, USA
-
-^3^Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan
-
-^4^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
-
-^5^Author for correspondence: joelnitta@gmail.com
+RESEARCH ARTICLE
+
+Short Title: Spatial phylogenetics of Japanese ferns
+
+__Spatial phylogenetics of Japanese ferns: patterns, processes, and implications for conservation__
+
+Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1^, and Atsushi Ebihara^3^
 
@@
 cat('
-Manuscript received _______; revision accepted _______.
-
-Short running title: Spatial phylogenetics of Japanese ferns
+Manuscript received 22 September 2021; revision accepted 11 March 2022
+')
+```
+
+^1^Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan
+
+^2^University and Jepson Herbaria, and Department of Integrative Biology, University of California, Berkeley, CA, USA
+
+^4^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
+
+__Correspondence__
+
+Joel H. Nitta, #601 Iwasaki Lab, Bioscience Bldg, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-0882, Japan
+
+Email: joelnitta@gmail.com
+
+ORCID ID: 0000-0003-4719-7472
+
+```{r author-page-doc-2, results = "asis", echo = FALSE, eval = params$doc_type == "doc"}
+# Author page, version for submission to journal
+cat('
+__Citation__: Nitta, J.
+H., B.
+D.
+Mishler, W.
+Iwasaki, and A.
+Ebihara.
+2022.
+Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation.
+*American Journal of Botany* 109(5): XXXX
+
+__DOI: XXXX__
 ')
@@
 
-**Premise:** Biodiversity is often only measured with species richness.
-However, this ignores evolutionary history and is not sufficient for making conservation decisions.
+**Premise:** Biodiversity is often only measured with species richness; however, this measure ignores evolutionary history and is not sufficient for making conservation decisions.
 Here, we characterize multiple facets and drivers of biodiversity to understand how these relate to bioregions and conservation status in the ferns of Japan.
 
-**Methods:** We compiled a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 20 km × 20 km grid-cells including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
-We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
+**Methods:** We compiled a community data set of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid cells (20 km × 20 km each) including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
+We combined the community data with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity and modeled biodiversity metrics in response to environmental factors and reproductive mode.
 Hierarchical clustering was used to delimit bioregions.
-Conservation status and threats were assessed by comparing the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer.
-
-**Results:** Taxonomic richness was highest at mid-latitudes.
+Conservation status and threats were assessed by comparing the overlap of significantly diverse grid cells with conservation zones and range maps of native Japanese deer.
+
+**Results:** Taxonomic richness was highest at mid latitudes.
 Phylogenetic and functional diversity and phylogenetic endemism were highest in small southern islands.
-Relative phylogenetic and functional diversity were high at high and low latitudes, and low at mid-latitudes.
-Grid-cells were grouped into `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
+Relative phylogenetic and functional diversity were high at high and low latitudes, and low at mid latitudes.
+Grid cells were grouped into `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
 Temperature and apomixis were identified as drivers of biodiversity patterns.
-Conservation status was generally high for grid-cells with significantly high biodiversity, but the threat due to herbivory by deer was greater for taxonomic richness than other metrics.
+Conservation status was generally high for grid cells with significantly high biodiversity, but the threat due to herbivory by deer was greater for taxonomic richness than other metrics.
 
@@
 
-Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it, and to conserve it.
+Characterizing the spatial distribution of biodiversity is a major goal of evolutionary biology with two equally important aspects: to understand the processes generating it and to conserve it.
 Until recently, the vast majority of studies seeking to understand the spatial distribution of biodiversity have focused on species richness.
@@
 
-Presumably, the main reason phylogeny has not been taken into account more prominently in biodiversity studies is because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently [@Soltis2016a; @Folk2021].
-These datasets and tools now make it possible to analyze other dimensions of biodiversity, such as phylogenetic diversity [@Faith1992] and phylogenetic endemism [@Rosauer2009].
-Better ways to characterize biological regions ("bioregions"; `r ie`, areas defined by their taxonomic composition or evolutionary history) are also available [@Laffan2016; @White2019; @Daru2020a], rather than relying on ad-hoc characterizations.
+Presumably, phylogeny has not been taken into account more prominently in biodiversity studies mainly because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently [@Soltis2016a; @Folk2021].
+These datasets and tools now make it possible to analyze other dimensions of biodiversity such as phylogenetic diversity [@Faith1992] and phylogenetic endemism [@Rosauer2009].
+Better ways to characterize biological regions ("bioregions"; `r ie`, areas defined by their taxonomic composition or evolutionary history) are also available [@Laffan2016; @White2019; @Daru2020a], rather than relying on ad hoc characterizations.
 Furthermore, incorporating these two frameworks---the categorization of areas into phyloregions with analysis of over/under dispersion of biodiversity---can provide powerful insights into the processes structuring biodiversity and suggest conservation priorities.
 
-However, a comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, and such datasets are rare on the regional (country) scale.
-The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed in @Ebihara2019b].
-The ferns of Japan include `r total_fern_taxa` native, non-hybrid taxa (including species and varieties) and hundreds of hybrids [nothotaxa\; @Ebihara2019b].
+A comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, but such datasets are rare on the regional (country) scale.
+The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed by @Ebihara2019b].
+The ferns of Japan include `r total_fern_taxa` native, nonhybrid taxa (including species and varieties) and hundreds of hybrids [nothotaxa\; @Ebihara2019b].
 The availability of detailed distribution data [distribution maps at the ca.
@@
 One particularly valuable characteristic of the Japanese fern flora is the availability of data on reproductive mode, which is known for `r repro_count$known$total` native fern taxa excluding hybrids [`r repro_count$known$percent`\; @Ebihara2019b].
-Reproductive mode is likely to affect population-level genetic diversity [@Bengtsson2003], and thereby higher-level biodiversity [@Krueger-Hadfield2019].
+Reproductive mode is likely to affect population-level genetic diversity [@Bengtsson2003] and thereby higher-level biodiversity [@Krueger-Hadfield2019].
 In ferns, clonally reproducing apomictic species are often polyploid hybrids that share identical plastid genotypes among taxa and within populations [@Grusz2009; @Chao2012a; @Hori2014; @Hori2019].
@@
 
-Here we leverage this exceptionally rich dataset to analyze the geographic distribution of biodiversity in detail.
-We ask the following questions in our study system, the ferns of Japan: 1.
+Here, we leveraged this exceptionally rich data set to analyze the geographic distribution of biodiversity in detail.
+We asked the following questions in our study system, the ferns of Japan: 1.
 How is biodiversity distributed?
@@
 
-We used a list of native, non-hybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
+We used a list of native, nonhybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
 We chose to use this over other sources such as the Global Biodiversity Information Facility (GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
@@
 We treated infraspecific taxa as distinct species during analysis and hereafter use "taxon" and "species" interchangeably to refer to entities at this rank, unless otherwise indicated.
-All names are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/).
+All names are standardized to a common taxonomy, the Japan Green List v. 1.01 [@Ebihara2019c].
 
@@
 
-We further cleaned the list prior to analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
+We further cleaned the list before analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
 Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, CoordinateCleaner\; @Zizka2019] have the potential to erroneously exclude true occurrence points [`r ie`, false positives\; @Zizka2020], we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].
@@
 Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling.
-We created sets of 10 km, 20 km, 30 km, and 40 km grid-cells covering Japan using a Mollweide equal-area projection (size refers to the length of the side of each square grid-cell).
-At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell).
-We calculated sampling redundancy [1 - richness/number of specimens\; @Garcillan2003] to quantify adequacy of sampling.
-Preliminary analysis indicated that 20 km grid-cells are optimal for our dataset: there is a sudden improvement in redundancy values from 10 km to 20 km, but much less improvement as grain size is increased beyond 20 km (`r s_figure("grain-size")`).
-Although the grid-cells are defined with equal area, actual land area of each grid-cell varies due to coastlines.
-Strictly filtering out all grid-cells with less than complete land area would result in a large loss of data, as Japan has many coastlines and small islands.
-Therefore, we instead filtered the grid-cells by sampling completeness, removing any grid-cells with redundancy < 0.1; this dataset was used for all subsequent analyses.
+We created sets of 10-km, 20-km, 30-km, and 40-km grid cells covering Japan using a Mollweide equal-area projection (size refers to the length of the side of each square grid cell).
+At each grain size, species occurrences were converted to a presence–absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell).
+We calculated sampling redundancy [1 - Richness/Number of specimens\; @Garcillan2003] to quantify adequacy of sampling.
+Preliminary analysis indicated that 20-km grid cells are optimal for our data set; redundancy values greatly increased from 10 to 20 km, but much less improvement as grain size is increased beyond 20 km (`r s_figure("grain-size")`).
+Although the grid cells are defined with equal area, actual land area of each grid cell varies due to coastlines.
+Strictly filtering out all grid cells with less than complete land area would result in a large loss of data, as Japan has many coastlines and small islands.
+Therefore, we instead filtered the grid cells by sampling completeness, removing any grid cells with redundancy < 0.1; this dataset was used for all subsequent analyses.
 The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].
@@
 We used traits originally compiled for identification of ferns and lycophytes of Japan [@Ebihara2016b; @Ebihara2017], which were formatted to be used with Lucid software (https://www.lucidcentral.org/).
-Continuous traits were measured on 10 randomly selected specimens per species, then four values per species were obtained from these measurements following Lucid format: outside (`r ie`, outside the typical range) minimum, typical minimum, typical maximum, and outside maximum.
+Continuous traits were measured on 10 randomly selected specimens per species, then four values per species were obtained from these measurements following the Lucid format: outside (`r ie`, outside the typical range) minimum, typical minimum, typical maximum, and outside maximum.
 For species with dimorphic fertile (`r ie`, spore-bearing) and sterile fronds, fertile and sterile fronds were measured separately.
@@
 We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown.
-We calculated the percentage of apomictic taxa in each grid-cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).
+We calculated the percentage of apomictic taxa in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).
 
@@
 We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`).
-We identified the intersection of the climate data at the 2.5 minute scale with the 20 km grid-cells, then calculated the mean of each climatic variable for each grid-cell.
-We calculated area as rolling mean area (km^2^) across 100 km latitudinal bands.
-This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).
+We identified the intersection of the climate data at the 2.5-minute scale with the 20-km grid cells, then calculated the mean of each climatic variable for each grid cell.
+We calculated area as rolling mean area (km^2^) across 100-km latitudinal bands.
+This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context because there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).
 
@@
 Furthermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
-Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree.
-Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (FTOL v0.0.1; Nitta et al., in prep.; https://github.com/fernphy/ftol/) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa).
+Because one goal of the study is to understand the distribution of paleo- vs. neoendemic areas (defined in units of time), we require an ultrametric tree.
+Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes [FTOL v0.0.1\; @Nitta2022b] (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` operational taxonomic units [OTUs] total, including outgroup taxa).
 This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across ca.
@@
 We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].
-Molecular dating was conducted with `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
+Molecular dating was conducted with the program `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
 The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
 We trimmed the resulting ultrametric tree to only Japanese taxa.
-`r software("treePL")` requires branch lengths to be > 0, and it sets extremely small branches to an arbitrary minimum positive length.
-This may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
-We converted any clades in the dated tree consisting of identical sequences to polytomies with the node at 0 mya (`r ie`, the present).
+`r software("treePL")` requires branch lengths to be >0, and it sets extremely small branches to an arbitrary minimum positive length.
+This behavior may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
+We converted any clades in the dated tree consisting of identical sequences to polytomies with the node at 0 million years ago (i.e., the present; Ma).
 This tree was used for all subsequent analysis unless indicated otherwise.
@@
 
-We tested for phylogenetic signal in continuous traits using two alternative metrics, Blomberg's *K* [@Blomberg2003] and Pagel's lambda [@Pagel1999].
+We tested for phylogenetic signal in continuous traits using two alternative metrics, Blomberg's *K* [@Blomberg2003] and Pagel's $\lambda$ [@Pagel1999].
 *K* describes the ratio between the observed variance in a trait vs.
-the amount of variance expected under Brownian Motion (BM); lambda is a scaling parameter between 0 and 1 that transforms the phylogeny such that trait values fit those most closely expected under BM.
-Both lambda and *K* = 1 under BM; relatively lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering).
-We measured *K* and lambda using the `r func("phylosig")` function in the R package `r pack("phytools")` [@Revell2012], which also calculates significance based on a randomization test for *K* and a likelihood ratio test for lambda.
+the amount of variance expected under brownian motion (BM); $\lambda$ is a scaling parameter between 0 and 1 that transforms the phylogeny such that trait values fit those most closely expected under BM.
+Both $\lambda$ and *K* = 1 under BM; relatively lower values indicate less signal (overdispersion) and higher values indicate more signal (clustering).
+We measured *K* and $\lambda$ using the `r func("phylosig")` function in the R package `r pack("phytools")` [@Revell2012], which also calculates significance based on a randomization test for *K* and a likelihood ratio test for $\lambda$.
 We tested for phylogenetic signal in qualitative (coded as binary) traits using Fritz and Purvis' *D* [@Fritz2010] with the `r func("phylo.d")` function in the R package `r pack("caper")` [@Orme2018], which also assesses significance by comparing the observed value to a null distribution of random values and a simulated distribution of values expected under BM.
@@
 
-We measured taxonomic diversity using taxonomic richness (number of taxa in each community).
+We measured taxonomic diversity using taxonomic richness (`r ie`, number of taxa in each community).
 
@@
 
-We measured functional diversity (FD) following the method of @Petchey2002.
+We measured functional diversity (FD) using the method of @Petchey2002.
 We first calculated functional distances using Gower's method [@Gower1971] on the trait data with the `r func("gowdis")` function in the R package `r pack("FD")` [@Laliberte2010].
@@
 
-Observed values of PD, RPD, FD, RFD, and PE are known to be related to species richness: as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others.
+Observed values of PD, RPD, FD, RFD, and PE are known to be related to species richness; as more taxa are drawn at random without replacement from a given set, it becomes increasingly unlikely to draw a taxon that is distantly related (or functionally dissimilar) to the others.
 Therefore, we determined whether the observed metrics are more or less diverse than random for a given number of taxa by conducting a randomization test.
-We generated 999 random communities using the "curveball" algorithm of @Strona2014, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values.
-Statistical significance was assessed using a one-tailed test for PE or a two-tailed test for other metrics with alpha = 0.5; observed values in the extreme 5% of the null distribution were considered significant.
+We generated 999 random communities using the curveball algorithm of @Strona2014, which conserves richness per site and species occurrence frequencies while randomizing species identities, then compared the observed value to the distribution of random values.
+Statistical significance was assessed using a one-tailed test for PE or a two-tailed test for other metrics with $\alpha$ = 0.5; observed values in the extreme 5% of the null distribution were considered significant.
 We also calculated the standard effect size (SES) of each biodiversity metric.
-SES measures how extreme the observed value (*obs*) is relative to the null distribution (*null*):
-
-$SES = \frac{obs - mean(null)}{sd(null)}$
-
-We tested whether centers of endemism have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleo-endemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
+SES measures how extreme the observed value (Obs) is relative to the null distribution (Null):
+
+$SES = \frac{Obs - Mean(null)}{SD(Null)}$
+
+We tested whether centers of endemism have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleoendemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
 Low RPE indicates short range-restricted branches (neoendemism) and high RPE indicates long range-restricted branches (paleoendemism).
@@
 It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (a "border effect").
-This is especially expected for southern, sub-tropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country.
-To partially account for this, we also calculated all endemism-related metrics using a dataset including species found only in Japan.
+This is especially expected for southern, subtropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country.
+To partially account for the border effect, we also calculated all endemism-related metrics using a data set including species found only in Japan.
 
@@
 
-Two grid-cells lacking environmental data, one grid-cell occupied by a single species lacking reproductive mode data, and an outlier with 100% apomictic taxa due to extremely low species richness were excluded from modeling analyses (Appendix S2; see Supplemental Data with this article).
-
-We tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) using Moran's I with the `r func("moran.mc")` function in the R package `r pack("spdep")` [@Bivand2013].
+Two grid cells lacking environmental data, one grid cell occupied by a single species lacking reproductive mode data, and an outlier with 100% apomictic taxa due to extremely low species richness were excluded from modeling analyses (Appendix S2; see Supplemental Data with this article).
+
+We tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and percentage apomictic taxa [hereafter, % apomictic taxa]) using Moran's I with the `r func("moran.mc")` function in the R package `r pack("spdep")` [@Bivand2013].
 In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05) (`r s_table("moran")`).
@@
 We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the R package `r pack("SpatialPack")` [@SpatialPack2020].
-This indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature (`r s_table("corr")`), so for the main modeling analysis we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
+This test indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature (`r s_table("corr")`), so for the main modeling analysis we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
 
@@
 Richness was fit with the negative binomial response family; others (SES of PD, SES of FD, SES of RPD, SES of RFD) were Gaussian.
-We fit models including a Matérn correlation matrix based on the geographic centroids of each grid-cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014].
+We fit models including a Matérn correlation matrix based on the geographic centroids of each grid cell as a random effect with the R package `r pack("spaMM")` [@Rousset2014].
 We verified the significance of each term of the model by conducting likelihood ratio tests (LRTs) between the full model and models each with a single term of interest removed, in sequence.
@@
 These models included biodiversity metrics with a phylogenetic component (SES of PD and SES of RPD) as dependent on % apomictic taxa, precipitation, and precipitation seasonality, with random spatial effects as described above.
-We compared the fit between environmental models and reproductive models for SES of PD and SES of RPD using conditional Akaike Information Criterion (cAIC) [@Courtiol2017; @Vaida2005].
+We compared the fit between environmental models and reproductive models for SES of PD and SES of RPD using conditional Akaike information criterion (cAIC) [@Courtiol2017; @Vaida2005].
 
@@
 We analyzed bioregions using the framework of @Daru2020 and their R package `r pack("phyloregion")`.
-Briefly, this involves first calculating beta diversity (change in composition between communities) using either species names (taxonomic bioregions) or phylogenetic distances (phylogenetic bioregions).
+Briefly, the analysis involves first calculating beta diversity (change in composition between communities) using either species names (taxonomic bioregions) or phylogenetic distances (phylogenetic bioregions).
 For taxonomic bioregions, we calculated the turnover in species between sites due to species replacement [@Baselga2012] with the Sørensen index.
@@
 The distances are used to construct a dendrogram, which is then split into into *k* bioregions.
-The optimal value of *k* cannot be known *a-priori*.
-Previous vegetation zone schemes in Japan have typically included 5--6 zones [@Shimizu2014]; to enable approximate comparison with these, we tested values of *k* from 1 to 10, and selected the optimal value using the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")`.
+The optimal value of *k* cannot be known a priori.
+Previous vegetation zone schemes in Japan have typically included five or six zones [@Shimizu2014]; to enable approximate comparison with these, we tested values of *k* from 1 to 10, and selected the optimal value using the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")`.
 
@@
 We excluded marine zones and those that do not protect plants.
-We categorized protected areas as either "high" (no human activities allowed at all) or "medium" status [some economic activities allowed by permit\; @Kusumoto2017]; areas not afforded at least medium level of protection were not considered.
+We categorized protected areas as either high (no human activities allowed at all) or medium status [some economic activities allowed by permit\; @Kusumoto2017]; areas not afforded at least medium level of protection were not considered.
 Conservation status in Japan is administered by multiple laws, and protected areas frequently overlap [@Natori2012].
@@
 We downloaded distribution maps (SHP and TIF) of *Cervus nippon* from the Japan Ministry of the Environment (https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
-To assess change in threat due to herbivory by deer over time, we used three maps available in this dataset: range of Japanese deer surveyed in 1978, range surveyed in 2003, and estimated range inferred from a model including snow cover and forest type based on the 2003 survey data [@kankyosho2021].
+To assess change in threat due to herbivory by deer over time, we used three maps available in this data set: range of Japanese deer surveyed in 1978, range surveyed in 2003, and estimated range inferred from a model including snow cover and forest type based on the 2003 survey data [@kankyosho2021].
 We limited estimated range to areas considered "highly likely" to be occupied by deer [movement cost < 0.10\; @kankyosho2021].
 
-We cropped grid-cells to only include land area, then calculated the percent of protected area of each status type or percent of area occupied by deer for each survey period within grid-cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
-These values were compared with the overall percent of protected area or area occupied by deer across Japan (baseline rate).
-For PD, FD, and PE, significance was assessed with a one-tailed test at alpha = 0.05 against the 999 null communities described above.
-For taxonomic richness, grid-cells ranked in the top > 5% were considered highly diverse.
+We cropped grid cells to only include land area, then calculated the percentage of protected area of each status type or percentage of area occupied by deer for each survey period within grid cells with high levels of biodiversity as measured with PD, FD, PE, or taxonomic richness.
+These values were compared with the overall percentage of protected area or area occupied by deer across Japan (baseline rate).
+For PD, FD, and PE, significance was assessed with a one-tailed test at $\alpha$ = 0.05 against the 999 null communities described above.
+For taxonomic richness, grid cells ranked in the top > 5% were considered highly diverse.
 
@@
 
-The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, non-hybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
-The topology was in general agreement with other recent, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a\; `r s_figure("tree")`].
+The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, nonhybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
+The topology generally agreed with earlier, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a\; `r s_figure("tree")`].
 Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016].
-The age of the fern clade is estimated to be `r phylogeny_age` ma.
+The age of the fern clade is estimated to be `r phylogeny_age` Ma.
 
 The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`).
-The distribution of collection locations was highly skewed, with > 3,000 specimens per grid-cell northwest of Tokyo in Saitama prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
-Despite this, sampling redundancy was generally good [> 0.3\; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; errors are SD unless mentioned otherwise; `r s_figure("abundance")`B).
-The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid-cells.
-`r n_excluded_cells` grid-cells were excluded due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.
+The distribution of collection locations was highly skewed, with > 3000 specimens per grid cell northwest of Tokyo in Saitama Prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
+Despite this skew, sampling redundancy was generally good [> 0.3\; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; SDs are given unless mentioned otherwise; `r s_figure("abundance")`B).
+The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid cells.
+`r n_excluded_cells` grid cells were excluded due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.
 
@@
 
-All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and lambda: all lambda values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.1, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
+All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and $\lambda$: all $\lambda$ values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.1, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
 `r n_d_signif` out of `r nrow(binary_sig_results)` of binary traits (including qualitative traits coded as binary) showed evidence of phylogenetic signal (*D* ≤ 0.5, significantly different from random; `r s_table("phy-sig-binary")`).
@@
 
-Grid-cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`A).
-A smaller number of grid-cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (`r eg`, Kii Peninsula).
-Grid-cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`B).
-Grid-cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido.
+Grid cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`A).
+A smaller number of grid cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (`r eg`, Kii Peninsula).
+Grid cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`B).
+Grid cells with significantly high RPD are located both in the small southern islands and northern Honshu and Hokkaido.
 Results of the randomization test for functional diversity were broadly similar to phylogenetic diversity (`r figure("rand-diversity")`C, D).
@@
 
-Nearly all grid-cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`).
-The vast majority of grid-cells with significant endemism are mixed or super-endemic.
+Nearly all grid cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`).
+The vast majority of grid cells with significant endemism are mixed or super-endemic.
 Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands.
-Small clusters of grid-cells containing significant endemism were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`).
-When we ran CANAPE with a reduced dataset including only taxa absolutely restricted to Japan to account for the border effect, there were fewer significant grid-cells, but a similar pattern of significant endemism mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).
+Small clusters of grid cells containing significant endemism were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`).
+When we ran CANAPE with a reduced data set including only taxa absolutely restricted to Japan to account for the border effect, there were fewer significant grid cells, but a similar pattern of significant endemism mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).
 
@@
 
-Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid-cells fell into a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and high elevation areas of northern Honshu), bioregion 2 (low elevation areas of northern Honshu, southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
+Clustering analysis grouped grid cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
+The vast majority of grid cells fell into a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and high elevation areas of northern Honshu), bioregion 2 (low elevation areas of northern Honshu, southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
 Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 covered most of northern Honshu, while extending further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
-The other much smaller bioregions (including only one or two grid-cells each) are likely artifacts of insufficiently low taxon richness for robust clustering and are not discussed further.
+The other much smaller bioregions (including only one or two grid cells each) are likely artifacts of insufficiently low taxon richness for robust clustering and are not discussed further.
 
@@
 
-The percentage of protected area in grid-cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A).
+The percentage of protected area in grid cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A).
 Cells with significantly high biodiversity had a similar or greater percentage of area protected compared to the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
-In contrast, grid-cells with significantly high biodiversity had a similar or lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
-Only grid-cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.
+In contrast, grid cells with significantly high biodiversity had a similar or lower percentage of area protected than the baseline rate for Japan at the high protection status (`r scales::percent(total_percent_protected_high, accuracy = 0.1)`) for most biodiversity metrics.
+Only grid cells with significantly high PE had a higher percentage area protected (`r best_protected_high` high, `r best_protected_med` medium, `r best_protected_total` total) than the baseline rate at both levels of protection.
 
 The range of Japanese deer in Japan increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
-The percentage of area occupied by deer in grid-cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
-Grid-cells with high PD, FD, or PE tended to have less deer occupation compared to the baseline rate (except for PE relative to 1978 levels; `r figure("conserv-status")`B).
-However, grid-cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range dataset.
+The percentage of area occupied by deer in grid cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
+Grid cells with high PD, FD, or PE tended to have less deer occupation compared to the baseline rate (except for PE relative to 1978 levels; `r figure("conserv-status")`B).
+However, grid cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range data set.
 
@@
 Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
-A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or sub-tropical areas [@Kessler2011a; @Suissa2021].
+A hump-shaped pattern of richness along elevation (maximum richness at intermediate elevations) has been frequently observed in ferns on mountains in tropical or subtropical areas [@Kessler2011a; @Suissa2021].
 Hypotheses to explain such a pattern include purely geometrical constraints as well as biological hypotheses positing maximally favorable environmental conditions at mid-elevations [@Colwell2004a].
@@
 The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig-binary")`, `r s_table_num("phy-sig")`), and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
-Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid-cells with significant levels of PE, as well as paleoendemic grid-cells, are located in the southern islands (Figures `r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`).
+Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid cells with significant levels of PE, as well as paleoendemic grid cells, are located in the southern islands (Figures `r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`).
 This is clearly due to the small land area of these islands: small area combined with high PD is expected to lead to high PE.
@@
 
-The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group grid-cells with similar biodiversity values (`r figure("div-by-tax-region")`).
+The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group grid cells with similar biodiversity values (`r figure("div-by-tax-region")`).
 Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
@@
 
-We found that grid-cells with significantly high amounts of biodiversity are similarly or better protected compared to the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`A).
-The high protection status of grid-cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid-cells with high PE (`r s_figure("protected-map")`D).
+We found that grid cells with significantly high amounts of biodiversity are similarly or better protected compared to the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`A).
+The high protection status of grid cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid cells with high PE (`r s_figure("protected-map")`D).
 One possible reason for the generally high level of protection overall is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas [A.
@@
 
-Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid-cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate in recent surveys or modeled data, grid-cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`B).
+Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate in recent surveys or modeled data, grid cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`B).
 This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("deer-map")`A).
@@
 However, ferns are known to have high dispersal abilities [@Tryon1970; @Barrington1993; @Moran2001].
-In our dataset, the taxon with the greatest latitudinal distribution (`r taxon_with_widest_lat$text`) spans nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N).
+In our data set, the taxon with the greatest latitudinal distribution (`r taxon_with_widest_lat$text`) spans nearly the entire region (`r taxon_with_widest_lat$min_lat`°N to `r taxon_with_widest_lat$max_lat`°N), and `r n_widespread_taxa` taxa have latitudinal distributions spanning the Tokara strait in the south (29°N) to the Tsugaru strait in the north (41.5°N).
 Therefore, it seems reasonable to assume that ferns are capable of dispersing across the entire study area.
@@
 **`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.  
-**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km grid-cells across Japan.  
+**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km grid cells across Japan.  
 **`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
-**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20 km grid-cell in the ferns of Japan.  
+**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20 km grid cell in the ferns of Japan.  
 **`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
@@
 **`r s_figure("endemism-restricted")`**.
-Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted data set including only taxa endemic to Japan.
 **`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
@@
 Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
-**`r s_figure("bioregions-env")`**. Scatterplots of grid-cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
-**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.  
-**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).
+**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
+**`r s_figure("protected-map")`**. Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and protected areas.  
+**`r s_figure("deer-map")`**. Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).
 
@@
 (B) Phylogenetic bioregions.
-Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells.
-Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
+Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid cells.
+Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid cells each are lumped into the "Other" category.
 
@@
 Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan.
-(A) Percent of land area with medium or high protected status for grid-cells with significantly high biodiversity by protection status.
-(B) Percent of land area occupied by deer for grid-cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
+(A) Percent of land area with medium or high protected status for grid cells with significantly high biodiversity by protection status.
+(B) Percent of land area occupied by deer for grid cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
 Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
-Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).
+Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid cells in the top 5% considered highly diverse (see Materials and Methods).
 For (A), vertical lines indicate total percent area protected across Japan (baseline protection rate).

### 2022-04-18 R/ms_functions.R

@@ -95,7 +95,7 @@ eg <- "e.g."
 #'
 #' @return Name of R package formatted according to journal requirements
 #' 
-pack <- function(x) {glue::glue("'{x}' v.{packageVersion(x)}")}
+pack <- function(x) {glue::glue("{x} v.{packageVersion(x)}")}
 
 #' Format other software names
 #' 
@@ -114,7 +114,7 @@ software <- function(x) {x}
 #'
 #' @return Name of R function formatted according to journal requirements
 #' 
-func <- function(x) {glue::glue("'{x}'")}
+func <- function(x) {glue::glue("{x}")}
 
 # Pagebreaks ----
 

### 2022-04-18 ms/SI.Rmd
Update MS
As per editor comments

@@
 # Set knitr options
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE, dev="cairo_pdf")
 ```
@@
 
+### **`r s_table("moran")`**
+
+Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.
+Significant *I* > 0 indicates autocorrelation.
+
 ```{r moran, results = "as-is"}
@@
 
-### **`r s_table("moran")`**
-
-Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.
-Significant *I* > 0 indicates autocorrelation.
-
-\clearpage
+\clearpage
+### **`r s_table("corr")`**
+
+Results of modified *t*-test for correlation between variables while taking into account spatial position. Significantly correlated variables at *P* < 0.05 in bold.
 
@@
 
-### **`r s_table("corr")`**
-
-Results of modified *t*-test for correlation between variables while taking into account spatial position. Significantly correlated variables at *P* < 0.05 in bold.
-
-\clearpage
+\clearpage
+
+### **`r s_table("phy-sig-binary")`**
+
+Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.
+*D*, Fritz and Purvis' *D*; *N* (present), number of times a trait was observed present; *N* (absent), number of times a trait was observed absent; *P*(random), probability of obtaining observed trait distribution given random distribution of traits; *P*(Brownian), probability of obtaining observed trait distribution given random evolution by Brownian Motion.
+Presence or absence of indusium was originally a binary trait; other traits (margin, shape, texture) are qualitative traits scored in binary format.
+Values of *D* rounded to three decimal places.
 
@@
 
-### **`r s_table("phy-sig-binary")`**
-
-Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.
-*D*, Fritz and Purvis' *D*; *N* (present), number of times a trait was observed present; *N* (absent), number of times a trait was observed absent; *P*(random), probability of obtaining observed trait distribution given random distribution of traits; *P*(Brownian), probability of obtaining observed trait distribution given random evolution by Brownian Motion.
-Presence or absence of indusium was originally a binary trait; other traits (margin, shape, texture) are qualitative traits scored in binary format.
-Values of *D* rounded to three decimal places.
-
-\clearpage
+\clearpage
+
+### **`r s_table("phy-sig")`**
+
+Phylogenetic signal in continuous functional traits of the ferns of Japan. *K*, Blomberg's *K*; λ, Pagel's λ.
 
@@
 
-### **`r s_table("phy-sig")`**
-
-Phylogenetic signal in continuous functional traits of the ferns of Japan. *K*, Blomberg's *K*; λ, Pagel's λ.
-
-\clearpage
+\clearpage
+
+### **`r s_table("lrt-null")`**
+
+Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).
+Higher (less negative) log likelihood indicates better fit.
+Full environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
+Full reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
+PD, phylogenetic diversity; RPD, relative phylogenetic diversity; FD, functional diversity; RFD, relative functional diversity; SES, standard effect size.
 
@@
 
-### **`r s_table("lrt-null")`**
-
-Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).
-Higher (less negative) log likelihood indicates better fit.
-Full environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
-Full reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
-PD, phylogenetic diversity; RPD, relative phylogenetic diversity; FD, functional diversity; RFD, relative functional diversity; SES, standard effect size.
-
-\clearpage
+\clearpage
+
+### **`r s_table("aic")`**
+
+Model fit as measured with conditional Akaike information criterion (cAIC).
+Environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
+Reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
+Lower cAIC indicates better fit.
 
@@
 ```
-
-### **`r s_table("aic")`**
-
-Model fit as measured with conditional Akaike Information Criterion (cAIC).
-Environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
-Reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
-Lower cAIC indicates better fit.
 
@@
   as.numeric(x) %>% 
-    scales::number(suffix = "km", scale = 1e-3)
+    scales::number(suffix = "", scale = 1e-3)
 }
@@
   labs(
-    x = "Resolution",
+    x = "Resolution (km)",
     y = "Redundancy > 0.5"
@@
   labs(
-    x = "Resolution",
+    x = "Resolution (km)",
     y = "Redundancy = 0"
@@
 (A) Raw taxonomic diversity, (B) phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude.
-For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+For (B) and (C), raw branch lengths were transformed to relative values by dividing each branchlength by the sum of all branch lengths.
 Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].
@@
 
-Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted data set including only taxa endemic to Japan.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism), restricted data set including only taxa endemic to Japan.
 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
@@
 
-Relationship between (A) phylogenetic diversity (PD) and (B) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).
+Relationship between (A) phylogenetic diversity (PD) and (B) relative PD and % apomictic taxa, models inferred using phylogeny with branch lengths in units of expected genetic change (untransformed ML tree).
 Ribbon shows 95% confidence interval of model.

### 2022-04-18 ms/manuscript.Rmd
Update MS
As per editor comments

@@
 
-^4^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
+^3^Department of Botany, National Museum of Nature and Science, Tsukuba, Japan
 
@@
 
-**Methods:** We compiled a community data set of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid cells (20 km × 20 km each) including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
+**Methods:** We compiled a community data set of `r nrow(comm_ferns) %>% number()` grid cells (20 × 20 km each) including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number()` specimen records.
 We combined the community data with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity and modeled biodiversity metrics in response to environmental factors and reproductive mode.
@@
 <!-- may include title words -->
-**Keywords:** apomixis, biodiversity, biogeography, CANAPE, conservation, ferns, Japan, phylogenetic diversity.
+**KEYWORDS**  
+apomixis, biodiversity, biogeography, CANAPE, conservation, ferns, Japan, phylogenetic diversity
 
@@
 Presumably, phylogeny has not been taken into account more prominently in biodiversity studies mainly because the necessary data (DNA sequences and georeferenced occurrence records) and analytic tools have only become available relatively recently [@Soltis2016a; @Folk2021].
-These datasets and tools now make it possible to analyze other dimensions of biodiversity such as phylogenetic diversity [@Faith1992] and phylogenetic endemism [@Rosauer2009].
+These data sets and tools now make it possible to analyze other dimensions of biodiversity such as phylogenetic diversity [@Faith1992] and phylogenetic endemism [@Rosauer2009].
 Better ways to characterize biological regions ("bioregions"; `r ie`, areas defined by their taxonomic composition or evolutionary history) are also available [@Laffan2016; @White2019; @Daru2020a], rather than relying on ad hoc characterizations.
@@
 
-A comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, but such datasets are rare on the regional (country) scale.
+A comprehensive understanding of the relationships between richness and other metrics requires densely sampled data, but such data sets are rare on the regional (country) scale.
 The ferns of Japan are excellent model system because they have been the target of intense botanical interest for several decades and are densely sampled [reviewed by @Ebihara2019b].
@@
 Here, we leveraged this exceptionally rich data set to analyze the geographic distribution of biodiversity in detail.
-We asked the following questions in our study system, the ferns of Japan: 1.
-How is biodiversity distributed?
-2.
-How is biodiversity structured?
-3.
-What environmental and biological factors influence biodiversity?
-4.
-How well is biodiversity protected?
-
-## MATERIALS AND METHODS
+We asked the following questions in our study system, the ferns of Japan: (1) How is biodiversity distributed?
+(2) How is biodiversity structured?
+(3) What environmental and biological factors influence biodiversity?
+(4) How well is biodiversity protected?
+
+# MATERIALS AND METHODS
 
@@
 
-## Study Site
+## Study site
 
@@
 We used a list of native, nonhybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
-We chose to use this over other sources such as the Global Biodiversity Information Facility (GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
-Furthermore, the overall sampling completeness of this dataset is also high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article).
-The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety, excluding non-native taxa and nothotaxa.
+We chose to use this over other sources such as the Global Biodiversity Information Facility (GBIF) because of its high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available data sets.
+Furthermore, the overall sampling completeness of this data set is also high (`r s_figure("sampling-curve")` in Appendix S1).
+The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% number()` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety, excluding non-native taxa and nothotaxa.
 We treated infraspecific taxa as distinct species during analysis and hereafter use "taxon" and "species" interchangeably to refer to entities at this rank, unless otherwise indicated.
@@
 For specimens that lacked latitude and longitude data, georeferencing was done by mapping collection site names to a set of standard ca.
-10 km × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
+10 × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
 In the case that the collection site could not be mapped to a single second-degree mesh cell, it was excluded.
@@
 
-We further cleaned the list before analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
+We further cleaned the list before analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% number()` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
 Given the high quality of our occurrence data and the fact that automated occurrence cleaning algorithms [`r eg`, CoordinateCleaner\; @Zizka2019] have the potential to erroneously exclude true occurrence points [`r ie`, false positives\; @Zizka2020], we chose not to apply additional automated cleaning steps to our data as is often done with occurrence records obtained from GBIF [`r eg`, @Rice2019a; @Suissa2021].
@@
 At each grain size, species occurrences were converted to a presence–absence community matrix (a species was considered present if at least one specimen was recorded in that grid cell).
-We calculated sampling redundancy [1 - Richness/Number of specimens\; @Garcillan2003] to quantify adequacy of sampling.
-Preliminary analysis indicated that 20-km grid cells are optimal for our data set; redundancy values greatly increased from 10 to 20 km, but much less improvement as grain size is increased beyond 20 km (`r s_figure("grain-size")`).
+We calculated sampling redundancy [1 – Richness/Number of specimens\; @Garcillan2003] to quantify adequacy of sampling.
+Preliminary analysis indicated that 20-km grid cells are optimal for our data set; redundancy values greatly increased from 10 to 20 km, but much less as grain size is increased beyond 20 km (Appendix S1: `r s_figure("grain-size")`).
 Although the grid cells are defined with equal area, actual land area of each grid cell varies due to coastlines.
 Strictly filtering out all grid cells with less than complete land area would result in a large loss of data, as Japan has many coastlines and small islands.
-Therefore, we instead filtered the grid cells by sampling completeness, removing any grid cells with redundancy < 0.1; this dataset was used for all subsequent analyses.
+Therefore, we instead filtered the grid cells by sampling completeness, removing any grid cells with redundancy < 0.1; this data set was used for all subsequent analyses.
 The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].
@@
 We used data compiled by @Ebihara2019b, which classifies each fern taxon as sexual, apomictic, mixed (both sexual and apomictic modes known from the same taxon), or unknown.
-We calculated the percentage of apomictic taxa in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa (`r s_figure("envir")`B).
+We calculated the percentage of apomictic taxa in each grid cell as the sum of apomictic and mixed taxa divided by total number of taxa (Appendix S1: `r s_figure("envir")`B).
 
@@
 
-We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`).
+We downloaded climate data at the 2.5 minute-scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster] and extracted four variables related to our research questions: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (Appendix S1: `r s_figure("envir")`).
 We identified the intersection of the climate data at the 2.5-minute scale with the 20-km grid cells, then calculated the mean of each climatic variable for each grid cell.
 We calculated area as rolling mean area (km^2^) across 100-km latitudinal bands.
-This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context because there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).
+This approach reflects the idea that the amount of area with a given climate regime enclosing each grid cell may influence species richness of that cell [@Fahrig2013] and that using raw island areas would not be appropriate in this context because there are just four very large islands with many much smaller ones close by (`r figure("japan-map")`).
 
@@
   n_distinct() %>%
-  number(big.mark = ",")
+  number()
 
@@
 97% of the Japanese fern flora [@Ebihara2019b]; these sequences originate from samples that were identified using the same taxonomic system as the occurrence data [@Ebihara2010; @Ebihara2019b], so we used these preferentially over other data available on GenBank that might include misidentifications or different taxon concepts.
-However, using this dataset alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
+However, using this data set alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
 Furthermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
 Because one goal of the study is to understand the distribution of paleo- vs. neoendemic areas (defined in units of time), we require an ultrametric tree.
-Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes [FTOL v0.0.1\; @Nitta2022b] (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` operational taxonomic units [OTUs] total, including outgroup taxa).
-This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across ca.
-4,000 taxa using six plastid markers [@Testo2016a], and the addition of plastome data can be expected to increase support along the backbone.
-
-We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with 1,000 ultrafast bootstrap replicates [@Minh2013].
+Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid data set including all fern sequences on GenBank for four widely sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes [FTOL v0.0.1\; @Nitta2022b] (`r ape::Ntip(plastome_tree) %>% number()` operational taxonomic units [OTUs] total, including outgroup taxa).
+This gene sampling is comparable to a recent global fern phylogeny that resolved relationships across `r ca` `r number(4000)` taxa using six plastid markers [@Testo2016a], and the addition of plastome data can be expected to increase support along the backbone.
+
+We conducted maximum-likelihood phylogenetic analysis with `r software("IQ-TREE")` v1.6.12 [@Nguyen2015] under the GTR+I+G model (all sites concatenated into a single data matrix), and assessed node support with `r number(1000)` ultrafast bootstrap replicates [@Minh2013].
 Molecular dating was conducted with the program `r software("treePL")` v1.0 [@Smith2012] using `r n_fossil_points` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020].
-The root of the tree (most recent common ancestor of all land plants) was fixed to 475 mya [@Testo2016a].
+The root of the tree (most recent common ancestor of all land plants) was fixed at 475 million years ago [Ma\; @Testo2016a].
 We trimmed the resulting ultrametric tree to only Japanese taxa.
-`r software("treePL")` requires branch lengths to be >0, and it sets extremely small branches to an arbitrary minimum positive length.
-This behavior may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* dataset.
-We converted any clades in the dated tree consisting of identical sequences to polytomies with the node at 0 million years ago (i.e., the present; Ma).
+`r software("treePL")` requires branch lengths to be > 0, and it sets extremely small branches to an arbitrary minimum positive length.
+This behavior may artificially add branch length between taxa with identical DNA sequences, which occur multiple times in the Japanese fern *rbcL* data set.
+We converted any clades in the dated tree consisting of identical sequences to polytomies with the node at 0 Ma (i.e., the present).
 This tree was used for all subsequent analysis unless indicated otherwise.
@@
 
-We measured phylogenetic diversity (PD) using the Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny\; @Faith1992].
+We measured phylogenetic diversity (PD) using Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny\; @Faith1992].
 To detect areas with unusually long or short branches, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].
@@
 We tested whether centers of endemism have an over-representation of old (paleoendemic) or new (neoendemic) lineages using categorical analysis of neo- and paleoendemism (CANAPE), which involves measuring PE with an alternate tree where all branch lengths are equal, then comparing the ratio between raw PE and alternative PE (RPE) [@Mishler2014].
-Low RPE indicates short range-restricted branches (neoendemism) and high RPE indicates long range-restricted branches (paleoendemism).
+Low RPE indicates short range-restricted branches (neoendemism), and high RPE indicates long range-restricted branches (paleoendemism).
 Since all of these measures are affected by species richness, categorization of endemism is done using the rank of each observed variable relative to those calculated for the null distribution.
@@
 It is possible that endemism may be affected by species that are actually wide-ranging, but only have a small range in Japan (a "border effect").
-This is especially expected for southern, subtropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country.
+The border effect is especially expected for southern, subtropical islands, which harbor many tropical species with small ranges in Japan but wider ranges outside of the country.
 To partially account for the border effect, we also calculated all endemism-related metrics using a data set including species found only in Japan.
@@
 
-Two grid cells lacking environmental data, one grid cell occupied by a single species lacking reproductive mode data, and an outlier with 100% apomictic taxa due to extremely low species richness were excluded from modeling analyses (Appendix S2; see Supplemental Data with this article).
+Two grid cells lacking environmental data, one grid cell occupied by a single species lacking reproductive mode data, and an outlier with 100% apomictic taxa due to extremely low species richness were excluded from modeling analyses (Appendix S2).
 
 We tested for spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of FD, SES of RPD, SES of RFD) and independent variables (environmental variables and percentage apomictic taxa [hereafter, % apomictic taxa]) using Moran's I with the `r func("moran.mc")` function in the R package `r pack("spdep")` [@Bivand2013].
-In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05) (`r s_table("moran")`).
+In all cases, significant spatial autocorrelation was detected (Moran's I > 0, *P* value of permutation test < 0.05) (Appendix S1: `r s_table("moran")`).
 Therefore, we used spatially explicit methods for modeling.
@@
 We checked for correlation between independent variables while taking into account spatial autocorrelation using a modified *t*-test with the `r func("modified.ttest")` function in the R package `r pack("SpatialPack")` [@SpatialPack2020].
-This test indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature (`r s_table("corr")`), so for the main modeling analysis we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
+This test indicated that temperature seasonality and % apomictic taxa were correlated with mean annual temperature (Appendix S1: `r s_table("corr")`), so for the main modeling analysis, we only included the following environmental variables: mean annual temperature, precipitation, and precipitation seasonality.
 
@@
 
-## Datasets
+## Data sets
 
@@
 n_specimens <- occ_sum$filtered$n_specimens$value %>%
-  scales::number(big.mark = ",")
+  number()
 n_specimens_taxa <- occ_sum$filtered$n_taxa$value
@@
 The phylogeny included `r ape::Ntip(japan_fern_tree)` taxa (`r phylogeny_sampled_percent` of the total number of native, nonhybrid ferns [`r total_fern_taxa` taxa]) representing `r n_fern_phylo_genera` genera and `r n_fern_phylo_families` families.
-The topology generally agreed with earlier, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a\; `r s_figure("tree")`].
+The topology generally agreed with earlier, large plastid fern phylogenies [@Schuettpelz2007; @Lehtonen2011; @Testo2016a\; Appendix S1: `r s_figure("tree")`].
 Ferns were recovered as monophyletic, as were all families and genera except for *`r non_mono_genus`* [families follow @PteridophytePhylogenyGroupI2016].
@@
 The cleaned list of specimens included `r n_specimens` specimens representing `r n_specimens_taxa` taxa (`r n_specimens_taxa_percent`).
-The distribution of collection locations was highly skewed, with > 3000 specimens per grid cell northwest of Tokyo in Saitama Prefecture, and many fewer elsewhere (`r s_figure("abundance")`A).
-Despite this skew, sampling redundancy was generally good [> 0.3\; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; SDs are given unless mentioned otherwise; `r s_figure("abundance")`B).
-The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` grid cells.
-`r n_excluded_cells` grid cells were excluded due to low redundancy (< 0.1); these were mostly from Hokkaido and coastal areas.
-
-After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits; Tables `r table_num("traits-used")`, `r s_table_num("phy-sig-binary")`).
+The distribution of collection locations was highly skewed, with > `r number(3000)` specimens per grid cell northwest of Tokyo in Saitama Prefecture, and many fewer elsewhere (Appendix S1: `r s_figure("abundance")`A).
+Despite this skew, sampling redundancy was generally good [> 0.3\; @Gonzalez-Orozco2014b] throughout the country (mean `r mean_redundancy`, final community data matrix; SDs are given unless mentioned otherwise; Appendix S1: `r s_figure("abundance")`B).
+The final community data matrix included `r ncol(comm_ferns)` taxa and `r nrow(comm_ferns) %>% number()` grid cells after `r n_excluded_cells` grid cells (mostly from Hokkaido and coastal areas) were excluded due to low redundancy (< 0.1).
+
+After removing correlated and low-variance traits, the final trait set included `r n_binary_traits` binary trait, `r n_cont_traits` continuous traits, and `r n_qual_traits` qualitative traits (coded as `r n_qual_traits_as_binary` binary traits; Table `r table_num("traits-used")`; Appendix S1: `r s_table("phy-sig-binary")`).
 The trait matrix included `r n_taxa_traits_sampled` taxa (`r percent_taxa_traits_sampled`) and `r fern_traits %>% select(-taxon) %>% ncol()` traits, with `r percent_traits_missing` missing data.
@@
 
-All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and $\lambda$: all $\lambda$ values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.1, indicating less phylogenetic signal than expected under BM (`r s_table("phy-sig")`).
-`r n_d_signif` out of `r nrow(binary_sig_results)` of binary traits (including qualitative traits coded as binary) showed evidence of phylogenetic signal (*D* ≤ 0.5, significantly different from random; `r s_table("phy-sig-binary")`).
+All continuous traits showed significant phylogenetic signal, but estimates of signal strength differed between *K* and $\lambda$: all $\lambda$ values were > 0.95, indicating evolution by BM; however, values of *K* were < 0.1, indicating less phylogenetic signal than expected under BM (Appendix S1: `r s_table("phy-sig")`).
+`r n_d_signif` of `r nrow(binary_sig_results)` of binary traits (including qualitative traits coded as binary) showed evidence of phylogenetic signal (*D* ≤ 0.5, significantly different from random; Appendix S1: `r s_table("phy-sig-binary")`).
 
@@
 
-Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`A, `r s_figure("raw-lat")`A).
-Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (`r s_figure("raw-biplot")`A, C).
-Observed values of RPD and RFD were not linearly related to richness; however, they showed high variance at low species richness, and low variance at high species richness (`r s_figure("raw-biplot")`B, D).
+Taxonomic richness is lowest on the northernmost island of Hokkaido, then increases with decreasing latitude until reaching a maximum in southern Kyushu, then decreases again in smaller islands further south (`r figure("raw-diversity")`A; Appendix S1: `r s_figure("raw-lat")`A).
+Observed values of PD and FD were highly correlated with richness as expected (`r figure("raw-diversity")`B, C), showing an initial steep slope that gradually starts to level off at higher richness (Appendix S1: `r s_figure("raw-biplot")`A, C).
+Observed values of RPD and RFD were not linearly related to richness; however, they showed high variance at low species richness, and low variance at high species richness (Appendix S1: `r s_figure("raw-biplot")`B, D).
 
@@
 
-All models were significant relative to null models without any fixed effects (`r s_table("lrt-null")`).
+All models were significant relative to null models without any fixed effects (Appendix S1: `r s_table("lrt-null")`).
 The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
@@
 Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E).
-SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F).
-Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (`r s_table("aic")`).
-
-## Categorical analysis of neo- and paleo-endemism
-
-Nearly all grid cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (non-significant) endemism observed elsewhere (`r figure("endemism")`).
+Standard effect size of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F).
+Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (Appendix S1: `r s_table("aic")`).
+
+## Categorical analysis of neo- and paleoendemism
+
+Nearly all grid cells in islands south of Kyushu (Ryukyu and Ogasawara) have significant levels of phylogenetic endemism, with low (nonsignificant) endemism observed elsewhere (`r figure("endemism")`).
 The vast majority of grid cells with significant endemism are mixed or super-endemic.
-Concentrations of paleoendemism were observed in the Ryukyu archipelago, on Okinawa, Miyako, and Iriomote Islands.
-Small clusters of grid cells containing significant endemism were also observed northwest of Tokyo (Nagano prefecture), and in Hokkaido (`r figure("endemism")`).
-When we ran CANAPE with a reduced data set including only taxa absolutely restricted to Japan to account for the border effect, there were fewer significant grid cells, but a similar pattern of significant endemism mostly occurring in the southern islands was observed (`r s_figure("endemism-restricted")`).
+Concentrations of paleoendemism were observed in the Ryukyu archipelago, primarily on Okinawa, Miyako, and Iriomote Islands.
+Small clusters of grid cells containing significant endemism were also observed northwest of Tokyo (Nagano Prefecture) and on Hokkaido (`r figure("endemism")`).
+When we ran CANAPE with a reduced data set including only taxa absolutely restricted to Japan to account for the border effect, there were fewer significant grid cells, but a similar pattern of significant endemism mostly occurring in the southern islands was observed (Appendix S1: `r s_figure("endemism-restricted")`).
 
@@
 
-Clustering analysis grouped grid cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
-The vast majority of grid cells fell into a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and high elevation areas of northern Honshu), bioregion 2 (low elevation areas of northern Honshu, southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A).
-Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 covered most of northern Honshu, while extending further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B).
+Clustering analysis grouped grid cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (Appendix S1: `r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively.
+The vast majority of grid cells fell into a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and high elevation areas of northern Honshu), bioregion 2 (low elevation areas of northern Honshu, southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; Appendix S1: `r s_figure("dendrogram")`A).
+Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 covered most of northern Honshu, while extending further south (`r figure("bioregions")`B; Appendix S1: `r s_figure("dendrogram")`B).
 The other much smaller bioregions (including only one or two grid cells each) are likely artifacts of insufficiently low taxon richness for robust clustering and are not discussed further.
@@
 Bioregion 2 has lower RPD and RFD than bioregion 1.
-Comparison of biodiversity metrics across phylogenetic bioregions was similar, except that phylogenetic bioregion 3 mostly corresponds to taxonomic bioregions 3 and 4 combined (`r s_figure("div-by-phy-region")`).
+Comparison of biodiversity metrics across phylogenetic bioregions was similar, except that phylogenetic bioregion 3 mostly corresponds to taxonomic bioregions 3 and 4 combined (Appendix S1: `r s_figure("div-by-phy-region")`).
 
@@
       as.numeric %>%
-      scales::number(accuracy = 1, big.mark = ",") %>%
+      number(accuracy = 1) %>%
       paste("km^2^"),
@@
 
-The range of Japanese deer in Japan increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003, and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
+The range of Japanese deer in Japan increased from `r deer_overall[["1978"]]$area` (`r deer_overall[["1978"]]$percent`) in 1978 to `r deer_overall[["2003"]]$area` (`r deer_overall[["2003"]]$percent`) in 2003 and may be as high as `r deer_overall$estimated$area` (`r deer_overall$estimated$percent`) according to model estimates.
 The percentage of area occupied by deer in grid cells with significantly high biodiversity increased from `r deer_in_high_biodiv[["1978"]]$percent` in 1978 to `r deer_in_high_biodiv[["2003"]]$percent` in 2003 when averaged across biodiversity types.
-Grid cells with high PD, FD, or PE tended to have less deer occupation compared to the baseline rate (except for PE relative to 1978 levels; `r figure("conserv-status")`B).
-However, grid cells with high species richness consistently exceeded the baseline rate of deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range data set.
+Grid cells with high PD, FD, or PE tended to have fewer deer compared to the baseline rate (except for PE relative to 1978 levels; `r figure("conserv-status")`B).
+However, grid cells with high species richness consistently exceeded the baseline rate for deer presence, with up to `r deer_highest_danger$estimated$percent` occupied by deer in the estimated range data set.
 
@@
 
-The latitudinal hump-shaped pattern of species richness observed in ferns here (Figures `r figure_num("raw-diversity")`A, `r s_figure_num("raw-lat")`A) is well-documented [@Kusumoto2017; @Ebihara2019b].
+The latitudinal hump-shaped pattern of species richness observed in ferns here (`r figure("raw-diversity")`A; Appendix S1: `r s_figure("raw-lat")`A) is well documented [@Kusumoto2017; @Ebihara2019b].
 Similar patterns have been observed in deciduous broadleaved trees and conifers [@Shiono2015], and niche-modeled distributions of woody plants in Japan [@Fukaya2020].
-Our spatial modeling analysis indicates that this is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
+Our spatial modeling analysis indicates that this hump-shaped pattern is strongly linked to temperature, with maximum richness occurring at intermediate temperatures (`r figure("modelfit")`A).
 Previous studies have identified temperature, water availability [@Bickford2006; @Qian2012], or a combination of these [@Nagalingum2015; @Link-Perez2018; @Qian2021] to be important drivers of spatial fern diversity.
@@
 
-The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of southern subtropical islands (Ryukyu and Ogasawara Isl.) (`r figure("rand-diversity")`A, C).
-This is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country, which lack subtropical climate; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Isl.).
-The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Tables `r s_table_num("phy-sig-binary")`, `r s_table_num("phy-sig")`), and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
-Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid cells with significant levels of PE, as well as paleoendemic grid cells, are located in the southern islands (Figures `r figure_num("endemism")`, `r s_figure_num("endemism-restricted")`).
-This is clearly due to the small land area of these islands: small area combined with high PD is expected to lead to high PE.
-
-Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands, and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`B).
-The long branches in the southern islands are likely being driven by the tropical lineages restricted to this area; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high latitudes, such as *Equisetum* (`r s_figure("genus-map")`A).
-The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera there such as *Deparia* (`r s_figure("genus-map")`B) and *Dryopteris* (`r s_figure("genus-map")`D).
+The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of the southern subtropical islands (Ryukyu and Ogasawara Islands) (`r figure("rand-diversity")`A, C).
+This pattern is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country, which lack subtropical climate; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Island).
+The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Appendix S1: Tables `r s_table_num("phy-sig-binary")`, `r s_table_num("phy-sig")`) and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
+Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid cells with significant levels of PE, as well as paleoendemic grid cells, are located in the southern islands (`r figure("endemism")`; Appendix S1: `r s_figure("endemism-restricted")`).
+The high PE observed on the southern subtropical islands is clearly due to the small land area of these islands: small area combined with high PD is expected to lead to high PE.
+
+Another striking pattern revealed here is the predominance of long branches in the northern half of Honshu and the southern islands and the predominance of short branches in southern Honshu, Shikoku, and Kyushu (`r figure("rand-diversity")`B).
+The long branches in the southern islands are likely driven by the tropical lineages restricted to this area; those in the north may be due to refugial lineages that are distantly related to others and become more species rich at high latitudes (`r eg`, *Equisetum*; Appendix S1: `r s_figure("genus-map")`A).
+The preponderance of short branches in southern Honshu, Shikoku, and Kyushu may be due to the radiation of certain species-rich genera such as *Deparia* (Appendix S1: `r s_figure("genus-map")`B) and *Dryopteris* (Appendix S1: `r s_figure("genus-map")`D).
 
 Furthermore, we identified % apomictic taxa as a strong driver of SES of RPD (Figures `r figure_num("coeff")`G, `r figure_num("modelfit")`G), and to a slightly lesser extent SES of PD (Figures `r figure_num("coeff")`F, `r figure_num("modelfit")`F).
-Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (`r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (`r s_figure("genus-map")`).
-This is consistent with the hypothesis that communities with high proportions of asexual taxa have lower PD, and is the first time to our knowledge that this has been demonstrated in ferns.
-We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (`r s_figure("modelfit-not-ult")`).
-While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (`r s_table("aic")`).
-Ultimately, it is likely that temperature, in combination with other environmental conditions, is driving the predominance of apomictic taxa: @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca.
-34.8 °N latitude and increases with temperature in Japan (their study was limited to the main islands and did not include the smaller southern islands).
+Indeed, *Dryopteris* is one such genus with a high rate of apomictic taxa (Appendix S1: `r s_figure("apo-genus")`), and other genera with high rates of apomixis (*Cyrtomium*, *Pteris*; Appendix S1: `r s_figure("apo-genus")`) also reach greatest richness in the southern main islands (Appendix S1: `r s_figure("genus-map")`).
+The observation that % apomictic taxa drives SES of RPD is consistent with the hypothesis that communities with high proportions of asexual taxa have lower PD, and is the first time to our knowledge that this process has been demonstrated in ferns.
+We verified that this relationship is also observed when analyzed with a phylogeny with branch lengths in units of genetic change (`r ie`, the output of the maximum likelihood phylogenetic analysis, before dating), a more direct measure of genetic diversity (Appendix S1: `r s_figure("modelfit-not-ult")`).
+While temperature is also correlated with % apomictic taxa and therefore cannot be discounted, conditional AIC indicates a better fit of the reproductive model than the environmental model for SES of RPD (Appendix S1: `r s_table("aic")`).
+Ultimately, temperature, in combination with other environmental conditions, is likely driving the predominance of apomictic taxa; @Tanaka2014 also found that the proportion of apomictic taxa reaches a maximum at ca.
+34.8°N latitude and increases with temperature in Japan (their study was limited to the main islands and did not include the smaller southern islands).
 @Tanaka2014 asserted that one limiting factor affecting establishment of apomictic ferns at high latitudes or cold climates is the large cell size of apomicts, which are typically polyploid.
@@
 Water limitation is thought to play an important role in promoting apomixis in ferns especially in deserts [@Grusz2021], since ferns depend on water for transfer of sperm to egg, and apomictic plants would be able to reproduce in the absence of water.
-A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomixis in ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas including high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
-The monsoon hypothesis was not supported by our results, which did not show a particularly strong effect of precipitation seasonality in most models (`r figure("coeff")`), nor a correlation with % apomictic taxa (`r s_table("corr")`).
+A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomixis in ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas that include high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
+The monsoon hypothesis was not supported by our results, which showed neither a particularly strong effect of precipitation seasonality in most models (`r figure("coeff")`), nor a correlation with % apomictic taxa (Appendix S1: `r s_table("corr")`).
 
 The variation in distribution of different biodiversity metrics is reflected by the bioregions analysis, which tended to group grid cells with similar biodiversity values (`r figure("div-by-tax-region")`).
-Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized on an *ad-hoc* basis; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
+Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized ad hoc; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
 Furthermore, the border between bioregions 1 and 2 approximately corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
-This suggests that ferns are useful as bioindicators in Japan.
-Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], these did not play a major role in structuring fern bioregions.
-The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, there is some overlap between them on Yakushima and Amami Isl.
-(`r figure("bioregions")`B).
+The agreement between qualitative analysis of bioregions here with previous ad hoc schemes suggests that ferns are useful as bioindicators in Japan.
+Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], they did not play a major role in structuring fern bioregions.
+The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, the regions have some overlap on Yakushima and Amami Islands (`r figure("bioregions")`B).
 The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
-Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu islands; wetter) from bioregion 4 (mostly the Ogasawara islands; drier) (`r s_figure("bioregions-env")`).
+Rather, fern bioregions in Japan broadly separate along gradients of temperature and precipitation; in particular, precipitation splits taxonomic bioregion 3 (mostly the Ryukyu Islands; wetter) from bioregion 4 (mostly the Ogasawara Islands; drier) (Appendix S1: `r s_figure("bioregions-env")`).
 
@@
 We found that grid cells with significantly high amounts of biodiversity are similarly or better protected compared to the baseline protection rate at the medium protection level across all biodiversity metrics, and better than the baseline rate at the high protection level for PE (`r figure("conserv-status")`A).
-The high protection status of grid cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid cells with high PE (`r s_figure("protected-map")`D).
+The high protection status of grid cells with high PE may be due to an emphasis on conserving endemic plant and animal species in Japanese southern islands such as Iriomote and Amami [@kankyosho2021b], which include many grid cells with high PE (Appendix S1: `r s_figure("protected-map")`D).
 One possible reason for the generally high level of protection overall is that in the past, some ferns have been included in lists of rare and endemic plants for consideration in designation of conservation areas [A.
-Ebihara, pers.
-comm.\; @kankyosho2021b].
+Ebihara, personal communication\; @kankyosho2021b].
 Also, (as suggested previously) ferns may serve as bioindicators reflecting overall levels of biodiversity, particularly for other plant groups.
 
-Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats posed that conservation zones cannot ameliorate, the most obvious being climate change.
-One other such threat is herbivory by native Japanese deer, *Cervus nippon* [@Minamitani2005; @Yahara2006; @Hattori2010; @kankyosho2022].
+Despite the good news that ferns enjoy relatively strong protection status in Japan, we must bear in mind that there are other threats that conservation zones cannot ameliorate, the most obvious being climate change.
+One other such threat is herbivory by the native Japanese deer, *Cervus nippon* [@Minamitani2005; @Yahara2006; @Hattori2010; @kankyosho2022].
 The rapid increase in population size observed in surveys from 1978 to 2003 is likely due to a combination of climate change and a decrease in human hunters and natural predators [@Takatsuki2009].
@@
 Our analysis shows that the threat to fern biodiversity posed by Japanese deer differs strongly by biodiversity type; although grid cells with significantly high FD, PD, and PE tend to have lower rates of deer occupancy compared to the baseline rate in recent surveys or modeled data, grid cells with high richness experience much greater co-occurrence with deer than the baseline rate (`r figure("conserv-status")`B).
-This pattern is due to the fact that Japanese deer do not occur in the southern islands, which house high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (`r s_figure("deer-map")`A).
+This pattern is due to the fact that Japanese deer do not occur in the southern islands, which harbor high amounts of PD and PE; rather, areas with high species richness are primarily located in southern Honshu, Shikoku, and Kyushu, where deer expansion has been intense (Appendix S1: `r s_figure("deer-map")`A).
 This result highlights the importance of using multiple metrics of biodiversity for establishing conservation priorities.
@@
 
-Our study integrates diverse, thoroughly collected datasets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns.
+Our study integrates diverse, thoroughly collected data sets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns.
 We conclude by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.
@@
 First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan.
-Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (`r s_figure("abundance")`).
+Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (Appendix S1: `r s_figure("abundance")`).
 Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.
@@
 Therefore, it seems reasonable to assume that ferns are capable of dispersing across the entire study area.
-Furthermore, it should be noted that the southern islands are all low-altitude (not exceeding 1,000 m).
-If high elevation habitats were available, the southern islands would be expected to share more species in common with areas further north.
-For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (> 3,000 m), shares many fern species with the main Japanese islands that are not found in southern Japanese islands.
-This supports that the sharp taxonomic turnover between southern and main islands is more likely due to ecological factors than dispersal limitation.
+Furthermore, it should be noted that the southern islands are all low altitude (not exceeding `r number(1000)` m).
+If high-elevation habitats were available, the southern islands would be expected to share more species in common with areas farther north.
+For example, Taiwan, which is just east of the southernmost Japanese islands and harbors high mountains (> `r number(3000)` m), shares many fern species with the main Japanese islands that are not found in southern Japanese islands.
+This disjunct pattern supports that the sharp taxonomic turnover between the southern and main islands is more likely due to ecological factors than to dispersal limitation.
 
@@
 The Japanese fern flora includes `r n_hybrid_fern_taxa` nothotaxa [@Ebihara2019b].
-Although hybrids are often thought of as evolutionary "dead-ends", in some cases they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989].
+Although hybrids are often thought of as evolutionary "dead ends", in some cases, they are able to interbreed with diploids and thus contribute to diversification [@Barrington1989].
 Future studies should consider the distribution of hybrids in Japan and their possible effects on biodiversity.
@@
 Unlike seed plants, ferns have gametophytes that are capable of growing independently from the sporophyte.
-Futhermore, the two stages of the life cycle have vastly different morphology and ecophysiology, and may even occur over partially to completely disjunct ranges [reviewed in @Pinson2016a].
-Here, as in the vast majority of fern ecology studies, we consider only the sporophyte stage.
+Futhermore, the two stages of the life cycle have vastly different morphology and ecophysiology, and may even occur over partially to completely disjunct ranges [reviewed by @Pinson2016a].
+Here, as in the vast majority of ecological fern studies, we considered only the sporophytic stage.
 However, recent studies in Japan [@Ebihara2013] and elsewhere [@Nitta2017] have revealed different patterns in the community structure of gametophytes and sporophytes.
@@
 
-## ACKNOWLEDGEMENTS
-
-The authors thank Members of the Iwasaki Lab at the University of Tokyo and two anonymous reviewers for providing comments that improved the manuscript.
+# ACKNOWLEDGEMENTS
+
+The authors thank members of the Iwasaki Lab at the University of Tokyo and two anonymous reviewers for comments that improved the manuscript.
 Shawn Laffan provided advice about CANAPE.
 This study was supported by JSPS KAKENHI Grant Number 16H06279.
-The authors are grateful to members of the Nippon Fernist Club for their cooperation and efforts towards quantifying and conserving the biodiversity of the ferns of Japan.
-
-## AUTHOR CONTRIBUTIONS
-
-J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.
-
-## DATA AVAILABILITY STATEMENT
+The authors are grateful to members of the Nippon Fernist Club for their cooperation and efforts toward quantifying and conserving the biodiversity of the ferns of Japan.
+
+# AUTHOR CONTRIBUTIONS
+
+J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all coauthors.
+
+# DATA AVAILABILITY STATEMENT
 
@@
 
-## SUPPORTING INFORMATION
+# SUPPORTING INFORMATION
 
@@
 **`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
-**`r s_table("aic")`**. Model fit as measured with conditional Akaike Information Criterion (cAIC).  
+**`r s_table("aic")`**. Model fit as measured with conditional Akaike information criterion (cAIC).  
 **`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
 **`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.  
-**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km grid cells across Japan.  
+**`r s_figure("envir")`**. Reproductive mode and environmental data in 20-km grid cells across Japan.  
 **`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
-**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20 km grid cell in the ferns of Japan.  
+**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20-km grid cell in the ferns of Japan.  
 **`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
@@
 **`r s_figure("endemism-restricted")`**.
-Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted data set including only taxa endemic to Japan.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism), restricted data set including only taxa endemic to Japan.
 **`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
@@
 
+**`r table("traits-used")`**.
+Fern traits used in this study.
+Letters in parentheses indicate trait type: b, binary (trait present or absent); c, continuous; q, qualitative.
+Qualitative traits were coded as binary for analysis (see Materials and Methods).
+For complete list of trait states for qualitative traits, see Appendix S1: `r s_table("phy-sig-binary")`.
+
 | Trait               | Significance                                               | Reference       |
@@
 | Frond width (c)     | Scales with overall plant size                             | @Creese2011     |
-| No. pinna pairs (c) | Less divided fronds have lower rates of evapotranspiration | @Kluge2007      |
+| No. pinna pairs (c) | Less-divided fronds have lower rates of evapotranspiration | @Kluge2007      |
 | Stipe length (c)    | Shorter lengths minimize resistance                        | @WatkinsJr.2010 |
 | Frond margin (q)    | Thermoregulation                                           | @Nicotra2011    |
-| Frond shape (q)     | Less divided fronds have lower rates of evapotranspiration | @Kluge2007      |
+| Frond shape (q)     | Less-divided fronds have lower rates of evapotranspiration | @Kluge2007      |
 | Frond texture (q)   | Thicker leaves have lower rates of evapotranspiration      | @Givnish1987    |
-
-**`r table("traits-used")`**.
-Fern traits used in this study.
-Letters in parentheses indicate trait type: b, binary (trait present or absent); c, continuous; q, qualitative.
-Qualitative traits were coded as binary for analysis (see Materials and Methods).
-For complete list of trait states for qualitative traits, see `r s_table("phy-sig-binary")`.
 
@@
 (C) Raw functional diversity.
-For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths.
+For (B) and (C), raw branch lengths were transformed to relative values by dividing each branch length by the sum of all branch lengths.
 
@@
     hjust = "inward", vjust = "inward", color = "white") +
-  scale_fill_manual(values = signif_cols) +
+  scale_fill_manual(
+    name = "Significance",
+    values = signif_cols,
+    labels = signif_labels) +
   map_theme() +
@@
     axis.text.x = element_blank(),
-    legend.title = element_blank()
+    legend.text = element_markdown()
   )
@@
   labs(fill = "") +
-  scale_fill_manual(values = signif_cols) +
+  scale_fill_manual(
+    name = "Significance",
+    values = signif_cols,
+    labels = signif_labels) +
   map_theme() +
@@
     axis.text.y = element_blank(),
-    legend.title = element_blank()
+    legend.text = element_markdown()
   )
@@
   labs(fill = "") +
-  scale_fill_manual(values = signif_cols) +
+  scale_fill_manual(
+    name = "Significance",
+    values = signif_cols,
+    labels = signif_labels) +
   scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
   map_theme() +
-  theme(legend.title = element_blank())
+  theme(legend.text = element_markdown())
 
@@
   labs(fill = "") +
-  scale_fill_manual(values = signif_cols) +
+  scale_fill_manual(
+    name = "Significance",
+    values = signif_cols,
+    labels = signif_labels) +
   scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
@@
 **`r figure("rand-diversity") %>% str_to_upper()`**.
-Results of randomization tests for phylogenetic and functional diversity of the ferns of Japan.
+Results of randomization tests for phylogenetic and functional diversity of ferns of Japan.
 (A) Phylogenetic diversity (PD).
@@
 (D) Relative functional diversity (RFD).
-For each metric, raw values were compared to a null distribution of 999 random values, and the *P*-value calculated using a two-tailed test.
+For each metric, raw values were compared to a null distribution of 999 random values, and the *p*-value calculated using a two-tailed test.
 For details, see Materials and Methods.
@@
 Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models for the effect of environment and reproductive mode on biodiversity metrics in the ferns of Japan.
-(A)--(E) Environmental models (models including temperature).
+(A--E) Environmental models (models including temperature).
 (F, G) Reproductive models (models including % apomictic taxa).
-Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *P*-value indicates probability of no difference between the full and null model.
+Statistical significance was assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-values indicate probability of no difference between the full and null model.
 Name of response variable and model type indicated above each subplot.
@@
 Relationship between biodiversity metrics and selected predictor variables in the ferns of Japan.
-(A)--(E) Environmental models (models including temperature).
-(F, G) Reproductive models (models including % apomictic taxa).
+(A--E) Environmental (env.) models (models including temperature).
+Precip., precipitation; Temp., temperature; Temp^2^, temperature squared.
+(F, G) Reproductive (repro.) models (models including % apomictic taxa).
 Ribbon shows 95% confidence interval of model.
 Line fit with the focal predictor variable while averaging over other predictors.
-For phylogenetic diversity (PD), relative PD (RPD), functional diversity (FD), and relative FD (RFD), the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*P*-value) calculated using a two-tailed test (see Materials and Methods).
+For phylogenetic diversity (PD), relative PD (RPD), functional diversity (FD), and relative FD (RFD), the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*p*-value) determined using a two-tailed test (see Materials and Methods).
 
@@
   geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
-  scale_fill_manual(values = canape_cols) +
+  scale_fill_manual(
+    values = canape_cols,
+    breaks = names(canape_cols)[names(canape_cols) != "not significant"],
+    name = "Endemism type") +
   annotation_scale(location = "br", width_hint = 0.2) +
-  map_theme() +
-  theme(legend.title = element_blank())
+  map_theme()
 ```
@@
 **`r figure("endemism") %>% str_to_upper()`**.
-Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism).
-'neo' indicates areas with an overabundance of rare short branches; 'paleo' indicates areas with an overabundance of rare long branches; 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism).
+neo, grid cells with an overabundance of rare short branches; paleo, grid cells with an overabundance of rare long branches; mixed, significantly endemic grid cells that have neither an overabundance of short nor long branches; super, highly significantly endemic grid cells that have neither an overabundance of short nor long branches.
+Grid cells without significant endemism in grey.
 For details, see Materials and Methods.
@@
 Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid cells.
-Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid cells each are lumped into the "Other" category.
+Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid cells each were lumped into category "Other".
 
@@
 (D) SES of relative FD (RFD).
-(E) *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities.
+(E) The *p*-score for observed phylogenetic endemism (PE) relative to 999 random communities.
 
@@
 Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan.
-(A) Percent of land area with medium or high protected status for grid cells with significantly high biodiversity by protection status.
-(B) Percent of land area occupied by deer for grid cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
+(A) Percentage of land area with medium or high protected status for grid cells with significantly high biodiversity by protection status.
+(B) Percentage of land area occupied by deer for grid cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data).
 Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE).
 Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid cells in the top 5% considered highly diverse (see Materials and Methods).
-For (A), vertical lines indicate total percent area protected across Japan (baseline protection rate).
+For (A), vertical lines indicate total percentage area protected across Japan (baseline protection rate).
 For (B), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).

### 2022-04-18 R/ms_functions.R

@@ -32,6 +32,14 @@ signif_cols <-
     "> 0.975" = colorb_paired[[1]], # Light blue 
     "> 0.99" = colorb_paired[[2]] # Dark blue 
   )
+signif_labels <-
+  c(
+    "< 0.01" = "*p* < 0.01", #  Dark red
+    "< 0.025" = "*p* < 0.025", # Light red
+    "not significant" = "not significant", # Beige
+    "> 0.975" = "*p* > 0.975", # Light blue 
+    "> 0.99" = "*p* > 0.99" # Dark blue 
+  )
 
 # - CANAPE (Okabe-Ito CVD safe)
 canape_cols <-
@@ -86,6 +94,46 @@ deer_lines <- c(
 # AJB: Common Latin words (e.g., in vivo, sensu lato) are not italicized.
 ie <- "i.e."
 eg <- "e.g."
+ca <- "ca."
+
+# Helper function to count decimal places in a number
+# https://stackoverflow.com/questions/5173692/how-to-return-number-of-decimal-places-in-r
+decimalplaces <- function(x) {
+  # We only care about decimal places, not actual number so this is OK:
+  # Add 1 so as.character() doesn't result in scientific notation
+  if (x < 1) x <- 1 + x
+  if (abs(x - round(x)) > .Machine$double.eps^0.5) {
+    as.character(x) %>%
+      sub('0+$', '', .) %>%
+      strsplit(".", fixed = TRUE) %>%
+      purrr::pluck(1) %>%
+      purrr::pluck(2) %>%
+      nchar()
+  } else {
+    # If number of decimals exceeds machine precision, return
+    # default number of digits for printing
+    getOption("digits")
+  }
+}
+
+# Format numbers
+# AJB rules:
+# no commas with four-digit numbers (e.g., 1000 instead of 1,000)
+# 0.13 instead of .13.
+number <- function(x, accuracy = NULL, ...) {
+  
+  # Set default accuracy to number of decimal places
+  # in the number, if precision is below machine precision
+  if (is.null(accuracy) & abs(x - round(x)) > .Machine$double.eps^0.5) {
+    accuracy <- 1/10^decimalplaces(x)
+  }
+  
+  if (as.integer(x) > 9999) {
+    scales::number(x, big.mark = ",", accuracy = accuracy, ...)
+  } else {
+    scales::number(x, big.mark = "", accuracy = accuracy, ...)
+  }
+}
 
 #' Format R packages names
 #' 

### 2022-04-27 ms/data_exploration.Rmd
Fix URL for data in SI2

@@
 
-# Alternatively, unzip `results.zip`
-# (FIXME: add DOI when available)
-# and read in the data from there
+# Alternatively, download `results.zip` from
+# https://figshare.com/ndownloader/files/34354472
+# unzip it, and read in the data from there
 # biodiv_ferns_cent_figshare <- read_csv(here::here("japan_ferns_biodiv.csv"))

### 2022-04-27 ms/SI.Rmd
Update SI1 as per copy-editor changes

@@
 Significant *I* > 0 indicates autocorrelation.
+SES, standard effect size; PD, phylogenetic diversity; RPD, relative PD; FD, functional diversity; RFD, relative FD.
 
@@
 
-Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).
+Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix) for spatial models describing the effect of environmental or biological variables on the biodiversity of the ferns of Japan.
 Higher (less negative) log likelihood indicates better fit.
@@
 Full reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
-PD, phylogenetic diversity; RPD, relative phylogenetic diversity; FD, functional diversity; RFD, relative functional diversity; SES, standard effect size.
+PD, phylogenetic diversity; RPD, relative PD; FD, functional diversity; RFD, relative FD; SES, standard effect size.
 
@@
 
-Model fit as measured with conditional Akaike information criterion (cAIC).
+Model fit as measured with conditional Akaike information criterion (cAIC) for spatial models describing the effect of environmental or biological variables on the biodiversity of the ferns of Japan.
 Environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
@@
 Lower cAIC indicates better fit.
+PD, phylogenetic diversity; RPD, relative PD; SES, standard effect size.
 
@@
     x = "Redundancy",
-    y = "Resolution"
+    y = "Resolution (km)"
   )
@@
 
-Effect of grain size on sampling redundancy.
-(A) Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km).
-(B) Percent of grid cells with redundancy = 0 by grain size.
+Effect of grain size on sampling redundancy of the ferns of Japan.
+(A) Percentage of grid cells with redundancy >0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km).
+(B) Percentage of grid cells with redundancy = 0 by grain size.
 (C) Distribution of redundancy values by grain size.
@@
 a <- plot_env("A", "Area", "lat_area", "km^2")
-b <- plot_env("B", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
+b <- plot_env("B", "% Apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
 c <- plot_env("C", "Precipitation", "precip", "mm")
@@
 
-Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan.
-(A) rolling mean area (km^2^) in 20 km latitudinal bands.
-(B) % apomictic taxa.
-(C) annual precipitation.
-(D) percipitation seasonality (coefficient of variation).
-(E) mean annual temperature.
-(F) temperature seasonality (standard deviation × 100).
+Reproductive mode and environmental data in 20 km × 20 km grid cells for the ferns of Japan.
+(A) Rolling mean area (square km) in 20 km latitudinal bands.
+(B) Percentage apomictic taxa.
+(C) Annual precipitation.
+(D) Percipitation seasonality (coefficient of variation).
+(E) Mean annual temperature.
+(F) Temperature seasonality (standard deviation × 100).
 Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.
@@
 
-Time tree of native Japanese ferns and lycophytes, excluding hybrids. Points at nodes indicate bootstrap support. Scalebar 100 my. Tree visualized with the `r pack("ggtree")` R package [@Yu2017].
+Time tree of native Japanese ferns and lycophytes, excluding hybrids.
+Points at nodes indicate bootstrap support.
+Scalebar = 100 Myr.
+Tree visualized with the `r pack("ggtree")` R package [@Yu2017].
 
@@
 
-Observed number of specimens (A) and sampling redundancy (B) per 20 km × 20 km grid cell in the ferns of Japan.
+Observed number of specimens (A) and sampling redundancy (B) per 20 km × 20 km grid cell for ferns of Japan.
 
@@
 
-(A) Raw taxonomic diversity, (B) phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude.
+(A) Raw taxonomic diversity, (B) phylogenetic diversity (PD), and (C) functional diversity (FD) of the ferns of Japan plotted by latitude.
 For (B) and (C), raw branch lengths were transformed to relative values by dividing each branchlength by the sum of all branch lengths.
@@
     values = canape_cols,
-    na.value = "transparent") +
+    breaks = names(canape_cols)[names(canape_cols) != "not significant"],
+    name = "Endemism type") +
   annotation_scale(location = "tl", width_hint = 0.2) +
-  map_theme() +
-  theme(legend.title = element_blank())
+  map_theme()
 ```
@@
 Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism), restricted data set including only taxa endemic to Japan.
-'neo' indicates neodendemic areas (those with an overabundance of short branches); 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches.
+neo, neodendemic areas (those with an overabundance of short branches); paleo, paleoendemic areas (those with an overabundance of long branches); mixed, significantly endemic sites that have neither an overabundance of short nor long branches; super, highly significantly endemic sites that have neither an overabundance of short nor long branches.
+Grid cells without significant endemism in grey.
 For details, see Materials and Methods.
@@
 
-Selection of *k* for bioregions.
-(A) taxonomic bioregions.
-(B) phylogenetic bioregions.
+Selection of *k* for bioregions in the ferns of Japan.
+(A) Taxonomic bioregions.
+(B) Phylogenetic bioregions.
 Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).
@@
 
-Bioregion dendrograms.
-Tips are sites (names not shown).
-Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods).
-Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
-(A) taxonomic bioregions.
-(B) phylogenetic bioregions.
+Bioregion dendrograms for the ferns of Japan.
+Tips are grid cells (names not shown).
+Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid cells (see Materials and Methods).
+Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid cells each are lumped into the "Other" category.
+(A) Taxonomic bioregions.
+(B) Phylogenetic bioregions.
 Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
+
 \clearpage
@@
   scale_x_continuous(labels = percent, limits = c(0, .85)) +
-  labs(x = "% apomictic taxa") +
+  labs(x = "% Apomictic taxa") +
   theme(
@@
 
-Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon. Numbers next to each bar show number of apomictic taxa out of total richness per genus.
+Rate of apomixis in genera of native, non-hybrid Japanese ferns with >1 apomictic taxon.
+Numbers next to each bar show number of apomictic taxa out of total richness per genus.
 
@@
 
-Relationship between (A) phylogenetic diversity (PD) and (B) relative PD and % apomictic taxa, models inferred using phylogeny with branch lengths in units of expected genetic change (untransformed ML tree).
+Relationship between (A) phylogenetic diversity (PD) and (B) relative PD (RPD) and % apomictic taxa in the ferns of Japan, models inferred using phylogeny with branch lengths in units of expected genetic change (untransformed ML tree).
 Ribbon shows 95% confidence interval of model.
@@
 
-Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions arranged by mean annual temperature and annual precipitation.
+Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions of the ferns of Japan arranged by mean annual temperature and annual precipitation.
 Points shown with partial transparency; darker areas indicate overlapping points.
@@
 
-Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and protected areas.
+Maps showing overlap of grid cells with significantly high biodiversity and protected areas for the ferns of Japan .
 Grid cells with black outlines indicate significantly high biodiversity.

### 2022-04-27 ms/data_exploration.Rmd

@@
 
-So we will conservatively only remove the grid-cell with unusually high % apomictic taxa, then run Grubb's test again.
+So we will conservatively only remove the single grid-cell with most % apomictic taxa, and inspect the data.
 
@@
 biodiv_ferns_cent_no_outliers %>%
-  pull(percent_apo) %>%
-  grubbs_tidy()
+  arrange(desc(percent_apo)) %>%
+  select(percent_apo, richness)
 ```
 
-OK, no more outliers in `percent_apo`.
+That seems more reasonable.
 

### 2022-04-28 ms/SI.Rmd
Update SI captions and endem figure

@@
 
-(A) Raw taxonomic diversity, (B) phylogenetic diversity (PD), and (C) functional diversity (FD) of the ferns of Japan plotted by latitude.
+Raw biodiversity metrics of the ferns of Japan plotted by latitude.
+(A) Raw taxonomic diversity. 
+(B) Phylogenetic diversity (PD).
+(C) Functional diversity (FD).
 For (B) and (C), raw branch lengths were transformed to relative values by dividing each branchlength by the sum of all branch lengths.
@@
     values = canape_cols,
-    breaks = names(canape_cols)[names(canape_cols) != "not significant"],
     name = "Endemism type") +
@@
 neo, neodendemic areas (those with an overabundance of short branches); paleo, paleoendemic areas (those with an overabundance of long branches); mixed, significantly endemic sites that have neither an overabundance of short nor long branches; super, highly significantly endemic sites that have neither an overabundance of short nor long branches.
-Grid cells without significant endemism in grey.
+Grid cells without significant endemism in light grey.
+Grid cells without at least two taxa endemic to Japan in dark grey.
+
 For details, see Materials and Methods.

### 2022-04-28 ms/SI.Rmd
Fix richness of endemic Japanese ferns data
- previously included richness of all ferns

@@
 ```{r endemism-restricted, fig.width = 5, fig.height = 5}
+# Drop one grid cell with NA for endem type
+biodiv_ferns_endemic_spatial_no_na <- biodiv_ferns_endemic_spatial %>%
+  filter(!is.na(endem_type))
+
 ggplot() +
-  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
-  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent") +
-  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
+  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
+  geom_sf(data = biodiv_ferns_endemic_spatial_no_na, aes(fill = endem_type), color = "transparent") + #nolint
+  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
   scale_fill_manual(
     values = canape_cols,
+    breaks = names(canape_cols)[names(canape_cols) != "not significant"],
     name = "Endemism type") +
-  annotation_scale(location = "tl", width_hint = 0.2) +
+  annotation_scale(location = "br", width_hint = 0.2) +
   map_theme()
@@
 Grid cells without significant endemism in light grey.
-Grid cells without at least two taxa endemic to Japan in dark grey.
 

### 2022-04-28 ms/manuscript.Rmd
Update MS
According to corrections made to proof

@@
 
-Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1^, and Atsushi Ebihara^3^
+Joel H. Nitta^1,4^, Brent D. Mishler^2^, Wataru Iwasaki^1^, and Atsushi Ebihara^3^
 
@@
 
+^4^Current affiliation: Department of Integrated Biosciences, Graduate School of Frontier Sciences, The University of Tokyo, Chiba, Japan
+
 __Correspondence__
 
-Joel H. Nitta, #601 Iwasaki Lab, Bioscience Bldg, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-0882, Japan
+Joel H. Nitta, #601 Bioscience Bldg, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-0882, Japan
 
@@
 
-**Premise:** Biodiversity is often only measured with species richness; however, this measure ignores evolutionary history and is not sufficient for making conservation decisions.
+**Premise:** Biodiversity is often only measured with species richness; however, this metric ignores evolutionary history and is not sufficient for making conservation decisions.
 Here, we characterize multiple facets and drivers of biodiversity to understand how these relate to bioregions and conservation status in the ferns of Japan.
@@
 
-**Results:** Taxonomic richness was highest at mid latitudes.
+**Results:** Taxonomic richness was highest at mid-latitudes.
 Phylogenetic and functional diversity and phylogenetic endemism were highest in small southern islands.
-Relative phylogenetic and functional diversity were high at high and low latitudes, and low at mid latitudes.
+Relative phylogenetic and functional diversity were high at high and low latitudes, and low at mid-latitudes.
 Grid cells were grouped into `r maj_cluster_count$phylo_cluster$n %>% english2` (phylogenetic) or `r maj_cluster_count$taxonomic_cluster$n %>% english2` (taxonomic) major bioregions.
@@
 However, using this data set alone for phylogenetic analysis suffers two drawbacks: it is not densely sampled enough to resolve some internal nodes, and it lacks many lineages that have fossils available for molecular dating.
-Furthermore, use of community phylogenies generated by only by sampling the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
-Because one goal of the study is to understand the distribution of paleo- vs. neoendemic areas (defined in units of time), we require an ultrametric tree.
+Furthermore, use of community phylogenies generated by sampling only the species present in the local community has been shown to produce spurious results in simulation studies [@Park2018].
+Because one goal of the study is to understand the distribution of paleo- vs. neoendemic areas (defined in units of time), we required an ultrametric tree.
 Therefore, to obtain a robust, ultrametric tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid data set including all fern sequences on GenBank for four widely sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes [FTOL v0.0.1\; @Nitta2022b] (`r ape::Ntip(plastome_tree) %>% number()` operational taxonomic units [OTUs] total, including outgroup taxa).
@@
 We measured phylogenetic diversity (PD) using Faith's PD [the total branch length connecting all terminal taxa in a community, including the root of the phylogeny\; @Faith1992].
-To detect areas with unusually long or short branches, we calculated relative phylogenetic diversity (RPD), which is a ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].
+To detect areas with unusually long or short branches, we calculated relative phylogenetic diversity (RPD), which is the ratio of observed PD to PD measured on a tree where all branch lengths have been transformed to equal length [@Mishler2014].
 
@@
 Grid cells with significantly low PD were almost all located on the main islands of Japan (Honshu, Kyushu, and Shikoku; `r figure("rand-diversity")`A).
-A smaller number of grid cells had significantly high PD, and these were located mostly on small southern (Ryukyu, Izu, Ogasawara) islands or coastal areas (`r eg`, Kii Peninsula).
+A smaller number of grid cells had significantly high PD, and these were located mostly on small southern islands (Ryukyu, Izu, Ogasawara) or coastal areas (`r eg`, Kii Peninsula).
 Grid cells with significantly low RPD were mostly observed in the southern main islands, particularly in and around Kyushu (`r figure("rand-diversity")`B).
@@
 The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models.
-Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A).
+Richness showed a hump-shaped relationship with temperature, reaching maximum values at intermediate temperatures (`r figure("modelfit")`A).
 Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E).
@@
 
-The percentage of protected area in grid cells with significantly high biodiversity range from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A).
+The percentage of protected area in grid cells with significantly high biodiversity ranges from `r min_protected_total` to `r max_protected_total` total, including `r min_protected_high` to `r max_protected_high` with high status and `r min_protected_med` to `r max_protected_med` with medium status (`r figure("conserv-status")`A).
 Cells with significantly high biodiversity had a similar or greater percentage of area protected compared to the baseline rate for Japan at the medium protection status (`r scales::percent(total_percent_protected_medium, accuracy = 0.1)`) across all biodiversity metrics.
@@
 The overwhelming pattern characterizing phylogenetic and functional diversity (PD and FD) is the high diversity of the southern subtropical islands (Ryukyu and Ogasawara Islands) (`r figure("rand-diversity")`A, C).
-This pattern is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country, which lack subtropical climate; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima Island).
+This pattern is almost certainly due to the presence of primarily tropical lineages that do not occur in other parts of the country, which lack subtropical climate; `r n_southern_genera` genera and `r n_southern_families` families only occur south of 30.1° latitude (just south of Yakushima).
 The similarity in patterns of PD and FD is likely due to the moderate degree of phylogenetic signal present in the traits used in our analysis (Appendix S1: Tables `r s_table_num("phy-sig-binary")`, `r s_table_num("phy-sig")`) and suggests that at least in the ferns of Japan, phylogeny can be used as a reasonable stand-in for functional diversity.
-Given their high PD, it is perhaps not surprising that southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid cells with significant levels of PE, as well as paleoendemic grid cells, are located in the southern islands (`r figure("endemism")`; Appendix S1: `r s_figure("endemism-restricted")`).
-The high PE observed on the southern subtropical islands is clearly due to the small land area of these islands: small area combined with high PD is expected to lead to high PE.
+Given their high PD, it is perhaps not surprising that the southern islands also host high amounts of phylogenetic endemism (PE); indeed, the vast majority of grid cells with significant levels of PE, as well as paleoendemic grid cells, are located in the southern islands (`r figure("endemism")`; Appendix S1: `r s_figure("endemism-restricted")`).
+The high PE observed on the southern islands is clearly due to their small land area: small area combined with high PD is expected to lead to high PE.
 
@@
 A similar argument has been made to assert that areas with seasonal monsoons are linked to increased rates of apomixis in ferns [@Liu2012; @Tanaka2014; @Picard2021], but in Japan, although the areas that include high percentages of apomictic taxa are monsoonal, the degree of seasonal water limitation does not approach that of a desert.
-The monsoon hypothesis was not supported by our results, which showed neither a particularly strong effect of precipitation seasonality in most models (`r figure("coeff")`), nor a correlation with % apomictic taxa (Appendix S1: `r s_table("corr")`).
+The monsoon hypothesis was not supported by our results, which showed neither a particularly strong effect of precipitation seasonality in most models (`r figure("coeff")`), nor a correlation of precipitation seasonality with % apomictic taxa (Appendix S1: `r s_table("corr")`).
 
@@
 Interestingly, the bioregions identified here correspond roughly to major vegetation zones previously categorized ad hoc; bioregion 1 corresponds to the cool or subartic deciduous type, bioregion 2 to the warm evergreen type, and bioregions 3 and 4 to the subtropical rainy type [@Shimizu2014].
-Furthermore, the border between bioregions 1 and 2 approximately corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he proposed based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
+Furthermore, the border between bioregions 1 and 2 approximately corresponds to the "*Gleichenia japonica* Line" proposed by @Nakaike1983, which he based on the distribution of *Diplopterygium glaucum* (Houtt.) Nakai (= *Gleichenia japonica* Spreng.) as an indicator of evergreen broad-leaf forests in Japan.
 The agreement between qualitative analysis of bioregions here with previous ad hoc schemes suggests that ferns are useful as bioindicators in Japan.
 Although some studies have emphasized the importance of deep-sea straits (the Tsugaru Strait separating Hokkaido from Honshu and the Tokara Strait separating the southern islands from Kyushu) in structuring biological communities in Japan [@Millien-Parra1999; @Kubota2014], they did not play a major role in structuring fern bioregions.
-The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, the regions have some overlap on Yakushima and Amami Islands (`r figure("bioregions")`B).
+The Tsugaru Strait had no relation to any of the bioregions, and although the Tokara Strait somewhat splits phylogenetic bioregions 2 and 3, these regions overlap near Yakushima (`r figure("bioregions")`B).
 The weak effect of these straits as barriers makes sense given the presumably high dispersal abilities of ferns, which disperse via tiny spores carried on the wind [@Tryon1970].
@@
 First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan.
-Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (Appendix S1: `r s_figure("abundance")`).
+Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (Appendix S1: `r s_figure("abundance")`A).
 Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.
@@
 
-Additional supporting information may be found online in the Supporting Information section at the end of the article.
+This article includes online-only Supplemental Data.
 
@@
 **`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20-km grid cell in the ferns of Japan.  
-**`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
+**`r s_figure("raw-lat")`**. Raw biodiversity metrics of the ferns of Japan plotted by latitude.  
 **`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
@@
 Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models for the effect of environment and reproductive mode on biodiversity metrics in the ferns of Japan.
-(A--E) Environmental models (models including temperature).
-(F, G) Reproductive models (models including % apomictic taxa).
+(A--E) Environmental (env.) models (models including temperature).
+(F, G) Reproductive (repro.) models (models including % apomictic taxa).
+Precip., precipitation; Temp., temperature; SES, standard effect size; PD, phylogenetic diversity; RPD, relative PD; FD, functional diversity; RFD, relative FD.
 Statistical significance was assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-values indicate probability of no difference between the full and null model.
@@
 Relationship between biodiversity metrics and selected predictor variables in the ferns of Japan.
-(A--E) Environmental (env.) models (models including temperature).
-Precip., precipitation; Temp., temperature; Temp^2^, temperature squared.
-(F, G) Reproductive (repro.) models (models including % apomictic taxa).
+(A--E) Environmental models (models including temperature).
+(F, G) Reproductive models (models including % apomictic taxa).
 Ribbon shows 95% confidence interval of model.
@@
 neo, grid cells with an overabundance of rare short branches; paleo, grid cells with an overabundance of rare long branches; mixed, significantly endemic grid cells that have neither an overabundance of short nor long branches; super, highly significantly endemic grid cells that have neither an overabundance of short nor long branches.
-Grid cells without significant endemism in grey.
+Grid cells without significant endemism in light grey.
 For details, see Materials and Methods.

### 2022-04-28 ms/SI.Rmd
Print out SI tree fig to PDF first
fixes problem of jumbled text in tip labels

@@
 
-```{r tree, fig.width = 8, fig.height = 8}
+```{r tree_make}
 phy <- japan_pterido_tree
@@
 
-rotate_tree(p, 3)
+p_rot <- rotate_tree(p, 3)
+
+tree_fig_file <- glue::glue("results/{s_figure('tree')}.pdf") %>%
+  str_replace_all(" ", "_") %>%
+  here::here()
+
+ggsave(
+ file = tree_fig_file, plot = p_rot, height = 7, width = 7, units = "in")
+```
+
+```{r, tree-print, fig.width = 7, fig.height = 7, results = "show"}
+knitr::include_graphics(tree_fig_file)
 ```

### 2022-04-28 ms/SI.Rmd
Update SI

@@
 Reproductive mode and environmental data in 20 km × 20 km grid cells for the ferns of Japan.
-(A) Rolling mean area (square km) in 20 km latitudinal bands.
+(A) Rolling mean area (square km) in 100-km latitudinal bands.
 (B) Percentage apomictic taxa.
@@
 
-Relationships between observed phylogenetic and functional diversity and taxonomic richness in the ferns of Japan.
+Relationships between raw phylogenetic and functional diversity and taxonomic richness in the ferns of Japan.
 (A) Phylogenetic diversity (PD).
@@
 neo, neodendemic areas (those with an overabundance of short branches); paleo, paleoendemic areas (those with an overabundance of long branches); mixed, significantly endemic sites that have neither an overabundance of short nor long branches; super, highly significantly endemic sites that have neither an overabundance of short nor long branches.
-Grid cells without significant endemism in light grey.
-
-For details, see Materials and Methods.
+Grid cells without significant endemism in light grey. For details, see Materials and Methods.
 

### 2022-04-28 ms/data_exploration.Rmd
Update data exploration

@@
 ```{r plot-raw-data, fig.height = 8}
-ggplot(biodiv_ferns_cent_no_outliers, aes(x = .panel_x, y = .panel_y)) + 
+biodiv_ferns_cent_no_outliers %>%
+  # Convert precipitation to meters so 
+  # the numbers don't overlap on the x-axis
+  mutate(precip = precip/1000) %>%
+  ggplot(aes(x = .panel_x, y = .panel_y)) + 
   geom_point(alpha = 0.5, shape = 16, size = 0.5) + 

### 2022-04-28 ms/data_readme.Rmd
Update data README
Add DOIs

@@
 (2022) Data from: Spatial phylogenetics of Japanese ferns: Patterns, processes,
-and implications for conservation FIXME: add DOI when available
+and implications for conservation. FigShare. 
+http://doi.org/10.6084/m9.figshare.16655263
 
@@
 Patterns, processes, and implications for conservation.
-American Journal of Botany. FIXME: add DOI, page numbers when available
+American Journal of Botany. https://doi.org/10.1002/ajb2.1848
 
@@
 MD5 checksum `r tools::md5sum(here::here(ebihara_2019_zip_file))`
-- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment 
-and phylogenetic tree (Nitta et al., in prep.) downloaded from 
-https://figshare.com/s/87b8ec102303bf9f3212.
+- ftol_data_release_v0.0.1.zip: Fern Tree of Life (FTOL) project global DNA 
+alignment and phylogenetic tree v0.0.1 (Nitta et al., in prep 
+https://doi.org/10.1101/2022.03.31.486640) downloaded from 
+https://doi.org/10.6084/m9.figshare.13251290.v1.
 MD5 checksum `r tools::md5sum(here::here(ftol_zip_file))`
-FIXME: replace with DOI when available
 
@@
 phylogenetics of Japanese ferns: Patterns, processes, and implications for
-conservation. American Journal of Botany. 
-FIXME: add DOI, page numbers when available
+conservation. American Journal of Botany. https://doi.org/10.1002/ajb2.1848
 

### 2023-03-28 ms/manuscript.Rmd
Update MS
Update ms use standard fig dpi throughout

@@
 
-Joel H. Nitta, #601 Bioscience Bldg, The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-0882, Japan
+Joel H. Nitta, #601 Bioscience Bldg., The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa, Chiba 277-0882, Japan
 
@@
   st_transform(4612)
+
+fig_dpi <- 600
 ```
@@
 ggsave(
-  result_file(figure("japan-map"), "pdf"),
-  japan_map_plot, width = 5, height = 5, units = "in")
+  result_file(figure("japan-map"), "png"),
+  japan_map_plot, width = 5, height = 5, units = "in", dpi = fig_dpi)
 ```
@@
 ggsave(
-  result_file(figure("raw-diversity"), "pdf"),
-  width = 7.25, height = 3.4, units = "in", raw_diversity_plot)
+  result_file(figure("raw-diversity"), "png"),
+  width = 7.25, height = 3.4, units = "in", raw_diversity_plot, dpi = fig_dpi)
 ```
@@
 ggsave(
-  result_file(figure("rand-diversity"), "pdf"),
-  rand_diversity_plot, width = 7.25, height = 8, units = "in")
+  result_file(figure("rand-diversity"), "png"),
+  rand_diversity_plot, width = 7.25, height = 8, units = "in", dpi = fig_dpi)
 ```
@@
 ggsave(
-  result_file(figure("coeff"), "pdf"),
-  coeff_plot, width = 7.25, height = 5, units = "in")
+  result_file(figure("coeff"), "png"),
+  coeff_plot, width = 7.25, height = 5, units = "in", dpi = fig_dpi)
 ```
@@
 ggsave(
-  result_file(figure("modelfit"), "pdf"),
-  modelfit_plot, width = 7.25, height = 5, units = "in")
+  result_file(figure("modelfit"), "png"),
+  modelfit_plot, width = 7.25, height = 5, units = "in", dpi = fig_dpi)
 ```
@@
 ggsave(
-  result_file(figure("endemism"), "pdf"),
-  endemism_plot, width = 5, height = 5, units = "in")
+  result_file(figure("endemism"), "png"),
+  endemism_plot, width = 5, height = 5, units = "in", dpi = fig_dpi)
 ```
@@
 ggsave(
-  result_file(figure("bioregions"), "pdf"),
-  bioregions_plot, width = 7.25, height = 4.3, units = "in")
+  result_file(figure("bioregions"), "png"),
+  bioregions_plot, width = 7.25, height = 4.3, units = "in", dpi = fig_dpi)
 ```
@@
 ggsave(
-  result_file(figure("div-by-tax-region"), "pdf"),
-  div_by_tax_region_plot, width = 7.25, height = 7, units = "in")
+  result_file(figure("div-by-tax-region"), "png"),
+  div_by_tax_region_plot, width = 7.25, height = 7, units = "in", dpi = fig_dpi)
 ```
@@
 ggsave(
-  result_file(figure("conserv-status"), "pdf"),
-  conserv_status_plot, width = 7.25, height = 4, units = "in")
+  result_file(figure("conserv-status"), "png"),
+  conserv_status_plot, width = 7.25, height = 4, units = "in", dpi = fig_dpi)
 ```

### 2023-04-17 ms/shida-no-kai-figs.Rmd
Add Rmd to produce images with Japanese

---
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions:
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output:
  bookdown::word_document2:
    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
    toc: no
    number_sections: no
    reference_docx: american-journal-of-botany.docx
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "american-journal-of-botany.csl"]
params:
  doc_type: doc
editor_options:
  chunk_output_type: console
---

---
bibliography:
  - `r here::here("ms/references.yaml")`
  - `r here::here("ms/references_other.bib")`
---

```{r manual, include = FALSE, evaluate = FALSE}
# Run these lines if rendering outside of targets plan
library(targets)
library(tarchetypes)
source(here::here("R/packages.R"))
source(here::here("R/functions.R"))
```

```{r setup, include = FALSE, cache = FALSE}
# Load libraries used for rendering Rmds
library(flextable)
library(kableExtra)
library(ftExtra)
library(scales)
library(patchwork)
library(RColorBrewer)
library(ggspatial) # for scalebars on maps

# Set knitr options
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  results = "hide", cache = FALSE)

# Load functions and variables only used when rendering Rmds
source(here::here("R/ms_functions.R"), local = TRUE)
```

```{r load-data}
# Load objects from targets workflow
tar_load(c(
  aic_env_repro,
  binary_sig_results,
  biodiv_ferns_cent,
  biodiv_ferns_spatial,
  broad_alignment_list,
  comm_ferns,
  comm_ferns_full,
  csl_file,
  deer_range,
  fern_traits,
  green_list,
  japan_fern_tree,
  japan_map_points,
  japan_pterido_tree,
  japan_rbcL,
  japan_shp,
  k_phylogeny,
  k_taxonomy,
  lat_span_summary,
  lrt_results,
  model_fits,
  model_params,
  ms_functions,
  occ_point_data_summary,
  phy_sig_results,
  plastome_calibration_dates,
  plastome_tree,
  ppgi,
  protected_areas,
  refs_other_yaml,
  refs_yaml,
  refs_other_bib,
  repro_data,
  signif_cells_protected_area,
  signif_cells_deer_danger,
  template_file,
  traits_summary,
  biodiv_ferns_spatial_lumped,
  bioregion_cutoff
))
```

```{r plot-prep}
# (color palettes are set in ms_functions.R)

# Define nicely formatted terms for printing variable names
resp_var_lookup <- tribble(
  ~resp_var, ~resp_var_pretty,
  "fd_obs_z", "SES of FD",
  "pd_obs_z", "SES of PD",
  "pe_obs_p_upper", "PE *p*-value",
  "rfd_obs_z", "SES of RFD",
  "richness", "Richness",
  "rpd_obs_z", "SES of RPD",
)

indep_var_lookup <- tribble(
  ~term, ~term_pretty,
  "temp", "Temperature",
  "precip", "Precipitation",
  "precip_season", "Precipitation seasonality",
  "I(temp^2)", "Temperature^2^",
  "percent_apo", "% apomictic taxa"
)

# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial)) %>%
  st_transform(4612)

# Format biodiv_ferns_spatial data for plotting
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>%
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(
      endem_type,
      levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all
    # variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x,
      "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  ) %>%
  st_transform(4612)

fig_dpi <- 600

# For powerpoint presentations, increase base font size
theme_set(theme_grey(base_size = 20))
```

```{r raw-diversity-make}
trend_line_theme <- function () {
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )
}

# Convert diversity metrics dataframe to centroids for plotting trends
ses_div_centroids <-
  biodiv_ferns_spatial %>%
  mutate(
    longitude = sf::st_centroid(geom) %>% st_coordinates %>% magrittr::extract(,1),
    latitude = sf::st_centroid(geom) %>% st_coordinates %>% magrittr::extract(,2)) %>%
  as_tibble %>%
  dplyr::select(-geom)
# Get min and max latitude for setting common limits between
# maps and trendlines
ja_lat_min <- ses_div_centroids %>% pull(latitude) %>% min()
ja_lat_max <- ses_div_centroids %>% pull(latitude) %>% max()

# Fern richness
# - map
a <-
  ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
  geom_sf() +
  scale_fill_scico(palette = "lajolla") +
  labs(fill = "種数") +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  theme(
    legend.title = element_text(size = 30/.pt),
    legend.text = element_text(size = 24/.pt),
    legend.justification=c(0,1), legend.position=c(0,1)
  )
# - latitudinal trends
b <- ggplot(ses_div_centroids, aes(x = latitude, y = richness)) +
  geom_point(color = "grey50", shape = 1 ) +
  geom_smooth(se = FALSE) +
  xlim(ja_lat_min, ja_lat_max) +
  coord_flip() +
  labs(y = "種数") +
  trend_line_theme()
a + b + plot_layout(widths = c(3,1), heights = c(1,1))

ggsave("working/shida-no-kai/richness.png")
```

```{r rand-diversity-make}
# PD
pd_rand <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent") + #nolint
  # label
  labs(title = "PD") +
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  scale_fill_manual(
    name = "有意性",
    values = signif_cols,
    labels = signif_labels) +
  map_theme() +
  theme(
    legend.position = "right",
    legend.text = element_markdown()
  )

ggsave("working/shida-no-kai/pd_rand.png")

# FD
fd_rand <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent") + #nolint
  # label
  labs(title = "FD") +
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  scale_fill_manual(
    name = "有意性",
    values = signif_cols,
    labels = signif_labels) +
  map_theme() +
  theme(
    legend.position = "right",
    legend.text = element_markdown()
  )

ggsave("working/shida-no-kai/fd_rand.png")

# RPD
rpd_rand <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent") + #nolint
  labs(title = "RPD") +
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  scale_fill_manual(
    name = "有意性",
    values = signif_cols,
    labels = signif_labels) +
  map_theme() +
  theme(
    legend.position = "right",
    legend.text = element_markdown()
  )

ggsave("working/shida-no-kai/rpd_rand.png")
```

```{r bioregions-make}
# Taxonomic bioregions
a <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = taxonomic_cluster), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  labs(title = "分類学的距離") +
  labs(fill = "バイオリージョン") +
  scale_fill_manual(
    values = bioregion_cols,
    guide = guide_legend(
      title.position = "top"
    )) +
  map_theme()

# Phylogenetic bioregions
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = phylo_cluster), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  labs(title = "系統的距離") +
  labs(fill = "バイオリージョン") +
  scale_fill_manual(
    values = bioregion_cols,
    guide = guide_legend(
      title.position = "top"
    )) +
  annotation_scale(location = "br", width_hint = 0.2) +
  map_theme() +
  theme(axis.text.y = element_blank())

bioregions_plot <- a + b +
  plot_layout(ncol = 2)

ggsave("working/shida-no-kai/bioregions.png")

# Taxonomic, plain version
ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) + #nolint
  geom_sf(data = biodiv_ferns_spatial_lumped, aes(fill = taxonomic_cluster), color = "transparent") + #nolint
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) + #nolint
  labs(fill = "バイオリージョン") +
  scale_fill_manual(
    values = bioregion_cols,
    guide = guide_legend(
      title.position = "top"
    )) +
  map_theme()

ggsave("working/shida-no-kai/bioregions_taxonomic.png")
```

```{r find-top-species}
# Extract clusters
clusters <-
  biodiv_ferns_spatial_lumped %>%
  as_tibble() %>%
  select(site = "grids", cluster = taxonomic_cluster, -geom)
# Convert community df to long format
comm_ferns_long <-
  comm_ferns %>%
  rownames_to_column("site") %>%
  pivot_longer(names_to = "taxon", values_to = "abun", -site)
# Find top-scoring species (with high occurrence inside one cluster, and low elsewhere)
top_species <-
  comm_ferns_long %>%
  left_join(clusters, by = "site") %>%
  filter(cluster %in% 1:4) %>%
  group_by(taxon, cluster) %>%
  summarize(
    abun = sum(abun),
    total = n()
  ) %>%
  mutate(
    # percent abundance within a bioregion
    perc_abun = abun / total
  ) %>%
  arrange(taxon, desc(perc_abun)) %>%
  select(taxon, cluster, perc_abun) %>%
  group_by(taxon) %>%
  # take top two highest percent abundances per species, calculate their difference
  slice_max(n = 2, order_by = perc_abun, with_ties = FALSE) %>%
  mutate(
    top_cluster = cluster[perc_abun == max(perc_abun)]
  ) %>%
  summarize(
    max_dif = max(perc_abun) - min(perc_abun),
    cluster = unique(top_cluster)
  ) %>%
  # keep only those with greater than 65% difference between top 1st and 2nd 
  # most abundant bioregions
  filter(max_dif > 0.5) %>%
  group_by(cluster) %>%
  # Choose top three per cluster for plotting
  slice_max(n = 3, order_by = max_dif, with_ties = FALSE)
```

```{r plot-top-species-prep}
# make helper function to plot top-scoring species
plot_top_taxon <- function (taxon_select, ses_div_ferns_spatial, comm_ferns_long) {
ses_div_ferns_spatial %>%
  left_join(
    filter(comm_ferns_long, taxon == taxon_select),
    by = c(grids = "site")
  ) %>%
  mutate(abun = factor(abun)) %>%
  ggplot(aes(fill = abun)) +
  scale_fill_manual(values = c("#404040", "#d01c8b")) +
  geom_sf(color = "transparent") +
  labs(
    subtitle = str_replace_all(taxon_select, "_", " ")
  ) +
  map_theme() +
  theme(
    plot.subtitle = element_text(face = "italic"),
    legend.position = "none"
  )
}
```

```{r plot-top-species-cluster-1, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial_lumped %>%
  mutate(focus = ifelse(taxonomic_cluster == 1, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_manual(values = bioregion_cols) +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  labs(subtitle = "リージョン 1") +
  theme(legend.position = "none")

focus_region +
plot_top_taxon(top_species$taxon[[1]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[2]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[3]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave("working/shida-no-kai/region_1.png")
```

```{r plot-top-species-cluster-2, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial_lumped %>%
  mutate(focus = ifelse(taxonomic_cluster == 2, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_manual(values = bioregion_cols) +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  labs(subtitle = "リージョン 2") +
  theme(legend.position = "none")

focus_region +
plot_top_taxon(top_species$taxon[[4]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[5]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[6]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave("working/shida-no-kai/region_2.png")
```

```{r plot-top-species-cluster-3, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial_lumped %>%
  mutate(focus = ifelse(taxonomic_cluster == 3, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_manual(values = bioregion_cols) +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  labs(subtitle = "リージョン 3") +
  theme(legend.position = "none")

focus_region +
plot_top_taxon(top_species$taxon[[7]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[8]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[9]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave("working/shida-no-kai/region_3.png")
```

```{r plot-top-species-cluster-4, out.width = "120%", fig.asp = 0.65}
focus_region <-
biodiv_ferns_spatial_lumped %>%
  mutate(focus = ifelse(taxonomic_cluster == 4, "yes", "no")) %>%
ggplot(aes(fill = taxonomic_cluster, alpha = focus)) +
  geom_sf(color = "transparent") +
  scale_fill_manual(values = bioregion_cols) +
  scale_alpha_manual(values = c(yes = 1, no = 0.25)) +
  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  map_theme() +
  labs(subtitle = "リージョン 4") +
  theme(legend.position = "none")

focus_region +
plot_top_taxon(top_species$taxon[[10]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[11]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_top_taxon(top_species$taxon[[12]], biodiv_ferns_spatial_lumped, comm_ferns_long) +
plot_layout(ncol = 2, nrow = 2)

ggsave("working/shida-no-kai/region_4.png")
```

```{r rpd-vs-apo}
# Extract dataset used for model, including significance columns
biodiv_ferns_used_for_model <- biodiv_ferns_spatial %>%
  # drop geometry (only need centroids for modeling)
  st_set_geometry(NULL) %>%
  as_tibble() %>%
  # filter to only sites used in model
  inner_join(
    biodiv_ferns_cent %>% as_tibble() %>% select(grids),
    by = "grids"
  )

rpd_vs_apo_data <-
  model_fits %>%
  mutate(
    signif_var = str_replace_all(resp_var, "_obs_z", "_signif"),
    resp_var_print = case_when(
      resp_var == "richness" ~ "Richness (no. spp.)",
      TRUE ~ str_to_upper(resp_var) %>% word(1, 1, "_") %>% paste("SES of", .)
    ),
    model_data = list(biodiv_ferns_used_for_model),
    signif_cols = list(signif_cols)) %>%
  filter(resp_var == "rpd_obs_z", indep_var == "percent_apo")

ggplot(rpd_vs_apo_data$model_data[[1]], aes(x = percent_apo, y = rpd_obs_z)) +
  labs(
    y = "RPDの効果量",
    x = "無配生殖種の割合") +
  geom_point(aes(color = rpd_signif), shape = 16, size = 3) +
  scale_color_manual(values = signif_cols) +
  scale_x_continuous(labels = function(x) paste0(round(x * 100), "%")) +
  geom_line(data = rpd_vs_apo_data$fit[[1]], alpha = 0.35, color = "blue") +
  geom_ribbon(
    data = rpd_vs_apo_data$fit[[1]],
    aes(ymin = low, ymax = up), 
    alpha = 0.15
  ) +
  theme(legend.position = "none")

ggsave("working/shida-no-kai/rpd_vs_apo.png")
```

```{r protected-areas, out.width = "120%", fig.asp = 0.65, dev="png", dpi=200}
japan_box <- sf::st_bbox(biodiv_ferns_spatial)

protection_cols_ja_map <- protection_cols
names(protection_cols_ja_map) <- str_to_lower(names(protection_cols_ja_map))
protection_cols_ja_map <- protection_cols_ja_map[c("medium", "high")]

status_labels <-
protected_areas %>%
  # area is in sq m, so convert to sq kilom first
  mutate(area = st_area(.) / 1e6) %>%
  mutate(
    area = as.numeric(area),
    area_chr = area %>% round %>% scales::number(big.mark = ","),
    area_percent = area %>% magrittr::divide_by(377900) %>% scales::percent(),
    label = glue::glue("{str_to_sentence(status)} ({area_percent})") %>% set_names(status)
) %>%
  pull(label) %>%
  str_replace_all("High", "高") %>%
  str_replace_all("Medium", "中") %>%
  set_names(c("high", "medium"))

ggplot() +
  geom_sf(data = japan) +
  geom_sf(data = protected_areas, aes(fill = status), color = "transparent") +
  coord_sf(xlim = c(japan_box$xmin, japan_box$xmax), ylim = c(japan_box$ymin, japan_box$ymax)) +
  labs(
    fill = "保全レベル"
  ) +
  scale_fill_manual(values = protection_cols_ja_map, labels = status_labels[names(protection_cols_ja_map)]) +
  theme(
    axis.ticks = element_blank()
  ) + 
  labs(
    caption = "環境省\nhttps://www.biodic.go.jp/biodiversity/activity/policy/map/map17/"
  )

ggsave("working/shida-no-kai/cons_areas.png")
```

```{r conserv-stats}
# Convert signif_cells_protected_area to wide format
signif_cells_protected_area_wide <- #nolint
  signif_cells_protected_area %>%
  select(status, metric, percent_protected) %>%
  pivot_wider(names_from = status, values_from = percent_protected) %>%
  rowwise %>%
  mutate(total = sum(medium, high)) %>%
  ungroup

# Calculate min and max percent area protected overall
# - define helper functions
calc_min_pa <- function(signif_cells_protected_area_wide, ...) { # nolint
  signif_cells_protected_area_wide %>%
    pull(...) %>%
    min() %>%
    scales::percent(accuracy = 0.1)
}
calc_max_pa <- function(signif_cells_protected_area_wide, ...) { # nolint
  signif_cells_protected_area_wide %>%
    pull(...) %>%
    max() %>%
    scales::percent(accuracy = 0.1)
}

min_protected_med <- calc_min_pa(signif_cells_protected_area_wide, medium)
max_protected_med <- calc_max_pa(signif_cells_protected_area_wide, medium)
min_protected_high <- calc_min_pa(signif_cells_protected_area_wide, high)
max_protected_high <- calc_max_pa(signif_cells_protected_area_wide, high)
min_protected_total <- calc_min_pa(signif_cells_protected_area_wide, total)
max_protected_total <- calc_max_pa(signif_cells_protected_area_wide, total)

# Calculate min and max protected area for the best protected category
signif_cells_protected_area_best <- # nolint
  signif_cells_protected_area_wide %>%
  slice_max(total)

# Verify that the best-protected biodiv type is PE
assertthat::assert_that(signif_cells_protected_area_best$metric == "pe")

best_protected_med <- signif_cells_protected_area_best %>%
  pull(medium) %>%
  scales::percent(accuracy = 0.1)
best_protected_high <- signif_cells_protected_area_best %>%
  pull(high) %>%
  scales::percent(accuracy = 0.1)
best_protected_total <- signif_cells_protected_area_best %>%
  pull(total) %>%
  scales::percent(accuracy = 0.1)

# Calculate the total area of Japan
area_of_japan <-
  japan_shp %>%
  mutate(
    area = st_area(.) %>%
      units::set_units(km^2)
  ) %>%
  pull(area)

# Calculate baseline protection percentage:
# percent protected areas across the whole country
total_percent_protected <-
  protected_areas %>%
  st_make_valid %>%
  mutate(
    area = st_area(.) %>%
      units::set_units(km^2)
  ) %>%
  st_drop_geometry %>%
  as_tibble() %>%
  filter(status %in% c("medium", "high")) %>%
  group_by(status) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_protected = total_area / area_of_japan) %>%
  mutate(percent_protected = as.numeric(percent_protected)) %>%
  # Add sum for high + medium
  bind_rows(
    summarize(
      .,
      percent_protected = sum(percent_protected),
      total_area = sum(total_area)
    ) %>% mutate(status = "total")
  )

# Extract values for medium and high protection level
total_percent_protected_medium <- total_percent_protected %>%
  filter(status == "medium") %>%
  pull(percent_protected)
total_percent_protected_high <- total_percent_protected %>%
  filter(status == "high") %>%
  pull(percent_protected)
total_percent_protected_all <- sum(
  total_percent_protected_medium, total_percent_protected_high)

### Deer ###

# Calculate baseline protection percentage:
# percent protected areas across the whole country
total_percent_deer_danger <-
  deer_range %>%
  mutate(
    area = st_area(.) %>% units::set_units(km^2)
  ) %>%
  st_drop_geometry %>%
  as_tibble() %>%
  group_by(range) %>%
  summarize(
    total_area = sum(area),
    .groups = "drop"
  ) %>%
  mutate(percent_danger = total_area / area_of_japan) %>%
  mutate(percent_danger = as.numeric(percent_danger))

# Convert to a list for easy indexing
deer_overall <- total_percent_deer_danger %>%
  mutate(
    # Format numbers for nice printing
    area = total_area %>%
      as.numeric %>%
      number(accuracy = 1) %>%
      paste("km^2^"),
    percent = scales::percent(percent_danger, accuracy = 0.1)
  ) %>%
  split(.$range)

deer_in_high_biodiv <-
signif_cells_deer_danger %>%
  group_by(range) %>%
  summarize(percent_danger = mean(percent_danger), .groups = "drop") %>%
  mutate(
    # Format numbers for nice printing
    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
  split(.$range)

deer_highest_danger <- signif_cells_deer_danger %>%
  slice_max(percent_danger) %>%
  # Make sure biodiv type with greatest danger is richness in estimated dataset
  verify(metric == "richness") %>%
  verify(range == "estimated") %>%
  mutate(
    # Format numbers for nice printing
    percent = scales::percent(percent_danger, accuracy = 0.1)) %>%
  split(.$range)
```

```{r conserv-status-make}
# Add total protected area (medium + high)
signif_cells_protected_area_total <- #nolint
  signif_cells_protected_area %>%
  group_by(metric) %>%
  summarize(percent_protected = sum(percent_protected), .groups = "drop") %>%
  mutate(status = "total")

total_percent_protected_for_plot <- #nolint
  total_percent_protected %>%
  select(status, percent_protected) %>%
  mutate(status = str_to_sentence(status))

# Plot protected areas
signif_cells_protected_area %>%
  select(metric, status, percent_protected) %>%
  bind_rows(signif_cells_protected_area_total) %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "種数",
    ),
    status = str_to_sentence(status)) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "種数"))) %>%
  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
  geom_col(position = position_dodge()) +
  geom_vline(
    data = total_percent_protected_for_plot,
    aes(xintercept = percent_protected, linetype = status),
    color = "grey30"
  ) +
  scale_linetype_manual(
    values = protection_lines,
    labels = c("中", "高", "合計"),
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      label.position = "left"
    )) +
  scale_fill_manual(
    values = protection_cols,
    labels = c("中", "高", "合計"),
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      label.position = "left"
    )) +
  scale_x_continuous(
    labels = function(x) scales::percent(x, accuracy = 1),
    expand = expansion(mult = c(0, .05))) +
  labs(
    x = "保全されている割合",
    y = "生物多様性の指標",
    fill = "保全レベル",
    linetype = "保全レベル"
  ) +
  theme_bw(base_size = 24) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "right"
  )

ggsave("working/shida-no-kai/protection_levels.png")

names(deer_lines) <- str_replace_all(
  names(deer_lines),
  "Estimated",
  "予想"
)

names(deer_cols) <- str_replace_all(
  names(deer_cols),
  "Estimated",
  "予想"
)

# Plot areas in danger from herbivory by deer
signif_cells_deer_danger %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "種数",
    ),
    range = str_to_sentence(range) %>%
      str_replace_all("Estimated", "予想")
  ) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "種数"))) %>%
  ggplot(aes(x = percent_danger, y = metric, fill = range)) +
  geom_col(position = position_dodge()) +
  geom_vline(
    data = total_percent_deer_danger %>% mutate(
      range = str_to_sentence(range) %>%
      str_replace_all("Estimated", "予想")),
    aes(linetype = range, xintercept = percent_danger),
    color = "grey30"
  ) +
  scale_linetype_manual(
    values = deer_lines,
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      label.position = "left"
    )) +
  scale_fill_manual(
    values = deer_cols,
    guide = guide_legend(
      direction = "horizontal",
      title.position = "top",
      label.position = "left"
    )) +
  scale_x_continuous(
    labels = function(x) scales::percent(x, accuracy = 1),
    expand = expansion(mult = c(0, .05))) +
  labs(
    x = "シカの面積",
    y = "生物多様性の指標",
    fill = "時代",
    linetype = "時代"
  ) +
  theme_bw(base_size = 20) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "right"
  )

ggsave("working/shida-no-kai/deer_levels.png")
```

```{r conserv-status-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(
  result_file(figure("conserv-status"), "png"),
  conserv_status_plot, width = 7.25, height = 4, units = "in", dpi = fig_dpi)
```

```{r conserv-status-print, fig.width = 7.25, fig.height = 4, eval = params$doc_type == "pdf"}
# Print plot if rendering to PDF
conserv_status_plot
```

**`r figure("conserv-status") %>% str_to_upper()`**. Conservation status and threat due to herbivory by deer in areas with high levels of biodiversity for ferns of Japan. (A) Percentage of land area with medium or high protected status for grid cells with significantly high biodiversity by protection status. (B) Percentage of land area occupied by deer for grid cells with significantly high biodiversity by deer range data source (1978 survey, 2003 survey, or range estimated from model based on 2003 survey data). Biodiversity metrics include taxon richness, phylogenetic diversity (PD), functional diversity (FD), and phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid cells in the top 5% considered highly diverse (see Materials and Methods). For (A), vertical lines indicate total percentage area protected across Japan (baseline protection rate). For (B), vertical lines indicate total percentage of land area occupied by deer across Japan (baseline threat rate).

### 2023-04-20 ms/shida-no-kai-figs.Rmd
Add PE plot

@@
 ggsave("working/shida-no-kai/richness.png")
+
+# Fern PE
+# - map
+a <-
+  ggplot() +
+  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
+  geom_sf(data = biodiv_ferns_spatial, aes(fill = pe_obs_p_upper), color = "transparent", alpha = 0.8) +
+  geom_sf() +
+  scale_fill_scico(palette = "lajolla") +
+  labs(fill = "PE P-value") +
+  scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1), limits = c(ja_lat_min, ja_lat_max)) +
+  scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
+  map_theme() +
+  theme(
+    legend.title = element_text(size = 30/.pt),
+    legend.text = element_text(size = 24/.pt),
+    legend.justification=c(0,1), legend.position=c(0,1)
+  )
+# - latitudinal trends
+b <- ggplot(ses_div_centroids, aes(x = latitude, y = pe_obs_p_upper)) +
+  geom_point(color = "grey50", shape = 1 ) +
+  geom_smooth(se = FALSE) +
+  xlim(ja_lat_min, ja_lat_max) +
+  coord_flip() +
+  labs(y = "PE P-value") +
+  trend_line_theme()
+a + b + plot_layout(widths = c(3,1), heights = c(1,1))
+
+ggsave("working/shida-no-kai/pe.png")
+
 ```
